### 2012-03-15 Examples/OPT/StatPDE/Example1/content.tex
Initial Transfer From OLD CVS Reprository

This example solves the distributed minimization problem
\begin{gather*}
\min J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2\\
\text{s.t.} (\nabla u,\nabla \phi) = (q+f,\phi)\;\;\forall\,\phi \in H^1_0(\Omega)
\end{gather*}
on the domain $\Omega = [0,1]^2$, and the data is chosen as follows:
\begin{gather*}
 f = \left(20\pi^2  \sin(4 \pi x) - \frac{1}{\alpha}  \sin(\pi x)\right) \sin(2 \pi y)\\
 u^d = \left( 5 \pi^2 \sin(\pi x) + \sin(4 \pi x)\right)  \sin(2\pi y)
\end{gather*}
and $\alpha = 10^{-3}$.
Hence its solution is given by:
\begin{gather*}
 \overline{q} = \frac{1}{\alpha} \sin(\pi x) \sin(2 \pi y)\\
 \overline{u} = \sin(4 \pi x) \sin(2 \pi y)
\end{gather*}

In addition the following functionals are evaluated:
\begin{gather*}
  \text{MidPoint: } u(0.5 ; 0.5)\\[2mm]
  \text{MeanValue: }\int_\Omega u
\end{gather*}

### 2012-03-15 Examples/OPT/StatPDE/Example2/content.tex
Initial Transfer From OLD CVS Reprository

This example solves the distributed minimization problem
\begin{gather*}
\min J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2\\
\text{s.t.} (\nabla u,\nabla \phi) = (f(q),\phi)\;\;\forall\,\phi \in H^1_0(\Omega; \mathbb R^2)
\end{gather*}
on the domain $\Omega = [0,1]^2$, and the data is chosen as follows:
\begin{align*}
 f(q) &= q_0 \left(\begin{matrix}2\pi^2  \sin( \pi x) \sin(\pi y)\\0 \end{matrix}\right)\\
      &+ q_1 \left(\begin{matrix}5\pi^2  \sin( \pi x) \sin(2\pi y)\\0 \end{matrix}\right)\\
      &+ q_2 \left(\begin{matrix}0 \\8\pi^2  \sin(2\pi x) \sin(2\pi y)\end{matrix}\right)\\
\end{align*}
with $\alpha = 0$ and $u^d$ such that the solution is given by:
\begin{gather*}
 \overline{q} = \bigl(1;0.5;1\bigr)\\
 \overline{u} = \left(\begin{matrix} \sin( \pi x)( \sin(\pi y)+0.5\sin(2\pi y))\\\sin(2\pi x) \sin(2\pi y) \end{matrix}\right)
\end{gather*}

### 2012-03-15 Examples/OPT/StatPDE/Example3/content.tex
Initial Transfer From OLD CVS Reprository

In this example, we solve a optrimization problems from 
fluid dynamcis. The configuration is similar to 
the fluid optimization problem proposed by Roland Becker 
in 2000 ``Mesh adaption for stationary flow control''.

We describe the underlying equations later...

### 2012-03-15 Examples/OPT/StatPDE/Example4/content.tex
Initial Transfer From OLD CVS Reprository

This example solves the minimization problem
\begin{gather*}
\min J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2\\
\text{s.t. } (\nabla u,\nabla \phi) = (f,\phi)\;\;\forall\,\phi \in H^1_0(\Omega; \mathbb R^2)\\
\end{gather*}
on the domain $\Omega = [0,1]^2$. In addition we set the dirichlet data of the state on the boundary as follows
\begin{gather*}
 u_0(0,y) = q_0,\quad u_0(1,y) = q_1,\quad u_0(x,0) = q_2,\quad u_0(x,1) = q_3,\\
 u_1 = q_4^3.
\end{gather*}
The data is chosen as follows:
\begin{align*}
 f &= \left(\begin{matrix}20\pi^2  \sin( \pi x) \sin(\pi y)\\1 \end{matrix}\right)\\
 u^d&= \left(\begin{matrix}\sin( \pi x) \sin(\pi y)*x\\x \end{matrix}\right)
\end{align*}
with $\alpha = 10$.

### 2012-03-15 Examples/OPT/StatPDE/Example6/content.tex
Initial Transfer From OLD CVS Reprository

This example implements the minimum compliance problem for the thickness optimization 
of an MBB-Beam. Using the MMA-Method of K. Svanberg together with an augmented Lagrangian 
approach for the subproblems following M. Stingl.

### 2012-03-15 Examples/OPT/StatPDE/Example7/content.tex
Initial Transfer From OLD CVS Reprository

This example solves the distributed minimization problem
\begin{gather*}
\min J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2\\
\text{s.t.} (\nabla u,\nabla \phi) = (q+f,\phi)\;\;\forall\,\phi \in H^1_0(\Omega)\\
\text{s.t.} -500 \le q \le 500\;\text{a.e. in }\Omega
\end{gather*}
on the domain $\Omega = [0,1]^2$, and the data is chosen as follows:
\begin{gather*}
 f = \left(20\pi^2  \sin(4 \pi x) - \frac{1}{\alpha}  \sin(\pi x)\right) \sin(2 \pi y)\\
 u^d = \left( 5 \pi^2 \sin(\pi x) + \sin(4 \pi x)\right)  \sin(2\pi y)
\end{gather*}
and $\alpha = 10^{-3}$.

In addition the following functionals are evaluated:
\begin{gather*}
  \text{MidPoint: } u(0.5 ; 0.5)\\[2mm]
  \text{MeanValue: }\int_\Omega u
\end{gather*}

The Problem is similar to that of {\tt OPT/StatPDE/Example1} except for the box control constraints.
Another new feature is the use of the commercial optimization library SNOPT. 
In order to use this library you need to install SNOPT on your computer and then generate a symlink to 
the snopt directory (where you have the libs and the header files) 
in the {\tt DOpE/ThirdPartyLibs} directory named {\tt snopt}, i.e., you should have the file
{\tt DOpE/ThirdPartyLibs/snopt} pointing to the snopt directory. If you have not done this you can compile the
example but when running the example you will only get an error message like\\
{\tt Warning: During execution of `Reduced\underline{ }SnoptAlgorithm::Solve`\\ the following Problem occurred!\\
To use this algorithm you need to have SNOPT installed! \\To use this set the WITH\underline{ }SNOP CompilerFlag.}

### 2012-03-15 Examples/OPT/StatPDE/Example8/content.tex
Initial Transfer From OLD CVS Reprository

This example implements the topology optimization of an MBB-Beam given in \\
{\tt OPT/StatPDE/Example6} using the SIMP method.

The solution is computed using the commercial optimization library SNOPT as in 
{\tt OPT/StatPDE/Example7}. This Example demonstrate how global constraints on
the control variable can be included into the optimization call.

### 2012-03-15 Examples/PDE/InstatPDE/Example1/content.tex
Initial Transfer From OLD CVS Reprository

In this example we consider the nonstationary incompressible Stokes equation.

\textbf{General problem description}\\
As in the stationary PDE Example 3 \ref{PDE_Stat_FSI_STVK},
we use the symmetric fluid stress tensor.  
In strong formulation we deal with 
\begin{align*}
\partial_t v -\nabla\cdot (\nabla v + \nabla v^{T}) + \nabla p &= f \\
div\, v &= 0 
\end{align*}
on time interval $I=[0,T]$ and the domain 
$\Omega = [-6,6]\times [0,2]$. For simplicity,  we set $f=0$.

As introduced earlier, 
we formulate the time stepping scheme as \textit{One-step-$\theta$ scheme},
which are based on finite difference schemes. 
The time interval is given by $I=[0,T]$. Let $v^n,p^n$ and the time
step $k=t^{n+1}-t^n$ be given. Find $v=v^{n+1}, p^{n+1}$ such that:

\begin{align*}
v - k\theta (\nabla\cdot (\nabla v + \nabla v^{T}) + \nabla p) &=
k\theta f^{n+1} + k(1-\theta)f^{n+1}\\
&+  v^n + k(1-\theta) (\nabla\cdot (\nabla v^n + \nabla (v^n)^{T}) 
+ \nabla p^n) \\
div \,v &= 0 
\end{align*}
In the case of the BE-scheme, $\theta = 1$, and the equation is reduced to
\begin{align*}
v - k\theta (\nabla\cdot (\nabla v + \nabla v^{T}) + \nabla p) &=
k\theta f^{n+1} +  v^n  \\
div \,v &= 0
\end{align*}
Note, that one should prefer a complete implicit treatment of the
pressure $p$. Instead of using $\theta p^{n+1} + (1-\theta)p^n$, the pressure
appears only with $\theta p^{n+1}$.

After discretization in time, the space is treated, as ususally, with 
a Galerkin finite element scheme, here based on the Taylor-Hood element 
$Q_2^c / Q_1^c$.

The variational formulation reads:

\begin{Problem}[Backward Euler (BE) timestepping problem]
Find $v:= v^{n+1}\in V$ and $p:= p^{n+1}\in L$:
\begin{align*}
(v,\phi^v) + k\theta (\nabla v + \nabla v^{T}, \nabla\phi^v) 
- k  (p, \nabla\cdot\phi^v) &=
k\theta (f^{n+1},\phi^v) +  (v^n,\phi^v)\\
(div \,v,\phi^p) = 0  
\end{align*}
\end{Problem}
for all suitable test functions ${\phi^v , \phi^p} \in V\times L$. 
The parameter $\theta$ is chosen as $\theta =1$.

Derivation of the other timestepping problems is analogous.

\textbf{Specific features for 
solving nonstationary problems}\\

In the following, we explain in more detail the 
different member functions that are required to
implement nonstationary equations.

\begin{verbatim}
void CellEquation (..., double scale, double scale_ico)
\end{verbatim}
The two arguments 
are used to distigunish between explicit 
components and fully implicit components. 
For standard equations (such as the heat 
equation and the wave equation), there is 
no special treatment required needed.

However, solving the Navier-Stokes equations are 
multi-physics problems (like fluid-structure
interaction), parts of the equations are treated
with a fully implicit time-stepping scheme.

Thus, the argument
\begin{verbatim}
double scale
\end{verbatim}
is used to indicate that the present term 
can be used for implicit/explicit or mixed 
discretization (such as time discretization with
the Crank-Nicolson.

The other argument
\begin{verbatim}
double scale_ico
\end{verbatim} 
is used to indicate that the present term 
only is treated in a fully implicit manner. 
For example, the pressure term (which is of course
a Lagrange multiplier of the incompressibility 
term of the fluid). It is recommended to treat 
this term in a time discretization in a fully 
implicit manner.

\begin{verbatim}
void CellMatrix (..., double scale, double scale_ico)
\end{verbatim}
The directional derivatives of the state equation
are implemented in the present function. As before,
the last two parameters
\begin{verbatim}
double scale, double scale_ico
\end{verbatim} 
are used to distigungish between fully implicit 
are other behavior.

\begin{verbatim}
void CellTimeEquation (...) 
\end{verbatim}
This function is used to implement the 
time derivative in weak formulation 
\begin{equation*}
(\partial_t v, \phi)_{\Omega}.
\end{equation*}
This term is time discretized via
\begin{equation*}
k^{-1} (v^n - v^{n-1} , \phi)_{\Omega}. 
\end{equation*}
Here, it suffices to implement the term
\begin{equation*}
(v^n, \phi)_{\Omega},
\end{equation*}
because the already known term $v^{n-1}$ is 
automatically treated by the specific time
stepping scheme.

In contrast to this behavior, the user 
has the possibility to write all terms of 
$\partial_t v$ explicitly. In this case, we 
use the 
\begin{verbatim}
void CellTimeEquationExplicit (...) 
\end{verbatim}
and we write 
\begin{equation*}
(v^n - v^{n-1}, \phi)_{\Omega}.
\end{equation*}
This behavior is useful for multi-phyics problems
where other solutions variables have to be considered
around $\partial_t v$. The user should have a look 
in the second Example 
\ref{PDE_Instat_FSI} for nonstationary problems 
for an illustration
of this function.

Consequently, the directional derivatives of the 
cell terms are implemented in the corresponding
matrix functions, i.e., 
\begin{verbatim}
void CellTimeMatrix (...), void CellTimeMatrixExplicit 
\end{verbatim}

### 2012-03-15 Examples/PDE/InstatPDE/Example2/content.tex
Initial Transfer From OLD CVS Reprository

\textbf{General problem description}

In the present example, we solve a 
nonstationary fluid-structure interaction problem. 
The underlying equations are stated in the following:
\begin{Problem}[Variational fluid-structure interaction framework]
  \label{eq:fsi:ale:harmonic}
  Find $\{\hat v,\hat u,\hat p\} \in \{ \hat v^D + \hat {\cal V}^0\} 
\times \{ \hat u^D + \hat {\cal V}^0 \}\times \hat {\cal L}$, 
  such that $\hat v (0) = \hat v^0$ and $\hat u(0) = \hat u^0$, for almost all
  time steps $t$, and
  \begin{eqnarray*}
    \begin{aligned}
      (\hat J\hat\rho_f  \partial_t \hat v,\hat\psi^v)_{\hat\Omega_f}  
      +(\hat\rho_f \hat J  (\hat F^{-1}(\hat
      v-\partial_t \hat u)\cdot\hat\nabla) \hat v),
      \hat\psi^v)_{\hat\Omega_f} &\\
      + (\hat J\hat\sigma_f\hat
      F^{-T},\hat\nabla\hat\psi^v)_{\hat\Omega_f}
      %
      + (\hat\rho_s \partial_t \hat v,\hat\psi^v)_{\Omega_s}  
      + (\hat J\hat\sigma_s\hat F^{-T},\hat\nabla\hat\psi^v)_{\hat\Omega_s}&\\
      - \langle \hat g, \hat\psi^v \rangle_{\hat\Gamma_N} -
      (\hat\rho_f \hat J\hat f_f, \hat\psi^v)_{\hat\Omega_f}
      - (\hat\rho_s\hat f_s, \hat\psi^v)_{\hat\Omega_s}
      &=0&&\forall\hat\psi^v\in \hat V^0,
      \\
      %%%%%%%%%%%% 
      (\partial_t\hat u-\hat v,\hat\psi^u)_{\hat\Omega_s} 
      + 
      (\hat\sigma_g ,\hat \nabla\hat\psi^u)_{\hat\Omega_f}
      -\langle \hat\sigma_g \hat n_f ,\hat\psi^u\rangle_{\hat \Gamma_i}
      &=0&&\forall\hat\psi^u\in \hat V^0,\\
      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
      (\widehat{\text{div}}\,(\hat J\hat F^{-1}
      \hat v_f),\hat\psi^p)_{\hat\Omega_f} 
      + (\hat p_s ,\hat \psi^p)_{\hat\Omega_s}
      &=0&&\forall\hat\psi^p\in \hat L,
    \end{aligned}
  \end{eqnarray*}  
  with $\hat\rho_f$, $\hat\rho_s$, $\nu_f$, $\mu_s$, $\lambda_s$, $\hat F$, and $\hat J$. 
  The stress tensors for the fluid and structure are
implemented in  
  $\hat\sigma_f$, $\hat\sigma_s$, and $\hat\sigma_g$ 
\end{Problem}

{\bf Code validation for fluid problems with ALE}

With the ALE code implemented in Example 
\ref{PDE_Stat_FSI_STVK} it is possible to treat fluid 
problems as well as FSI computations. 
In the case of fluid problems the deformation
gradient and its determinant become:
\begin{equation*}
F:= I , \quad \det F = J = 1.
\end{equation*}

The code is validated by the well-known 
fluid- and FSI benchmark problems. Here, 
the results have been summarized in the table below.

 \begin{table}[h]
   \small
   \centering
     \begin{tabular}{llllcccc}    
       \hline
       Configuration & Date         & Refinement & Time dis. & $k$ & $\Delta p$& $F_D$ & $F_L$   \\ \hline\hline
       BFAC 2D-1     & Mar 10, 2010 & 2 global   & BE        & 1.0    & 0.117412  &       &         \\
       BFAC 2D-1     & Mar 10, 2010 & 2 global   & FS        & 1.0    & 0.117412  &       &         \\
       BFAC 2D-1     & Mar 10, 2010 & 2 global   & CN(k)     & 1.0    & 0.117412  &       &         \\ 
       BFAC 2D-1     & Mar 10, 2010 & 2 global   & CN        & 1.0    & 0.117383  &       &         \\ 
       BFAC 2D-2     & Mar 10, 2010 & 2 global   & CN(k)     & 1.0    & 2.31541   &       &         \\ 
       BFAC 2D-2     & Mar 10, 2010 & 3 global   & CN(k)     & 1.0    & 2.32011   &       &         \\   
       BFAC 2D-2     & Mar 12, 2010 & 3 global   & CN(k)     & 1.0e-2 & 2.50288   &       &         \\   
     \end{tabular}
  \end{table}

\subsubsection{Code validation for FSI with ALE}

The results are summarized below:

 \begin{table}[h]
   \small
   \centering
     \begin{tabular}{llllccccc}    
       \hline
       Config. & Date & Refinement & Time dis. & $k$ &  $u_x(A) [\times 10^{-5}]$ &$u_y(A) [\times 10^{-4}]$ & $F_D$ & $F_L$  \\ \hline
       FSI 1   & Mar 12, 2010 & 3 global & BE   & 1.0 & 8.2003  & 2.2732  & &        \\
       FSI 1   & Apr 08, 2010 & 2 global & BE   & 1.0 & 8.2258  & 2.2813  & &        \\
       FSI 1   & Apr 08, 2010 & 2 global & CN(k)& 0.5 & 8.2268  & 2.2813  & &        \\
       FSI 1   & Apr 08, 2010 & 2 global & CN   & 0.5 & 8.2268  & 2.2813  & &        \\
     \end{tabular}
  \end{table}

\vspace{0.2cm}

\textbf{Program description}

\vspace{0.2cm}

### 2012-03-15 Examples/PDE/InstatPDE/Example3/content.tex
Initial Transfer From OLD CVS Reprository

\textbf{General problem description}

\vspace{0.2cm}

The problem under consideration is the so called multivariate Black-Scholes equation arising from pricing European style options in finance. \index{instationary PDE} \index{instationary PDE!Black-Scholes equation}\\
To state the general form of the equation we need some nomenclature: We consider an option on $d$ risky assets with {\it maturity} $T>0$ and {\it strikeprice} $K>0$. For the sake of simplicity we assume the {\it interest rate} $r>0$ and the {\it volatility} of the i-th asset $\sigma_i>0$, $1\leq i\leq d$, to be constant. Besides, we assume the matrix $\rho=(\rho_{ij})$ of the {\it correlation factors} $\rho_{ij}$ with  $-1\leq \rho_{ij}\leq 1$ for  $1\leq i,j \leq d$, to be positive definite. Of course $\rho$ is symmetric with $\rho_{ii}=1$.\\ 
With $(t,x) \in I=(0,T] \times \mathbb{R}^d_+$ denoting the prices of the underlying assets at time $t$, the problem of determining the fair price $u$ of such an option is (after a time reversal) given by the following equation:
\begin{subequations}
\begin{align} 
   \partial_t{u}-\frac12\sum_{i,j=1}^d \sigma_i\sigma_j\rho_{ij}x_ix_j \partial_{{x_i}}\partial_{x_j}u-r\sum_{i=1}^dx_i\partial_{x_i}{u}+ru&=0 &&\mbox{in } (0,T]\times \mathbb{R}_+^d,\\
   u(0)&=u_0 &&\mbox{in } \mathbb{R}_+^d.
\end{align}
\end{subequations}
The initial condition $u_0\in C^{0}\left({\mathbb{R}_+^d}\right) $ (i.e. the {\it payoff}) is given depending of the type of the option. For example
\begin{equation}
u_0 :=\begin{cases} 
\max(\sum_{i=1}^d \lambda_i x_i-K,0),&  u\text{ is a \textit{Call},}\\
\max(K-\sum_{i=1}^d\lambda_i x_i,0),& u\text{ is a \textit{Put},}
\end{cases}
\end{equation}  
for a plain vanilla European option on a basket of assets containing a share of $0<\lambda_i\leq 1$ of the $i$-th asset. For the computation, we truncate the domain, i.e. we choose $\overline x \in \mathbb{R}^d_+$ and consider the computational domain $\Omega := (x_1, \overline x_1)\times \dots \times (x_d, \overline x_d)$. On the new part of the boundary $\Gamma$ with $\Gamma :=\{x\in \partial\Omega |\exists_{1\leq i\leq d}  x_i=\overline x_i\}$) we impose asymptotic values as dirichlet conditions. For a put, we take $u|_\Gamma=0$. We emphasize that no boundary conditions will be imposed on $\partial \Omega \setminus \Gamma$.\\
In this particular example we examine the case of two uncorrelated stocks (with $\lambda_1=\lambda_2=\frac12$) and the following parameters:
 \begin{table}[hb]
    \centering
    \begin{tabular}{lr}
      \toprule 
 & 2d-Put)\\
       \cmidrule(l){2-2}
	 actual asset value $x_0$ 	&  (25,25)\\
	 strikeprice $K$			&  25 \\
	 maturity date $T$ 			& 1 \\
	 volatility $\sigma$		&  ($\frac12$, $\frac 3{10}$) \\
	 cutoff $\overline x$		& $(100,100)$\\
	 interest rate $r$ 				&  $0{,}05$\\
	 option value $u(T,x_0$)			& ca.  2{,}269172389 \\
      \bottomrule
    \end{tabular}
  \end{table}

\vspace{0.2cm}

\textbf{Program description}

\vspace{0.2cm}

### 2012-03-15 Examples/PDE/InstatPDE/Example4/content.tex
Initial Transfer From OLD CVS Reprository

\textbf{General problem description}

\vspace{0.2cm}

In this example we consider one of the prototypical instationary equations, the parabolic heat equation
\begin{align*}
\partial_t u(t,x) - \Delta u(t,x) &= f(t,x),\\
u(t,x)|_{\partial \Omega} &= g(t,x),\\
u(0,x) &= u_0(x)
\end{align*}
with unknown solution $u:\Omega \rightarrow \mathbb{R}$, where $\Omega \subset \mathbb{R}^d$. In our example, we consider the simplest case $d=1$, where the Laplacian $\Delta$ reduces to $\partial_x^2$. The computational domain is $\Omega \times I = [0,1] \times [0,1]$. For further simplification, we choose the righthand side as $f=0$ as well as homogeneous Dirichlet boundary conditions ($g=0$). The initial condition is given by $u_0(x) = \min(x,1-x)$.
\vspace{0.2cm}

\textbf{Program description}

\vspace{0.2cm}

There are few new things compared to the first two nonstationary examples. This is the first time we solve an equation in one spatial dimension. In most cases, the dimension dependence is covered by the \texttt{LOCALDOPEDIM} and \texttt{LOCALDEALDIM} variables (which are defined at the beginning of the \textit{main.cc} file), but there might be some places in the code (especially your own code) where a concrete dimension number is given to an object. There you have to replace it manually. Do not forget to insert the correct dimension in the \textit{Makefile}!\\
The most important feature of this example is the serial application of several time-stepping schemes. At the moment, the following schemes are available (see also example \ref{PDE_Instat_Stokes}):
\begin{enumerate}
\item
Forward Euler scheme (FE)
\item
Backward Euler scheme (BE)
\item
Crank-Nicolson scheme (CN)
\item
shifted Crank-Nicolson scheme (sCN)
\item
Fractional-Step-$\theta$ scheme (FS)
\end{enumerate}
All these time-stepping methods are applied in the current example in order to check them and to compare their characteristics. To keep the computing time acceptable, we choose a one dimensional example.\\
One more innovation is the output format. We want to represent the output at single timepoints as a function graph on the space interval $[0,1]$; this can be done using \textit{GNUPLOT}, for example, so instead of \textit{.vtk} files as in all former examples, we now write out \textit{.gpl} files.

### 2012-03-15 Examples/PDE/InstatPDE/Example5/content.tex
Initial Transfer From OLD CVS Reprository

\textbf{General problem description}

\vspace{0.2cm}

This example differs only slightly from the previous one. Again, we consider the heat equation, this time with an additional nonlinear term
\begin{equation*}
\partial_t u(t,x,y) - \Delta u(t,x,y) + u(t,x,y)^2 = f(t,x,y),
\end{equation*}
but now in two space dimensions and with known solution 
\begin{equation*}
u(t,x,y) = e^{t-t^2} \sin(x) \sin(y).
\end{equation*}
The computational domain is $\Omega \times I = [0,\pi]^2 \times [0,1]$. From the known solution, we can compute the appropriate data 
\begin{align*}
f(t,x,y) &= (3-2t)e^{t-t^2} \sin(x) \sin(y) + e^{(t-t^2)^2} \sin^2(x) \sin^2(y),\\
u_0(x,y) &= \sin(x) \sin(y).
\end{align*}
Furthermore, we have to prescribe homogeneous Dirichlet boundary conditions.

\vspace{0.2cm}

\textbf{Program description}

\vspace{0.2cm}

The new feature of this example is the nonhomogeneous righthand side. In examples \ref{PDE_Stat_Laplace_2D} and \ref{PDE_Stat_Laplace_3D}, we regarded stationary problems with nonhomogeneous righthand sides, but up to now, we never involved the time variable into the nonhomogeneity. To do this, DOpE yields a \texttt{SetTime()} function which has to be applied in the \textit{localpde.h} file as well as at the place where the \texttt{RightHandSideFunction} class is declared (here the \textit{myfunctions.h} file.

### 2012-03-15 Examples/PDE/InstatPDE/content.tex
Initial Transfer From OLD CVS Reprository

Until now, the DOpE provides various time-stepping schemes
that are based on finite differences. Specifically, the 
user can choose between the 
\begin{itemize}
\item Forward Euler scheme (FE), which is an explicit timestepping scheme. Here, 
one as to take into account that $k\leq ch^2$ where $k$ denotes the timestep size
and $h$ the local mesh cell diameter.
\item Backward Euler scheme (BE), which is an implicit timestepping scheme.
It is strongly A-stable but only from first order and very dissipative. 
The BE-scheme is well suited for stationary numerical examples.  
\item Crank-Nicolson scheme (CN), which is of second order, 
A-stable, has very little dissipation but suffers from case to case from instabilities caused by rough
initial- and/or boundary data. These properties are due to weak stability (it is not \textit{strongly}
A-stable).
\item Shifted (or stabilized) Crank-Nicolson scheme (CN shifted), which is
  also of second order, but provides global stability.
\item Fractional-step-$\theta$ scheme (FS). 
It has second-order accuracy and is strongly A-stable, and therefore
well-suited for computing solutions with rough data. 
\end{itemize}

### 2012-03-15 Examples/PDE/StatPDE/Example1/content.tex
Initial Transfer From OLD CVS Reprository

\textbf{General problem description}\\

\vspace{0.2cm}
\index{stationary PDE}
In this example we consider the stationary incompressible Stokes equation \index{stationary PDE!Stokes equation}. Here,
we use the symmetric stress tensor which has a little consequence when using 
the do-nothing outflow condition. In strong formulation we have
\begin{align} \label{eq:strong}
-\frac{1}{2}\nabla\cdot (\nabla v + \nabla v^{T}) + \nabla p &= f \\ 
\nabla \cdot v &= 0 \notag
\end{align} 
on the domain $\Omega = [-6,6]\times [0,2]$. We choose for simplicity $f=0$.\\

\vspace{0.2cm}

\textbf{Program structure}\\

\vspace{0.2cm}

In all examples, the whole program is split up into several files for the sake of readability. These files are always denoted in the same way, so we only have to explain the general structure in this first example, whereas in the following examples, we will only point out differences to the current one. The content of the single files will be described in more detail below.\\

If we do not use one of the standard grids given in the deal.II library, we can read a grid from an input file. In our example, the domain $\Omega = [-6,6]\times [0,2]$ is given in the \textit{channel.inp} file, where all nodes, cells and boundary lines are listed explicitly and the boundary is divided into disjoint parts by attributing different colors to the boundary lines.\index{grid}\\

Certain parameters occurring during the solution process, e.g. error tolerances or the maximum number of iterations in an iterative solution procedure, are fixed in a parameter file called \textit{dope.prm}. This parameter file comprises several subsections corresponding to different solver components.\index{parameters}\\

In the \textit{functionals.h} file we declare classes for different scalar quantities of interest (described mathematically as functionals) which we want to evaluate during the solution process.\index{functional}\\

The \textit{localfunctional.h} file is relevant only if we want to solve an optimal control problem. In this case, it contains the cost functional, whereas the file is not needed for the forward solution of PDEs. We will get back to this later in the context of optimal control problems.\index{cost functional}\\

All information about the PDE problem (in the optimal control case about the constraining PDE) is included in the \textit{localpde.h} file. In a class called \texttt{LocalPDE}, we build up the cell equation, the cell matrix and cell righthand side as well as the boundary equation, boundary matrix and boundary righthand side. Later on, the integrator collects this local information and creates the global vectors and matrices.\\

The most important part of each example is the \textit{main.cc} file which contains the \texttt{int main()} function. Here we create objects of all classes described above and actually solve the respective problem.\\

\vspace{0.2cm}

\textbf{The} functionals.h \textbf{file}\\

\vspace{0.2cm}
\index{functional evaluation}
Here, we declare all quantities of interest (functionals), e.g. point values, drag, lift, mean values of certain quantities over a subdomain etc. \\
Each of these functionals is declared as a class of its own, but in DOpE all classes are derived from a so-called \texttt{FunctionalInterface} class.\\
In the current example we declare functionals for point values of the velocity and for the flux at the outflow boundary of the channel.\index{functional!point value} \index{functional!boundary flux}\\

\vspace{0.2cm}

\textbf{The} localpde.h \textbf{file}\\

\vspace{0.2cm}

The \textit{LocalPDE} is derived from a \texttt{PDEInterface} class. It comprises several functions which build up the cell and boundary equations, matrices and righthand sides. The weak formulation of problem \eqref{eq:strong} with $f=0$ is
\begin{equation} \label{eq:weak}
   \frac{1}{2}(\nabla v, \nabla \varphi) + \frac{1}{2}(\nabla v^{T}, \nabla \varphi) - (p, \nabla \cdot \varphi) + (\nabla \cdot v, \psi) = 0.
\end{equation}
This problem is vector valued, i.e. the velocity variable $v$ has two components and the pressure variable $p$ is a scalar. For the implementation, we use a vector valued solution variable with three components, where the distinction between velocity and pressure is done by use of the deal.II \texttt{FEValuesExtractors} class. \index{vector-valued problem}\\
Furthermore, in DOpE we always interpret the problems in the context of a Newton method. Usually, a PDE in its weak formulation is given as
\begin{equation*}
   a(u;\varphi) = f(\varphi).
\end{equation*}
The lefthand side is implemented in the \texttt{CellEquation} function, the righthand side is implemented in the \texttt{CellRightHandSide} function (which is unused in this example, because $f=0$).\\
To apply Newton's method, this problem is linearized: on the lefthand side, we have the derivative of the (semilinear) form $a(\cdot;\cdot)$ with respect to the solution variable $u$, and the righthand side is the residual of the weak formulation:
\begin{equation*}
   a_u'(u;u^+,\varphi) = -a(u;\varphi) + f(\varphi).
\end{equation*}
In the \texttt{CellMatrix} function, we implement the following matrix $A$ as representation of the derivative on the lefthand side:
\begin{equation*}
  A = (a_u'(u;\varphi_i,\varphi_j))_{j,i=1}^N
\end{equation*}
with the number $N$ of the degrees of freedom. Similarly, the \texttt{CellEquation} contains the vector
\begin{equation*}
  a = a(u;\varphi_i)_{i=1}^{N},
\end{equation*}
and the \texttt{CellRightHandSide} in the case $f \neq 0$ would contain a vector
\begin{equation*}
  \tilde{f} = (f;\varphi_i)_{i=1}^N.
\end{equation*}
The system of equations which is then actually solved is
\begin{equation*}
  A\tilde{u}^+ = -a + \tilde{f}.
\end{equation*}
Because of the linearity of equation \eqref{eq:weak}, there is almost no difference between the two functions.\index{Newton's method}\\

\vspace{0.1cm}

\textit{At this point, it is important to note that DOpE interprets any given problem as a nonlinear one which is solved by Newton's method; the special case of linear problems is included into this general framework.}\\

\vspace{0.2cm}

\textbf{The} main.cc \textbf{file}\\

\vspace{0.2cm}

First of all, several header files have to be included that are needed during the solution process. We divide these includes into blocks corresponding to DOpE headers, deal.II headers, C++ headers and header files of the example itself (like the ones mentioned above).\\
Furthermore, we define names for certain objects via \texttt{typedef} which act as abbreviations in order to keep the code readable. In our case, these are \texttt{OP, INTEGRATOR, LINEARSOLVER, NLS} and \texttt{SSOLVER}.\\
In the \texttt{int main()} function, we first create a possibility to read the parameter values from the \textit{dope.prm} file. Then there are several standard steps for finite element codes like
\begin{itemize}
\item
definition of a triangulation and create a grid object (which we read from the \textit{channel.inp} file)
\item
creation of finite element objects for the state and the control and of quadrature formula objects
\end{itemize}
and in addition, we 
\begin{itemize}
\item
create objects of the \texttt{LocalPDE} class and of the different functional classes declared in the \textit{functionals.h} file (including an object of the \texttt{LocalFunctional} class, even if we do not regard an optimal control problem here).
\end{itemize}
\textit{REMARK:}\\
Up to now  we have to create a pseudo time even for stationary problems. The \\\texttt{MethodOfLines\underline{ }StateSpaceTimeHandler} object (\texttt{DOFH}) which is needed for the initialization of \texttt{OP} requires a vector in which timepoints are specified. However, this is again merely a dummy variable, for we do not actually apply a time stepping method in the stationary case. This will also be removed in future versions of DOpE.\\

Before we initialize the \texttt{SSolver} object and actually solve the problem, we have to set the correct boundary conditions. Via the \texttt{compmask} vector, we ensure that the boundary conditions are set only for the velocity components of our solution vector. We set homogeneous Dirichlet values at the upper and lower boundaries of the channel. The inflow is described by a parabolic profile at the left boundary (the corresponding function class is declared in the \textit{myfunctions.cc} file), whereas we do not prescribe anything at the outflow boundary (so-called do-nothing condition).\\
The output of the program (the two functional values) is rather unspectacular; as the problem is linear, the solution is computed within one Newton step.

### 2012-03-15 Examples/PDE/StatPDE/Example2/content.tex
Initial Transfer From OLD CVS Reprository

\textbf{General problem description}

\vspace{0.2cm}

In this example we consider a simple stationary FSI problem. \index{stationary PDE} \index{stationary PDE!FSI problem} The fluid
is given as an incompressible Newtonian fluid modelled by the Stokes equation. Here, we use the symmetric stress tensor which has a little consequence when using the do-nothing outflow condition. \index{fluid} \index{fluid!incompressible} \index{fluid!Newtonian}
The computation domain is $\Omega = [-6,6]\times [0,2]$ and we choose for simplicity $f=0$.

The fluid reads: \index{fluid!Eulerian framework}
\begin{Problem}[Variational fluid problem, Eulerian framework]
 Find $\{v_f,p_f\} \in \{ v_f^D + V\} \times L_f$, such that,
  \begin{equation*}
    \begin{aligned} 
      (\rho_f v\cdot\nabla v_f,\phi^v)_{\Omega_f} +
      (\sigma_f,\nabla\phi^v)_{\Omega_f} 
      &= \langle n_f\cdot g_s^\sigma ,\phi^v \rangle_{\Gamma_i}
      &&\forall \phi^v\in V_f, \\
      %%%
      (\operatorname{div}\, v_f,\phi^p)_{\Omega_f} &= 0
      &&\forall\phi^p\in L_f.
    \end{aligned}
  \end{equation*}
  The Cauchy stress tensor $\sigma_f$ is given by \index{Cauchy stress tensor}
  \begin{eqnarray}
    \sigma_f:=-p_f I + \rho_f\nu_f (\nabla v_f+\nabla v_f^T),
  \end{eqnarray}
  with the fluid's density $\rho_f$  and the kinematic viscosity
  $\nu_f$. By $n_f$ we denote the outer normal vector on $\Gamma_i$
  and by $g^\sigma_f$ is a function which describes forces acting on the interface. These
  will be specified in the context of fluid-structure interaction
  models. \index{fluid-structure interaction (FSI)} \index{fluid-structure interaction (FSI)!interface}
\end{Problem}
We define:
\begin{equation*}
\hat T:=\text{id}+\hat u,\quad
\hat F:=I+\hat\nabla \hat u,\quad
\hat J:=\text{det}(I+\hat\nabla \hat u).
\end{equation*}

The structure equations are given by incompressible neo-Hookean material
\index{structure} \index{structure!incompressible neo-Hookean (INH) material} \index{structure!Lagrangian framework}
\begin{Problem}[Incompressible neo-Hookean Model (Lagrangian)]
  \begin{equation*}
    \begin{aligned}
      (\hat J_s \hat\sigma_s \hat F_s^{-T},\hat\nabla\hat\phi^v)_{\hat\Omega_s}
      &= \langle\hat J_s\hat n_s\cdot \hat g_s^\sigma\hat
      F_s^{-T},\hat\phi^v  \rangle_{\hat\Gamma_i}
      &&\quad\forall\hat \phi^v\in\hat V_s\\
      %%%%%%%%%%%%%%%% 
      (\hat v_s,\hat\phi^u)_{\hat\Omega_s}&=0
      &&\quad\forall\hat \phi^u\in\hat V_s    ,\\
      (\text{det}\,\hat F_s-1,\hat\phi^p)_{\hat\Omega_s} &=0
      &&\quad\forall\hat \phi^p\in\hat L_s,\\
    \end{aligned}
  \end{equation*}
  where $\rho_s$ is the solid's density, $\mu_s$ the Lam\'e
  coefficient, $\hat n_s$ the outer normal vector at $\hat\Gamma_i$, 
  $\hat g_s^\sigma$ the force on the interface and with  
  \[
  \hat\sigma_s:=-\hat p_sI+\mu_s(\hat F_s\hat F_s^T-I).
  \]
\end{Problem}

The resulting FSI problem is then given by: \index{fluid-structure interaction (FSI)!ALE model}
\begin{Problem}[Stationary Fluid-Structure Interaction (ALE)]
  \begin{equation*}
    \begin{aligned}
      (\hat J \rho_f \hat F^{-1}\hat  v\cdot\hat\nabla \hat v,
      \hat\phi^v)_{\hat\Omega_f}
      + (\hat J\hat\sigma_f\hat F^{-T},\hat \nabla\hat\phi^v)_{\hat\Omega_f}\\
      + (\hat J\hat\sigma_s\hat F^{-T},\hat \nabla\hat\phi^v)_{\hat\Omega_s}
      &= 0&&\forall\hat\phi^v\in \hat V,\\
      %%%%%%%%%%%% 
      (\hat v,\hat\phi^u)_{\hat\Omega_s}
      + (\alpha_u \hat \nabla \hat u,\hat \nabla\hat\phi^u)_{\hat\Omega_f}
      &=0&&\forall\hat\phi^u\in \hat V,\\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      (\widehat{\text{div}}\,(\hat J\hat F^{-1}\hat
      v_f),\hat\phi^p)_{\hat\Omega_f} 
      + (\hat J - 1,\hat \phi^p)_{\hat\Omega_s}
      &=0&&\forall\hat \phi^p\in \hat L,
    \end{aligned}
  \end{equation*}  
\end{Problem}

\vspace{0.2cm}

\textbf{Program description}

\vspace{0.2cm}

There is not much difference to the prior example, so we will keep our remarks rather short. The \textit{functionals.h, localfunctional.h} and \textit{myfunctions.cc} files did not change at all, so we simply refer to the corresponding sections in the last example.\\
In the \textit{localpde.h} file, all functions of the \texttt{LocalPDE} class have to be adjusted to the current FSI problem. This only makes the equations and matrices a little more complicated, and our solution vector now consists of five components (two velocity components of the fluid, the pressure component, and two additional displacement components for the structure variables). Otherwise, everything is analogous to the former example.\\
In the \textit{main.cc} we only have to add two components to the \texttt{compmask} vector and prescribe boundary conditions for the structure variables. Apart from that, we define objects for the same classes as before that are even named equally and use the same solvers.\\ 
Again, the solution is reached within one Newton step, and all we see from the program output is the values of the functionals.

### 2012-03-15 Examples/PDE/StatPDE/Example3/content.tex
Initial Transfer From OLD CVS Reprository

\textbf{General problem description}

\vspace{0.2cm}

This example is an extension the previous one.
We solve an stationary FSI problem either with 
INH material (see Problem definition before) 
or St. Venant Kirchhoff material STVK: \index{structure} \index{structure!compressible St.Venant-Kirchhoff (STVK) material}
\begin{Problem}[Compressible Saint Venant-Kirchhoff, Lagrangian framework]
  \label{prob:stvk:lagrange}
  Find $\{\hat v_f,\hat u_f\} \in \{ \hat v_f^D + \hat V\} 
  \times \{ \hat u_f^D + \hat V\}$, 
  such that 
  \begin{eqnarray}
    \begin{aligned}   
      (\hat J_s \hat\sigma_s \hat F_s^{-T},\hat\nabla\hat\phi^v)_{\hat\Omega_s}
      &= \langle\hat J_s\hat n_s\cdot \hat g_s^\sigma\hat
      F_s^{-T},\hat\phi^v  \rangle_{\hat\Gamma_i}
      &&\quad\forall\hat \phi^v\in\hat V_s\\
      %%%%%%%%%%%%%%%% 
      (\hat v_s,\hat\phi^u)_{\hat\Omega_s}&=0
      &&\quad\forall\hat \phi^u\in\hat V_s    ,\\
    \end{aligned}
  \end{eqnarray}
  where $\rho_s$ is the density of the structure, $\mu_s$ and $\lambda_s$ 
  the Lam\'e coefficients, $\hat n_s$ the outer normal vector at $\hat\Gamma_i$, 
  $\hat g_s^\sigma$ some forces on the interface.\index{Lam\'e coefficients} The properties of the STVK material
  is specified by the constitutive law
  \begin{eqnarray}
    \hat\sigma_s:=\hat J^{-1} \hat F (\lambda_s (tr\hat E) I + 2\mu_s \hat E)\hat F^{-T}.
  \end{eqnarray}
\end{Problem}
Often, the elasticity properties of structure materials is characterized by 
Poisson's ratio $\nu_s$ ($\nu_s < \frac{1}{2}$ \index{Poisson's ratio} for 
compressible materials) and the Young modulus $E$ \index{Young modulus}. The 
relationship to the Lam\'e coefficients $\mu_s$ and $\lambda_s$ is
given by:
\begin{eqnarray}
\nu_s = \frac{\lambda_s}{2(\lambda_s + \mu_s)}, \qquad 
E = \frac{\mu_s (\lambda_s + 2\mu_s)}{(\lambda_s + \mu_s )} . 
\end{eqnarray}
The whole equation system is solved on the benchmark configuration domain. For 
details on parameters and geometry, we refer to the 
numerical FSI benchmark \index{fluid-structure interaction (FSI)} \index{fluid-structure interaction (FSI)!FSI benchmark} proposal from Hron and Turek [2006].

The code is established by computing the stationary FSI benchmark example FSI 1 with 
the following values of interest: $x$-displacement, $y$-displacement, drag, and lift.\index{functional} \index{functional!drag} \index{functional!lift} \index{functional!deflection}

\vspace{0.2cm}

\textbf{Program description}

\vspace{0.2cm}

Compared to the two former examples, there are some differences which we will briefly discuss in the following. First of all, the problem is nonlinear in contrast to the former ones. We work on a different domain (given in the \textit{benchfst0100tw.inp} file), namely a channel with a cylinder put at half height near the inflow boundary; further \textit{.inp} files yield the possibility to vary the domain.\\
Furthermore, in the \textit{dope.prm} parameter file there are two additional subsections which are added only for the current problem. From the denotation of these subsections one can immediately see where in the code the parameters are used.\\
As we want to compute certain benchmark quantities, we have to regard corresponding functionals in the \textit{functionals.h} file. The pressure at a point as well as the displacement in $x$- and $y$-directions are point values; furthermore we implement the drag and lift functionals (for which we need the additionally defined problem parameters).\\
As before, we build up the cell and boundary equations and matrices in the \textit{localpde.h} file. Apart from using the additionally defined problem parameters and modelling compressible STVK material instead of INH material (which leads to changes in the weak formulation of the equations), there are no major differences to the corresponding file in the last example.\\
In the \textit{main.cc} file, we have to include additional header files from the deal.II library concerning error estimation and grid refinement. Further on, everything is pretty much the same as in the last example, but we have to use the \texttt{SetBoundaryFunctionalColors} function of the \texttt{PDEProblemContainer} class to be able to compute drag and lift in the respective functional classes in \textit{functionals.h}.\\
The main innovation in contrast to the preceding examples is the refinement of the grid combined with a simple error estimator given in the deal.II \texttt{KellyErrorEstimator} class. If we look at the output of our program, everything is computed several times (once on each refinement level). Furthermore, we see that several Newton steps are needed on each refinement level; this is due to the nonlinearity of the current problem.

### 2012-03-15 Examples/PDE/StatPDE/Example4/content.tex
Initial Transfer From OLD CVS Reprository

\textbf{General problem description}

\vspace{0.2cm}

In this example we consider the following benchmark problem from elasticity theory: \index{stationary PDE} \index{stationary PDE!elasticity equations}
\begin{equation}
   (\sigma(u),\varepsilon(\varphi)) = (g,\varphi)_{\Gamma_N}.
\end{equation}
Here $\tilde{\Omega}$ is a quadratic domain with side length $200$ mm, where a circular hole with radius $10$ mm around the center is cut out. Using symmetries of the domain, we restrict our actual computational domain $\Omega$ to the upper left quarter of $\tilde{\Omega}$.\\
In the above equation, $\varepsilon(v) := \frac{1}{2}(\nabla v + \nabla v^T)$ is the symmetric strain tensor, and \index{strain tensor}
\begin{equation*}
   \sigma(v) := 2\mu \varepsilon(v)^D + \kappa \cdot tr (\varepsilon(v)) I
\end{equation*}
denotes the symmetric stress tensor. Here $\tau^D$ is the deviatoric part of a tensor $\tau$, in two dimensions defined as \index{deviator}
\begin{equation*}
   \tau^D := \tau - \frac{1}{2} tr(\tau) I,
\end{equation*}
and the parameters $\mu$ and $\kappa$ are chosen as $\mu = 80193.800283$ resp. $\kappa = 190937.589172$.\\
The corner points of our computational domain are in anticlockwise order: $(0,0), (90,0)$, $(100,10), (100,100)$ and $(0,100)$. We prescribe homogeneous Dirichlet boundary conditions in $y$-direction between $(0,0)$ and $(90,0)$ (lower boundary part), homogeneous Dirichlet boundary conditions in $x$-direction between $(100,10)$ and $(100,100)$ (right boundary part), and we interpret the righthand side of equation $(1)$ with $g = 450$ as a boundary condition between $(0,100)$ and $(100,100)$ (upper boundary part).\\
The goal of our computations is to match the following functional reference values taken from \textit{E. Stein (editor), Error-controlled Adaptive Finite Elements in Solid Mechanics, Wiley (2003), pp. 386 - 387}:
\begin{table}[h]
\centering
\begin{tabular}{lccc}    
\hline
 Functional & $u_1$ at $(90,0)$ & $\sigma_{22}$ at $(90,0)$ & $u_2$ at $(100,100)$\\
\hline 
 Reference value & 0.021290 & 1388.732343 & 0.20951\\
\hline
\end{tabular}
\end{table}
\begin{table}[h]
\centering
\begin{tabular}{lcc}    
\hline
 Functional & $u_1$ at $(0,100)$ & $\int_{(100,100)}^{(0,100)} u_2$\\
\hline 
 Reference value & 0.076758 & 20.40344\\
\hline
\end{tabular}
\end{table}

\vspace{0.2cm}

\textbf{Program description}

\vspace{0.2cm}

From the previous examples we know how to read a grid from an \textit{.inp} file. The grid of our current example comes from the above mentioned benchmark problem. The parameter file (\textit{dope.prm}) contains the same variables as in the first two examples.\\
Apart from different point values of derivatives of the solution, we want to evaluate an integral over part of the boundary. \index{functional} \index{functional!boundary integral} This is newly implemented in \textit{functionals.h}.\\
The (linear) elasticity equation is much simpler than the FSI problems we treated before, which can be seen in \textit{localpde.h} where the \texttt{CellEquation} and \texttt{CellMatrix} are implemented. In principle, everything is clear from the preceding examples.\\
The \textit{main.cc} file does not contain any innovation, either. Here, we refine the grid globally instead of using an error estimator for local refinement.\\
The output of the program reflects again the linearity of the problem (only one Newton step is needed for solution).\\

### 2012-03-15 Examples/PDE/StatPDE/Example5/content.tex
Initial Transfer From OLD CVS Reprository

\textbf{General problem description}

\vspace{0.2cm}

Similar to the previous example, we consider the following benchmark problem from plasticity theory:
\begin{equation}
   (\Pi(\sigma(u)),\varepsilon(\varphi)) = (g,\varphi)_{\Gamma_N}.
\end{equation}
Here $\tilde{\Omega}$ is again the quadratic domain with a circular hole around the center cut out. Again, we restrict our actual computational domain $\Omega$ to the upper left quarter of $\tilde{\Omega}$ for reasons of symmetry.\\
We use the symmetric strain tensor $\varepsilon(v) := \frac{1}{2}(\nabla v + \nabla v^T)$, and the symmetric stress tensor $\sigma$ is defined as
\begin{equation*}
   \sigma(v) := 2\mu \varepsilon(v)^D + \kappa \cdot tr (\varepsilon(v)) I
\end{equation*}
where $\tau^D$ is the deviatoric part of a tensor $\tau$, in two dimensions defined as
\begin{equation*}
   \tau^D := \tau - \frac{1}{2} tr(\tau) I.
\end{equation*}
The main difference with respect to the elastic case is the projection operator $\Pi$ in equation $(1)$. It is defined as follows:
\begin{equation*}
\Pi(\tau) = \left\{
            \begin{array}{lr}
            \tau & |\tau^D| \leq \sigma_0\\
            \sigma_0 |\tau^D|^{-1} \tau^D + \frac{1}{2} tr(\tau) I & |\tau^D| > \sigma_0\\ 
            \end{array}
            \right. 
\end{equation*}
In our computations, we choose $\sigma_0 = \sqrt{\frac{2}{3}}\cdot 450$, and the above parameters $\mu$ and $\kappa$ as $\mu = 80193.800283$ resp. $\kappa = 190937.589172$. The corner points of our computational domain are the same as before, and the boundary conditions are not altered, either.\\
The goal of our computations is to detect a subdomain in $\Omega$ where plastic behaviour occurs (compare \textit{E. Stein (editor), Error-controlled Adaptive Finite Elements in Solid Mechanics, Wiley (2003), pp. 386 - 389}). This subdomain depends on the righthand side $g$ in equation $(1)$ which we write as $g = \lambda \cdot p$ with $p = 100$ and $\lambda \in [1.5;4.5]$.\\

\vspace{0.2cm}

\textbf{Program description}

\vspace{0.2cm}

The code of the current example is nearly identical to the code of the previous one. The only difference worth mentioning is the change of the equations which leads to different implementations of the \texttt{CellEquation, CellMatrix} and \texttt{BoundaryEquations} functions in \textit{localpde.h}.\\
Furthermore, the elasticity equations solved in the last example are linear, whereas the plasticity equations are nonlinear; this difference is evident also from the output (here, we need several Newton steps until convergence).\\
The functionals that appear in the output yield additional information and are not required in the above problem setting. The subdomain with plastic behaviour we want to detect can be visualized from the \textit{.vtk} files written to the \texttt{Results/Mesh} subfolders.

### 2012-03-15 Examples/PDE/StatPDE/Example6/content.tex
Initial Transfer From OLD CVS Reprository

todo. short: this example shows how to use the constraintsmaker class by enforcing periodic boundary conditions in 2d.

### 2012-03-15 Examples/PDE/StatPDE/Example7/content.tex
Initial Transfer From OLD CVS Reprository

In this example we consider the stationary incompressible Stokes equation. Here,
we use the symmetric stress tensor which has a little consequence when using 
the do-nothing outflow condition. In strong formulation we have
\begin{align*}
-\nabla\cdot (\nabla v + \nabla v^{T}) + \nabla p &= f \\
\operatorname{div}\, v &= 0 
\end{align*}
on the domain $\Omega = [-6,6]\times [0,2]$. We choose for simplicity $f=0$.

### 2012-03-15 Examples/PDE/StatPDE/Example8/content.tex
Initial Transfer From OLD CVS Reprository

Laplace in 3D
on locally refined grids with Kelly estimator
with iterative solver (CG without preconditioning)

### 2012-03-15 Examples/description.tex
Initial Transfer From OLD CVS Reprository

\documentclass[a4paper,cleardoubleempty]{scrreprt}
\usepackage{amsmath,amssymb}
\usepackage[pdftex,bookmarks=false,pdfpagelabels=true,plainpages=false]{hyperref}
\usepackage{booktabs}
\usepackage{makeidx}
%\usepackage{sfmath}
\newtheorem{Problem}{Problem}[section]

\newcommand{\todo}[1]{\textbf{\textsc{\textcolor{black}{TODO: #1}}}}

\makeindex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\title{DOpE - Deal Optimization Environment}
\author{
  Michael Geiger\thanks{University of Heidelberg, 
    {michael.geiger@iwr.uni-heidelberg.de}}, 
  Christian Goll\thanks{University of Heidelberg, 
    {christian.goll@iwr.uni-heidelberg.de}}, 
  Thomas Wick\thanks{University of Heidelberg, 
    {thomas.wick@iwr.uni-heidelberg.de}}, 
  Winnifried Wollner\thanks{University of Hamburg, 
    {winnifried.wollner@math.uni-hamburg.de}}
}

\date{\today}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\maketitle
\cleardoublepage\pagestyle{headings}
\tableofcontents
\cleardoublepage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Introduction}
In this report, we describe the DOpE
\textit{Deal Optimization Environment} project. 
Originally, the project was initiated in the year 2009 at the 
University of Heidelberg (Germany) in the numerical analysis
group of Rolf Rannacher.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The \textit{Deal Optimization Environment} (DOpE) project is mainly 
based on the
\textit{deal.II} finite element library which has been developed initially
by W. Bangerth, R. Hartmann, and G. Kanschat. 
In addition, some algorithmic aspects to compute optimization problems
in which the state is given in terms of a partial
differential equations are similar to algorithms of the
Gascoigne/RoDoBo 
project, which was initiated by 
Roland Becker, Dominik Meidner,  and Boris Vexler. 
Both projects have also been 
initiated at the University of Heidelberg.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The aim of DOpE is to provide a software toolkit to solve forward PDE
problems as well as optimal control problems constrained by PDE. The
solution of a broad variety of PDE is possible in deal.II as well, but
DOpE concentrates on a unified approach for both linear and nonlinear
problems by interpreting every PDE problem as nonlinear and applying a
Newton method to solve it. While deal.II leaves much of the work and many
decisions to the user, DOpE intends to be user-optimized by delivering
prefabricated tools which require from the user only adjustments connected
to his specific problem. The solution of optimal control problems with PDE
constraints is an innovation in the DOpE framework.
The focus is on the numerical solution of both stationary and nonstationary
problems which come from different application fields, like elasticity and
plasticity, fluid dynamics, and fluid-structure interactions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
At the present stage, we are able to provide the following features and 
problems:
\begin{itemize}
\item Computation of forward problems in 1d, 2d, and 3d.
\item Solution of stationary and nonstationary forward problems
\item Various timestepping schemes (based on finite differences), 
  such as forward Euler, backward Euler,
  Crank-Nicolson, shifted Crank-Nicolson, and Fractional-Step-$\Theta$ scheme.
\item Various finite elements taken from deal.ii.
\item Computation of stationary optimal control problems.
\item Solving of standard Laplace equation, Navier-Stokes, and 
fluid-structure interaction. 
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the present manual we give a short introduction to all example programs
which can be found currently in DOpE. The organization of the examples is
as follows: we distinguish between the mere solution of a PDE and the
solution of an optimal control problem, and in each case we differentiate
further between stationary and nonstationary problems. This leads to the
following structure (we list problems treated in the examples in
brackets):
\begin{itemize}
\item
solution of PDE problem
 \begin{itemize}
   \item
     stationary PDE (Laplace equation, Stokes equations, FSI problems,
elasticity and plasticity problems)
   \item
     nonstationary PDE (Navier-Stokes equations, FSI problem, Black-Scholes
equation)
 \end{itemize}
\item
solution of PDE constrained optimal control problem
 \begin{itemize}
   \item
     stationary constraining PDE
   \item
     nonstationary constraining PDE
 \end{itemize}
\end{itemize}
The seventh stationary PDE example \ref{PDE_Stat_Laplace_2D} 
(solution of the Laplace equation) is
implemented analogously to the corresponding tutorial step in deal.II. By
comparing the respective program codes, deal.II users should be able to
catch the differences between the two software packages.
In later examples we also provide algorithms for 'goal-oriented' automatic
mesh adaption and error analysis.

{\bf Comment on deal.II}

The deal.II library is open
source and can be downloaded from \texttt{http://www.dealii.org/}. On this
homepage, one also finds lots of further information on deal.II as well as
an extensive tutorial where many features of deal.II are discussed in a
well-documented example framework. In order to use DOpE, it is highly
recommendable to be roughly acquainted with deal.II.

\cleardoublepage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Example Handling, Creating new Examples}
To implement new examples or to use existing examples 
from the library for own research, the user 
can simply copy an existing example. In this 
new example, own code and changes can be compiled.

Adding a new regression test to the repository 
needs some concentration of the user. In the 
following, we explain how to do this and what 
\textit{must} be taken into account:
\begin{verbatim}
cp -r Example1 ExampleNew
cd ExampleNew
rm -r CVS
cd Test
rm -r CVS
\end{verbatim}
It is important to remove the repository information,
which is stored in the directory CVS.

Then, we have to consider the following step. In order to 
run the regression tests, we start the executable file
of the present example. This executable comes from the 
Makefile. For example in the Makefile of 
Example \ref{PDE_Stat_Stokes}, 
\begin{verbatim}
target   = $(BINDIR)/DOpE-PDE-StatPDE-Example1-$
           (dope_dimension)d-$(deal_II_dimension)d
\end{verbatim}
and concretely:
\begin{verbatim}
DOpE-PDE-StatPDE-Example1-2d-2d
\end{verbatim}
The aforementioned expression is the name of 
the executable that is required to run the 
regression test. Hence, in 
\begin{verbatim}
PDE/StatPDE/Example1/Test> emacs test.sh
\end{verbatim}
you find two times the line
\begin{verbatim}
echo "Running Program 
     ../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm"
     (../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm 2>&1) 
     > /dev/null
\end{verbatim} 
It is really important when copying an existing example,
to change this information. 
First, we have to change the target in the Makefile
\begin{verbatim}
target   = $(BINDIR)/DOpE-PDE-StatPDE-ExampleNew-$
           (dope_dimension)d-$(deal_II_dimension)d
\end{verbatim}
Second, we have to replace 
Example1 by ExampleNew such that 
\begin{verbatim}
echo "Running Program 
     ../../../../../bin/DOpE-PDE-StatPDE-ExampleNew-2d-2d test.prm"
     (../../../../../bin/DOpE-PDE-StatPDE-ExampleNew-2d-2d test.prm 2>&1) 
     > /dev/null
\end{verbatim} 
Please, also change the dimension parameters 
\begin{verbatim}
-2d-2d
\end{verbatim} 
Now, the user is prepared to change any information in the 
\begin{verbatim}
main.cc, localpde.h, functionals.h, localfunctional.h, etc
\end{verbatim}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Examples for PDE Solution}
\label{PDE}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Stationary PDEs} 
\label{PDE_Stat}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stationary Stokes Equations} 
\label{PDE_Stat_Stokes}
\input{PDE/StatPDE/Example1/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stationary FSI with INH Material} 
\label{PDE_Stat_FSI_INH}
\input{PDE/StatPDE/Example2/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stationary FSI with STVK Material} 
\label{PDE_Stat_FSI_STVK}
\input{PDE/StatPDE/Example3/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stationary Elasticity Benchmark} 
\label{PDE_Stat_Elasticity}
\input{PDE/StatPDE/Example4/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stationary Plasticity Benchmark} 
\label{PDE_Stat_Plasticity}
\input{PDE/StatPDE/Example5/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Stationary Stokes Equations with periodic BC} 
\label{PDE_Stat_Periodic_Stokes}
\input{PDE/StatPDE/Example6/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Laplace Equation in 2D} 
\label{PDE_Stat_Laplace_2D}
\input{PDE/StatPDE/Example7/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Laplace Equation in 3D}
\label{PDE_Stat_Laplace_3D}
\input{PDE/StatPDE/Example8/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Nonstationary PDEs}
\label{PDE_Instat}
\input{PDE/InstatPDE/content.tex}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Nonstationary Stokes Equations}
\label{PDE_Instat_Stokes}
\input{PDE/InstatPDE/Example1/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Nonstationary FSI Problem}
\label{PDE_Instat_FSI}
\input{PDE/InstatPDE/Example2/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Black-Scholes Equation}
\label{PDE_Instat_Black_Scholes}
\input{PDE/InstatPDE/Example3/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Heat Equation in 1D}
\label{PDE_Instat_Heat_1D}
\input{PDE/InstatPDE/Example4/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Heat Equation in 2D with nonlinearity}
\label{PDE_Instat_Heat_2D}
\input{PDE/InstatPDE/Example5/content.tex}
\cleardoublepage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Examples with Optimization}
\label{OPT}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Subject to a Stationary PDE}
\label{OPT_Stat}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Distributed control with a linear elliptic PDE}
\label{OPT_Stat_Distrib_Lin_Ellipt}
\input{OPT/StatPDE/Example1/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Parameter control with a linear elliptic PDE}
\label{OPT_Stat_Param_Lin_Ellipt}
\input{OPT/StatPDE/Example2/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Parameter control with a nonlinear PDE from fluid dynamcis}
\label{OPT_Stat_Param_Nonlin_Fluid}
\input{OPT/StatPDE/Example3/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Control in the dirichlet boundary values}
\label{OPT_Stat_Dirichlet_Boundary}
\input{OPT/StatPDE/Example4/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Distributed Control with Different Meshes for Control and State}
\label{OPT_Stat_Distributed_MultiMesh}
\input{OPT/StatPDE/Example5/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Compliance Minimization of a variable Thickness MBB-Beam}
\label{OPT_Stat_MBB-Beam}
\input{OPT/StatPDE/Example6/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Distributed control with a linear elliptic PDE using SNOPT}
\label{OPT_Stat_Box_controlconstraints_SNOPT}
\input{OPT/StatPDE/Example7/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Topology optimization of an MBB-Beam using SNOPT}
\label{OPT_Stat_TopOpt_MBB_SNOPT}
\input{OPT/StatPDE/Example8/content.tex}
\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Parameter control with a non-linear PDE from FSI dynamics}
\label{OPT_Stat_Param_Nonlin_FSI}
%\input{OPT/StatPDE/Example5/content.tex}
\clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Testing examples}
The DOpE testsuite consists of regression tests. They are run to compare 
the output to previous outputs. This is useful (necessary) after 
changing programming code anywhere in the library. If a test
succeeds, everything is fine in the library. If not, you should not
check in your code into DOpE. Please make sure what is going wrong and WHY!

Every command is computed via a Makefile. In the basic example 
directory

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Where can I find the tests}
In each example directory you find a subdirectory `Test'. Herein, you find 
the parameter files for meshes (*.inp) and a param file (test.prm).
Moreover, the executable is denoted by `test.sh'. Please make sure, that 
the 
\begin{verbatim}
set never_write_list
\end{verbatim}
contains every possible output
\begin{verbatim}
Gradient;Hessian;Tangent;Residual;Update;Control;State
\end{verbatim}
That means, no solution files are written to the output.
Recall, that we are just interested in terminal output that 
is of course sufficient to verify the things.

Hence, the results directory should be empty
\begin{verbatim}
set results_dir   = ./
\end{verbatim}
The rest in the param file must be identically the same as 
in the dope.prm file in the parent directory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{How to start testing?}
You start testing by typing 
\begin{verbatim}
> ./test.sh Store
\end{verbatim}
in the terminal.

After the run, you have to call 
\begin{verbatim}
> ./test.sh Test
\end{verbatim}
to compare your stored output. Of course, there should be no differences.

The useful point is now the following. After implemention of 
new pieces of code in the DOpE library or in the examples, you can
run 
\begin{verbatim}
> ./test.sh Test
\end{verbatim}
Hereby, you compare your `new' output with the previous stored output.

Attention: After changes you should NOT run again
\begin{verbatim}
> ./test.sh Store
\end{verbatim}
In that case, you overwrite your previous output.
\cleardoublepage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\addcontentsline{toc}{chapter}{Index}
\printindex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2012-03-15 DOpEsrc/basic/spacetimehandler_base.h

@@ -1,0 +1,361 @@
+#ifndef _SPACE_TIME_HANDLER_BASE_H_
+#define _SPACE_TIME_HANDLER_BASE_H_
+
+#include <lac/vector.h>
+#include <lac/block_vector_base.h>
+#include <lac/block_vector.h>
+
+#include <vector>
+#include <iostream>
+#include <sstream>
+#include <limits>
+
+#include "timedofhandler.h"
+#include "timeiterator.h"
+
+namespace DOpE
+{
+/**
+ * Interface to the dimension independent functionality of a
+ * SpaceTimeDoFHandler
+ */
+template<typename VECTOR>//has to be a class template because you can no have virtual member function templates.
+class SpaceTimeHandlerBase
+{
+  public:
+
+    SpaceTimeHandlerBase()
+    {
+      _state_ticket = 1;
+      _control_ticket = 1;
+    }
+
+    SpaceTimeHandlerBase(const dealii::Triangulation<1> & times) :
+      _tdfh(times), _interval(_tdfh.first_interval())
+    {
+      _state_ticket = 1;
+      _control_ticket = 1;
+    }
+
+    SpaceTimeHandlerBase(const dealii::Triangulation<1> & times,
+        const dealii::FiniteElement<1>& fe) :
+          _tdfh(times, fe), _interval(_tdfh.first_interval())
+    {
+      _state_ticket = 1;
+      _control_ticket = 1;
+    }
+
+
+    virtual ~SpaceTimeHandlerBase()
+    {
+      _tdfh.clear();
+    }
+
+    /**
+     * This function has to get called after temporal refinement.
+     */
+    void ReInitTime()
+    {
+      IncrementStateTicket();
+      _tdfh.distribute_dofs();
+      //FIXME When we have temporal discretization also for control and constraint,
+      //one has to increment here the _control_ticket!
+    }
+
+    /**
+     * Returns the index of the latest time point (i.e. DoF) in the time mesh.
+     * Counting starts at zero.
+     *
+     * @return The maximal feasible time point.
+     *
+     */
+    unsigned int GetMaxTimePoint() const
+    {
+      return (_tdfh.GetNbrOfDoFs()-1);//because we start counting at 0.
+    }
+
+    /**
+     * Returns the number of intervals in the actual time triangulation.
+     */
+    unsigned int GetNbrOfIntervals() const
+    {
+      return _tdfh.GetNbrOfIntervals();
+    }
+
+
+    /**
+     * Sets the current interval
+     *
[…]

### 2012-03-15 DOpEsrc/include/functionalinterface.h

@@ -1,0 +1,443 @@
+#ifndef _FUNCTIONAL_INTERFACE_H_
+#define _FUNCTIONAL_INTERFACE_H_
+
+#include <map>
+#include <string>
+
+#include <fe/fe_system.h>
+#include <fe/fe_values.h>
+#include <fe/mapping.h>
+
+#include "fevalues_wrapper.h"
+#include "dofhandler_wrapper.h"
+#include "celldatacontainer.h"
+#include "facedatacontainer.h"
+#include "multimesh_celldatacontainer.h"
+#include "multimesh_facedatacontainer.h"
+
+namespace DOpE
+{
+/**
+ * A template for an arbitrary Functional J to be used as cost functional for an optimization problem.
+ * Or any Functional that should be evaluated. For evaluation only *Value routines are required, but none  of
+ * the derivatives thereof.
+ */
+template<template<typename DOFHANDLER, typename VECTOR, int dealdim> class CDC, template<typename DOFHANDLER, typename VECTOR, int dealdim> class FDC, typename DOFHANDLER, typename VECTOR, int dopedim, int dealdim=dopedim>
+class FunctionalInterface
+{
+  public:
+    FunctionalInterface();
+    ~FunctionalInterface();
+
+    /**
+     * This evaluates the Cost Functional J(q,u) = \int_\Omega j(q(x),u(x)) \dx on a given element T.
+     *
+     * @param control_fe_values        The DOpEWrapper::FEValues preinitialized to the given element for the control_variable
+     * @param state_fe_values          The DOpEWrapper::FEValues preinitialized to the given element for the state_variable
+     * @param param_values             A std::map containing parameter data (e.g. non space dependent data). If the control
+     *                                 is done by parameters, it is contained in this map at the position "control".
+     * @param domain_values            A std::map containing domain data (e.g. nodal vectors for FE-Functions). If the control
+     *                                 is distributed, it is contained in this map at the position "control". The state may always
+     *                                 be found in this map at the position "state"
+     * @param n_q_points               The number of quadrature points to be considered on this element.
+     * @param material_id              An unsigned integer that may indicate where we are in the domain.
+     * @param cell_diameter            The diameter of the current element.
+     *
+     * @return                         A number which is \int_T j(q(x),u(x)) \dx
+     */
+    virtual double Value(const CDC<DOFHANDLER, VECTOR, dealdim>& cdc);
+
+
+    /**
[…]

### 2012-03-15 DOpEsrc/opt_algorithms/reduced_snopt_algorithm.h

@@ -1,0 +1,461 @@
+#ifndef _REDUCED_SNOPT__ALGORITHM_H_
+#define _REDUCED_SNOPT__ALGORITHM_H_
+
+#include "reducedalgorithm.h"
+#include "parameterreader.h"
+
+#include "snopt_wrapper.h"
+
+#include <iostream>
+#include <assert.h>
+#include <iomanip>
+
+namespace DOpE
+{
+
+  template <typename PROBLEM, typename VECTOR, int dopedim,  int dealdim>
+    class Reduced_SnoptAlgorithm : public ReducedAlgorithm<PROBLEM, VECTOR, dopedim,dealdim>
+  {
+  public:
+    Reduced_SnoptAlgorithm(PROBLEM* OP,
+			   ReducedProblemInterface<PROBLEM, VECTOR,dopedim,dealdim>* S,
+			   std::string vector_behavior,
+			   ParameterReader &param_reader,
+			   DOpEExceptionHandler<VECTOR>* Except=NULL,
+			   DOpEOutputHandler<VECTOR>* Output=NULL,
+			   int base_priority=0);
+    ~Reduced_SnoptAlgorithm();
+
+    static void declare_params(ParameterReader &param_reader);
+
+    /**
+     * This solves an Optimizationproblem in only the control variable
+     * using the commercial optimization library snopt. 
+     * To use it you need to define the compiler flag 
+     * WITH_SNOPT and ensure that all required snopt headers and 
+     * libraries are within the path or otherwise known.
+     *
+     * @param q           The initial point.
+     * @param global_tol  An optional parameter specifying the required  tolerance.
+     *                    The actual tolerance is the maximum of this and the one specified in the param
+     *                    file. Its default value is negative, so that it has no influence if not specified.
+     */
+    virtual int Solve(ControlVector<VECTOR>& q,double global_tol=-1.);
+
+  protected:
+
+  private:
+#ifdef WITH_SNOPT
+    int rsa_func_(DOpEWrapper::SNOPT_FUNC_DATA& data);
+#endif
+    std::string _postindex;
+    std::string _vector_behavior;
+
+    double _func_prec, _feas_tol, _opt_tol;
+    int _max_inner_iter, _max_outer_iter;
+    bool _capture_out;
+  };
+
+  /***************************************************************************************/
[…]

### 2012-03-15 DOpEsrc/source/dopeexceptionhandler.cc

@@ -1,0 +1,60 @@
+#include "dopeexceptionhandler.h"
+#include "outputhandler.h"
+
+#include <iostream>
+#include <string>
+#include <cstdlib>
+
+
+namespace DOpE
+{
+
+/*******************************************************/
+
+template<typename VECTOR>
+DOpEExceptionHandler<VECTOR>::DOpEExceptionHandler(DOpEOutputHandler<VECTOR>* OutputHandler)
+{
+  _OutputHandler = OutputHandler;
+}
+
+/*******************************************************/
+
+template<typename VECTOR>
+DOpEExceptionHandler<VECTOR>::~DOpEExceptionHandler()
+{
+
+}
+
+/*******************************************************/
+
+template<typename VECTOR>
+void DOpEExceptionHandler<VECTOR>::HandleException(DOpEException& e,std::string reporter)
+{
+  GetOutputHandler()->WriteError("The following `" + e.GetName() + "` reports!");
+  GetOutputHandler()->WriteError("Warning: During execution of `" + e.GetThrowingInstance()
+      + "` the following Problem occurred!");
+  GetOutputHandler()->WriteError(e.GetErrorMessage());
+  GetOutputHandler()->WriteError("Reported by `"+reporter+"`");
+}
+
+/*******************************************************/
+
+template<typename VECTOR>
+void DOpEExceptionHandler<VECTOR>::HandleCriticalException(DOpEException& e,std::string reporter)
+{
+  GetOutputHandler()->WriteError("The following `" + e.GetName() + "` reports!");
+  GetOutputHandler()->WriteError("Error: During execution of `" + e.GetThrowingInstance()
+      + "` the following fatal Problem occurred!");
+  GetOutputHandler()->WriteError(e.GetErrorMessage());
+  GetOutputHandler()->WriteError("Reported by `"+reporter+"`");
+  abort();
+}
+
+}//Endof namespace
+/******************************************************/
+
+/******************************************************/
+/******************************************************/
+template class DOpE::DOpEExceptionHandler<dealii::Vector<double> >;
+template class DOpE::DOpEExceptionHandler<dealii::BlockVector<double> >;
+

### 2012-03-15 DOpEsrc/source/statpdeproblem.cc

@@ -1,0 +1,77 @@
+#include "statpdeproblem.h"
+#include "preconditioner_wrapper.h"
+#include "integratordatacontainer.h"
+
+
+#if dope_dimension == deal_II_dimension
+
+/******************************************************/
+#define _IDC     DOpE::IntegratorDataContainer<dealii::DoFHandler<deal_II_dimension>, dealii::Quadrature<deal_II_dimension>, dealii::Quadrature<deal_II_dimension-1>,dealii::BlockVector<double>, deal_II_dimension  >
+#define _DOP     DOpE::PDEProblemContainer<DOpE::PDEInterface<DOpE::CellDataContainer,DOpE::FaceDataContainer,dealii::DoFHandler<deal_II_dimension>,dealii::BlockVector<double>,deal_II_dimension>,DOpE::DirichletDataInterface<dealii::BlockVector<double>,deal_II_dimension>,dealii::BlockSparsityPattern,dealii::BlockVector<double>,deal_II_dimension>
+#define _DIN     DOpE::Integrator<_IDC,dealii::BlockVector<double>,double,deal_II_dimension>
+#define _DCGS    DOpE::CGLinearSolverWithMatrix<DOpEWrapper::PreconditionIdentity_Wrapper<dealii::BlockSparseMatrix<double> >, dealii::BlockSparsityPattern,dealii::BlockSparseMatrix<double>,dealii::BlockVector<double>,deal_II_dimension>
+#define _DGMRESS DOpE::GMRESLinearSolverWithMatrix<DOpEWrapper::PreconditionIdentity_Wrapper<dealii::BlockSparseMatrix<double> >,dealii::BlockSparsityPattern,dealii::BlockSparseMatrix<double>,dealii::BlockVector<double>,deal_II_dimension>
+#define _DDS     DOpE::DirectLinearSolverWithMatrix<dealii::BlockSparsityPattern,dealii::BlockSparseMatrix<double>,dealii::BlockVector<double>,deal_II_dimension>
+
+/******************************************************/
+template class DOpE::StatPDEProblem<
+  DOpE::NewtonSolver<_DIN,_DCGS,dealii::BlockVector<double>,deal_II_dimension>,
+  _DIN,
+  _DOP,dealii::BlockVector<double>,deal_II_dimension>;
+
+/******************************************************/
+template class DOpE::StatPDEProblem<
[…]

### 2012-03-15 Examples/OPT/StatPDE/Example1/Test/test.sh

@@ -1,0 +1,49 @@
+#!/bin/bash
+if [ $# -ne 1 ]
+    then
+    echo "Usage: "$0" [Test|Store]"
+    exit 1
+fi
+
+if [ $1 == "Test" ]
+then
+    if [ -f test.log ]
+    then
+	echo "Running Program ../../../../../bin/DOpE-OPT-StatPDE-Example1-2d-2d test.prm"
+	(../../../../../bin/DOpE-OPT-StatPDE-Example1-2d-2d test.prm 2>&1) > /dev/null
+	echo "Comparing Results:"
+	(diff dope.log test.log 2>&1) > /dev/null
+	if [ $? -eq 0 ]
+	then
+	    echo "No differences found."
+	    rm dope.log
+	    rm -r Mesh0/
+	    rm -r Mesh1/
+	    exit 0
+	else
+	    echo "There where discrepancies in the Output."
+	    diff dope.log test.log
+	    rm -r Mesh0/
+	    rm -r Mesh1/
+	    exit 1
+	fi
+    else
+	echo "No File test.log found for comparisson. Run '"$0" Store' to create one."
+	exit 1
+    fi
+else
+    if [ $1 == "Store" ]
+    then
+	echo "Running Program ../../../../../bin/DOpE-OPT-StatPDE-Example1-2d-2d test.prm"
+	(../../../../../bin/DOpE-OPT-StatPDE-Example1-2d-2d test.prm 2>&1) > /dev/null
+	echo "Run completed. Cleaning up ..."
+	mv dope.log test.log
+	rm -r Mesh0/
+	rm -r Mesh1/
+	exit 0;
+    else
+	echo "Unknown Option: "$1
+	exit 1
+    fi
+fi
+    

### 2012-03-15 Examples/OPT/StatPDE/Example1/main.cc

@@ -1,0 +1,176 @@
+
+#include "reducednewtonalgorithm.h"
+#include "reducedtrustregionnewton.h"
+#include "optproblem.h"
+#include "functionalinterface.h"
+#include "pdeinterface.h"
+#include "statreducedproblem.h"
+#include "newtonsolver.h"
+#include "cglinearsolver.h"
+#include "directlinearsolver.h"
+#include "integrator.h"
+#include "parameterreader.h"
+#include "mol_spacetimehandler.h"
+#include "simpledirichletdata.h"
+#include "noconstraints.h"
+#include "sparsitymaker.h"
+#include "constraintsmaker.h"
+#include "preconditioner_wrapper.h"
+#include "integratordatacontainer.h"
+
+#include <iostream>
+
+#include <grid/tria.h>
+#include <dofs/dof_handler.h>
+#include <grid/grid_generator.h>
+#include <fe/fe_q.h>
+#include <dofs/dof_tools.h>
+#include <base/quadrature_lib.h>
+#include <base/function.h>
+#include <lac/precondition_block.h>
+
+#include "localpde.h"
+#include "localfunctional.h"
+#include "functionals.h"
+
+using namespace std;
+using namespace dealii;
+using namespace DOpE;
+
+#define VECTOR dealii::BlockVector<double>
+#define DOFHANDLER dealii::DoFHandler<2>
+#define FE DOpEWrapper::FiniteElement<2>
+
+typedef OptProblem<FunctionalInterface<CellDataContainer,FaceDataContainer,DOFHANDLER, VECTOR, 2,2>,
+		   FunctionalInterface<CellDataContainer,FaceDataContainer,DOFHANDLER, VECTOR, 2,2>,
+		   PDEInterface<CellDataContainer,FaceDataContainer,DOFHANDLER, VECTOR,2,2>,
+		   DirichletDataInterface<VECTOR,2,2>,
+		   ConstraintInterface<CellDataContainer,FaceDataContainer,DOFHANDLER, VECTOR, 2,2>,
+		   BlockSparsityPattern, VECTOR, 2,2> OP;
+
+typedef IntegratorDataContainer<DOFHANDLER, dealii::Quadrature<2>, dealii::Quadrature<1>, VECTOR, 2 > IDC;
+
[…]

### 2012-03-15 Examples/OPT/StatPDE/Example3/localpde.h

@@ -1,0 +1,1237 @@
+#ifndef _LOCALPDE_
+#define _LOCALPDE_
+
+#include "pdeinterface.h"
+
+using namespace std;
+using namespace dealii;
+using namespace DOpE;
+
+template<typename VECTOR, int dopedim, int dealdim>
+  class LocalPDE : public PDEInterface<CellDataContainer,FaceDataContainer,dealii::DoFHandler<dealdim>, VECTOR, dopedim,dealdim>
+  {
+  public:
+
+    static void declare_params(ParameterReader &param_reader)
+    {
+      param_reader.SetSubsection("Local PDE parameters");
+      param_reader.declare_entry("density_fluid", "0.0",
+				 Patterns::Double(0));
+      param_reader.declare_entry("viscosity", "0.0",
+				 Patterns::Double(0));
+      param_reader.declare_entry("alpha_u", "0.0",
+				 Patterns::Double(0));
+      param_reader.declare_entry("mu", "0.0",
+				 Patterns::Double(0));
+      param_reader.declare_entry("poisson_ratio_nu", "0.0",
+				 Patterns::Double(0));
+    }
+
+
+  LocalPDE(ParameterReader &param_reader) : _control_block_components(2,0), _state_block_components(5,0)
+      {
+	// control block components
+	_control_block_components[0]= 0;
+	_control_block_components[1]= 1;
+
+	// state block components
+	_state_block_components[2]= 1;  // pressure
+	_state_block_components[3]= 2;  // displacement
+	_state_block_components[4]= 2;
+
+	param_reader.SetSubsection("Local PDE parameters");
+	_density_fluid = param_reader.get_double ("density_fluid");
+	_viscosity = param_reader.get_double ("viscosity");
+	_alpha_u = param_reader.get_double ("alpha_u");
+
+	_lame_coefficient_mu = param_reader.get_double ("mu");
+	_poisson_ratio_nu = param_reader.get_double ("poisson_ratio_nu");
+	_lame_coefficient_lambda =  (2 * _poisson_ratio_nu * _lame_coefficient_mu)/
+	  (1.0 - 2 * _poisson_ratio_nu);
+      }
+
+
+  // Domain values for cells
[…]

### 2012-03-15 Examples/OPT/StatPDE/Example3/main.cc

@@ -1,0 +1,292 @@
+#include "reducednewtonalgorithm.h"
+#include "optproblem.h"
+#include "functionalinterface.h"
+#include "pdeinterface.h"
+#include "statreducedproblem.h"
+#include "newtonsolver.h"
+#include "gmreslinearsolver.h"
+#include "cglinearsolver.h"
+#include "directlinearsolver.h"
+#include "voidlinearsolver.h" // for mixed dim opt. control
+#include "integrator.h"
+#include "noconstraints.h"
+#include "solutionextractor.h"
+
+#include "integratormixeddims.h"    // for mixed dim opt. control
+#include "newtonsolvermixeddims.h"  // for mixed dim opt. control
+#include "finiteelement_wrapper.h"
+
+#include "parameterreader.h"
+#include "mol_spacetimehandler.h"
+#include "simpledirichletdata.h"
+#include "sparsitymaker.h"
+#include "constraintsmaker.h"
+#include "integratordatacontainer.h"
+
+// C++
+#include <iostream>
+#include <fstream>
+
+
+// deal.II
+#include <grid/tria.h>
+#include <grid/grid_in.h>
+#include <grid/tria_boundary_lib.h>
+#include <dofs/dof_handler.h>
+#include <grid/grid_generator.h>
+#include <fe/fe_q.h>
+#include <fe/fe_dgp.h>
+#include <fe/fe_nothing.h>
+#include <dofs/dof_tools.h>
+#include <base/quadrature_lib.h>
+#include <base/function.h>
+
+// for grid local refinement
+#include <numerics/error_estimator.h>
+#include <grid/grid_refinement.h>
+
+
+#include "localpde.h"
+#include "localfunctional.h"
+#include "functionals.h"
+
+#include "my_functions.h"
+
+
+
+using namespace std;
+using namespace dealii;
+using namespace DOpE;
+
+#define VECTOR dealii::BlockVector<double>
+#define DOFHANDLER dealii::DoFHandler<2>
+#define FE DOpEWrapper::FiniteElement<2>
+
+typedef OptProblem<FunctionalInterface<CellDataContainer,FaceDataContainer,DOFHANDLER, VECTOR,0,2>,
+		   FunctionalInterface<CellDataContainer,FaceDataContainer,DOFHANDLER, VECTOR,0,2>,
[…]

### 2012-03-15 Examples/OPT/StatPDE/Example5/localfunctional.h

@@ -1,0 +1,173 @@
+#ifndef _LOCALFunctional_
+#define _LOCALFunctional_
+
+#include "functionalinterface.h"
+
+using namespace std;
+using namespace dealii;
+using namespace DOpE;
+
+template<typename VECTOR, int dopedim, int dealdim>
+  class LocalFunctional : public FunctionalInterface<Multimesh_CellDataContainer,Multimesh_FaceDataContainer,dealii::DoFHandler<dealdim>, VECTOR, dopedim,dealdim>
+  {
+  public:
+    LocalFunctional()
+    {
+      _alpha = 1.e-3;
+    }
+
+    double Value(const Multimesh_CellDataContainer<dealii::DoFHandler<dealdim>, VECTOR, dealdim>& cdc)
+    {
+      const DOpEWrapper::FEValues<dealdim> & state_fe_values = cdc.GetFEValuesState();
+      unsigned int n_q_points = cdc.GetNQPoints();
+
+      {
+	_qvalues.resize(n_q_points);
+	_fvalues.resize(n_q_points);
+	_uvalues.resize(n_q_points);
+
+	cdc.GetValuesControl("control",_qvalues);
+  	cdc.GetValuesState("state",_uvalues);
+      }
+
+      double r = 0.;
+
+      for(unsigned int q_point = 0; q_point< n_q_points; q_point++)
+      {
+	_fvalues[q_point] = ( 1.*sin(4 * M_PI * state_fe_values.quadrature_point(q_point)(0)) +
+			      5.*M_PI*M_PI*sin(M_PI * state_fe_values.quadrature_point(q_point)(0))) *
+	  sin(2 * M_PI * state_fe_values.quadrature_point(q_point)(1));
+
+	r += 0.5* (_uvalues[q_point]-_fvalues[q_point])*(_uvalues[q_point]-_fvalues[q_point])
+	  * state_fe_values.JxW(q_point);
+	r += 0.5*_alpha* (_qvalues[q_point]*_qvalues[q_point])
+	  * state_fe_values.JxW(q_point);
+      }
+      return r;
+    }
+
+    void Value_U(const Multimesh_CellDataContainer<dealii::DoFHandler<dealdim>, VECTOR, dealdim>& cdc,
+		 dealii::Vector<double> &local_cell_vector,
+		 double scale)
+    {
+      const DOpEWrapper::FEValues<dealdim> & state_fe_values = cdc.GetFEValuesState();
[…]

### 2012-03-15 Examples/OPT/StatPDE/Example7/localconstraints.h

@@ -1,0 +1,151 @@
+#ifndef _LOCAL_CONSTRAINT_H_
+#define _LOCAL_CONSTRAINT_H_
+
+#include "constraintinterface.h"
+#include "localconstraintaccessor.h"
+
+namespace DOpE
+{
+  /**
+   * A template for an arbitrary Constraints.
+   * GlobalConstraints are dealt with as a Functional, hence all functions from Functionals are inherited.
+   */
+  template<template<typename DOFHANDLER, typename VECTOR, int dealdim> class CDC, template<typename DOFHANDLER, typename VECTOR, int dealdim> class FDC, typename DOFHANDLER, typename VECTOR,int dopedim,int dealdim>
+    class LocalConstraint : public ConstraintInterface<CDC,FDC,DOFHANDLER,VECTOR,dopedim,dealdim>
+  {
+  public:
+    LocalConstraint(LocalConstraintAccessor& CA) : LCA(CA) 
+    {
+    }
+    ~LocalConstraint()  {}
+ 
+    void EvaluateLocalControlConstraints(const dealii::BlockVector<double>& control,
+					 dealii::BlockVector<double>& constraints)
+    {
+      assert(constraints.block(0).size() == 2*control.block(0).size());
+
+      for(unsigned int i=0; i < control.block(0).size(); i++)
+      {
+	//Add Control Constraints, such that if control is feasible all  entries are not positive!
+	// _rho_min <= control <= _rho_max
+	LCA.ControlToLowerConstraint(control,constraints);
+	LCA.ControlToUpperConstraint(control,constraints);
+      }
+    }
+    void GetControlBoxConstraints(VECTOR& lb, VECTOR& ub) const
+    {
+      LCA.GetControlBoxConstraints(lb,ub);
+    }
+   
+   std::string GetType() const 
+   { 
+       throw DOpEException("Unknown problem_type "+this->GetProblemType(),"LocalConstraints::GetType");
+   }
+   std::string GetName() const 
+   { 
+       throw DOpEException("Unknown problem_type "+this->GetProblemType(),"LocalConstraints::GetName");
+   } 
+        
+   dealii::UpdateFlags GetUpdateFlags() const
+   {
+     return update_values | update_quadrature_points;
+   }
+       
+   bool IsFeasible(const ConstraintVector<VECTOR>& g) const
+   {
[…]

### 2012-03-15 Examples/OPT/StatPDE/Example8/localconstraints.h

@@ -1,0 +1,265 @@
+#ifndef _LOCAL_CONSTRAINT_H_
+#define _LOCAL_CONSTRAINT_H_
+
+#include "constraintinterface.h"
+#include "localconstraintaccessor.h"
+
+namespace DOpE
+{
+  /**
+   * A template for an arbitrary Constraints.
+   * GlobalConstraints are dealt with as a Functional, hence all functions from Functionals are inherited.
+   */
+  template<template<typename DOFHANDLER, typename VECTOR, int dealdim> class CDC, template<typename DOFHANDLER, typename VECTOR, int dealdim> class FDC, typename DOFHANDLER, typename VECTOR,int dopedim,int dealdim>
+    class LocalConstraint : public ConstraintInterface<CDC,FDC,DOFHANDLER,VECTOR,dopedim,dealdim>
+  {
+  public:
+    LocalConstraint(LocalConstraintAccessor& CA) : LCA(CA) 
+    {
+      _vol_max = 0.5;
+    }
+    ~LocalConstraint()  {}
+ 
+    void EvaluateLocalControlConstraints(const dealii::BlockVector<double>& control,
+					 dealii::BlockVector<double>& constraints)
+    {
+      assert(constraints.block(0).size() == 2*control.block(0).size());
+
+      for(unsigned int i=0; i < control.block(0).size(); i++)
+      {
+	//Add Control Constraints, such that if control is feasible all  entries are not positive!
+	// _rho_min <= control <= _rho_max
+	LCA.ControlToLowerConstraint(control,constraints);
+	LCA.ControlToUpperConstraint(control,constraints);
+      }
+    }
+    void GetControlBoxConstraints(VECTOR& lb, VECTOR& ub) const
+    {
+      LCA.GetControlBoxConstraints(lb,ub);
+    }
+
+    double Value(const CDC<DOFHANDLER,VECTOR,dealdim>& cdc)
+    {
+      if(this->GetProblemType() == "global_constraints" && this->GetProblemTypeNum()==0)
+      {
+	const DOpEWrapper::FEValues<dealdim> & control_fe_values = cdc.GetFEValuesControl();
+	unsigned int n_q_points = cdc.GetNQPoints();
+ 	
+	double ret = 0.;
+	{
+	  _qvalues.resize(n_q_points,Vector<double>(1));
+	  cdc.GetValuesControl("control",_qvalues);
+	}
+	for(unsigned int q_point = 0; q_point < n_q_points; q_point++)
+	{
[…]

### 2012-03-15 Examples/PDE/InstatPDE/Example2/localpde.h

@@ -1,0 +1,1570 @@
+#ifndef _LOCALPDE_
+#define _LOCALPDE_
+
+#include "pdeinterface.h"
+#include "ale_transformations.h"
+#include "celldatacontainer.h"
+#include "facedatacontainer.h"
+
+
+using namespace std;
+using namespace dealii;
+using namespace DOpE;
+
+template<typename VECTOR, int dopedim, int dealdim>
+  class LocalPDE : public PDEInterface<CellDataContainer,FaceDataContainer,dealii::DoFHandler<dealdim>, VECTOR,dopedim,dealdim>
+  {
+  public:
+
+     static void declare_params(ParameterReader &param_reader)
+    {
+      param_reader.SetSubsection("Local PDE parameters");
+      param_reader.declare_entry("density_fluid", "1.0",
+				 Patterns::Double(0));
+      param_reader.declare_entry("density_structure", "1.0",
+				 Patterns::Double(0));
+      param_reader.declare_entry("viscosity", "1.0",
+				 Patterns::Double(0));
+      param_reader.declare_entry("alpha_u", "1.0",
+				 Patterns::Double(0));
+      param_reader.declare_entry("mu", "1.0",
+				 Patterns::Double(0));
+      param_reader.declare_entry("poisson_ratio_nu", "0.4",
+				 Patterns::Double(0));
+    }
+
+  LocalPDE(ParameterReader &param_reader) : _state_block_components(5,0)
+      {
+	_alpha = 1.e-3;
+	_state_block_components[2]= 1;
+	_state_block_components[3]= 1;
+	_state_block_components[4]= 2;
+
+	param_reader.SetSubsection("Local PDE parameters");
+	density_fluid = param_reader.get_double ("density_fluid");
+	density_structure = param_reader.get_double ("density_structure");
+	viscosity = param_reader.get_double ("viscosity");
+	alpha_u = param_reader.get_double ("alpha_u");
+	lame_coefficient_mu = param_reader.get_double ("mu");
+	poisson_ratio_nu = param_reader.get_double ("poisson_ratio_nu");
+
+	lame_coefficient_lambda =  (2 * poisson_ratio_nu * lame_coefficient_mu)/
+	  (1.0 - 2 * poisson_ratio_nu);
+
+	// variables for mesh-dependent alpha_u
+	structure_continuation_E = 1.0e-6;
+	scale_distance_to_interface = 5.0e-6;
+	scale_d_exp = 100.0;
+
[…]

### 2012-03-15 Examples/PDE/StatPDE/Example7/localpde.h

@@ -1,0 +1,142 @@
+#ifndef _LOCALPDE_
+#define _LOCALPDE_
+
+#include "pdeinterface.h"
+#include "celldatacontainer.h"
+#include "facedatacontainer.h"
+
+using namespace std;
+using namespace dealii;
+using namespace DOpE;
+
+template<typename VECTOR, int dealdim>
+class LocalPDE: public PDEInterface<CellDataContainer,FaceDataContainer,dealii::DoFHandler<dealdim>, VECTOR, dealdim>
+{
+  public:
+    LocalPDE() :
+      _state_block_components(2, 0)
+    {
+
+    }
+
+    // Domain values for cells
+    void CellEquation(const CellDataContainer<dealii::DoFHandler<dealdim>, VECTOR, dealdim>& cdc,
+                      dealii::Vector<double> &local_cell_vector,
+                      double scale, double)
+    {
+      assert(this->_problem_type == "state");
+
+      const DOpEWrapper::FEValues<dealdim> & state_fe_values = cdc.GetFEValuesState();
+      unsigned int n_dofs_per_cell = cdc.GetNDoFsPerCell();
+      unsigned int n_q_points = cdc.GetNQPoints();
+
+      _ugrads.resize(n_q_points, vector<Tensor<1, 2> > (2));
+
+      cdc.GetGradsState("last_newton_solution", _ugrads);
+
+      const FEValuesExtractors::Vector displacements(0);
+
+      for (unsigned int q_point = 0; q_point < n_q_points; q_point++)
+      {
+        Tensor<2, 2> ugrads;
+        ugrads.clear();
+        ugrads[0][0] = _ugrads[q_point][0][0];
+        ugrads[0][1] = _ugrads[q_point][0][1];
+        ugrads[1][0] = _ugrads[q_point][1][0];
+        ugrads[1][1] = _ugrads[q_point][1][1];
+
+	for (unsigned int i = 0; i < n_dofs_per_cell; i++)
+        {
+	  const Tensor<2, 2> phi_i_grads_u = state_fe_values[displacements].gradient(i, q_point);
+
+          local_cell_vector(i) += scale * scalar_product(ugrads, phi_i_grads_u) 
+	    * state_fe_values.JxW(q_point);
+        }
+      }
+
+    }
+
+    void CellMatrix(const CellDataContainer<dealii::DoFHandler<dealdim>, VECTOR, dealdim>& cdc, FullMatrix<double> &local_entry_matrix, double, double)
+    {
+
[…]

### 2012-03-16 Examples/Documentation/foreword.tex

--- 2012-03-15 Examples/description.tex
@@
-\documentclass[a4paper,cleardoubleempty]{scrreprt}
-\usepackage{amsmath,amssymb}
-\usepackage[pdftex,bookmarks=false,pdfpagelabels=true,plainpages=false]{hyperref}
-\usepackage{booktabs}
-\usepackage{makeidx}
-%\usepackage{sfmath}
-\newtheorem{Problem}{Problem}[section]
-
-\newcommand{\todo}[1]{\textbf{\textsc{\textcolor{black}{TODO: #1}}}}
-
-\makeindex
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-\title{DOpE - Deal Optimization Environment}
-\author{
-  Michael Geiger\thanks{University of Heidelberg, 
-    {michael.geiger@iwr.uni-heidelberg.de}}, 
-  Christian Goll\thanks{University of Heidelberg, 
-    {christian.goll@iwr.uni-heidelberg.de}}, 
-  Thomas Wick\thanks{University of Heidelberg, 
-    {thomas.wick@iwr.uni-heidelberg.de}}, 
-  Winnifried Wollner\thanks{University of Hamburg, 
-    {winnifried.wollner@math.uni-hamburg.de}}
-}
-
-\date{\today}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\maketitle
-\cleardoublepage\pagestyle{headings}
-\tableofcontents
-\cleardoublepage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Introduction}
+\chapter{Foreword}
 In this report, we describe the DOpE
@@
 recommendable to be roughly acquainted with deal.II.
-
-\cleardoublepage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Example Handling, Creating new Examples}
-To implement new examples or to use existing examples 
-from the library for own research, the user 
-can simply copy an existing example. In this 
-new example, own code and changes can be compiled.
-
-Adding a new regression test to the repository 
-needs some concentration of the user. In the 
-following, we explain how to do this and what 
-\textit{must} be taken into account:
-\begin{verbatim}
-cp -r Example1 ExampleNew
-cd ExampleNew
-rm -r CVS
-cd Test
-rm -r CVS
-\end{verbatim}
-It is important to remove the repository information,
-which is stored in the directory CVS.
-
-Then, we have to consider the following step. In order to 
-run the regression tests, we start the executable file
-of the present example. This executable comes from the 
-Makefile. For example in the Makefile of 
-Example \ref{PDE_Stat_Stokes}, 
-\begin{verbatim}
-target   = $(BINDIR)/DOpE-PDE-StatPDE-Example1-$
-           (dope_dimension)d-$(deal_II_dimension)d
-\end{verbatim}
-and concretely:
-\begin{verbatim}
-DOpE-PDE-StatPDE-Example1-2d-2d
-\end{verbatim}
-The aforementioned expression is the name of 
-the executable that is required to run the 
-regression test. Hence, in 
-\begin{verbatim}
-PDE/StatPDE/Example1/Test> emacs test.sh
-\end{verbatim}
-you find two times the line
-\begin{verbatim}
-echo "Running Program 
-     ../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm"
-     (../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm 2>&1) 
-     > /dev/null
-\end{verbatim} 
-It is really important when copying an existing example,
-to change this information. 
-First, we have to change the target in the Makefile
-\begin{verbatim}
-target   = $(BINDIR)/DOpE-PDE-StatPDE-ExampleNew-$
-           (dope_dimension)d-$(deal_II_dimension)d
-\end{verbatim}
-Second, we have to replace 
-Example1 by ExampleNew such that 
-\begin{verbatim}
-echo "Running Program 
-     ../../../../../bin/DOpE-PDE-StatPDE-ExampleNew-2d-2d test.prm"
-     (../../../../../bin/DOpE-PDE-StatPDE-ExampleNew-2d-2d test.prm 2>&1) 
-     > /dev/null
-\end{verbatim} 
-Please, also change the dimension parameters 
-\begin{verbatim}
--2d-2d
-\end{verbatim} 
-Now, the user is prepared to change any information in the 
-\begin{verbatim}
-main.cc, localpde.h, functionals.h, localfunctional.h, etc
-\end{verbatim}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Examples for PDE Solution}
-\label{PDE}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Stationary PDEs} 
-\label{PDE_Stat}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary Stokes Equations} 
-\label{PDE_Stat_Stokes}
-\input{PDE/StatPDE/Example1/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary FSI with INH Material} 
-\label{PDE_Stat_FSI_INH}
-\input{PDE/StatPDE/Example2/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary FSI with STVK Material} 
-\label{PDE_Stat_FSI_STVK}
-\input{PDE/StatPDE/Example3/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary Elasticity Benchmark} 
-\label{PDE_Stat_Elasticity}
-\input{PDE/StatPDE/Example4/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary Plasticity Benchmark} 
-\label{PDE_Stat_Plasticity}
-\input{PDE/StatPDE/Example5/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary Stokes Equations with periodic BC} 
-\label{PDE_Stat_Periodic_Stokes}
-\input{PDE/StatPDE/Example6/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Laplace Equation in 2D} 
-\label{PDE_Stat_Laplace_2D}
-\input{PDE/StatPDE/Example7/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Laplace Equation in 3D}
-\label{PDE_Stat_Laplace_3D}
-\input{PDE/StatPDE/Example8/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Nonstationary PDEs}
-\label{PDE_Instat}
-\input{PDE/InstatPDE/content.tex}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Nonstationary Stokes Equations}
-\label{PDE_Instat_Stokes}
-\input{PDE/InstatPDE/Example1/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Nonstationary FSI Problem}
-\label{PDE_Instat_FSI}
-\input{PDE/InstatPDE/Example2/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Black-Scholes Equation}
-\label{PDE_Instat_Black_Scholes}
-\input{PDE/InstatPDE/Example3/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Heat Equation in 1D}
-\label{PDE_Instat_Heat_1D}
-\input{PDE/InstatPDE/Example4/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Heat Equation in 2D with nonlinearity}
-\label{PDE_Instat_Heat_2D}
-\input{PDE/InstatPDE/Example5/content.tex}
-\cleardoublepage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Examples with Optimization}
-\label{OPT}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Subject to a Stationary PDE}
-\label{OPT_Stat}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Distributed control with a linear elliptic PDE}
-\label{OPT_Stat_Distrib_Lin_Ellipt}
-\input{OPT/StatPDE/Example1/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Parameter control with a linear elliptic PDE}
-\label{OPT_Stat_Param_Lin_Ellipt}
-\input{OPT/StatPDE/Example2/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Parameter control with a nonlinear PDE from fluid dynamcis}
-\label{OPT_Stat_Param_Nonlin_Fluid}
-\input{OPT/StatPDE/Example3/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Control in the dirichlet boundary values}
-\label{OPT_Stat_Dirichlet_Boundary}
-\input{OPT/StatPDE/Example4/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Distributed Control with Different Meshes for Control and State}
-\label{OPT_Stat_Distributed_MultiMesh}
-\input{OPT/StatPDE/Example5/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Compliance Minimization of a variable Thickness MBB-Beam}
-\label{OPT_Stat_MBB-Beam}
-\input{OPT/StatPDE/Example6/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Distributed control with a linear elliptic PDE using SNOPT}
-\label{OPT_Stat_Box_controlconstraints_SNOPT}
-\input{OPT/StatPDE/Example7/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Topology optimization of an MBB-Beam using SNOPT}
-\label{OPT_Stat_TopOpt_MBB_SNOPT}
-\input{OPT/StatPDE/Example8/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Parameter control with a non-linear PDE from FSI dynamics}
-\label{OPT_Stat_Param_Nonlin_FSI}
-%\input{OPT/StatPDE/Example5/content.tex}
-\clearpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Testing examples}
-The DOpE testsuite consists of regression tests. They are run to compare 
-the output to previous outputs. This is useful (necessary) after 
-changing programming code anywhere in the library. If a test
-succeeds, everything is fine in the library. If not, you should not
-check in your code into DOpE. Please make sure what is going wrong and WHY!
-
-Every command is computed via a Makefile. In the basic example 
-directory
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Where can I find the tests}
-In each example directory you find a subdirectory `Test'. Herein, you find 
-the parameter files for meshes (*.inp) and a param file (test.prm).
-Moreover, the executable is denoted by `test.sh'. Please make sure, that 
-the 
-\begin{verbatim}
-set never_write_list
-\end{verbatim}
-contains every possible output
-\begin{verbatim}
-Gradient;Hessian;Tangent;Residual;Update;Control;State
-\end{verbatim}
-That means, no solution files are written to the output.
-Recall, that we are just interested in terminal output that 
-is of course sufficient to verify the things.
-
-Hence, the results directory should be empty
-\begin{verbatim}
-set results_dir   = ./
-\end{verbatim}
-The rest in the param file must be identically the same as 
-in the dope.prm file in the parent directory.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{How to start testing?}
-You start testing by typing 
-\begin{verbatim}
-> ./test.sh Store
-\end{verbatim}
-in the terminal.
-
-After the run, you have to call 
-\begin{verbatim}
-> ./test.sh Test
-\end{verbatim}
-to compare your stored output. Of course, there should be no differences.
-
-The useful point is now the following. After implemention of 
-new pieces of code in the DOpE library or in the examples, you can
-run 
-\begin{verbatim}
-> ./test.sh Test
-\end{verbatim}
-Hereby, you compare your `new' output with the previous stored output.
-
-Attention: After changes you should NOT run again
-\begin{verbatim}
-> ./test.sh Store
-\end{verbatim}
-In that case, you overwrite your previous output.
-\cleardoublepage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\addcontentsline{toc}{chapter}{Index}
-\printindex
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\end{document}

### 2012-03-16 Examples/description.tex

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Introduction}
-In this report, we describe the DOpE
-\textit{Deal Optimization Environment} project. 
-Originally, the project was initiated in the year 2009 at the 
-University of Heidelberg (Germany) in the numerical analysis
-group of Rolf Rannacher.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The \textit{Deal Optimization Environment} (DOpE) project is mainly 
-based on the
-\textit{deal.II} finite element library which has been developed initially
-by W. Bangerth, R. Hartmann, and G. Kanschat. 
-In addition, some algorithmic aspects to compute optimization problems
-in which the state is given in terms of a partial
-differential equations are similar to algorithms of the
-Gascoigne/RoDoBo 
-project, which was initiated by 
-Roland Becker, Dominik Meidner,  and Boris Vexler. 
-Both projects have also been 
-initiated at the University of Heidelberg.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The aim of DOpE is to provide a software toolkit to solve forward PDE
-problems as well as optimal control problems constrained by PDE. The
-solution of a broad variety of PDE is possible in deal.II as well, but
-DOpE concentrates on a unified approach for both linear and nonlinear
-problems by interpreting every PDE problem as nonlinear and applying a
-Newton method to solve it. While deal.II leaves much of the work and many
-decisions to the user, DOpE intends to be user-optimized by delivering
-prefabricated tools which require from the user only adjustments connected
-to his specific problem. The solution of optimal control problems with PDE
-constraints is an innovation in the DOpE framework.
-The focus is on the numerical solution of both stationary and nonstationary
-problems which come from different application fields, like elasticity and
-plasticity, fluid dynamics, and fluid-structure interactions.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-At the present stage, we are able to provide the following features and 
-problems:
-\begin{itemize}
-\item Computation of forward problems in 1d, 2d, and 3d.
-\item Solution of stationary and nonstationary forward problems
-\item Various timestepping schemes (based on finite differences), 
-  such as forward Euler, backward Euler,
-  Crank-Nicolson, shifted Crank-Nicolson, and Fractional-Step-$\Theta$ scheme.
-\item Various finite elements taken from deal.ii.
-\item Computation of stationary optimal control problems.
-\item Solving of standard Laplace equation, Navier-Stokes, and 
-fluid-structure interaction. 
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In the present manual we give a short introduction to all example programs
-which can be found currently in DOpE. The organization of the examples is
-as follows: we distinguish between the mere solution of a PDE and the
-solution of an optimal control problem, and in each case we differentiate
-further between stationary and nonstationary problems. This leads to the
-following structure (we list problems treated in the examples in
-brackets):
-\begin{itemize}
-\item
-solution of PDE problem
- \begin{itemize}
-   \item
-     stationary PDE (Laplace equation, Stokes equations, FSI problems,
-elasticity and plasticity problems)
-   \item
-     nonstationary PDE (Navier-Stokes equations, FSI problem, Black-Scholes
-equation)
- \end{itemize}
-\item
-solution of PDE constrained optimal control problem
- \begin{itemize}
-   \item
-     stationary constraining PDE
-   \item
-     nonstationary constraining PDE
- \end{itemize}
-\end{itemize}
-The seventh stationary PDE example \ref{PDE_Stat_Laplace_2D} 
-(solution of the Laplace equation) is
-implemented analogously to the corresponding tutorial step in deal.II. By
-comparing the respective program codes, deal.II users should be able to
-catch the differences between the two software packages.
-In later examples we also provide algorithms for 'goal-oriented' automatic
-mesh adaption and error analysis.
-
-{\bf Comment on deal.II}
-
-The deal.II library is open
-source and can be downloaded from \texttt{http://www.dealii.org/}. On this
-homepage, one also finds lots of further information on deal.II as well as
-an extensive tutorial where many features of deal.II are discussed in a
-well-documented example framework. In order to use DOpE, it is highly
-recommendable to be roughly acquainted with deal.II.
+\include{Documentation/foreword.tex}
 

### 2012-03-16 Examples/Documentation/examples.tex
More structure for the documentation. And some more information on software requirements

--- 2012-03-16 Examples/description.tex
@@
-\documentclass[a4paper,cleardoubleempty]{scrreprt}
-\usepackage{amsmath,amssymb}
-\usepackage[pdftex,bookmarks=false,pdfpagelabels=true,plainpages=false]{hyperref}
-\usepackage{booktabs}
-\usepackage{makeidx}
-%\usepackage{sfmath}
-\newtheorem{Problem}{Problem}[section]
-
-\newcommand{\todo}[1]{\textbf{\textsc{\textcolor{black}{TODO: #1}}}}
-
-\makeindex
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-\title{DOpE - Deal Optimization Environment}
-\author{
-  Michael Geiger\thanks{University of Heidelberg, 
-    {michael.geiger@iwr.uni-heidelberg.de}}, 
-  Christian Goll\thanks{University of Heidelberg, 
-    {christian.goll@iwr.uni-heidelberg.de}}, 
-  Thomas Wick\thanks{University of Heidelberg, 
-    {thomas.wick@iwr.uni-heidelberg.de}}, 
-  Winnifried Wollner\thanks{University of Hamburg, 
-    {winnifried.wollner@math.uni-hamburg.de}}
-}
-
-\date{\today}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\maketitle
-\cleardoublepage\pagestyle{headings}
-\tableofcontents
-\cleardoublepage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\include{Documentation/foreword.tex}
-
-\cleardoublepage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \chapter{Example Handling, Creating new Examples}
+\label{chap:howtoex}
 To implement new examples or to use existing examples 
@@
 \end{verbatim}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Examples for PDE Solution}
-\label{PDE}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Stationary PDEs} 
-\label{PDE_Stat}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary Stokes Equations} 
-\label{PDE_Stat_Stokes}
-\input{PDE/StatPDE/Example1/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary FSI with INH Material} 
-\label{PDE_Stat_FSI_INH}
-\input{PDE/StatPDE/Example2/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary FSI with STVK Material} 
-\label{PDE_Stat_FSI_STVK}
-\input{PDE/StatPDE/Example3/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary Elasticity Benchmark} 
-\label{PDE_Stat_Elasticity}
-\input{PDE/StatPDE/Example4/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary Plasticity Benchmark} 
-\label{PDE_Stat_Plasticity}
-\input{PDE/StatPDE/Example5/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary Stokes Equations with periodic BC} 
-\label{PDE_Stat_Periodic_Stokes}
-\input{PDE/StatPDE/Example6/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Laplace Equation in 2D} 
-\label{PDE_Stat_Laplace_2D}
-\input{PDE/StatPDE/Example7/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Laplace Equation in 3D}
-\label{PDE_Stat_Laplace_3D}
-\input{PDE/StatPDE/Example8/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Nonstationary PDEs}
-\label{PDE_Instat}
-\input{PDE/InstatPDE/content.tex}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Nonstationary Stokes Equations}
-\label{PDE_Instat_Stokes}
-\input{PDE/InstatPDE/Example1/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Nonstationary FSI Problem}
-\label{PDE_Instat_FSI}
-\input{PDE/InstatPDE/Example2/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Black-Scholes Equation}
-\label{PDE_Instat_Black_Scholes}
-\input{PDE/InstatPDE/Example3/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Heat Equation in 1D}
-\label{PDE_Instat_Heat_1D}
-\input{PDE/InstatPDE/Example4/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Heat Equation in 2D with nonlinearity}
-\label{PDE_Instat_Heat_2D}
-\input{PDE/InstatPDE/Example5/content.tex}
-\cleardoublepage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Examples with Optimization}
-\label{OPT}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Subject to a Stationary PDE}
-\label{OPT_Stat}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Distributed control with a linear elliptic PDE}
-\label{OPT_Stat_Distrib_Lin_Ellipt}
-\input{OPT/StatPDE/Example1/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Parameter control with a linear elliptic PDE}
-\label{OPT_Stat_Param_Lin_Ellipt}
-\input{OPT/StatPDE/Example2/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Parameter control with a nonlinear PDE from fluid dynamcis}
-\label{OPT_Stat_Param_Nonlin_Fluid}
-\input{OPT/StatPDE/Example3/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Control in the dirichlet boundary values}
-\label{OPT_Stat_Dirichlet_Boundary}
-\input{OPT/StatPDE/Example4/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Distributed Control with Different Meshes for Control and State}
-\label{OPT_Stat_Distributed_MultiMesh}
-\input{OPT/StatPDE/Example5/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Compliance Minimization of a variable Thickness MBB-Beam}
-\label{OPT_Stat_MBB-Beam}
-\input{OPT/StatPDE/Example6/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Distributed control with a linear elliptic PDE using SNOPT}
-\label{OPT_Stat_Box_controlconstraints_SNOPT}
-\input{OPT/StatPDE/Example7/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Topology optimization of an MBB-Beam using SNOPT}
-\label{OPT_Stat_TopOpt_MBB_SNOPT}
-\input{OPT/StatPDE/Example8/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Parameter control with a non-linear PDE from FSI dynamics}
-\label{OPT_Stat_Param_Nonlin_FSI}
-%\input{OPT/StatPDE/Example5/content.tex}
-\clearpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Testing examples}
-The DOpE testsuite consists of regression tests. They are run to compare 
-the output to previous outputs. This is useful (necessary) after 
-changing programming code anywhere in the library. If a test
-succeeds, everything is fine in the library. If not, you should not
-check in your code into DOpE. Please make sure what is going wrong and WHY!
-
-Every command is computed via a Makefile. In the basic example 
-directory
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Where can I find the tests}
-In each example directory you find a subdirectory `Test'. Herein, you find 
-the parameter files for meshes (*.inp) and a param file (test.prm).
-Moreover, the executable is denoted by `test.sh'. Please make sure, that 
-the 
-\begin{verbatim}
-set never_write_list
-\end{verbatim}
-contains every possible output
-\begin{verbatim}
-Gradient;Hessian;Tangent;Residual;Update;Control;State
-\end{verbatim}
-That means, no solution files are written to the output.
-Recall, that we are just interested in terminal output that 
-is of course sufficient to verify the things.
-
-Hence, the results directory should be empty
-\begin{verbatim}
-set results_dir   = ./
-\end{verbatim}
-The rest in the param file must be identically the same as 
-in the dope.prm file in the parent directory.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{How to start testing?}
-You start testing by typing 
-\begin{verbatim}
-> ./test.sh Store
-\end{verbatim}
-in the terminal.
-
-After the run, you have to call 
-\begin{verbatim}
-> ./test.sh Test
-\end{verbatim}
-to compare your stored output. Of course, there should be no differences.
-
-The useful point is now the following. After implemention of 
-new pieces of code in the DOpE library or in the examples, you can
-run 
-\begin{verbatim}
-> ./test.sh Test
-\end{verbatim}
-Hereby, you compare your `new' output with the previous stored output.
-
-Attention: After changes you should NOT run again
-\begin{verbatim}
-> ./test.sh Store
-\end{verbatim}
-In that case, you overwrite your previous output.
-\cleardoublepage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\addcontentsline{toc}{chapter}{Index}
-\printindex
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\end{document}

### 2012-03-16 Examples/Documentation/foreword.tex
More structure for the documentation. And some more information on software requirements

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The \textit{Deal Optimization Environment} (DOpE) project is mainly 
-based on the
-\textit{deal.II} finite element library which has been developed initially
-by W. Bangerth, R. Hartmann, and G. Kanschat. 
-In addition, some algorithmic aspects to compute optimization problems
-in which the state is given in terms of a partial
-differential equations are similar to algorithms of the
-Gascoigne/RoDoBo 
-project, which was initiated by 
-Roland Becker, Dominik Meidner,  and Boris Vexler. 
-Both projects have also been 
-initiated at the University of Heidelberg.
+The \textit{Deal Optimization Environment} (DOpE) project is 
+based on the \textit{deal.II} finite element library which has been developed
+ initially by W. Bangerth, R. Hartmann, and G. Kanschat. \marginpar{Citation}
+Its main feature is to give a unified interface to highlevel algorithms such as 
+time stepping methods, nonlinear solvers and optimization routines. 
+We aim that the user should only need to write those parts
+of the code that are problem dependent while all invariant parts of the algorithms
+should be reusable without any need for further coding.
+In particular, the user should be able to switch between various different 
+algorithms without the need to rewrite the problem dependent code, though he or she will
+have to replace the algorithm object with an other one.
+
+The authors acknwoledge their past experience as well as discussions with 
+the authors of the libraries 
+Gascoigne/RoDoBo project, which was initiated by 
+Roland Becker, Dominik Meidner,  and Boris Vexler. \marginpar{Citation} 
+From which some of the ideas to modularize the algorithms have arisen.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-At the present stage, we are able to provide the following features and 
-problems:
+At the present stage the following features are supported by the library
 \begin{itemize}
-\item Computation of forward problems in 1d, 2d, and 3d.
-\item Solution of stationary and nonstationary forward problems
+\item Solution of stationary and nonstationary PDEs in 1d, 2d, and 3d.
 \item Various timestepping schemes (based on finite differences), 
@@
   Crank-Nicolson, shifted Crank-Nicolson, and Fractional-Step-$\Theta$ scheme.
-\item Various finite elements taken from deal.ii.
-\item Computation of stationary optimal control problems.
-\item Solving of standard Laplace equation, Navier-Stokes, and 
-fluid-structure interaction. 
+\item All finite elements of from deal.ii including hp-support.
+\item Several examples showing the solution of several PDEs including
+   Poisson, Navier-Stokes, Plasticity and fluid-structure interaction problems. 
+\item Self written linesearch and trustregion newton algorithms for the 
+   solution of optimization problems with PDEs. (cp. Nocedal \& Wright\marginpar{Citation})
+\item Interface to SNOPT for the solution of optimization problems with PDEs and
+  additional other constraints.
+\item Several examples showing how to solve various kinds of optimization problems
+  involving stationary PDE constraints.
+\item Mesh adaptation.
+\item Different spatial triangulations for control and state variables.
 \end{itemize}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In the present manual we give a short introduction to all example programs
-which can be found currently in DOpE. The organization of the examples is
-as follows: we distinguish between the mere solution of a PDE and the
-solution of an optimal control problem, and in each case we differentiate
-further between stationary and nonstationary problems. This leads to the
-following structure (we list problems treated in the examples in
-brackets):
-\begin{itemize}
-\item
-solution of PDE problem
- \begin{itemize}
-   \item
-     stationary PDE (Laplace equation, Stokes equations, FSI problems,
-elasticity and plasticity problems)
-   \item
-     nonstationary PDE (Navier-Stokes equations, FSI problem, Black-Scholes
-equation)
- \end{itemize}
-\item
-solution of PDE constrained optimal control problem
- \begin{itemize}
-   \item
-     stationary constraining PDE
-   \item
-     nonstationary constraining PDE
- \end{itemize}
-\end{itemize}
-The seventh stationary PDE example \ref{PDE_Stat_Laplace_2D} 
-(solution of the Laplace equation) is
-implemented analogously to the corresponding tutorial step in deal.II. By
-comparing the respective program codes, deal.II users should be able to
-catch the differences between the two software packages.
-In later examples we also provide algorithms for 'goal-oriented' automatic
-mesh adaption and error analysis.
+The rest of this document is structured as follows: We start with an introduction in
+Chapter~\ref{chap:intro} where you will learn what is needed to run {\tt DOpElib}. 
+Further you will learn what problems we can solve and how all the different classes 
+work together for this purpose. This should help you figure out what the different classes
+do if you are in need of writing your own algorithm.
 
-{\bf Comment on deal.II}
+Then assuming that you can work to your satisfaction with the algorithms already implemented
+we will show you how to create your own running example in Chapter~\ref{chap:howtoex}.
+This will be followed by a detailed description of all examples already shipped with 
+the library. You can find the examples for the solution of PDEs in Chaper~\ref{PDE}
+and those for the solution of optimization problems with PDEs in Chapter~\ref{OPT}.
 
-The deal.II library is open
-source and can be downloaded from \texttt{http://www.dealii.org/}. On this
-homepage, one also finds lots of further information on deal.II as well as
-an extensive tutorial where many features of deal.II are discussed in a
-well-documented example framework. In order to use DOpE, it is highly
-recommendable to be roughly acquainted with deal.II.
+These notes conclude with a section that explains how we do automated testing of the 
+implementation in Chapter~\ref{chap:test}. This chapter will be of interest only if you 
+are trying to implement some new features to the library so that you can check that 
+the new code did not break anything.

### 2012-03-16 Examples/Documentation/introduction.tex
More structure for the documentation. And some more information on software requirements

\chapter{Introduction}\label{chap:intro}
\section{Developers}
The library is currently maintained by 
\begin{itemize}
  \item Christian Goll (University of Heidelberg)
  \item Thomas Wick (University of Heidelberg)
  \item Winnifried Wollner (University of Hamburg)
\end{itemize}

However, there are more highly appreciated contributions
made by (in alphabetical order)
\begin{itemize}
  \item Michael Geiger (Examples for Plasticity, and Documentation of several PDE-Examples)
\end{itemize}

\section{Software requirements}

\subsection{deal.ii}
This library is based upon {\tt deal.ii} hence in order to run 
{\tt DOpElib} you need a running copy of {\tt deal.ii}.

The deal.II library is open source and is freely availiable for noncommercial project.
It can be downloaded from \url{http://www.dealii.org/}. On this
homepage, one also finds lots of further information on deal.II as well as
an extensive tutorial where many features of deal.II are discussed in a
well-documented example framework. In order to use DOpE, it is highly
recommendable to be roughly acquainted with deal.II.

\subsection{SNOPT}
If you would like to use the features offered in our SNOPT wrapper. You will 
need to obtain a licence for {\tt SNOPT} 
\url{http://www.sbsi-sol-optimize.com/asp/sol_product_snopt.htm}.
Unfortuately this is at present not availiable for free, but you should check if 
there is a department license already availiable.

\section{How this library is struktured}
\marginpar{TODO!}

### 2012-03-16 Examples/Documentation/tests.tex
More structure for the documentation. And some more information on software requirements

--- 2012-03-16 Examples/description.tex
@@
-\documentclass[a4paper,cleardoubleempty]{scrreprt}
-\usepackage{amsmath,amssymb}
-\usepackage[pdftex,bookmarks=false,pdfpagelabels=true,plainpages=false]{hyperref}
-\usepackage{booktabs}
-\usepackage{makeidx}
-%\usepackage{sfmath}
-\newtheorem{Problem}{Problem}[section]
-
-\newcommand{\todo}[1]{\textbf{\textsc{\textcolor{black}{TODO: #1}}}}
-
-\makeindex
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{document}
-\title{DOpE - Deal Optimization Environment}
-\author{
-  Michael Geiger\thanks{University of Heidelberg, 
-    {michael.geiger@iwr.uni-heidelberg.de}}, 
-  Christian Goll\thanks{University of Heidelberg, 
-    {christian.goll@iwr.uni-heidelberg.de}}, 
-  Thomas Wick\thanks{University of Heidelberg, 
-    {thomas.wick@iwr.uni-heidelberg.de}}, 
-  Winnifried Wollner\thanks{University of Hamburg, 
-    {winnifried.wollner@math.uni-hamburg.de}}
-}
-
-\date{\today}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\maketitle
-\cleardoublepage\pagestyle{headings}
-\tableofcontents
-\cleardoublepage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\include{Documentation/foreword.tex}
-
-\cleardoublepage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Example Handling, Creating new Examples}
-To implement new examples or to use existing examples 
-from the library for own research, the user 
-can simply copy an existing example. In this 
-new example, own code and changes can be compiled.
-
-Adding a new regression test to the repository 
-needs some concentration of the user. In the 
-following, we explain how to do this and what 
-\textit{must} be taken into account:
-\begin{verbatim}
-cp -r Example1 ExampleNew
-cd ExampleNew
-rm -r CVS
-cd Test
-rm -r CVS
-\end{verbatim}
-It is important to remove the repository information,
-which is stored in the directory CVS.
-
-Then, we have to consider the following step. In order to 
-run the regression tests, we start the executable file
-of the present example. This executable comes from the 
-Makefile. For example in the Makefile of 
-Example \ref{PDE_Stat_Stokes}, 
-\begin{verbatim}
-target   = $(BINDIR)/DOpE-PDE-StatPDE-Example1-$
-           (dope_dimension)d-$(deal_II_dimension)d
-\end{verbatim}
-and concretely:
-\begin{verbatim}
-DOpE-PDE-StatPDE-Example1-2d-2d
-\end{verbatim}
-The aforementioned expression is the name of 
-the executable that is required to run the 
-regression test. Hence, in 
-\begin{verbatim}
-PDE/StatPDE/Example1/Test> emacs test.sh
-\end{verbatim}
-you find two times the line
-\begin{verbatim}
-echo "Running Program 
-     ../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm"
-     (../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm 2>&1) 
-     > /dev/null
-\end{verbatim} 
-It is really important when copying an existing example,
-to change this information. 
-First, we have to change the target in the Makefile
-\begin{verbatim}
-target   = $(BINDIR)/DOpE-PDE-StatPDE-ExampleNew-$
-           (dope_dimension)d-$(deal_II_dimension)d
-\end{verbatim}
-Second, we have to replace 
-Example1 by ExampleNew such that 
-\begin{verbatim}
-echo "Running Program 
-     ../../../../../bin/DOpE-PDE-StatPDE-ExampleNew-2d-2d test.prm"
-     (../../../../../bin/DOpE-PDE-StatPDE-ExampleNew-2d-2d test.prm 2>&1) 
-     > /dev/null
-\end{verbatim} 
-Please, also change the dimension parameters 
-\begin{verbatim}
--2d-2d
-\end{verbatim} 
-Now, the user is prepared to change any information in the 
-\begin{verbatim}
-main.cc, localpde.h, functionals.h, localfunctional.h, etc
-\end{verbatim}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Examples for PDE Solution}
-\label{PDE}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Stationary PDEs} 
-\label{PDE_Stat}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary Stokes Equations} 
-\label{PDE_Stat_Stokes}
-\input{PDE/StatPDE/Example1/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary FSI with INH Material} 
-\label{PDE_Stat_FSI_INH}
-\input{PDE/StatPDE/Example2/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary FSI with STVK Material} 
-\label{PDE_Stat_FSI_STVK}
-\input{PDE/StatPDE/Example3/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary Elasticity Benchmark} 
-\label{PDE_Stat_Elasticity}
-\input{PDE/StatPDE/Example4/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary Plasticity Benchmark} 
-\label{PDE_Stat_Plasticity}
-\input{PDE/StatPDE/Example5/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary Stokes Equations with periodic BC} 
-\label{PDE_Stat_Periodic_Stokes}
-\input{PDE/StatPDE/Example6/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Laplace Equation in 2D} 
-\label{PDE_Stat_Laplace_2D}
-\input{PDE/StatPDE/Example7/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Laplace Equation in 3D}
-\label{PDE_Stat_Laplace_3D}
-\input{PDE/StatPDE/Example8/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Nonstationary PDEs}
-\label{PDE_Instat}
-\input{PDE/InstatPDE/content.tex}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Nonstationary Stokes Equations}
-\label{PDE_Instat_Stokes}
-\input{PDE/InstatPDE/Example1/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Nonstationary FSI Problem}
-\label{PDE_Instat_FSI}
-\input{PDE/InstatPDE/Example2/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Black-Scholes Equation}
-\label{PDE_Instat_Black_Scholes}
-\input{PDE/InstatPDE/Example3/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Heat Equation in 1D}
-\label{PDE_Instat_Heat_1D}
-\input{PDE/InstatPDE/Example4/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Heat Equation in 2D with nonlinearity}
-\label{PDE_Instat_Heat_2D}
-\input{PDE/InstatPDE/Example5/content.tex}
-\cleardoublepage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Examples with Optimization}
-\label{OPT}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Subject to a Stationary PDE}
-\label{OPT_Stat}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Distributed control with a linear elliptic PDE}
-\label{OPT_Stat_Distrib_Lin_Ellipt}
-\input{OPT/StatPDE/Example1/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Parameter control with a linear elliptic PDE}
-\label{OPT_Stat_Param_Lin_Ellipt}
-\input{OPT/StatPDE/Example2/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Parameter control with a nonlinear PDE from fluid dynamcis}
-\label{OPT_Stat_Param_Nonlin_Fluid}
-\input{OPT/StatPDE/Example3/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Control in the dirichlet boundary values}
-\label{OPT_Stat_Dirichlet_Boundary}
-\input{OPT/StatPDE/Example4/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Distributed Control with Different Meshes for Control and State}
-\label{OPT_Stat_Distributed_MultiMesh}
-\input{OPT/StatPDE/Example5/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Compliance Minimization of a variable Thickness MBB-Beam}
-\label{OPT_Stat_MBB-Beam}
-\input{OPT/StatPDE/Example6/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Distributed control with a linear elliptic PDE using SNOPT}
-\label{OPT_Stat_Box_controlconstraints_SNOPT}
-\input{OPT/StatPDE/Example7/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Topology optimization of an MBB-Beam using SNOPT}
-\label{OPT_Stat_TopOpt_MBB_SNOPT}
-\input{OPT/StatPDE/Example8/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Parameter control with a non-linear PDE from FSI dynamics}
-\label{OPT_Stat_Param_Nonlin_FSI}
-%\input{OPT/StatPDE/Example5/content.tex}
-\clearpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Testing examples}
+\chapter{Testing examples}\label{chap:test}
 The DOpE testsuite consists of regression tests. They are run to compare 
@@
 In that case, you overwrite your previous output.
-\cleardoublepage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\addcontentsline{toc}{chapter}{Index}
-\printindex
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\end{document}

### 2012-03-16 Examples/description.tex
More structure for the documentation. And some more information on software requirements

@@
 \usepackage{makeidx}
+\usepackage{url}
 %\usepackage{sfmath}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\include{Documentation/foreword.tex}
-
-\cleardoublepage
+\input{Documentation/foreword.tex}
+\input{Documentation/introduction.tex}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Example Handling, Creating new Examples}
-To implement new examples or to use existing examples 
-from the library for own research, the user 
-can simply copy an existing example. In this 
-new example, own code and changes can be compiled.
-
-Adding a new regression test to the repository 
-needs some concentration of the user. In the 
-following, we explain how to do this and what 
-\textit{must} be taken into account:
-\begin{verbatim}
-cp -r Example1 ExampleNew
-cd ExampleNew
-rm -r CVS
-cd Test
-rm -r CVS
-\end{verbatim}
-It is important to remove the repository information,
-which is stored in the directory CVS.
-
-Then, we have to consider the following step. In order to 
-run the regression tests, we start the executable file
-of the present example. This executable comes from the 
-Makefile. For example in the Makefile of 
-Example \ref{PDE_Stat_Stokes}, 
-\begin{verbatim}
-target   = $(BINDIR)/DOpE-PDE-StatPDE-Example1-$
-           (dope_dimension)d-$(deal_II_dimension)d
-\end{verbatim}
-and concretely:
-\begin{verbatim}
-DOpE-PDE-StatPDE-Example1-2d-2d
-\end{verbatim}
-The aforementioned expression is the name of 
-the executable that is required to run the 
-regression test. Hence, in 
-\begin{verbatim}
-PDE/StatPDE/Example1/Test> emacs test.sh
-\end{verbatim}
-you find two times the line
-\begin{verbatim}
-echo "Running Program 
-     ../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm"
-     (../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm 2>&1) 
-     > /dev/null
-\end{verbatim} 
-It is really important when copying an existing example,
-to change this information. 
-First, we have to change the target in the Makefile
-\begin{verbatim}
-target   = $(BINDIR)/DOpE-PDE-StatPDE-ExampleNew-$
-           (dope_dimension)d-$(deal_II_dimension)d
-\end{verbatim}
-Second, we have to replace 
-Example1 by ExampleNew such that 
-\begin{verbatim}
-echo "Running Program 
-     ../../../../../bin/DOpE-PDE-StatPDE-ExampleNew-2d-2d test.prm"
-     (../../../../../bin/DOpE-PDE-StatPDE-ExampleNew-2d-2d test.prm 2>&1) 
-     > /dev/null
-\end{verbatim} 
-Please, also change the dimension parameters 
-\begin{verbatim}
--2d-2d
-\end{verbatim} 
-Now, the user is prepared to change any information in the 
-\begin{verbatim}
-main.cc, localpde.h, functionals.h, localfunctional.h, etc
-\end{verbatim}
+\input{Documentation/examples.tex}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Testing examples}
-The DOpE testsuite consists of regression tests. They are run to compare 
-the output to previous outputs. This is useful (necessary) after 
-changing programming code anywhere in the library. If a test
-succeeds, everything is fine in the library. If not, you should not
-check in your code into DOpE. Please make sure what is going wrong and WHY!
-
-Every command is computed via a Makefile. In the basic example 
-directory
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Where can I find the tests}
-In each example directory you find a subdirectory `Test'. Herein, you find 
-the parameter files for meshes (*.inp) and a param file (test.prm).
-Moreover, the executable is denoted by `test.sh'. Please make sure, that 
-the 
-\begin{verbatim}
-set never_write_list
-\end{verbatim}
-contains every possible output
-\begin{verbatim}
-Gradient;Hessian;Tangent;Residual;Update;Control;State
-\end{verbatim}
-That means, no solution files are written to the output.
-Recall, that we are just interested in terminal output that 
-is of course sufficient to verify the things.
-
-Hence, the results directory should be empty
-\begin{verbatim}
-set results_dir   = ./
-\end{verbatim}
-The rest in the param file must be identically the same as 
-in the dope.prm file in the parent directory.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{How to start testing?}
-You start testing by typing 
-\begin{verbatim}
-> ./test.sh Store
-\end{verbatim}
-in the terminal.
-
-After the run, you have to call 
-\begin{verbatim}
-> ./test.sh Test
-\end{verbatim}
-to compare your stored output. Of course, there should be no differences.
-
-The useful point is now the following. After implemention of 
-new pieces of code in the DOpE library or in the examples, you can
-run 
-\begin{verbatim}
-> ./test.sh Test
-\end{verbatim}
-Hereby, you compare your `new' output with the previous stored output.
-
-Attention: After changes you should NOT run again
-\begin{verbatim}
-> ./test.sh Store
-\end{verbatim}
-In that case, you overwrite your previous output.
-\cleardoublepage
+\input{Documentation/tests.tex}
 

### 2012-03-16 Examples/Documentation/examples.tex
Augmented the Documentation. Please read carefully to see if this is fine

@@
 cd ExampleNew
-rm -r CVS
+rm -rf .svn
 cd Test
-rm -r CVS
+rm -rf .svn
 \end{verbatim}
 It is important to remove the repository information,
-which is stored in the directory CVS.
+which is stored in the directory .svn.
 

### 2012-03-16 Examples/Documentation/foreword.tex
Augmented the Documentation. Please read carefully to see if this is fine

@@
  initially by W. Bangerth, R. Hartmann, and G. Kanschat. \marginpar{Citation}
-Its main feature is to give a unified interface to highlevel algorithms such as 
+Its main feature is to give a unified interface to high level algorithms such as 
 time stepping methods, nonlinear solvers and optimization routines. 
@@
 
-The authors acknwoledge their past experience as well as discussions with 
+The authors acknowledge their past experience as well as discussions with 
 the authors of the libraries 
@@
 \item Solution of stationary and nonstationary PDEs in 1d, 2d, and 3d.
-\item Various timestepping schemes (based on finite differences), 
+\item Various time stepping schemes (based on finite differences), 
   such as forward Euler, backward Euler,
@@
    Poisson, Navier-Stokes, Plasticity and fluid-structure interaction problems. 
-\item Self written linesearch and trustregion newton algorithms for the 
-   solution of optimization problems with PDEs. (cp. Nocedal \& Wright\marginpar{Citation})
+\item Self written line search and trust region newton algorithms for the 
+   solution of optimization problems with PDEs. (cf. Nocedal \& Wright\marginpar{Citation})
 \item Interface to SNOPT for the solution of optimization problems with PDEs and
@@
 This will be followed by a detailed description of all examples already shipped with 
-the library. You can find the examples for the solution of PDEs in Chaper~\ref{PDE}
+the library. You can find the examples for the solution of PDEs in Chapter~\ref{PDE}
 and those for the solution of optimization problems with PDEs in Chapter~\ref{OPT}.

### 2012-03-16 Examples/Documentation/introduction.tex
Augmented the Documentation. Please read carefully to see if this is fine

@@
 \section{Software requirements}
+The library DOpE has been tested to work on both Linux and 
+MAC OSX. It requires a recent {\tt g++} (at least 4.4.3) due to 
+some new {\tt C++} features implemented in {\tt C++0x} and {\tt C++11}.
 
@@
 
-The deal.II library is open source and is freely availiable for noncommercial project.
+The deal.II library is open source and is freely available for noncommercial project.
 It can be downloaded from \url{http://www.dealii.org/}. On this
@@
 If you would like to use the features offered in our SNOPT wrapper. You will 
-need to obtain a licence for {\tt SNOPT} 
+need to obtain a license for {\tt SNOPT} 
 \url{http://www.sbsi-sol-optimize.com/asp/sol_product_snopt.htm}.
-Unfortuately this is at present not availiable for free, but you should check if 
-there is a department license already availiable.
+Unfortunately this is at present not available for free, but you should 
+check if 
+there is a department license already available.
 
-\section{How this library is struktured}
-\marginpar{TODO!}
+\section{How this library is structured}
+This library is designed to allow easy implementation and numerical solutions 
+of problems involving partial differential equations (PDEs). The easiest case 
+is that of a PDE in weak form to find some $u$
+\[
+a(u)(\phi) = 0 \quad \forall \phi \in V
+\]
+with some appropriate space $V$.
+More complex cases involve optimization problems given in the form (OPT)
+\begin{align*}
+\min\;&J(q,u) \\
+  &\text{s.t.}\; a(q,u) = 0 \quad \forall \phi\\
+  &a \le q \le b\\
+  &g(q,u) \le 0  
+\end{align*}
+where $u$ is a FE-function and $q$ can either be a FE-function or some 
+fixed number of parameters.
+
+\subsection{Problem description}
+In order to allow our algorithms the automatic assembly of all required 
+data we need to have some container which contains the complete problem 
+description in a common data format. For this we have the following 
+classes in {\tt DOpEsrc/container}
+\begin{itemize}
+  \item {\tt pdeproblemcontainer.h} Is used to describe  stationary PDE problems.
+  \item {\tt instatpdeproblemcontainer.h} {\bf TODO is still missing but will be used for nonstationary problems.} This will be implemented once we have nonstationary optimization problems running to avoid error duplication in the coding process.
+  \item {\tt optproblem.h} Is used to describe  OPT problems governed by 
+    stationary PDEs. {\bf TODO should be renamed to optproblemcontainer.h including the class name!}
+  \item {\tt instatoptproblemcontainer.h} Is used to describe  OPT problems governed by nonstationary PDEs.
+The only difference to the stationary case is that we need to specify a time-stepping method.
+\end{itemize}
+In order to fill these containers there are two things to be done,
+first we need to actually write some data like $a$, $J$, $\ldots$
+which describe the problem. Then we have to select some numerical 
+algorithm components like finite elements, linear solvers $\ldots$.
+The latter ones should be written such that when exchanging these components
+none of the problem descriptions should require changes. 
+Note that it still may be necessary to write some additional descriptions, 
+e.g., if you solve the PDE with a fix point iteration you don't need derivatives
+but if you want to use Newton's method you then have to provide some.
+
+We will start by discussing the problem description components implemented so far
+
+\subsubsection{Problem specific classes}
+The following classes are used to describe the problem and will usually require 
+some implementation.
+
+\begin{itemize}
+  \item {\tt basic/constraints.h} This is used by the spacetimehandlers to 
+    compute the number of constraints from the control and state vectors. 
+    It must not be reimplemented by the user, but needs to be properly 
+    initialized if OPT is used with box control constraints or $g(q,u) \le 0$.
+  \item {\tt interfaces/functionalinterface.h} This gives an interface 
+    for the functional $J$ and any other functional you may want to evaluate.
+    In general this can be used as a base class to write your own functionals 
+    in examples. We note that we only need to write the integrands on 
+    elements or faces the loop over elements will be taken care of else where.
+  \item {\tt interfaces/constraintinterface.h} This gives an interface for both 
+    the control box constraints as well as the general constraint $g \le 0$. This 
+    needs to be specified if constraints are to be used. If they are not needed 
+    a default class {\tt problemdata/noconstraints.h} can be used. We note that we only 
+    need to write the integrands on 
+    elements or faces the loop over elements will be taken care of else where.
+  \item {\tt interfaces/pdeinterface.h} This defines an interface for the 
+    partial differential equation $a(q,u)(\phi) = 0$. This needs to be written
+    by the user. We note that we only need to write the integrands on 
+    elements or faces the loop over elements will be taken care of else where.
+  \item {\tt interfaces/dirichletdatainterface.h} This gives an interface to the 
+    Dirichlet data for a problem. If the Dirichlet data are simply a function 
+    (and do not depend on the control $q$ one can use the default class
+    {\tt problemdata/simpledirichletdata.h}.
+\end{itemize}
+
+\subsubsection{Numeric components}
+These are the components from which a user needs to select some in order to actually 
+solve the given problem. They will not require any rewriting, but sometimes it is 
+advisable to write other than the default parameter into the param file for the 
+solution.
+
+First we need to select a method how to handle all dofs in space and time.
+\begin{itemize}
+\item {\tt basic/spacetimehandler\underline{ }base.h} This class is used to define 
+  an interface to the dimension independent functionality of all space time dof handlers. 
+\item {\tt basic/statespacetimehandler.h} Another intermediate interface class which adds 
+  the dimension dependent functionality if only the variable $u$ is considered, i.e., a 
+  PDE problem.
+\item {\tt basic/spacetimehandler.h } Same as above but with both $q$ and $u$, i.e., for
+  OPT problems.
+\item {\tt basic/mol\underline{ }statespacetimehandler.h} Implementation of a method of 
+  line space time dof handler for PDE problems. It has only one spatial 
+  dofhandler which is used for all time intervals.
+\item {\tt basic/mol\underline{ }spacetimehandler.h} Same as above for OPT problems.
+  A separate spatial dof handler for each of the variables $q$ and $u$ is maintained 
+  but only one triangulation.
+\item {\tt basic/mol\underline{ }multimesh\underline{ }spacetimehandler.h}
+  Same as above, but now in addition the triangulations for $q$ and $u$ can be refined
+  separately from one common initial coarse triangulation. Note that this will
+  in addition require the use of the multimesh version for integrator and 
+  face- as well as celldatacontainer.
+\end{itemize}
+Note that we use these for stationary problems as well, but then you don't have to specify
+any time information.
+
+Second you will need to specify some container classes to be used to 
+pass data between objects. At present you don't have much choice, but you may wish 
+to reimplement some of these if you need data that is not currently included in 
+the containers.
+\begin{itemize}
+\item {\tt container/celldatacontainer.h} This object is used to pass data 
+  given on the current element (cell) of the mesh to the functions in PDE, functional, 
+  $\ldots$. 
+\item {\tt container/facedatacontainer.h} This object is used to pass data 
+  given on the current face of the mesh to the functions in PDE, functional, 
+  $\ldots$. 
+\item {\tt container/multimesh\underline{ }celldatacontainer.h} This is the same as the 
+  celldatacontainer, but it
+  is capable to handle data defined on an alternative triangulation.
+\item {\tt container/multimesh\underline{ }facedatacontainer.h} This is the same as the
+  facedatacontainer, but it
+  is capable to handle data defined on an alternative triangulation.
+\item {\tt container/integratordatacontainer.h} This contains some data that 
+  should be passed to the integrator like quadrature formulas and the above cell and 
+  face data container.
+\end{itemize}
+
+Third, at least for nonstationary PDEs we need to select a time stepping scheme
+the file names of which are mostly self explanatory:
+\begin{itemize}
+\item {\tt include/forward\underline{ }euler\underline{ }problem.h}
+\item {\tt include/shifted\underline{ }crank\underline{ }nicolson\underline{ }problem.h}
+\item {\tt include/backward\underline{ }euler\underline{ }problem.h}
+\item {\tt include/fractional\underline{ }step\underline{ }theta\underline{ }problem.h} Note that the use of this scheme requires a special Newton solver!
+\item {\tt include/crank\underline{ }nicolson\underline{ }problem.h}
+\end{itemize}
+
+Finally, we need to select a way how to integrate and solve linear and nonlinear equations
+\begin{itemize}
+\item {\tt templates/integrator.h} This class computes integrals over a given 
+  triangulation (including its faces).
+\item {\tt templates/integrator\underline{ }multimesh.h} The same as above but it is 
+  possible that some of the FE functions are defined on an other triangulation 
+  as long as the have a common coarse triangulation.
+\item {\tt templates/integratormixeddims.h} This is used to compute integral which 
+  are given in another (larger) dimension than the current variable. (This is exclusively
+  used if the control variable is given by some parameters. Which means {\tt dopedim == 0}). 
+\item {\tt templates/cglinearsolver.h} This is a wrapper for the cg solver implemented in 
+  {\tt deal.ii}. The solver will build and store the stiffness matrix for the PDE.
+\item {\tt templates/gmreslinearsolver.h} This is a wrapper for the gmres solver 
+  implemented in {\tt deal.ii}. The solver will build and store the stiffness matrix 
+  for the PDE.
+\item {\tt templates/directlinearsolver.h} This is a wrapper for the direct solver 
+  implemented in {\tt deal.ii} using {\tt UMFPACK}. 
+  The solver will build and store the stiffness matrix for the PDE.
+\item {\tt templates/voidlinearsolver.h} This is a wrapper for certain cases when we 
+  know that the matrix to be inverted is the identity. It simply copies the rhs to the
+  lhs. This is only needed for compatibility reasons some other components.
+\item {\tt templates/newtonsolver.h} This solves some nonlinear equation using a 
+  line-search Newton method.
+\item {\tt templates/newtonsolvermixeddims.h} The same but in the case when there is 
+  another variable in a (larger) dimension is involved. See 
+  {\tt integratormixeddims.h}.
+\item {\tt templates/instat\underline{ }step\underline{ }newtonsolver.h} This is a 
+  Newton method as above to invert the next time-step. It differs from the plain vanilla
+  version in that it computes certain data from the previous time step only once 
+  and not in every Newton iteration.
+\item {\tt templates/fractional\underline{ }step\underline{ }theta\underline{ }step\underline{ }newtonsolver.h} This is the Newton solver for the time step in a 
+  fractional-step-theta scheme. It combines the computation of all three sub steps.
+\end{itemize}
+
+\subsection{Reduced problems (Solve the PDE)}
+At times it is nice to remove the PDE constraint. This is handled by so called reduced 
+problems. This means that the reduced problem implicitly solves the PDE whenever required
+and eliminates the variable $u$ from the problem.
+\begin{itemize}
+\item {\tt reducedproblems/statpdeproblem.h} This is used to remove the variable $u$ in 
+  a stationary PDE problem. This means that call the method \\
+  {\tt StatPDEProblem::ComputeReducedFunctionals} will evaluate the functionals 
+  defined in the problem description, i.e., in {\tt PDEProblemContainer}, in the 
+  solution of the given PDE.
+\item {\tt reducedproblems/statreducedproblem.h} This eliminates $u$ from the OPT
+  problem with a stationary PDE.
+\item {\tt reducedproblems/instatreducedproblem.h} The same as above but for a
+  nonstationary PDE. {\bf FIXME there is something wrong in this file see FIXME 
+    comment in the source.}
+\item {\tt reducedproblems/voidreducedproblem.h} A wrapper file that eliminates $u$ 
+  if it is not present anyways. This is used so that we can use the same routines to 
+  solve problems that have no PDE constraint.
+\end{itemize}
+
+\subsection{Optimization algorithms}
+Now, in order to solve optimization algorithms we need to define some algorithms.
+At present we offer a selection of algorithms that solve the reduced optimization 
+problem where the PDE constraint has been eliminated as explained in the previous section
+\begin{itemize}
+\item {\tt opt\underline{ }algorithms/reducedalgorithm.h} An interface for all 
+  optimization problems in the reduced formulation. It offers some test functionality
+  to assert that the derivatives of the problem are computed correctly.
+\item {\tt opt\underline{ }algorithms/reducednewtonalgorithm.h}
+  A line-search Newton algorithm using a cg method to invert the reduced hessian. 
+  Implementation ignores any additional constraints.
+\item {\tt opt\underline{ }algorithms/reducedtrustregionnewton.h}
+  A trust region Newton algorithm using a cg method to invert the reduced hessian.
+  Implementation ignores any additional constraints.
+\item {\tt opt\underline{ }algorithms/reduced\underline{ }snopt\underline{ }algorithm.h}
+  An algorithm to solve reduced optimization problems with additional control constraints.
+  ((reduced) state constraints are not yet implemented.)
+\item {\tt opt\underline{ }algorithms/reducednewtonalgorithmwithinverse.h}
+  Line-search Newton algorithm that assumes there exists a method in the reduced problem
+  that can invert the reduced hessian. (This usually makes sense only if there is no 
+  PDE constraint.)
+\item {\tt opt\underline{ }algorithms/generalized\underline{ }mma\underline{ }algorithm.h}
+  An implementation of the MMA-Algorithm for structural optimization using an augmented
+  Lagrangian formulation for the subproblems. The subproblem is implemented using the 
+  special purpose file
+  {\tt include/augmentedlagrangianproblem.h}.
+\end{itemize}
+
+\subsection{Other Components}
+Beyond these clearly structured groups before there are some classes remaining that
+do not fit the above but are important for the user to know:
+\begin{itemize}
+\item {\tt include/statevector.h} This stores all dofs in space and time for the state 
+  variable $u$. It is possible to select whether all this should be kept in memory or 
+  or unused parts can be written to the hard disk.
+\item {\tt include/controlvector.h} This stores all dofs in space and time for the 
+  control variable $q$. At present no time dependence is implemented.
+\item {\tt include/constraintvector.h} This stores all dofs in space and time for the 
+  non PDE constraints (and corresponding multipliers). 
+  At present no time dependence is implemented.
+\item {\tt include/parameterreader.h} This file is used to define a parameter reader
+  that is used to read run time parameters from a given file.
+\item {\tt include/dopeexception.h} Defines some Exceptions that are thrown by the program
+  should it encounter any unexpected errors.
+\item {\tt include/dopeexceptionhandler.h} This class is used to write information 
+  contained in the exceptions to the output in a uniform manner.
+\item {\tt include/outputhandler.h} This file defines an outputhandler object which 
+  can be used to decide whether some information should be written to screen or file.
+  In addition it can format output according to some run time parameters given by a 
+  parameter file.
+\end{itemize}
+
+\marginpar{Write the rest! WW}
+\subsection{Internal structures}
+interfaces:
+transposeddirichletdatainterface.h
+reducedprobleminterface.h
+            pdeprobleminterface.h
+
+problemdata:
+noconstraints.h        stateproblem.h          transposedgradientdirichletdata.h
+primaldirichletdata.h  tangentdirichletdata.h  transposedhessiandirichletdata.h
+simpledirichletdata.h
+\subsection{Wrapper classes}
+
+wrapper:
+dofhandler\underline{ }wrapper.h  finiteelement\underline{ }wrapper.h  preconditioner\underline{ }wrapper.h
+fevalues\underline{ }wrapper.h    function\underline{ }wrapper.h       snopt\underline{ }wrapper.h
+
+{\bf TODO! The following files need some appropriate description here 
+This should be done by someone who knows what they are used for. Pleas move the description to the right place above.}
+\marginpar{TODO}
+
+basic:
+      sth\underline{ }internals.h                     
+      DOpEtypes.h 
+include:   
+solutionextractor.h                 
+constraintsmaker.h            helloworld.h                     sparsitymaker.h                     
+            helper.h                                           timedofhandler.h                    
+              timeiterator.h
+
+reducedproblems:
+ forwardtimestepreducedproblem.h
+interfaces:
+active\underline{ }fe\underline{ }index\underline{ }setter\underline{ }interface.h

### 2012-03-16 Examples/Documentation/tests.tex
Augmented the Documentation. Please read carefully to see if this is fine

@@
 \chapter{Testing examples}\label{chap:test}
-The DOpE testsuite consists of regression tests. They are run to compare 
+The DOpE test suite consists of regression tests. They are run to compare 
 the output to previous outputs. This is useful (necessary) after 
@@
 \section{Where can I find the tests}
-In each example directory you find a subdirectory `Test'. Herein, you find 
+In each example directory you find a sub directory `Test'. Herein, you find 
 the parameter files for meshes (*.inp) and a param file (test.prm).
@@
 
-The useful point is now the following. After implemention of 
+The useful point is now the following. After implementation of 
 new pieces of code in the DOpE library or in the examples, you can

### 2012-03-16 Examples/description.tex
Augmented the Documentation. Please read carefully to see if this is fine

@@
 \begin{document}
-\title{DOpE - Deal Optimization Environment}
+\title{DOpElib - Deal Optimization Environment}
 \author{
-  Michael Geiger\thanks{University of Heidelberg, 
-    {michael.geiger@iwr.uni-heidelberg.de}}, 
   Christian Goll\thanks{University of Heidelberg, 

### 2012-03-19 Examples/Documentation/introduction.tex
Added some lines to the how the library is structured part.

--- 2012-03-16 Examples/Documentation/introduction.tex
@@
 \section{Software requirements}
+The library DOpE has been tested to work on both Linux and 
+MAC OSX. It requires a recent \texttt{ g++} (at least 4.4.3) due to 
+some new \texttt{ C++} features implemented in \texttt{ C++0x} and \texttt{ C++11}.
 
-\subsection{deal.ii}
-This library is based upon {\tt deal.ii} hence in order to run 
-{\tt DOpElib} you need a running copy of {\tt deal.ii}.
+\subsection{deal.II}
+This library is based upon \texttt{deal.II} hence in order to run 
+\texttt{ DOpElib} you need a running copy of \texttt{deal.II}.
 
-The deal.II library is open source and is freely availiable for noncommercial project.
+The \texttt{deal.II} library is open source and is freely available for noncommercial project.
 It can be downloaded from \url{http://www.dealii.org/}. On this
@@
 If you would like to use the features offered in our SNOPT wrapper. You will 
-need to obtain a licence for {\tt SNOPT} 
+need to obtain a license for \texttt{ SNOPT} 
 \url{http://www.sbsi-sol-optimize.com/asp/sol_product_snopt.htm}.
-Unfortuately this is at present not availiable for free, but you should check if 
-there is a department license already availiable.
+Unfortunately this is at present not available for free, but you should 
+check if 
+there is a department license already available.
 
-\section{How this library is struktured}
-\marginpar{TODO!}
+\section{How this library is structured}
+This library is designed to allow easy implementation and numerical solutions 
+of problems involving partial differential equations (PDEs). The easiest case 
+is that of a PDE in weak form to find some $u$
+\[
+a(u)(\phi) = 0 \quad \forall \phi \in V
+\]
+with some appropriate space $V$.
+More complex cases involve optimization problems given in the form (OPT)
+\begin{align*}
+\min\;&J(q,u) \\
+  &\text{s.t.}\; a(q,u)(\phi) = 0 \quad \forall \phi\\
+  &a \le q \le b\\
+  &g(q,u) \le 0  
+\end{align*}
+where $u$ is a FE-function and $q$ can either be a FE-function or some 
+fixed number of parameters.
+
+\subsection{Problem description}
+In order to allow our algorithms the automatic assembly of all required 
+data we need to have some container which contains the complete problem 
+description in a common data format. For this we have the following 
+classes in \texttt{ DOpEsrc/container}
+\begin{itemize}
+  \item \texttt{ pdeproblemcontainer.h} Is used to describe  stationary PDE problems.
+  \item \texttt{ instatpdeproblemcontainer.h} {\bf TODO is still missing but will be used for nonstationary problems.} This will be implemented once we have nonstationary optimization problems running to avoid error duplication in the coding process.
+  \item \texttt{ optproblem.h} Is used to describe  OPT problems governed by 
+    stationary PDEs. {\bf TODO should be renamed to optproblemcontainer.h including the class name!}
+  \item \texttt{ instatoptproblemcontainer.h} Is used to describe  OPT problems governed by nonstationary PDEs.
+The only difference to the stationary case is that we need to specify a time-stepping method.
+\end{itemize}
+In order to fill these containers there are two things to be done,
+first we need to actually write some data like $a$, $J$, $\ldots$
+which describe the problem. Then we have to select some numerical 
+algorithm components like finite elements, linear solvers $\ldots$.
+The latter ones should be written such that when exchanging these components
+none of the problem descriptions should require changes. 
+Note that it still may be necessary to write some additional descriptions, 
+e.g., if you solve the PDE with a fix point iteration you don't need derivatives
+but if you want to use Newton's method you then have to provide some.
+
+We will start by discussing the problem description components implemented so far
+
+\subsubsection{Problem specific classes}
+The following classes are used to describe the problem and will usually require 
+some implementation.
+
+\begin{itemize}
+  \item \texttt{ basic/constraints.h} This is used by the spacetimehandlers to 
+    compute the number of constraints from the control and state vectors. 
+    It must not be reimplemented by the user, but needs to be properly 
+    initialized if OPT is used with box control constraints or $g(q,u) \le 0$.
+  \item \texttt{ interfaces/functionalinterface.h} This gives an interface 
+    for the functional $J$ and any other functional you may want to evaluate.
+    In general this can be used as a base class to write your own functionals 
+    in examples. We note that we only need to write the integrands on 
+    elements or faces the loop over elements will be taken care of else where.
+  \item \texttt{ interfaces/constraintinterface.h} This gives an interface for both 
+    the control box constraints as well as the general constraint $g \le 0$. This 
+    needs to be specified if constraints are to be used. If they are not needed 
+    a default class \texttt{ problemdata/noconstraints.h} can be used. We note that we only 
+    need to write the integrands on 
+    elements or faces the loop over elements will be taken care of else where.
+  \item \texttt{ interfaces/pdeinterface.h} This defines an interface for the 
+    partial differential equation $a(q,u)(\phi) = 0$. This needs to be written
+    by the user. We note that we only need to write the integrands on 
+    elements or faces the loop over elements will be taken care of else where.
+  \item \texttt{ interfaces/dirichletdatainterface.h} This gives an interface to the 
+    Dirichlet data for a problem. If the Dirichlet data are simply a function 
+    (and do not depend on the control $q$ one can use the default class
+    \texttt{ problemdata/simpledirichletdata.h}.
+\end{itemize}
+
+\subsubsection{Numeric components}
+These are the components from which a user needs to select some in order to actually 
+solve the given problem. They will not require any rewriting, but sometimes it is 
+advisable to write other than the default parameter into the param file for the 
+solution.
+
+First we need to select a method how to handle all dofs in space and time.
+\begin{itemize}
+\item \texttt{ basic/spacetimehandler\underline{ }base.h} This class is used to define 
+  an interface to the dimension independent functionality of all space time dof handlers. 
+\item \texttt{ basic/statespacetimehandler.h} Another intermediate interface class which adds 
+  the dimension dependent functionality if only the variable $u$ is considered, i.e., a 
+  PDE problem.
+\item \texttt{ basic/spacetimehandler.h } Same as above but with both $q$ and $u$, i.e., for
+  OPT problems.
+\item \texttt{ basic/mol\underline{ }statespacetimehandler.h} Implementation of a method of 
+  line space time dof handler for PDE problems. It has only one spatial 
+  dofhandler which is used for all time intervals.
+\item \texttt{ basic/mol\underline{ }spacetimehandler.h} Same as above for OPT problems.
+  A separate spatial dof handler for each of the variables $q$ and $u$ is maintained 
+  but only one triangulation.
+\item \texttt{ basic/mol\underline{ }multimesh\underline{ }spacetimehandler.h}
+  Same as above, but now in addition the triangulations for $q$ and $u$ can be refined
+  separately from one common initial coarse triangulation. Note that this will
+  in addition require the use of the multimesh version for integrator and 
+  face- as well as celldatacontainer.
+\end{itemize}
+Note that we use these for stationary problems as well, but then you don't have to specify
+any time information.
+
+Second you will need to specify some container classes to be used to 
+pass data between objects. At present you don't have much choice, but you may wish 
+to reimplement some of these if you need data that is not currently included in 
+the containers.
+\begin{itemize}
+\item \texttt{ container/celldatacontainer.h} This object is used to pass data 
+  given on the current element (cell) of the mesh to the functions in PDE, functional, 
+  $\ldots$. 
+\item \texttt{ container/facedatacontainer.h} This object is used to pass data 
+  given on the current face of the mesh to the functions in PDE, functional, 
+  $\ldots$. 
+\item \texttt{ container/multimesh\underline{ }celldatacontainer.h} This is the same as the 
+  celldatacontainer, but it
+  is capable to handle data defined on an alternative triangulation.
+\item \texttt{ container/multimesh\underline{ }facedatacontainer.h} This is the same as the
+  facedatacontainer, but it
+  is capable to handle data defined on an alternative triangulation.
+\item \texttt{ container/integratordatacontainer.h} This contains some data that 
+  should be passed to the integrator like quadrature formulas and the above cell and 
+  face data container.
+\end{itemize}
+
+Third, at least for nonstationary PDEs we need to select a time stepping scheme
+the file names of which are mostly self explanatory:
+\begin{itemize}
+\item \texttt{ include/forward\underline{ }euler\underline{ }problem.h}
+\item \texttt{ include/shifted\underline{ }crank\underline{ }nicolson\underline{ }problem.h}
+\item \texttt{ include/backward\underline{ }euler\underline{ }problem.h}
+\item \texttt{ include/fractional\underline{ }step\underline{ }theta\underline{ }problem.h} Note that the use of this scheme requires a special Newton solver!
+\item \texttt{ include/crank\underline{ }nicolson\underline{ }problem.h}
+\end{itemize}
+
+Finally, we need to select a way how to integrate and solve linear and nonlinear equations
+\begin{itemize}
+\item \texttt{ templates/integrator.h} This class computes integrals over a given 
+  triangulation (including its faces).
+\item \texttt{ templates/integrator\underline{ }multimesh.h} The same as above but it is 
+  possible that some of the FE functions are defined on an other triangulation 
+  as long as the have a common coarse triangulation.
+\item \texttt{ templates/integratormixeddims.h} This is used to compute integral which 
+  are given in another (larger) dimension than the current variable. (This is exclusively
+  used if the control variable is given by some parameters. Which means \texttt{ dopedim == 0}). 
+\item \texttt{ templates/cglinearsolver.h} This is a wrapper for the cg solver implemented in 
+  \texttt{deal.II}. The solver will build and store the stiffness matrix for the PDE.
+\item \texttt{ templates/gmreslinearsolver.h} This is a wrapper for the gmres solver 
+  implemented in \texttt{deal.II}. The solver will build and store the stiffness matrix 
+  for the PDE.
+\item \texttt{ templates/directlinearsolver.h} This is a wrapper for the direct solver 
+  implemented in \texttt{deal.II} using \texttt{ UMFPACK}. 
+  The solver will build and store the stiffness matrix for the PDE.
+\item \texttt{ templates/voidlinearsolver.h} This is a wrapper for certain cases when we 
+  know that the matrix to be inverted is the identity. It simply copies the rhs to the
+  lhs. This is only needed for compatibility reasons some other components.
+\item \texttt{ templates/newtonsolver.h} This solves some nonlinear equation using a 
+  line-search Newton method.
+\item \texttt{ templates/newtonsolvermixeddims.h} The same but in the case when there is 
+  another variable in a (larger) dimension is involved. See 
+  \texttt{ integratormixeddims.h}.
+\item \texttt{ templates/instat\underline{ }step\underline{ }newtonsolver.h} This is a 
+  Newton method as above to invert the next time-step. It differs from the plain vanilla
+  version in that it computes certain data from the previous time step only once 
+  and not in every Newton iteration.
+\item \texttt{ templates/fractional\underline{ }step\underline{ }theta\underline{ }step\underline{ }newtonsolver.h} This is the Newton solver for the time step in a 
+  fractional-step-theta scheme. It combines the computation of all three sub steps.
+\end{itemize}
+
+\subsection{Reduced problems (Solve the PDE)}
+At times it is nice to remove the PDE constraint. This is handled by so called reduced 
+problems. This means that the reduced problem implicitly solves the PDE whenever required
+and eliminates the variable $u$ from the problem.
+\begin{itemize}
+\item \texttt{ reducedproblems/statpdeproblem.h} This is used to remove the variable $u$ in 
+  a stationary PDE problem. This means that call the method \\
+  \texttt{ StatPDEProblem::ComputeReducedFunctionals} will evaluate the functionals 
+  defined in the problem description, i.e., in \texttt{ PDEProblemContainer}, in the 
+  solution of the given PDE.
+\item \texttt{ reducedproblems/statreducedproblem.h} This eliminates $u$ from the OPT
+  problem with a stationary PDE.
+\item \texttt{ reducedproblems/instatreducedproblem.h} The same as above but for a
+  nonstationary PDE. {\bf FIXME there is something wrong in this file see FIXME 
+    comment in the source.}
+\item \texttt{ reducedproblems/voidreducedproblem.h} A wrapper file that eliminates $u$ 
+  if it is not present anyways. This is used so that we can use the same routines to 
+  solve problems that have no PDE constraint.
+\end{itemize}
+
+\subsection{Optimization algorithms}
+Now, in order to solve optimization algorithms we need to define some algorithms.
+At present we offer a selection of algorithms that solve the reduced optimization 
+problem where the PDE constraint has been eliminated as explained in the previous section
+\begin{itemize}
+\item \texttt{ opt\underline{ }algorithms/reducedalgorithm.h} An interface for all 
+  optimization problems in the reduced formulation. It offers some test functionality
+  to assert that the derivatives of the problem are computed correctly.
+\item \texttt{ opt\underline{ }algorithms/reducednewtonalgorithm.h}
+  A line-search Newton algorithm using a cg method to invert the reduced hessian. 
+  Implementation ignores any additional constraints.
+\item \texttt{ opt\underline{ }algorithms/reducedtrustregionnewton.h}
+  A trust region Newton algorithm using a cg method to invert the reduced hessian.
+  Implementation ignores any additional constraints.
+\item \texttt{ opt\underline{ }algorithms/reduced\underline{ }snopt\underline{ }algorithm.h}
+  An algorithm to solve reduced optimization problems with additional control constraints.
+  ((reduced) state constraints are not yet implemented.)
+\item \texttt{ opt\underline{ }algorithms/reducednewtonalgorithmwithinverse.h}
+  Line-search Newton algorithm that assumes there exists a method in the reduced problem
+  that can invert the reduced hessian. (This usually makes sense only if there is no 
+  PDE constraint.)
+\item \texttt{ opt\underline{ }algorithms/generalized\underline{ }mma\underline{ }algorithm.h}
+  An implementation of the MMA-Algorithm for structural optimization using an augmented
+  Lagrangian formulation for the subproblems. The subproblem is implemented using the 
+  special purpose file
+  \texttt{ include/augmentedlagrangianproblem.h}.
+\end{itemize}
+
+\subsection{Other Components}
+Beyond these clearly structured groups before there are some classes remaining that
+do not fit the above but are important for the user to know:
+\begin{itemize}
+\item \texttt{ include/statevector.h} This stores all dofs in space and time for the state 
+  variable $u$. It is possible to select whether all this should be kept in memory or 
+  or unused parts can be written to the hard disk.
+\item \texttt{ include/controlvector.h} This stores all dofs in space and time for the 
+  control variable $q$. At present no time dependence is implemented.
+\item \texttt{ include/constraintvector.h} This stores all dofs in space and time for the 
+  non PDE constraints (and corresponding multipliers). 
+  At present no time dependence is implemented.
+\item \texttt{ include/parameterreader.h} This file is used to define a parameter reader
+  that is used to read run time parameters from a given file.
+\item \texttt{ include/dopeexception.h} Defines some Exceptions that are thrown by the program
+  should it encounter any unexpected errors.
+\item \texttt{ include/dopeexceptionhandler.h} This class is used to write information 
+  contained in the exceptions to the output in a uniform manner.
+\item \texttt{ include/outputhandler.h} This file defines an outputhandler object which 
+  can be used to decide whether some information should be written to screen or file.
+  In addition it can format output according to some run time parameters given by a 
+  parameter file.
+\item \texttt{include/constraintsmaker.h} This class sets the constraints on the DOFs of the state FE solution.
+In the standard implementation, only the hanging-node-constraints are built, but the user can reimplement this class and then include other constraints as well (for example periodic BC).
+\item \texttt{include/sparsitymaker.h} This class sets the sparsity pattern for the state FE solution.
+The standard implementiation is just a wrapper for \texttt{dealii::DoFTools::} \texttt{make\_sparsity\_pattern}, but the user can reimplement this class to allow for more sophisticated sparsity patterns.
+\item \texttt{interfaces/active\underline{ }fe\underline{ }index\underline{ }setter\underline{ }interface.h} In the case of hp finite elements, one has to specify for each cell which finite element to use.
+This is done via this interface.
+\end{itemize}
+
+\marginpar{Write the rest! WW}
+\subsection{Internal structures}
+interfaces:
+transposeddirichletdatainterface.h
+reducedprobleminterface.h
+            pdeprobleminterface.h
+
+problemdata:
+noconstraints.h        stateproblem.h          transposedgradientdirichletdata.h
+primaldirichletdata.h  tangentdirichletdata.h  transposedhessiandirichletdata.h
+simpledirichletdata.h
+\begin{itemize}
+\item \texttt{include/timedofhandler.h} DoFHandler responsible for the management of the timedofs (this is a part of the \texttt{SpaceTimeDoFHandler}-classes).
+Basically a wrapper for a $1d$ \texttt{deal.II}-DoFHandler.
+\item \texttt{include/timeiterator.h} This class works as an iterator on the \texttt{TimeDoFHandler}.                 
+\end{itemize}
+\subsection{Wrapper classes}
+
+wrapper:
+dofhandler\underline{ }wrapper.h  finiteelement\underline{ }wrapper.h  preconditioner\underline{ }wrapper.h
+fevalues\underline{ }wrapper.h    function\underline{ }wrapper.h       snopt\underline{ }wrapper.h
+
+{\bf TODO! The following files need some appropriate description here 
+This should be done by someone who knows what they are used for. Please move the description to the right place above.}
+\marginpar{TODO}
+
+basic:
+      sth\underline{ }internals.h                     
+      DOpEtypes.h 
+include:   
+solutionextractor.h                 
+            helloworld.h                               
+            helper.h
+
+reducedproblems:
+ forwardtimestepreducedproblem.h

### 2012-03-22 Examples/Documentation/foreword.tex
Add Literature and update foreword

@@
 based on the \textit{deal.II} finite element library which has been developed
- initially by W. Bangerth, R. Hartmann, and G. Kanschat. \marginpar{Citation}
+ initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{deal}.
 Its main feature is to give a unified interface to high level algorithms such as 
@@
 Gascoigne/RoDoBo project, which was initiated by 
-Roland Becker, Dominik Meidner,  and Boris Vexler. \marginpar{Citation} 
+Roland Becker, Dominik Meidner,  and Boris Vexler \cite{rodobo}. 
 From which some of the ideas to modularize the algorithms have arisen.
@@
   Crank-Nicolson, shifted Crank-Nicolson, and Fractional-Step-$\Theta$ scheme.
-\item All finite elements of from deal.ii including hp-support.
+\item All finite elements of from deal.II including hp-support.
 \item Several examples showing the solution of several PDEs including
@@
 \item Self written line search and trust region newton algorithms for the 
-   solution of optimization problems with PDEs. (cf. Nocedal \& Wright\marginpar{Citation})
+   solution of optimization problems with PDEs \cite{NoWr00}
 \item Interface to SNOPT for the solution of optimization problems with PDEs and

### 2012-03-22 Examples/Documentation/introduction.tex
Introduction ueberarbeitet

@@
 \chapter{Introduction}\label{chap:intro}
+\section{License information}
+{\bf TODO:} Unter welcher License soll DOpE laufen?
+
 \section{Developers}
@@
 However, there are more highly appreciated contributions
-made by (in alphabetical order)
+made by %(in alphabetical order)
 \begin{itemize}
@@
 \[
-a(u)(\phi) = 0 \quad \forall \phi \in V
+a(u)(\phi) = 0 \quad \forall \phi \in V,
 \]
@@
 \min\;&J(q,u) \\
-  &\text{s.t.}\; a(q,u)(\phi) = 0 \quad \forall \phi\\
-  &a \le q \le b\\
-  &g(q,u) \le 0  
+  &\text{s.t.}\; a(q,u)(\phi) = 0 \quad \forall \phi\in V,\\
+  &a \le q \le b,\\
+  &g(q,u) \le 0,  
 \end{align*}
 where $u$ is a FE-function and $q$ can either be a FE-function or some 
-fixed number of parameters.
+fixed number of parameters, $a$ and $b$ are constraint bounds for the control $q$,
+and $g(\cdot)$ is some state constraint.
 
@@
 In order to fill these containers there are two things to be done,
-first we need to actually write some data like $a$, $J$, $\ldots$
+first we need to actually write some data, for instance,
+the semilinear form $a(\cdot)(\cdot)$, a target functional $J(\cdot)$, etc.,
 which describe the problem. Then we have to select some numerical 
@@
 e.g., if you solve the PDE with a fix point iteration you don't need derivatives
-but if you want to use Newton's method you then have to provide some.
+but if you want to use Newton's method, derivatives are needed.
 
@@
 
-\subsubsection{Problem specific classes}
-The following classes are used to describe the problem and will usually require 
-some implementation.
-
-\begin{itemize}
-  \item \texttt{ basic/constraints.h} This is used by the spacetimehandlers to 
-    compute the number of constraints from the control and state vectors. 
-    It must not be reimplemented by the user, but needs to be properly 
-    initialized if OPT is used with box control constraints or $g(q,u) \le 0$.
-  \item \texttt{ interfaces/functionalinterface.h} This gives an interface 
-    for the functional $J$ and any other functional you may want to evaluate.
-    In general this can be used as a base class to write your own functionals 
-    in examples. We note that we only need to write the integrands on 
-    elements or faces the loop over elements will be taken care of else where.
-  \item \texttt{ interfaces/constraintinterface.h} This gives an interface for both 
-    the control box constraints as well as the general constraint $g \le 0$. This 
-    needs to be specified if constraints are to be used. If they are not needed 
-    a default class \texttt{ problemdata/noconstraints.h} can be used. We note that we only 
-    need to write the integrands on 
-    elements or faces the loop over elements will be taken care of else where.
-  \item \texttt{ interfaces/pdeinterface.h} This defines an interface for the 
-    partial differential equation $a(q,u)(\phi) = 0$. This needs to be written
-    by the user. We note that we only need to write the integrands on 
-    elements or faces the loop over elements will be taken care of else where.
-  \item \texttt{ interfaces/dirichletdatainterface.h} This gives an interface to the 
-    Dirichlet data for a problem. If the Dirichlet data are simply a function 
-    (and do not depend on the control $q$ one can use the default class
-    \texttt{ problemdata/simpledirichletdata.h}.
-\end{itemize}
-
-\subsubsection{Numeric components}
+\subsection{Numeric components}
 These are the components from which a user needs to select some in order to actually 
@@
 
+\subsubsection{Space-time handler}
 First we need to select a method how to handle all dofs in space and time.
@@
   an interface to the dimension independent functionality of all space time dof handlers.
+  {\bf TODO: Beispiele geben}
 \item \texttt{ basic/statespacetimehandler.h} Another intermediate interface class which adds 
@@
   line space time dof handler for PDE problems. It has only one spatial 
-  dofhandler which is used for all time intervals.
+  dofhandler that is used for all time intervals.
 \item \texttt{ basic/mol\underline{ }spacetimehandler.h} Same as above for OPT problems.
@@
 
+\subsubsection{Container classes}
 Second you will need to specify some container classes to be used to 
@@
 
+\subsubsection{Time stepping schemes}
 Third, at least for nonstationary PDEs we need to select a time stepping scheme
@@
 \item \texttt{ include/backward\underline{ }euler\underline{ }problem.h}
-\item \texttt{ include/fractional\underline{ }step\underline{ }theta\underline{ }problem.h} Note that the use of this scheme requires a special Newton solver!
+\item \texttt{ include/fractional\underline{ }step\underline{ }theta\underline{ }problem.h} Note that the use of this scheme requires a special Newton solver, which is, however, already
+implemented for the convenience of the user!
 \item \texttt{ include/crank\underline{ }nicolson\underline{ }problem.h}
@@
 
+\subsubsection{Integrator routines}
 Finally, we need to select a way how to integrate and solve linear and nonlinear equations
@@
   used if the control variable is given by some parameters. Which means \texttt{ dopedim == 0}). 
-\item \texttt{ templates/cglinearsolver.h} This is a wrapper for the cg solver implemented in 
-  \texttt{deal.II}. The solver will build and store the stiffness matrix for the PDE.
-\item \texttt{ templates/gmreslinearsolver.h} This is a wrapper for the gmres solver 
-  implemented in \texttt{deal.II}. The solver will build and store the stiffness matrix 
-  for the PDE.
-\item \texttt{ templates/directlinearsolver.h} This is a wrapper for the direct solver 
-  implemented in \texttt{deal.II} using \texttt{ UMFPACK}. 
-  The solver will build and store the stiffness matrix for the PDE.
-\item \texttt{ templates/voidlinearsolver.h} This is a wrapper for certain cases when we 
-  know that the matrix to be inverted is the identity. It simply copies the rhs to the
-  lhs. This is only needed for compatibility reasons some other components.
+\end{itemize}
+
+\subsubsection{Nonlinear solvers}
+\begin{itemize}
 \item \texttt{ templates/newtonsolver.h} This solves some nonlinear equation using a 
@@
 
+\subsubsection{Linear solvers}
+\begin{itemize}
+\item \texttt{ templates/cglinearsolver.h} This is a wrapper for the cg solver implemented in 
+  \texttt{deal.II}. The solver will build and store the stiffness matrix for the PDE.
+\item \texttt{ templates/gmreslinearsolver.h} This is a wrapper for the gmres solver 
+  implemented in \texttt{deal.II}. The solver will build and store the stiffness matrix 
+  for the PDE.
+\item \texttt{ templates/directlinearsolver.h} This is a wrapper for the direct solver 
+  implemented in \texttt{deal.II} using \texttt{ UMFPACK}. 
+  The solver will build and store the stiffness matrix for the PDE.
+\item \texttt{ templates/voidlinearsolver.h} This is a wrapper for certain cases when we 
+  know that the matrix to be inverted is the identity. It simply copies the rhs to the
+  lhs. This is only needed for compatibility reasons some other components.
+\end{itemize}
+
+\subsection{Problem specific classes}
+The following classes are used to describe the problem and will usually require 
+some implementation.
+
+\begin{itemize}
+  \item \texttt{ basic/constraints.h} This is used by the spacetimehandlers to 
+    compute the number of constraints from the control and state vectors. 
+    It must not be reimplemented by the user, but needs to be properly 
+    initialized if OPT is used with box control constraints or $g(q,u) \le 0$.
+  \item \texttt{ interfaces/functionalinterface.h} This gives an interface 
+    for the functional $J(\cdot)$ and any other functional you may want to evaluate.
+    In general this can be used as a base class to write your own functionals 
+    in examples. We note that we only need to write the integrands on 
+    elements or faces the loop over elements will be taken care of in the integrator.
+    Specifically, derivatives are written therein, too.
+  \item \texttt{ interfaces/constraintinterface.h} This gives an interface for both 
+    the control box constraints as well as the general constraint $g \le 0$. This 
+    needs to be specified if constraints are to be used. If they are not needed 
+    a default class \texttt{ problemdata/noconstraints.h} can be used. We note that we only 
+    need to write the integrands on 
+    elements or faces the loop over elements will be taken care of in the integrator.
+  \item \texttt{ interfaces/pdeinterface.h} This defines an interface for the 
+    partial differential equation $a(q,u)(\phi) = 0$. This needs to be written
+    by the user. We note that we only need to write the integrands on 
+    elements or faces the loop over elements will be taken care of in the integrator.
+    Specifically, derivatives are written therein, too.
+  \item \texttt{ interfaces/dirichletdatainterface.h} This gives an interface to the 
+    Dirichlet data for a problem. If the Dirichlet data are simply a function 
+    (and do not depend on the control $q$) one can use the default class\\
+    \texttt{ problemdata/simpledirichletdata.h}.
+\end{itemize}
+
 \subsection{Reduced problems (Solve the PDE)}
-At times it is nice to remove the PDE constraint. This is handled by so called reduced 
-problems. This means that the reduced problem implicitly solves the PDE whenever required
+At times it is nice to remove the PDE constraint in (OPT). 
+This is handled by so called reduced 
+problems (for algorithmis aspects we refer the reader to 
+\cite{BeMeVe06}). 
+This means that the reduced problem implicitly solves the PDE whenever required
 and eliminates the variable $u$ from the problem.
@@
 At present we offer a selection of algorithms that solve the reduced optimization 
-problem where the PDE constraint has been eliminated as explained in the previous section
+problem where the PDE constraint has been eliminated as explained in the previous section.
 \begin{itemize}
@@
 Beyond these clearly structured groups before there are some classes remaining that
-do not fit the above but are important for the user to know:
+do not fit the above but are important for the user to know.
+
+\subsubsection{Vectors}
 \begin{itemize}
@@
   At present no time dependence is implemented.
+\end{itemize}
+
+\subsubsection{Parameter handling}
+\begin{itemize}
 \item \texttt{ include/parameterreader.h} This file is used to define a parameter reader
   that is used to read run time parameters from a given file.
+\end{itemize}
+
+\subsubsection{Exception handling}
+\begin{itemize}
 \item \texttt{ include/dopeexception.h} Defines some Exceptions that are thrown by the program
@@
   contained in the exceptions to the output in a uniform manner.
+\end{itemize}
+
+\subsubsection{Output handling}
+\begin{itemize}
 \item \texttt{ include/outputhandler.h} This file defines an outputhandler object which 
@@
   parameter file.
+\end{itemize}
+
+\subsubsection{Constraints and system matrix}
+\begin{itemize}
 \item \texttt{include/constraintsmaker.h} This class sets the constraints on the DOFs of the state FE solution.
 In the standard implementation, only the hanging-node-constraints are built, but the user can reimplement this class and then include other constraints as well (for example periodic BC).
+{\bf TODO: to rename the file?? (CG)}
 \item \texttt{include/sparsitymaker.h} This class sets the sparsity pattern for the state FE solution.
 The standard implementiation is just a wrapper for \texttt{dealii::DoFTools::} \texttt{make\_sparsity\_pattern}, but the user can reimplement this class to allow for more sophisticated sparsity patterns.
+\end{itemize}
+
+\subsubsection{HP components}
+\begin{itemize}
 \item \texttt{interfaces/active\underline{ }fe\underline{ }index\underline{ }setter\underline{ }interface.h} In the case of hp finite elements, one has to specify for each cell which finite element to use.
@@
 \subsection{Internal structures}
+{\bf TODO: ab hier habe ich noch nichts gemacht (TW)}
 interfaces:

### 2012-03-22 Examples/Documentation/introduction.tex
More documentation including installation hints

@@
 The library DOpE has been tested to work on both Linux and 
-MAC OSX. It requires a recent \texttt{ g++} (at least 4.4.3) due to 
+MAC OSX. See also the \texttt{README.OSX} file
+
+\texttt{DOpElib} requires a recent \texttt{ g++} (at least 4.4.3) due to 
 some new \texttt{ C++} features implemented in \texttt{ C++0x} and \texttt{ C++11}.
@@
 
-\subsection{SNOPT}
+When installing \texttt{deal.II} you should take care to configure 
+it to use UMFPACK, i.e.,\\
+\texttt{./configure --with-umfpack}
+
+\subsection{ThirdPartyLibraries}
+In order for \texttt{DOpE} to be able to auto-detect some of the installed 
+Third Party Libraries you should generate according links in the 
+\texttt{ThirdPartyLibs}. See also \texttt{ThirdPartyLibs/README}.
+
+\subsubsection{SNOPT}
 If you would like to use the features offered in our SNOPT wrapper. You will 
@@
 there is a department license already available.
+For further information you should consult the file 
+\texttt{ThirdPartyLibs/SNOPT.INSTALLNOTES}. In particular you need to configure 
+\texttt{deal.II} with at least the following options:\\
+\texttt{./configure --with-umfpack --with-blas --with-lapack}
+
+\section{Installation}
+I) Get a copy of \texttt{DOpElib} from the svn repository using\\
+\texttt{svn co --username=[your username] \textbackslash \\ \hspace*{5mm}https://ganymed.iwr.uni-heidelberg.de/svn/dope}\\
+(The backslash in the first line means that the line will be continued as one!)
+Note that you need a valid username and password. If you have none you can contact the 
+maintainer by sending an EMail to \url{dope@ganymed.iwr.uni-heidelberg.de}.\\[3mm]
+%
+II) Install \texttt{deal.II} to your home directory, i.e., it should be located in \\
+\texttt{\textasciitilde/deal.II}. \\[1mm]
+If you would like to have another path you will either have to manually edit the files
+\texttt{DOpEsrc/source/Makefile} and \texttt{Examples/Make.global\underline{ }options} and replace the 
+line \\
+\texttt{D = \$(HOME)/deal.II} with the appropriate path.\\[1mm]
+Configure it properly and make the whole \texttt{deal.II} library.\\[3mm]
+%
+III) If you want to use some of the supported third party libraries install them and follow 
+ the instructions in \texttt{ThirdPartyLibs/README}. There may be further information 
+ in some \texttt{ThirdPartyLibs/*.INSTALLNOTES} that you may want to consider.\\[3mm]
+%
+IV) To build the \texttt{DOpElib} you have to change to the directory 
+\texttt{DOpEsrc} where you can call\\ 
+\texttt{make all}\\
+to build the library. If you want to generate some documentation you may use
+\texttt{make pdf-doc} or \texttt{make html-doc} which will require either 
+latex or doxygen to be installed on your computer.\\[3mm]
+%
+V) To build the Examples change to the \texttt{Examples} directory and \texttt{make all}.
+%
+VI) Finally you may wish to test if everything worked. To do so you can 
+change to the \texttt{Examples} directory and \texttt{make tests} which will give you a 
+list of all the examples and whether they behave as expected by the library, see also 
+Chapter~\ref{chap:test}.
 
@@
 This is handled by so called reduced 
-problems (for algorithmis aspects we refer the reader to 
+problems (for algorithmic aspects we refer the reader to 
 \cite{BeMeVe06}). 
@@
 
+\subsection{Data Access}
+\begin{itemize}
+\item \texttt{include/solutionextractor.h} This class is used to gain access to the finite element 
+  solutions stored in the reduced problems.
+\end{itemize}
+
 \subsubsection{Constraints and system matrix}
@@
 \item \texttt{include/sparsitymaker.h} This class sets the sparsity pattern for the state FE solution.
-The standard implementiation is just a wrapper for \texttt{dealii::DoFTools::} \texttt{make\_sparsity\_pattern}, but the user can reimplement this class to allow for more sophisticated sparsity patterns.
+The standard implementation is just a wrapper for \texttt{dealii::DoFTools::} \texttt{make\_sparsity\_pattern}, but the user can reimplement this class to allow for more sophisticated sparsity patterns.
 \end{itemize}
@@
 
-\marginpar{Write the rest! WW}
 \subsection{Internal structures}
 {\bf TODO: ab hier habe ich noch nichts gemacht (TW)}
-interfaces:
-transposeddirichletdatainterface.h
-reducedprobleminterface.h
-            pdeprobleminterface.h
+\subsubsection{Interface Classes}
+\begin{itemize}
+  \item \texttt{interfaces/transposeddirichletdatainterface.h} This provides an interface to 
+    the functionality required by {\em transposed Dirichlet data}. Usually when one applies Dirichlet 
+    data $g$ to a function one has to calculate a continuation $Bg$ which is defined on the whole domain.
+    In optimization problems when the Dirichlet data depends on the control one has to evaluate the 
+    dual operator $B^*$ in order to obtain a representation for the reduced gradient of the objective $J$.
+    This is done using the {\em transposed Dirichlet data}.
+  \item \texttt{interfaces/reducedprobleminterface.h} In order to allow all algorithms to be written independent
+    of the given (OPT) problem 
+    (and not requiring the problem as template argument) there is a common base class which 
+    defines the required interfaces. 
+  \item \texttt{interfaces/pdeprobleminterface.h} The same as above but for (PDE) problems.
+\end{itemize}
 
-problemdata:
-noconstraints.h        stateproblem.h          transposedgradientdirichletdata.h
-primaldirichletdata.h  tangentdirichletdata.h  transposedhessiandirichletdata.h
-simpledirichletdata.h
+\subsubsection{Default Classes}
+\begin{itemize}
+  \item \texttt{problemdata/noconstraints.h} A class that can be used for optimization problems 
+    having only a PDE constraint but no further constraints.
+  \item \texttt{problemdata/simpledirichletdata.h} A class that can be used to implement Dirichlet
+    data that are given as a fixed function (independent of the control).
+  \end{itemize}
+
+\subsubsection{Auto-generated Problem Descriptions}
+\begin{itemize}
+  \item \texttt{problemdata/stateproblem.h} This is the problem description for the (forward/primal) PDE constraint.
+    {\bf Similar descriptors will be build for the other problems (adjoint, tangent, $\ldots$) when time allows.}
+  \item \texttt{problemdata/primaldirichletdata.h} This class contains the Dirichlet data for the 
+    forward/primal PDE.
+  \item \texttt{problemdata/tangentdirichletdata.h} This class contains the Dirichlet data for the tangent PDE, i.e.,
+    the first derivative of the Dirichlet data.
+  \item \texttt{problemdata/transposedgradientdirichletdata.h} This contains the transposed Dirichlet data needed 
+    to calculate the gradient of the reduced objective functional, 
+    for detail see \texttt{interfaces/transposeddirichletdatainterface.h}.
+  \item \texttt{problemdata/transposedhessiandirichletdata.h} This contains the transposed Dirichlet data needed 
+    to calculate the hessian of the reduced objective functional, 
+    for detail see \texttt{interfaces/transposeddirichletdatainterface.h}.
+\end{itemize}
+\subsubsection{Management of Time Dependent Problems}
 \begin{itemize}
@@
 \subsection{Wrapper classes}
+\begin{itemize}
+  \item \texttt{wrapper/dofhandler\underline{ }wrapper.h} A wrapper class for the \texttt{deal.II} DoFHandlers. This 
+    class is needed to provide support for the \texttt{dim $= 0$} case and to have a uniform interface to 
+    DoFHandler and HPDoFHandler.
+  \item \texttt{wrapper/preconditioner\underline{ }wrapper.h} Contains wrappers for several of the preconditioners
+    in \texttt{deal.II}. This is required since unfortunately the preconditioners in \texttt{deal.II} have different
+    interfaces for their initialization.
+  \item \texttt{wrapper/function\underline{ }wrapper.h} An interface that allows to use functions that depend 
+    not only on space but also on time.
+  \item \texttt{wrapper/snopt\underline{ }wrapper.h} An interface to the SNOPT fortran library. This is an additional
+    wrapper to the one provided by SNOPT to allow automatic construction of the functions required 
+    by SNOPT using our library.
+  \item \texttt{wrapper/finiteelement\underline{ }wrapper.h} {\bf Will be removed soon!}
+  \item \texttt{wrapper/fevalues\underline{ }wrapper.h} {\bf Will be removed soon!}
+\end{itemize}
 
-wrapper:
-dofhandler\underline{ }wrapper.h  finiteelement\underline{ }wrapper.h  preconditioner\underline{ }wrapper.h
-fevalues\underline{ }wrapper.h    function\underline{ }wrapper.h       snopt\underline{ }wrapper.h
+\subsection{Other}
+\begin{itemize}
+  \item \texttt{basic/dopetypes.h} This file contains type definitions used in the library. 
+\end{itemize}
 
@@
 This should be done by someone who knows what they are used for. Please move the description to the right place above.}
+
 \marginpar{TODO}
@@
       sth\underline{ }internals.h
-      DOpEtypes.h 
+
 include:   
-solutionextractor.h                 
-            helloworld.h                               
             helper.h

### 2012-04-14 DOpEsrc/container/integratordatacontainer.h
First draft of implemntation of stationry DWR. Only implemented for PDEProblem, will extended to the optimization case. Lacking or incomplet documentation. Will taken care of next week.

@@ -33,8 +33,9 @@ namespace DOpE
     {
       public:
         IntegratorDataContainer(const QUADRATURE& quad,
-            const FACEQUADRATURE & face_quad) :
-      _quad(quad), _face_quad(face_quad), _fdc(NULL), _cdc(NULL), _mm_fdc(NULL), _mm_cdc(NULL)
+            const FACEQUADRATURE & face_quad)
+            : _quad(&quad), _face_quad(&face_quad), _fdc(NULL), _cdc(NULL), _mm_fdc(
+                NULL), _mm_cdc(NULL)
         {
         }
 
@@ -67,8 +68,7 @@ namespace DOpE
          */
         template<typename STH>
           void
-          InitializeFDC(
-              UpdateFlags update_flags,
+          InitializeFDC(const FACEQUADRATURE& fquad, UpdateFlags update_flags,
               STH& sth,
               const std::vector<
                   typename DOpEWrapper::DoFHandler<dim, DOFHANDLER>::active_cell_iterator>& cell,
@@ -78,14 +78,34 @@ namespace DOpE
           {
             if (_fdc != NULL)
               delete _fdc;
-            _fdc = new FaceDataContainer<DOFHANDLER, VECTOR, dim> (_face_quad,
-                update_flags, sth, cell, param_values, domain_values, need_interfaces);
+            _fdc = new FaceDataContainer<DOFHANDLER, VECTOR, dim>(fquad,
+                update_flags, sth, cell, param_values, domain_values,
+                need_interfaces);
           }
 
+        /**
+         * Initializes the FaceDataContainer. See the documentation there.
+         * This one uses the previously given facequadrature.
+         */
         template<typename STH>
           void
-          InitializeCDC(
-              UpdateFlags update_flags,
+          InitializeFDC(UpdateFlags update_flags, STH& sth,
+              const std::vector<
+                  typename DOpEWrapper::DoFHandler<dim, DOFHANDLER>::active_cell_iterator>& cell,
+              const std::map<std::string, const Vector<double>*> &param_values,
+              const std::map<std::string, const VECTOR*> &domain_values,
+              bool need_interfaces = false)
+          {
+            InitializeFDC(GetFaceQuad(), update_flags, sth, cell, param_values,
+                domain_values, need_interfaces);
+          }
+
+        /**
+         * Initializes the CellDataContainer. See the documentation there.
+         */
+        template<typename STH>
+          void
[…]

### 2012-04-16 Examples/OPT/StatPDE/Example5/content.tex
Initial Transfer From OLD CVS Reprository
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9

--- 2012-03-15 Examples/OPT/StatPDE/Example1/content.tex
@@
+\subsubsection{General problem description}
 This example solves the distributed minimization problem
@@
 \begin{gather*}
-  \text{MidPoint: } u(0.5 ; 0.5)\\[2mm]
-  \text{MeanValue: }\int_\Omega u
+  \text{MidPoint: } u_h(0.5 ; 0.5)\\[2mm]
+  \text{L1-Value: }\int_\Omega |u_h|
+  \text{QError: }\int_\Omega |q_h-\overline{q}|^2
+  \text{UError: }\int_\Omega |u_h-\overline{u}|^2
 \end{gather*}
+
+The important new feature is that we can now use two different meshes for control and state variable.
+This is tested first for globally refined meshes, and then for locally refined meshes with different refinements 
+for the control and state variable.

### 2012-04-16 Examples/PDE/StatPDE/Example9/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9

\subsubsection{General problem description}
This example shows the use of the adaptive grid refinement and error estimation by the DWR-method (For a description of the method, see \cite{BR03}.) applied to the laplace equation 
\begin{align*}
-\Delta u &= f && \text{ in } \Omega
\end{align*}
with the analytical solution
\begin{align*}
u = \sin\left(\frac{\pi}{x^2 + y^2}\right),
\end{align*}
 the corresponding right hand side $f = -\Delta u$ and appropriate Dirichlet Conditions on $\partial \Omega$, where the domain is given by
\begin{align*}
\Omega = [-2,2]^2\setminus \overline{B}_{0.5}(0).
\end{align*}
We want to estimate the error in the Functional 
$$\fullfunction{J}{H^1(\Omega)}{\mathbb{R}}{u}{\frac 1 {\abs{\Gamma}} \int_\Gamma u \,dx}$$
where
$\Gamma = \left\{(x,y)\in \mathbb R ^2\,\big | \,x=0,-2<y<0.5 \right \}$.

For this setting, we have the error representation
\begin{align*}
J(e)=\sum_{K\in \mathbb T _h} \left\{(R_h, z-\psi_h)_K + (r_h, z-\psi_h)_{\partial_K}\right\}
\end{align*}
with the error $e=u-u_h$, the Triangulation $\mathbb T_h$, the dual solution $z$, arbitraty function $\phi_h \in V_h$ (the ansatz space) and the cell and edge residuals $\eval[1]{R_h}_{K} = f+\Delta u_h$ resp. 
\begin{align*}
\eval[1]{r_h}_{\Sigma} =
\begin{cases}
\frac 12 [\partial_n u_h] ,&\text{if } \Sigma \subset \partial_K \setminus \partial \Omega,\\
-0,& \text{ if } \Sigma \subset\partial \Omega.
\end{cases}
\end{align*}
\subsubsection{Program description}
Additionally to all the things one has to do when just solving the equation, we have to include {\verb higher_order_dwrc.h }. In this program, we approximate the so called 'weights' $z-\phi_h$ by a patchwise higher order interpolation of $z_h$ (the computed dual solution).

The beginning of the program is basically as before, note however that we enforce patch-wise refinement of the grid by giving the flag {\verb Triangulation<2>:: } {\verb MeshSmoothing:: } {\verb patch_level_1 }
to the triangulation. To be able to solve the adjoint equation for the error estimation, we need to implement {\verb FaceValue_U } in \texttt{functionals.h} as well as {\verb CellEquation_U }, {\verb CellMatrix_T }, {\verb CellEquation_U } and {\verb FaceEquation_U } in {\verb localpde.h }. In the latter, we also have to implement {\verb StrongCellResidual }, {\verb StrongFaceResidual } and {\verb StrongBoundaryResidual }.

After everything is set up for the solution of the equation, we have to define some  higher order FiniteElements, some (face)quadrature formulas (and build a \texttt{IntegratorDataContainer} with them.) and a higher order \texttt{SpaceTimeHandler}. We use these to define an object of the  \texttt{HigherOrderDWRC} class. The last step is to initialize the \texttt{HigherOrderDWRC} via \texttt{solver.InitializeHigherOrderDWRC(dwrc)}.

Succeding the solution of the state equation ({\verb solver.ComputeReducedFunctionals() }), we compute the error indicators by calling { \verb solver.ComputeRefinementIndicators(dwrc) }. With these indicators, we are now able to refine our grid adaptively (there are several mesh adaption strategies implemented, like 'optimized', 'fixednumber' or 'fixedfraction').

### 2012-04-17 Examples/PDE/InstatPDE/Example1/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
Added some documentation regarding the DWR-Example.

@@
-In this example we consider the nonstationary incompressible Stokes equation.
-
-\textbf{General problem description}\\
-As in the stationary PDE Example 3 \ref{PDE_Stat_FSI_STVK},
+\subsubsection{General problem description}
+In this example we consider the nonstationary incompressible Stokes equation. As in the stationary PDE Example 3 \ref{PDE_Stat_FSI_STVK},
 we use the symmetric fluid stress tensor.  

### 2012-04-17 Examples/PDE/StatPDE/Example9/content.tex
Added some documentation regarding the DWR-Example.

@@
 \subsubsection{General problem description}
-This example shows the use of the adaptive grid refinement and error estimation by the DWR-method (For a description of the method, see \cite{BR03}.) applied to the laplace equation 
+This example shows the use of the adaptive grid refinement and error estimation by the \textit{DWR method} (For a description of the method, see \cite{BR03}.) applied to the laplace equation 
 \begin{align*}
@@
 \end{align*}
-We want to estimate the error in the Functional 
+We want to estimate the error in the following functional of interest
 $$\fullfunction{J}{H^1(\Omega)}{\mathbb{R}}{u}{\frac 1 {\abs{\Gamma}} \int_\Gamma u \,dx}$$
@@
 For this setting, we have the error representation
-\begin{align*}
+\begin{align}\label{dwr.error_identity}
 J(e)=\sum_{K\in \mathbb T _h} \left\{(R_h, z-\psi_h)_K + (r_h, z-\psi_h)_{\partial_K}\right\}
-\end{align*}
-with the error $e=u-u_h$, the Triangulation $\mathbb T_h$, the dual solution $z$, arbitraty function $\phi_h \in V_h$ (the ansatz space) and the cell and edge residuals $\eval[1]{R_h}_{K} = f+\Delta u_h$ resp.
-\begin{align*}
+\end{align}
+with the error $e=u-u_h$, the Triangulation $\mathbb T_h$, the dual solution $z$, arbitraty function $\phi_h \in V_h$ (the ansatz space) and the cell- and edge-residuals: 
+\begin{align}\label{dwr.cell_residual}
+\eval[1]{R_h}_{K} = f+\Delta u_h
+\end{align}
+resp. 
+\begin{align}\label{dwr.edge_residual}
 \eval[1]{r_h}_{\Sigma} =
@@
 \frac 12 [\partial_n u_h] ,&\text{if } \Sigma \subset \partial_K \setminus \partial \Omega,\\
--0,& \text{ if } \Sigma \subset\partial \Omega.
+0,& \text{if } \Sigma \subset\partial \Omega.
 \end{cases}
-\end{align*}
+\end{align}
+
+It holds $J(u) \approx 0.441956231972232$.
 \subsubsection{Program description}
-Additionally to all the things one has to do when just solving the equation, we have to include {\verb higher_order_dwrc.h }.
-In this program, we approximate the so called 'weights' $z-\phi_h$ by a patchwise higher order interpolation of $z_h$ (the computed dual solution).
+In this section we want to focus on what you have to do if you want to enhance your existing code to use the \textit{DWR method}.
 
-The beginning of the program is basically as before, note however that we enforce patch-wise refinement of the grid by giving the flag {\verb Triangulation<2>:: } {\verb MeshSmoothing:: } {\verb patch_level_1 }
+First, additionally to all the things one has to do when just solving the equation, we have to include the file
+\begin{verbatim}
+higher_order_dwrc.h
+\end{verbatim}
+
+As we approximate the so called 'weights' $z-\phi_h$ in the error representation by a patchwise higher order interpolation of $z_h$ (the computed dual solution), we have to enforce patch-wise refinement of the grid by giving the flag
+\begin{verbatim}
+Triangulation<2>::MeshSmoothing::patch_level_1
+\end{verbatim}
 to the triangulation.
-To be able to solve the adjoint equation for the error estimation, we need to implement {\verb FaceValue_U } in \texttt{functionals.h} as well as {\verb CellEquation_U }, {\verb CellMatrix_T }, {\verb CellEquation_U } and {\verb FaceEquation_U } in {\verb localpde.h }.
-In the latter, we also have to implement {\verb StrongCellResidual }, {\verb StrongFaceResidual } and {\verb StrongBoundaryResidual }.
 
-After everything is set up for the solution of the equation, we have to define some  higher order FiniteElements, some (face)quadrature formulas (and build a \texttt{IntegratorDataContainer} with them.) and a higher order \texttt{SpaceTimeHandler}.
-We use these to define an object of the  \texttt{HigherOrderDWRC} class.
-The last step is to initialize the \texttt{HigherOrderDWRC} via \texttt{solver.InitializeHigherOrderDWRC(dwrc)}.
+To be able to solve the adjoint equation for the error estimation one needs to implement some methods regarding the equation as well es the functional of interest:
+\begin{itemize}
+\item In {\verb pdeinterface.h }
+\begin{itemize}
+\item {\verb CellEquation_U }: Weak form of the adjoint equation.
+\item {\verb CellMatrix_T }: The FE matrix for the adjoint problem.
+\item {\verb FaceEquation_U }: This one is needed in this case here because we have a functional of interest that lives on faces. 
+\end{itemize}
+\item {\verb functionalinterface.h }
+\begin{itemize}
+\item {\verb FaceValue_U }: This is the right hand side of the adjoint equation.
+\end{itemize}
+\end{itemize}
+During the  evaluation of \eqref{dwr.error_identity}, the following methods are needed
+\begin{itemize}
+\item {\verb StrongCellResidual }: The cell residual, see \eqref{dwr.cell_residual}.
+\item {\verb StrongFaceResidual }: The terms in \eqref{dwr.error_identity} that lies in the interiour (i.e. the jumps).
+\item {\verb StrongBoundaryResidual }:The terms in \eqref{dwr.error_identity} that lies on the boundary (There are none in this case).
+\end{itemize}
 
-Succeding the solution of the state equation ({\verb solver.ComputeReducedFunctionals() }), we compute the error indicators by calling { \verb solver.ComputeRefinementIndicators(dwrc) }.
-With these indicators, we are now able to refine our grid adaptively (there are several mesh adaption strategies implemented, like 'optimized', 'fixednumber' or 'fixedfraction').
+After this, we tell the problem which functional we want to use for the error estimation, this is done via
+\begin{verbatim}
+P.SetFunctionalForErrorEstimation(LFF.GetName())
+\end{verbatim}
+where \texttt{P} is of type \texttt{PDEProblemContainer} and \texttt{LFF} is the desired functional of interest.
+
+The next thing we need is an object of the type
+\begin{verbatim}
+HigherOrderDWRContainer
+\end{verbatim}
+This container takes care of the computation of the weights.
+
+To build this, we need the following:
+\begin{itemize}
+\item {\verb DOFH_higher_order }: With some higher order Finite Elemnts and the already defined triangulation, we build this \texttt{SpaceTimeHandler}.
+This is needed because  we want to use the patch-wise higher order interpolation of the weights.
+\item {\verb idc_high }: A \texttt{IntegratorDataContainer} in which we put some (face)quadrature formulas for the evaluation of the error Identity.
+\item A string which indicates how we want to store the weight-vectors (here: { \verb "fullmem" }).
+\item \texttt{pr}: The \texttt{ParameterReader} which we have already defined.
+\item A enum of type \texttt{EETerms} that tells the container, which error terms we want to compute (primal error indicators vs. dual error indicators, see \cite{BR03}).  
+\end{itemize}
+The last preparation step is now to initialize the \texttt{DWRDataContainer} with the problem in use:
+\begin{verbatim}
+  solver.InitializeHigherOrderDWRC(dwrc);
+\end{verbatim}
+
+Succeding the solution of the state equation 
+\begin{verbatim}
+solver.ComputeReducedFunctionals(); 
+\end{verbatim}
+we compute the error indicators by calling 
+\begin{verbatim}
+solver.ComputeRefinementIndicators(dwrc);
+\end{verbatim}
+We can now get the error indicators (with signs!) out of \texttt{dwrc} by 
+\begin{verbatim}
+dwrc.GetErrorIndicators();
+\end{verbatim}
+With these indicators, we are now able to refine our grid adaptively (there are several mesh adaption strategies implemented, like 'optimized', 'fixednumber' or 'fixedfraction')
+\begin{verbatim}
+DOFH.RefineSpace("optimized", &error_ind);
+\end{verbatim}
+Note, that one has to take the norm of each entry in the vector of the error indicators before feeding them into the \texttt{RefineSpace} method.

### 2012-05-03 Examples/Documentation/introduction.tex
Fixed the FIXME comment in instatreducedproblem.h, i.e., the forwardtimeloop now uses the given problem instead of recalling a new one. This also means that forwardtimestepreducedproblem is now superflous and has been removed, accordingly
Adapted the documentation and examples to the previous update regarding DoFConstraints.

@@
 \begin{itemize}
-  \item Christian Goll (University of Heidelberg)
-  \item Thomas Wick (University of Heidelberg)
+  \item Christian Goll (Heidelberg University)
+  \item Thomas Wick (Heidelberg University)
   \item Winnifried Wollner (University of Hamburg)
@@
 
-\texttt{DOpElib} requires a recent \texttt{ g++} (at least 4.4.3) due to 
-some new \texttt{ C++} features implemented in \texttt{ C++0x} and \texttt{ C++11}.
+\texttt{DOpElib} requires a recent \texttt{g++} (at least 4.4.3) due to 
+some new \texttt{C++} features implemented in \texttt{C++0x} and \texttt{C++11}.
 
@@
 This library is based upon \texttt{deal.II} hence in order to run 
-\texttt{ DOpElib} you need a running copy of \texttt{deal.II}.
+\texttt{DOpElib} you need a running copy of \texttt{deal.II}.
 
@@
 If you would like to use the features offered in our SNOPT wrapper. You will 
-need to obtain a license for \texttt{ SNOPT} 
+need to obtain a license for \texttt{SNOPT} 
 \url{http://www.sbsi-sol-optimize.com/asp/sol_product_snopt.htm}.
@@
 description in a common data format. For this we have the following 
-classes in \texttt{ DOpEsrc/container}
+classes in \texttt{DOpEsrc/container}
 \begin{itemize}
-  \item \texttt{ pdeproblemcontainer.h} Is used to describe  stationary PDE problems.
-  \item \texttt{ instatpdeproblemcontainer.h} {\bf TODO is still missing but will be used for nonstationary problems.} This will be implemented once we have nonstationary optimization problems running to avoid error duplication in the coding process.
-  \item \texttt{ optproblem.h} Is used to describe  OPT problems governed by 
+  \item \texttt{pdeproblemcontainer.h} Is used to describe  stationary PDE problems.
+  \item \texttt{instatpdeproblemcontainer.h} {\bf TODO is still missing but will be used for nonstationary problems.} This will be implemented once we have nonstationary optimization problems running to avoid error duplication in the coding process.
+  \item \texttt{optproblem.h} Is used to describe  OPT problems governed by 
     stationary PDEs. {\bf TODO should be renamed to optproblemcontainer.h including the class name!}
-  \item \texttt{ instatoptproblemcontainer.h} Is used to describe  OPT problems governed by nonstationary PDEs.
+  \item \texttt{instatoptproblemcontainer.h} Is used to describe  OPT problems governed by nonstationary PDEs.
 The only difference to the stationary case is that we need to specify a time-stepping method.
@@
 \begin{itemize}
-\item \texttt{ basic/spacetimehandler\underline{ }base.h} This class is used to define 
+\item \texttt{basic/spacetimehandler\underline{ }base.h} This class is used to define 
   an interface to the dimension independent functionality of all space time dof handlers.
   {\bf TODO: Beispiele geben}
-\item \texttt{ basic/statespacetimehandler.h} Another intermediate interface class which adds 
+\item \texttt{basic/statespacetimehandler.h} Another intermediate interface class which adds 
   the dimension dependent functionality if only the variable $u$ is considered, i.e., a 
   PDE problem.
-\item \texttt{ basic/spacetimehandler.h } Same as above but with both $q$ and $u$, i.e., for
+\item \texttt{basic/spacetimehandler.h } Same as above but with both $q$ and $u$, i.e., for
   OPT problems.
-\item \texttt{ basic/mol\underline{ }statespacetimehandler.h} Implementation of a method of 
+\item \texttt{basic/mol\underline{ }statespacetimehandler.h} Implementation of a method of 
   line space time dof handler for PDE problems. It has only one spatial 
   dofhandler that is used for all time intervals.
-\item \texttt{ basic/mol\underline{ }spacetimehandler.h} Same as above for OPT problems.
+\item \texttt{basic/mol\underline{ }spacetimehandler.h} Same as above for OPT problems.
   A separate spatial dof handler for each of the variables $q$ and $u$ is maintained 
   but only one triangulation.
-\item \texttt{ basic/mol\underline{ }multimesh\underline{ }spacetimehandler.h}
+\item \texttt{basic/mol\underline{ }multimesh\underline{ }spacetimehandler.h}
   Same as above, but now in addition the triangulations for $q$ and $u$ can be refined
@@
 \begin{itemize}
-\item \texttt{ container/celldatacontainer.h} This object is used to pass data 
+\item \texttt{container/celldatacontainer.h} This object is used to pass data 
   given on the current element (cell) of the mesh to the functions in PDE, functional, 
   $\ldots$. 
-\item \texttt{ container/facedatacontainer.h} This object is used to pass data 
+\item \texttt{container/facedatacontainer.h} This object is used to pass data 
   given on the current face of the mesh to the functions in PDE, functional, 
   $\ldots$. 
-\item \texttt{ container/multimesh\underline{ }celldatacontainer.h} This is the same as the 
+\item \texttt{container/multimesh\underline{ }celldatacontainer.h} This is the same as the 
   celldatacontainer, but it
   is capable to handle data defined on an alternative triangulation.
-\item \texttt{ container/multimesh\underline{ }facedatacontainer.h} This is the same as the
+\item \texttt{container/multimesh\underline{ }facedatacontainer.h} This is the same as the
   facedatacontainer, but it
   is capable to handle data defined on an alternative triangulation.
-\item \texttt{ container/integratordatacontainer.h} This contains some data that 
+\item \texttt{container/integratordatacontainer.h} This contains some data that 
   should be passed to the integrator like quadrature formulas and the above cell and 
@@
 \begin{itemize}
-\item \texttt{ include/forward\underline{ }euler\underline{ }problem.h}
-\item \texttt{ include/shifted\underline{ }crank\underline{ }nicolson\underline{ }problem.h}
-\item \texttt{ include/backward\underline{ }euler\underline{ }problem.h}
-\item \texttt{ include/fractional\underline{ }step\underline{ }theta\underline{ }problem.h} Note that the use of this scheme requires a special Newton solver, which is, however, already
+\item \texttt{include/forward\underline{ }euler\underline{ }problem.h}
+\item \texttt{include/shifted\underline{ }crank\underline{ }nicolson\underline{ }problem.h}
+\item \texttt{include/backward\underline{ }euler\underline{ }problem.h}
+\item \texttt{include/fractional\underline{ }step\underline{ }theta\underline{ }problem.h} Note that the use of this scheme requires a special Newton solver, which is, however, already
 implemented for the convenience of the user!
-\item \texttt{ include/crank\underline{ }nicolson\underline{ }problem.h}
+\item \texttt{include/crank\underline{ }nicolson\underline{ }problem.h}
 \end{itemize}
@@
 \begin{itemize}
-\item \texttt{ templates/integrator.h} This class computes integrals over a given 
+\item \texttt{templates/integrator.h} This class computes integrals over a given 
   triangulation (including its faces).
-\item \texttt{ templates/integrator\underline{ }multimesh.h} The same as above but it is 
+\item \texttt{templates/integrator\underline{ }multimesh.h} The same as above but it is 
   possible that some of the FE functions are defined on an other triangulation 
   as long as the have a common coarse triangulation.
-\item \texttt{ templates/integratormixeddims.h} This is used to compute integral which 
+\item \texttt{templates/integratormixeddims.h} This is used to compute integral which 
   are given in another (larger) dimension than the current variable. (This is exclusively
-  used if the control variable is given by some parameters. Which means \texttt{ dopedim == 0}). 
+  used if the control variable is given by some parameters. Which means \texttt{dopedim == 0}). 
 \end{itemize}
@@
 \begin{itemize}
-\item \texttt{ templates/newtonsolver.h} This solves some nonlinear equation using a 
+\item \texttt{templates/newtonsolver.h} This solves some nonlinear equation using a 
   line-search Newton method.
-\item \texttt{ templates/newtonsolvermixeddims.h} The same but in the case when there is 
+\item \texttt{templates/newtonsolvermixeddims.h} The same but in the case when there is 
   another variable in a (larger) dimension is involved. See 
-  \texttt{ integratormixeddims.h}.
-\item \texttt{ templates/instat\underline{ }step\underline{ }newtonsolver.h} This is a 
+  \texttt{integratormixeddims.h}.
+\item \texttt{templates/instat\underline{ }step\underline{ }newtonsolver.h} This is a 
   Newton method as above to invert the next time-step. It differs from the plain vanilla
@@
   and not in every Newton iteration.
-\item \texttt{ templates/fractional\underline{ }step\underline{ }theta\underline{ }step\underline{ }newtonsolver.h} This is the Newton solver for the time step in a 
+\item \texttt{templates/fractional\underline{ }step\underline{ }theta\underline{ }step\underline{ }newtonsolver.h} This is the Newton solver for the time step in a 
   fractional-step-theta scheme. It combines the computation of all three sub steps.
@@
 \begin{itemize}
-\item \texttt{ templates/cglinearsolver.h} This is a wrapper for the cg solver implemented in 
+\item \texttt{templates/cglinearsolver.h} This is a wrapper for the cg solver implemented in 
   \texttt{deal.II}. The solver will build and store the stiffness matrix for the PDE.
-\item \texttt{ templates/gmreslinearsolver.h} This is a wrapper for the gmres solver 
+\item \texttt{templates/gmreslinearsolver.h} This is a wrapper for the gmres solver 
   implemented in \texttt{deal.II}. The solver will build and store the stiffness matrix 
   for the PDE.
-\item \texttt{ templates/directlinearsolver.h} This is a wrapper for the direct solver 
-  implemented in \texttt{deal.II} using \texttt{ UMFPACK}. 
+\item \texttt{templates/directlinearsolver.h} This is a wrapper for the direct solver 
+  implemented in \texttt{deal.II} using \texttt{UMFPACK}. 
   The solver will build and store the stiffness matrix for the PDE.
-\item \texttt{ templates/voidlinearsolver.h} This is a wrapper for certain cases when we 
+\item \texttt{templates/voidlinearsolver.h} This is a wrapper for certain cases when we 
   know that the matrix to be inverted is the identity. It simply copies the rhs to the
@@
 \begin{itemize}
-  \item \texttt{ basic/constraints.h} This is used by the spacetimehandlers to 
+  \item \texttt{basic/constraints.h} This is used by the spacetimehandlers to 
     compute the number of constraints from the control and state vectors. 
@@
     initialized if OPT is used with box control constraints or $g(q,u) \le 0$.
-  \item \texttt{ interfaces/functionalinterface.h} This gives an interface 
+  \item \texttt{interfaces/functionalinterface.h} This gives an interface 
     for the functional $J(\cdot)$ and any other functional you may want to evaluate.
@@
     Specifically, derivatives are written therein, too.
-  \item \texttt{ interfaces/constraintinterface.h} This gives an interface for both 
+  \item \texttt{interfaces/constraintinterface.h} This gives an interface for both 
     the control box constraints as well as the general constraint $g \le 0$. This 
     needs to be specified if constraints are to be used. If they are not needed 
-    a default class \texttt{ problemdata/noconstraints.h} can be used. We note that we only 
+    a default class \texttt{problemdata/noconstraints.h} can be used. We note that we only 
     need to write the integrands on 
     elements or faces the loop over elements will be taken care of in the integrator.
-  \item \texttt{ interfaces/pdeinterface.h} This defines an interface for the 
+  \item \texttt{interfaces/pdeinterface.h} This defines an interface for the 
     partial differential equation $a(q,u)(\phi) = 0$. This needs to be written
@@
     Specifically, derivatives are written therein, too.
-  \item \texttt{ interfaces/dirichletdatainterface.h} This gives an interface to the 
+  \item \texttt{interfaces/dirichletdatainterface.h} This gives an interface to the 
     Dirichlet data for a problem. If the Dirichlet data are simply a function 
     (and do not depend on the control $q$) one can use the default class\\
-    \texttt{ problemdata/simpledirichletdata.h}.
+    \texttt{problemdata/simpledirichletdata.h}.
 \end{itemize}
@@
 \begin{itemize}
-\item \texttt{ reducedproblems/statpdeproblem.h} This is used to remove the variable $u$ in 
+\item \texttt{reducedproblems/statpdeproblem.h} This is used to remove the variable $u$ in 
   a stationary PDE problem. This means that call the method \\
-  \texttt{ StatPDEProblem::ComputeReducedFunctionals} will evaluate the functionals 
-  defined in the problem description, i.e., in \texttt{ PDEProblemContainer}, in the 
+  \texttt{StatPDEProblem::ComputeReducedFunctionals} will evaluate the functionals 
+  defined in the problem description, i.e., in \texttt{PDEProblemContainer}, in the 
   solution of the given PDE.
-\item \texttt{ reducedproblems/statreducedproblem.h} This eliminates $u$ from the OPT
+\item \texttt{reducedproblems/statreducedproblem.h} This eliminates $u$ from the OPT
   problem with a stationary PDE.
-\item \texttt{ reducedproblems/instatreducedproblem.h} The same as above but for a
+\item \texttt{reducedproblems/instatreducedproblem.h} The same as above but for a
   nonstationary PDE. {\bf FIXME there is something wrong in this file see FIXME 
     comment in the source.}
-\item \texttt{ reducedproblems/voidreducedproblem.h} A wrapper file that eliminates $u$ 
+\item \texttt{reducedproblems/voidreducedproblem.h} A wrapper file that eliminates $u$ 
   if it is not present anyways. This is used so that we can use the same routines to 
@@
 \begin{itemize}
-\item \texttt{ opt\underline{ }algorithms/reducedalgorithm.h} An interface for all 
+\item \texttt{opt\underline{ }algorithms/reducedalgorithm.h} An interface for all 
   optimization problems in the reduced formulation. It offers some test functionality
   to assert that the derivatives of the problem are computed correctly.
-\item \texttt{ opt\underline{ }algorithms/reducednewtonalgorithm.h}
+\item \texttt{opt\underline{ }algorithms/reducednewtonalgorithm.h}
   A line-search Newton algorithm using a cg method to invert the reduced hessian. 
   Implementation ignores any additional constraints.
-\item \texttt{ opt\underline{ }algorithms/reducedtrustregionnewton.h}
+\item \texttt{opt\underline{ }algorithms/reducedtrustregionnewton.h}
   A trust region Newton algorithm using a cg method to invert the reduced hessian.
   Implementation ignores any additional constraints.
-\item \texttt{ opt\underline{ }algorithms/reduced\underline{ }snopt\underline{ }algorithm.h}
+\item \texttt{opt\underline{ }algorithms/reduced\underline{ }snopt\underline{ }algorithm.h}
   An algorithm to solve reduced optimization problems with additional control constraints.
   ((reduced) state constraints are not yet implemented.)
-\item \texttt{ opt\underline{ }algorithms/reducednewtonalgorithmwithinverse.h}
+\item \texttt{opt\underline{ }algorithms/reducednewtonalgorithmwithinverse.h}
   Line-search Newton algorithm that assumes there exists a method in the reduced problem
@@
   PDE constraint.)
-\item \texttt{ opt\underline{ }algorithms/generalized\underline{ }mma\underline{ }algorithm.h}
+\item \texttt{opt\underline{ }algorithms/generalized\underline{ }mma\underline{ }algorithm.h}
   An implementation of the MMA-Algorithm for structural optimization using an augmented
@@
   special purpose file
-  \texttt{ include/augmentedlagrangianproblem.h}.
+  \texttt{include/augmentedlagrangianproblem.h}.
 \end{itemize}
@@
 \begin{itemize}
-\item \texttt{ include/statevector.h} This stores all dofs in space and time for the state 
+\item \texttt{include/statevector.h} This stores all dofs in space and time for the state 
   variable $u$. It is possible to select whether all this should be kept in memory or 
   or unused parts can be written to the hard disk.
-\item \texttt{ include/controlvector.h} This stores all dofs in space and time for the 
+\item \texttt{include/controlvector.h} This stores all dofs in space and time for the 
   control variable $q$. At present no time dependence is implemented.
-\item \texttt{ include/constraintvector.h} This stores all dofs in space and time for the 
+\item \texttt{include/constraintvector.h} This stores all dofs in space and time for the 
   non PDE constraints (and corresponding multipliers). 
@@
 \begin{itemize}
-\item \texttt{ include/parameterreader.h} This file is used to define a parameter reader
+\item \texttt{include/parameterreader.h} This file is used to define a parameter reader
   that is used to read run time parameters from a given file.
@@
 \begin{itemize}
-\item \texttt{ include/dopeexception.h} Defines some Exceptions that are thrown by the program
+\item \texttt{include/dopeexception.h} Defines some Exceptions that are thrown by the program
   should it encounter any unexpected errors.
-\item \texttt{ include/dopeexceptionhandler.h} This class is used to write information 
+\item \texttt{include/dopeexceptionhandler.h} This class is used to write information 
   contained in the exceptions to the output in a uniform manner.
@@
 \begin{itemize}
-\item \texttt{ include/outputhandler.h} This file defines an outputhandler object which 
+\item \texttt{include/outputhandler.h} This file defines an outputhandler object which 
   can be used to decide whether some information should be written to screen or file.
@@
 \begin{itemize}
-\item \texttt{include/constraintsmaker.h} This class sets the constraints on the DOFs of the state FE solution.
-In the standard implementation, only the hanging-node-constraints are built, but the user can reimplement this class and then include other constraints as well (for example periodic BC).
-{\bf TODO: to rename the file?? (CG)}
+\item \texttt{include/userdefineddofconstraints.h} This class sets the constraints on the DOFs of the state and/or control FE solution.
+DOpE itself builds the hanging-node-constraints, but the user can reimplement this class and thus include other constraints as well (for example periodic BC).
+Note, that the hanging-node-constraints come first (in case of conflicting constraints.)
 \item \texttt{include/sparsitymaker.h} This class sets the sparsity pattern for the state FE solution.
@@
             helper.h
-
-reducedproblems:
- forwardtimestepreducedproblem.h

### 2012-05-03 Examples/description.tex
Add Literature and update foreword
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
Adapted the documentation and examples to the previous update regarding DoFConstraints.

@@
 \documentclass[a4paper,cleardoubleempty]{scrreprt}
-\usepackage{amsmath,amssymb}
+\usepackage{amsmath,amssymb, commath}
 \usepackage[pdftex,bookmarks=false,pdfpagelabels=true,plainpages=false]{hyperref}
@@
 \author{
-  Christian Goll\thanks{University of Heidelberg, 
+  Christian Goll\thanks{Heidelberg University, 
     {christian.goll@iwr.uni-heidelberg.de}}, 
-  Thomas Wick\thanks{University of Heidelberg, 
+  Thomas Wick\thanks{Heidelberg University, 
     {thomas.wick@iwr.uni-heidelberg.de}}, 
@@
 \input{PDE/StatPDE/Example8/content.tex}
+\clearpage
+\subsection{Adaptive Solution of Laplace Equation in 2D}
+\label{PDE_adap_Stat_Laplace}
+\input{PDE/StatPDE/Example9/content.tex}
 \clearpage
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\bibliographystyle{plain}
+\bibliography{Documentation/lit}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \addcontentsline{toc}{chapter}{Index}

### 2012-05-03 Examples/OPT/StatPDE/Example5/main.cc

@@ -13,7 +13,7 @@
 #include "simpledirichletdata.h"
 #include "noconstraints.h"
 #include "sparsitymaker.h"
-#include "constraintsmaker.h"
+#include "userdefineddofconstraints.h"
 #include "preconditioner_wrapper.h"
 #include "integratordatacontainer.h"
 
@@ -107,7 +107,7 @@ int main(int argc, char **argv)
   dealii::GridGenerator::hyper_cube(times);
   triangulation.refine_global (3);
 
-  MethodOfLines_MultiMesh_SpaceTimeHandler<FE, DOFHANDLER, BlockSparsityPattern,VECTOR, SparsityMaker<DOFHANDLER,2>, ConstraintsMaker<DOFHANDLER,2>, 2> DOFH(triangulation,control_fe, state_fe);
+  MethodOfLines_MultiMesh_SpaceTimeHandler<FE, DOFHANDLER, BlockSparsityPattern,VECTOR, 2> DOFH(triangulation,control_fe, state_fe);
 
   NoConstraints<Multimesh_CellDataContainer,Multimesh_FaceDataContainer,DOFHANDLER,VECTOR, 2,2> Constraints;
 

### 2012-05-03 Examples/PDE/InstatPDE/Example1/main.cc

@@ -22,9 +22,8 @@
 #include "simpledirichletdata.h"
 #include "noconstraints.h"
 #include "sparsitymaker.h"
-#include "constraintsmaker.h"
+#include "userdefineddofconstraints.h"
 #include "integratordatacontainer.h"
-
 
 #include <iostream>
 #include <fstream>
@@ -42,14 +41,11 @@
 
 #include <fe/fe_q.h>
 
-
 #include "localpde.h"
 #include "localfunctional.h"
 #include "functionals.h"
 
 #include "my_functions.h"
-
-
 
 using namespace std;
 using namespace dealii;
@@ -67,31 +63,34 @@ using namespace DOpE;
 #define DD DOpE::DirichletDataInterface<VECTOR,LOCALDOPEDIM,LOCALDEALDIM>
 #define CONS DOpE::ConstraintInterface<CellDataContainer,FaceDataContainer,DOFHANDLER,VECTOR,LOCALDOPEDIM,LOCALDEALDIM>
 
-typedef OptProblem<FUNC,FUNC,PDE,DD,CONS,SPARSITYPATTERN, VECTOR, LOCALDOPEDIM,LOCALDEALDIM> OP_BASE;
+typedef OptProblem<FUNC, FUNC, PDE, DD, CONS, SPARSITYPATTERN, VECTOR,
+    LOCALDOPEDIM, LOCALDEALDIM> OP_BASE;
 
 #define PROB DOpE::StateProblem<OP_BASE,PDE,DD,SPARSITYPATTERN,VECTOR,LOCALDOPEDIM,LOCALDEALDIM>
 // Typedefs for timestep problem
-typedef ShiftedCrankNicolsonProblem<PROB,SPARSITYPATTERN, VECTOR, LOCALDOPEDIM,LOCALDEALDIM> TSP;
+typedef ShiftedCrankNicolsonProblem<PROB, SPARSITYPATTERN, VECTOR, LOCALDOPEDIM,
+    LOCALDEALDIM> TSP;
 
-typedef InstatOptProblemContainer<TSP,FUNC,FUNC,PDE,DD,CONS,SPARSITYPATTERN, VECTOR, LOCALDOPEDIM,LOCALDEALDIM> OP;
+typedef InstatOptProblemContainer<TSP, FUNC, FUNC, PDE, DD, CONS,
+    SPARSITYPATTERN, VECTOR, LOCALDOPEDIM, LOCALDEALDIM> OP;
 
[…]

### 2012-05-03 Examples/Documentation/introduction.tex
Updated the Documentation according to last ci

@@
     {\bf Similar descriptors will be build for the other problems (adjoint, tangent, $\ldots$) when time allows.}
+  \item \texttt{problemdata/initialproblem.h} This is the problem descriptor to compute the finite element representation
+    of the initial values. This is generated by the different time-stepping schemes based upon the defined 
+    representation by the pde, which is set to the component wise $L^2$ projection by default.
   \item \texttt{problemdata/primaldirichletdata.h} This class contains the Dirichlet data for the 

### 2012-05-04 Examples/PDE/InstatPDE/Example2/main.cc
Adaptions in the nonstationary case for the dual-timesteppingscheme

@@ -65,7 +65,9 @@ typedef OptProblem<FUNC, FUNC, PDE, DD, CONS, SPARSITYPATTERN, VECTOR,
 
 // Typedefs for timestep problem
 typedef BackwardEulerProblem<PROB, SPARSITYPATTERN, VECTOR, 2, 2> TSP;
-typedef InstatOptProblemContainer<TSP, FUNC, FUNC, PDE, DD, CONS,
+//FIXME: This should be a reasonable dual timestepping scheme
+typedef BackwardEulerProblem<OP_BASE, SPARSITYPATTERN, VECTOR, 2, 2> DTSP;
+typedef InstatOptProblemContainer<TSP, DTSP,FUNC, FUNC, PDE, DD, CONS,
     SPARSITYPATTERN, VECTOR, LOCALDOPEDIM, LOCALDEALDIM> OP;
 typedef IntegratorDataContainer<DOFHANDLER, dealii::Quadrature<LOCALDEALDIM>,
     dealii::Quadrature<LOCALDEALDIM - 1>, VECTOR, LOCALDEALDIM> IDC;

### 2012-05-04 DOpEsrc/include/crank_nicolson_problem.h
Removed some duplicated code from the ts-schemes and put it into ts_base.h resp. primal_ts_base.h. Moved the time stepping schemes to the folder tsschems

@@ -1,1026 +1,0 @@
-#ifndef _CRANKNICOLSONProblem_H_
-#define _CRANKNICOLSONProblem_H_
-
-#include "initialproblem.h" 
-
-namespace DOpE
-{
-  /**
-   * Class to compute time dependent problems with the Crank-Nicolson
-   * time stepping scheme.
-   *
-   * @tparam <OPTPROBLEM>       The problem to deal with.
-   * @tparam <SPARSITYPATTERN>  The sparsity pattern for control & state.
-   * @tparam <VECTOR>           The vector type for control & state (i.e. dealii::Vector<double> or dealii::BlockVector<double>)
-   * @tparam <FE>               The type of finite elements in use, must be compatible with the DOFHANDLER.
-   * @tparam <DOFHANDLER>       The type of the DoFHandler in use (to be more precise: The type of the dealii-DoFhandler which forms
-   *                            the base class of the DOpEWrapper::DoFHandler in use.)
-   * @tparam <dopedim>          The dimension for the control variable.
-   * @tparam <dealdim>          The dimension of the state variable.
-   *
-   */
-  template<typename OPTPROBLEM, typename SPARSITYPATTERN, typename VECTOR,
-      int dopedim, int dealdim, typename FE = DOpEWrapper::FiniteElement<
-          dealdim>, typename DOFHANDLER = dealii::DoFHandler<dealdim>>
-    class CrankNicolsonProblem
-    {
-    public:
-      CrankNicolsonProblem(OPTPROBLEM& OP) :
-        _OP(OP)
-      {
-      	_initial_problem = NULL;
-      }
-      ~CrankNicolsonProblem ()
-      {
-	if(_initial_problem != NULL)
-	  delete _initial_problem;
-      }
-
-      /******************************************************/
-
-      /**
-       * Returns the name of the time stepping scheme.
-       *
-       * @return A string containing the name of the time stepping scheme.
-       */
-      std::string
-      GetName()
-      {
-        return "Crank-Nicolson";
-      }
-      /******************************************************/
-
-      InitialProblem<CrankNicolsonProblem<OPTPROBLEM, SPARSITYPATTERN, VECTOR, dopedim,
-          dealdim, FE, DOFHANDLER>, VECTOR, dealdim>&
-      GetInitialProblem()
-      {
-	if (_initial_problem == NULL)
-	{
[…]

### 2012-05-05 Examples/OPT/InstatPDE/Example1/content.tex
Added an initial Example for the OPT-Instat case, it is not yet working, but will be soon...

\subsubsection{General problem description}

This example is a modified version of \texttt{PDE/InstatPDE/Example5}. 
Again, we consider the heat equation, this time with an additional nonlinear term
\begin{equation*}
\partial_t u(t,x,y) - \Delta u(t,x,y) + u(t,x,y)^2 = f(t,x,y),
\end{equation*}
with homogeneous dirichlet-data.
The computational domain is $\Omega \times I = [0,\pi]^2 \times [0,1]$. From the known solution, we can compute the appropriate data 
\begin{align*}
f(t,x,y) &= (3-2t)e^{t-t^2} \sin(x) \sin(y) + e^{(t-t^2)^2} \sin^2(x) \sin^2(y),\\
u_0(x,y) &= q(x,y).
\end{align*}

With the cost functional 
\[
 \min_{q,u} J(q,u) = \frac{1}{2}\int_{\Omega} (u(1,x,y) - \sin(x) \sin(y))^2\,d(x,y) + \frac{1}{2} \int_{\Omega} (q(x,y) - \sin(x) \sin(y))^2\,d(x,y).
\]
It has the optimal solution 
\begin{align*}
\overline{u}(t,x,y) &= e^{t-t^2} \sin(x) \sin(y),\\
\overline{q}(t,x,y) &= \sin(x) \sin(y)
\end{align*}
and $J(\overline{q},\overline{u}) = 0$.

### 2012-05-05 Examples/OPT/InstatPDE/Example1/functionals.h

@@ -1,0 +1,120 @@
+#ifndef _LOCALFunctionalS_
+#define _LOCALFunctionalS_
+
+#include "pdeinterface.h"
+
+#define PI 3.14159265359
+
+using namespace std;
+using namespace dealii;
+using namespace DOpE;
+
+
+/****************************************************************************************/
+
+template<typename VECTOR, int dopedim, int dealdim>
+  class LocalPointFunctional : public FunctionalInterface<CellDataContainer,FaceDataContainer,dealii::DoFHandler<dealdim>, VECTOR, dopedim,dealdim>
+  {
+  private:
+    mutable double time;
+
+  public:
+
+    void SetTime(double t) const
+    {
+      time = t;
+    }
+
+    bool NeedTime() const
+    {
+      if (time==0.)
+    		return true;
+    	else
+    		return false;
+    }
+
+  double PointValue(const DOpEWrapper::DoFHandler<dopedim, dealii::DoFHandler<dealdim> > &/* control_dof_handler*/,
+		    const DOpEWrapper::DoFHandler<dealdim, dealii::DoFHandler<dealdim> > & state_dof_handler,
+		    const std::map<std::string, const dealii::Vector<double>* > &/*param_values*/,
+		    const std::map<std::string, const VECTOR* > &domain_values)
+  {
+
+    Point<2> evaluation_point(0.5*PI,0.5*PI);
+
+    typename map<string, const VECTOR* >::const_iterator it = domain_values.find("state");
+
+    double point_value = VectorTools::point_value(state_dof_handler, *(it->second), evaluation_point);
+
+    // pressure analysis
+    return point_value;
+  }
+
+  string GetType() const
+  {
+    return "point timelocal";
+    // 1) point domain boundary face
+    // 2) timelocal timedistributed
+  }
+  string GetName() const
+  {
+    return "Start-Time-Point evaluation";
+  }
+
+  };
+
+/************************************************************************************************************************************************/
+
+template<typename VECTOR, int dopedim, int dealdim>
+  class LocalPointFunctional2 : public FunctionalInterface<CellDataContainer,FaceDataContainer,dealii::DoFHandler<dealdim>, VECTOR, dopedim,dealdim>
+  {
+  private:
+    mutable double time;
+
+  public:
+
+    void SetTime(double t) const
+    {
+      time = t;
+    }
+
+    bool NeedTime() const
+    {
+      if (time==1.)
+    		return true;
+    	else
+    		return false;
[…]

### 2012-05-07 Paper_May_2012/main.tex
Vorlage und ersten Text fuer das Paper

--- 2012-05-03 Examples/Documentation/introduction.tex
@@
-\chapter{Introduction}\label{chap:intro}
-\section{License information}
-{\bf TODO:} Unter welcher License soll DOpE laufen?
+\documentclass[twoside, 11pt]{article}
+\usepackage{jnm}
+%  
+\author{Christian\,Goll\footnotemark[1]\comma Thomas\,Wick\footnotemark[1] 
+\comma Winnifried\,Wollner\footnotemark[1]\thanks{%
+Institut f\"ur Angewandte Mathematik, Universit\"at Heidelberg, Im Neuenheimer Feld 293/294,
+			D-69120 Heidelberg, Germany.}
+}
 
-\section{Developers}
-The library is currently maintained by 
+\sauthor{C. Goll, T. Wick and  W. Wollner}
+
+\title{DOpE: A Goal Oriented Software Library for Computing PDEs and Optimization Problems}
+\stitle{DOpE: A Goal Oriented Software Library}
+\received{Received \hfill}
+%
+%%% User-requested packages placed after this line %%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{amsfonts,amssymb,amsmath}
+\usepackage{booktabs,dcolumn}
+\usepackage[T1]{fontenc}
+\usepackage{mathtools}
+\usepackage{enumerate,graphicx}
+\usepackage{subfig}
+\usepackage{multirow}
+% \usepackage[center]{subfigure}
+\numberwithin{equation}{section}
+\usepackage[noend]{algorithmic}                        %Algorithm typesetting environment
+\usepackage[ruled]{algorithm}
+\usepackage{braket}                             %Provides \Set for typing sets
+\usepackage{color}
+\usepackage[square,sort&compress,numbers]{natbib}
+\usepackage{float}
+\usepackage{subfig}
+
+%%% User's macros placed after this line %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\IfFileExists{dsfont.sty}%
+%\usepackage{dsfont}
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\N}{\mathbb{N}}
+\renewcommand{\phi}{\varphi}
+\renewcommand{\epsilon}{\varepsilon}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\abstract{%
+In this article, we describe the software library 
+package \textit{Deal Optimization Environment} (DOpE).
+Its main feature is to give a unified interface to high level algorithms such as 
+time stepping methods, nonlinear solvers and optimization routines. The performance and features 
+of the software package are demonstrated by computing 
+four numerical tests.
+
+}
+
+\keywords{%
+
+}
+
+%
+\begin{document}
+\maketitle
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
+\label{introduction}
+The \textit{Deal Optimization Environment} (DOpE) project is 
+based on the \textit{deal.II} \citep{deal} finite element library which has been developed
+ initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{deal}.
+Its main feature is to give a unified interface to high level algorithms such as 
+time stepping methods, nonlinear solvers and optimization routines. 
+We aim that the user should only need to write those parts
+of the code that are problem dependent while all invariant parts of the algorithms
+should be reusable without any need for further coding.
+In particular, the user should be able to switch between various different 
+algorithms without the need to rewrite the problem dependent code, though he or she will
+have to replace the algorithm object with an other one.
+
+The authors acknowledge their past experience as well as discussions with 
+the authors of the libraries 
+Gascoigne/RoDoBo project, which was initiated by 
+Roland Becker, Dominik Meidner,  and Boris Vexler \cite{rodobo}. 
+From which some of the ideas to modularize the algorithms have arisen.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The aim of DOpE is to provide a software toolkit to solve forward PDE
+problems as well as optimal control problems constrained by PDE. The
+solution of a broad variety of PDE is possible in deal.II as well, but
+DOpE concentrates on a unified approach for both linear and nonlinear
+problems by interpreting every PDE problem as nonlinear and applying a
+Newton method to solve it. While deal.II leaves much of the work and many
+decisions to the user, DOpE intends to be user-optimized by delivering
+prefabricated tools which require from the user only adjustments connected
+to his specific problem. The solution of optimal control problems with PDE
+constraints is an innovation in the DOpE framework.
+The focus is on the numerical solution of both stationary and nonstationary
+problems which come from different application fields, like elasticity and
+plasticity, fluid dynamics, and fluid-structure interactions.
+
+At the present stage the following features are supported by the library
 \begin{itemize}
-  \item Christian Goll (Heidelberg University)
-  \item Thomas Wick (Heidelberg University)
-  \item Winnifried Wollner (University of Hamburg)
+\item Solution of stationary and nonstationary PDEs in 1d, 2d, and 3d.
+\item Various time stepping schemes (based on finite differences), 
+  such as forward Euler, backward Euler,
+  Crank-Nicolson, shifted Crank-Nicolson, and Fractional-Step-$\Theta$ scheme.
+\item All finite elements of from deal.II including hp-support.
+\item Several examples showing the solution of several PDEs including
+   Poisson, Navier-Stokes, Plasticity and fluid-structure interaction problems. 
+\item Self written line search and trust region newton algorithms for the 
+   solution of optimization problems with PDEs \cite{NoWr00}
+\item Interface to SNOPT for the solution of optimization problems with PDEs and
+  additional other constraints.
+\item Several examples showing how to solve various kinds of optimization problems
+  involving stationary PDE constraints.
+\item Mesh adaptation.
+\item Different spatial triangulations for control and state variables.
 \end{itemize}
 
-However, there are more highly appreciated contributions
-made by %(in alphabetical order)
-\begin{itemize}
-  \item Michael Geiger (Examples for Plasticity, and Documentation of several PDE-Examples)
-\end{itemize}
+The rest of this document is structured as follows: We start with an introduction in
+Chapter~\ref{chap:intro} where you will learn what is needed to run {\tt DOpElib}. 
+Further you will learn what problems we can solve and how all the different classes 
+work together for this purpose. This should help you figure out what the different classes
+do if you are in need of writing your own algorithm.
 
-\section{Software requirements}
-The library DOpE has been tested to work on both Linux and 
-MAC OSX. See also the \texttt{README.OSX} file
+Then assuming that you can work to your satisfaction with the algorithms already implemented
+we will show you how to create your own running example in Chapter~\ref{chap:howtoex}.
+This will be followed by a detailed description of all examples already shipped with 
+the library. You can find the examples for the solution of PDEs in Chapter~\ref{PDE}
+and those for the solution of optimization problems with PDEs in Chapter~\ref{OPT}.
 
-\texttt{DOpElib} requires a recent \texttt{g++} (at least 4.4.3) due to 
-some new \texttt{C++} features implemented in \texttt{C++0x} and \texttt{C++11}.
+These notes conclude with a section that explains how we do automated testing of the 
+implementation in Chapter~\ref{chap:test}. This chapter will be of interest only if you 
+are trying to implement some new features to the library so that you can check that 
+the new code did not break anything.
 
-\subsection{deal.II}
-This library is based upon \texttt{deal.II} hence in order to run 
-\texttt{DOpElib} you need a running copy of \texttt{deal.II}.
-
-The \texttt{deal.II} library is open source and is freely available for noncommercial project.
-It can be downloaded from \url{http://www.dealii.org/}. On this
-homepage, one also finds lots of further information on deal.II as well as
-an extensive tutorial where many features of deal.II are discussed in a
-well-documented example framework. In order to use DOpE, it is highly
-recommendable to be roughly acquainted with deal.II.
-
-When installing \texttt{deal.II} you should take care to configure 
-it to use UMFPACK, i.e.,\\
-\texttt{./configure --with-umfpack}
-
-\subsection{ThirdPartyLibraries}
-In order for \texttt{DOpE} to be able to auto-detect some of the installed 
-Third Party Libraries you should generate according links in the 
-\texttt{ThirdPartyLibs}. See also \texttt{ThirdPartyLibs/README}.
-
-\subsubsection{SNOPT}
-If you would like to use the features offered in our SNOPT wrapper. You will 
-need to obtain a license for \texttt{SNOPT} 
-\url{http://www.sbsi-sol-optimize.com/asp/sol_product_snopt.htm}.
-Unfortunately this is at present not available for free, but you should 
-check if 
-there is a department license already available.
-For further information you should consult the file 
-\texttt{ThirdPartyLibs/SNOPT.INSTALLNOTES}. In particular you need to configure 
-\texttt{deal.II} with at least the following options:\\
-\texttt{./configure --with-umfpack --with-blas --with-lapack}
-
-\section{Installation}
-I) Get a copy of \texttt{DOpElib} from the svn repository using\\
-\texttt{svn co --username=[your username] \textbackslash \\ \hspace*{5mm}https://ganymed.iwr.uni-heidelberg.de/svn/dope}\\
-(The backslash in the first line means that the line will be continued as one!)
-Note that you need a valid username and password. If you have none you can contact the 
-maintainer by sending an EMail to \url{dope@ganymed.iwr.uni-heidelberg.de}.\\[3mm]
-%
-II) Install \texttt{deal.II} to your home directory, i.e., it should be located in \\
-\texttt{\textasciitilde/deal.II}. \\[1mm]
-If you would like to have another path you will either have to manually edit the files
-\texttt{DOpEsrc/source/Makefile} and \texttt{Examples/Make.global\underline{ }options} and replace the 
-line \\
-\texttt{D = \$(HOME)/deal.II} with the appropriate path.\\[1mm]
-Configure it properly and make the whole \texttt{deal.II} library.\\[3mm]
-%
-III) If you want to use some of the supported third party libraries install them and follow 
- the instructions in \texttt{ThirdPartyLibs/README}. There may be further information 
- in some \texttt{ThirdPartyLibs/*.INSTALLNOTES} that you may want to consider.\\[3mm]
-%
-IV) To build the \texttt{DOpElib} you have to change to the directory 
-\texttt{DOpEsrc} where you can call\\ 
-\texttt{make all}\\
-to build the library. If you want to generate some documentation you may use
-\texttt{make pdf-doc} or \texttt{make html-doc} which will require either 
-latex or doxygen to be installed on your computer.\\[3mm]
-%
-V) To build the Examples change to the \texttt{Examples} directory and \texttt{make all}.
-%
-VI) Finally you may wish to test if everything worked. To do so you can 
-change to the \texttt{Examples} directory and \texttt{make tests} which will give you a 
-list of all the examples and whether they behave as expected by the library, see also 
-Chapter~\ref{chap:test}.
-
-\section{How this library is structured}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Detailed Description of the Features}
 This library is designed to allow easy implementation and numerical solutions 
@@
     by SNOPT using our library.
-  \item \texttt{wrapper/finiteelement\underline{ }wrapper.h} {\bf Will be removed soon!}
   \item \texttt{wrapper/fevalues\underline{ }wrapper.h} {\bf Will be removed soon!}
@@
 {\bf TODO! The following files need some appropriate description here 
-This should be done by someone who knows what they are used for. Please move the description to the right place above.}
+This should be done by someone who knows what they are used for. Please move the description to the right place above.} tsschemes.
 
-\marginpar{TODO}
+\subsection{Interfaces to other software packages}
+deal.II and SNOPT
 
-basic:
-      sth\underline{ }internals.h
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Numerical Examples}
 
-include:   
-            helper.h
+\subsection{Nonstationary Fluid-Structure Interaction}
+The results are already obtained
+
+\subsection{Compliance minimization}
+Results already obtained.
+
+\subsection{Goal-oriented mesh refinement for Navier-Stokes}
+Has to be computed.
+
+\subsection{Dirichlet Velocity, Neumann Pressure, and
+Periodic Boundary Conditions}
+I do not know.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Acknowledgment}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\bibliographystyle{abbrvnat}
+\bibliographystyle{plainnat}
+\bibliography{lit}
+%
+
+\end{document}

### 2012-05-07 Examples/OPT/StatPDE/Example2/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
OPT StatPDE Example2 uses now a pointfunctional as cost functional.

@@
-This example solves the distributed minimization problem
-\begin{gather*}
-\min J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2\\
-\text{s.t.} (\nabla u,\nabla \phi) = (f(q),\phi)\;\;\forall\,\phi \in H^1_0(\Omega; \mathbb R^2)
-\end{gather*}
-on the domain $\Omega = [0,1]^2$, and the data is chosen as follows:
+\subsubsection{General problem description}
+This example solves the following pointwise minimization problem
+\begin{align*}
+\min_{(q,u)\in \mathbb R^3 \times H_0^1(\Omega; \mathbb R^2)} J(q,u) &= \frac{1}{2} \sum_{i=0}^2\abs{(u-\overline u)(x_i)}^2 + \frac{\alpha}{2}\|q\|^2\\
+\text{s.t.} (\nabla u,\nabla \phi) &= (f(q),\phi)\;\;\forall\,\phi \in H^1_0(\Omega; \mathbb R^2)
+\end{align*}
+on the domain $\Omega = [0,1]^2$, with
+\begin{itemize}
+\item the observation points
+\begin{align*}
+x_0 = (0.5, 0.5), \quad x_1 = (0.5, 0.25),\quad x_2 = (0.25, 0.25),
+\end{align*}
+\item the regularization parameter $\alpha = 0$, 
+\item the right hand side
 \begin{align*}
@@
 \end{align*}
-with $\alpha = 0$ and $u^d$ such that the solution is given by:
-\begin{gather*}
- \overline{q} = \bigl(1;0.5;1\bigr)\\
- \overline{u} = \left(\begin{matrix} \sin( \pi x)( \sin(\pi y)+0.5\sin(2\pi y))\\\sin(2\pi x) \sin(2\pi y) \end{matrix}\right)
-\end{gather*}
+\item and the exact solution given by 
+\begin{align*}
+ \overline{q} &= \bigl(1;0.5;1\bigr)\\
+ \overline{u}& = \left(\begin{matrix} \sin( \pi x)( \sin(\pi y)+0.5\sin(2\pi y))\\\sin(2\pi x) \sin(2\pi y) \end{matrix}\right).
+\end{align*}
+\end{itemize}
+\subsubsection{Program description}
+Notice that the cost functional is of mixed type (from our computational point of view), i.e.
+the first part is a pointfunctional whereas the regularization part requires the evaluation of a domain integral.
+So our \texttt{LocalFunctional} returns as his type:
+\begin{verbatim}
+      string
+      GetType() const
+      {
+        return "point domain";
+      }
+\end{verbatim}
+This brings along that we have not only to implement 
+\begin{itemize} 
+\item \texttt{Value},
+\item  \texttt{Value\_U},
+\item \texttt{Value\_Q}, 
+\item \texttt{Value\_UU},
+\item \texttt{Value\_UQ},
+\item \texttt{Value\_QU},
+\item \texttt{Value\_QQ}
+\end{itemize}
+ from \texttt{FunctionalInterface}, but also all the aforementioned methods with a preceeding \text{Point} (\texttt{PointValue} etc.).

### 2012-05-07 Examples/OPT/StatPDE/Example2/localfunctional.h

@@ -1,185 +1,360 @@
 #ifndef _LOCALFunctional_
 #define _LOCALFunctional_
 
+#include <numerics/vectors.h>
+#include <base/quadrature.h>
+#include <fe/fe_values.h>
+#include <grid/grid_tools.h>
+#include <fe/mapping_q1.h>
+
 #include "functionalinterface.h"
+#include "myfunctions.h"
 
 using namespace std;
 using namespace dealii;
 using namespace DOpE;
 
 template<typename VECTOR, int dopedim, int dealdim>
-  class LocalFunctional : public FunctionalInterface<CellDataContainer,FaceDataContainer,dealii::DoFHandler<dealdim>, VECTOR, dopedim,dealdim>
+  class LocalFunctional : public FunctionalInterface<CellDataContainer,
+      FaceDataContainer, dealii::DoFHandler<dealdim>, VECTOR, dopedim, dealdim>
   {
-  public:
-  LocalFunctional()
+    public:
+      LocalFunctional()
       {
-	_alpha = 0.;
+        _alpha = 0.;
+        _eval_points.resize(3);
+        //for q0
+        _eval_points[0][0] = 0.5;
+        _eval_points[0][1] = 0.5;
+        //for q1
+        _eval_points[1][0] = 0.5;
+        _eval_points[1][1] = 0.25;
+        //for q2
+        _eval_points[2][0] = 0.25;
+        _eval_points[2][1] = 0.25;
       }
 
-      double Value(const CellDataContainer<dealii::DoFHandler<dealdim>, VECTOR, dealdim>& cdc)
+      double
+      Value(
+          const CellDataContainer<dealii::DoFHandler<dealdim>, VECTOR, dealdim>& cdc)
       {
-	const DOpEWrapper::FEValues<dealdim> & state_fe_values = cdc.GetFEValuesState();
-	unsigned int n_q_points = cdc.GetNQPoints();
-	{
-	  _qvalues.reinit(3);
-	  _fvalues.resize(n_q_points,Vector<double>(2));
-	  _uvalues.resize(n_q_points,Vector<double>(2));
+        const DOpEWrapper::FEValues<dealdim> & state_fe_values =
+            cdc.GetFEValuesState();
+        unsigned int n_q_points = cdc.GetNQPoints();
+        {
+          _qvalues.reinit(3);
+          cdc.GetParamValues("control", _qvalues);
+        }
+        double r = 0.;
 
-	  cdc.GetParamValues("control",_qvalues);
[…]

### 2012-05-08 Examples/Documentation/introduction.tex
Removed some duplicated code from the ts-schemes and put it into ts_base.h resp. primal_ts_base.h. Moved the time stepping schemes to the folder tsschems
FiniteElementWrapper class from description removed
Update of the documentation.

@@
   \item \texttt{instatpdeproblemcontainer.h} {\bf TODO is still missing but will be used for nonstationary problems.} This will be implemented once we have nonstationary optimization problems running to avoid error duplication in the coding process.
-  \item \texttt{optproblem.h} Is used to describe  OPT problems governed by 
-    stationary PDEs. {\bf TODO should be renamed to optproblemcontainer.h including the class name!}
+  \item \texttt{optproblemcontainer.h} Is used to describe  OPT problems governed by 
+    stationary PDEs. 
   \item \texttt{instatoptproblemcontainer.h} Is used to describe  OPT problems governed by nonstationary PDEs.
@@
 \begin{itemize}
-\item \texttt{include/forward\underline{ }euler\underline{ }problem.h}
-\item \texttt{include/shifted\underline{ }crank\underline{ }nicolson\underline{ }problem.h}
-\item \texttt{include/backward\underline{ }euler\underline{ }problem.h}
-\item \texttt{include/fractional\underline{ }step\underline{ }theta\underline{ }problem.h} Note that the use of this scheme requires a special Newton solver, which is, however, already
+\item \texttt{tsschemes/forward\underline{ }euler\underline{ }problem.h}
+\item \texttt{tsschemes/shifted\underline{ }crank\underline{ }nicolson\underline{ }problem.h}
+\item \texttt{tsschemes/backward\underline{ }euler\underline{ }problem.h}
+\item \texttt{tsschemes/fractional\underline{ }step\underline{ }theta\underline{ }problem.h} Note that the use of this scheme requires a special Newton solver, which is, however, already
 implemented for the convenience of the user!
-\item \texttt{include/crank\underline{ }nicolson\underline{ }problem.h}
+\item \texttt{tsschemes/crank\underline{ }nicolson\underline{ }problem.h}
 \end{itemize}
@@
   \item \texttt{basic/dopetypes.h} This file contains type definitions used in the library. 
+\item \item \texttt{basic/sth\underline{ }internals.h} Wrapper for the \texttt{MapDoFsToSupportPoints} function. The implementation of this changes with the \texttt{deal.II} version in use.
+\item \texttt{include/helper.h} Collection of various small helper functions.
+\item \texttt{tsschemes/primal\_ts\_base.h}  This class contains the methods which all primal time stepping schemes share.
+\item \texttt{tsschemes/ts\_base.h}  This class contains the methods which all time stepping schemes share.
 \end{itemize}
@@
 {\bf TODO! The following files need some appropriate description here 
-This should be done by someone who knows what they are used for. Please move the description to the right place above.}
+This should be done by someone who knows what they are used for. Please move the description to the right place above.}.
 
 \marginpar{TODO}
-
-basic:
-      sth\underline{ }internals.h
-
-include:   
-            helper.h

### 2012-05-16 Examples/Documentation/introduction.tex
Now the Example for Control of Nonstationary PDEs via the Initial Values is running. However there seems to be a bug in the hessian computation that hopefully will be fixed soon.

@@
 
-However, there are more highly appreciated contributions
+Furthermore, there are more highly appreciated contributions
 made by %(in alphabetical order)
@@
 \subsection{Internal structures}
-{\bf TODO: ab hier habe ich noch nichts gemacht (TW)}
 \subsubsection{Interface Classes}
@@
 \end{itemize}
-
-{\bf TODO! The following files need some appropriate description here 
-This should be done by someone who knows what they are used for. Please move the description to the right place above.}.
-
-\marginpar{TODO}

### 2012-05-16 Examples/OPT/InstatPDE/Example1/content.tex
Now the Example for Control of Nonstationary PDEs via the Initial Values is running. However there seems to be a bug in the hessian computation that hopefully will be fixed soon.

@@
 \end{align*}
-and $J(\overline{q},\overline{u}) = 0$.
+and $J(\overline{q},\overline{u}) = 0$.\\[4mm]
+
+The following new things differ from the PDE-Eamples and the optimization examples with stationary 
+PDEs:\\[2mm]
+In the file \texttt{main.cc} the SpaceTimeHandler now gets an additional argument.
+In this case \texttt{DOpEtypes::initial} which specifies that the control is entering in the 
+initial value.\\[2mm]
+In the file \texttt{localpde.h} we now have to specify the Methods \texttt{Init\underline{ }CellRhs}
+and \texttt{Init\underline{ }CellRhs\underline{ }Q}. They need to be adapted, since usualy the InitialValue for the 
+PDE is auto generated from a deal.ii function in the ProblemContainer. This Value is set in the 
+\texttt{Init\underline{ }CellRhs} hence we need to change this function to use the control instead. 
+Correspondingly we need to implement the first derivative of this with respect to the control. 
+--We don't need the second derivative since it is zero anyways.--\\[2mm]
+
+{\bf Note:} In contrast to the ,,normal'' Cell-terms in the PDE we assume in the program 
+that the \texttt{Init\underline{ }CellEquation} is linear in the state and no other solution variables are present.
+Thus no derivatives of the \texttt{Init\underline{ }CellEquation} need to be implemented.

### 2012-05-16 Examples/OPT/InstatPDE/Example1/main.cc

@@ -2,13 +2,11 @@
 #include "instatoptproblemcontainer.h"
 #include "forward_euler_problem.h"
 #include "backward_euler_problem.h"
-#include "crank_nicolson_problem.h"
-#include "shifted_crank_nicolson_problem.h"
 #include "functionalinterface.h"
 #include "pdeinterface.h"
 #include "instatreducedproblem.h"
 #include "instat_step_newtonsolver.h"
-#include "newtonsolvermixeddims.h"
+#include "newtonsolver.h"
 #include "gmreslinearsolver.h"
 #include "cglinearsolver.h"
 #include "directlinearsolver.h"
@@ -78,6 +76,7 @@ typedef OptProblemContainer<FUNC,FUNC,PDE,DD,CONS,SPARSITYPATTERN, VECTOR, LOCAL
 //FIXME: This should be a reasonable dual timestepping scheme
 #define DTSP BackwardEulerProblem
 
+
 typedef InstatOptProblemContainer<TSP,DTSP,FUNC,FUNC,PDE,DD,CONS,SPARSITYPATTERN, VECTOR, LOCALDOPEDIM,LOCALDEALDIM> OP;
 
 #undef TSP
@@ -90,15 +89,14 @@ typedef Integrator<IDC , VECTOR , double, LOCALDEALDIM> INTEGRATOR;
 typedef DirectLinearSolverWithMatrix<SPARSITYPATTERN, MATRIX , VECTOR,
     LOCALDEALDIM> LINEARSOLVER;
 
+typedef NewtonSolver<INTEGRATOR, LINEARSOLVER, VECTOR , LOCALDEALDIM>
+    CNLS;
 typedef InstatStepNewtonSolver<INTEGRATOR, LINEARSOLVER, VECTOR , LOCALDEALDIM>
     NLS;
 
-typedef FractionalStepThetaStepNewtonSolver<INTEGRATOR, LINEARSOLVER, VECTOR , LOCALDEALDIM>
-    NLS2;
-
 typedef ReducedNewtonAlgorithm<OP, VECTOR, LOCALDOPEDIM, LOCALDEALDIM> RNA;
 
-typedef InstatReducedProblem<NLS, NLS, INTEGRATOR, INTEGRATOR, OP,
+typedef InstatReducedProblem<CNLS, NLS, INTEGRATOR, INTEGRATOR, OP,
     VECTOR, LOCALDOPEDIM, LOCALDEALDIM> SSolver;
[…]

### 2012-05-24 Paper_May_2012/main.tex
Modifications in the paper

@@
-\documentclass[twoside, 11pt]{article}
-\usepackage{jnm}
+\documentclass[prodmode,acmtoms]{acmsmall}
 %
-\author{Christian\,Goll\footnotemark[1]\comma Thomas\,Wick\footnotemark[1] 
-\comma Winnifried\,Wollner\footnotemark[1]\thanks{%
-Institut f\"ur Angewandte Mathematik, Universit\"at Heidelberg, Im Neuenheimer Feld 293/294,
-			D-69120 Heidelberg, Germany.}
-}
 
-\sauthor{C. Goll, T. Wick and  W. Wollner}
+% Package to generate and customize Algorithm as per ACM style
+\usepackage[ruled]{algorithm2e}
+\renewcommand{\algorithmcfname}{ALGORITHM}
+\SetAlFnt{\small}
+\SetAlCapFnt{\small}
+\SetAlCapNameFnt{\small}
+\SetAlCapHSkip{0pt}
+\IncMargin{-\parindent}
 
-\title{DOpE: A Goal Oriented Software Library for Computing PDEs and Optimization Problems}
-\stitle{DOpE: A Goal Oriented Software Library}
-\received{Received \hfill}
+% Metadata Information
+\acmVolume{0}
+\acmNumber{0}
+\acmArticle{00}
+\acmYear{0000}
+\acmMonth{0}
+
 %
@@
 \numberwithin{equation}{section}
-\usepackage[noend]{algorithmic}                        %Algorithm typesetting environment
-\usepackage[ruled]{algorithm}
 \usepackage{braket}                             %Provides \Set for typing sets
-\usepackage{color}
-\usepackage[square,sort&compress,numbers]{natbib}
 \usepackage{float}
@@
 
+%
+\begin{document}
+\markboth{C. Goll, T. Wick, and W. Wollner}{DOpElib: A Goal Oriented Software Library}
+
+\title{DOpElib: A Goal Oriented Software Library for Computing PDEs and Optimization Problems}
+
+\author{CHRISTIAN GOLL
+\affil{University of Heidelberg}
+THOMAS WICK
+\affil{University of Heidelberg}
+WINNIFRIED WOLLNER
+\affil{University of Hamburg}}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\abstract{%
+\begin{abstract}
 In this article, we describe the software library 
-package \textit{Deal Optimization Environment} (DOpE).
-Its main feature is to give a unified interface to high level algorithms such as 
-time stepping methods, nonlinear solvers and optimization routines. The performance and features 
-of the software package are demonstrated by computing 
-four numerical tests.
+\textit{Deal.II Optimization Environment} (DOpElib).
+The main feature of DOpElib is that it provides a unified interface to high level algorithms 
+such as time stepping methods, nonlinear solvers and optimization routines. This structure ensures 
+that first of all the user is required to write those sections of code that are specific to 
+the considered problem. Second, the exchange of parts of the used routines is possible 
+with the need for just a few lines of code to change.
+The article illustrates performance and features 
+of the software package by four numerical tests. 
+\end{abstract}
 
-}
+%\category{C.2.2}{Computer-Communication Networks}{Network Protocols}
 
-\keywords{%
+%\terms{Design, Algorithms, Performance}
 
-}
+%\keywords{Wireless sensor networks, media access control,
+%multi-channel, radio interference, time synchronization}
 
-%
-\begin{document}
+%\acmformat{Zhou, G., Wu, Y., Yan, T., He, T., Huang, C., Stankovic,
+%J. A., and Abdelzaher, T. F.  2010. A multifrequency MAC specially
+%designed for  wireless sensor network applications.}
+
+\begin{bottomstuff}
+Author's adresses: C. Goll {and} T. Wick, Institut f\"ur Angewandte Mathematik, Universit\"at Heidelberg;
+W. Wollner, Fachbereich Mathematik, Universit\"at Hamburg.
+\end{bottomstuff}
+
 \maketitle
@@
 \label{introduction}
-The \textit{Deal Optimization Environment} (DOpE) project is 
-based on the \textit{deal.II} \citep{deal} finite element library which has been developed
+
+The \textit{Deal Optimization Environment} (DOpElib) project is 
+based on the \textit{deal.II} \cite{deal} finite element library which has been developed
  initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{deal}.
@@
 %\bibliographystyle{abbrvnat}
-\bibliographystyle{plainnat}
+\bibliographystyle{acmsmall}
 \bibliography{lit}

### 2012-06-02 Paper_May_2012/main.tex
Ausbau des Artikels

@@
 \renewcommand{\epsilon}{\varepsilon}
+
+\newcommand{\todo}[1]{\textbf{\textsc{\textcolor{black}{TODO: #1}}}}
 
@@
 
-The \textit{Deal Optimization Environment} (DOpElib) project is 
+\todo{Wie sehr wollen wir deal.II im ersten Satz hervorheben?}
+\todo{Vorschlag von TW: 
+The aim of the \textit{Deal Optimization Environment} (DOpElib) 
+is to provide a software toolkit to solve forward PDE
+problems as well as optimal control problems constrained by PDE. The
+solution of a broad variety of PDE is possible in other software
+libraries (like in deal.II, dune, or commercial solvers Ansys) 
+as well, but
+DOpE concentrates on a unified approach for both linear and nonlinear
+problems by interpreting every PDE problem as nonlinear and applying a
+Newton method to solve it. While deal.II leaves much of the work and many
+decisions to the user, DOpE intends to be user-optimized by delivering
+prefabricated tools which require from the user only adjustments connected
+to his specific problem. The solution of optimal control problems with PDE
+constraints is an innovation in the DOpE framework.
+The focus is on the numerical solution of both stationary and nonstationary
+problems which come from different application fields, like elasticity and
+plasticity, fluid dynamics, and fluid-structure interactions.
+}
+
+The DOpElib project is 
 based on the \textit{deal.II} \cite{deal} finite element library which has been developed
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The aim of DOpE is to provide a software toolkit to solve forward PDE
-problems as well as optimal control problems constrained by PDE. The
-solution of a broad variety of PDE is possible in deal.II as well, but
-DOpE concentrates on a unified approach for both linear and nonlinear
-problems by interpreting every PDE problem as nonlinear and applying a
-Newton method to solve it. While deal.II leaves much of the work and many
-decisions to the user, DOpE intends to be user-optimized by delivering
-prefabricated tools which require from the user only adjustments connected
-to his specific problem. The solution of optimal control problems with PDE
-constraints is an innovation in the DOpE framework.
-The focus is on the numerical solution of both stationary and nonstationary
-problems which come from different application fields, like elasticity and
-plasticity, fluid dynamics, and fluid-structure interactions.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%The aim of DOpE is to provide a software toolkit to solve forward PDE
+%problems as well as optimal control problems constrained by PDE. The
+%solution of a broad variety of PDE is possible in deal.II as well, but
+%DOpE concentrates on a unified approach for both linear and nonlinear
+%problems by interpreting every PDE problem as nonlinear and applying a
+%Newton method to solve it. While deal.II leaves much of the work and many
+%decisions to the user, DOpE intends to be user-optimized by delivering
+%prefabricated tools which require from the user only adjustments connected
+%to his specific problem. The solution of optimal control problems with PDE
+%constraints is an innovation in the DOpE framework.
+%The focus is on the numerical solution of both stationary and nonstationary
+%problems which come from different application fields, like elasticity and
+%plasticity, fluid dynamics, and fluid-structure interactions.
+%
 
-At the present stage the following features are supported by the library
+At the present stage the following key features are supported by the library
 \begin{itemize}
@@
 
-The rest of this document is structured as follows: We start with an introduction in
-Chapter~\ref{chap:intro} where you will learn what is needed to run {\tt DOpElib}. 
-Further you will learn what problems we can solve and how all the different classes 
-work together for this purpose. This should help you figure out what the different classes
-do if you are in need of writing your own algorithm.
+The article is organized as follows. In Section
+\ref{detailed_description}
 
-Then assuming that you can work to your satisfaction with the algorithms already implemented
-we will show you how to create your own running example in Chapter~\ref{chap:howtoex}.
-This will be followed by a detailed description of all examples already shipped with 
-the library. You can find the examples for the solution of PDEs in Chapter~\ref{PDE}
-and those for the solution of optimization problems with PDEs in Chapter~\ref{OPT}.
-
-These notes conclude with a section that explains how we do automated testing of the 
-implementation in Chapter~\ref{chap:test}. This chapter will be of interest only if you 
-are trying to implement some new features to the library so that you can check that 
-the new code did not break anything.
+%The rest of this document is structured as follows: We start with an introduction in
+%Chapter~\ref{chap:intro} where you will learn what is needed to run {\tt DOpElib}. 
+%Further you will learn what problems we can solve and how all the different classes 
+%work together for this purpose. This should help you figure out what the different classes
+%do if you are in need of writing your own algorithm.
+%
+%Then assuming that you can work to your satisfaction with the algorithms already implemented
+%we will show you how to create your own running example in Chapter~\ref{chap:howtoex}.
+%This will be followed by a detailed description of all examples already shipped with 
+%the library. You can find the examples for the solution of PDEs in Chapter~\ref{PDE}
+%and those for the solution of optimization problems with PDEs in Chapter~\ref{OPT}.
+%
+%These notes conclude with a section that explains how we do automated testing of the 
+%implementation in Chapter~\ref{chap:test}. This chapter will be of interest only if you 
+%are trying to implement some new features to the library so that you can check that 
+%the new code did not break anything.
+%
 
@@
 \section{Detailed Description of the Features}
+\label{detailed_description}
 This library is designed to allow easy implementation and numerical solutions 
@@
 
+\section{Documentation, Code Development, Webpage}
+\label{documentation}
+At the present step, the DOpElib project comes with 
+an detailed documentation of all features and examples 
+in pdf format and an detailed programming code 
+documentation.
+
+\todo{License? How to access to DOpE?}
+\todo{Name of the website}
+
+In addition, the DOpElib test suite provides some regression tests. 
+They are run to compare 
+the output to previous outputs. This is useful (necessary) after 
+changing programming code anywhere in the library. If a test
+succeeds, everything is fine in the library. If not, you should not
+check in your code into DOpE. Please make sure what is going wrong and WHY!
+Every command is computed via a Makefile.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Numerical Examples}
+\section{Applications}
+\label{applications}
+We present some numerical tests that have been already computed 
+with DOpElib. These tests demonstrate the performance 
+of DOpElib by covering different computational aspects. 
+\begin{itemize}
+\item In detail, the first numerical test, validates various time-stepping schemes
+and the numerical solution of nonlinear coupled partial differential
+equations, formulated in a monolithic solution algorithm. 
+Specifically, the equations are solved by the Rothe method, i.e,
+first temporal discretization and then spatial discretization 
+with a Galerkin finite element scheme. The nonlinaer problem
+is solved by Newton's method. The Jacobian is derived 
+by analytical expressions of the derivatives. 
+In DOpElib this example is implemented in 
+\texttt{Example/PDE/InstatPDE/Example2}
+\item The second example presents an
+optimization problem and its numerical solution.
+\item In the third example, we consider the numerical
+solution of fluid flows governed by the Navier-Stokes
+equations. In contrast to the previous tests, we 
+use finite elements on locally adapted meshes 
+with goal-oriented mesh refinement with the help
+of the dual-weighted residual method. Using 
+this method of local mesh refinement, the error estimator
+is derived with respect to some goal functional and 
+provides optimal meshes to the accurate measurement 
+of the considered functional value. Consequently,
+this is an effiecient method to reduce the computational 
+cost. Most other 
+software libraries does not provide goal-oriented 
+mesh refinement.  
+\item In the last numerical test, we consider 
+different types of boundary conditions to drive
+our problems. 
+\end{itemize}
 
 \subsection{Nonstationary Fluid-Structure Interaction}
-The results are already obtained
+In this example, the three proposed mesh motion models are applied 
+to an unsteady fluid-structure interaction problem. We consider 
+the numerical benchmark test FSI 2, which was proposed in \cite{HrTu06b}. 
+The configuration is 
+sketched in Figure \ref{configuration_csm_and_fsi_2D}.
+New results can be found in \cite{BuSc06, TuHrMaRaWoAc10, DeHaeAnnBrVie10}. 
+The Fractional-Step-$\theta$ scheme was used for time discretization with
+different time step sizes $k$.
+
+% Bilder von FSI 2 
+\begin{figure}[h]
+\centering
+{\includegraphics[width=6cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_mesh8070_scale.png}}
+{\includegraphics[width=6cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_x_velo8070_scale.png}}
+\caption{FSI 2 test case: mesh (left) and velocity profile in vertical 
+direction (right) at time $t=16.14s$.}
+\label{res:fsi_2_mesh_and_x_velo}
+\end{figure}
+
+\begin{figure}
+\centering
+{\includegraphics[width=5.8cm]{Pictures/ux_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
+{\includegraphics[width=5.8cm]{Pictures/uy_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
+{\includegraphics[width=5.8cm]{Pictures/Drag_fluid_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
+{\includegraphics[width=5.8cm]{Pictures/Lift_fluid_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
+\caption{FSI 2: the deflections of the beam, $u_x(A)$ and $u_y(A)$ (in $cm$), and 
+the drag and the lift evaluation (in $kg/m\,s^2$) are displayed versus
+time (in $s$).
+} 
+\label{res:results_ux_and_uy_fsi_2}
+\end{figure}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions and Outlook}
+\label{conclusions}
+In this article, we described the features 
+and applications of the DOpElib project, which 
+was initiated at the Heidelberg University in 2010.
+Specifically, the applications cover a broad 
+spectrum of numerical exmamples motivated 
+by different disciplines. 
+Currently, we are developing concepts for the efficient 
+numerical solution of time-dependent optimization
+problems governed by PDEs. 
+\todo{Was wollen wir noch in den nächsten 1-2 Jahren erreichen?}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Acknowledgment}
+The second author thanks Rolf Rannacher from
+the Institute for Applied mathematics at the 
+Heidelberg University for financial suppport.
 

### 2012-06-09 Paper_May_2012/main.tex
Modifikationen am Artikel: Kommentare und Ideen von der Diskussion mit Christian eingefuegt.

@@
 \label{introduction}
+The \textit{Deal Optimization Environment} (DOpElib) 
+provides a software toolkit to solve forward PDE
+problems as well as optimal control problems constrained by a PDE. 
+Its main feature is to give a unified interface to high level algorithms such as 
+time stepping methods, nonlinear solvers and optimization routines. 
+We aim that the user should only need to write those parts
+of the code that are problem dependent while all invariant parts of the algorithms
+should be reusable without any need for further coding.
+In particular, the user should be able to switch between various different 
+algorithms without the need to rewrite the problem dependent code, though he or she will
+have to replace the algorithm object with an other one.
 
-\todo{Wie sehr wollen wir deal.II im ersten Satz hervorheben?}
-\todo{Vorschlag von TW: 
-The aim of the \textit{Deal Optimization Environment} (DOpElib) 
-is to provide a software toolkit to solve forward PDE
-problems as well as optimal control problems constrained by PDE. The
+The
 solution of a broad variety of PDE is possible in other software
-libraries (like in deal.II, dune, or commercial solvers Ansys) 
+libraries (like in deal.II \cite{deal}, dune \cite{dune} , 
+or commercial solvers like Ansys \cite{ansys}) 
 as well, but
@@
 plasticity, fluid dynamics, and fluid-structure interactions.
-}
 
@@
  initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{deal}.
-Its main feature is to give a unified interface to high level algorithms such as 
-time stepping methods, nonlinear solvers and optimization routines. 
-We aim that the user should only need to write those parts
-of the code that are problem dependent while all invariant parts of the algorithms
-should be reusable without any need for further coding.
-In particular, the user should be able to switch between various different 
-algorithms without the need to rewrite the problem dependent code, though he or she will
-have to replace the algorithm object with an other one.
-
 The authors acknowledge their past experience as well as discussions with 
@@
   involving stationary PDE constraints.
-\item Mesh adaptation.
+\item Goal-oriented mesh adaptation with the dual-weighted residual method.
 \item Different spatial triangulations for control and state variables.
@@
 \label{detailed_description}
+\todo{Aus der Sicht des Users anhand eines konkreten Beispiels beschreiben}
+\todo{Einfaches Beispiel nehmen}
+\todo{erwaehnen, dass wir alles immer nicht-linear betrachten}
+\todo{Aussehen eines typischen PDE und Opt loop und deren Gemeinsamkeiten
+erklaeren. Vielleicht mir Pseudocode}
 This library is designed to allow easy implementation and numerical solutions 
@@
   \item \texttt{optproblem.h} Is used to describe  OPT problems governed by 
-    stationary PDEs. {\bf TODO should be renamed to optproblemcontainer.h including the class name!}
-  \item \texttt{instatoptproblemcontainer.h} Is used to describe  OPT problems governed by nonstationary PDEs.
-The only difference to the stationary case is that we need to specify a time-stepping method.
+    stationary PDEs. 
+  \item \texttt{instatoptproblemcontainer.h} Is used to describe  OPT problems
+    governed by nonstationary PDEs. The only difference to the stationary case
+    is that we need to specify a time-stepping method.  
+  \item \texttt{interfaces/functionalinterface.h} This gives an interface 
+    for the functional $J(\cdot)$ and any other functional you may want to evaluate.
+    In general this can be used as a base class to write your own functionals 
+    in examples. We note that we only need to write the integrands on 
+    elements or faces the loop over elements will be taken care of in the integrator.
+    Specifically, derivatives are written therein, too.
 \end{itemize}
@@
 \subsubsection{Time stepping schemes}
+\todo{Kurz halten}
 Third, at least for nonstationary PDEs we need to select a time stepping scheme
@@
 \subsubsection{Linear solvers}
-\begin{itemize}
-\item \texttt{templates/cglinearsolver.h} This is a wrapper for the cg solver implemented in 
-  \texttt{deal.II}. The solver will build and store the stiffness matrix for the PDE.
-\item \texttt{templates/gmreslinearsolver.h} This is a wrapper for the gmres solver 
-  implemented in \texttt{deal.II}. The solver will build and store the stiffness matrix 
-  for the PDE.
-\item \texttt{templates/directlinearsolver.h} This is a wrapper for the direct solver 
-  implemented in \texttt{deal.II} using \texttt{UMFPACK}. 
-  The solver will build and store the stiffness matrix for the PDE.
-\item \texttt{templates/voidlinearsolver.h} This is a wrapper for certain cases when we 
+To solve the linear equations within Newton's method, we use 
+standard solvers from iterative type 
+such as the method of the conjugate gradients (CG method), 
+GMRES, or direct solvers such as \texttt{UMFPACK}.
+In addition, we DOpE provides  a wrapper class 
+\texttt{templates/voidlinearsolver.h} for certain cases when we 
   know that the matrix to be inverted is the identity. It simply copies the rhs to the
   lhs. This is only needed for compatibility reasons some other components.
-\end{itemize}
 
-\subsection{Problem specific classes}
-The following classes are used to describe the problem and will usually require 
-some implementation.
+%\begin{itemize}
+%\item \texttt{templates/cglinearsolver.h} This is a wrapper for the cg solver implemented in 
+%  \texttt{deal.II}. The solver will build and store the stiffness matrix for the PDE.
+%\item \texttt{templates/gmreslinearsolver.h} This is a wrapper for the gmres solver 
+%  implemented in \texttt{deal.II}. The solver will build and store the stiffness matrix 
+%  for the PDE.
+%\item \texttt{templates/directlinearsolver.h} This is a wrapper for the direct solver 
+%  implemented in \texttt{deal.II} using \texttt{UMFPACK}. 
+%  The solver will build and store the stiffness matrix for the PDE.
+%\item \texttt{templates/voidlinearsolver.h} This is a wrapper for certain cases when we 
+%  know that the matrix to be inverted is the identity. It simply copies the rhs to the
+%  lhs. This is only needed for compatibility reasons some other components.
+%\end{itemize}
+%
 
-\begin{itemize}
-  \item \texttt{basic/constraints.h} This is used by the spacetimehandlers to 
-    compute the number of constraints from the control and state vectors. 
-    It must not be reimplemented by the user, but needs to be properly 
-    initialized if OPT is used with box control constraints or $g(q,u) \le 0$.
-  \item \texttt{interfaces/functionalinterface.h} This gives an interface 
-    for the functional $J(\cdot)$ and any other functional you may want to evaluate.
-    In general this can be used as a base class to write your own functionals 
-    in examples. We note that we only need to write the integrands on 
-    elements or faces the loop over elements will be taken care of in the integrator.
-    Specifically, derivatives are written therein, too.
-  \item \texttt{interfaces/constraintinterface.h} This gives an interface for both 
-    the control box constraints as well as the general constraint $g \le 0$. This 
-    needs to be specified if constraints are to be used. If they are not needed 
-    a default class \texttt{problemdata/noconstraints.h} can be used. We note that we only 
-    need to write the integrands on 
-    elements or faces the loop over elements will be taken care of in the integrator.
-  \item \texttt{interfaces/pdeinterface.h} This defines an interface for the 
-    partial differential equation $a(q,u)(\phi) = 0$. This needs to be written
-    by the user. We note that we only need to write the integrands on 
-    elements or faces the loop over elements will be taken care of in the integrator.
-    Specifically, derivatives are written therein, too.
-  \item \texttt{interfaces/dirichletdatainterface.h} This gives an interface to the 
-    Dirichlet data for a problem. If the Dirichlet data are simply a function 
-    (and do not depend on the control $q$) one can use the default class\\
-    \texttt{problemdata/simpledirichletdata.h}.
-\end{itemize}
+%\subsection{Problem specific classes}
+%The following classes are used to describe the problem and will usually require 
+%some implementation.
+%
+%\begin{itemize}
+%  \item \texttt{basic/constraints.h} This is used by the spacetimehandlers to 
+%    compute the number of constraints from the control and state vectors. 
+%    It must not be reimplemented by the user, but needs to be properly 
+%    initialized if OPT is used with box control constraints or $g(q,u) \le 0$.
+%
+%  \item \texttt{interfaces/constraintinterface.h} This gives an interface for both 
+%    the control box constraints as well as the general constraint $g \le 0$. This 
+%    needs to be specified if constraints are to be used. If they are not needed 
+%    a default class \texttt{problemdata/noconstraints.h} can be used. We note that we only 
+%    need to write the integrands on 
+%    elements or faces the loop over elements will be taken care of in the integrator.
+%  \item \texttt{interfaces/pdeinterface.h} This defines an interface for the 
+%    partial differential equation $a(q,u)(\phi) = 0$. This needs to be written
+%    by the user. We note that we only need to write the integrands on 
+%    elements or faces the loop over elements will be taken care of in the integrator.
+%    Specifically, derivatives are written therein, too.
+%  \item \texttt{interfaces/dirichletdatainterface.h} This gives an interface to the 
+%    Dirichlet data for a problem. If the Dirichlet data are simply a function 
+%    (and do not depend on the control $q$) one can use the default class\\
+%    \texttt{problemdata/simpledirichletdata.h}.
+%\end{itemize}
+%
 
@@
 
-\subsection{Other Components}
-Beyond these clearly structured groups before there are some classes remaining that
-do not fit the above but are important for the user to know.
-
-\subsubsection{Vectors}
-\begin{itemize}
-\item \texttt{include/statevector.h} This stores all dofs in space and time for the state 
-  variable $u$. It is possible to select whether all this should be kept in memory or 
-  or unused parts can be written to the hard disk.
-\item \texttt{include/controlvector.h} This stores all dofs in space and time for the 
-  control variable $q$. At present no time dependence is implemented.
-\item \texttt{include/constraintvector.h} This stores all dofs in space and time for the 
-  non PDE constraints (and corresponding multipliers). 
-  At present no time dependence is implemented.
-\end{itemize}
-
-\subsubsection{Parameter handling}
-\begin{itemize}
-\item \texttt{include/parameterreader.h} This file is used to define a parameter reader
-  that is used to read run time parameters from a given file.
-\end{itemize}
-
-\subsubsection{Exception handling}
-\begin{itemize}
-\item \texttt{include/dopeexception.h} Defines some Exceptions that are thrown by the program
-  should it encounter any unexpected errors.
-\item \texttt{include/dopeexceptionhandler.h} This class is used to write information 
-  contained in the exceptions to the output in a uniform manner.
-\end{itemize}
-
-\subsubsection{Output handling}
-\begin{itemize}
-\item \texttt{include/outputhandler.h} This file defines an outputhandler object which 
-  can be used to decide whether some information should be written to screen or file.
-  In addition it can format output according to some run time parameters given by a 
-  parameter file.
-\end{itemize}
-
-\subsection{Data Access}
-\begin{itemize}
-\item \texttt{include/solutionextractor.h} This class is used to gain access to the finite element 
-  solutions stored in the reduced problems.
-\end{itemize}
-
-\subsubsection{Constraints and system matrix}
-\begin{itemize}
-\item \texttt{include/userdefineddofconstraints.h} This class sets the constraints on the DOFs of the state and/or control FE solution.
-DOpE itself builds the hanging-node-constraints, but the user can reimplement this class and thus include other constraints as well (for example periodic BC).
-Note, that the hanging-node-constraints come first (in case of conflicting constraints.)
-\item \texttt{include/sparsitymaker.h} This class sets the sparsity pattern for the state FE solution.
-The standard implementation is just a wrapper for \texttt{dealii::DoFTools::} \texttt{make\_sparsity\_pattern}, but the user can reimplement this class to allow for more sophisticated sparsity patterns.
-\end{itemize}
-
-\subsubsection{HP components}
-\begin{itemize}
-\item \texttt{interfaces/active\underline{ }fe\underline{ }index\underline{ }setter\underline{ }interface.h} In the case of hp finite elements, one has to specify for each cell which finite element to use.
-This is done via this interface.
-\end{itemize}
-
-\subsection{Internal structures}
-{\bf TODO: ab hier habe ich noch nichts gemacht (TW)}
-\subsubsection{Interface Classes}
-\begin{itemize}
-  \item \texttt{interfaces/transposeddirichletdatainterface.h} This provides an interface to 
-    the functionality required by {\em transposed Dirichlet data}. Usually when one applies Dirichlet 
-    data $g$ to a function one has to calculate a continuation $Bg$ which is defined on the whole domain.
-    In optimization problems when the Dirichlet data depends on the control one has to evaluate the 
-    dual operator $B^*$ in order to obtain a representation for the reduced gradient of the objective $J$.
-    This is done using the {\em transposed Dirichlet data}.
-  \item \texttt{interfaces/reducedprobleminterface.h} In order to allow all algorithms to be written independent
-    of the given (OPT) problem 
-    (and not requiring the problem as template argument) there is a common base class which 
-    defines the required interfaces. 
-  \item \texttt{interfaces/pdeprobleminterface.h} The same as above but for (PDE) problems.
-\end{itemize}
-
-\subsubsection{Default Classes}
-\begin{itemize}
-  \item \texttt{problemdata/noconstraints.h} A class that can be used for optimization problems 
-    having only a PDE constraint but no further constraints.
-  \item \texttt{problemdata/simpledirichletdata.h} A class that can be used to implement Dirichlet
-    data that are given as a fixed function (independent of the control).
-  \end{itemize}
-
-\subsubsection{Auto-generated Problem Descriptions}
-\begin{itemize}
-  \item \texttt{problemdata/stateproblem.h} This is the problem description for the (forward/primal) PDE constraint.
-    {\bf Similar descriptors will be build for the other problems (adjoint, tangent, $\ldots$) when time allows.}
-  \item \texttt{problemdata/initialproblem.h} This is the problem descriptor to compute the finite element representation
-    of the initial values. This is generated by the different time-stepping schemes based upon the defined 
-    representation by the pde, which is set to the component wise $L^2$ projection by default.
-  \item \texttt{problemdata/primaldirichletdata.h} This class contains the Dirichlet data for the 
-    forward/primal PDE.
-  \item \texttt{problemdata/tangentdirichletdata.h} This class contains the Dirichlet data for the tangent PDE, i.e.,
-    the first derivative of the Dirichlet data.
-  \item \texttt{problemdata/transposedgradientdirichletdata.h} This contains the transposed Dirichlet data needed 
-    to calculate the gradient of the reduced objective functional, 
-    for detail see \texttt{interfaces/transposeddirichletdatainterface.h}.
-  \item \texttt{problemdata/transposedhessiandirichletdata.h} This contains the transposed Dirichlet data needed 
-    to calculate the hessian of the reduced objective functional, 
-    for detail see \texttt{interfaces/transposeddirichletdatainterface.h}.
-\end{itemize}
-\subsubsection{Management of Time Dependent Problems}
-\begin{itemize}
-\item \texttt{include/timedofhandler.h} DoFHandler responsible for the management of the timedofs (this is a part of the \texttt{SpaceTimeDoFHandler}-classes).
-Basically a wrapper for a $1d$ \texttt{deal.II}-DoFHandler.
-\item \texttt{include/timeiterator.h} This class works as an iterator on the \texttt{TimeDoFHandler}.                 
-\end{itemize}
-\subsection{Wrapper classes}
-\begin{itemize}
-  \item \texttt{wrapper/dofhandler\underline{ }wrapper.h} A wrapper class for the \texttt{deal.II} DoFHandlers. This 
-    class is needed to provide support for the \texttt{dim $= 0$} case and to have a uniform interface to 
-    DoFHandler and HPDoFHandler.
-  \item \texttt{wrapper/preconditioner\underline{ }wrapper.h} Contains wrappers for several of the preconditioners
-    in \texttt{deal.II}. This is required since unfortunately the preconditioners in \texttt{deal.II} have different
-    interfaces for their initialization.
-  \item \texttt{wrapper/function\underline{ }wrapper.h} An interface that allows to use functions that depend 
-    not only on space but also on time.
-  \item \texttt{wrapper/snopt\underline{ }wrapper.h} An interface to the SNOPT fortran library. This is an additional
-    wrapper to the one provided by SNOPT to allow automatic construction of the functions required 
-    by SNOPT using our library.
-  \item \texttt{wrapper/fevalues\underline{ }wrapper.h} {\bf Will be removed soon!}
-\end{itemize}
-
-\subsection{Other}
-\begin{itemize}
-  \item \texttt{basic/dopetypes.h} This file contains type definitions used in the library. 
-\end{itemize}
-
-{\bf TODO! The following files need some appropriate description here 
-This should be done by someone who knows what they are used for. Please move the description to the right place above.} tsschemes.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \subsection{Interfaces to other software packages}
-deal.II and SNOPT
+Since DOpE uses as basis the software library deal.II, all its interfaces
+to other packages can be accessed to. For instance, such libararies 
+are trilinos for with an algebraic multigrid method, MPI for a parallel
+solution of the problem.
+
+In addition DOpE itself has an interface to SNOPT.
 
@@
 \begin{itemize}
-\item In detail, the first numerical test, validates various time-stepping schemes
+\item The first numerical test shows various time-stepping schemes
 and the numerical solution of nonlinear coupled partial differential
 equations, formulated in a monolithic solution algorithm. 
-Specifically, the equations are solved by the Rothe method, i.e,
-first temporal discretization and then spatial discretization 
+Specifically, the equations are solved by the method of lines, i.e,
+first spatial discretization and then temporal  discretization 
 with a Galerkin finite element scheme. The nonlinaer problem
@@
 optimization problem and its numerical solution.
-\item In the third example, we consider the numerical
+\item The main purpose of the third example is the application 
+of refinement techniques for local mesh refinement for 
+two and three dimensional problems. We consider the numerical
 solution of fluid flows governed by the Navier-Stokes
@@
 mesh refinement.  
-\item In the last numerical test, we consider 
-different types of boundary conditions to drive
-our problems. 
+\item Here, we show a time-dependent optimization problem or 
+FSI optimization. 
 \end{itemize}
@@
 \subsection{Goal-oriented mesh refinement for Navier-Stokes}
-Has to be computed.
+Has to be computed in two and three dimensions.
 
-\subsection{Dirichlet Velocity, Neumann Pressure, and
-Periodic Boundary Conditions}
-I do not know.
+\subsection{Time-dependent Opimization or FSI optimization}
 

### 2012-06-19 Examples/Documentation/introduction.tex
Documentation updated
Updated the manual to include information on the installation on Mac OSX

@@
 
+\subsection{g++}
 \texttt{DOpElib} requires a recent \texttt{g++} (at least 4.4.3) due to 
 some new \texttt{C++} features implemented in \texttt{C++0x} and \texttt{C++11}. 
+You can check the version number using the command-line argument \texttt{g++ -v}.
+
+Under Linux-systems this typically means that you have to do nothing if you have a recent 
+version number. Otherwise you can either install the required version of g++ using the 
+appropriate software installation tool, or you can build the required version from 
+source \texttt{gcc.gnu.org}.
+
+Under MAC OSX, you need to install the \texttt{XCode} tools delivered with the operating system 
+or availiable for free \texttt{developer.apple.com/xcode}. Unfortunately, the delivered 
+version of \texttt{g++} is too old, so you need to install the real thing. To do so, 
+download \texttt{MacPorts} from \texttt{macports.org}. Once you have installed MacPorts 
+you can use it to install aditional linux software. 
+First, update the MacPorts installation \texttt{sudo port selfupdate}
+after that you can install a new version of \texttt{g++} using forinstance 
+\texttt{sudo port install gcc45} to install version $4.5$ of the compiler.
+Afterwards, you need to set the search path appropriate to find the macports version
+of \texttt{g++}, to check if this has been done use \texttt{g++ -v}.
 
@@
 \item \texttt{reducedproblems/instatreducedproblem.h} The same as above but for a
-  nonstationary PDE. {\bf FIXME there is something wrong in this file see FIXME 
-    comment in the source.}
+  nonstationary PDE.
 \item \texttt{reducedproblems/voidreducedproblem.h} A wrapper file that eliminates $u$ 

### 2012-06-19 Examples/PDE/StatPDE/Example9/content.tex
Modifications in the program to allow evaluation of residual type error estimators, like the one for L2 and H1 error. To allow that this can be done in the same framework as the dwr-estimator there is a new method ResidualModifier in the PDEs. This should always be applied to the terms residual and jump terms in Strong...Residual. It is modified by the error estimators to switch between the calculation of the integral over the residual and the norm of the residual. PDE/StatPDE/Example9 has been modified to calculate these new values in addition to the dwr estimates.

@@
 
+Note that in the above three functions we always apply the method \texttt{ResidualModifier} both to the 
+residual as well as to the jumps on the faces. This is done to assert that we can apply both a 
+DWR-error estimator where the residual should be multiplied with the computed weights (then this function 
+does not do anything) as well as Residual Type error estimator for the $L^2$ or $H^1$ norm where we need to calculate element wise norms of the residual and the jumps.
+Then this function calculates the appropriate
+local terms, e.g., the square of the residual scaled with appropriate powers of the local mesh size.
+
 After this, we tell the problem which functional we want to use for the error estimation, this is done via

### 2012-06-27 Examples/OPT/StatPDE/Example9/content.tex
Beispiel Optimierung mit FSI eingecheckt

In this example we solve an optimization problem where 
the equations come from fluid-structure interaction (FSI). The idea
is to extend the steady FSI benchmark problem (FSI 1) to an optimization 
problem where the drag is minimized over the cylinder plus beam.

The example has many similarities to Opt Example 3.

Detailed description will follow later.

### 2012-07-11 Examples/Documentation/introduction.tex
Removed Makefile.dep from svn

@@
 Chapter~\ref{chap:test}.
+
+\section{FAQs}
+\paragraph{1.) When building the library I get an error message:}
+\begin{itemize}
+\item {\bf unrecognized command line option "-std=gnu++0x"}\\
+  This means that your compiler is too old. You can check the 
+  version of your compiler using \texttt{g++ -v}. If the version is lower than
+  \texttt{4.5} you need to get a newer compiler version.
+\end{itemize}
+
+\paragraph{2.) I have installed a new g++ compiler but g++ -v still finds the old one :}
+\ \\
+  This means that your computer does not find the new compiler. Try
+  \texttt{which g++} to see whether it appears in the list of availiable 
+  compilers (but is maybe too far in the back of the list.) Then you should 
+  modify your \texttt{\$PATH} environment variable so that the new g++ compiler
+  appears.
+
+  If \texttt{which g++} only returns one g++ compiler, then probably you need to
+  set an appropriate symlink.
+
+  For example on Mac OSX you will find only one g++ compiler \texttt{/usr/bin/g++}
+  which is in fact a symlink to \texttt{/usr/bin/g++-4.2}. Thus to change, the 
+  desired behavior you should add a new link named \texttt{g++} in a directory 
+  in front of \texttt{/usr/bin} or replace the symlink in \texttt{/usr/bin}
+  with a link to the newer compiler which may be for instance located (macports 
+  g++ 4.5 on OSX Lion) in \texttt{/opt/local/bin/g++-mp-4.5}. Make shure that 
+  you add additional appropriate links to \texttt{gcc, cc, }and \texttt{c++}.
 

### 2012-07-20 Examples/Documentation/introduction.tex
Added a dof_constraints_maker to have homogeneous Dirichletdata that are enforced on single vertices(=support-points)
Adjusted the pde examples 5 and 9: in example 5 I reduced the precision of the newton output, example 9 was adapted to the new DWR-routine regarding computing the errors on faces with hanging nodes.

@@
 The standard implementation is just a wrapper for \texttt{dealii::DoFTools::} \texttt{make\_sparsity\_pattern}, but the user can reimplement this class to allow for more sophisticated sparsity patterns.
+\item \texttt{include/pointconstraintsmaker.h} This class is derived from \texttt{include/userdefineddofconstraints.h}
+  and allows to set homogeneous dirichlet values at given points/components.
 \end{itemize}
@@
   \item \texttt{basic/dopetypes.h} This file contains type definitions used in the library. 
-\item \item \texttt{basic/sth\underline{ }internals.h} Wrapper for the \texttt{MapDoFsToSupportPoints} function. The implementation of this changes with the \texttt{deal.II} version in use.
+\item \texttt{basic/sth\underline{ }internals.h} Wrapper for the \texttt{MapDoFsToSupportPoints} function. The implementation of this changes with the \texttt{deal.II} version in use.
 \item \texttt{include/helper.h} Collection of various small helper functions.
+\item \texttt{reducedproblems/problemcontainer\underline{ }internal.h} Houses some functions und variables common in the various problemcontainer.
 \item \texttt{tsschemes/primal\_ts\_base.h}  This class contains the methods which all primal time stepping schemes share.

### 2012-08-20 Examples/Documentation/introduction.tex
Corrected Docu for MacOSX

@@
   If \texttt{which g++} only returns one g++ compiler, then probably you need to
-  set an appropriate symlink.
+  set an appropriate symlink. Or more robust, you can configure \texttt{dealii}
+  to use the compiler you intend by configuring deal with the right compiler. To
+  do so adjust  the \texttt{CC} and \texttt{CXX} environment variable appropriately
+  before configuring \texttt{dealii}
 
   For example on Mac OSX you will find only one g++ compiler \texttt{/usr/bin/g++}
-  which is in fact a symlink to \texttt{/usr/bin/g++-4.2}. Thus to change, the 
-  desired behavior you should add a new link named \texttt{g++} in a directory 
-  in front of \texttt{/usr/bin} or replace the symlink in \texttt{/usr/bin}
-  with a link to the newer compiler which may be for instance located (macports 
-  g++ 4.5 on OSX Lion) in \texttt{/opt/local/bin/g++-mp-4.5}. Make shure that 
-  you add additional appropriate links to \texttt{gcc, cc, }and \texttt{c++}.
+  which is in fact a symlink to \texttt{/usr/bin/g++-4.2}. 
+  So that you need to insall a newer compiler this may then be found ,e.g., using macports 
+  g++ 4.5 on OSX Lion, in \texttt{/opt/local/bin/g++-mp-4.5}.
 

### 2012-09-06 Examples/PDE/InstatPDE/Example1/content.tex
Docu of PDE/InstatPDE/Example1 has some flaws.

@@
 \subsubsection{General problem description}
-In this example we consider the nonstationary incompressible Stokes equation. As in the stationary PDE Example 3 \ref{PDE_Stat_FSI_STVK},
+In this example we consider the nonstationary incompressible Navier-Stokes equation. As in the stationary PDE Example 3 \ref{PDE_Stat_FSI_STVK},
 we use the symmetric fluid stress tensor.
+
+\textbf{The rest of this parte is written for Stokes instead of Navier-Stokes. Needs to be adjusted.}
+
 In strong formulation we deal with 
@@
 
-However, solving the Navier-Stokes equations are 
+However, solving the Navier-Stokes equations or 
 multi-physics problems (like fluid-structure

### 2012-10-04 Paper_May_2012/main.tex
Modifikationen am Artikel: Kommentare und Ideen von der Diskussion mit Christian eingefuegt.
Paper: Kapitel 2 mit Beschreibung der main features begonnen. Teil fuer stationaere Probleme fast fertig. Kapitel 3: FSI Example fertig gestellt. D.h. kurze Beschreibung und Darstellung der wesentlichen Ergebnisse.

@@
 \usepackage{multirow}
+\usepackage{listings}
+
 % \usepackage[center]{subfigure}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Detailed Description of the Features}
+\section{Detailed Description of the Main Features}
 \label{detailed_description}
@@
 This library is designed to allow easy implementation and numerical solutions 
-of problems involving partial differential equations (PDEs). The easiest case 
+of problems involving partial differential equations (PDEs). 
+\subsection{Solving of PDEs}
+\subsubsection{Stationary problems}
+The easiest case 
 is that of a PDE in weak form to find some $u$
@@
 \]
-with some appropriate space $V$.
+with some appropriate space $V$. As example, we take 
+the Laplace problem in weak form:
+\[
+a(u)(\phi) = (\nabla u, \nabla \phi) \quad \forall \phi \in V.
+\]
+To solve the problem, we need a configuration (a mesh), eventually 
+a diffusion parameter, a finite element, a quadrature rule,
+a linear solver (here the CG solver), and some output such as
+a graphical solution and eventually some measurement functional.
+Each of those routines are written in the \texttt{main.cc} file. 
+We start by including necessary files both the DOpElib library and
+the deal.II software:
+\begin{lstlisting}
+#include "pdeproblemcontainer.h" etc. ...
+\end{lstlisting}
+In the \texttt{main} function, the following steps need to be done. First, 
+the parameter reader for runtime parameters is initialized:
+\begin{lstlisting}
+  ParameterReader pr;
+  SSolver::declare_params(pr);
+  DOpEOutputHandler<VECTOR>::declare_params(pr);
+  pr.read_parameters(paramfile);
+\end{lstlisting}
+A linear finite element in two dimensions and two components is 
+initialized:
+\begin{lstlisting}
+  FESystem<2> state_fe(FE_Q<2>(1), 2);
+\end{lstlisting}
+and a quadrature rule of order three is used to approximate the 
+integrals:
+\begin{lstlisting}
+  QGauss<2> quadrature_formula(3);
+\end{lstlisting}
+Then, the \texttt{LocalPDE} object comes into play. This is an quite 
+important feature of DOpElib because it allows for easy implementation 
+of the variational formulation. Specific explication is given below because
+the implementation is seperated in \texttt{localpde.h}.
+Afterwards, a \texttt{LocalPointFunctional} is declared. As for the PDE, the 
+implementation is done in another file, namely \texttt{localfunctional.h}.
+
+Generically, each problem in DOpElib is considered to be time-dependent and nonlinear because
+these are the 'real' world problems. Therefore, (even for the linear and stationary Laplace equation),
+we have to declare a (pseudo) time:
+\begin{lstlisting}
+  std::vector<double> times(1, 0.);
+\end{lstlisting}
+Then, the spatial mesh (here the unit square) is created and three-times globally-refined:
+\begin{lstlisting}
+  GridGenerator::hyper_cube(triangulation, 0, 1);
+  triangulation.refine_global(3);
+\end{lstlisting}
+A major component which organizes the whole solution process (independent from 
+the problem!) is the 
+\begin{lstlisting}
+MethodOfLines_StateSpaceTimeHandler<FE, DOFHANDLER, SPARSITYPATTERN, VECTOR, 2> DOFH(
+      triangulation, state_fe);
+\end{lstlisting}
+Then, the problem (here the Laplace PDE) is given to the \texttt{DOFH} via
+\begin{lstlisting}
+  OP P(LPDE, DOFH);
+\end{lstlisting}
+It remains to declare boundary conditions for the problem under consideration. 
+Let's simply take zero Dirichlet conditions on the boundary:
+\begin{lstlisting}
+  std::vector<bool> comp_mask(2);
+
+  comp_mask[0] = true;
+  comp_mask[1] = true;
+
+  DOpEWrapper::ZeroFunction<2> zf(2);
+  SimpleDirichletData<VECTOR, 2> DD1(zf);
+
+  P.SetDirichletBoundaryColors(0, comp_mask, &DD1);
+\end{lstlisting}
+Next, the solver for the linear equations needs to be specified:
+\begin{lstlisting}
+  SSolver solver(&P, "fullmem", pr, idc);
+\end{lstlisting}
+Finally, the ouput handler is initialized via
+\begin{lstlisting}
+  P.RegisterOutputHandler(&out);
+  solver.RegisterOutputHandler(&out);
+\end{lstlisting}
+Finally, the programm is started with:
+\begin{lstlisting}
+      solver.ReInit();
+      out.ReInit();
+      out.Write(outp, 1, 1, 1);
+
+      solver.ComputeReducedFunctionals();
+\end{lstlisting}
+After explaining the basic compononents that are necessary for all  
+(linear and stationary) partial differential equations, we shall give 
+an overview to the implementation of the specific form of the PDE. 
+This step is done in the \texttt{localpde.h} file. As already explained,
+each PDE is treatet in a nonlinear manner such that for solving 
+the Laplace problem the following nonlinear step is necessary to recapitulate:
+Find $\delta u\in V$:
+\[
+(\nabla \delta u, \nabla \phi) = -(\nabla u, \nabla \phi) 
+\]
+and $u^{n+1} = u^n + \delta u$ for $n=1,2,3,ldots$.
+Therefore, the implementation reads for the right hand side:
+\begin{lstlisting}
+    void CellEquation(const CellDataContainer<dealii::DoFHandler<dealdim>, VECTOR, dealdim>& cdc,
+                      dealii::Vector<double> &local_cell_vector,
+                      double scale, double)
+    {
+      // implement (\nabla u, \nabla \phi) 
+    }
+\end{lstlisting}
+Specifically,
+\begin{lstlisting}
+      for (unsigned int q_point = 0; q_point < n_q_points; q_point++)
+      {
+        Tensor<2, 2> ugrads;
+        ugrads.clear();
+        ugrads[0][0] = _ugrads[q_point][0][0];
+        ugrads[0][1] = _ugrads[q_point][0][1];
+        ugrads[1][0] = _ugrads[q_point][1][0];
+        ugrads[1][1] = _ugrads[q_point][1][1];
+
+	for (unsigned int i = 0; i < n_dofs_per_cell; i++)
+        {
+	  const Tensor<2, 2> phi_i_grads_u = state_fe_values[displacements].gradient(i, q_point);
+
+          local_cell_vector(i) += scale * scalar_product(ugrads, phi_i_grads_u) 
+	    * state_fe_values.JxW(q_point);
+        }
+      }
+\end{lstlisting}
+For the left hand side, we need to implement the Jacobian matrix:
+\begin{lstlisting}
+    void CellMatrix(const CellDataContainer<dealii::DoFHandler<dealdim>, VECTOR, dealdim>& cdc, FullMatrix<double> &local_entry_matrix, double, double)
+    {
+// implement  (\nabla \delta u, \nabla \phi)
+}
+\end{lstlisting}
+Specifically,
+\begin{lstlisting}
+      for (unsigned int q_point = 0; q_point < n_q_points; q_point++)
+      {
+        for (unsigned int k = 0; k < n_dofs_per_cell; k++)
+        {
+          phi_grads_u[k] = state_fe_values[displacements].gradient(k, q_point);
+        }
+
+        for (unsigned int i = 0; i < n_dofs_per_cell; i++)
+        {
+          for (unsigned int j = 0; j < n_dofs_per_cell; j++)
+          {
+            local_entry_matrix(i, j) += scalar_product(phi_grads_u[j], phi_grads_u[i]) 
+                * state_fe_values.JxW(q_point);
+          }
+        }
+      }
+\end{lstlisting}
+The cost functional is now implemented in \todo{hier morgen weiter machen}
+
+\subsubsection{Nonstationary problems}
+In this section, we focus on the extension to 
+the numerical solution of nonstationary processes. 
+As example, we consider the heat equation:
+
+...
+
+Because the \texttt{main} file 
+is basically the same as for stationary problems
+which is one major goal of the software library,
+we only focus on the implementation of the PDE in the 
+\texttt{localpde.h} file. Here, the stationary part, i.e.,
+the diffusion is implemented as before. The only thing we add 
+is the time-depdendent part $\partial_t u$:
+
+\begin{lstlisting}
+ void CellTimeEquation(const CellDataContainer<dealii::DoFHandler<dealdim>, VECTOR, dealdim>& cdc,
+                           dealii::Vector<double> &local_cell_vector __attribute__((unused)),
+		       double scale __attribute__((unused)))
+    {
+// implement \partial_t u in weak form
+}
+\end{lstlisting}
+and it is done in a straight-forward way:
+\begin{lstlisting}
+ for (unsigned int q_point=0;q_point<n_q_points;q_point++)
+	{
+	  Tensor<1,2> v;
+	  u[0] = _uvalues[q_point](0);
+	  u[1] = _uvalues[q_point](1);
+
+	  for (unsigned int i=0;i<n_dofs_per_cell;i++)
+	    {
+	      const Tensor<1,2> phi_i_u = state_fe_values.value (i, q_point);
+
+	      local_cell_vector(i) +=  scale * (u * phi_i_u) * state_fe_values.JxW(q_point);
+	    }
+	}
+\end{lstlisting}
+
+\subsection{Solving of optimization problems}
 More complex cases involve optimization problems given in the form (OPT)
@@
 
-\subsection{Problem description}
-In order to allow our algorithms the automatic assembly of all required 
-data we need to have some container which contains the complete problem 
-description in a common data format. For this we have the following 
-classes in \texttt{DOpEsrc/container}
-\begin{itemize}
-  \item \texttt{pdeproblemcontainer.h} Is used to describe  stationary PDE problems.
-  \item \texttt{instatpdeproblemcontainer.h} {\bf TODO is still missing but will be used for nonstationary problems.} This will be implemented once we have nonstationary optimization problems running to avoid error duplication in the coding process.
-  \item \texttt{optproblem.h} Is used to describe  OPT problems governed by 
-    stationary PDEs. 
-  \item \texttt{instatoptproblemcontainer.h} Is used to describe  OPT problems
-    governed by nonstationary PDEs. The only difference to the stationary case
-    is that we need to specify a time-stepping method.  
-  \item \texttt{interfaces/functionalinterface.h} This gives an interface 
-    for the functional $J(\cdot)$ and any other functional you may want to evaluate.
-    In general this can be used as a base class to write your own functionals 
-    in examples. We note that we only need to write the integrands on 
-    elements or faces the loop over elements will be taken care of in the integrator.
-    Specifically, derivatives are written therein, too.
-\end{itemize}
-In order to fill these containers there are two things to be done,
-first we need to actually write some data, for instance,
-the semilinear form $a(\cdot)(\cdot)$, a target functional $J(\cdot)$, etc.,
-which describe the problem. Then we have to select some numerical 
-algorithm components like finite elements, linear solvers $\ldots$.
-The latter ones should be written such that when exchanging these components
-none of the problem descriptions should require changes. 
-Note that it still may be necessary to write some additional descriptions, 
-e.g., if you solve the PDE with a fix point iteration you don't need derivatives
-but if you want to use Newton's method, derivatives are needed.
+\subsection{Independent features}
+In this final section, we shall give a brief summary of features 
+that are applicable (in general) to all types of problems. Specifically,
+DOpElib provides goal-oriented mesh refinement with the help of the 
+Dual-Weighted-Residual (DWR) method \cite{BeRa96}. Here, a certain 
+cost functional such as a point evaluation, line intergration or domain
+integration is considered. Determining the lowest error between the 
+(unknown) exact solution and its numerical approximation for lowest computational
+cost is the main goal of the DWR method.
 
-We will start by discussing the problem description components implemented so far
+Moreover, DOpElib provides standard error estimators based on residual evaluation and 
+a ZZ error estimator, i.e., measureing the smoothness of the discrete first derivatives.
+The latter one is accessed via deal.II.
 
-\subsection{Numeric components}
-These are the components from which a user needs to select some in order to actually 
-solve the given problem. They will not require any rewriting, but sometimes it is 
-advisable to write other than the default parameter into the param file for the 
-solution.
+In addition, DOpElib provides multi-mesh procedures 
+\todo{kann Winni hier etwas schreiben?}
 
-\subsubsection{Space-time handler}
-First we need to select a method how to handle all dofs in space and time.
-\begin{itemize}
-\item \texttt{basic/spacetimehandler\underline{ }base.h} This class is used to define 
-  an interface to the dimension independent functionality of all space time dof handlers.
-  {\bf TODO: Beispiele geben}
-\item \texttt{basic/statespacetimehandler.h} Another intermediate interface class which adds 
-  the dimension dependent functionality if only the variable $u$ is considered, i.e., a 
-  PDE problem.
-\item \texttt{basic/spacetimehandler.h } Same as above but with both $q$ and $u$, i.e., for
-  OPT problems.
-\item \texttt{basic/mol\underline{ }statespacetimehandler.h} Implementation of a method of 
-  line space time dof handler for PDE problems. It has only one spatial 
-  dofhandler that is used for all time intervals.
-\item \texttt{basic/mol\underline{ }spacetimehandler.h} Same as above for OPT problems.
-  A separate spatial dof handler for each of the variables $q$ and $u$ is maintained 
-  but only one triangulation.
-\item \texttt{basic/mol\underline{ }multimesh\underline{ }spacetimehandler.h}
-  Same as above, but now in addition the triangulations for $q$ and $u$ can be refined
-  separately from one common initial coarse triangulation. Note that this will
-  in addition require the use of the multimesh version for integrator and 
-  face- as well as celldatacontainer.
-\end{itemize}
-Note that we use these for stationary problems as well, but then you don't have to specify
-any time information.
+Concerning temporal integration schemes, the library comes 
+with various time stepping schemes (based on finite differences), 
+such as the first-order forward Euler and first-order backward Euler scheme. 
+To compute fully nonstationary processes, the second-order schemes
+Crank-Nicolson (trapozoidal rule), 
+shifted Crank-Nicolson (slight shift to the implicit side to maintain
+global stability), 
+and the Fractional-Step-$\Theta$ scheme, are implemented.
 
-\subsubsection{Container classes}
-Second you will need to specify some container classes to be used to 
-pass data between objects. At present you don't have much choice, but you may wish 
-to reimplement some of these if you need data that is not currently included in 
-the containers.
-\begin{itemize}
-\item \texttt{container/celldatacontainer.h} This object is used to pass data 
-  given on the current element (cell) of the mesh to the functions in PDE, functional, 
-  $\ldots$. 
-\item \texttt{container/facedatacontainer.h} This object is used to pass data 
-  given on the current face of the mesh to the functions in PDE, functional, 
-  $\ldots$. 
-\item \texttt{container/multimesh\underline{ }celldatacontainer.h} This is the same as the 
-  celldatacontainer, but it
-  is capable to handle data defined on an alternative triangulation.
-\item \texttt{container/multimesh\underline{ }facedatacontainer.h} This is the same as the
-  facedatacontainer, but it
-  is capable to handle data defined on an alternative triangulation.
-\item \texttt{container/integratordatacontainer.h} This contains some data that 
-  should be passed to the integrator like quadrature formulas and the above cell and 
-  face data container.
-\end{itemize}
-
-\subsubsection{Time stepping schemes}
-\todo{Kurz halten}
-Third, at least for nonstationary PDEs we need to select a time stepping scheme
-the file names of which are mostly self explanatory:
-\begin{itemize}
-\item \texttt{include/forward\underline{ }euler\underline{ }problem.h}
-\item \texttt{include/shifted\underline{ }crank\underline{ }nicolson\underline{ }problem.h}
-\item \texttt{include/backward\underline{ }euler\underline{ }problem.h}
-\item \texttt{include/fractional\underline{ }step\underline{ }theta\underline{ }problem.h} Note that the use of this scheme requires a special Newton solver, which is, however, already
-implemented for the convenience of the user!
-\item \texttt{include/crank\underline{ }nicolson\underline{ }problem.h}
-\end{itemize}
-
-\subsubsection{Integrator routines}
-Finally, we need to select a way how to integrate and solve linear and nonlinear equations
-\begin{itemize}
-\item \texttt{templates/integrator.h} This class computes integrals over a given 
-  triangulation (including its faces).
-\item \texttt{templates/integrator\underline{ }multimesh.h} The same as above but it is 
-  possible that some of the FE functions are defined on an other triangulation 
-  as long as the have a common coarse triangulation.
-\item \texttt{templates/integratormixeddims.h} This is used to compute integral which 
-  are given in another (larger) dimension than the current variable. (This is exclusively
-  used if the control variable is given by some parameters. Which means \texttt{dopedim == 0}). 
-\end{itemize}
-
-\subsubsection{Nonlinear solvers}
-\begin{itemize}
-\item \texttt{templates/newtonsolver.h} This solves some nonlinear equation using a 
-  line-search Newton method.
-\item \texttt{templates/newtonsolvermixeddims.h} The same but in the case when there is 
-  another variable in a (larger) dimension is involved. See 
-  \texttt{integratormixeddims.h}.
-\item \texttt{templates/instat\underline{ }step\underline{ }newtonsolver.h} This is a 
-  Newton method as above to invert the next time-step. It differs from the plain vanilla
-  version in that it computes certain data from the previous time step only once 
-  and not in every Newton iteration.
-\item \texttt{templates/fractional\underline{ }step\underline{ }theta\underline{ }step\underline{ }newtonsolver.h} This is the Newton solver for the time step in a 
-  fractional-step-theta scheme. It combines the computation of all three sub steps.
-\end{itemize}
-
-\subsubsection{Linear solvers}
-To solve the linear equations within Newton's method, we use 
-standard solvers from iterative type 
-such as the method of the conjugate gradients (CG method), 
-GMRES, or direct solvers such as \texttt{UMFPACK}.
-In addition, we DOpE provides  a wrapper class 
-\texttt{templates/voidlinearsolver.h} for certain cases when we 
-  know that the matrix to be inverted is the identity. It simply copies the rhs to the
-  lhs. This is only needed for compatibility reasons some other components.
-
+%\subsection{Problem description}
+%In order to allow our algorithms the automatic assembly of all required 
+%data we need to have some container which contains the complete problem 
+%description in a common data format. For this we have the following 
+%classes in \texttt{DOpEsrc/container}
 %\begin{itemize}
-%\item \texttt{templates/cglinearsolver.h} This is a wrapper for the cg solver implemented in 
-%  \texttt{deal.II}. The solver will build and store the stiffness matrix for the PDE.
-%\item \texttt{templates/gmreslinearsolver.h} This is a wrapper for the gmres solver 
-%  implemented in \texttt{deal.II}. The solver will build and store the stiffness matrix 
-%  for the PDE.
-%\item \texttt{templates/directlinearsolver.h} This is a wrapper for the direct solver 
-%  implemented in \texttt{deal.II} using \texttt{UMFPACK}. 
-%  The solver will build and store the stiffness matrix for the PDE.
-%\item \texttt{templates/voidlinearsolver.h} This is a wrapper for certain cases when we 
+%  \item \texttt{pdeproblemcontainer.h} Is used to describe  stationary PDE problems.
+%  \item \texttt{instatpdeproblemcontainer.h} {\bf TODO is still missing but will be used for nonstationary problems.} This will be implemented once we have nonstationary optimization problems running to avoid error duplication in the coding process.
+%  \item \texttt{optproblem.h} Is used to describe  OPT problems governed by 
+%    stationary PDEs. 
+%  \item \texttt{instatoptproblemcontainer.h} Is used to describe  OPT problems
+%    governed by nonstationary PDEs. The only difference to the stationary case
+%    is that we need to specify a time-stepping method.  
+%  \item \texttt{interfaces/functionalinterface.h} This gives an interface 
+%    for the functional $J(\cdot)$ and any other functional you may want to evaluate.
+%    In general this can be used as a base class to write your own functionals 
+%    in examples. We note that we only need to write the integrands on 
+%    elements or faces the loop over elements will be taken care of in the integrator.
+%    Specifically, derivatives are written therein, too.
+%\end{itemize}
+%In order to fill these containers there are two things to be done,
+%first we need to actually write some data, for instance,
+%the semilinear form $a(\cdot)(\cdot)$, a target functional $J(\cdot)$, etc.,
+%which describe the problem. Then we have to select some numerical 
+%algorithm components like finite elements, linear solvers $\ldots$.
+%The latter ones should be written such that when exchanging these components
+%none of the problem descriptions should require changes. 
+%Note that it still may be necessary to write some additional descriptions, 
+%e.g., if you solve the PDE with a fix point iteration you don't need derivatives
+%but if you want to use Newton's method, derivatives are needed.
+%
+%We will start by discussing the problem description components implemented so far
+%
+%
+%\subsection{Numeric components}
+%These are the components from which a user needs to select some in order to actually 
+%solve the given problem. They will not require any rewriting, but sometimes it is 
+%advisable to write other than the default parameter into the param file for the 
+%solution.
+%
+%\subsubsection{Space-time handler}
+%First we need to select a method how to handle all dofs in space and time.
+%\begin{itemize}
+%\item \texttt{basic/spacetimehandler\underline{ }base.h} This class is used to define 
+%  an interface to the dimension independent functionality of all space time dof handlers.
+%  {\bf TODO: Beispiele geben}
+%\item \texttt{basic/statespacetimehandler.h} Another intermediate interface class which adds 
+%  the dimension dependent functionality if only the variable $u$ is considered, i.e., a 
+%  PDE problem.
+%\item \texttt{basic/spacetimehandler.h } Same as above but with both $q$ and $u$, i.e., for
+%  OPT problems.
+%\item \texttt{basic/mol\underline{ }statespacetimehandler.h} Implementation of a method of 
+%  line space time dof handler for PDE problems. It has only one spatial 
+%  dofhandler that is used for all time intervals.
+%\item \texttt{basic/mol\underline{ }spacetimehandler.h} Same as above for OPT problems.
+%  A separate spatial dof handler for each of the variables $q$ and $u$ is maintained 
+%  but only one triangulation.
+%\item \texttt{basic/mol\underline{ }multimesh\underline{ }spacetimehandler.h}
+%  Same as above, but now in addition the triangulations for $q$ and $u$ can be refined
+%  separately from one common initial coarse triangulation. Note that this will
+%  in addition require the use of the multimesh version for integrator and 
+%  face- as well as celldatacontainer.
+%\end{itemize}
+%Note that we use these for stationary problems as well, but then you don't have to specify
+%any time information.
+%
+%\subsubsection{Container classes}
+%Second you will need to specify some container classes to be used to 
+%pass data between objects. At present you don't have much choice, but you may wish 
+%to reimplement some of these if you need data that is not currently included in 
+%the containers.
+%\begin{itemize}
+%\item \texttt{container/celldatacontainer.h} This object is used to pass data 
+%  given on the current element (cell) of the mesh to the functions in PDE, functional, 
+%  $\ldots$. 
+%\item \texttt{container/facedatacontainer.h} This object is used to pass data 
+%  given on the current face of the mesh to the functions in PDE, functional, 
+%  $\ldots$. 
+%\item \texttt{container/multimesh\underline{ }celldatacontainer.h} This is the same as the 
+%  celldatacontainer, but it
+%  is capable to handle data defined on an alternative triangulation.
+%\item \texttt{container/multimesh\underline{ }facedatacontainer.h} This is the same as the
+%  facedatacontainer, but it
+%  is capable to handle data defined on an alternative triangulation.
+%\item \texttt{container/integratordatacontainer.h} This contains some data that 
+%  should be passed to the integrator like quadrature formulas and the above cell and 
+%  face data container.
+%\end{itemize}
+%
+%
+%\subsubsection{Time stepping schemes}
+%\todo{Kurz halten}
+%Third, at least for nonstationary PDEs we need to select a time stepping scheme
+%the file names of which are mostly self explanatory:
+%\begin{itemize}
+%\item \texttt{include/forward\underline{ }euler\underline{ }problem.h}
+%\item \texttt{include/shifted\underline{ }crank\underline{ }nicolson\underline{ }problem.h}
+%\item \texttt{include/backward\underline{ }euler\underline{ }problem.h}
+%\item \texttt{include/fractional\underline{ }step\underline{ }theta\underline{ }problem.h} Note that the use of this scheme requires a special Newton solver, which is, however, already
+%implemented for the convenience of the user!
+%\item \texttt{include/crank\underline{ }nicolson\underline{ }problem.h}
+%\end{itemize}
+%
+%\subsubsection{Integrator routines}
+%Finally, we need to select a way how to integrate and solve linear and nonlinear equations
+%\begin{itemize}
+%\item \texttt{templates/integrator.h} This class computes integrals over a given 
+%  triangulation (including its faces).
+%\item \texttt{templates/integrator\underline{ }multimesh.h} The same as above but it is 
+%  possible that some of the FE functions are defined on an other triangulation 
+%  as long as the have a common coarse triangulation.
+%\item \texttt{templates/integratormixeddims.h} This is used to compute integral which 
+%  are given in another (larger) dimension than the current variable. (This is exclusively
+%  used if the control variable is given by some parameters. Which means \texttt{dopedim == 0}). 
+%\end{itemize}
+%
+%\subsubsection{Nonlinear solvers}
+%\begin{itemize}
+%\item \texttt{templates/newtonsolver.h} This solves some nonlinear equation using a 
+%  line-search Newton method.
+%\item \texttt{templates/newtonsolvermixeddims.h} The same but in the case when there is 
+%  another variable in a (larger) dimension is involved. See 
+%  \texttt{integratormixeddims.h}.
+%\item \texttt{templates/instat\underline{ }step\underline{ }newtonsolver.h} This is a 
+%  Newton method as above to invert the next time-step. It differs from the plain vanilla
+%  version in that it computes certain data from the previous time step only once 
+%  and not in every Newton iteration.
+%\item \texttt{templates/fractional\underline{ }step\underline{ }theta\underline{ }step\underline{ }newtonsolver.h} This is the Newton solver for the time step in a 
+%  fractional-step-theta scheme. It combines the computation of all three sub steps.
+%\end{itemize}
+%
+%\subsubsection{Linear solvers}
+%To solve the linear equations within Newton's method, we use 
+%standard solvers from iterative type 
+%such as the method of the conjugate gradients (CG method), 
+%GMRES, or direct solvers such as \texttt{UMFPACK}.
+%In addition, we DOpE provides  a wrapper class 
+%\texttt{templates/voidlinearsolver.h} for certain cases when we 
 %  know that the matrix to be inverted is the identity. It simply copies the rhs to the
 %  lhs. This is only needed for compatibility reasons some other components.
+%
+%
+%
+%
+%%\begin{itemize}
+%%\item \texttt{templates/cglinearsolver.h} This is a wrapper for the cg solver implemented in 
+%%  \texttt{deal.II}. The solver will build and store the stiffness matrix for the PDE.
+%%\item \texttt{templates/gmreslinearsolver.h} This is a wrapper for the gmres solver 
+%%  implemented in \texttt{deal.II}. The solver will build and store the stiffness matrix 
+%%  for the PDE.
+%%\item \texttt{templates/directlinearsolver.h} This is a wrapper for the direct solver 
+%%  implemented in \texttt{deal.II} using \texttt{UMFPACK}. 
+%%  The solver will build and store the stiffness matrix for the PDE.
+%%\item \texttt{templates/voidlinearsolver.h} This is a wrapper for certain cases when we 
+%%  know that the matrix to be inverted is the identity. It simply copies the rhs to the
+%%  lhs. This is only needed for compatibility reasons some other components.
+%%\end{itemize}
+%%
+%
+%
+%%\subsection{Problem specific classes}
+%%The following classes are used to describe the problem and will usually require 
+%%some implementation.
+%%
+%%\begin{itemize}
+%%  \item \texttt{basic/constraints.h} This is used by the spacetimehandlers to 
+%%    compute the number of constraints from the control and state vectors. 
+%%    It must not be reimplemented by the user, but needs to be properly 
+%%    initialized if OPT is used with box control constraints or $g(q,u) \le 0$.
+%%
+%%  \item \texttt{interfaces/constraintinterface.h} This gives an interface for both 
+%%    the control box constraints as well as the general constraint $g \le 0$. This 
+%%    needs to be specified if constraints are to be used. If they are not needed 
+%%    a default class \texttt{problemdata/noconstraints.h} can be used. We note that we only 
+%%    need to write the integrands on 
+%%    elements or faces the loop over elements will be taken care of in the integrator.
+%%  \item \texttt{interfaces/pdeinterface.h} This defines an interface for the 
+%%    partial differential equation $a(q,u)(\phi) = 0$. This needs to be written
+%%    by the user. We note that we only need to write the integrands on 
+%%    elements or faces the loop over elements will be taken care of in the integrator.
+%%    Specifically, derivatives are written therein, too.
+%%  \item \texttt{interfaces/dirichletdatainterface.h} This gives an interface to the 
+%%    Dirichlet data for a problem. If the Dirichlet data are simply a function 
+%%    (and do not depend on the control $q$) one can use the default class\\
+%%    \texttt{problemdata/simpledirichletdata.h}.
+%%\end{itemize}
+%%
+%
+%
+%
+%\subsection{Reduced problems (Solve the PDE)}
+%At times it is nice to remove the PDE constraint in (OPT). 
+%This is handled by so called reduced 
+%problems (for algorithmic aspects we refer the reader to 
+%\cite{BeMeVe06}). 
+%This means that the reduced problem implicitly solves the PDE whenever required
+%and eliminates the variable $u$ from the problem.
+%\begin{itemize}
+%\item \texttt{reducedproblems/statpdeproblem.h} This is used to remove the variable $u$ in 
+%  a stationary PDE problem. This means that call the method \\
+%  \texttt{StatPDEProblem::ComputeReducedFunctionals} will evaluate the functionals 
+%  defined in the problem description, i.e., in \texttt{PDEProblemContainer}, in the 
+%  solution of the given PDE.
+%\item \texttt{reducedproblems/statreducedproblem.h} This eliminates $u$ from the OPT
+%  problem with a stationary PDE.
+%\item \texttt{reducedproblems/instatreducedproblem.h} The same as above but for a
+%  nonstationary PDE. {\bf FIXME there is something wrong in this file see FIXME 
+%    comment in the source.}
+%\item \texttt{reducedproblems/voidreducedproblem.h} A wrapper file that eliminates $u$ 
+%  if it is not present anyways. This is used so that we can use the same routines to 
+%  solve problems that have no PDE constraint.
 %\end{itemize}
 %
-
-%\subsection{Problem specific classes}
-%The following classes are used to describe the problem and will usually require 
-%some implementation.
-%
+%\subsection{Optimization algorithms}
+%Now, in order to solve optimization algorithms we need to define some algorithms.
+%At present we offer a selection of algorithms that solve the reduced optimization 
+%problem where the PDE constraint has been eliminated as explained in the previous section.
 %\begin{itemize}
-%  \item \texttt{basic/constraints.h} This is used by the spacetimehandlers to 
-%    compute the number of constraints from the control and state vectors. 
-%    It must not be reimplemented by the user, but needs to be properly 
-%    initialized if OPT is used with box control constraints or $g(q,u) \le 0$.
-%
-%  \item \texttt{interfaces/constraintinterface.h} This gives an interface for both 
-%    the control box constraints as well as the general constraint $g \le 0$. This 
-%    needs to be specified if constraints are to be used. If they are not needed 
-%    a default class \texttt{problemdata/noconstraints.h} can be used. We note that we only 
-%    need to write the integrands on 
-%    elements or faces the loop over elements will be taken care of in the integrator.
-%  \item \texttt{interfaces/pdeinterface.h} This defines an interface for the 
-%    partial differential equation $a(q,u)(\phi) = 0$. This needs to be written
-%    by the user. We note that we only need to write the integrands on 
-%    elements or faces the loop over elements will be taken care of in the integrator.
-%    Specifically, derivatives are written therein, too.
-%  \item \texttt{interfaces/dirichletdatainterface.h} This gives an interface to the 
-%    Dirichlet data for a problem. If the Dirichlet data are simply a function 
-%    (and do not depend on the control $q$) one can use the default class\\
-%    \texttt{problemdata/simpledirichletdata.h}.
+%\item \texttt{opt\underline{ }algorithms/reducedalgorithm.h} An interface for all 
+%  optimization problems in the reduced formulation. It offers some test functionality
+%  to assert that the derivatives of the problem are computed correctly.
+%\item \texttt{opt\underline{ }algorithms/reducednewtonalgorithm.h}
+%  A line-search Newton algorithm using a cg method to invert the reduced hessian. 
+%  Implementation ignores any additional constraints.
+%\item \texttt{opt\underline{ }algorithms/reducedtrustregionnewton.h}
+%  A trust region Newton algorithm using a cg method to invert the reduced hessian.
+%  Implementation ignores any additional constraints.
+%\item \texttt{opt\underline{ }algorithms/reduced\underline{ }snopt\underline{ }algorithm.h}
+%  An algorithm to solve reduced optimization problems with additional control constraints.
+%  ((reduced) state constraints are not yet implemented.)
+%\item \texttt{opt\underline{ }algorithms/reducednewtonalgorithmwithinverse.h}
+%  Line-search Newton algorithm that assumes there exists a method in the reduced problem
+%  that can invert the reduced hessian. (This usually makes sense only if there is no 
+%  PDE constraint.)
+%\item \texttt{opt\underline{ }algorithms/generalized\underline{ }mma\underline{ }algorithm.h}
+%  An implementation of the MMA-Algorithm for structural optimization using an augmented
+%  Lagrangian formulation for the subproblems. The subproblem is implemented using the 
+%  special purpose file
+%  \texttt{include/augmentedlagrangianproblem.h}.
 %\end{itemize} 
 %
-
-\subsection{Reduced problems (Solve the PDE)}
-At times it is nice to remove the PDE constraint in (OPT). 
-This is handled by so called reduced 
-problems (for algorithmic aspects we refer the reader to 
-\cite{BeMeVe06}). 
-This means that the reduced problem implicitly solves the PDE whenever required
-and eliminates the variable $u$ from the problem.
-\begin{itemize}
-\item \texttt{reducedproblems/statpdeproblem.h} This is used to remove the variable $u$ in 
-  a stationary PDE problem. This means that call the method \\
-  \texttt{StatPDEProblem::ComputeReducedFunctionals} will evaluate the functionals 
-  defined in the problem description, i.e., in \texttt{PDEProblemContainer}, in the 
-  solution of the given PDE.
-\item \texttt{reducedproblems/statreducedproblem.h} This eliminates $u$ from the OPT
-  problem with a stationary PDE.
-\item \texttt{reducedproblems/instatreducedproblem.h} The same as above but for a
-  nonstationary PDE. {\bf FIXME there is something wrong in this file see FIXME 
-    comment in the source.}
-\item \texttt{reducedproblems/voidreducedproblem.h} A wrapper file that eliminates $u$ 
-  if it is not present anyways. This is used so that we can use the same routines to 
-  solve problems that have no PDE constraint.
-\end{itemize}
-
-\subsection{Optimization algorithms}
-Now, in order to solve optimization algorithms we need to define some algorithms.
-At present we offer a selection of algorithms that solve the reduced optimization 
-problem where the PDE constraint has been eliminated as explained in the previous section.
-\begin{itemize}
-\item \texttt{opt\underline{ }algorithms/reducedalgorithm.h} An interface for all 
-  optimization problems in the reduced formulation. It offers some test functionality
-  to assert that the derivatives of the problem are computed correctly.
-\item \texttt{opt\underline{ }algorithms/reducednewtonalgorithm.h}
-  A line-search Newton algorithm using a cg method to invert the reduced hessian. 
-  Implementation ignores any additional constraints.
-\item \texttt{opt\underline{ }algorithms/reducedtrustregionnewton.h}
-  A trust region Newton algorithm using a cg method to invert the reduced hessian.
-  Implementation ignores any additional constraints.
-\item \texttt{opt\underline{ }algorithms/reduced\underline{ }snopt\underline{ }algorithm.h}
-  An algorithm to solve reduced optimization problems with additional control constraints.
-  ((reduced) state constraints are not yet implemented.)
-\item \texttt{opt\underline{ }algorithms/reducednewtonalgorithmwithinverse.h}
-  Line-search Newton algorithm that assumes there exists a method in the reduced problem
-  that can invert the reduced hessian. (This usually makes sense only if there is no 
-  PDE constraint.)
-\item \texttt{opt\underline{ }algorithms/generalized\underline{ }mma\underline{ }algorithm.h}
-  An implementation of the MMA-Algorithm for structural optimization using an augmented
-  Lagrangian formulation for the subproblems. The subproblem is implemented using the 
-  special purpose file
-  \texttt{include/augmentedlagrangianproblem.h}.
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Interfaces to other software packages}
-Since DOpE uses as basis the software library deal.II, all its interfaces
-to other packages can be accessed to. For instance, such libararies 
-are trilinos for with an algebraic multigrid method, MPI for a parallel
-solution of the problem.
-
-In addition DOpE itself has an interface to SNOPT.
-
-\section{Documentation, Code Development, Webpage}
-\label{documentation}
-At the present step, the DOpElib project comes with 
-an detailed documentation of all features and examples 
-in pdf format and an detailed programming code 
-documentation.
-
-\todo{License? How to access to DOpE?}
-\todo{Name of the website}
-
-In addition, the DOpElib test suite provides some regression tests. 
-They are run to compare 
-the output to previous outputs. This is useful (necessary) after 
-changing programming code anywhere in the library. If a test
-succeeds, everything is fine in the library. If not, you should not
-check in your code into DOpE. Please make sure what is going wrong and WHY!
-Every command is computed via a Makefile.
-
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%\subsection{Interfaces to other software packages}
+%Since DOpE uses as basis the software library deal.II, all its interfaces
+%to other packages can be accessed to. For instance, such libararies 
+%are trilinos for with an algebraic multigrid method, MPI for a parallel
+%solution of the problem.
+%
+%In addition DOpE itself has an interface to SNOPT.
+%
+%%\section{Documentation, Code Development, Webpage}
+%%\label{documentation}
+%%At the present step, the DOpElib project comes with 
+%%an detailed documentation of all features and examples 
+%%in pdf format and an detailed programming code 
+%%documentation. 
+%%
+%%\todo{License? How to access to DOpE?}
+%%\todo{Name of the website}
+%%
+%%In addition, the DOpElib test suite provides some regression tests. 
+%%They are run to compare 
+%%the output to previous outputs. This is useful (necessary) after 
+%%changing programming code anywhere in the library. If a test
+%%succeeds, everything is fine in the library. If not, you should not
+%%check in your code into DOpE. Please make sure what is going wrong and WHY!
+%%Every command is computed via a Makefile. 
+%%
+%
+%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{applications}
-We present some numerical tests that have been already computed 
-with DOpElib. These tests demonstrate the performance 
-of DOpElib by covering different computational aspects. 
+\todo{In den Beispielen in Pseudo code evtl. zeigen, was der 
+Nutzer aendern kann.}
+We present three numerical tests which demonstrate different
+features of DOpElib:
 \begin{itemize}
-\item The first numerical test shows various time-stepping schemes
-and the numerical solution of nonlinear coupled partial differential
-equations, formulated in a monolithic solution algorithm. 
-Specifically, the equations are solved by the method of lines, i.e,
-first spatial discretization and then temporal  discretization 
-with a Galerkin finite element scheme. The nonlinaer problem
-is solved by Newton's method. The Jacobian is derived 
-by analytical expressions of the derivatives. 
-In DOpElib this example is implemented in 
-\texttt{Example/PDE/InstatPDE/Example2}
-\item The second example presents an
-optimization problem and its numerical solution.
-\item The main purpose of the third example is the application 
-of refinement techniques for local mesh refinement for 
-two and three dimensional problems. We consider the numerical
-solution of fluid flows governed by the Navier-Stokes
-equations. In contrast to the previous tests, we 
-use finite elements on locally adapted meshes 
-with goal-oriented mesh refinement with the help
-of the dual-weighted residual method. Using 
-this method of local mesh refinement, the error estimator
-is derived with respect to some goal functional and 
-provides optimal meshes to the accurate measurement 
-of the considered functional value. Consequently,
-this is an effiecient method to reduce the computational 
-cost. Most other 
-software libraries does not provide goal-oriented 
-mesh refinement.  
-\item Here, we show a time-dependent optimization problem or 
-FSI optimization. 
+\item In the first numerical example, we consider 
+goal-oriented mesh refinement with the help of the 
+Dual-Weighted-Residual (DWR) method for the stationary Navier-Stokes equations.
+\item The second example presents nonstationary fluid-structure 
+interaction. The challenges for those kind of problems are the multi-domain
+character of the coupled problem and the treatment of the coupling conditions.
+The problem is formulated in a variational monolithically-coupled way which 
+allows to consider goal-oriented mesh refinement and gradient-based optimization.
+\item In the third numerical test, an optmization problem for structural mechanics
+is discussed.
 \end{itemize}
 
+\subsection{Goal-oriented mesh refinement for Navier-Stokes}
+Christian.
+
 \subsection{Nonstationary Fluid-Structure Interaction}
-In this example, the three proposed mesh motion models are applied 
-to an unsteady fluid-structure interaction problem. We consider 
-the numerical benchmark test FSI 2, which was proposed in \cite{HrTu06b}. 
+In this example, the challenging benchmark FSI 2
+proposed by Hron and Turek \cite{HrTu06b} is considered.
+To solve this problem accurately, it is very important that 
+the coupling conditions
+\[
+v_f = v_s \quad \text{and} \quad \sigma_f n_f = \sigma_s n_s, 
+\]
+namely, the continuity of velocities and continuity of normal stresses
+are satisfied in each time step. This is achieved with the help of 
+a monolithic coupling scheme (strong coupling).
+
 The configuration is 
-sketched in Figure \ref{configuration_csm_and_fsi_2D}.
-New results can be found in \cite{BuSc06, TuHrMaRaWoAc10, DeHaeAnnBrVie10}. 
-The Fractional-Step-$\theta$ scheme was used for time discretization with
-different time step sizes $k$.
+sketched in Figure \ref{configuration_csm_and_fsi_2D}. Details 
+on parameters and evaluation functionals and other results 
+can be found in \cite{HrTu06b, BuSc06, TuHrMaRaWoAc10, DeHaeAnnBrVie10, Wi11}. 
+The Fractional-Step-$\theta$ scheme is used for time discretization with
+different time step sizes $k$. Since the problem is fully nonstationary, the 
+dynamics are presents in Figure \ref{res:fsi_2_mesh_and_x_velo}. 
+By comparison of our findings in Figure \ref{res:results_ux_and_uy_fsi_2}
+with the literature, it can be identified that 
+the implementation in DOpElib is correct.
 
@@
 \subsection{Compliance minimization}
+Winni
 Results already obtained.
-
-\subsection{Goal-oriented mesh refinement for Navier-Stokes}
-Has to be computed in two and three dimensions.
-
-\subsection{Time-dependent Opimization or FSI optimization}
 

### 2012-10-05 Paper_May_2012/main.tex
Paper: Nonstationary problems hinzugefuegt und mit Optimization problems begonnen und Ideen fuer dessen Realisierung im Paper hinzugefuegt

@@
 THOMAS WICK
-\affil{University of Heidelberg}
+\affil{The University of Texas at Austin}
 WINNIFRIED WOLLNER
@@
 \begin{bottomstuff}
-Author's adresses: C. Goll {and} T. Wick, Institut f\"ur Angewandte Mathematik, Universit\"at Heidelberg;
+Author's adresses: C. Goll, Institut f\"ur Angewandte Mathematik,
+Universit\"at Heidelberg;
+T. Wick, The Institute for Computational Engineering and Sciences;
 W. Wollner, Fachbereich Mathematik, Universit\"at Hamburg.
@@
 \label{detailed_description}
-\todo{Aus der Sicht des Users anhand eines konkreten Beispiels beschreiben}
-\todo{Einfaches Beispiel nehmen}
-\todo{erwaehnen, dass wir alles immer nicht-linear betrachten}
-\todo{Aussehen eines typischen PDE und Opt loop und deren Gemeinsamkeiten
-erklaeren. Vielleicht mir Pseudocode}
+%\todo{Aus der Sicht des Users anhand eines konkreten Beispiels beschreiben}
+%\todo{Einfaches Beispiel nehmen}
+%\todo{erwaehnen, dass wir alles immer nicht-linear betrachten}
+%\todo{Aussehen eines typischen PDE und Opt loop und deren Gemeinsamkeiten
+%erklaeren. Vielleicht mir Pseudocode}
 This library is designed to allow easy implementation and numerical solutions 
@@
 \end{lstlisting}
-Then, the problem (here the Laplace PDE) is given to the \texttt{DOFH} via
+This method handles all dofs in space and time and different versions exist
+for forward as well as optimization problems.
+
+Afterwards, the problem (here the Laplace PDE) is given to the \texttt{DOFH} via
 \begin{lstlisting}
@@
 \end{lstlisting}
+The integration of the equation for each DoF and each quadrature point on each
+cell is then done 
+in the \texttt{templates/integrator.h} function.
+
 The cost functional is now implemented in \todo{hier morgen weiter machen}
@@
 As example, we consider the heat equation:
-
-...
+Find $u$:
+\[
+\partial_t u - \Delta u = 0
+\]
+and in weak formulation:
+Find $u$:
+\[
+a(u)(\phi) = 0 \quad \forall \phi \in V,
+\]
+with some appropriate space $V$ and 
+\[
+a(u)(\phi) = (\partial_t u, \phi) + (\nabla u, \nabla \phi) \quad \forall \phi \in V.
+\]
+Temporal discretization using the One-step-$\theta$ scheme reads:
+Given the old timestep solution $u^n$, find $u^{n+1}$:
+\[
+\frac{(u^{n+1} - u^{n}, \phi)}{k} + \theta (\nabla u^{n+1}, \nabla \phi)
++ (1 - \theta) (\nabla u^{n}, \nabla \phi) = 0,
+\]
+Reordering in new and old terms and multiplication by the time step size $k$:
+\[
+(u^{n+1},\phi) + k \theta (\nabla u^{n+1}, \nabla \phi)
+= (u^n, \phi) - (1 - \theta) (\nabla u^{n}, \nabla \phi).
+\]
+In DOpElib, different time stepping schemes can be found 
+in the folder \texttt{DOpEsrc/tsschemes}.
 
@@
     {
-// implement \partial_t u in weak form
+// implement (u^{n+1},\phi) 
 }
@@
 \end{lstlisting}
+Here, the advantage is that DOpElib orders the terms in the timestepping
+problems
+such that only $(u^{n+1},\phi)$ must be implemented. However, if someone 
+needs explicit input of the old timestep term, then he can use 
+\begin{lstlisting}
+ void CellTimeEquationExplicit (const CellDataContainer<dealii::DoFHandler<dealdim>, VECTOR, dealdim>& cdc,
+                           dealii::Vector<double> &local_cell_vector,
+				   double scale)
+    {
+      // implement (u^{n+1},\phi) and (u^n,\phi)  
+    }
+    \end{lstlisting}
+    For instance, this explicit term is used for fluid-structure interaction in
+    which
+    the old timestep term needs explicit modification.
 
 \subsection{Solving of optimization problems}
-More complex cases involve optimization problems given in the form (OPT)
+Upgrading the framework for solving optimization problems
+is described in this section. A prototypical problem reads:
 \begin{align*}
@@
 and $g(\cdot)$ is some state constraint.
+As before, the main part is done in the \texttt{main} function there is no
+great difference to the previous explications. However, 
+since we utilize gradient-based optimization routines \cite{...},
+all derivatives must be implemented. All this information
+must given to the \texttt{localpde.h} file. Herein, one has to 
+implement the adjoint, tangent and adjoint Hession equations. 
+The algorithmic parts are described in the literature \cite{BeMeVe06}.
+
+Concretely, let us focus on the Laplace equation and ...
+
+this is taken from step-2 in the Opt examples.
+
+\todo{Optimality system}
+
+\todo{Two problems in which we identify the different terms and implementation 
+of these terms in DOpE}
 
@@
 %
-%
-%\subsubsection{Time stepping schemes}
-%\todo{Kurz halten}
-%Third, at least for nonstationary PDEs we need to select a time stepping scheme
-%the file names of which are mostly self explanatory:
-%\begin{itemize}
-%\item \texttt{include/forward\underline{ }euler\underline{ }problem.h}
-%\item \texttt{include/shifted\underline{ }crank\underline{ }nicolson\underline{ }problem.h}
-%\item \texttt{include/backward\underline{ }euler\underline{ }problem.h}
-%\item \texttt{include/fractional\underline{ }step\underline{ }theta\underline{ }problem.h} Note that the use of this scheme requires a special Newton solver, which is, however, already
-%implemented for the convenience of the user!
-%\item \texttt{include/crank\underline{ }nicolson\underline{ }problem.h}
-%\end{itemize}
-%
-%\subsubsection{Integrator routines}
-%Finally, we need to select a way how to integrate and solve linear and nonlinear equations
-%\begin{itemize}
-%\item \texttt{templates/integrator.h} This class computes integrals over a given 
-%  triangulation (including its faces).
-%\item \texttt{templates/integrator\underline{ }multimesh.h} The same as above but it is 
-%  possible that some of the FE functions are defined on an other triangulation 
-%  as long as the have a common coarse triangulation.
-%\item \texttt{templates/integratormixeddims.h} This is used to compute integral which 
-%  are given in another (larger) dimension than the current variable. (This is exclusively
-%  used if the control variable is given by some parameters. Which means \texttt{dopedim == 0}). 
-%\end{itemize}
-%
-%\subsubsection{Nonlinear solvers}
-%\begin{itemize}
-%\item \texttt{templates/newtonsolver.h} This solves some nonlinear equation using a 
-%  line-search Newton method.
-%\item \texttt{templates/newtonsolvermixeddims.h} The same but in the case when there is 
-%  another variable in a (larger) dimension is involved. See 
-%  \texttt{integratormixeddims.h}.
-%\item \texttt{templates/instat\underline{ }step\underline{ }newtonsolver.h} This is a 
-%  Newton method as above to invert the next time-step. It differs from the plain vanilla
-%  version in that it computes certain data from the previous time step only once 
-%  and not in every Newton iteration.
-%\item \texttt{templates/fractional\underline{ }step\underline{ }theta\underline{ }step\underline{ }newtonsolver.h} This is the Newton solver for the time step in a 
-%  fractional-step-theta scheme. It combines the computation of all three sub steps.
-%\end{itemize}
-%
-%\subsubsection{Linear solvers}
-%To solve the linear equations within Newton's method, we use 
-%standard solvers from iterative type 
-%such as the method of the conjugate gradients (CG method), 
-%GMRES, or direct solvers such as \texttt{UMFPACK}.
-%In addition, we DOpE provides  a wrapper class 
-%\texttt{templates/voidlinearsolver.h} for certain cases when we 
-%  know that the matrix to be inverted is the identity. It simply copies the rhs to the
-%  lhs. This is only needed for compatibility reasons some other components.
-%
-%
-%
-%
-%%\begin{itemize}
-%%\item \texttt{templates/cglinearsolver.h} This is a wrapper for the cg solver implemented in 
-%%  \texttt{deal.II}. The solver will build and store the stiffness matrix for the PDE.
-%%\item \texttt{templates/gmreslinearsolver.h} This is a wrapper for the gmres solver 
-%%  implemented in \texttt{deal.II}. The solver will build and store the stiffness matrix 
-%%  for the PDE.
-%%\item \texttt{templates/directlinearsolver.h} This is a wrapper for the direct solver 
-%%  implemented in \texttt{deal.II} using \texttt{UMFPACK}. 
-%%  The solver will build and store the stiffness matrix for the PDE.
-%%\item \texttt{templates/voidlinearsolver.h} This is a wrapper for certain cases when we 
-%%  know that the matrix to be inverted is the identity. It simply copies the rhs to the
-%%  lhs. This is only needed for compatibility reasons some other components.
-%%\end{itemize}
-%%
-%
-%
-%%\subsection{Problem specific classes}
-%%The following classes are used to describe the problem and will usually require 
-%%some implementation.
-%%
-%%\begin{itemize}
-%%  \item \texttt{basic/constraints.h} This is used by the spacetimehandlers to 
-%%    compute the number of constraints from the control and state vectors. 
-%%    It must not be reimplemented by the user, but needs to be properly 
-%%    initialized if OPT is used with box control constraints or $g(q,u) \le 0$.
-%%
-%%  \item \texttt{interfaces/constraintinterface.h} This gives an interface for both 
-%%    the control box constraints as well as the general constraint $g \le 0$. This 
-%%    needs to be specified if constraints are to be used. If they are not needed 
-%%    a default class \texttt{problemdata/noconstraints.h} can be used. We note that we only 
-%%    need to write the integrands on 
-%%    elements or faces the loop over elements will be taken care of in the integrator.
-%%  \item \texttt{interfaces/pdeinterface.h} This defines an interface for the 
-%%    partial differential equation $a(q,u)(\phi) = 0$. This needs to be written
-%%    by the user. We note that we only need to write the integrands on 
-%%    elements or faces the loop over elements will be taken care of in the integrator.
-%%    Specifically, derivatives are written therein, too.
-%%  \item \texttt{interfaces/dirichletdatainterface.h} This gives an interface to the 
-%%    Dirichlet data for a problem. If the Dirichlet data are simply a function 
-%%    (and do not depend on the control $q$) one can use the default class\\
-%%    \texttt{problemdata/simpledirichletdata.h}.
-%%\end{itemize}
-%%
 %
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions and Outlook}
+\section{Conclusions}
 \label{conclusions}
@@
 problems governed by PDEs.
+
+\section{Outlook}
 \todo{Was wollen wir noch in den nächsten 1-2 Jahren erreichen?}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Acknowledgment}
-The second author thanks Rolf Rannacher from
-the Institute for Applied mathematics at the 
-Heidelberg University for financial suppport.
 

### 2012-10-05 Paper_May_2012/main.tex
Paper: Im FSI einen Code-Schnipsel hinzugefuegt, der veranschaulicht, wie einfach andere Zeitschrittverfahren ausgewaehlt werden koennen

@@
 \label{applications}
-\todo{In den Beispielen in Pseudo code evtl. zeigen, was der 
-Nutzer aendern kann.}
 We present three numerical tests which demonstrate different
@@
 The Fractional-Step-$\theta$ scheme is used for time discretization with
-different time step sizes $k$. Since the problem is fully nonstationary, the 
+different time step sizes $k$. However, the time stepping scheme can be 
+very easily chosen in the \texttt{main.cc} function by declaring
+\begin{lstlisting}
+#define TSP1 ForwardEulerProblem
+#define TSP2 BackwardEulerProblem
+#define TSP3 CrankNicolsonProblem
+#define TSP4 ShiftedCrankNicolsonProblem
+#define TSP5 FractionalStepThetaProblem
+\end{lstlisting}
+and then using the designated scheme.
+
+Since the problem is fully nonstationary, the 
 dynamics are presents in Figure \ref{res:fsi_2_mesh_and_x_velo}. 

### 2012-10-05 Paper_May_2012/main.tex
Paper: Kapitel 2 fertig gestellt zum Gegenlesen

@@
 \begin{lstlisting}
-MethodOfLines_StateSpaceTimeHandler<FE, DOFHANDLER, SPARSITYPATTERN, VECTOR, 2> DOFH(
-      triangulation, state_fe);
+  MethodOfLines_StateSpaceTimeHandler<FE, DOFHANDLER, 
+    SPARSITYPATTERN, VECTOR, 2> DOFH(triangulation, state_fe);
 \end{lstlisting}
@@
   out.Write(outp, 1, 1, 1);
-
   solver.ComputeReducedFunctionals();
@@
 \begin{lstlisting}
-    void CellEquation(const CellDataContainer<dealii::DoFHandler<dealdim>, VECTOR, dealdim>& cdc,
-                      dealii::Vector<double> &local_cell_vector,
-                      double scale, double)
+   void CellEquation(...)
     {
@@
 \end{lstlisting}
-Specifically,
+Specifically, the inner loops reads
 \begin{lstlisting}
-      for (unsigned int q_point = 0; q_point < n_q_points; q_point++)
+   for (all quadrature points per cell q)
       {
-        Tensor<2, 2> ugrads;
-        ugrads.clear();
-        ugrads[0][0] = _ugrads[q_point][0][0];
-        ugrads[0][1] = _ugrads[q_point][0][1];
-        ugrads[1][0] = _ugrads[q_point][1][0];
-        ugrads[1][1] = _ugrads[q_point][1][1];
+        Tensor<2, 2> ugrads; // solution vector
 
-	for (unsigned int i = 0; i < n_dofs_per_cell; i++)
+	for (all dofs per cell i)
         {
-	  const Tensor<2, 2> phi_i_grads_u = state_fe_values[displacements].gradient(i, q_point);
+	  const Tensor<2, 2> phi_i_grads_u; // test function
 
-          local_cell_vector(i) += scale * scalar_product(ugrads, phi_i_grads_u) 
-	    * state_fe_values.JxW(q_point);
+          local_cell_vector(i) += scale * 
+            scalar_product(ugrads, phi_i_grads_u) *
+	    state_fe_values.JxW(q_point);
         }
@@
 \begin{lstlisting}
-    void CellMatrix(const CellDataContainer<dealii::DoFHandler<dealdim>, VECTOR, dealdim>& cdc, FullMatrix<double> &local_entry_matrix, double, double)
+   void CellMatrix(...)
     {
@@
 \begin{lstlisting}
-      for (unsigned int q_point = 0; q_point < n_q_points; q_point++)
+      for (all quadrature points per cell q)      
+        for (all dofs per cell i)    
+          for (all dofs per cell j)
           {
-        for (unsigned int k = 0; k < n_dofs_per_cell; k++)
-        {
-          phi_grads_u[k] = state_fe_values[displacements].gradient(k, q_point);
-        }
-
-        for (unsigned int i = 0; i < n_dofs_per_cell; i++)
-        {
-          for (unsigned int j = 0; j < n_dofs_per_cell; j++)
-          {
-            local_entry_matrix(i, j) += scalar_product(phi_grads_u[j], phi_grads_u[i]) 
-                * state_fe_values.JxW(q_point);
-          }
-        }
+            local_entry_matrix(i, j) += 
+              scalar_product(phi_grads_u[j], phi_grads_u[i]) * 
+              state_fe_values.JxW(q_point);
           }
@@
 
-The cost functional is now implemented in \todo{hier morgen weiter machen}
+The target functional 
+is now implemented in \texttt{functional.h} and 
+treats a point evaluation:
+\begin{lstlisting}
+  double PointValue(...)
+   {
+    Point<2> p1(0.5, 0.5);
+    Vector<double> tmp_vector(2);
+    // more stuff is necessary here
+    VectorTools::point_value (...);
+    double x = tmp_vector(0);
+
+    return x;
+   }
+\end{lstlisting}
 
@@
 \begin{lstlisting}
- void CellTimeEquation(const CellDataContainer<dealii::DoFHandler<dealdim>, VECTOR, dealdim>& cdc,
-                           dealii::Vector<double> &local_cell_vector __attribute__((unused)),
-		       double scale __attribute__((unused)))
+  void CellTimeEquation(...)
    {
@@
 \begin{lstlisting}
- for (unsigned int q_point=0;q_point<n_q_points;q_point++)
+  for (all quadrature points per cell q)
    {
-	  Tensor<1,2> v;
-	  u[0] = _uvalues[q_point](0);
-	  u[1] = _uvalues[q_point](1);
+     Tensor<1,2> u = ...;
 
-	  for (unsigned int i=0;i<n_dofs_per_cell;i++)
+     for (all dofs per cell i)
       {
-	      const Tensor<1,2> phi_i_u = state_fe_values.value (i, q_point);
+        const Tensor<1,2> phi_i_u = ...;
 
-	      local_cell_vector(i) +=  scale * (u * phi_i_u) * state_fe_values.JxW(q_point);
+        local_cell_vector(i) +=  
+          scale * (u * phi_i_u) * 
+          state_fe_values.JxW(q_point);
       }
@@
 \begin{lstlisting}
- void CellTimeEquationExplicit (const CellDataContainer<dealii::DoFHandler<dealdim>, VECTOR, dealdim>& cdc,
-                           dealii::Vector<double> &local_cell_vector,
-				   double scale)
+  void CellTimeEquationExplicit (...)
    {
@@
 
-Concretely, let us focus on the Laplace equation and ...
+Concretely, let us focus on the Laplace equation and 
+the following optimization problem (taken from 
+Example 2) in \texttt{Examples/OPT/StatPDE}:
+\begin{align*}
+\min_{(q,u)\in \mathbb R^3 \times H_0^1(\Omega; \mathbb R^2)} J(q,u) &=
+\frac{1}{2} \sum_{i=0}^2 |(u-\overline u)(x_i)|^2 + \frac{\alpha}{2}\|q\|^2\\
+\text{s.t.} (\nabla u,\nabla \phi) &= (f(q),\phi)\;\;\forall\,\phi \in H^1_0(\Omega; \mathbb R^2)
+\end{align*}
+on the domain $\Omega = [0,1]^2$, with
+\begin{itemize}
+\item the observation points
+\begin{align*}
+x_0 = (0.5, 0.5), \quad x_1 = (0.5, 0.25),\quad x_2 = (0.25, 0.25),
+\end{align*}
+\item the regularization parameter $\alpha = 0$, 
+\item the right hand side
+\begin{align*}
+ f(q) &= q_0 \left(\begin{matrix}2\pi^2  \sin( \pi x) \sin(\pi y)\\0 \end{matrix}\right)\\
+      &+ q_1 \left(\begin{matrix}5\pi^2  \sin( \pi x) \sin(2\pi y)\\0 \end{matrix}\right)\\
+      &+ q_2 \left(\begin{matrix}0 \\8\pi^2  \sin(2\pi x) \sin(2\pi y)\end{matrix}\right)\\
+\end{align*}
+\item and the exact solution given by 
+\begin{align*}
+ \overline{q} &= \bigl(1;0.5;1\bigr)\\
+ \overline{u}& = \left(\begin{matrix} \sin( \pi x)( \sin(\pi y)+0.5\sin(2\pi y))\\\sin(2\pi x) \sin(2\pi y) \end{matrix}\right).
+\end{align*}
+\end{itemize}
+\begin{defi}[Constrained optimization]
+\label{problem_constraint_optimization}
+Minimize the cost functional $J(q,U)$ subject to the state equation
+$A(q,U)(\Phi) = 0$ (as defined
+in~\eqref{state_form} and~\eqref{opt_state_equation}) for $(q,U) \in 
+Q\times  \{v_D+{\cal X}\}$. 
+\end{defi}
+%
+The constrained optimization problem  on the space $Q_d\times {\cal X}$ 
+is reformulated into an unconstrained optimization problem on the
+space $Q_d$. Therefore, we assume the existence of the solution operator  
+$S:Q_d\rightarrow {v_D+{\cal X}}$ with a unique solution
+$U=S(q)$. Herewith, we define the reduced cost functional
+$j:Q_d\rightarrow \mathbb{R}$ by 
+%
+\begin{equation}
+  \label{opt_reduced_functional}
+  j(q) := J(q,S(q)) .
+\end{equation}
+%
+Newton's method to solve Problem~\ref{opt_reduced_problem} reads:
+For $l=0,1,\ldots,$ solve
+\begin{equation}
+\label{opt_newton_method}
+\begin{aligned}
+j''(q^l)(\delta q , \tau q) &= -j'(q^l)(\delta q) 
+\quad\forall\tau q\in Q_d, \\
+q^{l+1} &= q^l + \omega\delta q,
+\end{aligned} 
+\end{equation}
+with a line search parameter $\omega\in (0,1]$ which 
+will be specified in Algorithm \ref{algo_opt}.
+Specifically, the residual and the Hessian of Newton's method 
+\eqref{opt_newton_method}
+can be computed with the help of the following two results:
+\begin{proposition}[Residual of Newton's method]
+\label{prop_newton_residual}
+Let $q\in Q_d, U=S(q)\in {\cal X}$
+and $Z\in {\cal X}$ the dual solution be obtained after solving 
+the first and second equation of the KKT system 
+\eqref{opt_KKT_system}. 
+Then the residual of Newton's method \eqref{opt_newton_method} 
+is defined as
+\begin{equation*}
+j'(q)(\delta q) := {\cal L}'_{q}(q,U,Z)(\delta q),
+\end{equation*}
+i.e., in explicit representation
+\begin{equation*}
+j'(q)(\delta q) := \alpha_T (q , \tau q)_Q - 
+A_{q}'(q,U)(\tau q , Z).
+\end{equation*}
+\end{proposition}
+\begin{proof}
+The proof uses standard techniques. Details can be found in  
+\cite{BeMeVe06}.
+\end{proof}
+\begin{proposition}[Hessian of Newton's method]
+  \label{prop_newton_hessian}
+  Let $q\in Q_d, U=S(q)\in {\cal X}$ and $Z\in {\cal X}$ and  $\delta
+  q\in Q_d$ be given. Further, let $\delta U\in {\cal X}$ be the
+  solution of the tangent problem 
+  \[
+  \delta U\in{\cal X},\quad 
+  A_{U}'(q,U)(\delta U , \Phi) 
+  = -A_{q}'(q,U)(\delta q , \Phi)  \quad\forall \Phi\in {\cal
+    X},
+  \]
+  and $\delta Z\in {\cal X}$ the solution of the adjoint Hessian 
+  problem  
+  \[
+  A_{U}'(q,U)(\Phi, \delta Z)
+  = J''_{UU}(q,U)(\delta U, \Phi)
+  - A_{UU}''(q,U)(\delta U, \Phi, Z)
+  - A_{qU}''(q,U)(\delta q , \Phi, Z)  \quad\forall \Phi\in
+  {\cal X}. 
+  \]
+ Then it holds  for all $\tau q\in Q_d$:
+  \[
+  j''(q)(\delta q, \tau q):= 
+  \alpha_T (\delta q , \tau q)_Q -
+  A_{qq}''(q, U)(\delta q, \tau q , Z) -
+  A_{Uq}''(q,U)(\delta U, \tau q , Z) -
+  A_{q}'(q,U)(\tau q , \delta Z).
+  \]
+\end{proposition}
+\begin{proof}
+  For the proof this statement (including its time-dependent version), we refer to 
+  Becker et al. \cite{BeMeVe06}.
+  The stationary version of this proposition is easily derived by neglecting 
+  all time derivatives, which concludes the assertion. 
+\end{proof}
+The implementation of all previous terms 
+is realized in the \texttt{localpde.h} file.
+As an example, we examine the adjoint 
+\begin{equation}
+  \label{opt_KKT_system}
+  \begin{aligned}
+    A_{U}'(q,U)(\Phi , Z) &= J'_{U}(q,U)(\Phi) & \forall \Phi&\in {\cal X},
+  \end{aligned}
+\end{equation}
 
-this is taken from step-2 in the Opt examples.
+The implementation it split into equation terms, i.e., 
+\begin{lstlisting} 
+  void CellEquation_U(...)
+   {
+     // implement (\nabla \phi, \nabla z)
+   }
+\end{lstlisting}
+and functional terms, i.e.,     
+\begin{lstlisting} 
+  virtual void PointValue_U(...)
+      {
+        // implement here the derivative of 
+        // the point cost functional
+      }
+\end{lstlisting}
+All remaining terms of the previous two propositions 
+are implemented in the same manner.
 
-\todo{Optimality system}
-
-\todo{Two problems in which we identify the different terms and implementation 
-of these terms in DOpE}
+Afterwards, it remains to use an appropriate 
+optimization algorithm in order to solve the 
+problem. At present we offer a selection of algorithms that solve the reduced optimization 
+problem where the PDE constraint has been eliminated as explained previously.
+\begin{itemize}
+\item \texttt{opt\underline{ }algorithms/reducedalgorithm.h} An interface for all 
+  optimization problems in the reduced formulation. It offers some test functionality
+  to assert that the derivatives of the problem are computed correctly.
+\item \texttt{opt\underline{ }algorithms/reducednewtonalgorithm.h}
+  A line-search Newton algorithm using a cg method to invert the reduced hessian. 
+  Implementation ignores any additional constraints.
+\item \texttt{opt\underline{ }algorithms/reducedtrustregionnewton.h}
+  A trust region Newton algorithm using a cg method to invert the reduced hessian.
+  Implementation ignores any additional constraints.
+\item \texttt{opt\underline{ }algorithms/reduced\underline{ }snopt\underline{ }algorithm.h}
+  An algorithm to solve reduced optimization problems with additional control constraints.
+  ((reduced) state constraints are not yet implemented.)
+\item \texttt{opt\underline{ }algorithms/reducednewtonalgorithmwithinverse.h}
+  Line-search Newton algorithm that assumes there exists a method in the reduced problem
+  that can invert the reduced hessian. (This usually makes sense only if there is no 
+  PDE constraint.)
+\item \texttt{opt\underline{ }algorithms/generalized\underline{ }mma\underline{ }algorithm.h}
+  An implementation of the MMA-Algorithm for structural optimization using an augmented
+  Lagrangian formulation for the subproblems. The subproblem is implemented using the 
+  special purpose file
+  \texttt{include/augmentedlagrangianproblem.h}.
+\end{itemize}
 
@@
 %
-%\subsection{Optimization algorithms}
-%Now, in order to solve optimization algorithms we need to define some algorithms.
-%At present we offer a selection of algorithms that solve the reduced optimization 
-%problem where the PDE constraint has been eliminated as explained in the previous section.
-%\begin{itemize}
-%\item \texttt{opt\underline{ }algorithms/reducedalgorithm.h} An interface for all 
-%  optimization problems in the reduced formulation. It offers some test functionality
-%  to assert that the derivatives of the problem are computed correctly.
-%\item \texttt{opt\underline{ }algorithms/reducednewtonalgorithm.h}
-%  A line-search Newton algorithm using a cg method to invert the reduced hessian. 
-%  Implementation ignores any additional constraints.
-%\item \texttt{opt\underline{ }algorithms/reducedtrustregionnewton.h}
-%  A trust region Newton algorithm using a cg method to invert the reduced hessian.
-%  Implementation ignores any additional constraints.
-%\item \texttt{opt\underline{ }algorithms/reduced\underline{ }snopt\underline{ }algorithm.h}
-%  An algorithm to solve reduced optimization problems with additional control constraints.
-%  ((reduced) state constraints are not yet implemented.)
-%\item \texttt{opt\underline{ }algorithms/reducednewtonalgorithmwithinverse.h}
-%  Line-search Newton algorithm that assumes there exists a method in the reduced problem
-%  that can invert the reduced hessian. (This usually makes sense only if there is no 
-%  PDE constraint.)
-%\item \texttt{opt\underline{ }algorithms/generalized\underline{ }mma\underline{ }algorithm.h}
-%  An implementation of the MMA-Algorithm for structural optimization using an augmented
-%  Lagrangian formulation for the subproblems. The subproblem is implemented using the 
-%  special purpose file
-%  \texttt{include/augmentedlagrangianproblem.h}.
-%\end{itemize} 
+
 %
@@
 \end{lstlisting}
-and then using the designated scheme.
+and then using in the code the designated scheme, 
+for example here the Fractional-Step-$\theta$ scheme
+\texttt{TSP5}.
 

### 2012-10-05 Paper_May_2012/main.tex
Paper: Anmerkungen von Winni eingearbeitet sowie Referenzen verlinkt und nicht referenzierte labels entfernt

@@
 In the \texttt{main} function, the following steps need to be done. First, 
-the parameter reader for runtime parameters is initialized:
-\begin{lstlisting}
-  ParameterReader pr;
-  SSolver::declare_params(pr);
-  DOpEOutputHandler<VECTOR>::declare_params(pr);
-  pr.read_parameters(paramfile);
-\end{lstlisting}
+the parameter reader for runtime parameters might be initialized. In a parameter
+file the user might specifiy material paramters for the concrete PDE, but also
+general parameters for nonlinear and linear solvers, and output-related 
+aspects.
+
 A linear finite element in two dimensions and two components is 
@@
 \begin{lstlisting}
-  MethodOfLines_StateSpaceTimeHandler<FE, DOFHANDLER, 
-    SPARSITYPATTERN, VECTOR, 2> DOFH(triangulation, state_fe);
+  MethodOfLines_StateSpaceTimeHandler<...> DOFH (...);
 \end{lstlisting}
@@
 \end{lstlisting}
-After explaining the basic compononents that are necessary for all  
-(linear and stationary) partial differential equations, we shall give 
+After explaining the basic compononents that are necessary (and always the same in DOpElib) 
+for all  (linear and stationary) partial differential equations, we shall give 
 an overview to the implementation of the specific form of the PDE. 
+Beside specification of boundary conditions, the form of the local PDE depends 
+on the concrete problem under consideration. 
 This step is done in the \texttt{localpde.h} file. As already explained,
@@
 cell is then done 
-in the \texttt{templates/integrator.h} function.
+in the \texttt{templates/integrator.h} function. Finally, as usually in 
+finite element computations, this leads to a discrete system. Since
+we treat each problem with a Newton method, within each Newton step, the 
+following linear system must be solved:
+\[
+A\delta u = b
+\]
+in which $A$ is specified by \texttt{CellMatrix(...)} and 
+$b$ by \texttt{CellEquation(...)}. The vector $\delta u$ denotes the solution to the 
+present Newton step:
+\[
+u^n = u^{n-1} + \omega \delta u
+\]
+with a line search parameter $\omega\in (0,1]$.
 
@@
 great difference to the previous explications. However, 
-since we utilize gradient-based optimization routines \cite{...},
+since we utilize gradient-based optimization routines in the philosophy of
+\cite{Troe99,HiKu01},
 all derivatives must be implemented. All this information
@@
 \label{problem_constraint_optimization}
-Minimize the cost functional $J(q,U)$ subject to the state equation
-$A(q,U)(\Phi) = 0$ (as defined
-in~\eqref{state_form} and~\eqref{opt_state_equation}) for $(q,U) \in 
+Minimize the cost functional $J(q,u)$ subject to the state equation
+$a(q,u)(\phi) = 0$ for $(q,u) \in 
 Q\times  \{v_D+{\cal X}\}$. 
@@
 %
+Thus, the constrained optimization problem can be formulated by means
+of  
+%
+\begin{defi}[Unconstrained optimization]
+  \label{opt_reduced_problem}
+  Minimize $j(q)$ for $q\in Q_d$.
+\end{defi}
+%
 Newton's method to solve Problem~\ref{opt_reduced_problem} reads:
@@
 with a line search parameter $\omega\in (0,1]$ which 
-will be specified in Algorithm \ref{algo_opt}.
+will are implemented in \texttt{reducednewtonalgorithm.h}.
 Specifically, the residual and the Hessian of Newton's method 
@@
 \label{prop_newton_residual}
-Let $q\in Q_d, U=S(q)\in {\cal X}$
-and $Z\in {\cal X}$ the dual solution be obtained after solving 
+Let $q\in Q_d, u=S(q)\in {\cal X}$
+and $z\in {\cal X}$ the dual solution be obtained after solving 
 the first and second equation of the KKT system 
@@
 \begin{equation*}
-j'(q)(\delta q) := {\cal L}'_{q}(q,U,Z)(\delta q),
+j'(q)(\delta q) := {\cal L}'_{q}(q,u,z)(\delta q),
 \end{equation*}
@@
 j'(q)(\delta q) := \alpha_T (q , \tau q)_Q - 
-A_{q}'(q,U)(\tau q , Z).
+A_{q}'(q,u)(\tau q , z).
 \end{equation*}
@@
   \label{prop_newton_hessian}
-  Let $q\in Q_d, U=S(q)\in {\cal X}$ and $Z\in {\cal X}$ and  $\delta
-  q\in Q_d$ be given. Further, let $\delta U\in {\cal X}$ be the
+  Let $q\in Q_d, u=S(q)\in {\cal X}$ and $z\in {\cal X}$ and  $\delta
+  q\in Q_d$ be given. Further, let $\delta u\in {\cal X}$ be the
   solution of the tangent problem 
   \[
-  \delta U\in{\cal X},\quad 
-  A_{U}'(q,U)(\delta U , \Phi) 
-  = -A_{q}'(q,U)(\delta q , \Phi)  \quad\forall \Phi\in {\cal
+  \delta u\in{\cal X},\quad 
+  a_{u}'(q,u)(\delta u , \phi) 
+  = -a_{q}'(q,u)(\delta q , \phi)  \quad\forall \phi\in {\cal
     X},
   \]
-  and $\delta Z\in {\cal X}$ the solution of the adjoint Hessian 
+  and $\delta z\in {\cal X}$ the solution of the adjoint Hessian 
   problem  
   \[
-  A_{U}'(q,U)(\Phi, \delta Z)
-  = J''_{UU}(q,U)(\delta U, \Phi)
-  - A_{UU}''(q,U)(\delta U, \Phi, Z)
-  - A_{qU}''(q,U)(\delta q , \Phi, Z)  \quad\forall \Phi\in
+  a_{u}'(q,u)(\phi, \delta z)
+  = J''_{uu}(q,u)(\delta u, \phi)
+  - a_{uu}''(q,u)(\delta u, \phi, z)
+  - a_{qu}''(q,u)(\delta q , \phi, z)  \quad\forall \phi\in
   {\cal X}. 
@@
   \alpha_T (\delta q , \tau q)_Q -
-  A_{qq}''(q, U)(\delta q, \tau q , Z) -
-  A_{Uq}''(q,U)(\delta U, \tau q , Z) -
-  A_{q}'(q,U)(\tau q , \delta Z).
+  a_{qq}''(q, u)(\delta q, \tau q , z) -
+  a_{uq}''(q,u)(\delta u, \tau q , z) -
+  a_{q}'(q,u)(\tau q , \delta z).
   \]
@@
   \begin{aligned}
-    A_{U}'(q,U)(\Phi , Z) &= J'_{U}(q,U)(\Phi) & \forall \Phi&\in {\cal X},
+    a_{u}'(q,u)(\phi , z) &= J'_{u}(q,u)(\phi) & \forall \phi&\in {\cal X},
   \end{aligned}
@@
 The configuration is 
-sketched in Figure \ref{configuration_csm_and_fsi_2D}. Details 
+sketched in Figure \ref{configuration_csm_and_fsi_2D}. 
+The computational domain has length $L=2.5m$ and height $H=0.41m$. The circle center
+is positioned at $C=(0.2m,0.2m)$ with radius $r=0.05m$. The elastic beam has length
+$l=0.35m$ and height $h=0.02m$. The right lower end is positioned at 
+$(0.6m,0.19m)$, and
+the left end is attached to the circle.
+
+\begin{figure}[h]
+\centering
+\input{Pictures/pic_12}
+\caption{Flow around cylinder with elastic beam with 
+circle-center $C=(0.2,0.2)$ and radius $r=0.05$.}
+\label{configuration_csm_and_fsi_2D}
+\end{figure}
+
+Control points $A(t)$ (with $A(0) = (0.6,0.2)$) are fixed at the 
+trailing edge of the structure, measuring $x$- and $y$-deflections of the beam.
+
+Details 
 on parameters and evaluation functionals and other results 
-can be found in \cite{HrTu06b, BuSc06, TuHrMaRaWoAc10, DeHaeAnnBrVie10, Wi11}. 
+can be found in \cite{HrTu06b,BuSc06,DeHaeAnnBrVie10,Wi11}. 
 The Fractional-Step-$\theta$ scheme is used for time discretization with

### 2012-10-11 Examples/PDE/StatPDE/Example10/content.tex
Test of higher order mappings.

\subsubsection{General problem description}
This example shows the use higher order mappings in DOpE. To this end, we solve a simple laplace equation
\begin{align*}
-\Delta u &= -4 && \text{ in } \Omega
\end{align*}
with the analytical solution
\begin{align*}
u = {x^2 + y^2},
\end{align*}
and the Dirichlet Conditions on $u=1$, where the domain is given by a circle with radius 1 and the center located in the origin.

We compute the $L^2$-norm of the error and, additionally, we evaluate a functional which does not depend on the solution at all.
We integrate the constant $\frac 12$ once over the boundary of the domain. The result is an approximation of $\pi$.

All this is standard and would not justify an additional example, however we solve the equation and the functional not one but two
times. First with the standard Q1-mapping, the second time we use a higher order mapping. The exact order can be determined by the 
parameter file, the preset is Q2-mapping. At the end, we gather the errors and convergence rates over some refinement cycles
in a nice table and notice the higher order of convergence for the higher order mapping solutions. This is due to the fact that
we can approximate the circular boundary much better by a quadratic mapping.
\subsubsection{Program description}
In this section we want to focus on what you have to do if you want to enhance your existing code to use higher order mappings,
which is actually pretty simple.

You have to include the file
\begin{verbatim}
mapping_wrapper.h
\end{verbatim}

and create a mapping of the desired order by
\begin{verbatim}
  DOpEWrapper::Mapping<dim, DOFHANDLER> mapping(order_of_mapping);
\end{verbatim}

The last step is to give the mapping to the DoFHandler:
\begin{verbatim}
MethodOfLines_StateSpaceTimeHandler<FE, DOFHANDLER, SPARSITYPATTERN,VECTOR, 2> 
	  DOFH(triangulation, mapping, state_fe);
\end{verbatim}
The rest of the program is as usual.

### 2012-10-12 Paper_May_2012/main.tex
Wie gestern besprochen, hier die Einleitung zum Kapitel wie man StatPDE loest.

@@
 \renewcommand{\epsilon}{\varepsilon}
+\newcommand{\deal}{\texttt{deal.II}}
+\newcommand{\dope}{\texttt{DOpElib}}
 
@@
 In this article, we describe the software library 
-\textit{Deal.II Optimization Environment} (DOpElib).
-The main feature of DOpElib is that it provides a unified interface to high level algorithms 
+\textit{Deal.II Optimization Environment} (\dope).
+The main feature of \dope is that it provides a unified interface to high level algorithms 
 such as time stepping methods, nonlinear solvers and optimization routines. This structure ensures 
@@
 \label{introduction}
-The \textit{Deal Optimization Environment} (DOpElib) 
+The \textit{Deal Optimization Environment} (\dope) 
 provides a software toolkit to solve forward PDE
@@
 solution of a broad variety of PDE is possible in other software
-libraries (like in deal.II \cite{deal}, dune \cite{dune} , 
+libraries (like in \deal \cite{deal}, dune \cite{dune} , 
 or commercial solvers like Ansys \cite{ansys}) 
 as well, but
-DOpE concentrates on a unified approach for both linear and nonlinear
+\dope concentrates on a unified approach for both linear and nonlinear
 problems by interpreting every PDE problem as nonlinear and applying a
-Newton method to solve it. While deal.II leaves much of the work and many
-decisions to the user, DOpE intends to be user-optimized by delivering
+Newton method to solve it. While \deal leaves much of the work and many
+decisions to the user, \dope intends to be user-optimized by delivering
 prefabricated tools which require from the user only adjustments connected
 to his specific problem. The solution of optimal control problems with PDE
-constraints is an innovation in the DOpE framework.
+constraints is an innovation in the \dope framework.
 The focus is on the numerical solution of both stationary and nonstationary
@@
 
-The DOpElib project is 
-based on the \textit{deal.II} \cite{deal} finite element library which has been developed
+The \dope project is 
+based on the \deal \cite{deal} finite element library which has been developed
  initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{deal}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%The aim of DOpE is to provide a software toolkit to solve forward PDE
+%The aim of \dope is to provide a software toolkit to solve forward PDE
 %problems as well as optimal control problems constrained by PDE. The
-%solution of a broad variety of PDE is possible in deal.II as well, but
-%DOpE concentrates on a unified approach for both linear and nonlinear
+%solution of a broad variety of PDE is possible in \deal as well, but
+%\dope concentrates on a unified approach for both linear and nonlinear
 %problems by interpreting every PDE problem as nonlinear and applying a
-%Newton method to solve it. While deal.II leaves much of the work and many
-%decisions to the user, DOpE intends to be user-optimized by delivering
+%Newton method to solve it. While \deal leaves much of the work and many
+%decisions to the user, \dope intends to be user-optimized by delivering
 %prefabricated tools which require from the user only adjustments connected
 %to his specific problem. The solution of optimal control problems with PDE
-%constraints is an innovation in the DOpE framework.
+%constraints is an innovation in the \dope framework.
 %The focus is on the numerical solution of both stationary and nonstationary
@@
   Crank-Nicolson, shifted Crank-Nicolson, and Fractional-Step-$\Theta$ scheme.
-\item All finite elements of from deal.II including hp-support.
+\item All finite elements of from \deal including hp-support.
 \item Several examples showing the solution of several PDEs including
@@
 %The rest of this document is structured as follows: We start with an introduction in
-%Chapter~\ref{chap:intro} where you will learn what is needed to run {\tt DOpElib}. 
+%Chapter~\ref{chap:intro} where you will learn what is needed to run {\tt \dope}. 
 %Further you will learn what problems we can solve and how all the different classes 
@@
 of problems involving partial differential equations (PDEs). 
+In particular, we remove all tedious routine implementation needed in 
+\deal by introducing appropriate classes that take care of all the 
+problem independent implementation while leaving only the problem specific 
+part to the user. To show how this is designed we will, in the sequel, 
+describe the reasoning first for a simple stationary PDE and then show
+how this idea extends to non stationary problems as well as to optimization
+problems with PDE constraints.
+
 \subsection{Solving of PDEs}
 \subsubsection{Stationary problems}
-The easiest case 
-is that of a PDE in weak form to find some $u$
-\[
-a(u)(\phi) = 0 \quad \forall \phi \in V,
-\]
-with some appropriate space $V$. As example, we take 
-the Laplace problem in weak form:
-\[
-a(u)(\phi) = (\nabla u, \nabla \phi) \quad \forall \phi \in V.
-\]
+For the start we consider a prototypical problem given by the solution of 
+a stationary PDE, which we assume to be given in weak form.
+In abstract notation it reads; find some $u$ in an appropriate set 
+such that
+\[
+a(u)(\phi) = f(\phi) \quad \forall \phi \in V,
+\]
+with some appropriate space $V$ for the test functions and a 
+right hand side $f$.
+To illustrate this further, we take the stationary 
+incompressible Navier-Stokes equations in a domain $\Omega \subset \R^d$, 
+i.e., $u = (v,p)$ and $\phi = (\phi_u,\phi_p)$ then
+\[
+a(u)(\phi) = (\nabla v, \nabla\phi_u) + (v \nabla v,\phi_u) + (p, \nabla \cdot \phi_u) + (\nabla \cdot v ,\phi_p)
+\quad \forall \phi \in V = H^1_0(\Omega;\R^d) \times L^2_0(\Omega)
+\]
+Then the solution will be searched for in $V + (u_d,0)$ where $u_d$ is some 
+continuation of the given Dirichlet boundary data.
+Then typically we do not only want to calculate $u$ but more likely we are 
+interested in one (or several) numbers $J(u)$.
+
+In order to solve this problem, i.e., calculate $J(u)$, we need to discretize 
+the equation, here using finite elements. This means we need to pick a 
+subdivision $\mathcal T_h$, for instance into quadrilaterals,
+of the domain $\Omega$ and a finite element, in the concrete case we could, 
+for instance, take the $\mathcal Q_2/\mathcal Q_1$ Taylor-Hood element.
+As it is clear, that the initialization of the unknowns, the local shape 
+functions, removal of hanging node constraints, etc. always follows the 
+same procedure, we define an object which takes the finite element (FE)
+and the triangulation and takes care of all other things related with 
+handling the finite element space $V_h$. We will refer, to 
+this as the \texttt{SpaceTimeHandler}. It is then clear, that the simplest 
+form of the constructor should have the following interface:
+\begin{lstlisting}
+  template<typename Triangulation, typename FE>
+      void SpaceTimeHandler(Triangulation &, FE);
+\end{lstlisting}
+Next, one needs to solve the discrete equation
+\[
+a(u_h)(\phi_h) = f(\phi_h) \quad \forall \phi \in V_h.
+\]
+To this end one can employ a newton type method with some globalization 
+like line-search. To do so, one needs to calculate the residual of the 
+equation, as well as solving linearized equations for the calculation of 
+the newton method and then employing some kind of globalization strategy.
+This again will usually require the evaluation of the residual.
+
+To evaluate the residual will require the calculation of the integrals 
+involved in the weak formulation of the problem. For example 
+the $i$-th component of the residual, with respect to the given FE-basis 
+$\phi_i$, will read as follows
+\[
+ a(u_h)(\phi_i) - f(\phi_i) = \sum_{K\in \mathcal T_h} \int_K \nabla u, \nabla \phi_i\,dx + \ldots
+\]
+Again, we see that, except for the local element integrals, all 
+calculations are independent from the concrete form of the PDE. 
+Hence we define an object \texttt{Integrator}
+that offers a method to calculate the residual given a user-defined 
+object \texttt{PDE} that implements the local element integrals, i.e.,
+\begin{lstlisting}
+  template<typename PDE,typename VECTOR>
+      void Integrator
+             ::ComputeNonlinearResidual(PROBLEM& pde, 
+                                        VECTOR & residual);
+\end{lstlisting}
+where \texttt{VECTOR} is one of the possible vector types given by \deal.
+It is clear that we need to ensure that the integrator object has access
+to the information stored in the \texttt{SpaceTimeHandler}
+to correctly loop over all elements.
+To solve the linearized systems, again, all that is left is the calculation
+of the matrix corresponding to the linearized equation. To calculate this,
+again we can define a method in the integrator that does exactly this provided
+we have a user defined object \texttt{PDE} that can evaluate the local 
+integrals needed for assembling the matrix, i.e.,
+\begin{lstlisting}
+  template<typename PDE,typename MATRIX>
+      void Integrator
+             ::ComputeMatrix(PROBLEM& pde, 
+                             MATRIX & system_matrix);
+\end{lstlisting}
+where now \texttt{MATRIX} is one of the matrix types supported by \deal.
+With this we can now define a nonlinear solver \texttt{NewtonSolver} 
+for the solution of the discrete problem requiring the existence of the 
+following function
+\begin{lstlisting}
+  template<typename PDE, typename VECTOR>
+      void NewtonSolver
+             ::NonlinearSolve(PROBLEM& pde, 
+                              VECTOR & solution);
+\end{lstlisting}
+
+Finally, we note that to compute the requested functional value 
+all that is left is to evaluate the functional in the computed solution 
+which can be done again using an appropriate integrator method.
+Since the sequence of events sketched above is again independent of the 
+exact problem we introduce another container around it which we call 
+\texttt{StatPDEProblem} that then simply provides the user with the 
+method
+\begin{lstlisting}
+      void StatPDEProblem::ComputeReducedFunctionals();
+\end{lstlisting}
+That combines all the necessary steps to calculate $J(u_h)$.
+
+\begin{remark}
+In fact, the above constructors are not exactly what is implemented in \dope.
+This is due to the fact, that for instance 
+there are several different sparsity patterns available in 
+\deal which the user should be able chose. Thus it needs to be 
+given to the \texttt{SpaceTimeHandler}. However, to keep the 
+presentation clear we refrained from including these details above and in
+what follows.
+\end{remark}
+
+In the next section we will now sketch the way a user needs to implement the
+user-dependent data as the local cell integrals and the initialization of the
+objects described above.
+\subsubsection{Problem dependent implementation of stationary PDE problems}
+
 To solve the problem, we need a configuration (a mesh), eventually 
@@
 Each of those routines are written in the \texttt{main.cc} file. 
-We start by including necessary files both the DOpElib library and
-the deal.II software:
+We start by including necessary files both the \dope library and
+the \deal software:
 \begin{lstlisting}
@@
 the parameter reader for runtime parameters might be initialized. In a parameter
-file the user might specifiy material paramters for the concrete PDE, but also
+file the user might specify material parameters for the concrete PDE, but also
 general parameters for nonlinear and linear solvers, and output-related 
@@
 Then, the \texttt{LocalPDE} object comes into play. This is an quite 
-important feature of DOpElib because it allows for easy implementation 
+important feature of \dope because it allows for easy implementation 
 of the variational formulation. Specific explication is given below because
-the implementation is seperated in \texttt{localpde.h}.
+the implementation is separated in \texttt{localpde.h}.
 Afterwards, a \texttt{LocalPointFunctional} is declared. As for the PDE, the 
@@
 
-Generically, each problem in DOpElib is considered to be time-dependent and nonlinear because
+Generically, each problem in \dope is considered to be time-dependent and nonlinear because
 these are the 'real' world problems. Therefore, (even for the linear and stationary Laplace equation),
@@
 \end{lstlisting}
-Finally, the ouput handler is initialized via
+Finally, the output handler is initialized via
 \begin{lstlisting}
@@
 \end{lstlisting}
-After explaining the basic compononents that are necessary (and always the same in DOpElib) 
+After explaining the basic compononents that are necessary (and always the same in \dope) 
 for all  (linear and stationary) partial differential equations, we shall give 
@@
 \]
-In DOpElib, different time stepping schemes can be found 
+In \dope, different time stepping schemes can be found 
 in the folder \texttt{DOpEsrc/tsschemes}.
@@
 \end{lstlisting}
-Here, the advantage is that DOpElib orders the terms in the timestepping
+Here, the advantage is that \dope orders the terms in the timestepping
 problems
@@
 that are applicable (in general) to all types of problems. Specifically,
-DOpElib provides goal-oriented mesh refinement with the help of the 
+\dope provides goal-oriented mesh refinement with the help of the 
 Dual-Weighted-Residual (DWR) method \cite{BeRa96}. Here, a certain 
@@
 
-Moreover, DOpElib provides standard error estimators based on residual evaluation and 
+Moreover, \dope provides standard error estimators based on residual evaluation and 
 a ZZ error estimator, i.e., measureing the smoothness of the discrete first derivatives.
-The latter one is accessed via deal.II.
-
-In addition, DOpElib provides multi-mesh procedures 
+The latter one is accessed via \deal.
+
+In addition, \dope provides multi-mesh procedures 
 \todo{kann Winni hier etwas schreiben?}
@@
 %\subsection{Interfaces to other software packages}
-%Since DOpE uses as basis the software library deal.II, all its interfaces
+%Since DOpE uses as basis the software library \deal, all its interfaces
 %to other packages can be accessed to. For instance, such libararies 
@@
 %%\label{documentation}
-%%At the present step, the DOpElib project comes with 
+%%At the present step, the \dope project comes with 
 %%an detailed documentation of all features and examples 
@@
 %%
-%%In addition, the DOpElib test suite provides some regression tests. 
+%%In addition, the \dope test suite provides some regression tests. 
 %%They are run to compare 
@@
 We present three numerical tests which demonstrate different
-features of DOpElib:
+features of \dope:
 \begin{itemize}
@@
 with the literature, it can be identified that 
-the implementation in DOpElib is correct.
+the implementation in \dope is correct.
 
@@
 In this article, we described the features 
-and applications of the DOpElib project, which 
+and applications of the \dope project, which 
 was initiated at the Heidelberg University in 2010.

### 2012-10-15 Paper_May_2012/main.tex
Paper: Nonstationary Problem Beschreibung angepasst und auf Wunsch von Christian die Funktionalauswertung in dem FSI Beispiel hinzugefuegt, so dass klar ist, was die vier Graphen aussagen

@@
 \subsubsection{Nonstationary problems}
-In this section, we focus on the extension to 
-the numerical solution of nonstationary processes. 
-As example, we consider the heat equation:
-Find $u$:
+In this section, the 
+numerical solution of nonstationary processes
+is considered. We focus on details which are 
+beyond stationary examples as discussed in the 
+previous sections. Before, we introduce 
+the steps of implementation in DOpElib, we 
+derive the basic theoretical concepts.
+As an example, we consider the heat equation:
+Find $u\in V$:
 \[
@@
 and in weak formulation:
-Find $u$:
+Find $u\in V$:
 \[
@@
 \]
-In \dope, different time stepping schemes can be found 
-in the folder \texttt{DOpEsrc/tsschemes}.
-
-Because the \texttt{main} file 
-is basically the same as for stationary problems
-which is one major goal of the software library,
-we only focus on the implementation of the PDE in the 
-\texttt{localpde.h} file. Here, the stationary part, i.e.,
-the diffusion is implemented as before. The only thing we add 
-is the time-depdendent part $\partial_t u$:
-
+The concrete choice of $\theta$ leads to a
+specific time stepping scheme. For instance,
+$\theta = 0.5$ gives the Crank-Nicolson scheme,
+\begin{lstlisting}
+   CrankNicolsonProblem<...>
+\end{lstlisting}
+The main work is treated as for stationary problems
+within 
+\begin{lstlisting}
+   MethodOfLines_SpaceTimeHandler<...>
+\end{lstlisting}
+Again, the key issue is that this procedure is totally independent 
+of the specific form of the PDE. The user-related work 
+remains to provide the time derivative $\partial_t u$ in
+the implementation:
 \begin{lstlisting}
@@
 \end{lstlisting}
-For instance, this explicit term is used for fluid-structure interaction in
-which
-the old timestep term needs explicit modification.
 
@@
 
+The quantities of interest are evaluations of 
+$x$- and $y$ displacement at the point $A(0) = (0.6,0.2)$
+and the drag and lift forces acting on the cylinder and the elastic beam:
+\begin{align}
+\label{drag_lift_forces}
+(F_D , F_L) 
+= {\int_{S_f} \sigma_f \cdot n_f \, ds + 
+\int_{\Gamma_i} \sigma_s \cdot n_s \, ds},
+\end{align}
+where $S_{f}$ denotes the path over the cylinder in the fluid part and
+$\Gamma_i$ the interface between the elastic beam and the 
+fluid.
+
 Since the problem is fully nonstationary, the 
@@
 \caption{FSI 2: the deflections of the beam, $u_x(A)$ and $u_y(A)$ (in $cm$), and 
-the drag and the lift evaluation (in $kg/m\,s^2$) are displayed versus
+the drag $F_D$ and the lift $F_L$ evaluation (in $kg/m\,s^2$) are displayed versus
 time (in $s$).

### 2012-10-18 Paper_May_2012/main.tex
Habe meinen Part groesstenteils fertig, wollte nun erstmal Meinungen einholen, glattbuegeln kann ich dann immer noch. Habe auch ein Paar Sachen bei Winni geaendert. Anstonste: 1) Ist der Fluidbenchmark als Beispiel geeignet, oder ist das dann zu kompliziert? 2) Inwiefern soll/darf/muss ich vom tatsaechlichen Code abweichen? Ich bin dafuer, template-Argumente oder nicht benoetigte Parameter zu streichen, nur sollten wir am Ende darauf achten, dass wir das im ganzen Dokument einheitlich haben. Z.B.: DirichledData oder Integratordatacontainer -- den letzteren habe ich shcon weggelassen, wuerde ci

@@
 \newcommand{\dope}{\texttt{DOpElib}}
+\newcommand{\re}{\operatorname{Re}}
 
@@
 In this article, we describe the software library 
-\textit{Deal.II Optimization Environment} (\dope).
-The main feature of \dope is that it provides a unified interface to high level algorithms 
+\textit{Deal.II Optimization Environment} (\dope{}).
+The main feature of \dope{} is that it provides a unified interface to high level algorithms 
 such as time stepping methods, nonlinear solvers and optimization routines. This structure ensures 
@@
 \label{introduction}
-The \textit{Deal Optimization Environment} (\dope) 
+The \textit{Deal Optimization Environment} (\dope{}) 
 provides a software toolkit to solve forward PDE
@@
 solution of a broad variety of PDE is possible in other software
-libraries (like in \deal \cite{deal}, dune \cite{dune} , 
+libraries (like in \deal{} \cite{deal}, dune \cite{dune} , 
 or commercial solvers like Ansys \cite{ansys}) 
 as well, but
-\dope concentrates on a unified approach for both linear and nonlinear
+\dope{} concentrates on a unified approach for both linear and nonlinear
 problems by interpreting every PDE problem as nonlinear and applying a
-Newton method to solve it. While \deal leaves much of the work and many
-decisions to the user, \dope intends to be user-optimized by delivering
+Newton method to solve it. While \deal{} leaves much of the work and many
+decisions to the user, \dope{} intends to be user-optimized by delivering
 prefabricated tools which require from the user only adjustments connected
 to his specific problem. The solution of optimal control problems with PDE
-constraints is an innovation in the \dope framework.
+constraints is an innovation in the \dope{} framework.
 The focus is on the numerical solution of both stationary and nonstationary
@@
 
-The \dope project is 
-based on the \deal \cite{deal} finite element library which has been developed
+The \dope{} project is 
+based on the \deal{} \cite{deal} finite element library which has been developed
  initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{deal}.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%The aim of \dope is to provide a software toolkit to solve forward PDE
+%The aim of \dope{} is to provide a software toolkit to solve forward PDE
 %problems as well as optimal control problems constrained by PDE. The
-%solution of a broad variety of PDE is possible in \deal as well, but
-%\dope concentrates on a unified approach for both linear and nonlinear
+%solution of a broad variety of PDE is possible in \deal{} as well, but
+%\dope{} concentrates on a unified approach for both linear and nonlinear
 %problems by interpreting every PDE problem as nonlinear and applying a
-%Newton method to solve it. While \deal leaves much of the work and many
-%decisions to the user, \dope intends to be user-optimized by delivering
+%Newton method to solve it. While \deal{} leaves much of the work and many
+%decisions to the user, \dope{} intends to be user-optimized by delivering
 %prefabricated tools which require from the user only adjustments connected
 %to his specific problem. The solution of optimal control problems with PDE
-%constraints is an innovation in the \dope framework.
+%constraints is an innovation in the \dope{} framework.
 %The focus is on the numerical solution of both stationary and nonstationary
@@
   Crank-Nicolson, shifted Crank-Nicolson, and Fractional-Step-$\Theta$ scheme.
-\item All finite elements of from \deal including hp-support.
+\item All finite elements of from \deal{} including hp-support.
 \item Several examples showing the solution of several PDEs including
@@
 %The rest of this document is structured as follows: We start with an introduction in
-%Chapter~\ref{chap:intro} where you will learn what is needed to run {\tt \dope}. 
+%Chapter~\ref{chap:intro} where you will learn what is needed to run {\tt \dope{}}. 
 %Further you will learn what problems we can solve and how all the different classes 
@@
 In particular, we remove all tedious routine implementation needed in 
-\deal by introducing appropriate classes that take care of all the 
+\deal{} by introducing appropriate classes that take care of all the 
 problem independent implementation while leaving only the problem specific 
@@
 such that
-\[
+\begin{align}\label{eq:prototype_weak}
 a(u)(\phi) = f(\phi) \quad \forall \phi \in V,
-\]
-with some appropriate space $V$ for the test functions and a 
+\end{align}
+with some suitable space $V$ for the test functions and a 
 right hand side $f$.
@@
 i.e., $u = (v,p)$ and $\phi = (\phi_u,\phi_p)$ then
-\[
-a(u)(\phi) = (\nabla v, \nabla\phi_u) + (v \nabla v,\phi_u) + (p, \nabla \cdot \phi_u) + (\nabla \cdot v ,\phi_p)
-\quad \forall \phi \in V = H^1_0(\Omega;\R^d) \times L^2_0(\Omega)
-\]
-Then the solution will be searched for in $V + (u_d,0)$ where $u_d$ is some 
+\begin{align}\label{eq:ns}
+a(u)(\phi) = \nu(\nabla v, \nabla\phi_u) + (v \nabla v,\phi_u) + (p, \nabla \cdot \phi_u) + (\nabla \cdot v ,\phi_p)
+\end{align}
+for all $\phi \in V = H^1_0(\Omega;\R^d) \times L^2_0(\Omega)$.
+The solution will be searched for in $V + (u_d,0)$ where $u_d$ is some 
 continuation of the given Dirichlet boundary data.
 Then typically we do not only want to calculate $u$ but more likely we are 
-interested in one (or several) numbers $J(u)$.
+interested in one (or several) functionals depending on the solution, i.e., some numbers $J(u)$.
+%For example, if $u$ is the solution of the Navier-Stokes equations \eqref{eq:ns}, we look for the Drag 
+%\begin{align*}
+%J(u)=\int
+%\end{align*}
 
 In order to solve this problem, i.e., calculate $J(u)$, we need to discretize 
-the equation, here using finite elements. This means we need to pick a 
+the equation, in \dope, we are\marginpar{here using?} using finite elements to this end. This means we need to pick a 
 subdivision $\mathcal T_h$, for instance into quadrilaterals,
-of the domain $\Omega$ and a finite element, in the concrete case we could, 
+of the domain $\Omega$ and a finite element. In the concrete case we could, 
 for instance, take the $\mathcal Q_2/\mathcal Q_1$ Taylor-Hood element.
@@
 Next, one needs to solve the discrete equation
-\[
+\begin{align}\label{eq:discrete_equation}
 a(u_h)(\phi_h) = f(\phi_h) \quad \forall \phi \in V_h.
-\]
+\end{align}
 To this end one can employ a newton type method with some globalization 
@@
 
-To evaluate the residual will require the calculation of the integrals 
+The evaluation of the residual will require the calculation of the integrals 
 involved in the weak formulation of the problem. For example 
 the $i$-th component of the residual, with respect to the given FE-basis 
-$\phi_i$, will read as follows
+$\phi_{h,i}$, will read as follows
 \[
- a(u_h)(\phi_i) - f(\phi_i) = \sum_{K\in \mathcal T_h} \int_K \nabla u, \nabla \phi_i\,dx + \ldots
+ a(u_h)(\phi_{h,i}) - f(\phi_{h,i}) = \sum_{K\in \mathcal T_h} \nu\int_K \nabla u, \nabla \phi_i\,dx + \ldots
 \]
@@
 that offers a method to calculate the residual given a user-defined 
-object \texttt{PDE} that implements the local element integrals, i.e.,
-\begin{lstlisting}
-  template<typename PDE,typename VECTOR>
+object \texttt{pde} that implements the local element integrals, i.e.,
+\begin{lstlisting}
+  template<typename PROBLEM,typename VECTOR>
       void Integrator
@@
 \end{lstlisting}
-where \texttt{VECTOR} is one of the possible vector types given by \deal.
+where \texttt{VECTOR} is one of the possible vector types given by \deal{}.
 It is clear that we need to ensure that the integrator object has access
@@
 again we can define a method in the integrator that does exactly this provided
-we have a user defined object \texttt{PDE} that can evaluate the local 
+we have a user defined object \texttt{pde} that can evaluate the local 
 integrals needed for assembling the matrix, i.e.,
 \begin{lstlisting}
-  template<typename PDE,typename MATRIX>
+  template<typename PROBLEM,typename MATRIX>
       void Integrator
@@
 \end{lstlisting}
-where now \texttt{MATRIX} is one of the matrix types supported by \deal.
+where now \texttt{MATRIX} is one of the matrix types supported by \deal{}.
 With this we can now define a nonlinear solver \texttt{NewtonSolver} 
@@
 \begin{lstlisting}
-  template<typename PDE, typename VECTOR>
+  template<typename PROBLEM, typename VECTOR>
       void NewtonSolver
@@
 \begin{remark}
-In fact, the above constructors are not exactly what is implemented in \dope.
+In fact, the above constructors are not exactly what is implemented in \dope{}.
 This is due to the fact, that for instance 
 there are several different sparsity patterns available in 
-\deal which the user should be able chose. Thus it needs to be 
+\deal{} which the user should be able chose. Thus it needs to be 
 given to the \texttt{SpaceTimeHandler}. However, to keep the 
@@
 \subsubsection{Problem dependent implementation of stationary PDE problems}
-
-To solve the problem, we need a configuration (a mesh), eventually 
-a diffusion parameter, a finite element, a quadrature rule,
-a linear solver (here the CG solver), and some output such as
-a graphical solution and eventually some measurement functional.
-Each of those routines are written in the \texttt{main.cc} file. 
-We start by including necessary files both the \dope library and
-the \deal software:
-\begin{lstlisting}
-  #include "pdeproblemcontainer.h" etc. ...
-\end{lstlisting}
-In the \texttt{main} function, the following steps need to be done. First, 
-the parameter reader for runtime parameters might be initialized. In a parameter
-file the user might specify material parameters for the concrete PDE, but also
-general parameters for nonlinear and linear solvers, and output-related 
-aspects.
-
-A linear finite element in two dimensions and two components is 
-initialized:
-\begin{lstlisting}
-  FESystem<2> state_fe(FE_Q<2>(1), 2);
-\end{lstlisting}
-and a quadrature rule of order three is used to approximate the 
-integrals:
-\begin{lstlisting}
-  QGauss<2> quadrature_formula(3);
-\end{lstlisting}
-Then, the \texttt{LocalPDE} object comes into play. This is an quite 
-important feature of \dope because it allows for easy implementation 
-of the variational formulation. Specific explication is given below because
-the implementation is separated in \texttt{localpde.h}.
-Afterwards, a \texttt{LocalPointFunctional} is declared. As for the PDE, the 
-implementation is done in another file, namely \texttt{localfunctional.h}.
-
-Generically, each problem in \dope is considered to be time-dependent and nonlinear because
-these are the 'real' world problems. Therefore, (even for the linear and stationary Laplace equation),
-we have to declare a (pseudo) time:
-\begin{lstlisting}
-  std::vector<double> times(1, 0.);
-\end{lstlisting}
-Then, the spatial mesh (here the unit square) is created and three-times globally-refined:
-\begin{lstlisting}
-  GridGenerator::hyper_cube(triangulation, 0, 1);
-  triangulation.refine_global(3);
-\end{lstlisting}
-A major component which organizes the whole solution process (independent from 
-the problem!) is the 
-\begin{lstlisting}
-  MethodOfLines_StateSpaceTimeHandler<...> DOFH (...);
-\end{lstlisting}
-This method handles all dofs in space and time and different versions exist
-for forward as well as optimization problems.
-
-Afterwards, the problem (here the Laplace PDE) is given to the \texttt{DOFH} via
-\begin{lstlisting}
-  OP P(LPDE, DOFH);
-\end{lstlisting}
-It remains to declare boundary conditions for the problem under consideration. 
-Let's simply take zero Dirichlet conditions on the boundary:
-\begin{lstlisting}
-  std::vector<bool> comp_mask(2);
+To solve a stationary PDE problem, the user has to provide some code which can roughly be divided into the following main sections:
+\begin{description}
+\item[Discretization] Specify what kind of triangulation, finite element, quadrature rule, etc. we want to use.
+\item[Problem-data] Define the weak form, the corresponding matrices, the functionals as well as the boundary conditions.
+With exception of the boundary conditions, we have to specify all this only cell-wise, \dope{} handles the rest for us.
+\item[Solvers] Which (non-)linear solver should we use? \marginpar{besser algorithmen?}
+\item[Miscellaneous] ParameterReader? Outputhandler etc.
+\end{description} 
+[Hier eventuell auf Unterschied zw. blosser Objekterstellung und dem tatsaechlichen 'Programmieren' fuer localpde etc. eingehen.]
+
+We describe now a program designed to compute a value $J(u)$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in figure~\ref{fig:example_ns} with  $f=0$, and the functional $J$ is defined as [TODO: Definition von J.
+Nutze Benchmark case oder lenkt das zu sehr ab?].
+Additionally, for given inflow $v_{\operatorname{in}}$ we set  boundary conditions $
+v = v_{\operatorname{in}} \text{ on } \Gamma_{in}\text{ and }v = 0  \text{ on } \Gamma_{wall}$.
+\marginpar{Das Beispiel hier besser nochmal zusammen schreiben?}
+
+\begin{figure}[h]
+\centering
+\resizebox{0.5\textwidth}{!}{\input{Pictures/example_ns}}
+\caption{Flow around a cylinder with 
+circle-center $C=(0.2,0.2)$ and radius $r=0.05$.}
+\label{fig:example_ns}
+\end{figure}
+
+We split our program into the files \texttt{main.cc}, \texttt{functionals.h} and \texttt{localpde.h}.
+The \texttt{main.cc}-file is responsible for the declaration of various objects which have mostly to do with the \textit{discretization}, the \textit{solvers} and the general \textit{program sequence}, whereas the other two contain most of the problem-data -- especially the cell-wise contributions of the weak form, the right hand side and the functional.
+We give a brief explanation on every file in the following.
+
+\paragraph{\texttt{main.cc}}
+In this file we gather the basic components that are necessary for the solution process of a (stationary) partial differential equation.
+The structure of the program is basically the same for every equation under consideration, what changes is typically just the \textit{kind} of  finite element or the linear solver.\marginpar{Bin mit diesen Saetzen nicht so ganz zufrieden, mir faellt aber gerade nichts besseres ein.
+Mir geht es hier darum, dass ich nicht wirklich was programmieren, sondern eher etwas auswaehle, wohingegen in den anderen files tatsaechlich was zu tun ist.}
+%The implementation of problem related data happens somwhere else, here we basically just pick together things like th
+
+After the usual preamble, in which we include the necessary header-files from \deal{}, \dope{} and our own problem (i.e.
+\texttt{localfunctional.h} and \texttt{localpde.h}), we start by defining an instance of a \texttt{ParameterReader}-object, which handles run-time parameter.
+\begin{lstlisting}
+   ParameterReader pr;
+\end{lstlisting}
+In a text-file we might specify some problem-related parameters (for instance the viscosity $\nu$), but also general parameters for (non)-linear solvers and the output of computed data. 
+After that, we create a triangulation named \texttt{triangulation}.
+This is done exactly as in \deal{}, so we can choose between several possibilities: We could explicitly create a grid by providing a list of vertices and cells, read in a triangulation via \texttt{dealii::GridIn}-class or use the \texttt{dealii::GridGenerator}, which offers common domains like hypercubes or hyper-balls.
+
+[code, Grid einlesen im Benchmark-Fall, ansonsten GridGenerator something. Hier auch auf Randfarben kurz eingehen.]
+
+Next, we create  quadrature rules (for the evaluation of integrals over cells and faces) and a finite element appropriate for the PDE at hand.
+We choose a gaussian quadrature of order three and the previosly mentioned Tayler-Hood element.
+\marginpar{Wie klar machen, dass vieles von deal.II uebernommen und nicht selbst geschrieben?}
+\begin{lstlisting}
+  QGauss<2> quad_formula(3);
+  QGauss<1> face_quad_formula(3);
+  FESystem<2> finite_element(dealii::FE_Q<2>(2), 2,
+                             dealii::FE_Q(1), 1);
+\end{lstlisting}
+%IntegratorDataContainer<VECTOR, 2> idc(quad_formula,
+%                                         face_quad_formula);
+%The \texttt{IntegratorDataContainer} class just makes the constructor interface of 
+Then, the \texttt{LocalPDE} and \texttt{LocalFunctional} objects come into play.
+\begin{lstlisting}
+LocalPDE<VECTOR> local_pde;
+LocalFunctional<VECTOR> local_functional;
+\end{lstlisting}
+These hold the information of the PDE and functional under consideration. Further explication is given below because
+the definition of this classes is done spererately in \texttt{localpde.h} resp.   \texttt{localfunctional.h}.
+
+In the next step, we define an object of type \texttt{SpaceTimeHandler}, which handles all the things connected with the triangulation as well as the finite element space.
+This class is also capable of organizing degrees of freedom in time-dependent problems, hence the name.
+\begin{lstlisting}
+  SpaceTimeHandler<FE> dof_handler(triangulation, finite_element);
+\end{lstlisting}
+There exist different versions of this class for forward as well as optimization problems.
+
+Now we have everything at hand to create an instantiation of type \texttt{PDEProblemContainer}.
+This basically represents the disrectized problem, and bundles things like boundary conditions, the PDE, the functional and the discretization.
+First, we give the constructor the \texttt{local\_pde} and the \texttt{dof\_handler}.
+\begin{lstlisting}
+  PDEProblemContainer<LocalPDE> prob_container(local_pde,
+                                                  dof_handler);
+\end{lstlisting}
+After that, we add the functional under consideration and, because it is a boundaryfunctional, we specify the part of the boundary on which the functional operates (this is done via a 'boundarycolor').
+\begin{lstlisting}
+  prob_container.AddFunctional(&local_functional);
+  prob_container.SetBoundaryFunctionalColors(1);
+\end{lstlisting}
+Next, we incorporate the dirichlet data.
+First, we define two functions representing the dirichlet boundary conditions.
+These are given to an interface\marginpar{SimpleDirichletData weglassen?} class \texttt{SimpleDirichletData}.
+\begin{lstlisting}
+  ZeroFunction<2> zf;
+  InflowFunction<2> if;
+  SimpleDirichletData<VECTOR, 2> dd_zf(zf);
+  SimpleDirichletData<VECTOR, 2> dd_if(if);
+\end{lstlisting}
+Then we have to tell the \texttt{prob\_container} for which components and boundaries we prescribe the dirichlet conditions.
+\begin{lstlisting}
+  std::vector<bool> comp_mask(3);
 
@@
   comp_mask[1] = true;
-
-  DOpEWrapper::ZeroFunction<2> zf(2);
-  SimpleDirichletData<VECTOR, 2> DD1(zf);
-
-  P.SetDirichletBoundaryColors(0, comp_mask, &DD1);
-\end{lstlisting}
-Next, the solver for the linear equations needs to be specified:
-\begin{lstlisting}
-  SSolver solver(&P, "fullmem", pr, idc);
-\end{lstlisting}
-Finally, the output handler is initialized via
-\begin{lstlisting}
-  P.RegisterOutputHandler(&out);
-  solver.RegisterOutputHandler(&out);
-\end{lstlisting}
+  comp_mask[2] = false;
+
+  prob_container.SetDirichletBoundaryColors(0, comp_mask, &dd_if);
+  prob_container.SetDirichletBoundaryColors(2, comp_mask, &dd_zf);
+  prob_container.SetDirichletBoundaryColors(3, comp_mask, &dd_zf);
+\end{lstlisting}
+Now we create the object which steers the whole solution process, the previosly mentioned \texttt{StatPDEProblem}.
+It takes care of solving the discretized equation and evaluating the given functionals.
+The class depends actually on some template-parameters, so for the sake of abbreviation, we defined a typedef at the beginning of the \texttt{main.cc}-file.
+\begin{lstlisting}
+typedef StatPDEProblem<NEWTONSOLVER,
+           PDEProblemContainer<LocalPDE>, VECTOR, 2> SSOLVER;
+\end{lstlisting}
+Here, \texttt{NEWTONSOLVER} itself is an abbreviation of
+\begin{lstlisting}
+typedef NewtonSolver<LINEARSOLVER, VECTOR, 2> NEWTONSOLVER;
+\end{lstlisting}
+The first template parameter of \texttt{NewtonSolver} describes which linear solver we use in this example. We opted for a direct solver, so: 
+\begin{lstlisting}
+typedef DirectLinearSolverWithMatrix<MATRIX,
+            VECTOR, 2> LINEARSOLVER;
+\end{lstlisting}
+\marginpar{Kein IDC, dafuer quadformeln eingefuegt.}
+With the type settled, we create the object \texttt{solver}, which needs the discretization (represented here by the \texttt{prob\_container}), the parameter reader as well as the quadrature rules.
+\begin{lstlisting}
+  SSOLVER solver(&prob_container, pr, quad_formula, face_quad_formula);
+\end{lstlisting}
+\marginpar{outputhandler erlaeutern?}
+%Finally, the output handler is initialized via
+%\begin{lstlisting}
+%  P.RegisterOutputHandler(&out);
+%  solver.RegisterOutputHandler(&out);
+%\end{lstlisting}
 Finally, the programm is started with:
@@
   solver.ReInit();
-  out.ReInit();
-  out.Write(outp, 1, 1, 1);
   solver.ComputeReducedFunctionals();
 \end{lstlisting}
-After explaining the basic compononents that are necessary (and always the same in \dope) 
-for all  (linear and stationary) partial differential equations, we shall give 
-an overview to the implementation of the specific form of the PDE. 
-Beside specification of boundary conditions, the form of the local PDE depends 
-on the concrete problem under consideration. 
-This step is done in the \texttt{localpde.h} file. As already explained,
-each PDE is treatet in a nonlinear manner such that for solving 
-the Laplace problem the following nonlinear step is necessary to recapitulate:
-Find $\delta u\in V$:
-\[
-(\nabla \delta u, \nabla \phi) = -(\nabla u, \nabla \phi) 
-\]
-and $u^{n+1} = u^n + \delta u$ for $n=1,2,3,ldots$.
-Therefore, the implementation reads for the right hand side:
-\begin{lstlisting}
-   void CellEquation(...)
+
+\paragraph{\texttt{localpde.h}}
+We shall now give an overview to the implementation of the specific form of the PDE.
+
+\marginpar{Rekapitulierung des Newton-Verfahrens?}
+
+Let $N$ be the number of degrees of freedom of our discretization and $\Set{\phi_{h,i}|1\leq i \leq N}$ be a given FE-Basis.
+As stated in [sectionref], to solve the discrete equation \eqref{eq:discrete_equation} we need the residual, i.e.
+the vector
+\begin{align}\label{eq:discrete_residual}
+\left(a(u_h)(\phi_{h,i})-(f,\phi_{h,i})\right)_{i=1}^N,
+\end{align}
+and the Jacobi-matrix of the weak form, i.e.
+\begin{align}\label{eq:discrete_matrix}
+\left(\nabla a(u_h)(\phi_{h,j},\phi_{h,i})\right)_{i,j=1}^N.
+\end{align}
+Now all we have to supply here are the cell-wise contributions to these terms, the \texttt{Integrator} takes care of assembling \eqref{eq:discrete_residual} and \eqref{eq:discrete_matrix}.
+
+To this end, we define a class \texttt{LocalPDE} which provides the three methods
+\texttt{CellEquation}, \texttt{CellRightHandSide} and \texttt{CellMatrix}.
+The implementation of the terms works by and large as in \deal{}.
+We show the basic structure of these methods using the example of \texttt{CellEquation}.
+
+\begin{lstlisting}
+  void CellEquation(const CellDataContainer<2>& cdc,
+                    dealii::Vector<double> &local_cell_vector)
     {
-      // implement (\nabla u, \nabla \phi) 
-    }
-\end{lstlisting}
-Specifically, the inner loops reads
-\begin{lstlisting}
-   for (all quadrature points per cell q)
+      const auto& fe_values = cdc.GetFEValuesState();
+      unsigned int n_dofs_per_cell = cdc.GetNDoFsPerCell();
+      unsigned int n_q_points = cdc.GetNQPoints();
+\end{lstlisting}
+ The \texttt{CellDataContainer} contains all the information we need to compute the cell-contributions (things like number of quadrature points, number of degrees of freedom on the cell, the \texttt{dealii::FEValues}-objects, etc.).
+%  \begin{lstlisting}
+%      _uvalues.resize(n_q_points,Vector<double>(3));
+%      _ugrads.resize(n_q_points,vector<Tensor<1,2> >(3));
+%      cdc.GetValuesState("last_newton_solution", _uvalues);
+%      cdc.GetGradsState("last_newton_solution", _ugrads);
+%  \end{lstlisting}
+  \begin{lstlisting}
+     cdc.GetValuesState("last_newton_solution", _uvalues);
+     cdc.GetGradsState("last_newton_solution", _ugrads);
+ \end{lstlisting}
+\texttt{\_uvalues} and \texttt{\_ugrads} are of type \texttt{std::vector<...>}, with $\ldots$ symbolizing the appropriate data type.
+In these two lines, we load the (gradient of the) solution of the last newton iteration in all the quadrature points on the cell into the two \texttt{std::vector} objects.
+
+In the next linex, we loop over the quadrature points and the number of dofs of the cell.
+In the innermost loop, we evaluate the weak form tested with $\phi_{i,h}$ on the cell with the help of the previosly selected quadrature formula.
+%     const FEValuesExtractors::Vector velocities(0);
+%     const FEValuesExtractors::Scalar pressure(2); 
+ \begin{lstlisting}
+    for(unsigned int q_point = 0; q_point < n_q_points; q_point++)
     {
-        Tensor<2, 2> ugrads; // solution vector
-
-	for (all dofs per cell i)
-        {
-	  const Tensor<2, 2> phi_i_grads_u; // test function
-
-          local_cell_vector(i) += scale * 
-            scalar_product(ugrads, phi_i_grads_u) *
-	    state_fe_values.JxW(q_point);
-        }
-      }
-\end{lstlisting}
-For the left hand side, we need to implement the Jacobian matrix:
-\begin{lstlisting}
-   void CellMatrix(...)
-    {
-      // implement  (\nabla \delta u, \nabla \phi)
-    }
-\end{lstlisting}
-Specifically,
-\begin{lstlisting}
-      for (all quadrature points per cell q)      
-        for (all dofs per cell i)    
-          for (all dofs per cell j)
-          {
-            local_entry_matrix(i, j) += 
-              scalar_product(phi_grads_u[j], phi_grads_u[i]) * 
-              state_fe_values.JxW(q_point);
-          }
-\end{lstlisting}
-The integration of the equation for each DoF and each quadrature point on each
-cell is then done 
-in the \texttt{templates/integrator.h} function. Finally, as usually in 
-finite element computations, this leads to a discrete system. Since
-we treat each problem with a Newton method, within each Newton step, the 
-following linear system must be solved:
-\[
-A\delta u = b
-\]
-in which $A$ is specified by \texttt{CellMatrix(...)} and 
-$b$ by \texttt{CellEquation(...)}. The vector $\delta u$ denotes the solution to the 
-present Newton step:
-\[
-u^n = u^{n-1} + \omega \delta u
-\]
-with a line search parameter $\omega\in (0,1]$.
-
-The target functional 
-is now implemented in \texttt{functional.h} and 
-treats a point evaluation:
-\begin{lstlisting}
-  double PointValue(...)
-   {
-    Point<2> p1(0.5, 0.5);
-    Vector<double> tmp_vector(2);
-    // more stuff is necessary here
-    VectorTools::point_value (...);
-    double x = tmp_vector(0);
-
-    return x;
-   }
-\end{lstlisting}
+      Tensor<2,2> vgrads;
+      vgrads[0][0] = _ugrads[q_point][0][0]; 
+      //etc.
+      const Tensor<2,2> phi_i_grads_v = 
+          state_fe_values[velocities].gradient (i, q_point);
+      for (unsigned int i=0;i<n_dofs_per_cell;i++)
+        local_cell_vector(i) += ( _nu 
+                                *_vgrads[q_point] * phi_i_grads_v 
+                                + //all the other terms)
+                                * state_fe_values.JxW(q_point);
+\end{lstlisting}
+
+% [CellEquation]
+% 
+% \begin{lstlisting}
+%   void CellRightHandSide(const CellDataContainer<2>& cdc,
+%                          dealii::Vector<double> &local_cell_vector,
+%                          double scale);
+% \end{lstlisting}
+% 
+% \begin{lstlisting}
+%   void CellMatrix(const CellDataContainer<2>& cdc,
+%                   dealii::FullMatrix<double> &local_entry_matrix);
+% \end{lstlisting}
+% The PDE under consideration is nonlinear, so let us briefly recapitulate manner such that for solving 
+% the Laplace problem the following nonlinear step is necessary to recapitulate:
+% Find $\delta u\in V$:
+% \[
+% (\nabla \delta u, \nabla \phi) = -(\nabla u, \nabla \phi) 
+% \]
+% and $u^{n+1} = u^n + \delta u$ for $n=1,2,3,ldots$.
+% Therefore, the implementation reads for the right hand side:
+% \begin{lstlisting}
+%    void CellEquation(...)
+%     {
+%       // implement (\nabla u, \nabla \phi) 
+%     }
+% \end{lstlisting}
+% Specifically, the inner loops reads
+% \begin{lstlisting}
+%    for (all quadrature points per cell q)
+%       {
+%         Tensor<2, 2> ugrads; // solution vector
+%  
+% 	for (all dofs per cell i)
+%         {
+% 	  const Tensor<2, 2> phi_i_grads_u; // test function
+% 
+%           local_cell_vector(i) += scale * 
+%             scalar_product(ugrads, phi_i_grads_u) *
+% 	    state_fe_values.JxW(q_point);
+%         }
+%       }
+% \end{lstlisting}
+% For the left hand side, we need to implement the Jacobian matrix:
+% \begin{lstlisting}
+%    void CellMatrix(...)
+%     {
+%       // implement  (\nabla \delta u, \nabla \phi)
+%     }
+% \end{lstlisting}
+% Specifically,
+% \begin{lstlisting}
+%       for (all quadrature points per cell q)      
+%         for (all dofs per cell i)    
+%           for (all dofs per cell j)
+%           {
+%             local_entry_matrix(i, j) += 
+%               scalar_product(phi_grads_u[j], phi_grads_u[i]) * 
+%               state_fe_values.JxW(q_point);
+%           }
+% \end{lstlisting}
+% The integration of the equation for each DoF and each quadrature point on each
+% cell is then done 
+% in the \texttt{templates/integrator.h} function. Finally, as usually in 
+% finite element computations, this leads to a discrete system. Since
+% we treat each problem with a Newton method, within each Newton step, the 
+% following linear system must be solved:
+% \[
+% A\delta u = b
+% \]
+% in which $A$ is specified by \texttt{CellMatrix(...)} and 
+% $b$ by \texttt{CellEquation(...)}. The vector $\delta u$ denotes the solution to the 
+% present Newton step:
+% \[
+% u^n = u^{n-1} + \omega \delta u
+% \]
+% with a line search parameter $\omega\in (0,1]$.
+
+\paragraph{\texttt{localfunctional.h}}
+[TODO Christian]
+
+% The target functional 
+% is now implemented in \texttt{functional.h} and 
+% treats a point evaluation:
+% \begin{lstlisting}
+%   double PointValue(...)
+%    {
+%     Point<2> p1(0.5, 0.5);
+%     Vector<double> tmp_vector(2);
+%     // more stuff is necessary here
+%     VectorTools::point_value (...);
+%     double x = tmp_vector(0);
+% 
+%     return x;
+%    }
+% \end{lstlisting}
+
+% TWick:
+% To solve the problem, we need a configuration (a mesh), eventually 
+% a diffusion parameter, a finite element, a quadrature rule,
+% a linear solver (here the CG solver), and some output such as
+% a graphical solution and eventually some measurement functional.
+% Each of those routines are written in the \texttt{main.cc} file. 
+% We start by including necessary files both the \dope{} library and
+% the \deal{} software:
+% \begin{lstlisting}
+%   #include "pdeproblemcontainer.h" etc. ...
+% \end{lstlisting}
+% In the \texttt{main} function, the following steps need to be done. First, 
+% the parameter reader for runtime parameters might be initialized. In a parameter
+% file the user might specify material parameters for the concrete PDE, but also
+% general parameters for nonlinear and linear solvers, and output-related 
+% aspects.
+% 
+% A linear finite element in two dimensions and two components is 
+% initialized:
+% \begin{lstlisting}
+%   FESystem<2> state_fe(FE_Q<2>(1), 2);
+% \end{lstlisting}
+% and a quadrature rule of order three is used to approximate the 
+% integrals:
+% \begin{lstlisting}
+%   QGauss<2> quadrature_formula(3);
+% \end{lstlisting}
+% Then, the \texttt{LocalPDE} object comes into play. This is an quite 
+% important feature of \dope{} because it allows for easy implementation 
+% of the variational formulation. Specific explication is given below because
+% the implementation is separated in \texttt{localpde.h}.
+% Afterwards, a \texttt{LocalPointFunctional} is declared. As for the PDE, the 
+% implementation is done in another file, namely \texttt{localfunctional.h}.
+% 
+% Generically, each problem in \dope{} is considered to be time-dependent and nonlinear because
+% these are the 'real' world problems. Therefore, (even for the linear and stationary Laplace equation),
+% we have to declare a (pseudo) time:
+% \begin{lstlisting}
+%   std::vector<double> times(1, 0.);
+% \end{lstlisting}
+% Then, the spatial mesh (here the unit square) is created and three-times globally-refined:
+% \begin{lstlisting}
+%   GridGenerator::hyper_cube(triangulation, 0, 1);
+%   triangulation.refine_global(3);
+% \end{lstlisting}
+% A major component which organizes the whole solution process (independent from 
+% the problem!) is the 
+% \begin{lstlisting}
+%   MethodOfLines_StateSpaceTimeHandler<...> dof_handler (...);
+% \end{lstlisting}
+% This method handles all dofs in space and time and different versions exist
+% for forward as well as optimization problems.
+% 
+% Afterwards, the problem (here the Laplace PDE) is given to the \texttt{dof_handler} via
+% \begin{lstlisting}
+%   OP P(local_pde, dof_handler);
+% \end{lstlisting}
+% It remains to declare boundary conditions for the problem under consideration. 
+% Let's simply take zero Dirichlet conditions on the boundary:
+% \begin{lstlisting}
+%   std::vector<bool> comp_mask(2);
+% 
+%   comp_mask[0] = true;
+%   comp_mask[1] = true;
+% 
+%   DOpEWrapper::ZeroFunction<2> zf(2);
+%   SimpleDirichletData<VECTOR, 2> DD1(zf);
+% 
+%   P.SetDirichletBoundaryColors(0, comp_mask, &DD1);
+% \end{lstlisting}
+% Next, the solver for the linear equations needs to be specified:
+% \begin{lstlisting}
+%   SSolver solver(&P, "fullmem", pr, idc);
+% \end{lstlisting}
+% Finally, the output handler is initialized via
+% \begin{lstlisting}
+%   P.RegisterOutputHandler(&out);
+%   solver.RegisterOutputHandler(&out);
+% \end{lstlisting}
+% Finally, the programm is started with:
+% \begin{lstlisting}
+%   solver.ReInit();
+%   out.ReInit();
+%   out.Write(outp, 1, 1, 1);
+%   solver.ComputeReducedFunctionals();
+% \end{lstlisting}
+% After explaining the basic compononents that are necessary (and always the same in \dope{}) 
+% for all  (linear and stationary) partial differential equations, we shall give 
+% an overview to the implementation of the specific form of the PDE. 
+% Beside specification of boundary conditions, the form of the local PDE depends 
+% on the concrete problem under consideration. 
+% This step is done in the \texttt{localpde.h} file. As already explained,
+% each PDE is treatet in a nonlinear manner such that for solving 
+% the Laplace problem the following nonlinear step is necessary to recapitulate:
+% Find $\delta u\in V$:
+% \[
+% (\nabla \delta u, \nabla \phi) = -(\nabla u, \nabla \phi) 
+% \]
+% and $u^{n+1} = u^n + \delta u$ for $n=1,2,3,ldots$.
+% Therefore, the implementation reads for the right hand side:
+% \begin{lstlisting}
+%    void CellEquation(...)
+%     {
+%       // implement (\nabla u, \nabla \phi) 
+%     }
+% \end{lstlisting}
+% Specifically, the inner loops reads
+% \begin{lstlisting}
+%    for (all quadrature points per cell q)
+%       {
+%         Tensor<2, 2> ugrads; // solution vector
+%  
+% 	for (all dofs per cell i)
+%         {
+% 	  const Tensor<2, 2> phi_i_grads_u; // test function
+% 
+%           local_cell_vector(i) += scale * 
+%             scalar_product(ugrads, phi_i_grads_u) *
+% 	    state_fe_values.JxW(q_point);
+%         }
+%       }
+% \end{lstlisting}
+% For the left hand side, we need to implement the Jacobian matrix:
+% \begin{lstlisting}
+%    void CellMatrix(...)
+%     {
+%       // implement  (\nabla \delta u, \nabla \phi)
+%     }
+% \end{lstlisting}
+% Specifically,
+% \begin{lstlisting}
+%       for (all quadrature points per cell q)      
+%         for (all dofs per cell i)    
+%           for (all dofs per cell j)
+%           {
+%             local_entry_matrix(i, j) += 
+%               scalar_product(phi_grads_u[j], phi_grads_u[i]) * 
+%               state_fe_values.JxW(q_point);
+%           }
+% \end{lstlisting}
+% The integration of the equation for each DoF and each quadrature point on each
+% cell is then done 
+% in the \texttt{templates/integrator.h} function. Finally, as usually in 
+% finite element computations, this leads to a discrete system. Since
+% we treat each problem with a Newton method, within each Newton step, the 
+% following linear system must be solved:
+% \[
+% A\delta u = b
+% \]
+% in which $A$ is specified by \texttt{CellMatrix(...)} and 
+% $b$ by \texttt{CellEquation(...)}. The vector $\delta u$ denotes the solution to the 
+% present Newton step:
+% \[
+% u^n = u^{n-1} + \omega \delta u
+% \]
+% with a line search parameter $\omega\in (0,1]$.
+% 
+% 
+% The target functional 
+% is now implemented in \texttt{functional.h} and 
+% treats a point evaluation:
+% \begin{lstlisting}
+%   double PointValue(...)
+%    {
+%     Point<2> p1(0.5, 0.5);
+%     Vector<double> tmp_vector(2);
+%     // more stuff is necessary here
+%     VectorTools::point_value (...);
+%     double x = tmp_vector(0);
+% 
+%     return x;
+%    }
+% \end{lstlisting}
 
@@
 \end{lstlisting}
-Here, the advantage is that \dope orders the terms in the timestepping
+Here, the advantage is that \dope{} orders the terms in the timestepping
 problems
@@
 that are applicable (in general) to all types of problems. Specifically,
-\dope provides goal-oriented mesh refinement with the help of the 
+\dope{} provides goal-oriented mesh refinement with the help of the 
 Dual-Weighted-Residual (DWR) method \cite{BeRa96}. Here, a certain 
@@
 
-Moreover, \dope provides standard error estimators based on residual evaluation and 
+Moreover, \dope{} provides standard error estimators based on residual evaluation and 
 a ZZ error estimator, i.e., measureing the smoothness of the discrete first derivatives.
-The latter one is accessed via \deal.
-
-In addition, \dope provides multi-mesh procedures 
+The latter one is accessed via \deal{}.
+
+In addition, \dope{} provides multi-mesh procedures 
 \todo{kann Winni hier etwas schreiben?}
@@
 %\subsection{Interfaces to other software packages}
-%Since DOpE uses as basis the software library \deal, all its interfaces
+%Since DOpE uses as basis the software library \deal{}, all its interfaces
 %to other packages can be accessed to. For instance, such libararies 
@@
 %%\label{documentation}
-%%At the present step, the \dope project comes with 
+%%At the present step, the \dope{} project comes with 
 %%an detailed documentation of all features and examples 
@@
 %%
-%%In addition, the \dope test suite provides some regression tests. 
+%%In addition, the \dope{} test suite provides some regression tests. 
 %%They are run to compare 
@@
 We present three numerical tests which demonstrate different
-features of \dope:
+features of \dope{}:
 \begin{itemize}
@@
 with the literature, it can be identified that 
-the implementation in \dope is correct.
+the implementation in \dope{} is correct.
 
@@
 In this article, we described the features 
-and applications of the \dope project, which 
+and applications of the \dope{} project, which 
 was initiated at the Heidelberg University in 2010.

### 2012-10-20 Paper_May_2012/main.tex
deal.II -> Differential Equations and

@@
 In this article, we describe the software library 
-\textit{Deal.II Optimization Environment} (\dope{}).
+\textit{Differential Equations and Optimization Environment} (\dope{}).
 The main feature of \dope{} is that it provides a unified interface to high level algorithms 
@@
 \label{introduction}
-The \textit{Deal Optimization Environment} (\dope{}) 
+The \textit{Differential Equations and Optimization Environment} (\dope{}) 
 provides a software toolkit to solve forward PDE

### 2012-10-20 Examples/Documentation/foreword.tex
deal.II -> Differential Equations and

@@
 In this report, we describe the DOpE
-\textit{Deal Optimization Environment} project. 
+\textit{Differential Equations and  Optimization Environment} project. 
 Originally, the project was initiated in the year 2009 at the 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The \textit{Deal Optimization Environment} (DOpE) project is 
-based on the \textit{deal.II} finite element library which has been developed
+The \textit{Differential Equations and  Optimization Environment} (DOpE) project is 
+mainly based on the \textit{deal.II} finite element library which has been developed
  initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{deal}.

### 2012-10-20 Examples/Documentation/introduction.tex
Update of the documentation, includes now higher order mappings.
deal.II -> Differential Equations and

@@
 \subsection{deal.II}
-This library is based upon \texttt{deal.II} hence in order to run 
+This library is mainly based upon \texttt{deal.II} hence in order to run 
 \texttt{DOpElib} you need a running copy of \texttt{deal.II}.
@@
     DoFHandler and HPDoFHandler.
+  \item \texttt{wrapper/fevalues\underline{ }wrapper.h} {\bf Will be removed soon!}
+  \item \texttt{wrapper/function\underline{ }wrapper.h} An interface that allows to use functions that depend 
+    not only on space but also on time.
+  \item \texttt{wrapper/mapping\underline{ }wrapper.h} An interface that allows to use \texttt{deal.II}-mappings as well as  \texttt{deal.II}-mapping collections depending of the DoFHandler in use.
+To this end, the class has a template parameter \texttt{DOFHANDLER}.
   \item \texttt{wrapper/preconditioner\underline{ }wrapper.h} Contains wrappers for several of the preconditioners
@@
     interfaces for their initialization.
-  \item \texttt{wrapper/function\underline{ }wrapper.h} An interface that allows to use functions that depend 
-    not only on space but also on time.
   \item \texttt{wrapper/snopt\underline{ }wrapper.h} An interface to the SNOPT fortran library. This is an additional
@@
     by SNOPT using our library.
-  \item \texttt{wrapper/fevalues\underline{ }wrapper.h} {\bf Will be removed soon!}
 \end{itemize}

### 2012-10-22 Paper_May_2012/main.tex
update of part 2.1.2

@@
 In order to solve this problem, i.e., calculate $J(u)$, we need to discretize 
-the equation, in \dope, we are\marginpar{here using?} using finite elements to this end. This means we need to pick a 
+the equation. In \dope, we are using finite elements to do this. This means we need to pick a 
 subdivision $\mathcal T_h$, for instance into quadrilaterals,
@@
 $\phi_{h,i}$, will read as follows
-\[
+\begin{align}\label{eq:residual_vector}
  a(u_h)(\phi_{h,i}) - f(\phi_{h,i}) = \sum_{K\in \mathcal T_h} \nu\int_K \nabla u, \nabla \phi_i\,dx + \ldots
-\]
+ \end{align}
 Again, we see that, except for the local element integrals, all 
@@
 
-We describe now a program designed to compute a value $J(u)$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in figure~\ref{fig:example_ns} with  $f=0$, and the functional $J$ is defined as [TODO: Definition von J.
-Nutze Benchmark case oder lenkt das zu sehr ab?].
-Additionally, for given inflow $v_{\operatorname{in}}$ we set  boundary conditions $
-v = v_{\operatorname{in}} \text{ on } \Gamma_{in}\text{ and }v = 0  \text{ on } \Gamma_{wall}$.
-\marginpar{Das Beispiel hier besser nochmal zusammen schreiben?}
-
+We describe now a program designed to compute a value $J(u)$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in figure~\ref{fig:example_ns} with  $f=0$, and $J$ is the drag-force of the enclosed circle.
+We assume given Dirichlet data for the velocity $v$ on the boundary $\Gamma_{D}=\Gamma_{in}\cup\Gamma_{wall}\cup\Gamma_{circ}$.
+This is the Benchmark-Case 2D-1 from \cite{TuSchae96}, see there for further details.
+%Additionally, for given inflow $v_{\operatorname{in}}$ we set  boundary conditions 
+%v = v_{\operatorname{in}} \text{ on } \Gamma_{in}\text{ and }v = 0  \text{ on } \Gamma_{wall}.
 \begin{figure}[h]
@@
 
-We split our program into the files \texttt{main.cc}, \texttt{functionals.h} and \texttt{localpde.h}.
+We split the program into the parts \texttt{main.cc}, \texttt{functionals.h} and \texttt{localpde.h}.
 The \texttt{main.cc}-file is responsible for the declaration of various objects which have mostly to do with the \textit{discretization}, the \textit{solvers} and the general \textit{program sequence}, whereas the other two contain most of the problem-data -- especially the cell-wise contributions of the weak form, the right hand side and the functional.
@@
 In a text-file we might specify some problem-related parameters (for instance the viscosity $\nu$), but also general parameters for (non)-linear solvers and the output of computed data. 
-After that, we create a triangulation named \texttt{triangulation}.
-This is done exactly as in \deal{}, so we can choose between several possibilities: We could explicitly create a grid by providing a list of vertices and cells, read in a triangulation via \texttt{dealii::GridIn}-class or use the \texttt{dealii::GridGenerator}, which offers common domains like hypercubes or hyper-balls.
-
-[code, Grid einlesen im Benchmark-Fall, ansonsten GridGenerator something. Hier auch auf Randfarben kurz eingehen.]
-
-Next, we create  quadrature rules (for the evaluation of integrals over cells and faces) and a finite element appropriate for the PDE at hand.
+After that, we create a triangulation of the domain $\Omega$ named \texttt{triangulation}.
+%This is done exactly as in \deal{}, so we can choose between several possibilities: We could explicitly create a grid by providing a list of vertices and cells, read in a triangulation via \texttt{dealii::GridIn}-class or use the \texttt{dealii::GridGenerator}, which offers common domains like hypercubes or hyper-balls.
+
+Next, we create some quadrature rules (for the evaluation of integrals over cells and faces) and a finite element appropriate for the PDE at hand.
 We choose a gaussian quadrature of order three and the previosly mentioned Tayler-Hood element.
@@
 \begin{lstlisting}
-  SpaceTimeHandler<FE> dof_handler(triangulation, finite_element);
+  SpaceTimeHandler<TRIANGULATION, FE> dof_handler(triangulation,
+                                                  finite_element);
 \end{lstlisting}
@@
 \end{lstlisting}
-Next, we incorporate the dirichlet data.
-First, we define two functions representing the dirichlet boundary conditions.
-These are given to an interface\marginpar{SimpleDirichletData weglassen?} class \texttt{SimpleDirichletData}.
-\begin{lstlisting}
-  ZeroFunction<2> zf;
-  InflowFunction<2> if;
-  SimpleDirichletData<VECTOR, 2> dd_zf(zf);
-  SimpleDirichletData<VECTOR, 2> dd_if(if);
-\end{lstlisting}
+Next, we incorporate the dirichlet data, which is given by the class \texttt{DirichleData}.
 Then we have to tell the \texttt{prob\_container} for which components and boundaries we prescribe the dirichlet conditions.
-\begin{lstlisting}
-  std::vector<bool> comp_mask(3);
-
-  comp_mask[0] = true;
-  comp_mask[1] = true;
-  comp_mask[2] = false;
-
-  prob_container.SetDirichletBoundaryColors(0, comp_mask, &dd_if);
-  prob_container.SetDirichletBoundaryColors(2, comp_mask, &dd_zf);
-  prob_container.SetDirichletBoundaryColors(3, comp_mask, &dd_zf);
+This is done via a boundary-color (0 stands for $\Gamma_D$) and a \texttt{std::vector<bool> comp\_mask = (true, true, false)}.
+\begin{lstlisting}
+  DirichletData dd;
+  prob_container.SetDirichletBoundaryColors(0, comp_mask, dd);
 \end{lstlisting}
@@
 \end{lstlisting}
-\marginpar{Kein IDC, dafuer quadformeln eingefuegt.}
 With the type settled, we create the object \texttt{solver}, which needs the discretization (represented here by the \texttt{prob\_container}), the parameter reader as well as the quadrature rules.
 \begin{lstlisting}
-  SSOLVER solver(&prob_container, pr, quad_formula, face_quad_formula);
-\end{lstlisting}
-\marginpar{outputhandler erlaeutern?}
+  SSOLVER solver(prob_container, pr, quad_formula, face_quad_formula);
+\end{lstlisting}
 %Finally, the output handler is initialized via
@@
 %\end{lstlisting}
-Finally, the programm is started with:
+Finally, the programm is started by
 \begin{lstlisting}
@@
 
-\marginpar{Rekapitulierung des Newton-Verfahrens?}
-
+%\marginpar{Rekapitulierung des Newton-Verfahrens?}
 Let $N$ be the number of degrees of freedom of our discretization and $\Set{\phi_{h,i}|1\leq i \leq N}$ be a given FE-Basis.
@@
   void CellEquation(const CellDataContainer<2>& cdc,
-                    dealii::Vector<double> &local_cell_vector)
+                    VECTOR &local_cell_vector)
     {
@@
   \begin{lstlisting}
-     cdc.GetValuesState("last_newton_solution", _uvalues);
-     cdc.GetGradsState("last_newton_solution", _ugrads);
- \end{lstlisting}
-\texttt{\_uvalues} and \texttt{\_ugrads} are of type \texttt{std::vector<...>}, with $\ldots$ symbolizing the appropriate data type.
-In these two lines, we load the (gradient of the) solution of the last newton iteration in all the quadrature points on the cell into the two \texttt{std::vector} objects.
-
-In the next linex, we loop over the quadrature points and the number of dofs of the cell.
-In the innermost loop, we evaluate the weak form tested with $\phi_{i,h}$ on the cell with the help of the previosly selected quadrature formula.
+     cdc.GetValuesState("last_newton_solution", uvalues);
+     cdc.GetGradsState("last_newton_solution", ugrads);
+ \end{lstlisting}
+ In these two lines, we load the (gradient of the) solution of the last newton iteration in all the quadrature points on the cell into the two vectors \texttt{uvalues} and \texttt{ugrads}.
+
+In the next lines, we loop over the quadrature points and the number of DoFs of the cell.
+In the innermost loop, we evaluate the weak form tested with $\phi_{i,h}$ (represented by \texttt{phi\_i\_grads\_v}) on the cell with the help of the previosly selected quadrature formula.
 %     const FEValuesExtractors::Vector velocities(0);
@@
     {
+      //initialize phi_i_grads_v etc.
       Tensor<2,2> vgrads;
-      vgrads[0][0] = _ugrads[q_point][0][0]; 
-      //etc.
-      const Tensor<2,2> phi_i_grads_v = 
-          state_fe_values[velocities].gradient (i, q_point);
+      vgrads[0]= ugrads[q_point][0]; 
+      vgrads[1]= ugrads[q_point][1];
       for (unsigned int i=0;i<n_dofs_per_cell;i++)
         local_cell_vector(i) += ( _nu 
-                                *_vgrads[q_point] * phi_i_grads_v 
+                                * vgrads * phi_i_grads_v 
                                 + //all the other terms)
                                 * state_fe_values.JxW(q_point);
-\end{lstlisting}
-
+    }
+\end{lstlisting}
+\dope{} takes now care of the distribution of the local cell contributions into  the vector of the residual \eqref{eq:residual_vector}.
+The methods \texttt{CellRightHandSide} and \texttt{CellMatrix} follow the same structure as
+\texttt{CellEquation}, so we do not go into the details here.
 % [CellEquation]
@@
 \paragraph{\texttt{localfunctional.h}}
-[TODO Christian]
+The implementation of the boundary-functional $J$ follows the outline of the computation of the cell-residual in \texttt{localpde.h}.
+However, there are two main differences.
+First, we do not have to specify the method \texttt{CellEquation}, but \texttt{BoundaryValue} instead.
+This is due to the fact that we consider a boundary-functional in this example, and so we have to specify the value of the functional not cell-wise but face wise (for the faces that lie on $\Gamma_{circ}$).
+The second difference is that as we only want to evaluate a functional and not a residual-vector, we just loop over the quadrature points and not over the degrees of freedom on the face.
+\begin{lstlisting}
+ double
+ BoundaryValue(const FaceDataContainer<2>& fdc)
+  {
+    //same as before, get n_q_points etc.
+    fdc.GetFaceValuesState("state", ufacevalues);
+    fdc.GetFaceGradsState("state", ufacegrads);
+\end{lstlisting}
+Notice that we get here the values of the computed solution by giving the string "state".
+\begin{lstlisting}
+    for(unsigned int q_point = 0; q_point < n_q_points; q_point++)
+      erg += -1. * ( _nu * ufacegrads[q_point][0]
+                * state_fe_face_values.normal_vector(q_point)
+                + //other terms)        
+                * state_fe_face_values.JxW(q_point);
+    return erg;
+  }
+\end{lstlisting}
 

### 2012-10-22 Paper_May_2012/main.tex
Kleines Update of DWR-Beispiel

@@
 Christian.
+In this example we consider a laminar flow around a cylinder in 2d.
+To be more precise, we are computing the benchmark 2D-1 from \cite{TuSchae96}.
+The computational domain is as depicted in figure~\ref{fig:example_ns}.
+We want to compute the drag force on the enclosed cylinder
+\begin{equation}
+c_D = \frac 1 {20} \int_{\Gamma_{circ}} \nu\partial_nu _1 - pn,
+\end{equation}
+with the normal $n$ and $\nu = 0.001$. To achieve an efficient computation of the quantity $c_D$, we employ goal-oriented mesh refinement with the help of the DWR-method, see \cite{BeRa96}.
+
+\begin{itemize}
+\item Bild Freiheitsgrade vs Fehler, global und 
+\item Bild adaptive mesh, bild der Loesung (?)
+\item eventuell kurz auf duale gleichung eingehen.
+\end{itemize}
 

### 2012-10-22 Paper_May_2012/main.tex
Added comment on linear solver for newton alg, and explanation for PDEProblemContainer

@@
 \begin{lstlisting}
+ template<typename LINSOLVE>
   template<typename PROBLEM, typename VECTOR>
-      void NewtonSolver
+      void NewtonSolver<LINSOLVE>
              ::NonlinearSolve(PROBLEM& pde, 
@@
 \end{lstlisting}
+Where \texttt{LINSOLVE} denotes the solver that should be used to solve the 
+linearized equations during the newton iteration.
 
@@
 That combines all the necessary steps to calculate $J(u_h)$.
+
+Finally, we will need to have an object that is capable of providing all 
+the necessary functionality required from the template argument 
+\texttt{PROBLEM} in all of the above. We provide a wrapper named
+\texttt{PDEProblemContainer} collecting all the problem specific information 
+like the description of the PDE and the functional, the boundary values, etc. 
+Although, having such an object instead of directly giving these informations
+to the \texttt{StatPDEProblem} seems like a nuicance at the moment, we will 
+see that it is of greater interest for time-dependent problems or 
+optimization problems in what follows in Section~\ref{sec:timedep} and
+Section~\ref{sec:opt} respectively.
 
@@
 
-\subsubsection{Nonstationary problems}
+\subsubsection{Nonstationary problems}\label{sec:timedep}
 In this section, the 
@@
 
-\subsection{Solving of optimization problems}
+\subsection{Solving of optimization problems}\label{sec:opt}
 Upgrading the framework for solving optimization problems

### 2012-10-25 Paper_May_2012/main.tex
Paper: Section 2.1.3 und 2.1.4 fertig gestellt. Ich wusste nicht mehr ob ich auch 2.1.3 machen sollte, habe ich jetzt einfach mit gemacht.
Ausserdem in FSI Beispiel Config-Skizze geloescht und mich auf Christians
Skizze bezogen.

@@
 \newcommand{\todo}[1]{\textbf{\textsc{\textcolor{black}{TODO: #1}}}}
-
+%\newcommand{\mymarginpar}[1]{\marginpar{\textcolor{red}{\scriptsize{#1}}}}
+\newcommand{\mymarginpar}[1]{\marginpar{\textcolor{red}{\normalsize{#1}}}}
 %
@@
 With exception of the boundary conditions, we have to specify all this only cell-wise, \dope{} handles the rest for us.
-\item[Solvers] Which (non-)linear solver should we use? \marginpar{besser algorithmen?}
+\item[Solvers] Which (non-)linear solver should we use? 
+\mymarginpar{besser algorithmen?}
 \item[Miscellaneous] ParameterReader? Outputhandler etc.
@@
 \subsubsection{Nonstationary problems}\label{sec:timedep}
-In this section, the 
-numerical solution of nonstationary processes
-is considered. We focus on details which are 
-beyond stationary examples as discussed in the 
-previous sections. Before, we introduce 
-the steps of implementation in DOpElib, we 
-derive the basic theoretical concepts.
-As an example, we consider the heat equation:
-Find $u\in V$:
+To describe the solution of nonstationary 
+PDEs, we consider a time-dependent semi-linear form
+Let $I:=[0,T]$ be a time interval with end time point value $T$.
+To find some $u$ in a weak form, we define
 \[
-\partial_t u - \Delta u = 0
+\int_I a(u)(\phi) = \int_I f(phi) \quad \forall\phi\in X,
 \]
-and in weak formulation:
-Find $u\in V$:
+with some suitable Bochner space $X$.  As model problem,
+we consider the nonstationary Navier-Stokes equations
+in a domain $\Omega\in\mathbb{R}^d$. Find: 
+$(v,p)\in X$ with
 \[
-a(u)(\phi) = 0 \quad \forall \phi \in V,
+\int_I \bigl[ (\partial_t v, \phi)
++ \nu (\nabla v, \nabla \phi) + (v\cdot\nabla v, \phi)
+- (p,\nabla\cdot \phi)
++ (\nabla\cdot v, \chi)\bigr] \, dt
++ (v(0) - v^0, \phi(0))
+= \int_I (f,\phi) \, dt 
 \]
-with some appropriate space $V$ and 
+for all $\phi:= (\phi, \chi) \in X$ with
 \[
-a(u)(\phi) = (\partial_t u, \phi) + (\nabla u, \nabla \phi) \quad \forall \phi \in V.
+X:= \bigl\{ 
+(v,p) | v\in L^2(I,H_0^1(\Omega)^d , 
+\partial_t v\in L^2(I, L^2(\Omega)^d), 
+p\in L^2(I,L^2(\Omega)\setminus\mathbb{R} \bigr\}.
 \]
+As for stationary problems, the goal is to compute
+some target functional $J(\cdot)$ which is now possibly 
+time-dependent. To organize the discretization 
+process, we need to perform spatial and temporal 
+discretization. This is again done via
+a global object that treats the routines
+in DOpElib:
+\begin{lstlisting}
+  template<typename Triangulation, typename FE, typename times>
+      void MethodOfLines_SpaceTimeHandler(Triangulation &, FE, times);
+\end{lstlisting}
+Spatial discretization is the same
+as discussed previously, and thus, we briefly demonstrate
+routines to perform temporal integration in the following.
+
+\subsubsection{Problem-dependent 
+implementation of nonstationary problems}
+\label{sec:timedep:implementation}
+We explain in more detail the temporal discretization
+and specific issues of its implementation. 
+To keep the presentation easy, we abuse notation 
+by neglecting
+all material coefficients
+and give formal-related idea based on the weak 
+form of the Navier-Stokes equations:
+Find $(v,p)$ such that
+\begin{align*}
+(\partial_t v,\phi) 
++ (\nabla v, \nabla \phi)
++ (v\cdot\nabla v,\phi)
+-(p,\nabla\cdot \phi)
++(\nabla\cdot v, \chi)
+=(f,\phi).
+\end{align*}
 Temporal discretization using the One-step-$\theta$ scheme reads:
-Given the old timestep solution $u^n$, find $u^{n+1}$:
+Given the old timestep solution $(v^n,p^n)$, 
+find $(v^{n+1}, p^{n+1})$ (and multiplication 
+through by the timestep $k$):
 \[
-\frac{(u^{n+1} - u^{n}, \phi)}{k} + \theta (\nabla u^{n+1}, \nabla \phi)
-+ (1 - \theta) (\nabla u^{n}, \nabla \phi) = 0,
-\]
-Reordering in new and old terms and multiplication by the time step size $k$:
-\[
-(u^{n+1},\phi) + k \theta (\nabla u^{n+1}, \nabla \phi)
-= (u^n, \phi) - (1 - \theta) (\nabla u^{n}, \nabla \phi).
+(v^{n+1} - v^{n}, \phi)
++ k\theta (\nabla v^{n+1}, \nabla \phi)
++ k\theta (v^{n+1}\cdot\nabla v^n + 
+  v^{n}\cdot\nabla v^{n+1},\phi)
+- k (p^{n+1},\nabla\cdot \phi)
++ (\nabla\cdot v^{n+1}, \chi)
+= k\theta (f^{n+1},\phi) + k(1-\theta) (f^{n},\phi)
+- k(1-\theta) (\nabla v^{n}, \nabla \phi) 
 \]
@@
 \end{lstlisting}
-The main work is treated as for stationary problems
-within 
-\begin{lstlisting}
-   MethodOfLines_SpaceTimeHandler<...>
-\end{lstlisting}
-Again, the key issue is that this procedure is totally independent 
-of the specific form of the PDE. The user-related work 
-remains to provide the time derivative $\partial_t u$ in
-the implementation:
+
+The pressure and incompressibily terms should be 
+treated fully implicitly when time-discretizing 
+the Navier-Stokes equations.
+For this reason, the implementation is split up
+into several items to account for fully implicit 
+terms and remaining terms for which we need 
+old timestep and present timestep values. 
+Hence, in the local cell equation, a paramter
+is used to distigugish these terms:
+\begin{lstlisting}
+  void CellEquation(...)
+   {
+      local_cell_vector(i) +=  scale * (\nabla v^{n+1}, \nabla \phi) + etc.
+
+      local_cell_vector(i) +=  scale_ico * (-p^{n+1},\nabla\cdot v)
+   }
+\end{lstlisting}
+Analogoues procedure is performed in the local 
+cell matrix.
+
+In addition to stationary problems, the implementation
+of the time-derivative must be considered which is done
+in 
 \begin{lstlisting}
@@
 problems
-such that only $(u^{n+1},\phi)$ must be implemented. However, if someone 
+such that only $(u^{n+1},\phi)$ must be implemented.
+
+\begin{remark}[Nonlinear terms in the time derivative]
+However, if someone 
 needs explicit input of the old timestep term, then he can use 
@@
 \end{lstlisting}
+\end{remark}
+
+\begin{remark}[Fractional-Step-$\theta$-scheme]
+The temporal discretization in this section 
+is based on the One-Step-$\theta$-schemes. The library
+offers in addition an implementation 
+of the second-order and strongly $A$-stable 
+Fractional-Step-$\theta$-scheme in which 
+each time step is subdivided into three 
+sub-timeintervals. 
+\end{remark}
 
@@
 \subsection{Goal-oriented mesh refinement for Navier-Stokes}
-Christian.
 In this example we consider a laminar flow around a cylinder in 2d.
@@
 
-The configuration is 
-sketched in Figure \ref{configuration_csm_and_fsi_2D}. 
+The basic configuration is 
+sketched in Figure \ref{fig:example_ns} at which an elastic beam is attached 
+behind the rigid cylinder. 
 The computational domain has length $L=2.5m$ and height $H=0.41m$. The circle center
@@
 
-\begin{figure}[h]
-\centering
-\input{Pictures/pic_12}
-\caption{Flow around cylinder with elastic beam with 
-circle-center $C=(0.2,0.2)$ and radius $r=0.05$.}
-\label{configuration_csm_and_fsi_2D}
-\end{figure}
+%\begin{figure}[h]
+%\centering
+%\input{Pictures/pic_12}
+%\caption{Flow around cylinder with elastic beam with 
+%circle-center $C=(0.2,0.2)$ and radius $r=0.05$.}
+%\label{configuration_csm_and_fsi_2D}
+%\end{figure}
 

### 2012-11-05 Paper_May_2012/main.tex
Added comments on Stucture for Instat and OPT added Example on compliance minimization.

@@
 \[
-\int_I a(u)(\phi) = \int_I f(phi) \quad \forall\phi\in X,
+\int_I a(u)(\phi) = \int_I f(\phi) \quad \forall\phi\in X,
 \]
@@
 X:= \bigl\{ 
-(v,p) | v\in L^2(I,H_0^1(\Omega)^d , 
+(v,p) | v\in L^2(I,H_0^1(\Omega)^d) , 
 \partial_t v\in L^2(I, L^2(\Omega)^d), 
@@
 discretization. This is again done via
-a global object that treats the routines
-in DOpElib:
+an object that takes care of organizing the degrees of 
+freedom, as in the case of stationary problems.
+The only difference to the stationary case is that we need to 
+add the temporal subdivision \texttt{times}. In the simplest 
+case, when we have the same spatial triangulation 
+at all times, the constructor could look as follows:
 \begin{lstlisting}
   template<typename Triangulation, typename FE, typename times>
-      void MethodOfLines_SpaceTimeHandler(Triangulation &, FE, times);
-\end{lstlisting}
-Spatial discretization is the same
-as discussed previously, and thus, we briefly demonstrate
-routines to perform temporal integration in the following.
+      void SpaceTimeHandler(Triangulation &, FE, times&);
+\end{lstlisting}
+
+Now we may discretize the problem with some time-stepping scheme, which will 
+remove the temporal derivative and replace it by some difference quotients. 
+Since the exact way of doing this depends on the time-stepping scheme 
+we need a new object that takes the user-given description of the PDE and 
+generates the appropriate combinations of spatial integrals and scaling.
+Thus a generic problem may look as follows:
+\begin{lstlisting}
+  template<typename PROBLEM>
+      class TimeStepProblem;
+\end{lstlisting}
+At present we have prepared the following time stepping schemes in \dope:
+\begin{itemize}
+\item Forward-(Explicit)-Euler
+\item Backward-(Implicit)-Euler
+\item Crank-Nicolson scheme
+\item shifted Crank-Nicolson scheme (to deal with singular initial data)
+\item fractional-step-$\theta$ scheme
+\end{itemize}
+With this we can now run a loop over all timesteps, where in each of the 
+timesteps we can reuse the functionality to solve stationary PDEs 
+applied to the time-step problem.
+
+Now it is apparent, why we introduced the \texttt{StatPDEProblem}, because
+we now define the \texttt{InstatPDEProblem} to contain the necessary 
+loop over all time points. Next, we will describe the methods the user needs to 
+provide in addition to the stationary case.
 
@@
 find $(v^{n+1}, p^{n+1})$ (and multiplication 
-through by the timestep $k$):
-\[
+by the inverse of the timestep $k$):
+\begin{align*}
 (v^{n+1} - v^{n}, \phi)
-+ k\theta (\nabla v^{n+1}, \nabla \phi)
+&+ k\theta (\nabla v^{n+1}, \nabla \phi)
 + k\theta (v^{n+1}\cdot\nabla v^n + 
-  v^{n}\cdot\nabla v^{n+1},\phi)
-- k (p^{n+1},\nabla\cdot \phi)
-+ (\nabla\cdot v^{n+1}, \chi)
-= k\theta (f^{n+1},\phi) + k(1-\theta) (f^{n},\phi)
+  v^{n}\cdot\nabla v^{n+1},\phi)\\
+&- k (p^{n+1},\nabla\cdot \phi)
++ (\nabla\cdot v^{n+1}, \chi)\\
+=&\; k\theta (f^{n+1},\phi) + k(1-\theta) (f^{n},\phi)
 - k(1-\theta) (\nabla v^{n}, \nabla \phi) 
-\]
+\end{align*}
 The concrete choice of $\theta$ leads to a
@@
 and $g(\cdot)$ is some state constraint.
-As before, the main part is done in the \texttt{main} function there is no
-great difference to the previous explications. However, 
-since we utilize gradient-based optimization routines in the philosophy of
-\cite{Troe99,HiKu01},
-all derivatives must be implemented. All this information
-must given to the \texttt{localpde.h} file. Herein, one has to 
-implement the adjoint, tangent and adjoint Hession equations. 
-The algorithmic parts are described in the literature \cite{BeMeVe06}.
-
-Concretely, let us focus on the Laplace equation and 
-the following optimization problem (taken from 
-Example 2) in \texttt{Examples/OPT/StatPDE}:
+To keep the presentation brief we will neglect the additional bounds
+\[
+a \le q \le b,\qquad g(q,u) \le 0
+\]
+although they can be treated by \dope as well.
+Then, in order to minimize the functional $J$ one may apply the, so-called, 
+reduced approach, in which we eliminate the dependent variable $u$ by the solution 
+operator $S$ of the PDE, i.e., $a(q,S(q))(\phi) = 0 \quad \forall \phi\in V$. 
+This leads to the problem
 \begin{align*}
-\min_{(q,u)\in \mathbb R^3 \times H_0^1(\Omega; \mathbb R^2)} J(q,u) &=
-\frac{1}{2} \sum_{i=0}^2 |(u-\overline u)(x_i)|^2 + \frac{\alpha}{2}\|q\|^2\\
-\text{s.t.} (\nabla u,\nabla \phi) &= (f(q),\phi)\;\;\forall\,\phi \in H^1_0(\Omega; \mathbb R^2)
+\min\;&j(q) = J(q,S(q)). 
 \end{align*}
-on the domain $\Omega = [0,1]^2$, with
-\begin{itemize}
-\item the observation points
-\begin{align*}
-x_0 = (0.5, 0.5), \quad x_1 = (0.5, 0.25),\quad x_2 = (0.25, 0.25),
-\end{align*}
-\item the regularization parameter $\alpha = 0$, 
-\item the right hand side
-\begin{align*}
- f(q) &= q_0 \left(\begin{matrix}2\pi^2  \sin( \pi x) \sin(\pi y)\\0 \end{matrix}\right)\\
-      &+ q_1 \left(\begin{matrix}5\pi^2  \sin( \pi x) \sin(2\pi y)\\0 \end{matrix}\right)\\
-      &+ q_2 \left(\begin{matrix}0 \\8\pi^2  \sin(2\pi x) \sin(2\pi y)\end{matrix}\right)\\
-\end{align*}
-\item and the exact solution given by 
-\begin{align*}
- \overline{q} &= \bigl(1;0.5;1\bigr)\\
- \overline{u}& = \left(\begin{matrix} \sin( \pi x)( \sin(\pi y)+0.5\sin(2\pi y))\\\sin(2\pi x) \sin(2\pi y) \end{matrix}\right).
-\end{align*}
-\end{itemize}
-\begin{defi}[Constrained optimization]
-\label{problem_constraint_optimization}
-Minimize the cost functional $J(q,u)$ subject to the state equation
-$a(q,u)(\phi) = 0$ for $(q,u) \in 
-Q\times  \{v_D+{\cal X}\}$. 
-\end{defi}
-%
-The constrained optimization problem  on the space $Q_d\times {\cal X}$ 
-is reformulated into an unconstrained optimization problem on the
-space $Q_d$. Therefore, we assume the existence of the solution operator  
-$S:Q_d\rightarrow {v_D+{\cal X}}$ with a unique solution
-$U=S(q)$. Herewith, we define the reduced cost functional
-$j:Q_d\rightarrow \mathbb{R}$ by 
-%
-\begin{equation}
-  \label{opt_reduced_functional}
-  j(q) := J(q,S(q)) .
-\end{equation}
-%
-Thus, the constrained optimization problem can be formulated by means
-of  
-%
-\begin{defi}[Unconstrained optimization]
-  \label{opt_reduced_problem}
-  Minimize $j(q)$ for $q\in Q_d$.
-\end{defi}
-%
-Newton's method to solve Problem~\ref{opt_reduced_problem} reads:
-For $l=0,1,\ldots,$ solve
-\begin{equation}
-\label{opt_newton_method}
-\begin{aligned}
-j''(q^l)(\delta q , \tau q) &= -j'(q^l)(\delta q) 
-\quad\forall\tau q\in Q_d, \\
-q^{l+1} &= q^l + \omega\delta q,
-\end{aligned} 
-\end{equation}
-with a line search parameter $\omega\in (0,1]$ which 
-will are implemented in \texttt{reducednewtonalgorithm.h}.
-Specifically, the residual and the Hessian of Newton's method 
-\eqref{opt_newton_method}
-can be computed with the help of the following two results:
-\begin{proposition}[Residual of Newton's method]
-\label{prop_newton_residual}
-Let $q\in Q_d, u=S(q)\in {\cal X}$
-and $z\in {\cal X}$ the dual solution be obtained after solving 
-the first and second equation of the KKT system 
-\eqref{opt_KKT_system}. 
-Then the residual of Newton's method \eqref{opt_newton_method} 
-is defined as
-\begin{equation*}
-j'(q)(\delta q) := {\cal L}'_{q}(q,u,z)(\delta q),
-\end{equation*}
-i.e., in explicit representation
-\begin{equation*}
-j'(q)(\delta q) := \alpha_T (q , \tau q)_Q - 
-A_{q}'(q,u)(\tau q , z).
-\end{equation*}
-\end{proposition}
-\begin{proof}
-The proof uses standard techniques. Details can be found in  
-\cite{BeMeVe06}.
-\end{proof}
-\begin{proposition}[Hessian of Newton's method]
-  \label{prop_newton_hessian}
-  Let $q\in Q_d, u=S(q)\in {\cal X}$ and $z\in {\cal X}$ and  $\delta
-  q\in Q_d$ be given. Further, let $\delta u\in {\cal X}$ be the
-  solution of the tangent problem 
+Then, one can apply any algorithm to solve the, unconstraint, minimization 
+problem. This, will in any case require the evaluation of $j(q)$ which 
+we can do, e.g., using \texttt{StatPDEProblem::ComputeReducedFunctional} for 
+a stationary PDE. If one would like to use derivative based methods, one 
+needs in adition derivatives of $j$ and thus of $J$ as well as of $S$. 
+To do so, we remark, that one can calculate the derivative of 
+$j$ using the adjoint calculus. Then the directional derivative 
+of $j$ in direction $\delta q$ is given by
 \[
-  \delta u\in{\cal X},\quad 
-  a_{u}'(q,u)(\delta u , \phi) 
-  = -a_{q}'(q,u)(\delta q , \phi)  \quad\forall \phi\in {\cal
-    X},
+j'(q)(\delta q) = J'_q(q,S(q))(\delta q) + (S^* J_u'(q,S(q)),\delta q)
 \]
-  and $\delta z\in {\cal X}$ the solution of the adjoint Hessian 
-  problem  
-  \[
-  a_{u}'(q,u)(\phi, \delta z)
-  = J''_{uu}(q,u)(\delta u, \phi)
-  - a_{uu}''(q,u)(\delta u, \phi, z)
-  - a_{qu}''(q,u)(\delta q , \phi, z)  \quad\forall \phi\in
-  {\cal X}. 
-  \]
- Then it holds  for all $\tau q\in Q_d$:
-  \[
-  j''(q)(\delta q, \tau q):= 
-  \alpha_T (\delta q , \tau q)_Q -
-  a_{qq}''(q, u)(\delta q, \tau q , z) -
-  a_{uq}''(q,u)(\delta u, \tau q , z) -
-  a_{q}'(q,u)(\tau q , \delta z).
-  \]
-\end{proposition}
-\begin{proof}
-  For the proof this statement (including its time-dependent version), we refer to 
-  Becker et al. \cite{BeMeVe06}.
-  The stationary version of this proposition is easily derived by neglecting 
-  all time derivatives, which concludes the assertion. 
-\end{proof}
-The implementation of all previous terms 
-is realized in the \texttt{localpde.h} file.
-As an example, we examine the adjoint 
+where the index $q$ or $u$ denote directional derivatieves for the 
+respective argument. To calculate action of the adjoint operator $S^*$ 
+we note that $z =  S^* J_u'(q,S(q))$
+can be computed via the adjoint PDE
 \begin{equation}
@@
   \begin{aligned}
-    a_{u}'(q,u)(\phi , z) &= J'_{u}(q,u)(\phi) & \forall \phi&\in {\cal X},
+    a_{u}'(q,u)(\phi , z) &= J'_{u}(q,u)(\phi) & \forall \phi&\in {\cal X}.
   \end{aligned}
 \end{equation}
-
-The implementation it split into equation terms, i.e., 
+Similarly formulas for the second derivatives can be obtained.
+
+To do so, we define an object 
+\begin{lstlisting} 
+  template<typename PDE, typename COSTFUNCTIONAL>
+    class OPTProblemContainer;
+\end{lstlisting}
+that collects the user given PDE and functional descriptions and constructs the 
+required PDE and right hand side information needed for the calculations.
+
+This means for the user only the derivatives of the functional and the PDE 
+needs to be provided in addition to the 
 \begin{lstlisting} 
@@
    {
-     // implement (\nabla \phi, \nabla z)
+     // implement directional derivative of the CellEquation
    }
@@
 \begin{lstlisting} 
-  virtual void PointValue_U(...)
+  virtual void BoundaryValue_U(...)
       {
         // implement here the derivative of 
-        // the point cost functional
+        // the boundary cost functional
       }
 \end{lstlisting}
-All remaining terms of the previous two propositions 
-are implemented in the same manner.
-
-Afterwards, it remains to use an appropriate 
-optimization algorithm in order to solve the 
-problem. At present we offer a selection of algorithms that solve the reduced optimization 
-problem where the PDE constraint has been eliminated as explained previously.
-\begin{itemize}
-\item \texttt{opt\underline{ }algorithms/reducedalgorithm.h} An interface for all 
-  optimization problems in the reduced formulation. It offers some test functionality
-  to assert that the derivatives of the problem are computed correctly.
-\item \texttt{opt\underline{ }algorithms/reducednewtonalgorithm.h}
-  A line-search Newton algorithm using a cg method to invert the reduced hessian. 
-  Implementation ignores any additional constraints.
-\item \texttt{opt\underline{ }algorithms/reducedtrustregionnewton.h}
-  A trust region Newton algorithm using a cg method to invert the reduced hessian.
-  Implementation ignores any additional constraints.
-\item \texttt{opt\underline{ }algorithms/reduced\underline{ }snopt\underline{ }algorithm.h}
-  An algorithm to solve reduced optimization problems with additional control constraints.
-  ((reduced) state constraints are not yet implemented.)
-\item \texttt{opt\underline{ }algorithms/reducednewtonalgorithmwithinverse.h}
-  Line-search Newton algorithm that assumes there exists a method in the reduced problem
-  that can invert the reduced hessian. (This usually makes sense only if there is no 
-  PDE constraint.)
-\item \texttt{opt\underline{ }algorithms/generalized\underline{ }mma\underline{ }algorithm.h}
-  An implementation of the MMA-Algorithm for structural optimization using an augmented
-  Lagrangian formulation for the subproblems. The subproblem is implemented using the 
-  special purpose file
-  \texttt{include/augmentedlagrangianproblem.h}.
-\end{itemize}
-
-\subsection{Independent features}
+
+Due to the design of the classes it is easily possible to extend any given 
+and running code for a PDE both to the nonstationary counterpart, if the PDE is 
+stationary, as well as to an optimization problem where the constraint is 
+given by this PDE. Meaning that the use of our framework for the simulation of 
+PDEs will, due to the enforced interface, enable easy application of state-of-the-art 
+optimization algorithms to minimization problems governed by the PDE.
+
+\subsection{Additional features}
 In this final section, we shall give a brief summary of features 
@@
 (unknown) exact solution and its numerical approximation for lowest computational
-cost is the main goal of the DWR method.
-
-Moreover, \dope{} provides standard error estimators based on residual evaluation and 
-a ZZ error estimator, i.e., measureing the smoothness of the discrete first derivatives.
-The latter one is accessed via \deal{}.
-
-In addition, \dope{} provides multi-mesh procedures 
-\todo{kann Winni hier etwas schreiben?}
-
-Concerning temporal integration schemes, the library comes 
-with various time stepping schemes (based on finite differences), 
-such as the first-order forward Euler and first-order backward Euler scheme. 
-To compute fully nonstationary processes, the second-order schemes
-Crank-Nicolson (trapozoidal rule), 
-shifted Crank-Nicolson (slight shift to the implicit side to maintain
-global stability), 
-and the Fractional-Step-$\Theta$ scheme, are implemented.
-
-%\subsection{Problem description}
-%In order to allow our algorithms the automatic assembly of all required 
-%data we need to have some container which contains the complete problem 
-%description in a common data format. For this we have the following 
-%classes in \texttt{DOpEsrc/container}
-%\begin{itemize}
-%  \item \texttt{pdeproblemcontainer.h} Is used to describe  stationary PDE problems.
-%  \item \texttt{instatpdeproblemcontainer.h} {\bf TODO is still missing but will be used for nonstationary problems.} This will be implemented once we have nonstationary optimization problems running to avoid error duplication in the coding process.
-%  \item \texttt{optproblem.h} Is used to describe  OPT problems governed by 
-%    stationary PDEs. 
-%  \item \texttt{instatoptproblemcontainer.h} Is used to describe  OPT problems
-%    governed by nonstationary PDEs. The only difference to the stationary case
-%    is that we need to specify a time-stepping method.  
-%  \item \texttt{interfaces/functionalinterface.h} This gives an interface 
-%    for the functional $J(\cdot)$ and any other functional you may want to evaluate.
-%    In general this can be used as a base class to write your own functionals 
-%    in examples. We note that we only need to write the integrands on 
-%    elements or faces the loop over elements will be taken care of in the integrator.
-%    Specifically, derivatives are written therein, too.
-%\end{itemize}
-%In order to fill these containers there are two things to be done,
-%first we need to actually write some data, for instance,
-%the semilinear form $a(\cdot)(\cdot)$, a target functional $J(\cdot)$, etc.,
-%which describe the problem. Then we have to select some numerical 
-%algorithm components like finite elements, linear solvers $\ldots$.
-%The latter ones should be written such that when exchanging these components
-%none of the problem descriptions should require changes. 
-%Note that it still may be necessary to write some additional descriptions, 
-%e.g., if you solve the PDE with a fix point iteration you don't need derivatives
-%but if you want to use Newton's method, derivatives are needed.
-%
-%We will start by discussing the problem description components implemented so far
-%
-%
-%\subsection{Numeric components}
-%These are the components from which a user needs to select some in order to actually 
-%solve the given problem. They will not require any rewriting, but sometimes it is 
-%advisable to write other than the default parameter into the param file for the 
-%solution.
-%
-%\subsubsection{Space-time handler}
-%First we need to select a method how to handle all dofs in space and time.
-%\begin{itemize}
-%\item \texttt{basic/spacetimehandler\underline{ }base.h} This class is used to define 
-%  an interface to the dimension independent functionality of all space time dof handlers.
-%  {\bf TODO: Beispiele geben}
-%\item \texttt{basic/statespacetimehandler.h} Another intermediate interface class which adds 
-%  the dimension dependent functionality if only the variable $u$ is considered, i.e., a 
-%  PDE problem.
-%\item \texttt{basic/spacetimehandler.h } Same as above but with both $q$ and $u$, i.e., for
-%  OPT problems.
-%\item \texttt{basic/mol\underline{ }statespacetimehandler.h} Implementation of a method of 
-%  line space time dof handler for PDE problems. It has only one spatial 
-%  dofhandler that is used for all time intervals.
-%\item \texttt{basic/mol\underline{ }spacetimehandler.h} Same as above for OPT problems.
-%  A separate spatial dof handler for each of the variables $q$ and $u$ is maintained 
-%  but only one triangulation.
-%\item \texttt{basic/mol\underline{ }multimesh\underline{ }spacetimehandler.h}
-%  Same as above, but now in addition the triangulations for $q$ and $u$ can be refined
-%  separately from one common initial coarse triangulation. Note that this will
-%  in addition require the use of the multimesh version for integrator and 
-%  face- as well as celldatacontainer.
-%\end{itemize}
-%Note that we use these for stationary problems as well, but then you don't have to specify
-%any time information.
-%
-%\subsubsection{Container classes}
-%Second you will need to specify some container classes to be used to 
-%pass data between objects. At present you don't have much choice, but you may wish 
-%to reimplement some of these if you need data that is not currently included in 
-%the containers.
-%\begin{itemize}
-%\item \texttt{container/celldatacontainer.h} This object is used to pass data 
-%  given on the current element (cell) of the mesh to the functions in PDE, functional, 
-%  $\ldots$. 
-%\item \texttt{container/facedatacontainer.h} This object is used to pass data 
-%  given on the current face of the mesh to the functions in PDE, functional, 
-%  $\ldots$. 
-%\item \texttt{container/multimesh\underline{ }celldatacontainer.h} This is the same as the 
-%  celldatacontainer, but it
-%  is capable to handle data defined on an alternative triangulation.
-%\item \texttt{container/multimesh\underline{ }facedatacontainer.h} This is the same as the
-%  facedatacontainer, but it
-%  is capable to handle data defined on an alternative triangulation.
-%\item \texttt{container/integratordatacontainer.h} This contains some data that 
-%  should be passed to the integrator like quadrature formulas and the above cell and 
-%  face data container.
-%\end{itemize}
-%
-%
-%
-%
-%\subsection{Reduced problems (Solve the PDE)}
-%At times it is nice to remove the PDE constraint in (OPT). 
-%This is handled by so called reduced 
-%problems (for algorithmic aspects we refer the reader to 
-%\cite{BeMeVe06}). 
-%This means that the reduced problem implicitly solves the PDE whenever required
-%and eliminates the variable $u$ from the problem.
-%\begin{itemize}
-%\item \texttt{reducedproblems/statpdeproblem.h} This is used to remove the variable $u$ in 
-%  a stationary PDE problem. This means that call the method \\
-%  \texttt{StatPDEProblem::ComputeReducedFunctionals} will evaluate the functionals 
-%  defined in the problem description, i.e., in \texttt{PDEProblemContainer}, in the 
-%  solution of the given PDE.
-%\item \texttt{reducedproblems/statreducedproblem.h} This eliminates $u$ from the OPT
-%  problem with a stationary PDE.
-%\item \texttt{reducedproblems/instatreducedproblem.h} The same as above but for a
-%  nonstationary PDE. {\bf FIXME there is something wrong in this file see FIXME 
-%    comment in the source.}
-%\item \texttt{reducedproblems/voidreducedproblem.h} A wrapper file that eliminates $u$ 
-%  if it is not present anyways. This is used so that we can use the same routines to 
-%  solve problems that have no PDE constraint.
-%\end{itemize}
-%
-
-%
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%\subsection{Interfaces to other software packages}
-%Since DOpE uses as basis the software library \deal{}, all its interfaces
-%to other packages can be accessed to. For instance, such libararies 
-%are trilinos for with an algebraic multigrid method, MPI for a parallel
-%solution of the problem.
-%
-%In addition DOpE itself has an interface to SNOPT.
-%
-%%\section{Documentation, Code Development, Webpage}
-%%\label{documentation}
-%%At the present step, the \dope{} project comes with 
-%%an detailed documentation of all features and examples 
-%%in pdf format and an detailed programming code 
-%%documentation. 
-%%
-%%\todo{License? How to access to DOpE?}
-%%\todo{Name of the website}
-%%
-%%In addition, the \dope{} test suite provides some regression tests. 
-%%They are run to compare 
-%%the output to previous outputs. This is useful (necessary) after 
-%%changing programming code anywhere in the library. If a test
-%%succeeds, everything is fine in the library. If not, you should not
-%%check in your code into DOpE. Please make sure what is going wrong and WHY!
-%%Every command is computed via a Makefile. 
-%%
-%
-%
+cost is the main goal of the DWR method. The same routines 
+can also be used, without any modification, to use standard residual error estimators by exchanging the 
+error estimator class.
+
+Since in many applications it is desirable to use different discretizations
+for the state and the control \dope{} provides multi-mesh support, meaning that
+onec a problem has been written correctly using the same mesh for all finite element
+variables one can switch to a different SpaceTimeHandler that allows different meshes for all 
+variables without the need to change the user-given implementation of the PDE and functional.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 different time step sizes $k$. However, the time stepping scheme can be 
-very easily chosen in the \texttt{main.cc} function by declaring
-\begin{lstlisting}
-#define TSP1 ForwardEulerProblem
-#define TSP2 BackwardEulerProblem
-#define TSP3 CrankNicolsonProblem
-#define TSP4 ShiftedCrankNicolsonProblem
-#define TSP5 FractionalStepThetaProblem
-\end{lstlisting}
-and then using in the code the designated scheme, 
-for example here the Fractional-Step-$\theta$ scheme
-\texttt{TSP5}.
+very easily chosen in the \texttt{main.cc} function by choosing an appropriate 
+time stepping scheme, at present the following are availiable.
+\begin{lstlisting}
+ForwardEulerProblem
+BackwardEulerProblem
+CrankNicolsonProblem
+ShiftedCrankNicolsonProblem
+FractionalStepThetaProblem
+\end{lstlisting}
 
@@
 \subsection{Compliance minimization}
-Winni
-Results already obtained.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-\label{conclusions}
-In this article, we described the features 
-and applications of the \dope{} project, which 
-was initiated at the Heidelberg University in 2010.
-Specifically, the applications cover a broad 
-spectrum of numerical exmamples motivated 
-by different disciplines. 
-Currently, we are developing concepts for the efficient 
-numerical solution of time-dependent optimization
-problems governed by PDEs.
-
-\section{Outlook}
-\todo{Was wollen wir noch in den nächsten 1-2 Jahren erreichen?}
+In this example, we consider the compliance minimization of a standard MBB-Beam. 
+The thickness variable is allowed to take intermediate value apart from $0$ and $1$,
+i.e., we consider a variable thickness sheet. The discretization is done 
+using $Q_2$ elements for the displacement and discontinuous $P_0$ elements for the 
+thickness.
+\begin{figure}
+\centering
+\includegraphics[width=1.\textwidth, viewport=150 0 1024 500, clip]{Pictures/MBB.png}
+\caption{Thickness distribution for the MBB-beam minimum compliance problem.} 
+\label{res:mbb}
+\end{figure}
+A typical thickness distribution is shown in Figure~\ref{res:mbb}.
+
+\section*{Thanks}
+\todo{Wem wir so danken wollen, Boris, Dominik, Wolfgang ...?}
 

### 2012-11-05 Paper_May_2012/main.tex
Paper: Vorschlag fuer eine Einleitung und Danksagung

@@
 with the need for just a few lines of code to change.
-The article illustrates performance and features 
-of the software package by four numerical tests. 
+The article illustrates the performance and various features 
+of the software package and provides some 
+numerical results as demonstration.
 \end{abstract}
@@
 \label{introduction}
-The \textit{Differential Equations and Optimization Environment} (\dope{}) 
-provides a software toolkit to solve forward PDE
-problems as well as optimal control problems constrained by a PDE. 
+Today, there exist a broad variety of software libraries
+that provide numerical tools for solving partial differential
+equations (PDEs). Prominent examples are 
+\deal{} \cite{deal}, dune \cite{dune} , 
+or commercial solvers like Ansys \cite{ansys}). 
+In fact, these libraries provide very useful tools 
+for a huge variety of applications.
+
+However, often, the user is confronted with 
+certain bottlenecks. Either, standard high-level routines 
+which are necessary to solve even simplest PDEs,
+such as linear and nonlinear solvers or timestepping schemes must be implemented 
+by the user himself. However, such routines are problem-independent 
+and could be provided as standard tools. In fact, in many 
+numerical studies, the main task is to compute a  
+target functional which represents a certain error norm,
+point value, etc.
+Second, there is a ever increasing interest in computing 
+optimization problems in which the state equation is governed 
+by a PDE. Such optimization problems comprise optimal control,
+parameter estimation, inverse problems or optimal experimental design. 
+So far, there exist a very limited number of open source packages
+which provides such routines for scientific studies.
+
+On the other hand, 
+commercial toolkits provide these previously mentioned routines but limits user-interaction
+with the source code very often. This is quite unsatisfactory because
+developement of novel numerical algorithms should enable the programmer
+to change certain parameters or pieces of source code.
+
+The novel software toolkit 
+\textit{Differential Equations and Optimization Environment} (\dope{}) 
+trys to fill this gap. We present a software
+to solve forward PDE
+problems as well as optimization problems constrained by a PDE. 
 Its main feature is to give a unified interface to high level algorithms such as 
@@
 
-The
-solution of a broad variety of PDE is possible in other software
-libraries (like in \deal{} \cite{deal}, dune \cite{dune} , 
-or commercial solvers like Ansys \cite{ansys}) 
-as well, but
-\dope{} concentrates on a unified approach for both linear and nonlinear
-problems by interpreting every PDE problem as nonlinear and applying a
-Newton method to solve it. While \deal{} leaves much of the work and many
+While \deal{} (which might presently seen as the basis because 
+it provides various finite elements for discretization) 
+leaves much of the work and many
 decisions to the user, \dope{} intends to be user-optimized by delivering
 prefabricated tools which require from the user only adjustments connected
-to his specific problem. The solution of optimal control problems with PDE
+to his specific problem. The solution of optimization problems with PDE
 constraints is an innovation in the \dope{} framework.
@@
 
-The \dope{} project is 
-based on the \deal{} \cite{deal} finite element library which has been developed
- initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{deal}.
-The authors acknowledge their past experience as well as discussions with 
-the authors of the libraries 
-Gascoigne/RoDoBo project, which was initiated by 
-Roland Becker, Dominik Meidner,  and Boris Vexler \cite{rodobo}. 
-From which some of the ideas to modularize the algorithms have arisen.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%The aim of \dope{} is to provide a software toolkit to solve forward PDE
-%problems as well as optimal control problems constrained by PDE. The
-%solution of a broad variety of PDE is possible in \deal{} as well, but
-%\dope{} concentrates on a unified approach for both linear and nonlinear
-%problems by interpreting every PDE problem as nonlinear and applying a
-%Newton method to solve it. While \deal{} leaves much of the work and many
-%decisions to the user, \dope{} intends to be user-optimized by delivering
-%prefabricated tools which require from the user only adjustments connected
-%to his specific problem. The solution of optimal control problems with PDE
-%constraints is an innovation in the \dope{} framework.
-%The focus is on the numerical solution of both stationary and nonstationary
-%problems which come from different application fields, like elasticity and
-%plasticity, fluid dynamics, and fluid-structure interactions.
-%
-
-At the present stage the following key features are supported by the library
+Summarizing, 
+at the present stage the following key features are supported by the library
 \begin{itemize}
@@
 
-The article is organized as follows. In Section
-\ref{detailed_description}
+The article is organized as follows. In the main Section
+\ref{detailed_description}, a detailed description of 
+the main features is given. In fact, we first explain the solution 
+of stationary and nonstationary PDEs and give afterwards some 
+ideas how to treat optimization problems. Finally, 
+additional features are discussed. In the final Section
+\ref{applications}, we discuss three different numerical 
+examples which substantiates the features of \dope{}.
 
@@
 \section*{Thanks}
-\todo{Wem wir so danken wollen, Boris, Dominik, Wolfgang ...?}
+The \dope{} project makes use of various finite elements taken from  
+the \deal{} \cite{deal} finite element library which has been developed
+ initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{deal}.
+The authors acknowledge their past experience as well as discussions 
+on modularization of algorithms
+with 
+the authors of the libraries 
+Gascoigne/RoDoBo project, which was initiated by 
+Roland Becker, Dominik Meidner, and Boris Vexler \cite{rodobo}. 
+The second author thanks Mary F. Wheeler (ICES, Austin) for the 
+possibility to finish this work.
 

### 2012-11-05 Paper_May_2012/main.tex
Paper: Section 2.1.2: Ein wenig geändert, allerdings fand ich die Vorarbeiten von Christian schon recht gut.

@@
 \begin{description}
-\item[Discretization] Specify what kind of triangulation, finite element, quadrature rule, etc. we want to use.
+\item[Discretization] Specify what kind of triangulation, time stepping
+  scheme, finite element, quadrature rules, etc. we want to use.
 \item[Problem-data] Define the weak form, the corresponding matrices, the functionals as well as the boundary conditions.
@@
 \item[Solvers] Which (non-)linear solver should we use? 
-\mymarginpar{besser algorithmen?}
-\item[Miscellaneous] ParameterReader? Outputhandler etc.
+\item[Additional features] ParameterReader to read in the run-time parameters
+  such as material coefficients or adjustments for the solvers like tolerance
+  for Newton's method. Second, definition of Outputhandler etc.
 \end{description}
-[Hier eventuell auf Unterschied zw. blosser Objekterstellung und dem tatsaechlichen 'Programmieren' fuer localpde etc. eingehen.]
 
@@
 \paragraph{\texttt{main.cc}}
-In this file we gather the basic components that are necessary for the solution process of a (stationary) partial differential equation.
-The structure of the program is basically the same for every equation under consideration, what changes is typically just the \textit{kind} of  finite element or the linear solver.\marginpar{Bin mit diesen Saetzen nicht so ganz zufrieden, mir faellt aber gerade nichts besseres ein.
-Mir geht es hier darum, dass ich nicht wirklich was programmieren, sondern eher etwas auswaehle, wohingegen in den anderen files tatsaechlich was zu tun ist.}
-%The implementation of problem related data happens somwhere else, here we basically just pick together things like th
+In this file we gather the basic components that are necessary for the
+solution process of a (stationary) partial differential equation. The
+structure of the program is basically the same for every equation under
+consideration, what changes is typically just the \textit{kind} of  finite
+element or the linear solver. Here, a key issue is that we only need to pick 
+some pre-implemented routines from the \dope{} library rather than 
+implementing new features.
 
@@
 
-Next, we create some quadrature rules (for the evaluation of integrals over cells and faces) and a finite element appropriate for the PDE at hand.
-We choose a gaussian quadrature of order three and the previosly mentioned Tayler-Hood element.
-\marginpar{Wie klar machen, dass vieles von deal.II uebernommen und nicht selbst geschrieben?}
-\begin{lstlisting}
-  QGauss<2> quad_formula(3);
-  QGauss<1> face_quad_formula(3);
-  FESystem<2> finite_element(dealii::FE_Q<2>(2), 2,
-                             dealii::FE_Q(1), 1);
-\end{lstlisting}
+Next, we create some quadrature rules (for the evaluation of integrals over
+cells and faces) and a finite element appropriate for the PDE at hand. We
+choose a Gaussian quadrature of a certain degree and the previosly mentioned
+Tayler-Hood element. These routines are quite standard in each finite element
+software and we take advantage to take them from \deal{}.
+
 %IntegratorDataContainer<VECTOR, 2> idc(quad_formula,
@@
 %The \texttt{IntegratorDataContainer} class just makes the constructor interface of 
-Then, the \texttt{LocalPDE} and \texttt{LocalFunctional} objects come into play.
+Then, the two characteristic features \texttt{LocalPDE} and \texttt{LocalFunctional} objects come into play.
 \begin{lstlisting}
@@
 \end{lstlisting}
-These hold the information of the PDE and functional under consideration. Further explication is given below because
+These hold the information of the PDE and the functional under consideration. Further explication is given below because
 the definition of this classes is done spererately in \texttt{localpde.h} resp.   \texttt{localfunctional.h}.
@@
 
-In the next lines, we loop over the quadrature points and the number of DoFs of the cell.
+In the next lines, we finally implement our concrete PDE. Here, aim to provide 
+an interface which is closely related to the weak formulation which 
+has to be derived by the user. Thus,
+we loop over the quadrature points and the number of DoFs of the cell.
 In the innermost loop, we evaluate the weak form tested with $\phi_{i,h}$ (represented by \texttt{phi\_i\_grads\_v}) on the cell with the help of the previosly selected quadrature formula.

### 2012-11-07 Paper_May_2012/main.tex
Some comments to the new sections added.

@@
   involving stationary PDE constraints.
-\item Goal-oriented mesh adaptation with the dual-weighted residual method.
+\item Goal-oriented mesh adaptation with the Dual Weighted Residual method, see \cite{BR03}.
 \item Different spatial triangulations for control and state variables.
@@
 \subsection{Solving of PDEs}
-\subsubsection{Stationary problems}
+\subsubsection{Stationary problems}\label{subsubsec:stationary problems}
 For the start we consider a prototypical problem given by the solution of 
@@
 \begin{align}\label{eq:residual_vector}
- a(u_h)(\phi_{h,i}) - f(\phi_{h,i}) = \sum_{K\in \mathcal T_h} \nu\int_K \nabla u, \nabla \phi_i\,dx + \ldots
+ a(u_h)(\phi_{h,i}) - f(\phi_{h,i}) = \sum_{K\in \mathcal T_h} \nu\int_K \nabla u_h, \nabla \phi_{h,i}\,dx + \ldots
  \end{align}
@@
 of the matrix corresponding to the linearized equation. To calculate this,
-again we can define a method in the integrator that does exactly this provided
+ we can define a method in the integrator that does exactly this provided
 we have a user defined object \texttt{pde} that can evaluate the local 
@@
 \begin{description}
-\item[Discretization] Specify what kind of triangulation, time stepping
-  scheme, finite element, quadrature rules, etc. we want to use.
+\item[Discretization] Specify what kind of triangulation,
+% time steppingscheme,
+finite element, quadrature rules, etc. we want to use.
 \item[Problem-data] Define the weak form, the corresponding matrices, the functionals as well as the boundary conditions.
@@
 \end{lstlisting}
-After that, we add the functional under consideration and, because it is a boundaryfunctional, we specify the part of the boundary on which the functional operates (this is done via a 'boundarycolor').
+After that, we add the functional under consideration and, because it is a boundaryfunctional, we specify the part of the boundary on which the functional operates (this is done via a 'boundarycolor', $1$ represents $\Gamma_{circ}$).
 \begin{lstlisting}
@@
 \begin{lstlisting}
-  SSOLVER solver(prob_container, pr, quad_formula, face_quad_formula);
+  SSOLVER solver(prob_container, pr,
+                    quad_formula, face_quad_formula);
 \end{lstlisting}
@@
 Let $N$ be the number of degrees of freedom of our discretization and $\Set{\phi_{h,i}|1\leq i \leq N}$ be a given FE-Basis.
-As stated in [sectionref], to solve the discrete equation \eqref{eq:discrete_equation} we need the residual, i.e.
+As stated in subsection~\ref{subsubsec:stationary problems}, to solve the discrete equation \eqref{eq:discrete_equation} we need the residual, i.e.
 the vector
@@
 To describe the solution of nonstationary 
-PDEs, we consider a time-dependent semi-linear form
+PDEs, we consider a time-dependent semi-linear form {\color{red} Hier fehlt etwas? Vermutlich a? ~cg}
 Let $I:=[0,T]$ be a time interval with end time point value $T$.
-To find some $u$ in a weak form, we define
+To find some $u$ in a weak form, we define\marginpar{Inkonsistenz: $f(\phi)$ vs. $(f,\phi)$ etwas weiter unten. Ebenso Integrale mal mit $dt$, mal ohne. Waere dafuer, $dt$ wegzulassen.}
 \[
@@
 \dope{} provides goal-oriented mesh refinement with the help of the 
-Dual-Weighted-Residual (DWR) method \cite{BeRa96}. Here, a certain 
+Dual Weighted Residual (DWR) method \cite{BeRa96}. Here, a certain 
 cost functional such as a point evaluation, line intergration or domain
@@
 goal-oriented mesh refinement with the help of the 
-Dual-Weighted-Residual (DWR) method for the stationary Navier-Stokes equations.
+Dual Weighted Residual (DWR) method for the stationary Navier-Stokes equations.
 \item The second example presents nonstationary fluid-structure 

### 2012-11-07 Paper_May_2012/main.tex
Some comments to the new sections added.

@@
 \]
-although they can be treated by \dope as well.
+although they can be treated by \dope{} as well.
 Then, in order to minimize the functional $J$ one may apply the, so-called, 
 reduced approach, in which we eliminate the dependent variable $u$ by the solution 
-operator $S$ of the PDE, i.e., $a(q,S(q))(\phi) = 0 \quad \forall \phi\in V$. 
+operator $S$ of the PDE, i.e., 
+\begin{align*}
+a(q,S(q))(\phi) = 0 \quad \forall \phi\in V. 
+\end{align*}
 This leads to the problem

### 2012-11-07 Paper_May_2012/main.tex
Updated the deal reference to the new author.

@@
 equations (PDEs). Prominent examples are 
-\deal{} \cite{deal}, dune \cite{dune} , 
+\deal{} \cite{dealnew}, dune \cite{dune} , 
 or commercial solvers like Ansys \cite{ansys}). 
@@
 \dope{} provides goal-oriented mesh refinement with the help of the 
-Dual Weighted Residual (DWR) method \cite{BeRa96}. Here, a certain 
+Dual Weighted Residual (DWR) method, see \cite{BeRa96} or \cite{BR03}. Here, a certain 
 cost functional such as a point evaluation, line intergration or domain
@@
 The \dope{} project makes use of various finite elements taken from  
-the \deal{} \cite{deal} finite element library which has been developed
- initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{deal}.
+the \deal{} \cite{dealnew} finite element library which has been developed
+ initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{dealold}.
 The authors acknowledge their past experience as well as discussions 

### 2012-11-07 Paper_May_2012/main.tex
Paper: Einleitung etwas ueberarbeitet

@@
 \deal{} \cite{dealnew}, dune \cite{dune} , 
-or commercial solvers like Ansys \cite{ansys}). 
+or commercial solvers like ANSYS \cite{ansys} or COMSOL \cite{comsol}. 
 In fact, these libraries provide very useful tools 
@@
 However, often, the user is confronted with 
-certain bottlenecks. Either, standard high-level routines 
+certain bottlenecks. Either in open-source projects for academic research, 
+standard high-level routines 
 which are necessary to solve even simplest PDEs,
@@
 target functional which represents a certain error norm,
-point value, etc.
-Second, there is a ever increasing interest in computing 
+point value or traction, for example.
+Second, there is nowadays an increasing interest in computing 
 optimization problems in which the state equation is governed 
@@
 So far, there exist a very limited number of open source packages
-which provides such routines for scientific studies.
+which provides standard routines that serve as basis for scientific studies.
 
@@
 
-While \deal{} (which might presently seen as the basis because 
-it provides various finite elements for discretization) 
+While \deal{} (which might presently seen as the basis of DOpE  
+various finite elements are taken for the discretization) 
 leaves much of the work and many
@@
 
-Summarizing, 
+The project was initially initiated at the University of Heidelberg in 2010 and  
 at the present stage the following key features are supported by the library

### 2012-11-07 Paper_May_2012/main.tex
Paper: Section 2+3 Anmerkungen von Christian eingefuegt. Ich finde das Seite 6 aus typo-technischen Gruenden schwer zu lesen ist. Mir ist aber noch nichts eingefallen, wie man dies klarer aufbereiten koennte. Grundsaetzlich haben wir die wichtigsten Dinge erwaehnt

@@
 To describe the solution of nonstationary 
-PDEs, we consider a time-dependent semi-linear form {\color{red} Hier fehlt etwas? Vermutlich a? ~cg}
+PDEs, we consider a time-dependent semi-linear form $a(u)(\phi)$ and some
+right hand side functional $l(\phi)$.
 Let $I:=[0,T]$ be a time interval with end time point value $T$.
-To find some $u$ in a weak form, we define\marginpar{Inkonsistenz: $f(\phi)$ vs. $(f,\phi)$ etwas weiter unten. Ebenso Integrale mal mit $dt$, mal ohne. Waere dafuer, $dt$ wegzulassen.}
+To find some $u$ in a weak form, we define
 \[
-\int_I a(u)(\phi) = \int_I f(\phi) \quad \forall\phi\in X,
+\int_I a(u)(\phi)\, dt = \int_I l(\phi)\, dt \quad \forall\Phi\in X,
 \]
@@
 + (v(0) - v^0, \phi(0))
-= \int_I (f,\phi) \, dt 
+= \int_I l(\phi) \, dt 
 \]
-for all $\phi:= (\phi, \chi) \in X$ with
+for all $\Phi:= (\phi, \chi) \in X$ with
 \[
@@
 +(\nabla\cdot v, \chi)
-=(f,\phi).
+=(f,\phi),
 \end{align*}
+where $f$ describes some volume forces.
 Temporal discretization using the One-step-$\theta$ scheme reads:
@@
 old timestep and present timestep values. 
-Hence, in the local cell equation, a paramter
-is used to distigugish these terms:
+Hence, in the local cell equation, the parameters
+\texttt{scale} and \texttt{$scale_ico$}
+are used to distigugish these terms:
 \begin{lstlisting}
@@
 \subsection{Additional features}
-In this final section, we shall give a brief summary of features 
+Finally, we shall give a brief summary of features 
 that are applicable (in general) to all types of problems. Specifically,
@@
 \label{applications}
-We present three numerical tests which demonstrate different
+In this final section, 
+we present three numerical tests which demonstrate different
 features of \dope{}:
@@
 goal-oriented mesh refinement with the help of the 
-Dual Weighted Residual (DWR) method for the stationary Navier-Stokes equations.
+DWR method for the stationary Navier-Stokes equations.
 \item The second example presents nonstationary fluid-structure 
@@
 \subsection{Nonstationary Fluid-Structure Interaction}
-In this example, the challenging benchmark FSI 2
+In this example, a challenging benchmark FSI 2
 proposed by Hron and Turek \cite{HrTu06b} is considered.
@@
 the left end is attached to the circle. 
-
 %\begin{figure}[h]
@@
 %\end{figure}
-
 Control points $A(t)$ (with $A(0) = (0.6,0.2)$) are fixed at the 

### 2012-11-07 Paper_May_2012/main.tex
Habe versucht in den Seiten 6-8 etwas zu entzerren. Ausserdem gefallen mir die Paragraphen-Ueberschriften nicht, da die kursiv und sehr klein sind. Am liebsten wuerde ich die subsections aufwerten (Section 2.1: Solving of PDEs, Section 2.2: Solving of nonstationary PDEs, Section 2.3: solving of OPT. Dann hatten wir die subsection-Ueberschriften fuer localpde, etc. frei was momentan als Paragraph versteckt ist

@@
 %The \texttt{IntegratorDataContainer} class just makes the constructor interface of 
-Then, the two characteristic features \texttt{LocalPDE} and \texttt{LocalFunctional} objects come into play.
-\begin{lstlisting}
+Then, the two characteristic features \texttt{LocalPDE} and \texttt{LocalFunctional} objects come into play:
+\begin{lstlisting}
+
   LocalPDE<VECTOR> local_pde;
   LocalFunctional<VECTOR> local_functional;
+
 \end{lstlisting}
@@
 \begin{lstlisting}
+
   SpaceTimeHandler<TRIANGULATION, FE> dof_handler(triangulation,
                                                   finite_element);
+
 \end{lstlisting}
@@
 \begin{lstlisting}
+
   PDEProblemContainer<LocalPDE> prob_container(local_pde,
@@
 \begin{lstlisting}
+
   prob_container.AddFunctional(&local_functional);
   prob_container.SetBoundaryFunctionalColors(1);
+
 \end{lstlisting}
@@
 \begin{lstlisting}
+
   DirichletData dd;
   prob_container.SetDirichletBoundaryColors(0, comp_mask, dd);
+
 \end{lstlisting}
@@
 \begin{lstlisting}
+
   typedef StatPDEProblem<NEWTONSOLVER,
            PDEProblemContainer<LocalPDE>, VECTOR, 2> SSOLVER;
+
 \end{lstlisting}
@@
 \begin{lstlisting}
+
   typedef NewtonSolver<LINEARSOLVER, VECTOR, 2> NEWTONSOLVER;
+
 \end{lstlisting}
@@
 \begin{lstlisting}
+
   typedef DirectLinearSolverWithMatrix<MATRIX,
             VECTOR, 2> LINEARSOLVER;
+
 \end{lstlisting}
@@
 \begin{lstlisting}
+
   SSOLVER solver(prob_container, pr,
                     quad_formula, face_quad_formula);
+
 \end{lstlisting}
@@
 \begin{lstlisting}
+
   solver.ReInit();
   solver.ComputeReducedFunctionals();
+
 \end{lstlisting}
@@
    {
-      local_cell_vector(i) +=  scale * (\nabla v^{n+1}, \nabla \phi) + etc.
-
-      local_cell_vector(i) +=  scale_ico * (-p^{n+1},\nabla\cdot v)
+      local_cell_vector(i) +=  
+          scale * (\nabla v^{n+1}, \nabla \phi) + etc.
+
+      local_cell_vector(i) +=  
+          scale_ico * (-p^{n+1},\nabla\cdot v)
    }
@@
    {
-     // implement (u^{n+1},\phi) 
+     local_cell_vector(i) += 
+         scale * (u^{n+1},\phi) + etc. 
    }
@@
 
+\newpage
 \subsection{Compliance minimization}

### 2012-11-08 Paper_May_2012/main.tex
Introduction noch etwas umgeschrieben. Typos entfernt. Und die behandlung von f vereinheitlicht.

@@
 
-\title{DOpElib: A Goal Oriented Software Library for Computing PDEs and Optimization Problems}
+\title{DOpElib: Differential Equations and Optimization Environment; A Goal Oriented Software Library for Solving PDEs and Optimization Problems}
 
@@
 The main feature of \dope{} is that it provides a unified interface to high level algorithms 
-such as time stepping methods, nonlinear solvers and optimization routines. This structure ensures 
-that first of all the user is required to write those sections of code that are specific to 
+such as time-stepping methods, nonlinear solvers and optimization routines. This structure ensures 
+that first of all the user is required to write only those sections of code that are specific to 
 the considered problem. Second, the exchange of parts of the used routines is possible 
-with the need for just a few lines of code to change.
-The article illustrates the performance and various features 
-of the software package and provides some 
-numerical results as demonstration.
+with only a few lines of code to change instead of large reimplementations.
+The article illustrates the design principles and various features
+of the \dope{} and provides some 
+numerical results as demonstration for the versatility of the software.
 \end{abstract}
-
-%\category{C.2.2}{Computer-Communication Networks}{Network Protocols}
-
-%\terms{Design, Algorithms, Performance}
-
-%\keywords{Wireless sensor networks, media access control,
-%multi-channel, radio interference, time synchronization}
-
-%\acmformat{Zhou, G., Wu, Y., Yan, T., He, T., Huang, C., Stankovic,
-%J. A., and Abdelzaher, T. F.  2010. A multifrequency MAC specially
-%designed for  wireless sensor network applications.}
 
@@
 equations (PDEs). Prominent examples are 
-\deal{} \cite{dealnew}, dune \cite{dune} , 
+\deal{} \cite{dealnew}, dune \cite{dune}, 
+\marginpar{fenics ...} 
 or commercial solvers like ANSYS \cite{ansys} or COMSOL \cite{comsol}. 
+\marginpar{fluent ...} 
 In fact, these libraries provide very useful tools 
@@
 However, often, the user is confronted with 
-certain bottlenecks. Either in open-source projects for academic research, 
+certain bottlenecks. On the one hand, 
+in open-source projects for academic research, 
 standard high-level routines 
 which are necessary to solve even simplest PDEs,
-such as linear and nonlinear solvers or timestepping schemes must be implemented 
-by the user himself. However, such routines are problem-independent 
-and could be provided as standard tools. In fact, in many 
-numerical studies, the main task is to compute a  
-target functional which represents a certain error norm,
-point value or traction, for example.
-Second, there is nowadays an increasing interest in computing 
+such as linear and nonlinear solvers or time-stepping schemes must be implemented 
+by the users themselves. However, such routines are, at large,
+problem-independent and could be provided as standard tools. 
+As our guiding design principle, we note that in many 
+numerical studies, the main task is the computation of some  
+target functional, e.g., error norms,
+point values, or certain mean-values as drag or lift coefficients.
+As a second motivating principle, 
+there is nowadays an increasing interest in computing 
 optimization problems in which the state equation is governed 
@@
 parameter estimation, inverse problems or optimal experimental design. 
+To do this in complex applications usually requires tremendous work already
+on the PDE itself before considering optimization of the processes modeled 
+by the PDE. Thus it is advantageous to have an environment that allows to 
+focus on the development of an appropriate code for the PDE at first but 
+asserts due to the required interface, that the code written for the PDE
+is then reusable for the application of state-of-the-art optimization 
+routines.
 So far, there exist a very limited number of open source packages
-which provides standard routines that serve as basis for scientific studies.
-
-On the other hand, 
-commercial toolkits provide these previously mentioned routines but limits user-interaction
-with the source code very often. This is quite unsatisfactory because
-developement of novel numerical algorithms should enable the programmer
+which fulfill the requirements listed above. \marginpar{cite Gascoigne /RoDoBo?}
+
+On the other hand, commercial finite element toolkits provide these 
+previously mentioned routines but limits user-interaction
+with the source code to an extend that typically conflicts
+with the consideration of problems deviating from the intended 
+user-base for the commercial product. 
+This is quite unsatisfactory because development of novel numerical 
+algorithms should enable the researcher
 to change certain parameters or pieces of source code.
 
-The novel software toolkit 
-\textit{Differential Equations and Optimization Environment} (\dope{}) 
-trys to fill this gap. We present a software
-to solve forward PDE
-problems as well as optimization problems constrained by a PDE. 
+The presented toolkit \dope{}: 
+\textit{Differential Equations and Optimization Environment}
+fills this gap. We present a software
+to solve both PDE problems as well as optimization problems 
+constrained by a PDE. 
 Its main feature is to give a unified interface to high level algorithms such as 
-time stepping methods, nonlinear solvers and optimization routines. 
-We aim that the user should only need to write those parts
-of the code that are problem dependent while all invariant parts of the algorithms
-should be reusable without any need for further coding.
-In particular, the user should be able to switch between various different 
-algorithms without the need to rewrite the problem dependent code, though he or she will
+time-stepping methods, nonlinear solvers and optimization routines. 
+\dope{} is designed such that the user only needs to write those parts
+of the code that are problem dependent while all invariant 
+parts of the algorithms
+are reusable without any need for further coding.
+In particular, the user is enabled to switch between various different 
+algorithms without the need to rewrite the problem dependent code, 
+though obviously he or she will
 have to replace the algorithm object with an other one. 
-
-While \deal{} (which might presently seen as the basis of DOpE  
-various finite elements are taken for the discretization) 
-leaves much of the work and many
-decisions to the user, \dope{} intends to be user-optimized by delivering
+This replacement can be done by replacing the appropriate object at only
+one point in the code.
+
+While \deal{} --which at present is the only toolkit to which we provide an 
+interface-- leaves the implementation of all high-level algorithms to the user, 
+\dope{} is user-focused by delivering
 prefabricated tools which require from the user only adjustments connected
 to his specific problem. The solution of optimization problems with PDE
-constraints is an innovation in the \dope{} framework.
-The focus is on the numerical solution of both stationary and nonstationary
-problems which come from different application fields, like elasticity and
-plasticity, fluid dynamics, and fluid-structure interactions.
+constraints is an innovative feature in \dope{}.
 
@@
 \begin{itemize}
-\item Solution of stationary and nonstationary PDEs in 1d, 2d, and 3d.
-\item Various time stepping schemes (based on finite differences), 
+\item Solution of nonlinear stationary and nonstationary PDEs in 1d, 2d, and 3d.
+\item Various, easily exchangeable, 
+  time-stepping schemes (based on finite differences), 
   such as forward Euler, backward Euler,
   Crank-Nicolson, shifted Crank-Nicolson, and Fractional-Step-$\Theta$ scheme.
-\item All finite elements of from \deal{} including hp-support.
+\item Use of all finite elements of from \deal{} including hp-support.
 \item Several examples showing the solution of several PDEs including
    Poisson, Navier-Stokes, Plasticity and fluid-structure interaction problems. 
-\item Self written line search and trust region newton algorithms for the 
-   solution of optimization problems with PDEs \cite{NoWr00}
+\item Line search and trust region newton algorithms for the 
+   solution of optimization problems with PDEs by elimination of 
+   the state variable. \cite{NoWr00} 
+   \marginpar{spaeter zitieren}
 \item Interface to SNOPT for the solution of optimization problems with PDEs and
-  additional other constraints.
-\item Several examples showing how to solve various kinds of optimization problems
-  involving stationary PDE constraints.
-\item Goal-oriented mesh adaptation with the Dual Weighted Residual method, see \cite{BR03}.
-\item Different spatial triangulations for control and state variables.
+  additional other control and state constraints.
+\item Several examples showing how to solve various kinds of optimization 
+  problems involving PDE constraints.
+\item Goal-oriented mesh adaptation with the Dual Weighted Residual method, 
+  following \cite{BR03}.
+\item Different spatial triangulations for control and state variables without
+  the need to reimplement the user dependent code when switching from one common 
+  triangulation to different ones.
 \end{itemize}
@@
 the main features is given. In fact, we first explain the solution 
-of stationary and nonstationary PDEs and give afterwards some 
-ideas how to treat optimization problems. Finally, 
-additional features are discussed. In the final Section
+of stationary PDEs, nonstationary PDEs, and optimization problems. 
+Additional comments are given concerning the required implementation 
+provided by the user.
+Finally, additional features are discussed. In the final Section
 \ref{applications}, we discuss three different numerical 
-examples which substantiates the features of \dope{}.
-
-%The rest of this document is structured as follows: We start with an introduction in
-%Chapter~\ref{chap:intro} where you will learn what is needed to run {\tt \dope{}}. 
-%Further you will learn what problems we can solve and how all the different classes 
-%work together for this purpose. This should help you figure out what the different classes
-%do if you are in need of writing your own algorithm.
-%
-%Then assuming that you can work to your satisfaction with the algorithms already implemented
-%we will show you how to create your own running example in Chapter~\ref{chap:howtoex}.
-%This will be followed by a detailed description of all examples already shipped with 
-%the library. You can find the examples for the solution of PDEs in Chapter~\ref{PDE}
-%and those for the solution of optimization problems with PDEs in Chapter~\ref{OPT}.
-%
-%These notes conclude with a section that explains how we do automated testing of the 
-%implementation in Chapter~\ref{chap:test}. This chapter will be of interest only if you 
-%are trying to implement some new features to the library so that you can check that 
-%the new code did not break anything.
-%
+examples which illustrate the versatile features of \dope{}.
 
@@
 \label{detailed_description}
-%\todo{Aus der Sicht des Users anhand eines konkreten Beispiels beschreiben}
-%\todo{Einfaches Beispiel nehmen}
-%\todo{erwaehnen, dass wir alles immer nicht-linear betrachten}
-%\todo{Aussehen eines typischen PDE und Opt loop und deren Gemeinsamkeiten
-%erklaeren. Vielleicht mir Pseudocode}
 This library is designed to allow easy implementation and numerical solutions 
@@
 part to the user. To show how this is designed we will, in the sequel, 
-describe the reasoning first for a simple stationary PDE and then show
-how this idea extends to non stationary problems as well as to optimization
-problems with PDE constraints.
+describe the reasoning first for a simple stationary PDE in 
+Section~\ref{subsubsec:stationary problems} and then show
+how this idea extends to non stationary problems in 
+Section~\ref{sec:timedep} as well as to optimization
+problems with PDE constraints in Section~\ref{sec:opt}.
 
@@
 \begin{align}\label{eq:prototype_weak}
-a(u)(\phi) = f(\phi) \quad \forall \phi \in V,
+a(u)(\phi) = (f,\phi) \quad \forall \phi \in V,
 \end{align}
 with some suitable space $V$ for the test functions and a 
-right hand side $f$.
+right hand side function $f$.
 To illustrate this further, we take the stationary 
@@
 for all $\phi \in V = H^1_0(\Omega;\R^d) \times L^2_0(\Omega)$.
+Here $(\cdot,\cdot)$ denotes the $L^2$-inner-product.
 The solution will be searched for in $V + (u_d,0)$ where $u_d$ is some 
@@
 interested in one (or several) functionals depending on the solution, i.e., some numbers $J(u)$.
-%For example, if $u$ is the solution of the Navier-Stokes equations \eqref{eq:ns}, we look for the Drag 
-%\begin{align*}
-%J(u)=\int
-%\end{align*}
 
@@
 subdivision $\mathcal T_h$, for instance into quadrilaterals,
-of the domain $\Omega$ and a finite element. In the concrete case we could, 
+of the domain $\Omega$ and a finite element. In the concrete case, we could, 
 for instance, take the $\mathcal Q_2/\mathcal Q_1$ Taylor-Hood element.
@@
 same procedure, we define an object which takes the finite element (FE)
-and the triangulation and takes care of all other things related with 
+and the triangulation and takes care of all other things related to
 handling the finite element space $V_h$. We will refer, to 
-this as the \texttt{SpaceTimeHandler}. It is then clear, that the simplest 
+this as the \texttt{SpaceTimeHandler}. Then, it is clear that the simplest 
 form of the constructor should have the following interface:
@@
 \begin{align}\label{eq:discrete_equation}
-a(u_h)(\phi_h) = f(\phi_h) \quad \forall \phi \in V_h.
+a(u_h)(\phi_h) = (f,\phi_h) \quad \forall \phi \in V_h.
 \end{align}
-To this end one can employ a newton type method with some globalization 
+To this end, one can employ a newton type method with some globalization 
 like line-search. To do so, one needs to calculate the residual of the 
-equation, as well as solving linearized equations for the calculation of 
+equation, as well as to solve linearized equations for the calculation of 
 the newton method and then employing some kind of globalization strategy.
@@
 the $i$-th component of the residual, with respect to the given FE-basis 
-$\phi_{h,i}$, will read as follows
+$\phi_{h}^{i}$, will read as follows
 \begin{align}\label{eq:residual_vector}
- a(u_h)(\phi_{h,i}) - f(\phi_{h,i}) = \sum_{K\in \mathcal T_h} \nu\int_K \nabla u_h, \nabla \phi_{h,i}\,dx + \ldots
+ a(u_h)(\phi_{h}^{i}) - (f,\phi_{h}^{i}) = \sum_{K\in \mathcal T_h} \nu\int_K \nabla u_h\cdot \nabla \phi_{h}^{i}\,dx + \ldots
  \end{align}
@@
 like the description of the PDE and the functional, the boundary values, etc. 
-Although, having such an object instead of directly giving these informations
-to the \texttt{StatPDEProblem} seems like a nuicance at the moment, we will 
+Although, having such an object instead of directly giving these information
+to the \texttt{StatPDEProblem} seems like a nuisance at the moment, we will 
 see that it is of greater interest for time-dependent problems or 
@@
 \item[Discretization] Specify what kind of triangulation,
-% time steppingscheme,
 finite element, quadrature rules, etc. we want to use.
@@
 This is the Benchmark-Case 2D-1 from \cite{TuSchae96}, see there for further details.
-%Additionally, for given inflow $v_{\operatorname{in}}$ we set  boundary conditions 
-%v = v_{\operatorname{in}} \text{ on } \Gamma_{in}\text{ and }v = 0  \text{ on } \Gamma_{wall}.
 \begin{figure}[h]
@@
 After that, we create a triangulation of the domain $\Omega$ named \texttt{triangulation}.
-%This is done exactly as in \deal{}, so we can choose between several possibilities: We could explicitly create a grid by providing a list of vertices and cells, read in a triangulation via \texttt{dealii::GridIn}-class or use the \texttt{dealii::GridGenerator}, which offers common domains like hypercubes or hyper-balls.
 
@@
 cells and faces) and a finite element appropriate for the PDE at hand. We
-choose a Gaussian quadrature of a certain degree and the previosly mentioned
-Tayler-Hood element. These routines are quite standard in each finite element
+choose a Gaussian quadrature of a certain degree and the previously mentioned
+Taylor-Hood element. These routines are quite standard in each finite element
 software and we take advantage to take them from \deal{}.
 
-%IntegratorDataContainer<VECTOR, 2> idc(quad_formula,
-%                                         face_quad_formula);
-%The \texttt{IntegratorDataContainer} class just makes the constructor interface of 
 Then, the two characteristic features \texttt{LocalPDE} and \texttt{LocalFunctional} objects come into play:
@@
 These hold the information of the PDE and the functional under consideration. Further explication is given below because
-the definition of this classes is done spererately in \texttt{localpde.h} resp.   \texttt{localfunctional.h}.
+the definition of this classes is done separately in \texttt{localpde.h} resp.   \texttt{localfunctional.h}.
 
@@
 Now we have everything at hand to create an instantiation of type \texttt{PDEProblemContainer}.
-This basically represents the disrectized problem, and bundles things like boundary conditions, the PDE, the functional and the discretization.
+This basically represents the discretized problem, and bundles things like boundary conditions, the PDE, the functional and the discretization.
 First, we give the constructor the \texttt{local\_pde} and the \texttt{dof\_handler}.
@@
 \end{lstlisting}
-After that, we add the functional under consideration and, because it is a boundaryfunctional, we specify the part of the boundary on which the functional operates (this is done via a 'boundarycolor', $1$ represents $\Gamma_{circ}$).
+After that, we add the functional under consideration and, because it is a boundary functional, we specify the part of the boundary on which the functional operates (this is done via a 'boundary color', $1$ represents $\Gamma_{circ}$).
 \begin{lstlisting}
@@
 \end{lstlisting}
-Next, we incorporate the dirichlet data, which is given by the class \texttt{DirichleData}.
-Then we have to tell the \texttt{prob\_container} for which components and boundaries we prescribe the dirichlet conditions.
+Next, we incorporate the Dirichlet data, which is given by the class \texttt{DirichleData}.
+Then we have to tell the \texttt{prob\_container} for which components and boundaries we prescribe the Dirichlet conditions.
 This is done via a boundary-color (0 stands for $\Gamma_D$) and a \texttt{std::vector<bool> comp\_mask = (true, true, false)}.
@@
 \end{lstlisting}
-Now we create the object which steers the whole solution process, the previosly mentioned \texttt{StatPDEProblem}.
+Now we create the object which steers the whole solution process, the previously mentioned \texttt{StatPDEProblem}.
 It takes care of solving the discretized equation and evaluating the given functionals.
@@
 \end{lstlisting}
-%Finally, the output handler is initialized via
-%\begin{lstlisting}
-%  P.RegisterOutputHandler(&out);
-%  solver.RegisterOutputHandler(&out);
-%\end{lstlisting}
-Finally, the programm is started by
+Finally, the program is started by
 \begin{lstlisting}
@@
 
-%\marginpar{Rekapitulierung des Newton-Verfahrens?}
-Let $N$ be the number of degrees of freedom of our discretization and $\Set{\phi_{h,i}|1\leq i \leq N}$ be a given FE-Basis.
+Let $N$ be the number of degrees of freedom of our discretization and $\Set{\phi_{h}^{i}|1\leq i \leq N}$ be a given FE-Basis.
 As stated in subsection~\ref{subsubsec:stationary problems}, to solve the discrete equation \eqref{eq:discrete_equation} we need the residual, i.e.
@@
 \begin{align}\label{eq:discrete_residual}
-\left(a(u_h)(\phi_{h,i})-(f,\phi_{h,i})\right)_{i=1}^N,
+\left(a(u_h)(\phi_{h}^{i})-(f,\phi_{h}^{i})\right)_{i=1}^N,
 \end{align}
@@
 \begin{align}\label{eq:discrete_matrix}
-\left(\nabla a(u_h)(\phi_{h,j},\phi_{h,i})\right)_{i,j=1}^N.
+\left(\nabla a(u_h)(\phi_{h,j},\phi_{h}^{i})\right)_{i,j=1}^N.
 \end{align}
@@
  The \texttt{CellDataContainer} contains all the information we need to compute the cell-contributions (things like number of quadrature points, number of degrees of freedom on the cell, the \texttt{dealii::FEValues}-objects, etc.).
-%  \begin{lstlisting}
-%      _uvalues.resize(n_q_points,Vector<double>(3));
-%      _ugrads.resize(n_q_points,vector<Tensor<1,2> >(3));
-%      cdc.GetValuesState("last_newton_solution", _uvalues);
-%      cdc.GetGradsState("last_newton_solution", _ugrads);
-%  \end{lstlisting}
   \begin{lstlisting}
@@
 we loop over the quadrature points and the number of DoFs of the cell.
-In the innermost loop, we evaluate the weak form tested with $\phi_{i,h}$ (represented by \texttt{phi\_i\_grads\_v}) on the cell with the help of the previosly selected quadrature formula.
-%     const FEValuesExtractors::Vector velocities(0);
-%     const FEValuesExtractors::Scalar pressure(2); 
+In the innermost loop, we evaluate the weak form tested with $\phi_{i,h}$ (represented by \texttt{phi\_i\_grads\_v}) on the cell with the help of the previously selected quadrature formula.
  \begin{lstlisting}
@@
 \texttt{CellEquation}, so we do not go into the details here.
-% [CellEquation]
-% 
-% \begin{lstlisting}
-%   void CellRightHandSide(const CellDataContainer<2>& cdc,
-%                          dealii::Vector<double> &local_cell_vector,
-%                          double scale);
-% \end{lstlisting}
-% 
-% \begin{lstlisting}
-%   void CellMatrix(const CellDataContainer<2>& cdc,
-%                   dealii::FullMatrix<double> &local_entry_matrix);
-% \end{lstlisting}
-% The PDE under consideration is nonlinear, so let us briefly recapitulate manner such that for solving 
-% the Laplace problem the following nonlinear step is necessary to recapitulate:
-% Find $\delta u\in V$:
-% \[
-% (\nabla \delta u, \nabla \phi) = -(\nabla u, \nabla \phi) 
-% \]
-% and $u^{n+1} = u^n + \delta u$ for $n=1,2,3,ldots$.
-% Therefore, the implementation reads for the right hand side:
-% \begin{lstlisting}
-%    void CellEquation(...)
-%     {
-%       // implement (\nabla u, \nabla \phi) 
-%     }
-% \end{lstlisting}
-% Specifically, the inner loops reads
-% \begin{lstlisting}
-%    for (all quadrature points per cell q)
-%       {
-%         Tensor<2, 2> ugrads; // solution vector
-%  
-% 	for (all dofs per cell i)
-%         {
-% 	  const Tensor<2, 2> phi_i_grads_u; // test function
-% 
-%           local_cell_vector(i) += scale * 
-%             scalar_product(ugrads, phi_i_grads_u) *
-% 	    state_fe_values.JxW(q_point);
-%         }
-%       }
-% \end{lstlisting}
-% For the left hand side, we need to implement the Jacobian matrix:
-% \begin{lstlisting}
-%    void CellMatrix(...)
-%     {
-%       // implement  (\nabla \delta u, \nabla \phi)
-%     }
-% \end{lstlisting}
-% Specifically,
-% \begin{lstlisting}
-%       for (all quadrature points per cell q)      
-%         for (all dofs per cell i)    
-%           for (all dofs per cell j)
-%           {
-%             local_entry_matrix(i, j) += 
-%               scalar_product(phi_grads_u[j], phi_grads_u[i]) * 
-%               state_fe_values.JxW(q_point);
-%           }
-% \end{lstlisting}
-% The integration of the equation for each DoF and each quadrature point on each
-% cell is then done 
-% in the \texttt{templates/integrator.h} function. Finally, as usually in 
-% finite element computations, this leads to a discrete system. Since
-% we treat each problem with a Newton method, within each Newton step, the 
-% following linear system must be solved:
-% \[
-% A\delta u = b
-% \]
-% in which $A$ is specified by \texttt{CellMatrix(...)} and 
-% $b$ by \texttt{CellEquation(...)}. The vector $\delta u$ denotes the solution to the 
-% present Newton step:
-% \[
-% u^n = u^{n-1} + \omega \delta u
-% \]
-% with a line search parameter $\omega\in (0,1]$.
 
@@
 \end{lstlisting}
-
-% The target functional 
-% is now implemented in \texttt{functional.h} and 
-% treats a point evaluation:
-% \begin{lstlisting}
-%   double PointValue(...)
-%    {
-%     Point<2> p1(0.5, 0.5);
-%     Vector<double> tmp_vector(2);
-%     // more stuff is necessary here
-%     VectorTools::point_value (...);
-%     double x = tmp_vector(0);
-% 
-%     return x;
-%    }
-% \end{lstlisting}
-
-% TWick:
-% To solve the problem, we need a configuration (a mesh), eventually 
-% a diffusion parameter, a finite element, a quadrature rule,
-% a linear solver (here the CG solver), and some output such as
-% a graphical solution and eventually some measurement functional.
-% Each of those routines are written in the \texttt{main.cc} file. 
-% We start by including necessary files both the \dope{} library and
-% the \deal{} software:
-% \begin{lstlisting}
-%   #include "pdeproblemcontainer.h" etc. ...
-% \end{lstlisting}
-% In the \texttt{main} function, the following steps need to be done. First, 
-% the parameter reader for runtime parameters might be initialized. In a parameter
-% file the user might specify material parameters for the concrete PDE, but also
-% general parameters for nonlinear and linear solvers, and output-related 
-% aspects.
-% 
-% A linear finite element in two dimensions and two components is 
-% initialized:
-% \begin{lstlisting}
-%   FESystem<2> state_fe(FE_Q<2>(1), 2);
-% \end{lstlisting}
-% and a quadrature rule of order three is used to approximate the 
-% integrals:
-% \begin{lstlisting}
-%   QGauss<2> quadrature_formula(3);
-% \end{lstlisting}
-% Then, the \texttt{LocalPDE} object comes into play. This is an quite 
-% important feature of \dope{} because it allows for easy implementation 
-% of the variational formulation. Specific explication is given below because
-% the implementation is separated in \texttt{localpde.h}.
-% Afterwards, a \texttt{LocalPointFunctional} is declared. As for the PDE, the 
-% implementation is done in another file, namely \texttt{localfunctional.h}.
-% 
-% Generically, each problem in \dope{} is considered to be time-dependent and nonlinear because
-% these are the 'real' world problems. Therefore, (even for the linear and stationary Laplace equation),
-% we have to declare a (pseudo) time:
-% \begin{lstlisting}
-%   std::vector<double> times(1, 0.);
-% \end{lstlisting}
-% Then, the spatial mesh (here the unit square) is created and three-times globally-refined:
-% \begin{lstlisting}
-%   GridGenerator::hyper_cube(triangulation, 0, 1);
-%   triangulation.refine_global(3);
-% \end{lstlisting}
-% A major component which organizes the whole solution process (independent from 
-% the problem!) is the 
-% \begin{lstlisting}
-%   MethodOfLines_StateSpaceTimeHandler<...> dof_handler (...);
-% \end{lstlisting}
-% This method handles all dofs in space and time and different versions exist
-% for forward as well as optimization problems.
-% 
-% Afterwards, the problem (here the Laplace PDE) is given to the \texttt{dof_handler} via
-% \begin{lstlisting}
-%   OP P(local_pde, dof_handler);
-% \end{lstlisting}
-% It remains to declare boundary conditions for the problem under consideration. 
-% Let's simply take zero Dirichlet conditions on the boundary:
-% \begin{lstlisting}
-%   std::vector<bool> comp_mask(2);
-% 
-%   comp_mask[0] = true;
-%   comp_mask[1] = true;
-% 
-%   DOpEWrapper::ZeroFunction<2> zf(2);
-%   SimpleDirichletData<VECTOR, 2> DD1(zf);
-% 
-%   P.SetDirichletBoundaryColors(0, comp_mask, &DD1);
-% \end{lstlisting}
-% Next, the solver for the linear equations needs to be specified:
-% \begin{lstlisting}
-%   SSolver solver(&P, "fullmem", pr, idc);
-% \end{lstlisting}
-% Finally, the output handler is initialized via
-% \begin{lstlisting}
-%   P.RegisterOutputHandler(&out);
-%   solver.RegisterOutputHandler(&out);
-% \end{lstlisting}
-% Finally, the programm is started with:
-% \begin{lstlisting}
-%   solver.ReInit();
-%   out.ReInit();
-%   out.Write(outp, 1, 1, 1);
-%   solver.ComputeReducedFunctionals();
-% \end{lstlisting}
-% After explaining the basic compononents that are necessary (and always the same in \dope{}) 
-% for all  (linear and stationary) partial differential equations, we shall give 
-% an overview to the implementation of the specific form of the PDE. 
-% Beside specification of boundary conditions, the form of the local PDE depends 
-% on the concrete problem under consideration. 
-% This step is done in the \texttt{localpde.h} file. As already explained,
-% each PDE is treatet in a nonlinear manner such that for solving 
-% the Laplace problem the following nonlinear step is necessary to recapitulate:
-% Find $\delta u\in V$:
-% \[
-% (\nabla \delta u, \nabla \phi) = -(\nabla u, \nabla \phi) 
-% \]
-% and $u^{n+1} = u^n + \delta u$ for $n=1,2,3,ldots$.
-% Therefore, the implementation reads for the right hand side:
-% \begin{lstlisting}
-%    void CellEquation(...)
-%     {
-%       // implement (\nabla u, \nabla \phi) 
-%     }
-% \end{lstlisting}
-% Specifically, the inner loops reads
-% \begin{lstlisting}
-%    for (all quadrature points per cell q)
-%       {
-%         Tensor<2, 2> ugrads; // solution vector
-%  
-% 	for (all dofs per cell i)
-%         {
-% 	  const Tensor<2, 2> phi_i_grads_u; // test function
-% 
-%           local_cell_vector(i) += scale * 
-%             scalar_product(ugrads, phi_i_grads_u) *
-% 	    state_fe_values.JxW(q_point);
-%         }
-%       }
-% \end{lstlisting}
-% For the left hand side, we need to implement the Jacobian matrix:
-% \begin{lstlisting}
-%    void CellMatrix(...)
-%     {
-%       // implement  (\nabla \delta u, \nabla \phi)
-%     }
-% \end{lstlisting}
-% Specifically,
-% \begin{lstlisting}
-%       for (all quadrature points per cell q)      
-%         for (all dofs per cell i)    
-%           for (all dofs per cell j)
-%           {
-%             local_entry_matrix(i, j) += 
-%               scalar_product(phi_grads_u[j], phi_grads_u[i]) * 
-%               state_fe_values.JxW(q_point);
-%           }
-% \end{lstlisting}
-% The integration of the equation for each DoF and each quadrature point on each
-% cell is then done 
-% in the \texttt{templates/integrator.h} function. Finally, as usually in 
-% finite element computations, this leads to a discrete system. Since
-% we treat each problem with a Newton method, within each Newton step, the 
-% following linear system must be solved:
-% \[
-% A\delta u = b
-% \]
-% in which $A$ is specified by \texttt{CellMatrix(...)} and 
-% $b$ by \texttt{CellEquation(...)}. The vector $\delta u$ denotes the solution to the 
-% present Newton step:
-% \[
-% u^n = u^{n-1} + \omega \delta u
-% \]
-% with a line search parameter $\omega\in (0,1]$.
-% 
-% 
-% The target functional 
-% is now implemented in \texttt{functional.h} and 
-% treats a point evaluation:
-% \begin{lstlisting}
-%   double PointValue(...)
-%    {
-%     Point<2> p1(0.5, 0.5);
-%     Vector<double> tmp_vector(2);
-%     // more stuff is necessary here
-%     VectorTools::point_value (...);
-%     double x = tmp_vector(0);
-% 
-%     return x;
-%    }
-% \end{lstlisting}
 
@@
 \end{lstlisting}
-At present we have prepared the following time stepping schemes in \dope:
+At present we have prepared the following time-stepping schemes in \dope:
 \begin{itemize}
@@
 \end{itemize}
-With this we can now run a loop over all timesteps, where in each of the 
-timesteps we can reuse the functionality to solve stationary PDEs 
+With this we can now run a loop over all time-steps, where in each of the 
+time-steps we can reuse the functionality to solve stationary PDEs 
 applied to the time-step problem.
@@
 Temporal discretization using the One-step-$\theta$ scheme reads:
-Given the old timestep solution $(v^n,p^n)$, 
+Given the old time-step solution $(v^n,p^n)$, 
 find $(v^{n+1}, p^{n+1})$ (and multiplication 
-by the inverse of the timestep $k$):
+by the inverse of the time-step $k$):
 \begin{align*}
@@
 The concrete choice of $\theta$ leads to a
-specific time stepping scheme. For instance,
+specific time-stepping scheme. For instance,
 $\theta = 0.5$ gives the Crank-Nicolson scheme,
@@
 
-The pressure and incompressibily terms should be 
+The pressure and incompressibility terms should be 
 treated fully implicitly when time-discretizing 
@@
 terms and remaining terms for which we need 
-old timestep and present timestep values. 
+old time-step and present time-step values. 
 Hence, in the local cell equation, the parameters
 \texttt{scale} and \texttt{$scale_ico$}
-are used to distigugish these terms:
+are used to distinguish these terms:
 \begin{lstlisting}
@@
 \end{lstlisting}
-Analogoues procedure is performed in the local 
+Analogous procedure is performed in the local 
 cell matrix.
@@
 \end{lstlisting}
-Here, the advantage is that \dope{} orders the terms in the timestepping
+Here, the advantage is that \dope{} orders the terms in the time-stepping
 problems
@@
 However, if someone 
-needs explicit input of the old timestep term, then he can use 
+needs explicit input of the old time-step term, then he can use 
 \begin{lstlisting}
@@
 Fractional-Step-$\theta$-scheme in which 
-each time step is subdivided into three 
-sub-timeintervals. 
+each time-step is subdivided into three 
+sub-steps. 
 \end{remark}
@@
 a stationary PDE. If one would like to use derivative based methods, one 
-needs in adition derivatives of $j$ and thus of $J$ as well as of $S$. 
+needs in addition derivatives of $j$ and thus of $J$ as well as of $S$. 
 To do so, we remark, that one can calculate the derivative of 
@@
 \]
-where the index $q$ or $u$ denote directional derivatieves for the 
+where the index $q$ or $u$ denote directional derivatives for the 
 respective argument. To calculate action of the adjoint operator $S^*$ 
@@
 Dual Weighted Residual (DWR) method, see \cite{BeRa96} or \cite{BR03}. Here, a certain 
-cost functional such as a point evaluation, line intergration or domain
+cost functional such as a point evaluation, line integration or domain
 integration is considered. Determining the lowest error between the 
@@
 for the state and the control \dope{} provides multi-mesh support, meaning that
-onec a problem has been written correctly using the same mesh for all finite element
+once a problem has been written correctly using the same mesh for all finite element
 variables one can switch to a different SpaceTimeHandler that allows different meshes for all 
@@
 allows to consider goal-oriented mesh refinement and gradient-based optimization.
-\item In the third numerical test, an optmization problem for structural mechanics
+\item In the third numerical test, an optimization problem for structural mechanics
 is discussed.
@@
 namely, the continuity of velocities and continuity of normal stresses
-are satisfied in each time step. This is achieved with the help of 
+are satisfied in each time-step. This is achieved with the help of 
 a monolithic coupling scheme (strong coupling).
@@
 the left end is attached to the circle. 
-%\begin{figure}[h]
-%\centering
-%\input{Pictures/pic_12}
-%\caption{Flow around cylinder with elastic beam with 
-%circle-center $C=(0.2,0.2)$ and radius $r=0.05$.}
-%\label{configuration_csm_and_fsi_2D}
-%\end{figure}
 Control points $A(t)$ (with $A(0) = (0.6,0.2)$) are fixed at the 
@@
 The Fractional-Step-$\theta$ scheme is used for time discretization with
-different time step sizes $k$. However, the time stepping scheme can be 
+different time step sizes $k$. However, the time-stepping scheme can be 
 very easily chosen in the \texttt{main.cc} function by choosing an appropriate 
-time stepping scheme, at present the following are availiable.
+time-stepping scheme, at present the following are available.
 \begin{lstlisting}

### 2012-11-08 Paper_May_2012/main.tex
Paper: In der Einleitung Winnis Vorschlage zur Referenzierung anderer Software eingearbeitet und zitiert

@@
 \deal{} \cite{dealnew}, dune \cite{dune}, 
-\marginpar{fenics ...} 
-or commercial solvers like ANSYS \cite{ansys} or COMSOL \cite{comsol}. 
-\marginpar{fluent ...} 
+FEniCS \cite{fenics}
+or commercial solvers like ANSYS (with its Fluent package) \cite{ansys},
+COMSOL \cite{comsol} or simufact forming \cite{simufact}.
+Finally, often, a lot of work is (still) done in Matlab \cite{matlab}.
 In fact, these libraries provide very useful tools 
@@
 So far, there exist a very limited number of open source packages
-which fulfill the requirements listed above. \marginpar{cite Gascoigne /RoDoBo?}
+which fulfill the requirements listed above; 
+sucessful in-house code examples are Gascoigne \cite{gascoigne}
+and RoDoBo \cite{rodobo}. 
+Both have been initiated as well in 
+the Numerical Analysis Group in Heidelberg.
 

### 2012-11-08 Paper_May_2012/main.tex
Paper: Sections aufgewertet

@@
 
-\subsection{Solving of PDEs}
-\subsubsection{Stationary problems}\label{subsubsec:stationary problems}
+\subsection{Solving of Stationary PDEs}
+\label{subsubsec:stationary problems}
 For the start we consider a prototypical problem given by the solution of 
@@
 
-\paragraph{\texttt{main.cc}}
+\subsubsection{\texttt{main.cc}}
 In this file we gather the basic components that are necessary for the
@@
 
-\paragraph{\texttt{localpde.h}}
+\subsubsection{\texttt{localpde.h}}
 We shall now give an overview to the implementation of the specific form of the PDE.
@@
 
-\paragraph{\texttt{localfunctional.h}}
+\subsubsection{\texttt{localfunctional.h}}
 The implementation of the boundary-functional $J$ follows the outline of the computation of the cell-residual in \texttt{localpde.h}.
@@
 
-\subsubsection{Nonstationary problems}\label{sec:timedep}
+\subsection{Solving Nonstationary PDEs}\label{sec:timedep}
 To describe the solution of nonstationary 

### 2012-11-08 Paper_May_2012/main.tex
Templates und Code angeglichen, trotzdem noch unterschiede im coding-style, allerdings wird der Code von vorne nach hinten immer sparsamer, was ja durchaus zu begruenden ist. Eventuell vorne noch mehr kuerzen?

@@
 \begin{lstlisting}
-      void StatPDEProblem::ComputeReducedFunctionals();
-\end{lstlisting}
+  template<class NEWTON>
+      void StatPDEProblem<NEWTON>::ComputeReducedFunctionals();
+\end{lstlisting}
+The template argument \texttt{NEWTON} represents the choosen solver of the nonlinear problems.
 That combines all the necessary steps to calculate $J(u_h)$.
@@
 
-  LocalPDE<VECTOR> local_pde;
-  LocalFunctional<VECTOR> local_functional;
+  LocalPDE local_pde;
+  LocalFunctional local_functional;
 
@@
 
-  SpaceTimeHandler<TRIANGULATION, FE> dof_handler(triangulation,
-                                                  finite_element);
+  SpaceTimeHandler dof_handler(triangulation, finite_element);
 
@@
                                                dof_handler);
+
 \end{lstlisting}
@@
 
-  typedef StatPDEProblem<NEWTONSOLVER,
-           PDEProblemContainer<LocalPDE>, VECTOR, 2> SSOLVER;
+  typedef StatPDEProblem<NEWTONSOLVER> SSOLVER;
 
@@
 
-  typedef NewtonSolver<LINEARSOLVER, VECTOR, 2> NEWTONSOLVER;
-
-\end{lstlisting}
-The first template parameter of \texttt{NewtonSolver} describes which linear solver we use in this example. We opted for a direct solver, so: 
-\begin{lstlisting}
-
-  typedef DirectLinearSolverWithMatrix<MATRIX,
-            VECTOR, 2> LINEARSOLVER;
+  typedef NewtonSolver<LINEARSOLVER> NEWTONSOLVER;
+
+\end{lstlisting}
+The  template parameter \texttt{LINEARSOLVER} describes which linear solver we use in this example. We opted for a direct solver, so: 
+\begin{lstlisting}
+
+  typedef DirectLinearSolver<MATRIX, VECTOR> LINEARSOLVER;
 
@@
 Hence, in the local cell equation, the parameters
-\texttt{scale} and \texttt{$scale_ico$}
+\texttt{scale} and \texttt{scale\_ico}
 are used to distinguish these terms:
@@
    {
+   // Usual loops over DoFs and quadrature points
+     {
       local_cell_vector(i) +=  
-          scale * (\nabla v^{n+1}, \nabla \phi) + etc.
-
-      local_cell_vector(i) +=  
-          scale_ico * (-p^{n+1},\nabla\cdot v)
+          scale *// terms like (\nabla v^{n+1}, \nabla \phi) etc.
+          + scale_ico *// terms like (-p^{n+1},\nabla\cdot \phi)
+     }
    }
@@
    {
-     local_cell_vector(i) += 
-         scale * (u^{n+1},\phi) + etc. 
+      //implement (u^{n+1},\phi)
    }
 \end{lstlisting}
-and it is done in a straight-forward way:
-\begin{lstlisting}
-  for (all quadrature points per cell q)
+% \begin{lstlisting}
+%   void CellTimeEquation(...)
+%    {
+%      local_cell_vector(i) += 
+%          scale * (u^{n+1},\phi) + etc. 
+%    }
+% \end{lstlisting}
+and it is done in a straight-forward way:\marginpar{Mir gefaellt dieses Auseinanderziehen nicht so wirklich...ist dieser zweite Abschnitt heir notwendig?}
+\begin{lstlisting}
+  for (/*all quadrature points per cell q*/)
    {
@@
 
-     for (all dofs per cell i)
+     for (/*all dofs per cell i*/)
       {
-        const Tensor<1,2> phi_i_u = ...;
-
-        local_cell_vector(i) +=  
-          scale * (u * phi_i_u) * 
+        Tensor<1,2> phi_i_u = ...;
+        local_cell_vector(i) +=  scale * (u * phi_i_u) * 
                                   state_fe_values.JxW(q_point);    

### 2012-11-09 Paper_May_2012/main.tex
Kurzer Kommentar zu reusability von Code im stationaeren/instationaeren Fall am Beispiel der CellEquation eingefuegt, zudem den von Thomas in der Mail angesprochenen Abschnitt abgeaendert.

@@
 cell matrix.
+\begin{remark}
+Note that in the stationary case \dope{} automatically sets $ \texttt{scale\_ico} = \texttt{scale}$, so that in fact the user has to implement the \texttt{CellEquation} only once and can henceforth reuse it in all his stationary an instationary problems.
+This transfers of course also to other methods like \texttt{CellMatrix}.
+\end{remark}
 
 In addition to stationary problems, the implementation
-of the time-derivative must be considered which is done
-in 
+of the time-derivative, i.e. the term $ (v^{n+1},\phi)$, must be considered which is done
+in the method \texttt{CellTimeEquation} in a straight-forward way.
+After the initialization of $v$ and the shape function $\phi_i$ and the usual loops over the quadrature points and  degrees of freedom of the cell we fill the local cell vector by
 \begin{lstlisting}
-  void CellTimeEquation(...)
-   {
-      //implement (u^{n+1},\phi)
-   }
-\end{lstlisting}
-% \begin{lstlisting}
-%   void CellTimeEquation(...)
-%    {
-%      local_cell_vector(i) += 
-%          scale * (u^{n+1},\phi) + etc. 
-%    }
-% \end{lstlisting}
-and it is done in a straight-forward way:\marginpar{Mir gefaellt dieses Auseinanderziehen nicht so wirklich...ist dieser zweite Abschnitt heir notwendig?}
-\begin{lstlisting}
-  for (/*all quadrature points per cell q*/)
-   {
-     Tensor<1,2> u = ...;
-
-     for (/*all dofs per cell i*/)
-      {
-        Tensor<1,2> phi_i_u = ...;
-        local_cell_vector(i) +=  scale * (u * phi_i_u) * 
+        local_cell_vector(i) +=  scale * (v * phi_i) * 
                                   state_fe_values.JxW(q_point);    
-      }
-   }
 \end{lstlisting}
@@
 problems
-such that only $(u^{n+1},\phi)$ must be implemented.
+such that only $(v^{n+1},\phi)$ must be implemented.
 
@@
    {
-     // implement (u^{n+1},\phi) and (u^n,\phi)  
+     // implement (v^{n+1},\phi) and (v^n,\phi)  
    }

### 2012-11-09 Paper_May_2012/main.tex
Die abgesprochenen Zitierungsaenderungen und jetzt auch ein Zitat von der DOpElib Webseite

@@
 FEniCS \cite{fenics}
-or commercial solvers like ANSYS (with its Fluent package) \cite{ansys},
-COMSOL \cite{comsol} or simufact forming \cite{simufact}.
-Finally, often, a lot of work is (still) done in Matlab \cite{matlab}.
+or commercial solvers as for instance ANSYS (with its Fluent package),
+COMSOL, or simufact forming.
+Finally, often, a lot of work is (still) done in Matlab.
 In fact, these libraries provide very useful tools 
@@
 The presented toolkit \dope{}: 
-\textit{Differential Equations and Optimization Environment}
+\textit{Differential Equations and Optimization Environment} \cite{dope}
 fills this gap. We present a software
@@
    solution of optimization problems with PDEs by elimination of 
-   the state variable. \cite{NoWr00} 
-   \marginpar{spaeter zitieren}
+   the state variable.
 \item Interface to SNOPT for the solution of optimization problems with PDEs and
@@
   problems involving PDE constraints.
-\item Goal-oriented mesh adaptation with the Dual Weighted Residual method, 
-  following \cite{BR03}.
+\item Goal-oriented mesh adaptation with the Dual Weighted Residual method and
+  standard residual based error estimators.
 \item Different spatial triangulations for control and state variables without
@@
 \end{equation}
-Similarly formulas for the second derivatives can be obtained.
+Similarly formulas for the second derivatives can be obtained, see, e.g., 
+\cite{BeMeVe06}.
 
@@
 \end{lstlisting}
+
+Based on this classes we can now solve the optimization problem, for instance using 
+a Newton type method for the reduced problem, see, e.g., \cite{NoWr00}. This yields 
+an object \texttt{ReducedNewtonAlgorithm} which needs to be initialized 
+with the \texttt{OptProblemContainer} and then provides 
+the following method to be called by the user
+\begin{lstlisting}
+  ReducedNewtonAlgorithm::Solve(ControlVector<VECTOR>& )
+\end{lstlisting}
+where the argument provides an initial guess for the optimal control.
 

### 2012-11-12 Examples/Documentation/introduction.tex

@@
 \section{License information}
-{\bf TODO:} Unter welcher License soll DOpE laufen?
+Copyright (C) 2012 by the DOpElib authors\\[2mm]
+%
+This file is part of DOpElib\\[2mm]
+%
+DOpElib is free software: you can redistribute it
+and/or modify it under the terms of the GNU General Public
+License as published by the Free Software Foundation, either
+version 3 of the License, or (at your option) any later
+version.\\[2mm]
+%
+DOpElib is distributed in the hope that it will be
+useful, but WITHOUT ANY WARRANTY; without even the implied
+warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+PURPOSE.  See the GNU General Public License for more
+details.\\[2mm]
+%
+Please refer to the file LICENSE.TXT included in this distribution
+for further information on this license.
 
@@
   \item Christian Goll (Heidelberg University)
-  \item Thomas Wick (Heidelberg University)
+  \item Thomas Wick (The University of Texas at Austin)
   \item Winnifried Wollner (University of Hamburg)

### 2012-11-12 DOpEsrc/include/helloworld.h

@@ -1,3 +1,26 @@
+/**
+*
+* Copyright (C) 2012 by the DOpElib authors
+*
+* This file is part of DOpElib
+*
+* DOpElib is free software: you can redistribute it
+* and/or modify it under the terms of the GNU General Public
+* License as published by the Free Software Foundation, either
+* version 3 of the License, or (at your option) any later
+* version.
+*
+* DOpElib is distributed in the hope that it will be
+* useful, but WITHOUT ANY WARRANTY; without even the implied
+* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+* PURPOSE.  See the GNU General Public License for more
+* details.
+*
+* Please refer to the file LICENSE.TXT included in this distribution
+* for further information on this license.
+*
+**/
+
 namespace DOpE
 {
   class Hello

### 2012-11-12 DOpEsrc/include/sparsitymaker.h

@@ -1,3 +1,26 @@
+/**
+*
+* Copyright (C) 2012 by the DOpElib authors
+*
+* This file is part of DOpElib
+*
+* DOpElib is free software: you can redistribute it
+* and/or modify it under the terms of the GNU General Public
+* License as published by the Free Software Foundation, either
+* version 3 of the License, or (at your option) any later
+* version.
+*
+* DOpElib is distributed in the hope that it will be
+* useful, but WITHOUT ANY WARRANTY; without even the implied
+* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+* PURPOSE.  See the GNU General Public License for more
+* details.
+*
+* Please refer to the file LICENSE.TXT included in this distribution
+* for further information on this license.
+*
+**/
+
 /*
  * sparsitymaker.h
  *

### 2012-11-12 DOpEsrc/interfaces/active_fe_index_setter_interface.h

@@ -1,3 +1,26 @@
+/**
+*
+* Copyright (C) 2012 by the DOpElib authors
+*
+* This file is part of DOpElib
+*
+* DOpElib is free software: you can redistribute it
+* and/or modify it under the terms of the GNU General Public
+* License as published by the Free Software Foundation, either
+* version 3 of the License, or (at your option) any later
+* version.
+*
+* DOpElib is distributed in the hope that it will be
+* useful, but WITHOUT ANY WARRANTY; without even the implied
+* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+* PURPOSE.  See the GNU General Public License for more
+* details.
+*
+* Please refer to the file LICENSE.TXT included in this distribution
+* for further information on this license.
+*
+**/
+
 /*
  * active_fe_index_interface.h
  *

### 2012-11-12 DOpEsrc/interfaces/transposeddirichletdatainterface.h

@@ -1,3 +1,26 @@
+/**
+*
+* Copyright (C) 2012 by the DOpElib authors
+*
+* This file is part of DOpElib
+*
+* DOpElib is free software: you can redistribute it
+* and/or modify it under the terms of the GNU General Public
+* License as published by the Free Software Foundation, either
+* version 3 of the License, or (at your option) any later
+* version.
+*
+* DOpElib is distributed in the hope that it will be
+* useful, but WITHOUT ANY WARRANTY; without even the implied
+* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+* PURPOSE.  See the GNU General Public License for more
+* details.
+*
+* Please refer to the file LICENSE.TXT included in this distribution
+* for further information on this license.
+*
+**/
+
 #ifndef _TRANSPOSED_DIRICHLET_INTERFAC_H_
 #define _TRANSPOSED_DIRICHLET_INTERFAC_H_
 

### 2012-11-12 DOpEsrc/problemdata/tangentdirichletdata.h

@@ -1,3 +1,26 @@
+/**
+*
+* Copyright (C) 2012 by the DOpElib authors
+*
+* This file is part of DOpElib
+*
+* DOpElib is free software: you can redistribute it
+* and/or modify it under the terms of the GNU General Public
+* License as published by the Free Software Foundation, either
+* version 3 of the License, or (at your option) any later
+* version.
+*
+* DOpElib is distributed in the hope that it will be
+* useful, but WITHOUT ANY WARRANTY; without even the implied
+* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+* PURPOSE.  See the GNU General Public License for more
+* details.
+*
+* Please refer to the file LICENSE.TXT included in this distribution
+* for further information on this license.
+*
+**/
+
 #ifndef _TANGENT_DIRICHLET_DATA_H_
 #define _TANGENT_DIRICHLET_DATA_H_
 

### 2012-11-12 DOpEsrc/tsschemes/backward_euler_problem.h

@@ -1,3 +1,26 @@
+/**
+*
+* Copyright (C) 2012 by the DOpElib authors
+*
+* This file is part of DOpElib
+*
+* DOpElib is free software: you can redistribute it
+* and/or modify it under the terms of the GNU General Public
+* License as published by the Free Software Foundation, either
+* version 3 of the License, or (at your option) any later
+* version.
+*
+* DOpElib is distributed in the hope that it will be
+* useful, but WITHOUT ANY WARRANTY; without even the implied
+* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+* PURPOSE.  See the GNU General Public License for more
+* details.
+*
+* Please refer to the file LICENSE.TXT included in this distribution
+* for further information on this license.
+*
+**/
+
 #ifndef _BackwardEulerProblem_H_
 #define _BackwardEulerProblem_H_
 

### 2012-11-12 DOpEsrc/wrapper/preconditioner_wrapper.h

@@ -1,3 +1,26 @@
+/**
+*
+* Copyright (C) 2012 by the DOpElib authors
+*
+* This file is part of DOpElib
+*
+* DOpElib is free software: you can redistribute it
+* and/or modify it under the terms of the GNU General Public
+* License as published by the Free Software Foundation, either
+* version 3 of the License, or (at your option) any later
+* version.
+*
+* DOpElib is distributed in the hope that it will be
+* useful, but WITHOUT ANY WARRANTY; without even the implied
+* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+* PURPOSE.  See the GNU General Public License for more
+* details.
+*
+* Please refer to the file LICENSE.TXT included in this distribution
+* for further information on this license.
+*
+**/
+
 #ifndef _DOPE_PRECONDITIONER_H_
 #define _DOPE_PRECONDITIONER_H_
 

### 2012-11-12 Examples/OPT/StatPDE/Example6/main.cc

@@ -1,3 +1,26 @@
+/**
+*
+* Copyright (C) 2012 by the DOpElib authors
+*
+* This file is part of DOpElib
+*
+* DOpElib is free software: you can redistribute it
+* and/or modify it under the terms of the GNU General Public
+* License as published by the Free Software Foundation, either
+* version 3 of the License, or (at your option) any later
+* version.
+*
+* DOpElib is distributed in the hope that it will be
+* useful, but WITHOUT ANY WARRANTY; without even the implied
+* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+* PURPOSE.  See the GNU General Public License for more
+* details.
+*
+* Please refer to the file LICENSE.TXT included in this distribution
+* for further information on this license.
+*
+**/
+
 #include "generalized_mma_algorithm.h"
 #include "optproblemcontainer.h"
 #include "functionalinterface.h"

### 2012-11-12 Examples/PDE/StatPDE/Example9/myfunctions.h

@@ -1,3 +1,26 @@
+/**
+*
+* Copyright (C) 2012 by the DOpElib authors
+*
+* This file is part of DOpElib
+*
+* DOpElib is free software: you can redistribute it
+* and/or modify it under the terms of the GNU General Public
+* License as published by the Free Software Foundation, either
+* version 3 of the License, or (at your option) any later
+* version.
+*
+* DOpElib is distributed in the hope that it will be
+* useful, but WITHOUT ANY WARRANTY; without even the implied
+* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+* PURPOSE.  See the GNU General Public License for more
+* details.
+*
+* Please refer to the file LICENSE.TXT included in this distribution
+* for further information on this license.
+*
+**/
+
 /*
  * myfunctions.h
  *

### 2012-11-14 Paper_May_2012/main_tw.tex
Paper: Ich habe meine Anmerkungen zur jetztigen Version in ein eigens file eingecheckt. Alles ist im Text mit \todo{...} vermerkt

--- 2012-11-09 Paper_May_2012/main.tex
@@
 
-\newcommand{\todo}[1]{\textbf{\textsc{\textcolor{black}{TODO: #1}}}}
+\newcommand{\todo}[1]{\textbf{\textsc{\textcolor{blue}{TODO: #1}}}}
 %\newcommand{\mymarginpar}[1]{\marginpar{\textcolor{red}{\scriptsize{#1}}}}
@@
 there are several different sparsity patterns available in 
-\deal{} which the user should be able chose. Thus it needs to be 
+\todo{to choose}
+\deal{} which the user should be able to choose. Thus it needs to be 
 given to the \texttt{SpaceTimeHandler}. However, to keep the 
@@
 \item[Discretization] Specify what kind of triangulation,
-finite element, quadrature rules, etc. we want to use.
-\item[Problem-data] Define the weak form, the corresponding matrices, the functionals as well as the boundary conditions.
-With exception of the boundary conditions, we have to specify all this only cell-wise, \dope{} handles the rest for us.
-\item[Solvers] Which (non-)linear solver should we use? 
+finite element, quadrature rules, 
+\todo{etc. strange to me} etc. we want to use.
+\item[Problem-data] 
+Define the weak form, 
+the corresponding matrices, 
+the functionals \todo{$J(u)$} as well as the boundary conditions. With exception of the boundary conditions, we have to specify all this only cell-wise, \dope{} handles the rest for us.
+\item[Solvers] \todo{Choice of a (non-)linear solver.}
 \item[Additional features] ParameterReader to read in the run-time parameters
   such as material coefficients or adjustments for the solvers like tolerance
-  for Newton's method. Second, definition of Outputhandler etc.
+  for Newton's method. 
+\todo{Sentence strang to me} Second, definition of Outputhandler etc.
 \end{description}
 
-We describe now a program designed to compute a value $J(u)$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in figure~\ref{fig:example_ns} with  $f=0$, and $J$ is the drag-force of the enclosed circle.
+We describe now a program designed to compute a value $J(u)$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in Figure~\ref{fig:example_ns} with  $f=0$, and $J$ is the drag-force of the enclosed circle.
 We assume given Dirichlet data for the velocity $v$ on the boundary $\Gamma_{D}=\Gamma_{in}\cup\Gamma_{wall}\cup\Gamma_{circ}$.
@@
 \end{align}
+\todo{Hier klarer machen warum wir Jacobi brauchen!}
 and the Jacobi-matrix of the weak form, i.e.
+\todo{Notation: $\phi_{h,j},\phi_{h}^{i}$}
 \begin{align}\label{eq:discrete_matrix}
@@
 \texttt{CellEquation}, \texttt{CellRightHandSide} and \texttt{CellMatrix}. 
-The implementation of the terms works by and large as in \deal{}.
-We show the basic structure of these methods using the example of \texttt{CellEquation}.
+\todo{Wirklich??? Mich erinnert die Art der Implementierung eher an Gascoigne!!!}
+The implementation of the terms works by and large as in \deal{}. We show the basic structure of these methods using the example of \texttt{CellEquation}.
 
@@
 \subsubsection{\texttt{localfunctional.h}}
-The implementation of the boundary-functional $J$ follows the outline of the computation of the cell-residual in \texttt{localpde.h}.
+The implementation of the boundary-functional $J$ follows the outline of the
+computation of the cell-residual in \texttt{localpde.h}. 
+\todo{Fuer mich hoert sich dieser Satz so an, als ob es IMMER two main
+  differences gibt. Aber das ist ja nur in diesem speziellen Beispiel der Fall}
 However, there are two main differences.
@@
 with some suitable Bochner space $X$.  As model problem,
-we consider the nonstationary Navier-Stokes equations
+we 
+\todo{extend the previous example and }consider the nonstationary Navier-Stokes equations
 in a domain $\Omega\in\mathbb{R}^d$. Find: 
@@
 As for stationary problems, the goal is to compute
-some target functional $J(\cdot)$ which is now possibly 
+some target functional \todo{$J(u)$} which is now possibly 
 time-dependent. To organize the discretization 
@@
 \end{itemize}
-With this we can now run a loop over all time-steps, where in each of the 
+\todo{Strange sentence} With this we can now run a loop over all time-steps, where in each of the 
 time-steps we can reuse the functionality to solve stationary PDEs 
@@
 $\theta = 0.5$ gives the Crank-Nicolson scheme,
+
+\todo{Remove the following}
 \begin{lstlisting}
@@
 
+\todo{It is well-known in fluid mechanics that}
 The pressure and incompressibility terms should be 
@@
 
+\todo{Aussage des Satzes ist mir nicht klar! Warum PDEs hier?}
 Due to the design of the classes it is easily possible to extend any given 
@@
 optimization algorithms to minimization problems governed by the PDE.
+
+\todo{Sollten wir einen Remark zu nonstationary Optimization machen?}
 
@@
 
+\todo{Christians Bilder}
 \begin{itemize}
@@
 \subsection{Nonstationary Fluid-Structure Interaction}
-In this example, a challenging benchmark FSI 2
+In this example, a challenging benchmark (FSI 2)
 proposed by Hron and Turek \cite{HrTu06b} is considered.
@@
 very easily chosen in the \texttt{main.cc} function by choosing an appropriate 
-time-stepping scheme, at present the following are available.
+time-stepping scheme
+\todo{. Remove the following
+, at present the following are available.}
 \begin{lstlisting}
@@
 
+\todo{\section{Outlook}}
+\todo{Ich finde nach wir vor, dass wir kurz skizzieren sollten, was wir als
+  naechstes vor haben}
+\todo{In this work, we introduced
+the software DOpElib to solve PDEs
+and Optimization problems. In the future,
+we plan to implement efficient routines 
+for nonstationary optimization and different 
+types of control and state constraints. 
+Moreover, we plan to link to other 
+software packages, specifically, in order 
+to use parallel computing routines.  
+}
+
 \section*{Thanks}

### 2012-11-21 Paper_May_2012/main_tw.tex
main TW: Weitere Anmerkungen hinzugefuegt

@@
 \end{lstlisting}
+\todo{Finde diesen Schlusssatz komisch}
 Finally, the program is started by
@@
 \subsection{Compliance minimization}
+\todo{Was ist MBB? Konnest Du vielleicht auch eine Referenz zur Konfiguration angeben?}
 In this example, we consider the compliance minimization of a standard MBB-Beam. 

### 2012-11-22 Paper_May_2012/main.tex
Finished my example

@@
 To be more precise, we are computing the benchmark 2D-1 from \cite{TuSchae96}.
-The computational domain is as depicted in figure~\ref{fig:example_ns}.
+The computational domain is as depicted in figure~\ref{fig:example_ns} and the equation is discretized with the Taylor-Hood  element.
 We want to compute the drag force on the enclosed cylinder
 \begin{equation}
-c_D = \frac 1 {20} \int_{\Gamma_{circ}} \nu\partial_nu _1 - pn,
+\mathcal {J}( u) = \frac 1 {20} \int_{\Gamma_{circ}} \nu\partial_nv _1 - pn,
 \end{equation}
-with the normal $n$ and $\nu = 0.001$. To achieve an efficient computation of the quantity $c_D$, we employ goal-oriented mesh refinement with the help of the DWR-method, see \cite{BeRa96}.
-
-\begin{itemize}
-\item Bild Freiheitsgrade vs Fehler, global und 
-\item Bild adaptive mesh, bild der Loesung (?)
-\item eventuell kurz auf duale gleichung eingehen.
-\end{itemize}
+with the solution $u = (v,p)$, the normal $n$ and the viscosity $\nu = 0.001$.
+\begin{figure}[hbt]
+\centering
+{\includegraphics[width=0.8\textwidth]{Pictures/local_grid_NS.png}}
+\caption{Caption.}
+\label{fig:NS_local_grid}
+\end{figure}
+To achieve an efficient computation of the quantity $\mathcal J$, we employ goal-oriented mesh refinement with the help of the DWR-method, see \cite{BeRa96}.
+The method delivers error indicators and by this way computational grids tailored to minimize the error in the previously given functional $\mathcal J$.
+The price we have to pay ist the additional computation of a (linear) dual problem, which is implemented as described in section~\ref{sec:opt}.
+In figure~\ref{fig:NS_local_grid} we see a resulting triangulation after eight refinement steps with roughly $320\,000$ degrees of freedom.
+
+In figure~\ref{fig:NS_comparison} we compare the errors in the functional obtained by using global and local mesh refinement.
+We see clearly that the local mesh refinement allows for substantial savings with respect to degrees of freedom.
+
+\ref{fig:NS_local_grid}
+\begin{figure}[hbt]
+\centering
+\resizebox{0.5\textwidth}{!}{\input{Pictures/ConvergenceNS}}
+\caption{Comparison of global and local mesh refinement with respect to the error in the drag-functional $\mathcal J$.}
+\label{fig:NS_comparison}
+\end{figure}
 

### 2012-11-23 Paper_May_2012/main_tw.tex
Meine Kommentare eingefuegt

@@
 \newcommand{\todo}[1]{\textbf{\textsc{\textcolor{blue}{TODO: #1}}}}
+\newcommand{\todocg}[1]{\textbf{\textsc{\textcolor{red}{TODO: #1\textasciitilde cg}}}}
 %\newcommand{\mymarginpar}[1]{\marginpar{\textcolor{red}{\scriptsize{#1}}}}
@@
 The article illustrates the design principles and various features
-of the \dope{} and provides some 
+of the \dope{} \todocg{Wir sollten uns einigen ob the DOpElib oder nur DOpElib} and provides some 
 numerical results as demonstration for the versatility of the software.
@@
 COMSOL, or simufact forming.
-Finally, often, a lot of work is (still) done in Matlab.
+Finally, often\todocg{wdh, s.u.: not seldom}, a lot of work is (still) done in Matlab.
 In fact, these libraries provide very useful tools 
@@
 
-However, often, the user is confronted with 
+However, often\todocg{wdh, siehe oben}, the user is confronted with 
 certain bottlenecks. On the one hand, 
@@
 such as linear and nonlinear solvers or time-stepping schemes must be implemented 
-by the users themselves. However, such routines are, at large,
+by the users themselves \todocg{ich wuerde diesen Nebensatz nach vorne verschieben:On the one hand, 
+in open-source projects for academic research, 
+standard high-level routines such as linear and nonlinear solvers or time-stepping schemes,
+which are necessary to solve even simplest PDEs,
+must be implemented by the users themselves.} . However, such routines are, at large,
 problem-independent and could be provided as standard tools. 
@@
 Both have been initiated as well in 
-the Numerical Analysis Group in Heidelberg.
+the Numerical Analysis Group in Heidelberg.\todocg{Wird hier klar, dass die Arbeiten an DOpE in Hd begonnen haben?
+Falls nein: Like DOpElib, both have been initiated in the Numerical Analysis Group in Heidelberg.}
 
@@
 The presented toolkit \dope{}: 
-\textit{Differential Equations and Optimization Environment} \cite{dope}
+\textit{Differential Equations and Optimization Environment} \cite{dope} \todocg{Url in reference in dopelib.net geaendert}
 fills this gap. We present a software
-to solve both PDE problems as well as optimization problems 
+to solve both \todocg{pure} PDE problems as well as optimization problems 
 constrained by a PDE. 
@@
 time-stepping methods, nonlinear solvers and optimization routines. 
-\dope{} is designed such that the user only needs to write those parts
+\dope{} is designed such\todocg{in such a way} that the user only needs to write those parts
 of the code that are problem dependent while all invariant 
@@
 \begin{itemize}
-\item Solution of nonlinear stationary and nonstationary PDEs in 1d, 2d, and 3d.
+\item Solution \todocg{Solving?} of nonlinear stationary and nonstationary PDEs in 1d, 2d, and 3d.
 \item Various, easily exchangeable, 
@@
 \item Use of all finite elements of from \deal{} including hp-support.
-\item Several examples showing the solution of several PDEs including
+\item Several\todocg{wdh, Some} examples showing the solution of several PDEs including
    Poisson, Navier-Stokes, Plasticity and fluid-structure interaction problems. 
 \item Line search and trust region newton algorithms for the 
-   solution of optimization problems with PDEs by elimination of 
+   solution of optimization problems with PDEs \todocg{PDE constraints} by elimination of 
    the state variable.
@@
 \label{detailed_description}
-This library is designed to allow easy implementation and numerical solutions 
+This library is designed to allow easy implementation and numerical solutions \todocg{numerical solutions? strange} 
 of problems involving partial differential equations (PDEs). 
@@
 problem independent implementation while leaving only the problem specific 
-part to the user. To show how this is designed we will, in the sequel, 
+part \todocg{parts} to the user. To show how this is designed we will, in the sequel, 
 describe the reasoning first for a simple stationary PDE in 
@@
 
-In order to solve this problem, i.e., calculate $J(u)$, we need to discretize 
+In order to solve this problem, i.e., calculate $J(u)$,\todocg{i.e. calculate a series of values $J(u_h)$ approximating $J(u)$,} we need to discretize 
 the equation. In \dope, we are using finite elements to do this. This means we need to pick a 
@@
 form of the constructor should have the following interface:
+\todocg{(Triangulation\& triag, FE\& fe)}
+\todocg{TRIANGULATION. all other templates are in uppercase.}
 \begin{lstlisting}
@@
 
-We describe now a program designed to compute a value $J(u)$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in Figure~\ref{fig:example_ns} with  $f=0$, and $J$ is the drag-force of the enclosed circle.
+We describe now a program designed to compute a value $J(u)$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in Figure~\ref{fig:example_ns} with  $f=0$, and $J$ is the drag-force of the enclosed circle\todocg{Eigentlich berechnen wir nicht $J(u)$ sondern approximationen davon, siehe auch weiter oben.}.
 We assume given Dirichlet data for the velocity $v$ on the boundary $\Gamma_{D}=\Gamma_{in}\cup\Gamma_{wall}\cup\Gamma_{circ}$.
@@
 \end{figure}
-
 We split the program into the parts \texttt{main.cc}, \texttt{functionals.h} and \texttt{localpde.h}.
-The \texttt{main.cc}-file is responsible for the declaration of various objects which have mostly to do with the \textit{discretization}, the \textit{solvers} and the general \textit{program sequence}, whereas the other two contain most of the problem-data -- especially the cell-wise contributions of the weak form, the right hand side and the functional.
+The \texttt{main.cc}-file is responsible for the declaration of various objects which have mostly to do with the \todocg{textbf finde ich besser als textit, haengt aber vom stylefile ab.}\textit{discretization}, the \textit{solvers} and the general \textit{program sequence}, whereas the other two contain most of the problem-data -- especially the cell-wise contributions of the weak form, the right hand side and the functional.
 We give a brief explanation on every file in the following.
@@
 Then we have to tell the \texttt{prob\_container} for which components and boundaries we prescribe the Dirichlet conditions.
-This is done via a boundary-color (0 stands for $\Gamma_D$) and a \texttt{std::vector<bool> comp\_mask = (true, true, false)}.
+This is done via a boundary-color (0 stands for $\Gamma_D$) and a \todocg{comp mask representing the velocity components.} \texttt{std::vector<bool> comp\_mask = (true, true, false)}.
 \begin{lstlisting}
@@
 The class depends actually on some template-parameters, so for the sake of abbreviation, we defined a typedef at the beginning of the \texttt{main.cc}-file.
+\todocg{NEWTON statt NEWTONSOLVER}
 \begin{lstlisting}
@@
 \end{lstlisting}
-Here, \texttt{NEWTONSOLVER} itself is an abbreviation of
+Here, \texttt{NEWTONSOLVER}\todocg{NEWTON statt NEWTONSOLVER} itself is an abbreviation of
+\todocg{LINSOLVE statt LINEARSOLVER}
 \begin{lstlisting}
@@
 \end{lstlisting}
-The  template parameter \texttt{LINEARSOLVER} describes which linear solver we use in this example. We opted for a direct solver, so: 
+The  template parameter \texttt{LINEARSOLVER}\todocg{LINSOLVE statt LINEARSOLVER} describes which linear solver we use in this example. We opted for a direct solver, so: 
 \begin{lstlisting}
@@
 The implementation of the terms works by and large as in \deal{}. We show the basic structure of these methods using the example of \texttt{CellEquation}.
-
+\todocg{Remove template <2>}
 \begin{lstlisting}
@@
 we loop over the quadrature points and the number of DoFs of the cell.
-In the innermost loop, we evaluate the weak form tested with $\phi_{i,h}$ (represented by \texttt{phi\_i\_grads\_v}) on the cell with the help of the previously selected quadrature formula.
+In the innermost loop, we evaluate the weak form tested with $\phi_{i,h}$ \todocg{$\phi_h^i$}(represented by \texttt{phi\_i\_grads\_v}) on the cell with the help of the previously selected quadrature formula.
  \begin{lstlisting}
@@
   differences gibt. Aber das ist ja nur in diesem speziellen Beispiel der Fall}
+  \todocg{The main difference is that as we only want to evaluate a functional and not a residual-vector, we just loop over the quadrature points and not over the degrees of freedom on the face.
+Another deviation, special to the case at hand, is that we do not have to specify the method \texttt{CellEquation}, but \texttt{BoundaryValue} instead.
+This is due to the fact that we consider a boundary-functional in this example, and so we have to specify the value of the functional not cell-wise but face wise (for the faces that lie on $\Gamma_{circ}$).
+}
 However, there are two main differences.
@@
 The second difference is that as we only want to evaluate a functional and not a residual-vector, we just loop over the quadrature points and not over the degrees of freedom on the face.
+\todocg{Remove template <2>}
 \begin{lstlisting}
@@
 at all times, the constructor could look as follows:
+\todocg{typename TIMES (...) Triangulation\& triag, FE\& fe, TIMES\& times}
 \begin{lstlisting}
@@
 Thus a generic problem may look as follows:
+\todocg{arguments of the constructor?}
 \begin{lstlisting}
@@
 \begin{itemize}
-\item Forward-(Explicit)-Euler
+\item Forward-(Explicit)-Euler \todocg{Ist das nicht doppelt gemoppelt?}
 \item Backward-(Implicit)-Euler
@@
 in the method \texttt{CellTimeEquation} in a straight-forward way.
-After the initialization of $v$ and the shape function $\phi_i$ and the usual loops over the quadrature points and  degrees of freedom of the cell we fill the local cell vector by
+After the initialization of $v$ and the shape function $\phi_i$ \todocg{ich wuerde fuer v und phi hier die texttt function nutzen.} and the usual loops over the quadrature points and  degrees of freedom of the cell we fill the local cell vector by
 \begin{lstlisting}
@@
 respective argument. To calculate action of the adjoint operator $S^*$ 
-we note that $z =  S^* J_u'(q,S(q))$
+we note that $z =  S^* J_u'(q,S(q))$ \todocg{action of the ....?}
 can be computed via the adjoint PDE
@@
 
-To do so, we define an object 
+To do so, we define an object \todocg{Argumente?}
 \begin{lstlisting} 
@@
 behind the rigid cylinder. 
+\todocg{Die Masse, die aus dem Bild klarwerden, wuerde ich hier weglassen.}
 The computational domain has length $L=2.5m$ and height $H=0.41m$. The circle center
@@
 \subsection{Compliance minimization}
-\todo{Was ist MBB? Konnest Du vielleicht auch eine Referenz zur Konfiguration angeben?}
+\todo{Was ist MBB? Konnest Du vielleicht auch eine Referenz zur Konfiguration angeben?}\todocg{Wir sollten uns auf eine gemeinsame Darstellung der Bilder einigen (z.B. Achsenbeschriftung ja/nein).}
 In this example, we consider the compliance minimization of a standard MBB-Beam. 

### 2012-11-23 Paper_May_2012/main.tex
Update der DOpEwebadresse, minor typos.

@@
 After the usual preamble, in which we include the necessary header-files from \deal{}, \dope{} and our own problem (i.e.
-\texttt{localfunctional.h} and \texttt{localpde.h}), we start by defining an instance of a \texttt{ParameterReader}-object, which handles run-time parameter.
+\texttt{localfunctional.h} and \texttt{localpde.h}), we start by defining an instance of a \texttt{ParameterReader}-object, which handles run-time parameters.
 \begin{lstlisting}
@@
 \begin{lstlisting} 
-  virtual void BoundaryValue_U(...)
+  void BoundaryValue_U(...)
    {

### 2012-11-23 Paper_May_2012/main.tex
Todos eingefuegt (aber noch nicht style file). Wollte einmal speichern bevor ich alles im neuen style file umkopiere

@@
 
-\newcommand{\todo}[1]{\textbf{\textsc{\textcolor{black}{TODO: #1}}}}
+%\newcommand{\todo}[1]{\textbf{\textsc{\textcolor{black}{TODO: #1}}}}
+\newcommand{\todo}[1]{\textbf{\textsc{\textcolor{blue}{TODO: #1}}}}
+\newcommand{\todocg}[1]{\textbf{\textsc{\textcolor{red}{TODO: #1\textasciitilde cg}}}}
 %\newcommand{\mymarginpar}[1]{\marginpar{\textcolor{red}{\scriptsize{#1}}}}
@@
 \begin{abstract}
+\todo{Computing}
 In this article, we describe the software library 
@@
 The article illustrates the design principles and various features
-of the \dope{} and provides some 
+of \dope{} and provides some 
 numerical results as demonstration for the versatility of the software.
@@
 Universit\"at Heidelberg;
-T. Wick, The Institute for Computational Engineering and Sciences;
+T. Wick, The Institute for Computational Engineering and Sciences, The
+University of Texas at Austin;
 W. Wollner, Fachbereich Mathematik, Universit\"at Hamburg.
@@
 COMSOL, or simufact forming.
-Finally, often, a lot of work is (still) done in Matlab.
+Finally, a lot of work is (still) done in Matlab.
 In fact, these libraries provide very useful tools 
@@
 
-However, often, the user is confronted with 
+In many cases, the user is confronted with 
 certain bottlenecks. On the one hand, 
@@
 and RoDoBo \cite{rodobo}. 
-Both have been initiated as well in 
+Like DOpElib, both have been initiated as well in 
 the Numerical Analysis Group in Heidelberg.
@@
 fills this gap. We present a software
-to solve both PDE problems as well as optimization problems 
+to solve both pure PDE problems as well as optimization problems 
 constrained by a PDE. 
@@
 time-stepping methods, nonlinear solvers and optimization routines. 
-\dope{} is designed such that the user only needs to write those parts
+\dope{} is designed such a way that the user only needs to write those parts
 of the code that are problem dependent while all invariant 
@@
 \begin{itemize}
-\item Solution of nonlinear stationary and nonstationary PDEs in 1d, 2d, and 3d.
+\item Solving of nonlinear stationary and nonstationary PDEs in 1d, 2d, and 3d.
 \item Various, easily exchangeable, 
@@
 \item Use of all finite elements of from \deal{} including hp-support.
-\item Several examples showing the solution of several PDEs including
+\item Some examples showing the solution of several PDEs including
    Poisson, Navier-Stokes, Plasticity and fluid-structure interaction problems. 
 \item Line search and trust region newton algorithms for the 
-   solution of optimization problems with PDEs by elimination of 
+   solution of optimization problems with PDE constraints by elimination of 
    the state variable.
@@
 \label{detailed_description}
-This library is designed to allow easy implementation and numerical solutions 
+This library is designed to allow easy implementation for the numerical solution 
 of problems involving partial differential equations (PDEs). 
@@
 problem independent implementation while leaving only the problem specific 
-part to the user. To show how this is designed we will, in the sequel, 
+parts to the user. To show how this is designed we will, in the sequel, 
 describe the reasoning first for a simple stationary PDE in 
@@
 Then typically we do not only want to calculate $u$ but more likely we are 
-interested in one (or several) functionals depending on the solution, i.e., some numbers $J(u)$.
-
-In order to solve this problem, i.e., calculate $J(u)$, we need to discretize 
+interested in one (or several) functionals depending on the solution, i.e., some numbers $J$.
+
+In order to solve this problem, i.e., calculate a sequence of values $J(u_h)$ approximating $J(u)$,
+we need to discretize 
 the equation. In \dope, we are using finite elements to do this. This means we need to pick a 
@@
 \begin{lstlisting}
-  template<typename Triangulation, typename FE>
+  template<typename TRIANGULATION & triag, typename FE & fe>
       void SpaceTimeHandler(Triangulation &, FE);
@@
 there are several different sparsity patterns available in 
-\deal{} which the user should be able chose. Thus it needs to be 
+\deal{} which the user should be able to choose. Thus it needs to be 
 given to the \texttt{SpaceTimeHandler}. However, to keep the 
@@
 \begin{description}
-\item[Discretization] Specify what kind of triangulation,
-finite element, quadrature rules, etc. we want to use.
-\item[Problem-data] Define the weak form, the corresponding matrices, the functionals as well as the boundary conditions.
-With exception of the boundary conditions, we have to specify all this only cell-wise, \dope{} handles the rest for us.
-\item[Solvers] Which (non-)linear solver should we use? 
+\item[Discretization] Specify what kind of triangulation:
+in most cases this is the
+finite element and quadrature rules. Furthermore, other
+things might be adjusted as well.
+\item[Problem-data] Define the weak form, the corresponding matrices, the
+  functionals $J$ as well as the boundary conditions. With exception of the boundary conditions, we have to specify all this only cell-wise, \dope{} handles the rest for us.
+\item[Solvers] Choice of a (non-)linear solver. 
 \item[Additional features] ParameterReader to read in the run-time parameters
   such as material coefficients or adjustments for the solvers like tolerance
-  for Newton's method. Second, definition of Outputhandler etc.
+  for Newton's method. 
 \end{description}
 
-We describe now a program designed to compute a value $J(u)$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in figure~\ref{fig:example_ns} with  $f=0$, and $J$ is the drag-force of the enclosed circle.
+We describe now a program designed to compute a value $J$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in figure~\ref{fig:example_ns} with  $f=0$, and $J$ is the drag-force of the enclosed circle.
 We assume given Dirichlet data for the velocity $v$ on the boundary $\Gamma_{D}=\Gamma_{in}\cup\Gamma_{wall}\cup\Gamma_{circ}$.
@@
 
+\todocg{Haengt
+  von style-file ab: textbf finde ich besser als textit, haengt aber vom stylefile ab.}
 We split the program into the parts \texttt{main.cc}, \texttt{functionals.h} and \texttt{localpde.h}.
@@
 \end{lstlisting}
-Next, we incorporate the Dirichlet data, which is given by the class \texttt{DirichleData}.
-Then we have to tell the \texttt{prob\_container} for which components and boundaries we prescribe the Dirichlet conditions.
-This is done via a boundary-color (0 stands for $\Gamma_D$) and a \texttt{std::vector<bool> comp\_mask = (true, true, false)}.
+Next, we incorporate the Dirichlet data, which is given by the class
+\texttt{DirichleData}. Then we have to tell the \texttt{prob\_container} for
+which components and boundaries we prescribe the Dirichlet conditions. This is
+done via a boundary-color in a so-called component mask 
+(here, $0$ denotes $\Gamma_D$).
 \begin{lstlisting}
@@
 
-  typedef StatPDEProblem<NEWTONSOLVER> SSOLVER;
-
-\end{lstlisting}
-Here, \texttt{NEWTONSOLVER} itself is an abbreviation of
-\begin{lstlisting}
-
-  typedef NewtonSolver<LINEARSOLVER> NEWTONSOLVER;
-
-\end{lstlisting}
-The  template parameter \texttt{LINEARSOLVER} describes which linear solver we use in this example. We opted for a direct solver, so: 
-\begin{lstlisting}
-
-  typedef DirectLinearSolver<MATRIX, VECTOR> LINEARSOLVER;
+  typedef StatPDEProblem<NEWTON> PDE;
+
+\end{lstlisting}
+Here, \texttt{NEWTON} itself is an abbreviation of
+\begin{lstlisting}
+
+  typedef NewtonSolver<LINSOLVE> NEWTON;
+
+\end{lstlisting}
+The  template parameter \texttt{LINSOLVE} describes which linear solver we use in this example. We opted for a direct solver, so: 
+\begin{lstlisting}
+
+  typedef DirectLinearSolver<MATRIX, VECTOR> LINSOLVE;
 
@@
 
-  SSOLVER solver(prob_container, pr,
+  PDE pde(prob_container, pr,
                     quad_formula, face_quad_formula);
@@
 \end{lstlisting}
-Finally, the program is started by
-\begin{lstlisting}
-
-  solver.ReInit();
-  solver.ComputeReducedFunctionals();
+Finally, we compute the functional $J$ by calling:
+\begin{lstlisting}
+
+  pde.ComputeReducedFunctionals();
 
@@
 \begin{align}\label{eq:discrete_residual}
-\left(a(u_h)(\phi_{h}^{i})-(f,\phi_{h}^{i})\right)_{i=1}^N,
+\left(a(u_h)(\phi_{h}^{i})-(f,\phi_{h}^{i})\right)_{i=1}^N.
 \end{align}
-and the Jacobi-matrix of the weak form, i.e.
-\begin{align}\label{eq:discrete_matrix}
-\left(\nabla a(u_h)(\phi_{h,j},\phi_{h}^{i})\right)_{i,j=1}^N.
-\end{align}
+The corresponding Jacobian can be defined analogously.
 Now all we have to supply here are the cell-wise contributions to these terms, the \texttt{Integrator} takes care of assembling \eqref{eq:discrete_residual} and \eqref{eq:discrete_matrix}.
@@
 \texttt{CellEquation}, \texttt{CellRightHandSide} and \texttt{CellMatrix}. 
-The implementation of the terms works by and large as in \deal{}.
+The splitting of these terms is similar to Gascoigne
+\cite{gascoigne} whereas the content (i.e., the 
+implementation itself) of these
+terms works by and large as in \deal{}.
+
 We show the basic structure of these methods using the example of \texttt{CellEquation}.
@@
 \begin{lstlisting}
-  void CellEquation(const CellDataContainer<2>& cdc,
+  void CellEquation(const CellDataContainer &cdc,
                     VECTOR &local_cell_vector)
@@
 we loop over the quadrature points and the number of DoFs of the cell.
-In the innermost loop, we evaluate the weak form tested with $\phi_{i,h}$ (represented by \texttt{phi\_i\_grads\_v}) on the cell with the help of the previously selected quadrature formula.
+In the innermost loop, we evaluate the weak form tested with $\phi_{h}^i$ (represented by \texttt{phi\_i\_grads\_v}) on the cell with the help of the previously selected quadrature formula.
  \begin{lstlisting}
@@
 \subsubsection{\texttt{localfunctional.h}}
-The implementation of the boundary-functional $J$ follows the outline of the computation of the cell-residual in \texttt{localpde.h}.
-However, there are two main differences.
+The implementation of the boundary-functional $J$ follows the outline of the
+computation of the cell-residual in \texttt{localpde.h}. 
+\todo{CG and WW: Pruefen, ob der folgende Absatz nun gut ist:}
+The main difference is that as we only want to evaluate a functional and not a
+residual-vector, we just loop over the quadrature points and not over the
+degrees of freedom on the face. Another deviation, special to the case at
+hand, is that we do not have to specify the method \texttt{CellEquation}, but
+\texttt{BoundaryValue} instead because $J$ is just number to be computed.
+This is due to the fact that we consider a boundary-functional in this example, and so we have to specify the value of the functional not cell-wise but face wise (for the faces that lie on $\Gamma_{circ}$).
+
+%However, there are two main differences.
 First, we do not have to specify the method \texttt{CellEquation}, but \texttt{BoundaryValue} instead.
@@
  double
- BoundaryValue(const FaceDataContainer<2>& fdc)
+ BoundaryValue(const FaceDataContainer& fdc)
   {
@@
 with some suitable Bochner space $X$.  As model problem,
-we consider the nonstationary Navier-Stokes equations
+we extend the previous example and consider the nonstationary Navier-Stokes equations
 in a domain $\Omega\in\mathbb{R}^d$. Find: 
@@
 As for stationary problems, the goal is to compute
-some target functional $J(\cdot)$ which is now possibly 
+some target functional $J$ which is now possibly 
 time-dependent. To organize the discretization 
@@
 \begin{lstlisting}
-  template<typename Triangulation, typename FE, typename times>
-      void SpaceTimeHandler(Triangulation &, FE, times&);
+  template<typename TIMES & times, 
+           typename TRIANGULATION & triag, 
+           typename FE & fe>
+      void SpaceTimeHandler(Times &, Triangulation &, FE &);
 \end{lstlisting}
@@
   template<typename PROBLEM>
-      class TimeStepProblem;
+      class TimeStepProblem(Problem & P);
 \end{lstlisting}
@@
 \begin{itemize}
-\item Forward-(Explicit)-Euler
-\item Backward-(Implicit)-Euler
+\item Forward-(or Explicit)-Euler
+\item Backward-(or Implicit)-Euler
 \item Crank-Nicolson scheme
@@
 \end{itemize}
-With this we can now run a loop over all time-steps, where in each of the 
-time-steps we can reuse the functionality to solve stationary PDEs 
-applied to the time-step problem.
-
+\todo{CG and WW: bitte pruefen}
+After choosing the time-stepping scheme, the SpaceTimeHandler  
+loops over all time-steps, where in each of the 
+time-steps we can reuse the functionality to solve stationary PDEs. 
 Now it is apparent, why we introduced the \texttt{StatPDEProblem}, because
-we now define the \texttt{InstatPDEProblem} to contain the necessary 
-loop over all time points. Next, we will describe the methods the user needs to 
+we define the \texttt{InstatPDEProblem} to contain the necessary 
+loop over all time points. Consequently, the implementation 
+of nonstationary problems is quite similar to stationary problems,
+which was one major goal we claimed at the beginning.
+Next, we will describe the methods (i.e., changes) the user needs to 
 provide in addition to the stationary case.
@@
 specific time-stepping scheme. For instance,
-$\theta = 0.5$ gives the Crank-Nicolson scheme,
-\begin{lstlisting}
-   CrankNicolsonProblem<...>
-\end{lstlisting}
-
-The pressure and incompressibility terms should be 
+$\theta = 0.5$ gives the Crank-Nicolson scheme.
+
+It is well-known in fluid mechanics that the pressure and incompressibility terms should be 
 treated fully implicitly when time-discretizing 
@@
 in the method \texttt{CellTimeEquation} in a straight-forward way.
-After the initialization of $v$ and the shape function $\phi_i$ and the usual loops over the quadrature points and  degrees of freedom of the cell we fill the local cell vector by
+After the initialization of $\texttt{v}$ and the shape function $\texttt{phi\_i}$ and the usual loops over the quadrature points and  degrees of freedom of the cell we fill the local cell vector by
 \begin{lstlisting}
@@
 \subsection{Solving of optimization problems}\label{sec:opt}
+\todo{For WW: etwas klarer machen, dass wir instationare Optimierung koennen}
 Upgrading the framework for solving optimization problems
@@
 where the index $q$ or $u$ denote directional derivatives for the 
-respective argument. To calculate action of the adjoint operator $S^*$ 
+respective argument. To calculate the action of the adjoint operator $S^*$ 
 we note that $z =  S^* J_u'(q,S(q))$
@@
 
-Due to the design of the classes it is easily possible to extend any given 
+Summarizing, we learned from the previous sections 
+that 
+due to the design of the classes it is easily possible to extend any given 
 and running code for a PDE both to the nonstationary counterpart, if the PDE is 
@@
 with the solution $u = (v,p)$, the normal $n$ and the viscosity $\nu = 0.001$.
+\todo{For CG: Bildunterschrift und die $2$ am Ende und 
+das cal J}
 \begin{figure}[hbt]
@@
 behind the rigid cylinder. 
-The computational domain has length $L=2.5m$ and height $H=0.41m$. The circle center
-is positioned at $C=(0.2m,0.2m)$ with radius $r=0.05m$. The elastic beam has length
+The elastic beam has length
 $l=0.35m$ and height $h=0.02m$. The right lower end is positioned at 
@@
 very easily chosen in the \texttt{main.cc} function by choosing an appropriate 
-time-stepping scheme, at present the following are available.
-\begin{lstlisting}
-ForwardEulerProblem
-BackwardEulerProblem
-CrankNicolsonProblem
-ShiftedCrankNicolsonProblem
-FractionalStepThetaProblem
-\end{lstlisting}
+time-stepping scheme as previously explained.
 
@@
 \subsection{Compliance minimization}
+\todo{For Winni: Achsenbeschriftung, MBB ausschreiben, 
+Konfiguration citieren}
 In this example, we consider the compliance minimization of a standard MBB-Beam. 
@@
 
+\section{Conclusions}
+\todo{Frage an CG und Winni: Guter Abschluss?}
+In this article, we described the architecture 
+of the novel software library DOpElib for solving 
+PDEs and optimization problems. The key idea behind 
+the algorithms is the computation of some target 
+functional. This is achieved by unified interface 
+to state-of the art time-stepping methods, nonlinear solvers and optimization
+routines. The 
+capabilites of DOpElib are demonstrated
+wit the help of three examples.
+
 \section*{Thanks}

### 2012-11-23 Paper_May_2012/main_tw.tex
main tw: zwei Anmerkungen noch hinzugefuegt
Todos eingefuegt (aber noch nicht style file). Wollte einmal speichern bevor ich alles im neuen style file umkopiere

@@
 \begin{abstract}
+\todo{Computing. Anschrift tw}
+\todo{Welches Journal: Computing oder ATS?}
 In this article, we describe the software library 
@@
 The article illustrates the design principles and various features
-of the \dope{} \todocg{Wir sollten uns einigen ob the DOpElib oder nur DOpElib} and provides some 
+of the \dope{} \todocg{DOpElib} and provides some 
 numerical results as demonstration for the versatility of the software.
@@
 COMSOL, or simufact forming.
-Finally, often\todocg{wdh, s.u.: not seldom}, a lot of work is (still) done in Matlab.
+\todo{Finally}, a lot of work is (still) done in Matlab.
 In fact, these libraries provide very useful tools 
@@
 
-However, often\todocg{wdh, siehe oben}, the user is confronted with 
+\todo{In many cases}, the user is confronted with 
 certain bottlenecks. On the one hand, 
@@
 and RoDoBo \cite{rodobo}. 
+\todo{Alles ok auf Seite 2}
 Both have been initiated as well in 
@@
 \label{detailed_description}
-This library is designed to allow easy implementation and numerical solutions \todocg{numerical solutions? strange} 
+This library is designed to allow easy implementation \todo{for the} numerical solutions \todocg{numerical solutions? strange} 
 of problems involving partial differential equations (PDEs). 
@@
 problem independent implementation while leaving only the problem specific 
-part \todocg{parts} to the user. To show how this is designed we will, in the sequel, 
+part \todocg{Ja. parts} to the user. To show how this is designed we will, in the sequel, 
 describe the reasoning first for a simple stationary PDE in 
@@
 
-In order to solve this problem, i.e., calculate $J(u)$,\todocg{i.e. calculate a series of values $J(u_h)$ approximating $J(u)$,} we need to discretize 
+In order to solve this problem, i.e., calculate $J(u)$,\todocg{i.e. calculate a sequence of values $J(u_h)$ approximating $J(u)$,} we need to discretize 
 the equation. In \dope, we are using finite elements to do this. This means we need to pick a 
@@
 form of the constructor should have the following interface:
-\todocg{(Triangulation\& triag, FE\& fe)}
-\todocg{TRIANGULATION. all other templates are in uppercase.}
+\todocg{CG: (Triangulation\& triag, FE\& fe)}
+\todocg{CG: TRIANGULATION. all other templates are in uppercase.}
 \begin{lstlisting}
@@
 finite element, quadrature rules, 
-\todo{etc. strange to me} etc. we want to use.
+\todo{etc. strange to me: in most cases this is the 
+... and other things maybe adjusted as well} etc. we want to use.
 \item[Problem-data] 
@@
 the corresponding matrices, 
-the functionals \todo{$J(u)$} as well as the boundary conditions. With exception of the boundary conditions, we have to specify all this only cell-wise, \dope{} handles the rest for us.
-\item[Solvers] \todo{Choice of a (non-)linear solver.}
+the functionals \todo{Ok: $J$} as well as the boundary conditions. With exception of the boundary conditions, we have to specify all this only cell-wise, \dope{} handles the rest for us.
+\item[Solvers] \todo{Ok: Choice of a (non-)linear solver.}
 \item[Additional features] ParameterReader to read in the run-time parameters
@@
   for Newton's method. 
-\todo{Sentence strang to me} Second, definition of Outputhandler etc.
+\todo{Satz weg: Sentence strang to me} Second, definition of Outputhandler etc.
 \end{description}
 
-We describe now a program designed to compute a value $J(u)$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in Figure~\ref{fig:example_ns} with  $f=0$, and $J$ is the drag-force of the enclosed circle\todocg{Eigentlich berechnen wir nicht $J(u)$ sondern approximationen davon, siehe auch weiter oben.}.
+We describe now a program designed to compute a value $J(u)$, where $u=(v,p)$
+is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain
+$\Omega$ as seen in Figure~\ref{fig:example_ns} with  $f=0$, and $J$ is the
+drag-force of the enclosed circle\todocg{OK: Ueberall $J$ anstatt $J(u)$ Eigentlich berechnen wir nicht $J(u_h)$ sondern approximationen davon, siehe auch weiter oben.}.
 We assume given Dirichlet data for the velocity $v$ on the boundary $\Gamma_{D}=\Gamma_{in}\cup\Gamma_{wall}\cup\Gamma_{circ}$.
@@
 \end{figure}
-We split the program into the parts \texttt{main.cc}, \texttt{functionals.h} and \texttt{localpde.h}.
-The \texttt{main.cc}-file is responsible for the declaration of various objects which have mostly to do with the \todocg{textbf finde ich besser als textit, haengt aber vom stylefile ab.}\textit{discretization}, the \textit{solvers} and the general \textit{program sequence}, whereas the other two contain most of the problem-data -- especially the cell-wise contributions of the weak form, the right hand side and the functional.
+We split the program into the parts \texttt{main.cc}, \texttt{functionals.h}
+and \texttt{localpde.h}. The \texttt{main.cc}-file is responsible for the
+declaration of various objects which have mostly to do with the \todocg{Haengt
+  von style-file ab: textbf finde ich besser als textit, haengt aber vom stylefile ab.}\textit{discretization}, the \textit{solvers} and the general \textit{program sequence}, whereas the other two contain most of the problem-data -- especially the cell-wise contributions of the weak form, the right hand side and the functional.
 We give a brief explanation on every file in the following.
@@
 \end{lstlisting}
-Next, we incorporate the Dirichlet data, which is given by the class \texttt{DirichleData}.
-Then we have to tell the \texttt{prob\_container} for which components and boundaries we prescribe the Dirichlet conditions.
-This is done via a boundary-color (0 stands for $\Gamma_D$) and a \todocg{comp mask representing the velocity components.} \texttt{std::vector<bool> comp\_mask = (true, true, false)}.
+Next, we incorporate the Dirichlet data, which is given by the class
+\texttt{DirichleData}. Then we have to tell the \texttt{prob\_container} for
+which components and boundaries we prescribe the Dirichlet conditions. This is
+done via a boundary-color (0 stands for $\Gamma_D$) and a 
+\todocg{Aendern in Satz: component mask representing the velocity components.} \texttt{std::vector<bool> comp\_mask = (true, true, false)}.
 \begin{lstlisting}
@@
 The class depends actually on some template-parameters, so for the sake of abbreviation, we defined a typedef at the beginning of the \texttt{main.cc}-file.
-\todocg{NEWTON statt NEWTONSOLVER}
+\todocg{Alle Anerdungen accept auf Seite 7: NEWTON statt NEWTONSOLVER}
 \begin{lstlisting}
@@
 \end{lstlisting}
-With the type settled, we create the object \texttt{solver}, which needs the discretization (represented here by the \texttt{prob\_container}), the parameter reader as well as the quadrature rules.
+With the type settled, we create the object \texttt{solver}, which needs the
+discretization (represented here by the \texttt{prob\_container}), the
+parameter reader as well as the quadrature rules.
+\todo{SSolver zu PDE}
 \begin{lstlisting}
@@
 \end{lstlisting}
-\todo{Finde diesen Schlusssatz komisch}
+\todo{Finde diesen Schlusssatz komisch: Finally we compute 
+the functional $J$ und Reinit weg. We can calculate. solver zu pde}
 Finally, the program is started by
@@
 \end{align}
-\todo{Hier klarer machen warum wir Jacobi brauchen!}
+\todo{Jacobi explicit weg und Satz weg: Hier klarer machen warum wir Jacobi brauchen!}
 and the Jacobi-matrix of the weak form, i.e.
@@
 \texttt{CellEquation}, \texttt{CellRightHandSide} and \texttt{CellMatrix}. 
-\todo{Wirklich??? Mich erinnert die Art der Implementierung eher an Gascoigne!!!}
+\todo{Wirklich??? Mich erinnert die Art der Implementierung eher an
+  Gascoigne!!! Splitting is similar to Gascoigne. Todos von CG ok auf dieser Seite}
 The implementation of the terms works by and large as in \deal{}. We show the basic structure of these methods using the example of \texttt{CellEquation}.
@@
   differences gibt. Aber das ist ja nur in diesem speziellen Beispiel der Fall}
-  \todocg{The main difference is that as we only want to evaluate a functional and not a residual-vector, we just loop over the quadrature points and not over the degrees of freedom on the face.
+  \todocg{Satz von CG ist fast OK.
+Equation gegen Value tauschen.
+Todo belassen: The main difference is that as we only want to evaluate a functional and not a residual-vector, we just loop over the quadrature points and not over the degrees of freedom on the face.
 Another deviation, special to the case at hand, is that we do not have to specify the method \texttt{CellEquation}, but \texttt{BoundaryValue} instead.
@@
 The second difference is that as we only want to evaluate a functional and not a residual-vector, we just loop over the quadrature points and not over the degrees of freedom on the face.
-\todocg{Remove template <2>}
+\todocg{OK: Remove template <2>}
 \begin{lstlisting}
@@
 As for stationary problems, the goal is to compute
-some target functional \todo{$J(u)$} which is now possibly 
+some target functional \todo{alle Ssahcen auf Seite 10 uebernehmen $J$} which is now possibly 
 time-dependent. To organize the discretization 
@@
 \begin{itemize}
-\item Forward-(Explicit)-Euler \todocg{Ist das nicht doppelt gemoppelt?}
+\item Forward-(Explicit)-Euler \todocg{Doppelt belassen in klammern explicit Ist das nicht doppelt gemoppelt?}
 \item Backward-(Implicit)-Euler
@@
 \end{itemize}
-\todo{Strange sentence} With this we can now run a loop over all time-steps, where in each of the 
+\todo{Strange sentence: Alles etwas ausfuhrlicher} With this we can now run a loop over all time-steps, where in each of the 
 time-steps we can reuse the functionality to solve stationary PDEs 
@@
 
-\todo{Remove the following}
+\todo{Ok: Remove the following}
 \begin{lstlisting}
@@
 
-\todo{It is well-known in fluid mechanics that}
+\todo{Ok: It is well-known in fluid mechanics that}
 The pressure and incompressibility terms should be 
@@
 of the time-derivative, i.e. the term $ (v^{n+1},\phi)$, must be considered which is done
-in the method \texttt{CellTimeEquation} in a straight-forward way.
-After the initialization of $v$ and the shape function $\phi_i$ \todocg{ich wuerde fuer v und phi hier die texttt function nutzen.} and the usual loops over the quadrature points and  degrees of freedom of the cell we fill the local cell vector by
+in the method \texttt{CellTimeEquation} in a straight-forward way. After the
+initialization of $v$ and the shape function $\phi_i$ \todocg{ok: in ttt
+  aendern: ich wuerde fuer v und phi hier die texttt function nutzen.} and the usual loops over the quadrature points and  degrees of freedom of the cell we fill the local cell vector by
 \begin{lstlisting}
@@
 respective argument. To calculate action of the adjoint operator $S^*$ 
-we note that $z =  S^* J_u'(q,S(q))$ \todocg{action of the ....?}
+we note that $z =  S^* J_u'(q,S(q))$ \todocg{andern: the action of the ....?}
 can be computed via the adjoint PDE
@@
 
-To do so, we define an object \todocg{Argumente?}
+To do so, we define an object \todocg{weg: Argumente?}
 \begin{lstlisting} 
@@
 
-\todo{Aussage des Satzes ist mir nicht klar! Warum PDEs hier?}
+\todo{WW: Remark ob zeitabh oder nicht. 1-2 saetze. 
+check internet webseite}
+
+\todo{Summarzing, we have learned: Aussage des Satzes ist mir nicht klar! Warum PDEs hier?}
 Due to the design of the classes it is easily possible to extend any given 
@@
 \todo{Christians Bilder}
+\todo{Bildunterschrift CG und noch etwas anderes und die 2}
 \begin{itemize}
@@
 behind the rigid cylinder. 
-\todocg{Die Masse, die aus dem Bild klarwerden, wuerde ich hier weglassen.}
+\todocg{Die Masse und langen und Einheiten, die aus dem Bild klarwerden, wuerde ich hier weglassen.}
 The computational domain has length $L=2.5m$ and height $H=0.41m$. The circle center
@@
 time-stepping scheme
-\todo{. Remove the following
+\todo{OK: . Remove the following
 , at present the following are available.}
@@
 \subsection{Compliance minimization}
-\todo{Was ist MBB? Konnest Du vielleicht auch eine Referenz zur Konfiguration angeben?}\todocg{Wir sollten uns auf eine gemeinsame Darstellung der Bilder einigen (z.B. Achsenbeschriftung ja/nein).}
+\todo{Achsenbeschriftung Winni}
+\todo{Macht Winni: Was ist MBB?
+Konnest Du vielleicht auch eine Referenz zur Konfiguration angeben?}\todocg{Wir sollten uns auf eine gemeinsame Darstellung der Bilder einigen (z.B.
+Achsenbeschriftung ja/nein).}
 In this example, we consider the compliance minimization of a standard MBB-Beam. 
@@
 
-\todo{\section{Outlook}}
-\todo{Ich finde nach wir vor, dass wir kurz skizzieren sollten, was wir als
-  naechstes vor haben}
-\todo{In this work, we introduced
-the software DOpElib to solve PDEs
-and Optimization problems. In the future,
-we plan to implement efficient routines 
-for nonstationary optimization and different 
-types of control and state constraints. 
-Moreover, we plan to link to other 
-software packages, specifically, in order 
-to use parallel computing routines.  
-}
+\todo{Conclusions}
 

### 2012-11-23 Paper_May_2012/main.tex
neues style file von springer

@@
-\documentclass[prodmode,acmtoms]{acmsmall}
+%\documentclass{svjour3}                     % onecolumn (standard format)
+%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
+\documentclass[smallextended]{svjour3}       % onecolumn (second format)
+%\documentclass[twocolumn]{svjour3}          % twocolumn
 %
-
-% Package to generate and customize Algorithm as per ACM style
-\usepackage[ruled]{algorithm2e}
-\renewcommand{\algorithmcfname}{ALGORITHM}
-\SetAlFnt{\small}
-\SetAlCapFnt{\small}
-\SetAlCapNameFnt{\small}
-\SetAlCapHSkip{0pt}
-\IncMargin{-\parindent}
-
-% Metadata Information
-\acmVolume{0}
-\acmNumber{0} 
-\acmArticle{00}
-\acmYear{0000}
-\acmMonth{0}
-
+\smartqed  % flush right qed marks, e.g. at end of proof
 %
+\usepackage{graphicx}
+%
+% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
+%
+% insert here the call for the packages your document requires
+%\usepackage{latexsym}
+% etc.
+%
+% please place your own definitions here and don't use \def but
+% \newcommand{}{}
+%
+% Insert the name of "your journal" with
+ \journalname{Computing}
+%
+
 %%% User-requested packages placed after this line %%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \usepackage{listings}
+\usepackage{xcolor}
+\usepackage[numbers,sort&compress]{natbib}
 
@@
 %
+
 \begin{document}
-\markboth{C. Goll, T. Wick, and W. Wollner}{DOpElib:  A Goal Oriented Software Library}
 
@@
 
-\author{CHRISTIAN GOLL
-\affil{University of Heidelberg}
-THOMAS WICK
-\affil{The University of Texas at Austin}
-WINNIFRIED WOLLNER
-\affil{University of Hamburg}}
+\titlerunning{DOpElib}        % if too long for running head
+
+\author{C. Goll \and T. Wick \and W. Wollner}
+
+%\authorrunning{Short form of author list} % if too long for running head
+
+\institute{F. Author \at
+              first address \\
+              Tel.: +123-45-678910\\
+              Fax: +123-45-678910\\
+              \email{fauthor@example.com}           %  \\
+           \and
+Thomas Wick \at
+              The Institute for Computational Engineering and Sciences,\\
+The University of Texas at Austin,\\
+Austin, Texas 78712, USA \\
+              Tel.: +1 512-672-7763\\
+              \email{twick@ices.utexas.edu}   
+ \and
+F. Author \at
+              first address \\
+              Tel.: +123-45-678910\\
+              Fax: +123-45-678910\\
+              \email{fauthor@example.com}   
+}
+
+\date{Received: date / Accepted: date}
+% The correct dates will be entered by the editor
 
@@
 
+\maketitle
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{abstract}
-\todo{Computing}
 In this article, we describe the software library 
@@
 \end{abstract}
-
-\begin{bottomstuff}
-Author's adresses: C. Goll, Institut f\"ur Angewandte Mathematik,
-Universit\"at Heidelberg;
-T. Wick, The Institute for Computational Engineering and Sciences, The
-University of Texas at Austin;
-W. Wollner, Fachbereich Mathematik, Universit\"at Hamburg.
-\end{bottomstuff}
-
-\maketitle
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 equations (PDEs). Prominent examples are 
-\deal{} \cite{dealnew}, dune \cite{dune}, 
-FEniCS \cite{fenics}
+\deal{} \citep{dealnew}, dune \citep{dune}, 
+FEniCS \citep{fenics}
 or commercial solvers as for instance ANSYS (with its Fluent package),
@@
 \begin{lstlisting}
-  template<typename TRIANGULATION & triag, typename FE & fe>
+  template<typename TRIANGULATION & triag, 
+           typename FE & fe>
       void SpaceTimeHandler(Triangulation &, FE);
@@
   template<class NEWTON>
-      void StatPDEProblem<NEWTON>::ComputeReducedFunctionals();
+      void StatPDEProblem<NEWTON>
+             ::ComputeReducedFunctionals();
 \end{lstlisting}
@@
 
-\todocg{Haengt
-  von style-file ab: textbf finde ich besser als textit, haengt aber vom stylefile ab.}
 We split the program into the parts \texttt{main.cc}, \texttt{functionals.h} and \texttt{localpde.h}.
@@
  \begin{lstlisting}
-    for(unsigned int q_point = 0; q_point < n_q_points; q_point++)
+    for(unsigned int q_point = 0; 
+        q_point < n_q_points; 
+        q_point++)
     {
@@
       for (unsigned int i=0;i<n_dofs_per_cell;i++)
-        local_cell_vector(i) += ( _nu 
-                                * vgrads * phi_i_grads_v 
+        local_cell_vector(i) += 
+           ( _nu * vgrads * phi_i_grads_v 
              + //all the other terms)
@@
 \begin{lstlisting}
-    for(unsigned int q_point = 0; q_point < n_q_points; q_point++)
+    for(unsigned int q_point = 0; 
+        q_point < n_q_points; 
+        q_point++)
       erg += -1. * ( _nu * ufacegrads[q_point][0]
@@
 $(v,p)\in X$ with
-\[
-\int_I \bigl[ (\partial_t v, \phi)
+\begin{align*}
+&\int_I \bigl[ (\partial_t v, \phi)
 + \nu (\nabla v, \nabla \phi) + (v\cdot\nabla v, \phi)
 - (p,\nabla\cdot \phi)
-+ (\nabla\cdot v, \chi)\bigr] \, dt
-+ (v(0) - v^0, \phi(0))
++ (\nabla\cdot v, \chi)\bigr] \, dt\\
+&\quad + (v(0) - v^0, \phi(0))
 = \int_I l(\phi) \, dt 
-\]
+\end{align*}
 for all $\Phi:= (\phi, \chi) \in X$ with
@@
       local_cell_vector(i) += 
-          scale *// terms like (\nabla v^{n+1}, \nabla \phi) etc.
-          + scale_ico *// terms like (-p^{n+1},\nabla\cdot \phi)
+          scale * 
+          // terms like (\nabla v^{n+1}, \nabla \phi) etc.
+        + scale_ico *
+          // terms like (-p^{n+1},\nabla\cdot \phi)
      }
@@
    {
-     // implement directional derivative of the CellEquation
+     // implement directional derivative 
+     // of the CellEquation
    }
@@
 \centering
-{\includegraphics[width=6cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_mesh8070_scale.png}}
-{\includegraphics[width=6cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_x_velo8070_scale.png}}
+{\includegraphics[width=5cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_mesh8070_scale.png}}
+{\includegraphics[width=5cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_x_velo8070_scale.png}}
 \caption{FSI 2 test case: mesh (left) and velocity profile in vertical 
@@
 \end{document}
+% end of file template.tex

### 2012-11-26 Paper_May_2012/main.tex
Kommentare hinzugefuegt, mein Beispiel weiter aufgeraeumt, Einrueckung im Code an das neue, schmalere Design angepasst. Es gibt noch Probleme mit texttt im Text, TeX trennt diese Woerter nicht automatisch. Wollen wir Code-fragmente ueberhaupt trennen lassen?

@@
 
-\institute{F. Author \at
-              first address \\
-              Tel.: +123-45-678910\\
-              Fax: +123-45-678910\\
-              \email{fauthor@example.com}           %  \\
+\institute{Christian Goll \at
+              Institute of Applied Mathematics, University of Heidelberg
+              \\Im Neuenheimer Feld 293/294, 69120 Heidelberg, Germany\\
+              Tel.: +49 622-154-6154\\
+              \email{cgoll@iwr.uni-heidelberg.de}           %  \\
            \and
@@
 \begin{lstlisting}
-  template<typename TRIANGULATION & triag, 
-           typename FE & fe>
-      void SpaceTimeHandler(Triangulation &, FE);
+  template<typename TRIANGULATION, typename FE>
+     void SpaceTimeHandler(TRIANGULATION& triag, FE& fe);
 \end{lstlisting}
@@
 \begin{lstlisting}
-
   LocalPDE local_pde;
   LocalFunctional local_functional;
-
 \end{lstlisting}
@@
 \begin{lstlisting}
-
-  SpaceTimeHandler dof_handler(triangulation, finite_element);
-
+  SpaceTimeHandler dof_handler(triangulation,
+                               finite_element);
 \end{lstlisting}
@@
 \begin{lstlisting}
-
-  PDEProblemContainer<LocalPDE> prob_container(local_pde,
+  PDEProblemContainer<LocalPDE> 
+                      prob_container(local_pde,
                                      dof_handler);                                            
-
 \end{lstlisting}
@@
 \begin{lstlisting}
-
   prob_container.AddFunctional(&local_functional);
   prob_container.SetBoundaryFunctionalColors(1);
-
 \end{lstlisting}
@@
 \begin{lstlisting}
-
   DirichletData dd;
-  prob_container.SetDirichletBoundaryColors(0, comp_mask, dd);
-
+  prob_container.SetDirichletBoundaryColors(0,
+                                            comp_mask,
+                                            dd);
 \end{lstlisting}
@@
 \begin{lstlisting}
-
   typedef NewtonSolver<LINSOLVE> NEWTON;
-
 \end{lstlisting}
@@
 \begin{lstlisting}
-
   typedef DirectLinearSolver<MATRIX, VECTOR> LINSOLVE;
-
 \end{lstlisting}
@@
 \begin{lstlisting}
-
   PDE pde(prob_container, pr,
                     quad_formula, face_quad_formula);
-
 \end{lstlisting}
@@
 \begin{lstlisting}
-
   pde.ComputeReducedFunctionals();
-
 \end{lstlisting}
@@
     const auto& fe_values = cdc.GetFEValuesState();
-      unsigned int n_dofs_per_cell = cdc.GetNDoFsPerCell();
+    unsigned int dofs_per_cell = cdc.GetNDoFsPerCell();
     unsigned int n_q_points = cdc.GetNQPoints();
@@
       vgrads[1]= ugrads[q_point][1];
-      for (unsigned int i=0;i<n_dofs_per_cell;i++)
+      for (unsigned int i=0;i<dofs_per_cell;i++)
         local_cell_vector(i) += 
@@
 hand, is that we do not have to specify the method \texttt{CellEquation}, but
-\texttt{BoundaryValue} instead because $J$ is just number to be computed.
-This is due to the fact that we consider a boundary-functional in this example, and so we have to specify the value of the functional not cell-wise but face wise (for the faces that lie on $\Gamma_{circ}$).
+\texttt{BoundaryValue} instead because $J$ is just a number to be computed.
+This is due to the fact that we consider a boundary-functional in this example, and so we have to specify the value of the functional not cell-wise but face-wise (for the faces that lie on $\Gamma_{circ}$).
 
@@
 \begin{lstlisting}
-  template<typename TIMES & times, 
-           typename TRIANGULATION & triag, 
-           typename FE & fe>
-   void SpaceTimeHandler(Times &, Triangulation &, FE &);
+  template<typename TIMES, typename TRIANGULATION, 
+           typename FE>
+   void SpaceTimeHandler(TIMES& times,
+                         TRIANGULATION& triag,
+                         FE& fe);
 \end{lstlisting}
@@
 loops over all time-steps, where in each of the 
-time-steps we can reuse the functionality to solve stationary PDEs. 
+time-steps we can reuse the functionality \todocg{we employed} to solve stationary PDEs.\todocg{Hier Beispiele, was genau wir wiederverwenden? Oder auf den naechsten Abschnitt verweisen?} 
 Now it is apparent, why we introduced the \texttt{StatPDEProblem}, because
@@
          scale * 
-          // terms like (\nabla v^{n+1}, \nabla \phi) etc.
+         // terms like (\nabla v^{n+1}, \nabla\phi) etc.
         + scale_ico *
-          // terms like (-p^{n+1},\nabla\cdot \phi)
+         // terms like (-p^{n+1},\nabla\cdot\phi)
      }
@@
 \begin{equation}
-\mathcal {J}( u) = \frac 1 {20} \int_{\Gamma_{circ}} \nu\partial_nv _1 - pn,
+J(u) = \frac 1 {20} \int_{\Gamma_{circ}} \nu\partial_nv _1 - pn,
 \end{equation}
 with the solution $u = (v,p)$, the normal $n$ and the viscosity $\nu = 0.001$.
-\todo{For CG: Bildunterschrift und die $2$ am Ende und 
-das cal J}
 \begin{figure}[hbt]
 \centering
-{\includegraphics[width=0.8\textwidth]{Pictures/local_grid_NS.png}}
-\caption{Caption.}
+{\includegraphics[width=0.9\textwidth]{Pictures/local_grid_NS.png}}
+\caption{Locally adapted grid after eight refinement steps.}
 \label{fig:NS_local_grid}
 \end{figure}
-To achieve an efficient computation of the quantity $\mathcal J$, we employ goal-oriented mesh refinement with the help of the DWR-method, see \cite{BeRa96}.
-The method delivers error indicators and by this way computational grids tailored to minimize the error in the previously given functional $\mathcal J$.
+To achieve an efficient computation of the quantity $J$, we employ goal-oriented mesh refinement with the help of the DWR-method, see \cite{BeRa96}.
+The method delivers error indicators and by this way computational grids tailored to minimize the error in the previously given functional $J$.
 The price we have to pay ist the additional computation of a (linear) dual problem, which is implemented as described in section~\ref{sec:opt}.
@@
 
-\ref{fig:NS_local_grid}
 \begin{figure}[hbt]
@@
 \resizebox{0.5\textwidth}{!}{\input{Pictures/ConvergenceNS}}
-\caption{Comparison of global and local mesh refinement with respect to the error in the drag-functional $\mathcal J$.}
+\caption{Comparison of global and local mesh refinement with respect to the error in the drag-functional $J$.}
 \label{fig:NS_comparison}
@@
 \section{Conclusions}
-\todo{Frage an CG und Winni: Guter Abschluss?}
+\todo{Frage an CG und Winni: Guter Abschluss?}\todocg
 In this article, we described the architecture 
@@
 the algorithms is the computation of some target 
-functional. This is achieved by unified interface 
-to state-of the art time-stepping methods, nonlinear solvers and optimization
-routines. The 
-capabilites of DOpElib are demonstrated
-wit the help of three examples.
+functional \todocg{Ich wuerde eventuell die userfreundlichkeit mehr in den Fokus ruecken?
+The key idea of the software is to supply a toolkit which allows for fast implementation of the problems at hand, but at the same time gives the user full flexibility to change or alter the used algorithms}.
+This is achieved by a unified interface
+to state of the art time-stepping methods, nonlinear solvers and optimization
+routines. The capabilites of DOpElib are demonstrated with the help of three examples.
 

### 2012-11-26 Paper_May_2012/main.tex
Meine Beitraege wie besprochen.

@@
  \and
-F. Author \at
-              first address \\
-              Tel.: +123-45-678910\\
-              Fax: +123-45-678910\\
-              \email{fauthor@example.com}   
+W. Wollner \at
+              Universit\"at Hamburg\\
+              Department of Mathematics\\
+              \email{winnifried.wollner@math.uni-hamburg.de}   
 }
@@
 which fulfill the requirements listed above; 
-sucessful in-house code examples are Gascoigne \cite{gascoigne}
+successful in-house code examples are Gascoigne \cite{gascoigne}
 and RoDoBo \cite{rodobo}. 
@@
 \end{lstlisting}
-The template argument \texttt{NEWTON} represents the choosen solver of the nonlinear problems.
+The template argument \texttt{NEWTON} represents the chosen solver of the nonlinear problems.
 That combines all the necessary steps to calculate $J(u_h)$.
@@
 
-We split the program into the parts \texttt{main.cc}, \texttt{functionals.h} and \texttt{localpde.h}.
-The \texttt{main.cc}-file is responsible for the declaration of various objects which have mostly to do with the \textit{discretization}, the \textit{solvers} and the general \textit{program sequence}, whereas the other two contain most of the problem-data -- especially the cell-wise contributions of the weak form, the right hand side and the functional.
+We split the program into the parts \texttt{main.cc}, \texttt{func.h} and \texttt{pde.h}.
+The \texttt{main.cc}-file is responsible for the declaration of various objects which have mostly to do with the \textit{discretization}, the \textit{solvers} and the general \textit{program sequence}, whereas the other two contain most of the problem-data -- especially the cell-wise contributions of the weak form in \texttt{pdh.h} and the functional in \texttt{func.h}.
 We give a brief explanation on every file in the following.
@@
 
-Now we have everything at hand to create an instantiation of type \texttt{PDEProblemContainer}.
+Now we have everything at hand to create an instantiation of the container for the problem data \texttt{PDEProblemContainer}.
 This basically represents the discretized problem, and bundles things like boundary conditions, the PDE, the functional and the discretization.
@@
 \end{lstlisting}
-Next, we incorporate the Dirichlet data, which is given by the class
-\texttt{DirichleData}. Then we have to tell the \texttt{prob\_container} for
+Next, we incorporate the Dirichlet data. To this end, we instantiate a 
+class \texttt{DirichletData} containing the required description of the 
+boundary values. Then we have to tell the \texttt{prob\_container} for
 which components and boundaries we prescribe the Dirichlet conditions. This is
@@
 
-\subsubsection{\texttt{localpde.h}}
+\subsubsection{\texttt{pde.h}}
 We shall now give an overview to the implementation of the specific form of the PDE.
@@
 The corresponding Jacobian can be defined analogously.
-Now all we have to supply here are the cell-wise contributions to these terms, the \texttt{Integrator} takes care of assembling \eqref{eq:discrete_residual} and \eqref{eq:discrete_matrix}.
+Now all we have to supply here are the cell-wise contributions to these terms, the \texttt{Integrator} takes care of assembling \eqref{eq:discrete_residual}
+and similarly of the matrix for the linearized equation to be solved.
 
@@
 
-We show the basic structure of these methods using the example of \texttt{CellEquation}.
+We show the basic structure of these methods using the \texttt{CellEquation} 
+method for the implementation of the element-wise integrals.
 
@@
            ( _nu * vgrads * phi_i_grads_v 
-             + //all the other terms)
-             * state_fe_values.JxW(q_point);
+             + //all the other terms
+           ) * fe_values.JxW(q_point);
     }
@@
 
-\subsubsection{\texttt{localfunctional.h}}
+\subsubsection{\texttt{localfunc.h}}
 The implementation of the boundary-functional $J$ follows the outline of the
 computation of the cell-residual in \texttt{localpde.h}. 
-\todo{CG and WW: Pruefen, ob der folgende Absatz nun gut ist:}
-The main difference is that as we only want to evaluate a functional and not a
-residual-vector, we just loop over the quadrature points and not over the
-degrees of freedom on the face. Another deviation, special to the case at
-hand, is that we do not have to specify the method \texttt{CellEquation}, but
-\texttt{BoundaryValue} instead because $J$ is just a number to be computed.
-This is due to the fact that we consider a boundary-functional in this example, and so we have to specify the value of the functional not cell-wise but face-wise (for the faces that lie on $\Gamma_{circ}$).
-
-%However, there are two main differences.
-First, we do not have to specify the method \texttt{CellEquation}, but \texttt{BoundaryValue} instead.
-This is due to the fact that we consider a boundary-functional in this example, and so we have to specify the value of the functional not cell-wise but face wise (for the faces that lie on $\Gamma_{circ}$).
-The second difference is that as we only want to evaluate a functional and not a residual-vector, we just loop over the quadrature points and not over the degrees of freedom on the face.
+\todo{Habe ich mal umgeschrieben, bitte nochmal ansehen~WW.}
+The main difference, and common to all functionals is that as we only 
+want to evaluate a functional, i.e., calculate a single number, and not a
+residual-vector, i.e., calculate the elements of a vector. Thus, we 
+just loop over the quadrature points and not over the degrees of freedom 
+on the face. In the present situation, we want to evaluate a functional 
+that consists of a boundary integral. Thus we need to evaluate 
+integrals on faces, that lie on $\Gamma_{circ}$, instead of integrals over
+elements. 
+This face contributions are calculated in \texttt{BoundaryValue} and added to 
+sum up to the value of $J$.
+
 \begin{lstlisting}
@@
     //same as before, get n_q_points etc.
+    const auto& face_values = cdc.GetFEFaceValuesState();
     fdc.GetFaceValuesState("state", ufacevalues);
@@
         q_point++)
-      erg += -1. * ( _nu * ufacegrads[q_point][0]
-                * state_fe_face_values.normal_vector(q_point)
-                + //other terms)        
-                * state_fe_face_values.JxW(q_point);
+      erg -= ( _nu * ufacegrads[q_point][0]
+               * face_values.normal_vector(q_point)
+               + //other terms        
+             ) * face_values.JxW(q_point);
     return erg;
@@
 \begin{remark}
-Note that in the stationary case \dope{} automatically sets $ \texttt{scale\_ico} = \texttt{scale}$, so that in fact the user has to implement the \texttt{CellEquation} only once and can henceforth reuse it in all his stationary an instationary problems.
+Note that in the stationary case there is no such difference and accordingly 
+\dope{} automatically sets $ \texttt{scale\_ico} = \texttt{scale}$, so that in fact the user has to implement the \texttt{CellEquation} only once and can henceforth reuse it in all his stationary and nonstationary problems.
 This transfers of course also to other methods like \texttt{CellMatrix}.
@@
 
-In addition to stationary problems, the implementation
-of the time-derivative, i.e. the term $ (v^{n+1},\phi)$, must be considered which is done
+In comparison to stationary problems, we need to consider, in addition, 
+the implementation
+of the time-derivative, i.e., the term $ (v^{n+1},\phi)$. This is done
 in the method \texttt{CellTimeEquation} in a straight-forward way.
@@
 \begin{lstlisting}
-        local_cell_vector(i) +=  scale * (v * phi_i) * 
-                    state_fe_values.JxW(q_point);    
+    local_cell_vector(i) +=  scale * (v * phi_i)  
+                             * fe_values.JxW(q_point);    
 \end{lstlisting}
@@
 \subsection{Solving of optimization problems}\label{sec:opt}
-\todo{For WW: etwas klarer machen, dass wir instationare Optimierung koennen}
-Upgrading the framework for solving optimization problems
-is described in this section. A prototypical problem reads:
+Once we are able to solve stationary and nonstationary PDEs
+we can consider optimization problems constrained by these PDEs.
+How this is done will be described in this section.
+
+A prototypical PDE constrained optimization problem reads:
 \begin{align*}
@@
   &\text{s.t.}\; a(q,u)(\phi) = 0 \quad \forall \phi\in V,\\
-  &a \le q \le b,\\
+  &q_- \le q \le q_+,\\
   &g(q,u) \le 0,  
 \end{align*}
-where $u$ is a FE-function and $q$ can either be a FE-function or some 
-fixed number of parameters, $a$ and $b$ are constraint bounds for the control $q$,
-and $g(\cdot)$ is some state constraint.
+where $u$ is a function and $q$ can either be a function or some 
+fixed number of parameters. the numbers $q_-$ and $q_+$ are bounds for 
+the control $q$, and $g(\cdot)$ is some state constraint.
 To keep the presentation brief we will neglect the additional bounds
@@
 reduced approach, in which we eliminate the dependent variable $u$ by the solution 
-operator $S$ of the PDE, i.e., 
+operator $S$ of the (stationary or nonstationary) PDE, i.e., 
 \begin{align*}
@@
 \end{align*}
-Then, one can apply any algorithm to solve the, unconstraint, minimization 
-problem. This, will in any case require the evaluation of $j(q)$ which 
+This can be discretized using the finite element approximation $S_h$ of $S$ 
+and yields the problem 
+\begin{align*}
+\min\;&j_h(q) = J(q,S_h(q)). 
+\end{align*}
+After possibly discretizing the control $q$ as well using finite elements, 
+one can apply any algorithm to solve the, unconstraint, minimization 
+problem. This, will in any case require the evaluation of $j_h(q)$ which 
 we can do, e.g., using \texttt{StatPDEProblem::ComputeReducedFunctional} for 
-a stationary PDE. If one would like to use derivative based methods, one 
-needs in addition derivatives of $j$ and thus of $J$ as well as of $S$. 
+a stationary PDE or using \texttt{InstatPDEProblem::ComputeReducedFunctional} 
+for nonstationary PDEs.
+
+If one would like to use derivative based methods, one 
+needs in addition derivatives of $j_h$ and thus of $J$ as well as of $S_h$. 
 To do so, we remark, that one can calculate the derivative of 
@@
 \[
-j'(q)(\delta q) = J'_q(q,S(q))(\delta q) + (S^* J_u'(q,S(q)),\delta q)
+j'_h(q)(\delta q) = J'_q(q,S_h(q))(\delta q) + (S_h^* J_u'(q,S_h(q)),\delta q)
 \]
 where the index $q$ or $u$ denote directional derivatives for the 
-respective argument. To calculate the action of the adjoint operator $S^*$ 
-we note that $z =  S^* J_u'(q,S(q))$
+respective argument. To calculate the action of the adjoint operator $S_h^*$ 
+we note that $z =  S_h^* J_u'(q,S_h(q))$
 can be computed via the adjoint PDE
@@
 To be more precise, we are computing the benchmark 2D-1 from \cite{TuSchae96}.
-The computational domain is as depicted in figure~\ref{fig:example_ns} and the equation is discretized with the Taylor-Hood  element.
+The computational domain is as depicted in Fig.~\ref{fig:example_ns} and the equation is discretized with the Taylor-Hood  element.
 We want to compute the drag force on the enclosed cylinder
@@
 The method delivers error indicators and by this way computational grids tailored to minimize the error in the previously given functional $J$.
-The price we have to pay ist the additional computation of a (linear) dual problem, which is implemented as described in section~\ref{sec:opt}.
-In figure~\ref{fig:NS_local_grid} we see a resulting triangulation after eight refinement steps with roughly $320\,000$ degrees of freedom.
-
-In figure~\ref{fig:NS_comparison} we compare the errors in the functional obtained by using global and local mesh refinement.
+The price we have to pay is the additional computation of a (linear) dual problem, which is implemented as described in section~\ref{sec:opt}.
+In Fig.~\ref{fig:NS_local_grid} we see a resulting triangulation after eight refinement steps with roughly $320\,000$ degrees of freedom.
+
+In Fig.~\ref{fig:NS_comparison} we compare the errors in the functional obtained by using global and local mesh refinement.
 We see clearly that the local mesh refinement allows for substantial savings with respect to degrees of freedom.
@@
 The basic configuration is 
-sketched in Figure \ref{fig:example_ns} at which an elastic beam is attached 
+sketched in Fig.~\ref{fig:example_ns} at which an elastic beam is attached 
 behind the rigid cylinder. 
@@
 Since the problem is fully nonstationary, the 
-dynamics are presents in Figure \ref{res:fsi_2_mesh_and_x_velo}. 
-By comparison of our findings in Figure \ref{res:results_ux_and_uy_fsi_2}
+dynamics are presents in Fig.~\ref{res:fsi_2_mesh_and_x_velo}. 
+By comparison of our findings in Fig.~\ref{res:results_ux_and_uy_fsi_2}
 with the literature, it can be identified that 
@@
 \subsection{Compliance minimization}
-\todo{For Winni: Achsenbeschriftung, MBB ausschreiben, 
-Konfiguration citieren}
-In this example, we consider the compliance minimization of a standard MBB-Beam. 
+In this example, we consider the compliance minimization of a standard MBB-Beam (Messerschmidt-B\"olkow-Blohm-Beam), see, e.g.,~\cite{BendsoeSigmund:2003}.
 The thickness variable is allowed to take intermediate value apart from $0$ and $1$,
@@
 \centering
-\includegraphics[width=1.\textwidth, viewport=150 0 1024 500, clip]{Pictures/MBB.png}
+\includegraphics[width=1.\textwidth, viewport=205 88 975 500, clip]{Pictures/MBB.png}
 \caption{Thickness distribution for the MBB-beam minimum compliance problem.} 
@@
 \end{figure}
-A typical thickness distribution is shown in Figure~\ref{res:mbb}.
+A typical thickness distribution is shown in Fig.~\ref{res:mbb}.
 
 \section{Conclusions}
-\todo{Frage an CG und Winni: Guter Abschluss?}\todocg
+\todo{Frage an CG und Winni: Guter Abschluss?}
 In this article, we described the architecture 
@@
 functional \todocg{Ich wuerde eventuell die userfreundlichkeit mehr in den Fokus ruecken?
-The key idea of the software is to supply a toolkit which allows for fast implementation of the problems at hand, but at the same time gives the user full flexibility to change or alter the used algorithms}.
+The key idea of the software is to supply a toolkit which allows for fast implementation of the problems at hand, but at the same time gives the user full flexibility to change or alter the used algorithms}\todo{Bin daf\"ur~WW}.
 This is achieved by a unified interface
 to state of the art time-stepping methods, nonlinear solvers and optimization
-routines. The capabilites of DOpElib are demonstrated with the help of three examples.
+routines. The capabilities of DOpElib are demonstrated with the help of three examples.
 

### 2012-11-26 Paper_May_2012/main.tex
Antwort auf Winnis Kommentar.

@@
 \end{lstlisting}
-Finally, we compute the functional $J$ by calling:
+Finally, we compute the value $J(u)$ by calling:
 \begin{lstlisting}
@@
 \todo{Habe ich mal umgeschrieben, bitte nochmal ansehen~WW.}
-The main difference, and common to all functionals is that as we only 
+The main difference, and common to all functionals is that we only 
 want to evaluate a functional, i.e., calculate a single number, and not a
-residual-vector, i.e., calculate the elements of a vector. Thus, we 
+residual-vector\todocg{, i.e., calculate the elements of a vector Diesen Teil wuerde ich streichen.}. Thus, we 
 just loop over the quadrature points and not over the degrees of freedom 

### 2012-11-26 Paper_May_2012/main.tex
Frage von mir mit Christian geklaert und entsprechender Formulierungsvorschlag eingebaut.

@@
 \end{itemize}
-\todo{CG and WW: bitte pruefen}
+\todo{bitte gegenlesen}
 After choosing the time-stepping scheme, the SpaceTimeHandler  
 loops over all time-steps, where in each of the 
-time-steps we can reuse the functionality \todocg{we employed} to solve stationary PDEs.\todocg{Hier Beispiele, was genau wir wiederverwenden? Oder auf den naechsten Abschnitt verweisen?} 
+time-steps we can reuse the functionality we have provided to solve 
+stationary PDEs. 
+This means for instance, if we have implemented the 
+cell-integrals in the weak formulation for the stationary case, 
+then we can use the exact same 
+implementation again for the solution of the nonstationary version of 
+the equation. However, we will possibly need to provide an additional 
+term to cope with the integral for the time derivative.
+
 Now it is apparent, why we introduced the \texttt{StatPDEProblem}, because

### 2012-11-26 Paper_May_2012/main.tex
Antwort auf Winnis letzte Aenderung

@@
 \todo{Habe ich mal umgeschrieben, bitte nochmal ansehen~WW.}
-The main difference, and common to all functionals is that we only 
-want to evaluate a functional, i.e., calculate a single number, and not a
+The main difference, and common to all functionals is that we only want to
+\todocg{ evaluate a functional, i.e., <-wuerde ich streichen} calculate a single number, and not a
 residual-vector\todocg{, i.e., calculate the elements of a vector Diesen Teil wuerde ich streichen.}. Thus, we 

### 2012-11-26 Paper_May_2012/main.tex
Accept with last minor todo: in Christian's Drag Formal ist meiner Ansicht nach ein Fehler. Ich habe einen Vorschlag als Todo mit Literaturangabe. Bitte anschauen. Ansonsten sind wir nun fertig bis auf Rechtschreibpruefung

--- 2012-11-26 Paper_May_2012/main.tex
@@
 computation of the cell-residual in \texttt{localpde.h}. 
-\todo{Habe ich mal umgeschrieben, bitte nochmal ansehen~WW.}
-The main difference, and common to all functionals is that we only 
-want to evaluate a functional, i.e., calculate a single number, and not a
-residual-vector\todocg{, i.e., calculate the elements of a vector Diesen Teil wuerde ich streichen.}. Thus, we 
+The main difference, and common to all functionals is that we only want to
+\todocg{ evaluate a functional, i.e., <-wuerde ich streichen} calculate a single number, and not a
+whole 
+residual-vector. Thus, we 
 just loop over the quadrature points and not over the degrees of freedom 
@@
 \end{itemize}
-\todo{bitte gegenlesen}
 After choosing the time-stepping scheme, the SpaceTimeHandler  
@@
 implementation again for the solution of the nonstationary version of 
-the equation. However, we will possibly need to provide an additional 
+the equation. However, we need to provide an additional 
 term to cope with the integral for the time derivative.
@@
 \end{equation}
+\todo{
+\begin{equation}
+J(u) = \frac 1 {20} \int_{\Gamma_{circ}} (\nu\partial_nv _1 - pn)e_1,
+\end{equation}
+where $e_1$ denotes the unit vector in $x$-direction. Habil 
+Roland Becker S. 60 im pdf.
+}
 with the solution $u = (v,p)$, the normal $n$ and the viscosity $\nu = 0.001$.
@@
 \section{Conclusions}
-\todo{Frage an CG und Winni: Guter Abschluss?}
 In this article, we described the architecture 
 of the novel software library DOpElib for solving 
-PDEs and optimization problems.  The key idea behind 
-the algorithms is the computation of some target 
-functional \todocg{Ich wuerde eventuell die userfreundlichkeit mehr in den Fokus ruecken?
-The key idea of the software is to supply a toolkit which allows for fast implementation of the problems at hand, but at the same time gives the user full flexibility to change or alter the used algorithms}\todo{Bin daf\"ur~WW}.
+PDEs and optimization problems.  The key idea of the software is to supply a
+toolkit which allows for fast implementation of the problems at hand, but at
+the same time gives the user full flexibility to change or alter the used
+algorithms. 
 This is achieved by a unified interface 

### 2012-11-26 Paper_May_2012/main.tex
Todo beseitigt welches sich durch unser gleichzeitiges einchecken ergeben hat
Thomas' Verbesserung eingearbeitet

@@
 The main difference, and common to all functionals is that we only want to
-\todocg{ evaluate a functional, i.e., <-wuerde ich streichen} calculate a single number, and not a
-whole 
-residual-vector. Thus, we 
+evaluate a functional. Thus, we 
 just loop over the quadrature points and not over the degrees of freedom 
@@
 \begin{equation}
-J(u) = \frac 1 {20} \int_{\Gamma_{circ}} \nu\partial_nv _1 - pn,
+J(u) = \frac 1 {20} \int_{\Gamma_{circ}} \nu\partial_nv _1 - pn_1,
 \end{equation}
-\todo{
-\begin{equation}
-J(u) = \frac 1 {20} \int_{\Gamma_{circ}} (\nu\partial_nv _1 - pn)e_1,
-\end{equation}
-where $e_1$ denotes the unit vector in $x$-direction. Habil 
-Roland Becker S. 60 im pdf.
-}
 with the solution $u = (v,p)$, the normal $n$ and the viscosity $\nu = 0.001$.

### 2012-11-27 Paper_May_2012/main.tex
Aenderungen und Kommentare von cg bis zum Beginn von 2.2

@@
 \label{introduction}
-Today, there exist a broad variety of software libraries
+Today, there exists a broad variety of software libraries
 that provide numerical tools for solving partial differential
@@
 target functional, e.g., error norms,
-point values, or certain mean-values as drag or lift coefficients.
+point values, or certain mean-values such as drag or lift coefficients.
 As a second motivating principle, 
@@
 and RoDoBo \cite{rodobo}. 
-Like DOpElib, both have been initiated as well in 
+Like \dope{}, both have been initiated as well in 
 the Numerical Analysis Group in Heidelberg.
@@
 time-stepping methods, nonlinear solvers and optimization routines. 
-\dope{} is designed such a way that the user only needs to write those parts
+\dope{} is designed in such a way that the user only needs to write those parts
 of the code that are problem dependent while all invariant 
@@
 
-The project was initially initiated at the University of Heidelberg in 2010 and  
+The project was initially initiated \todocg{initially initiated - doppelt gemoppelt?} at the University of Heidelberg in 2010 and  
 at the present stage the following key features are supported by the library
@@
 This library is designed to allow easy implementation for the numerical solution 
-of problems involving partial differential equations (PDEs). 
+of problems involving partial differential equations (PDEs)\todocg{Die Abkuerzung PDE wird auf Seite 1 schon erlaeuternt}. 
 In particular, we remove all tedious routine implementation needed in 
@@
 Then typically we do not only want to calculate $u$ but more likely we are 
-interested in one (or several) functionals depending on the solution, i.e., some numbers $J$.
+interested in one (or several) functionals depending on the solution, i.e., some numbers $J(u)$.
 
@@
 \begin{description}
-\item[Discretization] Specify what kind of triangulation:
+\item[\textit{Discretization}]\todocg{Hier scheint etwas verlorengegangen zu sein, der Satz macht so keinen Sinn:} Specify what kind of triangulation:
 in most cases this is the
@@
 things might be adjusted as well.
-\item[Problem-data] Define the weak form, the corresponding matrices, the
+\item[\textit{Problem-data}] Define the weak form, the corresponding matrices, the
   functionals $J$ as well as the boundary conditions. With exception of the boundary conditions, we have to specify all this only cell-wise, \dope{} handles the rest for us.
-\item[Solvers] Choice of a (non-)linear solver. 
-\item[Additional features] ParameterReader to read in the run-time parameters
+\item[\textit{Solvers}] Choice of a (non-)linear solver. 
+\item[\textit{Additional features}] ParameterReader to read in the run-time parameters
   such as material coefficients or adjustments for the solvers like tolerance
@@
 
-We describe now a program designed to compute a value $J$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in figure~\ref{fig:example_ns} with  $f=0$, and $J$ is the drag-force of the enclosed circle.
+We describe now a program designed to compute a value $J(u)$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in figure~\ref{fig:example_ns} with  $f=0$, and $J$ is the drag-force of the enclosed circle.
 We assume given Dirichlet data for the velocity $v$ on the boundary $\Gamma_{D}=\Gamma_{in}\cup\Gamma_{wall}\cup\Gamma_{circ}$.
@@
 \subsubsection{\texttt{main.cc}}
-In this file we gather the basic components that are necessary for the
+In this file we gather all the basic components that are necessary for the
 solution process of a (stationary) partial differential equation. The
@@
 consideration, what changes is typically just the \textit{kind} of  finite
-element or the linear solver. Here, a key issue is that we only need to pick 
+element or linear solver we want to use. Here, a key issue is that we only need to pick 
 some pre-implemented routines from the \dope{} library rather than 
@@
   PDEProblemContainer<LocalPDE> 
-                      prob_container(local_pde,
-                                     dof_handler);                                            
+             prob_container(local_pde, dof_handler);                                            
 \end{lstlisting}
@@
 which components and boundaries we prescribe the Dirichlet conditions. This is
-done via a boundary-color in a so-called component mask 
+done via a boundary-color and a so-called component mask 
 (here, $0$ denotes $\Gamma_D$).
@@
 \begin{lstlisting}
-
   typedef StatPDEProblem<NEWTON> PDE;
-
 \end{lstlisting}
@@
 \end{lstlisting}
-Finally, we compute the value $J(u)$ by calling:
+Finally, we compute the value $J(u_h)$ by calling:
 \begin{lstlisting}
@@
 
-In the next lines, we finally implement our concrete PDE. Here, aim to provide 
+In the next lines, we finally implement our concrete PDE. Here, we aim to provide 
 an interface which is closely related to the weak formulation which 
@@
 The main difference, and common to all functionals is that we only want to
-evaluate a functional. Thus, we 
+evaluate a functional \todocg{Ist hier was verrutscht? der Satz passt im Moment nicht so wirklich, der Teil mit dem Vektor ist irgendwie verlorengegangen?}. Thus, we 
 just loop over the quadrature points and not over the degrees of freedom 
@@
 that consists of a boundary integral. Thus we need to evaluate 
-integrals on faces, that lie on $\Gamma_{circ}$, instead of integrals over
+integrals on faces, those who lie on $\Gamma_{circ}$, instead of integrals over
 elements. 
@@
 sum up to the value of $J$. 
-
 \begin{lstlisting}
@@
 In this article, we described the architecture 
-of the novel software library DOpElib for solving 
+of the novel software library \dope{} for solving 
 PDEs and optimization problems.  The key idea of the software is to supply a
@@
 to state of the art time-stepping methods, nonlinear solvers and optimization
-routines. The capabilities of DOpElib are demonstrated with the help of three examples.
+routines. The capabilities of \dope{} are demonstrated with the help of three examples.
 

### 2012-11-27 Paper_May_2012/main.tex
Aenderungen und Kommentare von cg ab 2.2 bis zum Ende

@@
 PDEs, we consider a time-dependent semi-linear form $a(u)(\phi)$ and some
-right hand side functional $l(\phi)$.
+right hand side functional $l$.\todocg{Warum hier l und nicht f?}
 Let $I:=[0,T]$ be a time interval with end time point value $T$.
@@
 \[
-X:= \bigl\{ 
+X:= \set{ 
 (v,p) | v\in L^2(I,H_0^1(\Omega)^d) , 
 \partial_t v\in L^2(I, L^2(\Omega)^d), 
-p\in L^2(I,L^2(\Omega)\setminus\mathbb{R} \bigr\}.
+p\in L^2(I,L^2(\Omega)\setminus\mathbb{R}}.
 \]
@@
 \item Crank-Nicolson scheme
-\item shifted Crank-Nicolson scheme (to deal with singular initial data)
-\item fractional-step-$\theta$ scheme
+\item Shifted Crank-Nicolson scheme (to deal with singular initial data)
+\item Fractional-Step-$\theta$ scheme
 \end{itemize}
-After choosing the time-stepping scheme, the SpaceTimeHandler  
+After choosing the time-stepping scheme, the \texttt{SpaceTimeHandler}  
 loops over all time-steps, where in each of the 

### 2012-11-27 Paper_May_2012/main.tex
Aenderungen und Kommentare von cg ab 2.2 bis zum Ende, hatte nicht gespeichert

@@
 This means for the user only the derivatives of the functional and the PDE 
-needs to be provided in addition to the 
+needs to be provided in addition to the  \todocg{Verstehe diesen Satz nicht. Das sind doch die Ableitungen des Funktionals und der PDE, warum addition?}
 \begin{lstlisting} 

### 2012-11-27 Paper_May_2012/main.tex
Paper: Meine Anmerkungen eingefuegt und teilweise Anmerkungen von CG bereits bearbeitet

@@
 \begin{abstract}
+\todo{Wird die Webseite genannt? ist das sinnvoll dies zu tun?}
 In this article, we describe the software library 
@@
 
+\todo{Das wir in Heidelberg angefangen haben wird schon vorher erwaehnt!
+  Wuerde den Satz einfach streichen}
 The project was initially initiated \todocg{initially initiated - doppelt gemoppelt?} at the University of Heidelberg in 2010 and  
@@
    Poisson, Navier-Stokes, Plasticity and fluid-structure interaction problems. 
-\item Line search and trust region newton algorithms for the 
+\item Line search and trust region Newton algorithms for the 
    solution of optimization problems with PDE constraints by elimination of 
@@
 This library is designed to allow easy implementation for the numerical solution 
-of problems involving partial differential equations (PDEs)\todocg{Die Abkuerzung PDE wird auf Seite 1 schon erlaeuternt}. 
+of problems involving PDEs.
 In particular, we remove all tedious routine implementation needed in 
@@
 of the domain $\Omega$ and a finite element. In the concrete case, we could, 
+\todo{Warum Tylor-Hood? Ist inf-sup stabil wuerde ich erwaehnen}
 for instance, take the $\mathcal Q_2/\mathcal Q_1$ Taylor-Hood element.
@@
 \end{align}
+\todo{wir schreiben newton immer klein. eigentlich ist das ein Eigenname: Newton}
 To this end, one can employ a newton type method with some globalization 
@@
 \begin{description}
-\item[\textit{Discretization}]\todocg{Hier scheint etwas verlorengegangen zu sein, der Satz macht so keinen Sinn:} Specify what kind of triangulation:
+\item[\textit{Discretization}]\todocg{Hier scheint etwas verlorengegangen zu
+    sein, der Satz macht so keinen Sinn:} 
+\todo{Vorschlag: Specification of triangulation issues such as finite elements
+  and quadrature rules}  
+Specify what kind of triangulation:
 in most cases this is the
@@
 things might be adjusted as well.
-\item[\textit{Problem-data}] Define the weak form, the corresponding matrices, the
+\item[\textit{Problem-data}] \todo{Definition of } Define the weak form, the corresponding matrices, the
   functionals $J$ as well as the boundary conditions. With exception of the boundary conditions, we have to specify all this only cell-wise, \dope{} handles the rest for us.
@@
 
+\todo{Ich wuerde localfunctional als volles Wort schreiben}
 \subsubsection{\texttt{localfunc.h}}
@@
 The main difference, and common to all functionals is that we only want to
-evaluate a functional \todocg{Ist hier was verrutscht? der Satz passt im Moment nicht so wirklich, der Teil mit dem Vektor ist irgendwie verlorengegangen?}. Thus, we 
+evaluate a functional \todocg{Ist hier was verrutscht? der Satz passt im
+  Moment nicht so wirklich, der Teil mit dem Vektor ist irgendwie
+  verlorengegangen?}. 
+\todo{Ich hatte Deine Anmerkung so verstanden, als ob Du alles nach dem i.e.,
+  geloescht 
+haben wolltest. Im main TW steht noch das Original}
+Thus, we 
 just loop over the quadrature points and not over the degrees of freedom 
@@
 right hand side functional $l$.\todocg{Warum hier l und nicht f?}
+\todo{Antwort: Weil $f$ die volumen kraft ist. Wir wollten 
+doch Funktional und Kraftvektor unterscheiden. Daher hier $l$. Es gilt :
+$l(\phi) = (f,\phi)$.}
 Let $I:=[0,T]$ be a time interval with end time point value $T$.
@@
 Roland Becker, Dominik Meidner, and Boris Vexler \cite{rodobo}. 
-The second author thanks Mary F. Wheeler (ICES, Austin) for the 
+Finally, the second author thanks Mary F. Wheeler (ICES, Austin) for the 
 possibility to finish this work.

### 2012-11-28 Paper_May_2012/main.tex
Antworten auf TW. Geklaerte TODOs entfernt.

@@
 \begin{abstract}
-\todo{Wird die Webseite genannt? ist das sinnvoll dies zu tun?}
+\todo{Wird die Webseite genannt? ist das sinnvoll dies zu tun?}\todocg{Seite 2, vorletzter Absatz. Da wird DOpElib referenziert, und da steht dann auch die Webseite.}
 In this article, we describe the software library 
@@
   and quadrature rules}  
+  \todocg{Gegenvorschlag: Specification of discretization issues...}
 Specify what kind of triangulation:
@@
 
-We split the program into the parts \texttt{main.cc}, \texttt{func.h} and \texttt{pde.h}.
-The \texttt{main.cc}-file is responsible for the declaration of various objects which have mostly to do with the \textit{discretization}, the \textit{solvers} and the general \textit{program sequence}, whereas the other two contain most of the problem-data -- especially the cell-wise contributions of the weak form in \texttt{pdh.h} and the functional in \texttt{func.h}.
+We split the program into the parts \texttt{main.cc}, \texttt{localfunctional.h} and \texttt{localpde.h}.
+The \texttt{main.cc}-file is responsible for the declaration of various objects which have mostly to do with the \textit{discretization}, the \textit{solvers} and the general \textit{program sequence}, whereas the other two contain most of the problem-data -- especially the cell-wise contributions of the weak form in \texttt{localpde.h} and the functional in \texttt{localfunctional.h}.
 We give a brief explanation on every file in the following.
@@
 
-\subsubsection{\texttt{pde.h}}
+\subsubsection{\texttt{localpde.h}}
 We shall now give an overview to the implementation of the specific form of the PDE.
@@
 
-\todo{Ich wuerde localfunctional als volles Wort schreiben}
-\subsubsection{\texttt{localfunc.h}}
+\subsubsection{\texttt{localfunctional.h}}
 The implementation of the boundary-functional $J$ follows the outline of the
 computation of the cell-residual in \texttt{localpde.h}. 
-The main difference, and common to all functionals is that we only want to
-evaluate a functional \todocg{Ist hier was verrutscht? der Satz passt im
+%The main difference, and common to all functionals is that we only want to evaluate a functional 
+\todocg{Ist hier was verrutscht? der Satz passt im
   Moment nicht so wirklich, der Teil mit dem Vektor ist irgendwie
@@
 haben wolltest. Im main TW steht noch das Original}
-Thus, we 
-just loop over the quadrature points and not over the degrees of freedom 
-on the face. In the present situation, we want to evaluate a functional 
+\todocg{Ich habe den Rest auskommentiert.
+Mein Vorschlag waere: The main difference is that as we only want to evaluate a functional, i.e.
+a number, and not a residual-vector, we just loop over the quadrature points and not over the degrees of freedom on the face.
+}
+%Thus, we just loop over the quadrature points and not over the degrees of freedom on the face.
+In the present situation, we want to evaluate a functional 
 that consists of a boundary integral. Thus we need to evaluate 
@@
 PDEs, we consider a time-dependent semi-linear form $a(u)(\phi)$ and some
-right hand side functional $l$.\todocg{Warum hier l und nicht f?}
-\todo{Antwort: Weil $f$ die volumen kraft ist. Wir wollten 
-doch Funktional und Kraftvektor unterscheiden. Daher hier $l$. Es gilt :
-$l(\phi) = (f,\phi)$.}
-Let $I:=[0,T]$ be a time interval with end time point value $T$.
+right hand side functional $l$. Let $I:=[0,T]$ be a time interval with end time point value $T$.
 To find some $u$ in a weak form, we define

### 2012-11-28 Paper_May_2012/main.tex
Antworten auf CGs Vorschlaege und am Anfang zwei neue Todos

@@
 Austin, Texas 78712, USA \\
-              Tel.: +1 512-672-7763\\
+              %Tel.: +1 512-672-7763\\
               \email{twick@ices.utexas.edu}   
@@
 \begin{abstract}
-\todo{Wird die Webseite genannt? ist das sinnvoll dies zu tun?}\todocg{Seite 2, vorletzter Absatz. Da wird DOpElib referenziert, und da steht dann auch die Webseite.}
+\todo{Ich finde die Adressen uneinheitlich. CG und ich koennten die
+  Telefonnummern wegstreichen, um uns an Winni anzupassen. Und Winni
+keonnte Hausnummer oder Postleitzahl hinzufuegen. Immerhin ist 
+das Standart in (fast) allen Journals}
 In this article, we describe the software library 
@@
 \label{introduction}
+\todo{Reihenfolge bei der Erwaehnung von Dopelib mit vollem Namen und 
+des Anfangs in 2010 passt nicht ueberein. 2010 wird am Ende erwahent und der
+volle Name von Dope wird erst genannt nachdem wir bereits die Abkuerzung
+benutzt haben. Das geht so nicht!}
 Today, there exists a broad variety of software libraries
@@
   \todocg{Gegenvorschlag: Specification of discretization issues...}
+\todo{Finde ich gut: Koennen wir machen}
 Specify what kind of triangulation:
@@
 haben wolltest. Im main TW steht noch das Original}
-\todocg{Ich habe den Rest auskommentiert.
-Mein Vorschlag waere: The main difference is that as we only want to evaluate a functional, i.e.
-a number, and not a residual-vector, we just loop over the quadrature points and not over the degrees of freedom on the face.
-}
+\todocg{Ich habe den Rest auskommentiert. Mein Vorschlag waere: The main
+  difference is that as we only want to evaluate a functional, i.e. a number,
+  and not a residual-vector, we just loop over the quadrature points and not
+  over the degrees of freedom on the face. }
+\todo{Finde ich gut. Kannst Du so machen!}
 %Thus, we just loop over the quadrature points and not over the degrees of freedom on the face.

### 2012-11-28 Paper_May_2012/main.tex
My Modifications

--- 2012-11-28 Paper_May_2012/main.tex
@@
 \newcommand{\todocg}[1]{\textbf{\textsc{\textcolor{red}{TODO: #1\textasciitilde cg}}}}
+\newcommand{\todoww}[1]{\textbf{\textsc{\textcolor{green}{TODO: #1\textasciitilde ww}}}}
 %\newcommand{\mymarginpar}[1]{\marginpar{\textcolor{red}{\scriptsize{#1}}}}
@@
 
-\title{DOpElib: Differential Equations and Optimization Environment; A Goal Oriented Software Library for Solving PDEs and Optimization Problems}
+\title{DOpElib: Differential Equations and Optimization Environment; A Goal Oriented Software Library for Solving PDEs and Optimization Problems with PDEs}
 
@@
               \\Im Neuenheimer Feld 293/294, 69120 Heidelberg, Germany\\
-              Tel.: +49 622-154-6154\\
+              %Tel.: +49 622-154-6154\\
               \email{cgoll@iwr.uni-heidelberg.de}           %  \\
@@
 Austin, Texas 78712, USA \\
-              Tel.: +1 512-672-7763\\
+              %Tel.: +1 512-672-7763\\
               \email{twick@ices.utexas.edu}   
  \and
-W. Wollner \at
+Winnifried Wollner \at
               Universit\"at Hamburg\\
               Department of Mathematics\\
+              Bundesstr. 55, 20146 Hamburg, Germany\\
               \email{winnifried.wollner@math.uni-hamburg.de}   
@@
 \begin{abstract}
-\todo{Wird die Webseite genannt? ist das sinnvoll dies zu tun?}\todocg{Seite 2, vorletzter Absatz. Da wird DOpElib referenziert, und da steht dann auch die Webseite.}
-In this article, we describe the software library 
+In this article, we describe the  
 \textit{Differential Equations and Optimization Environment} (\dope{}).
-The main feature of \dope{} is that it provides a unified interface to high level algorithms 
-such as time-stepping methods, nonlinear solvers and optimization routines. This structure ensures 
-that first of all the user is required to write only those sections of code that are specific to 
+\dope{} is a software library that provides a unified interface 
+to high level algorithms such as time-stepping methods, nonlinear solvers and 
+optimization routines. This structure ensures 
+that, first of all, the user is only required to write those sections of code 
+that are specific to 
 the considered problem. Second, the exchange of parts of the used routines is possible 
@@
 In many cases, the user is confronted with 
-certain bottlenecks. On the one hand, 
-in open-source projects for academic research, 
+certain bottlenecks when using such software. 
+On the one hand, in open-source projects for academic research, 
 standard high-level routines 
 which are necessary to solve even simplest PDEs,
-such as linear and nonlinear solvers or time-stepping schemes must be implemented 
-by the users themselves. However, such routines are, at large,
+such as linear and nonlinear solvers or time-stepping schemes must 
+be implemented by the users themselves. However, such routines are, at large,
 problem-independent and could be provided as standard tools. 
@@
 As a second motivating principle, 
-there is nowadays an increasing interest in computing 
+there is nowadays an increasing interest in solving
 optimization problems in which the state equation is governed 
@@
 parameter estimation, inverse problems or optimal experimental design. 
-To do this in complex applications usually requires tremendous work already
+To do this in complex applications requires tremendous work already
 on the PDE itself before considering optimization of the processes modeled 
@@
 routines.
-So far, there exist a very limited number of open source packages
-which fulfill the requirements listed above; 
+
+On the other hand, commercial finite element (FE) toolkits provide these 
+previously mentioned routines but limits user-interaction
+with the source code to an extend that typically conflicts
+with the consideration of problems deviating from the intended 
+user-base for the commercial product. 
+This is quite unsatisfactory because in the development of 
+novel numerical algorithms  or in the solution of more complex 
+applications the researcher needs to be enabled 
+to change certain parameters or pieces of source code.
+
+This article will provide information on who the {\em Differential 
+Equations and Optimization Environment} \dope{}  \cite{dope} 
+solves these problems by providing a well structured interface for high-level 
+routines integrated
+with tools to solve PDEs. At the present stage we provide an interface to 
+the finite element library \deal{} and a collection of algorithms to solve
+both pure PDE problems as well as optimization problems 
+constrained by a PDE. 
+This provides an extension to the 
+very few tools available doing so; 
 successful in-house code examples are Gascoigne \cite{gascoigne}
@@
 
-On the other hand, commercial finite element toolkits provide these 
-previously mentioned routines but limits user-interaction
-with the source code to an extend that typically conflicts
-with the consideration of problems deviating from the intended 
-user-base for the commercial product. 
-This is quite unsatisfactory because development of novel numerical 
-algorithms should enable the researcher
-to change certain parameters or pieces of source code.
-
-The presented toolkit \dope{}: 
-\textit{Differential Equations and Optimization Environment} \cite{dope}
-fills this gap. We present a software
-to solve both pure PDE problems as well as optimization problems 
-constrained by a PDE. 
-Its main feature is to give a unified interface to high level algorithms such as 
+The main feature of \dope is to give a unified interface to high level 
+algorithms such as 
 time-stepping methods, nonlinear solvers and optimization routines. 
@@
 
-While \deal{} --which at present is the only toolkit to which we provide an 
+While \deal{} --which at present is the only FE-toolkit to which we provide an 
 interface-- leaves the implementation of all high-level algorithms to the user, 
 \dope{} is user-focused by delivering
-prefabricated tools which require from the user only adjustments connected
-to his specific problem. The solution of optimization problems with PDE
-constraints is an innovative feature in \dope{}.
-
-\todo{Das wir in Heidelberg angefangen haben wird schon vorher erwaehnt!
-  Wuerde den Satz einfach streichen}
-The project was initially initiated \todocg{initially initiated - doppelt gemoppelt?} at the University of Heidelberg in 2010 and  
-at the present stage the following key features are supported by the library
+prefabricated tools which requires adjustments by the user only for parts
+connected to his specific problem. The possible solution of 
+optimization problems with PDE constraints is an innovative feature in \dope{}.
+
+To give a brief overview, at the present stage of the development 
+the following key features are supported by the library
 \begin{itemize}
-\item Solving of nonlinear stationary and nonstationary PDEs in 1d, 2d, and 3d.
+\item Solvers for nonlinear stationary and nonstationary PDEs in 1d, 2d, and 3d.
 \item Various, easily exchangeable, 
@@
 \item Use of all finite elements of from \deal{} including hp-support.
-\item Some examples showing the solution of several PDEs including
+\item Several examples showing the solution of PDEs including
    Poisson, Navier-Stokes, Plasticity and fluid-structure interaction problems. 
@@
 \item Several examples showing how to solve various kinds of optimization 
-  problems involving PDE constraints.
+  problems involving stationary and nonstationary PDE constraints.
 \item Goal-oriented mesh adaptation with the Dual Weighted Residual method and
@@
 \ref{detailed_description}, a detailed description of 
-the main features is given. In fact, we first explain the solution 
-of stationary PDEs, nonstationary PDEs, and optimization problems. 
-Additional comments are given concerning the required implementation 
-provided by the user.
-Finally, additional features are discussed. In the final Section
-\ref{applications}, we discuss three different numerical 
+the main features is given. First, we first explain the solution 
+of stationary PDEs, in Section~\ref{subsubsec:stationary problems}, nonstationary PDEs, in Section~\ref{sec:timedep}, and optimization problems, in 
+Section~\ref{sec:opt}. 
+In all three cases we first described how the interface is designed, and why 
+it is designed as such. Then the implementation required by the user is 
+demonstrated at an example.
+Finally, additional features are discussed in Section~\ref{sec:additional}. 
+In the final Section
+\ref{applications}, we give three different numerical 
 examples which illustrate the versatile features of \dope{}.
@@
 of problems involving PDEs.
-In particular, we remove all tedious routine implementation needed in 
+By providing interfaces to existing FE-toolkits, at present \deal{}, 
+we remove all tedious routine implementation needed in 
 \deal{} by introducing appropriate classes that take care of all the 
 problem independent implementation while leaving only the problem specific 
-parts to the user. To show how this is designed we will, in the sequel, 
-describe the reasoning first for a simple stationary PDE in 
-Section~\ref{subsubsec:stationary problems} and then show
-how this idea extends to non stationary problems in 
-Section~\ref{sec:timedep} as well as to optimization
-problems with PDE constraints in Section~\ref{sec:opt}.
-
-\subsection{Solving of Stationary PDEs}
+parts to the user. To show how this is designed, we will, in the sequel, 
+describe the reasoning, first, for a simple stationary PDE, in 
+Section~\ref{subsubsec:stationary problems}, and then show
+how this idea extends to non stationary problems, in 
+Section~\ref{sec:timedep}, as well as to optimization
+problems with PDE constraints in, Section~\ref{sec:opt}.
+
+\subsection{Solving Stationary PDEs}
 \label{subsubsec:stationary problems}
-For the start we consider a prototypical problem given by the solution of 
+For the start, we consider a prototypical problem given by the solution of 
 a stationary PDE, which we assume to be given in weak form.
@@
 continuation of the given Dirichlet boundary data.
-Then typically we do not only want to calculate $u$ but more likely we are 
+In a typical calculation, we do not only want to calculate $u$ but more likely we are 
 interested in one (or several) functionals depending on the solution, i.e., some numbers $J(u)$.
@@
 we need to discretize 
-the equation. In \dope, we are using finite elements to do this. This means we need to pick a 
+the equation. In \dope, we are using finite elements, as they are 
+provided by \deal{}, to do this. This means we need to pick a 
 subdivision $\mathcal T_h$, for instance into quadrilaterals,
 of the domain $\Omega$ and a finite element. In the concrete case, we could, 
-\todo{Warum Tylor-Hood? Ist inf-sup stabil wuerde ich erwaehnen}
-for instance, take the $\mathcal Q_2/\mathcal Q_1$ Taylor-Hood element.
+for instance, take the inf-sup stable  $\mathcal Q_2/\mathcal Q_1$ 
+Taylor-Hood element.
 As it is clear, that the initialization of the unknowns, the local shape 
@@
 handling the finite element space $V_h$. We will refer, to 
-this as the \texttt{SpaceTimeHandler}. Then, it is clear that the simplest 
+this object as the \texttt{SpaceTimeHandler} since it is also 
+capable of coping with unknowns for time dependent problems. Then, it is 
+clear that the simplest 
 form of the constructor should have the following interface:
@@
 \end{align}
-\todo{wir schreiben newton immer klein. eigentlich ist das ein Eigenname: Newton}
-To this end, one can employ a newton type method with some globalization 
-like line-search. To do so, one needs to calculate the residual of the 
-equation, as well as to solve linearized equations for the calculation of 
-the newton method and then employing some kind of globalization strategy.
-This again will usually require the evaluation of the residual.
+To this end, one can employ a Newton type method with some globalization 
+like line-search, see, e.g.,~\cite{NoWr00}. 
+To apply the Newton method, one needs to calculate the residual of the 
+equation, solve linearized equations for the calculation of 
+the update, and then employ some kind of globalization strategy.
+Again, this will usually require the evaluation of the residual.
 
@@
 \begin{align}\label{eq:residual_vector}
- a(u_h)(\phi_{h}^{i}) - (f,\phi_{h}^{i}) = \sum_{K\in \mathcal T_h} \nu\int_K \nabla u_h\cdot \nabla \phi_{h}^{i}\,dx + \ldots
+ a(u_h)(\phi_{h}^{i}) - (f,\phi_{h}^{i}) = \sum_{K\in \mathcal T_h} \int_K \nu\nabla u_h\cdot \nabla \phi_{h}^{i}\,dx + \ldots
  \end{align}
-Again, we see that, except for the local element integrals, all 
+Again, we see that, except for the local cell (element) integrals, all 
 calculations are independent from the concrete form of the PDE. 
@@
 that offers a method to calculate the residual given a user-defined 
-object \texttt{pde} that implements the local element integrals, i.e.,
+object \texttt{pde} that implements the local cell integrals, i.e.,
 \begin{lstlisting}
@@
 \end{lstlisting}
-where \texttt{VECTOR} is one of the possible vector types given by \deal{}.
+where \texttt{VECTOR} is some vector type in which the residual 
+should be stored.
 It is clear that we need to ensure that the integrator object has access
 to the information stored in the \texttt{SpaceTimeHandler}
-to correctly loop over all elements.
+to correctly loop over all cells.
 To solve the linearized systems, again, all that is left is the calculation
@@
 \end{lstlisting}
-where now \texttt{MATRIX} is one of the matrix types supported by \deal{}.
+where now \texttt{MATRIX} is an appropriate matrix type needed by the 
+linear solver.
 With this we can now define a nonlinear solver \texttt{NewtonSolver} 
@@
 Where \texttt{LINSOLVE} denotes the solver that should be used to solve the 
-linearized equations during the newton iteration.
+linearized equations during the Newton iteration.
 
@@
 Since the sequence of events sketched above is again independent of the 
-exact problem we introduce another container around it which we call 
+exact problem we introduce another container which we call 
 \texttt{StatPDEProblem} that then simply provides the user with the 
@@
 \end{lstlisting}
+that takes care of calling all other methods needed for the calculation of 
+$J(u_h)$.
 The template argument \texttt{NEWTON} represents the chosen solver of the nonlinear problems.
@@
 given to the \texttt{SpaceTimeHandler}. However, to keep the 
-presentation clear we refrained from including these details above and in
+presentation clear, and since default values are available for these arguments, 
+we refrained from including these details above and in
 what follows.
+\end{remark}
+
+\begin{remark}
+The reason to include the nonlinear solver as a template argument in 
+\texttt{StatPDEProblem} is that if a given problem can not be solved with 
+this algorithm, e.g., it has no derivatives, then the user should be able 
+to exchange only the nonlinear solver. Further, this makes it possible to
+compare different algorithms available for the solution of the PDE without 
+the need to recode other parts of the problem, thereby helping in doing fair 
+comparisons.
 \end{remark}
@@
 objects described above.
-\subsubsection{Problem dependent implementation of stationary PDE problems}
+\subsubsection{Problem Dependent Implementation of Stationary PDE Problems}
 To solve a stationary PDE problem, the user has to provide some code which can roughly be divided into the following main sections:
 \begin{description}
-\item[\textit{Discretization}]\todocg{Hier scheint etwas verlorengegangen zu
-    sein, der Satz macht so keinen Sinn:} 
-\todo{Vorschlag: Specification of triangulation issues such as finite elements
-  and quadrature rules}  
-  \todocg{Gegenvorschlag: Specification of discretization issues...}
-Specify what kind of triangulation:
-in most cases this is the
-finite element and quadrature rules. Furthermore, other
-things might be adjusted as well.
-\item[\textit{Problem-data}] \todo{Definition of } Define the weak form, the corresponding matrices, the
-  functionals $J$ as well as the boundary conditions. With exception of the boundary conditions, we have to specify all this only cell-wise, \dope{} handles the rest for us.
+\item[\textit{Discretization}] 
+Specification of discretization issues such as finite elements, 
+quadrature rules, and triangulation.
 \item[\textit{Solvers}] Choice of a (non-)linear solver. 
-\item[\textit{Additional features}] ParameterReader to read in the run-time parameters
+\item[\textit{Problem-data}] Definition of the weak form, 
+the corresponding matrices, the functionals $J$, as well as the boundary 
+conditions. 
+\item[\textit{Additional features}] ParameterReader to read in the 
+  run-time parameters
   such as material coefficients or adjustments for the solvers like tolerance
@@
 \end{description} 
-
-We describe now a program designed to compute a value $J(u)$, where $u=(v,p)$ is the solution of the Navier-Stokes equations \eqref{eq:ns} on the domain $\Omega$ as seen in figure~\ref{fig:example_ns} with  $f=0$, and $J$ is the drag-force of the enclosed circle.
-We assume given Dirichlet data for the velocity $v$ on the boundary $\Gamma_{D}=\Gamma_{in}\cup\Gamma_{wall}\cup\Gamma_{circ}$.
-This is the Benchmark-Case 2D-1 from \cite{TuSchae96}, see there for further details.
+The first two points will be done below in the discussion of the file 
+\texttt{main.cc}. The third point is done separately in the file \texttt{pde.h}
+and \texttt{func.h} for convenience. The specification of additional 
+features is skipped, since it consist of reading in a text-file only.
+
+Now, we describe a program designed to compute a value 
+$J(u)$, where $u=(v,p)$ is the solution of the Navier-Stokes 
+equations \eqref{eq:ns} on the domain $\Omega$ as seen in 
+Fig.~\ref{fig:example_ns} with  $f=0$, and $J$ is the drag-force 
+of the enclosed circle. We assume given Dirichlet data for the velocity 
+$v$ on the boundary $\Gamma_{D}=\Gamma_{in}\cup\Gamma_{wall}\cup\Gamma_{circ}$. 
+This is the Benchmark-Case 2D-1 described in \cite{TuSchae96}.
 \begin{figure}[h]
@@
 
-We split the program into the parts \texttt{main.cc}, \texttt{localfunctional.h} and \texttt{localpde.h}.
-The \texttt{main.cc}-file is responsible for the declaration of various objects which have mostly to do with the \textit{discretization}, the \textit{solvers} and the general \textit{program sequence}, whereas the other two contain most of the problem-data -- especially the cell-wise contributions of the weak form in \texttt{localpde.h} and the functional in \texttt{localfunctional.h}.
+We split the program into the parts \texttt{main.cc}, \texttt{func.h} and 
+\texttt{pde.h}. The \texttt{main.cc}-file is responsible for the declaration 
+of various objects which have mostly to do with the \textit{discretization}, 
+the \textit{solvers} and the general \textit{program sequence}, whereas the 
+other two contain most of the problem-data -- especially the cell-wise 
+contributions of the weak form in \texttt{pde.h} and the functional in 
+\texttt{func.h}. 
 We give a brief explanation on every file in the following.
@@
 \subsubsection{\texttt{main.cc}}
-In this file we gather all the basic components that are necessary for the
+In this file, we gather all the basic components that are necessary for the
 solution process of a (stationary) partial differential equation. The
@@
 some pre-implemented routines from the \dope{} library rather than 
-implementing new features.
-
-After the usual preamble, in which we include the necessary header-files from \deal{}, \dope{} and our own problem (i.e.
-\texttt{localfunctional.h} and \texttt{localpde.h}), we start by defining an instance of a \texttt{ParameterReader}-object, which handles run-time parameters.
+implementing these ourselves.
+
+After the usual preamble, in which we include the necessary header-files 
+from \deal{}, \dope{} and our own problem-data (i.e. \texttt{func.h} 
+and \texttt{pde.h}), we start by defining an instance of a 
+\texttt{ParameterReader}-object, which handles reading of run-time parameters.
 \begin{lstlisting}
@@
 \end{lstlisting}
-In a text-file we might specify some problem-related parameters (for instance the viscosity $\nu$), but also general parameters for (non)-linear solvers and the output of computed data. 
-After that, we create a triangulation of the domain $\Omega$ named \texttt{triangulation}.
+The object \texttt{pr} then reads a text-file in which we might specify some 
+problem-related parameters (for instance the viscosity $\nu$), but also general
+ parameters for (non)-linear solvers and the output-format of computed data. 
+After that, we create a triangulation of the domain $\Omega$ named 
+\texttt{triangulation} using the tools provided by the \deal{} FE-library.
 
 Next, we create some quadrature rules (for the evaluation of integrals over
-cells and faces) and a finite element appropriate for the PDE at hand. We
+cells \texttt{quad} and faces \texttt{face\_quad}) and a finite element \text{fe}
+appropriate for the PDE at hand. We
 choose a Gaussian quadrature of a certain degree and the previously mentioned
 Taylor-Hood element. These routines are quite standard in each finite element
-software and we take advantage to take them from \deal{}.
+software and thus we can use those provided by the FE-library \deal{}.
 
@@
 \begin{lstlisting}
-  LocalPDE local_pde;
-  LocalFunctional local_functional;
+  LocalPDE pde;
+  LocalFunctional functional;
 \end{lstlisting}
 These hold the information of the PDE and the functional under consideration. Further explication is given below because
-the definition of this classes is done separately in \texttt{localpde.h} resp.   \texttt{localfunctional.h}.
+the definition of this classes is done separately in \texttt{pde.h} 
+and \texttt{func.h} respectively.
 
 In the next step, we define an object of type \texttt{SpaceTimeHandler}, which handles all the things connected with the triangulation as well as the finite element space. 
-This class is also capable of organizing degrees of freedom in time-dependent problems, hence the name.
 \begin{lstlisting}
   SpaceTimeHandler dof_handler(triangulation,
-                               finite_element);
-\end{lstlisting}
-There exist different versions of this class for forward as well as optimization problems.
+                               fe);
+\end{lstlisting}
 
@@
 This basically represents the discretized problem, and bundles things like boundary conditions, the PDE, the functional and the discretization.
-First, we give the constructor the \texttt{local\_pde} and the \texttt{dof\_handler}.
+First, we give the constructor the \texttt{pde} and the \texttt{dof\_handler}.
 \begin{lstlisting}
   PDEProblemContainer<LocalPDE> 
-             prob_container(local_pde, dof_handler);                                            
-\end{lstlisting}
-After that, we add the functional under consideration and, because it is a boundary functional, we specify the part of the boundary on which the functional operates (this is done via a 'boundary color', $1$ represents $\Gamma_{circ}$).
-\begin{lstlisting}
-  prob_container.AddFunctional(&local_functional);
+             prob_container(pde, dof_handler);                                            
+\end{lstlisting}
+After that, we add the functional under consideration and, because it is a 
+functional acting on the boundary, specify the part of the boundary on which 
+the functional operates (this is done via a 'boundary color', where $1$ 
+represents $\Gamma_{circ}$).
+\begin{lstlisting}
+  prob_container.AddFunctional(&functional);
   prob_container.SetBoundaryFunctionalColors(1);
@@
 \end{lstlisting}
-Now we create the object which steers the whole solution process, the previously mentioned \texttt{StatPDEProblem}.
+Now, we create the object which steers the whole solution process, the previously mentioned \texttt{StatPDEProblem}.
 It takes care of solving the discretized equation and evaluating the given functionals.
@@
 \end{lstlisting}
-With the type settled, we create the object \texttt{solver}, which needs the discretization (represented here by the \texttt{prob\_container}), the parameter reader as well as the quadrature rules.
+With the types settled, we create the object \texttt{solver}, which needs the discretization (represented here by the \texttt{prob\_container}), the parameter reader as well as the quadrature rules.
 \begin{lstlisting}
   PDE pde(prob_container, pr,
-                    quad_formula, face_quad_formula);
+                    quad, face_quad);
 \end{lstlisting}
@@
 \end{lstlisting}
-
-\subsubsection{\texttt{localpde.h}}
+\todoww{Should we add a complete listing?}
+\subsubsection{\texttt{pde.h}}
 We shall now give an overview to the implementation of the specific form of the PDE.
@@
 The corresponding Jacobian can be defined analogously.
-Now all we have to supply here are the cell-wise contributions to these terms, the \texttt{Integrator} takes care of assembling \eqref{eq:discrete_residual}
+Now, all we have to supply here are the cell-wise contributions to these terms,
+then the \texttt{Integrator} object 
+takes care of assembling \eqref{eq:discrete_residual}
 and similarly of the matrix for the linearized equation to be solved.
@@
 implementation itself) of these
-terms works by and large as in \deal{}.
+terms works as it is required by the FE-library for which the integration is done, here this is \deal{}.
 
 We show the basic structure of these methods using the \texttt{CellEquation} 
-method for the implementation of the element-wise integrals.
+method for the implementation of the cell-wise integrals.
 
@@
   void CellEquation(const CellDataContainer &cdc,
-                    VECTOR &local_cell_vector)
+                    VECTOR &local_vector)
   {
@@
  The \texttt{CellDataContainer} contains all the information we need to compute the cell-contributions (things like number of quadrature points, number of degrees of freedom on the cell, the \texttt{dealii::FEValues}-objects, etc.).
-  \begin{lstlisting}
-    cdc.GetValuesState("last_newton_solution", uvalues);
-    cdc.GetGradsState("last_newton_solution", ugrads);
- \end{lstlisting}
- In these two lines, we load the (gradient of the) solution of the last newton iteration in all the quadrature points on the cell into the two vectors \texttt{uvalues} and \texttt{ugrads}.
-
-In the next lines, we finally implement our concrete PDE. Here, we aim to provide 
-an interface which is closely related to the weak formulation which 
-has to be derived by the user. Thus,
+To get access to the values and the gradient of the solution in the quadrature 
+points we need to extract them from the \texttt{CellDataContainer}:
+\todoww{Ich habe jetzt das Arg. in state ge\"andert, das wollten wir sowieso in 
+\dope{} umsetzen und ist dann nicht so umst\"andlich}
+  \begin{lstlisting}
+    cdc.GetValuesState("state", uvalues);
+    cdc.GetGradsState("state", ugrads);
+ \end{lstlisting}
+ In these two lines, we load the values and the gradient of the solution 
+during the last Newton iteration in all the quadrature points on the cell 
+into the two variables \texttt{uvalues} and \texttt{ugrads}.
+
+In the next lines, we finally implement our concrete 
+cell contribution of the PDE. Thus,
 we loop over the quadrature points and the number of DoFs of the cell.
-In the innermost loop, we evaluate the weak form tested with $\phi_{h}^i$ (represented by \texttt{phi\_i\_grads\_v}) on the cell with the help of the previously selected quadrature formula.
+In the innermost loop, we evaluate the weak form tested with $\phi_{h}^i$ (here \texttt{phi\_i\_grads\_v} denotes the matrix $\nabla \phi_h^i$)
+on the cell with the help of the previously selected quadrature formula. 
  \begin{lstlisting}
@@
       for (unsigned int i=0;i<dofs_per_cell;i++)
-        local_cell_vector(i) += 
+        local_vector(i) += 
            ( _nu * vgrads * phi_i_grads_v 
@@
 \end{lstlisting}
-\dope{} takes now care of the distribution of the local cell contributions into  the vector of the residual \eqref{eq:residual_vector}.
+\dope{} takes care of the distribution of the local cell contributions into  the vector of the residual \eqref{eq:residual_vector}.
 The methods \texttt{CellRightHandSide} and \texttt{CellMatrix} follow the same structure as
@@
 
-\subsubsection{\texttt{localfunctional.h}}
+\subsubsection{\texttt{func.h}}
 The implementation of the boundary-functional $J$ follows the outline of the
-computation of the cell-residual in \texttt{localpde.h}. 
-%The main difference, and common to all functionals is that we only want to evaluate a functional 
-\todocg{Ist hier was verrutscht? der Satz passt im
-  Moment nicht so wirklich, der Teil mit dem Vektor ist irgendwie
-  verlorengegangen?}. 
-\todo{Ich hatte Deine Anmerkung so verstanden, als ob Du alles nach dem i.e.,
-  geloescht 
-haben wolltest. Im main TW steht noch das Original}
-\todocg{Ich habe den Rest auskommentiert.
-Mein Vorschlag waere: The main difference is that as we only want to evaluate a functional, i.e.
-a number, and not a residual-vector, we just loop over the quadrature points and not over the degrees of freedom on the face.
-}
-%Thus, we just loop over the quadrature points and not over the degrees of freedom on the face.
+computation of the cell-residual in \texttt{pde.h}. 
+The main difference, and common to all functionals, is that as we only want 
+to evaluate a functional, i.e. a number, and not a residual-vector. 
+Hence, we loop over the quadrature points and not over the local 
+degrees of freedom.
+
 In the present situation, we want to evaluate a functional 
@@
 integrals on faces, those who lie on $\Gamma_{circ}$, instead of integrals over
-elements. 
+cells. 
 This face contributions are calculated in \texttt{BoundaryValue} and added to 
@@
     //same as before, get n_q_points etc.
-    const auto& face_values = cdc.GetFEFaceValuesState();
+    const auto& face_values = fdc.GetFEFaceValuesState();
     fdc.GetFaceValuesState("state", ufacevalues);
     fdc.GetFaceGradsState("state", ufacegrads);
-\end{lstlisting}
-Notice that we get here the values of the computed solution by giving the string "state".
-\begin{lstlisting}
+
+    double erg = 0.;
     for(unsigned int q_point = 0; 
@@
 right hand side functional $l$. Let $I:=[0,T]$ be a time interval with end time point value $T$.
-To find some $u$ in a weak form, we define
+The weak for of such a problem then consists of finding some $u$ such that
 \[
-\int_I a(u)(\phi)\, dt = \int_I l(\phi)\, dt \quad \forall\Phi\in X,
+\int_I \overline{a}(u)(\phi)\, dt = \int_I l(\phi)\, dt \quad \forall\Phi\in X,
 \]
@@
 we extend the previous example and consider the nonstationary Navier-Stokes equations
-in a domain $\Omega\in\mathbb{R}^d$. Find: 
-$(v,p)\in X$ with
+in a spatial domain $\Omega\in\mathbb{R}^d$. The weak for is then to find 
+$u = (v,p)\in X$ with
 \begin{align*}
@@
 &\quad + (v(0) - v^0, \phi(0))
-= \int_I l(\phi) \, dt 
+= \int_I (f,\phi) \, dt 
 \end{align*}
@@
 \partial_t v\in L^2(I, L^2(\Omega)^d), 
-p\in L^2(I,L^2(\Omega)\setminus\mathbb{R}}.
+p\in L^2(I,L^2(\Omega)\setminus\mathbb{R})}.
 \]
+We note that the only difference is the appearance of an additional
+time integration, the initial-value, and the term 
+\[
+\int_I (\partial_t v, \phi)\,dt
+\]
+so that we can reuse our stationary implementation for all the other 
+terms if we rescale them appropriately in the time integration.
+
 As for stationary problems, the goal is to compute
@@
 of nonstationary problems is quite similar to stationary problems,
+at least from the user implementation point of view,
 which was one major goal we claimed at the beginning.
-Next, we will describe the methods (i.e., changes) the user needs to 
+Next, we will describe the methods the user needs to 
 provide in addition to the stationary case.
 
-\subsubsection{Problem-dependent 
-implementation of nonstationary problems}
+\subsubsection{Problem-Dependent 
+Implementation of Nonstationary Problems}
 \label{sec:timedep:implementation}
@@
 and specific issues of its implementation. 
-To keep the presentation easy, we abuse notation 
-by neglecting
+To keep the presentation easy, we neglect
 all material coefficients
-and give formal-related idea based on the weak 
+to give the idea based on the weak 
 form of the Navier-Stokes equations:
-Find $(v,p)$ such that
+Find $(v,p)$ such that for all $(\phi,\chi)$ 
+and all $t \in I$
 \begin{align*}
@@
 &- k (p^{n+1},\nabla\cdot \phi)
-+ (\nabla\cdot v^{n+1}, \chi)\\
++ k(\nabla\cdot v^{n+1}, \chi)\\
 =&\; k\theta (f^{n+1},\phi) + k(1-\theta) (f^{n},\phi)
@@
 \texttt{scale} and \texttt{scale\_ico}
-are used to distinguish these terms:
+are used to distinguish these terms.
+I the present context this means that 
+\texttt{scale} $= \pm k\theta$ or \texttt{scale} $= \pm k(1-\theta)$ and
+\texttt{scale\_ico}~$= \pm k$.
+
+With this the \texttt{CellEquation} needs to contain the terms weighted
+with the appropriate scaling as it is indicated below but is otherwise the 
+same as it is in the stationary case. 
 \begin{lstlisting}
@@
 Note that in the stationary case there is no such difference and accordingly 
-\dope{} automatically sets $ \texttt{scale\_ico} = \texttt{scale}$, so that in fact the user has to implement the \texttt{CellEquation} only once and can henceforth reuse it in all his stationary and nonstationary problems.
+\dope{} automatically sets $ \texttt{scale\_ico} = \texttt{scale}$, so that in fact the user has to implement the \texttt{CellEquation} only once and can henceforth reuse it in all his stationary and nonstationary problems if
+\texttt{scale} and \texttt{scale\_ico} are properly used. 
 This transfers of course also to other methods like \texttt{CellMatrix}.
@@
 However, if someone 
-needs explicit input of the old time-step term, then he can use 
+needs to give an explicit formula for the difference quotient
+this can be written using the methods
 \begin{lstlisting}
@@
 \end{lstlisting}
+Such explicit statements may be useful for nonlinear terms to which time 
+derivatives are applied as it appears for instance in fluid-structure 
+interaction problems.
 \end{remark}
@@
 each time-step is subdivided into three 
-sub-steps. 
+sub-steps. However, no additional implementation is needed by the user 
+for the application of the Fractional-Step-$\theta$-scheme.
 \end{remark}
 
-\subsection{Solving of optimization problems}\label{sec:opt}
+\subsection{Solving Optimization Problems}\label{sec:opt}
 Once we are able to solve stationary and nonstationary PDEs
@@
 \end{align*}
-This leads to the problem
+Thus replacing $u$ by $S(q)$ one is left with the problem
 \begin{align*}
@@
 If one would like to use derivative based methods, one 
-needs in addition derivatives of $j_h$ and thus of $J$ as well as of $S_h$. 
-To do so, we remark, that one can calculate the derivative of 
-$j$ using the adjoint calculus. Then the directional derivative 
+needs, in addition, derivatives of $j_h$ and thus of $J$ as well as of $S_h$. 
+We remark, that one can calculate the derivative of 
+$j$ using adjoint calculus. Then the directional derivative 
 of $j$ in direction $\delta q$ is given by
@@
 
-To do so, we define an object 
+We define an object 
 \begin{lstlisting} 
@@
 
-This means for the user only the derivatives of the functional and the PDE 
-needs to be provided in addition to the  \todocg{Verstehe diesen Satz nicht. Das sind doch die Ableitungen des Funktionals und der PDE, warum addition?}
+This means, for solving optimization problems with PDEs, the user needs to
+provide the derivatives of the functional and the PDE 
+in addition to the already defined functions \texttt{CellEquation}, 
+\texttt{BoundaryValue}. For the above adjoint equation these are 
 \begin{lstlisting} 
@@
      // implement directional derivative 
-     // of the CellEquation
+     // of the CellEquation w.r.t. u
    }
@@
      // implement here the derivative of 
-     // the boundary cost functional
+     // the boundary cost functional w.r.t. u
    }
@@
 
-\subsection{Additional features}
+\subsection{Additional Features}\label{sec:additional}
 Finally, we shall give a brief summary of features 
@@
 \dope{} provides goal-oriented mesh refinement with the help of the 
-Dual Weighted Residual (DWR) method, see \cite{BeRa96} or \cite{BR03}. Here, a certain 
-cost functional such as a point evaluation, line integration or domain
-integration is considered. Determining the lowest error between the 
-(unknown) exact solution and its numerical approximation for lowest computational
-cost is the main goal of the DWR method. The same routines 
-can also be used, without any modification, to use standard residual error estimators by exchanging the 
+Dual Weighted Residual (DWR) method, see \cite{BeRa96} or \cite{BR03}. Here, a 
+given cost functional such as a point evaluation, line integration or domain
+integration is considered as target quantity. 
+The same routines 
+can also be used, without any modification by the user, to use standard 
+residual error estimators by exchanging the 
 error estimator class.
@@
 
-\subsection{Goal-oriented mesh refinement for Navier-Stokes}
-In this example we consider a laminar flow around a cylinder in 2d.
+\subsection{Goal-Oriented Mesh Refinement for Navier-Stokes}
+In this example, we consider a laminar flow around a cylinder in 2d.
 To be more precise, we are computing the benchmark 2D-1 from \cite{TuSchae96}.
@@
 The method delivers error indicators and by this way computational grids tailored to minimize the error in the previously given functional $J$.
-The price we have to pay is the additional computation of a (linear) dual problem, which is implemented as described in section~\ref{sec:opt}.
+The price we have to pay is the additional computation of a (linear) dual problem, which is implemented as described in Section~\ref{sec:opt}.
 In Fig.~\ref{fig:NS_local_grid} we see a resulting triangulation after eight refinement steps with roughly $320\,000$ degrees of freedom.
@@
 \newpage
-\subsection{Compliance minimization}
+\subsection{Compliance Minimization}
 In this example, we consider the compliance minimization of a standard MBB-Beam (Messerschmidt-B\"olkow-Blohm-Beam), see, e.g.,~\cite{BendsoeSigmund:2003}.

### 2012-11-29 Paper_May_2012/main.tex
Kleinere Aenderungen und ein bis zwei Anmerkungen.

@@
 \institute{Christian Goll \at
-              Institute of Applied Mathematics, University of Heidelberg
+              University of Heidelberg, Institute of Applied Mathematics, 
               \\Im Neuenheimer Feld 293/294, 69120 Heidelberg, Germany\\
               %Tel.: +49 622-154-6154\\
-              \email{cgoll@iwr.uni-heidelberg.de}           %  \\
+              \email{christian.goll@iwr.uni-heidelberg.de}           %  \\
            \and
@@
 
-This article will provide information on who the {\em Differential 
+This article will provide information on how the {\em Differential 
 Equations and Optimization Environment} \dope{}  \cite{dope} 
@@
 
-The main feature of \dope is to give a unified interface to high level 
+The main feature of \dope{} is to give a unified interface to high level 
 algorithms such as 
@@
 \dope{} is user-focused by delivering
-prefabricated tools which requires adjustments by the user only for parts
+prefabricated tools which requires\todocg{require? Ist plural} adjustments by the user only for parts
 connected to his specific problem. The possible solution of 
@@
 \ref{detailed_description}, a detailed description of 
-the main features is given. First, we first explain the solution 
+the main features is given. First, we explain the solution 
 of stationary PDEs, in Section~\ref{subsubsec:stationary problems}, nonstationary PDEs, in Section~\ref{sec:timedep}, and optimization problems, in 
@@
 Next, we create some quadrature rules (for the evaluation of integrals over
-cells \texttt{quad} and faces \texttt{face\_quad}) and a finite element \text{fe}
+cells \texttt{quad} and faces \texttt{face\_quad}) and a finite element \texttt{fe}
 appropriate for the PDE at hand. We
@@
 The main difference, and common to all functionals, is that as we only want 
-to evaluate a functional, i.e. a number, and not a residual-vector. 
-Hence, we loop over the quadrature points and not over the local 
+to evaluate a functional, i.e. a number, and not a residual-vector, we loop over the quadrature points and not over the local 
 degrees of freedom.
@@
 PDEs, we consider a time-dependent semi-linear form $a(u)(\phi)$ and some
-right hand side functional $l$. Let $I:=[0,T]$ be a time interval with end time point value $T$.
-The weak for of such a problem then consists of finding some $u$ such that
+right hand side functional $l$\todocg{Nochmal: Warum wird $l$ eingefuehrt, wenn unten doch wieder die l2-represenation $f$ verwandt wird?}.
+Let $I:=[0,T]$ be a time interval with end time point value $T$.
+The weak \todocg{formulation?} of such a problem then consists of finding some $u$ such that
 \[
@@
 we extend the previous example and consider the nonstationary Navier-Stokes equations
-in a spatial domain $\Omega\in\mathbb{R}^d$. The weak for is then to find 
+in a spatial domain $\Omega\in\mathbb{R}^d$. The weak \todocg{formulation?} is then to find 
 $u = (v,p)\in X$ with

### 2012-11-29 Paper_May_2012/main.tex
Corrected CGs comments, and fixed pagebreaks

@@
 \dope{} is user-focused by delivering
-prefabricated tools which requires\todocg{require? Ist plural} adjustments by the user only for parts
+prefabricated tools which require adjustments by the user only for parts
 connected to his specific problem. The possible solution of 
@@
 \end{lstlisting}
-\todoww{Should we add a complete listing?}
+
 \subsubsection{\texttt{pde.h}}
@@
 
-Let $N$ be the number of degrees of freedom of our discretization and $\Set{\phi_{h}^{i}|1\leq i \leq N}$ be a given FE-Basis.
-As stated in subsection~\ref{subsubsec:stationary problems}, to solve the discrete equation \eqref{eq:discrete_equation} we need the residual, i.e.
-the vector
+Let $N$ be the number of degrees of freedom (DoFs) of our discretization and $\Set{\phi_{h}^{i}|1\leq i \leq N}$ be a given FE-Basis.
+As stated in subsection~\ref{subsubsec:stationary problems}, to solve the discrete equation \eqref{eq:discrete_equation} we need the residual, i.e., the vector
 \begin{align}\label{eq:discrete_residual}
@@
 points we need to extract them from the \texttt{CellDataContainer}:
-\todoww{Ich habe jetzt das Arg. in state ge\"andert, das wollten wir sowieso in 
-\dope{} umsetzen und ist dann nicht so umst\"andlich}
   \begin{lstlisting}
@@
 PDEs, we consider a time-dependent semi-linear form $a(u)(\phi)$ and some
-right hand side functional $l$\todocg{Nochmal: Warum wird $l$ eingefuehrt, wenn unten doch wieder die l2-represenation $f$ verwandt wird?}.
-Let $I:=[0,T]$ be a time interval with end time point value $T$.
-The weak \todocg{formulation?} of such a problem then consists of finding some $u$ such that
+right hand side $f$. Let $I:=[0,T]$ be a time interval with end time point value $T$.
+The weak formulation of such a problem then consists of finding some $u$ such
+ that
 \[
-\int_I \overline{a}(u)(\phi)\, dt = \int_I l(\phi)\, dt \quad \forall\Phi\in X,
+\int_I \overline{a}(u)(\phi)\, dt = \int_I (f,\phi)\, dt \quad \forall\Phi\in X,
 \]
-with some suitable Bochner space $X$.  As model problem,
+with some suitable Bochner space $X$.  
+As model problem,
 we extend the previous example and consider the nonstationary Navier-Stokes equations
-in a spatial domain $\Omega\in\mathbb{R}^d$. The weak \todocg{formulation?} is then to find 
+in a spatial domain $\Omega\in\mathbb{R}^d$. The weak formulation is then to find 
 $u = (v,p)\in X$ with
@@
 \centering
-{\includegraphics[width=0.9\textwidth]{Pictures/local_grid_NS.png}}
+{\includegraphics[width=0.85\textwidth]{Pictures/local_grid_NS.png}}
 \caption{Locally adapted grid after eight refinement steps.}
@@
 \centering
-\resizebox{0.5\textwidth}{!}{\input{Pictures/ConvergenceNS}}
+\resizebox{0.45\textwidth}{!}{\input{Pictures/ConvergenceNS}}
 \caption{Comparison of global and local mesh refinement with respect to the error in the drag-functional $J$.}
@@
 fluid.
-
-Since the problem is fully nonstationary, the 
-dynamics are presents in Fig.~\ref{res:fsi_2_mesh_and_x_velo}. 
-By comparison of our findings in Fig.~\ref{res:results_ux_and_uy_fsi_2}
-with the literature, it can be identified that 
-the implementation in \dope{} is correct.
-
 % Bilder von FSI 2 
@@
 \centering
-{\includegraphics[width=5cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_mesh8070_scale.png}}
-{\includegraphics[width=5cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_x_velo8070_scale.png}}
+{\includegraphics[width=4.5cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_mesh8070_scale.png}}
+{\includegraphics[width=4.5cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_x_velo8070_scale.png}}
 \caption{FSI 2 test case: mesh (left) and velocity profile in vertical 
@@
 
-\begin{figure}
+Since the problem is fully nonstationary, only a snapshot of the solution
+and the mesh are presented in Fig.~\ref{res:fsi_2_mesh_and_x_velo}. 
+By comparison of our findings in Fig.~\ref{res:results_ux_and_uy_fsi_2}
+with the literature, it can be identified that 
+the implementation in \dope{} is correct.
+
+\begin{figure}[H]
 \centering
-{\includegraphics[width=5.8cm]{Pictures/ux_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
-{\includegraphics[width=5.8cm]{Pictures/uy_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
-{\includegraphics[width=5.8cm]{Pictures/Drag_fluid_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
-{\includegraphics[width=5.8cm]{Pictures/Lift_fluid_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
-\caption{FSI 2: the deflections of the beam, $u_x(A)$ and $u_y(A)$ (in $cm$), and 
+{\includegraphics[width=4.5cm]{Pictures/ux_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
+{\includegraphics[width=4.5cm]{Pictures/uy_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
+{\includegraphics[width=4.5cm]{Pictures/Drag_fluid_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
+{\includegraphics[width=4.5cm]{Pictures/Lift_fluid_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
+\caption{FSI 2: The deflections of the beam, $u_x(A)$ and $u_y(A)$ (in $cm$), and 
 the drag $F_D$ and the lift $F_L$ evaluation (in $kg/m\,s^2$) are displayed versus
@@
 \centering
-\includegraphics[width=1.\textwidth, viewport=205 88 975 500, clip]{Pictures/MBB.png}
+\includegraphics[width=0.75\textwidth, viewport=205 88 975 500, clip]{Pictures/MBB.png}
 \caption{Thickness distribution for the MBB-beam minimum compliance problem.} 
@@
 the authors of the libraries 
-Gascoigne/RoDoBo project, which was initiated by 
+Gascoigne, initiated by Malte Braack and  Roland Becker \cite{gascoigne}
+and RoDoBo, which where initiated by 
 Roland Becker, Dominik Meidner, and Boris Vexler \cite{rodobo}. 

### 2012-12-04 Paper_May_2012/main.tex
Some more modifications to the paper.

@@
 
+To obtain thickness values that are either $1$ or $0$, so that we can
+interpreter the results in terms of material or no material, we apply the 
+so called SIMP-model (Solid Isotropic Material with Penalization) 
+with parameter $p = 4$ 
+to obtain the distribution given in Fig.~\ref{res:mbb-top}.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.75\textwidth, viewport=205 88 975 500, clip]{Pictures/MBB-top.png}
+\caption{Thickness distribution for the MBB-beam minimum compliance topology optimization problem.} 
+\label{res:mbb-top}
+\end{figure}
+
 \section{Conclusions}

### 2012-12-05 Paper_May_2012/main.tex

--- 2012-11-29 Paper_May_2012/main.tex
@@
 numerical results as demonstration for the versatility of the software.
+
+\keywords{software environment \and goal oriented interface design \and numerical solution of optimization problems and partial differential equations}
+\subclass{49-04 \and  35-04}
 \end{abstract}
@@
 \subsection{Compliance Minimization}
-In this example, we consider the compliance minimization of a standard MBB-Beam (Messerschmidt-B\"olkow-Blohm-Beam), see, e.g.,~\cite{BendsoeSigmund:2003}.
-The thickness variable is allowed to take intermediate value apart from $0$ and $1$,
+In this example, we consider the compliance minimization of a standard MBB-Beam 
+(Messerschmidt-B\"olkow-Blohm-Beam), see, e.g.,~\cite{BendsoeSigmund:2003}. 
+In order to enforce thickness values to be either $1$ or $0$ we employ 
+the, so called, SIMP-model (Solid Isotropic Material with Penalization), 
+see as well,~\cite{BendsoeSigmund:2003}.
+
+If we pick the SIMP paramter $p=1$ 
+the thickness variable is allowed to take intermediate value apart from $0$ and $1$,
 i.e., we consider a variable thickness sheet. The discretization is done 
@@
 \centering
-\includegraphics[width=0.75\textwidth, viewport=205 88 975 500, clip]{Pictures/MBB.png}
-\caption{Thickness distribution for the MBB-beam minimum compliance problem.} 
+\hspace*{\fill}
+\includegraphics[width=0.44\textwidth, viewport=205 88 975 500, clip]{Pictures/MBB.png}
+\hspace{\fill}
+\includegraphics[width=0.44\textwidth, viewport=205 88 975 500, clip]{Pictures/MBB-top.png}
+\hspace{\fill}
+\caption{Thickness distribution for the MBB-beam minimum compliance problem using SIMP parameter $p=1$ (left) and $p=4$ right.} 
 \label{res:mbb}
 \end{figure}
-A typical thickness distribution is shown in Fig.~\ref{res:mbb}.
+A typical thickness distribution is shown in the left image in Fig.~\ref{res:mbb}.
+To obtain thickness values that are, mostly, either $1$ or $0$ we need 
+to pick a larger value for $p$. The distribution given in the right image in Fig.~\ref{res:mbb}
+has been obtained using $p=4$.
 

### 2012-12-05 Paper_May_2012/LetterToTheEditior.tex
Anschreiben an Editor

\documentclass{uhhbrief-fak6}    % Deutsch

\usepackage[utf8]{inputenc}

\usepackage{graphicx}
\usepackage{amsthm,amsmath,amsfonts,amssymb}
\usepackage{watermark}
\usepackage{url}

\selectlanguage{english} %For english Language

\name{Prof. Dr. Winnifried Wollner}
\signature{Winnifried Wollner}
\fbname{Fachbereich Mathematik}
\fbabkuerzung{FB Mathematik}
\strasse{Bundestr. 55}
\plzort{20146 Hamburg}
\telefonnummer{+49 (0)40 42838}
\telefondurchwahl{5988}
\faxnummer{+49 (0)40 42838}
\faxdurchwahl{5117}
\email{winnifried.wollner@math.uni-hamburg.de}
\http{www.math.uni-hamburg.de/home/wollner}

\date{\today}

\begin{document}
\thiswatermark{\begin{minipage}{\textwidth}\vspace*{43.5cm}\hspace*{5.8cm}\includegraphics[width=0.8\textwidth]{BG_logo}
  \end{minipage}
}
\begin{letter}{Editor-in-Chief of Computing\\Prof. Dr. Schahram Dustdar\\[\baselineskip]
    electronic}

\subject{Paper Submission}

\opening{Dear Professor Dustar,}

please find enclosed the manuscript entitled

{\em DOpElib: Differential Equations and Optimization Environment; A Goal Oriented Software Library for Solving PDEs and Optimization Problems with PDEs}

for publication in Computing by
C.Goll, T. Wick and W. Wollner.

The article outlines the guiding design principles of our software environment
DOpElib \url{dopelib.net} for the numerical solution of PDEs and optimization problems subject 
to PDEs with special emphasis on modular design of all components. 
We hope that you deem this publication worthy of publication in Computing.

\closing{Yours sincerely,}

%  \vspace*{-25mm}\includegraphics[width=.30\textwidth]{Signatur}\\[-2cm] $\mbox{}$\vspace{24mm}

%\ps{some ps}
%\cc{Some CC adresses}

%\encl{Some Anlagen}

\end{letter}
\end{document}

### 2012-12-05 Paper_May_2012/main.tex
Changed layout of cg add and cg pictures

@@
 \institute{Christian Goll \at
-              University of Heidelberg, Institute of Applied Mathematics, 
+              University of Heidelberg,\\ Institute of Applied Mathematics, 
               \\Im Neuenheimer Feld 293/294, 69120 Heidelberg, Germany\\
@@
 with the solution $u = (v,p)$, the normal $n$ and the viscosity $\nu = 0.001$.
+
 \begin{figure}[hbt]
 \centering
-{\includegraphics[width=0.85\textwidth]{Pictures/local_grid_NS.png}}
+{\includegraphics[trim = 50mm 50mm 50mm 50mm, clip, width=\textwidth]{Pictures/local_grid_NSv2.png}}
 \caption{Locally adapted grid after eight refinement steps.}
@@
 \end{figure}
+
 To achieve an efficient computation of the quantity $J$, we employ goal-oriented mesh refinement with the help of the DWR-method, see \cite{BeRa96}.
@@
 \centering
-\resizebox{0.45\textwidth}{!}{\input{Pictures/ConvergenceNS}}
+\resizebox{0.5\textwidth}{!}{\input{Pictures/ConvergenceNS}}
 \caption{Comparison of global and local mesh refinement with respect to the error in the drag-functional $J$.}

### 2012-12-05 Paper_May_2012/LetterToTheEditior.tex
Christians Comments, and resized a picture so its maximal big without changing the layout.

@@
 The article outlines the guiding design principles of our software environment
-DOpElib \url{dopelib.net} for the numerical solution of PDEs and optimization problems subject 
+DOpElib (see \url{dopelib.net}) for the numerical solution of PDEs and optimization problems subject 
 to PDEs with special emphasis on modular design of all components. 

### 2012-12-10 Paper_May_2012/main.tex
Changed layout of the examples of tw and ww. There was a big gap between the subsections after the last changes.

@@
-%\documentclass{svjour3}                     % onecolumn (standard format)
-%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
-\documentclass[smallextended]{svjour3}       % onecolumn (second format)
-%\documentclass[twocolumn]{svjour3}          % twocolumn
-%
-\smartqed  % flush right qed marks, e.g. at end of proof
-%
-\usepackage{graphicx}
-%
-% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
-%
-% insert here the call for the packages your document requires
-%\usepackage{latexsym}
-% etc.
-%
-% please place your own definitions here and don't use \def but
-% \newcommand{}{}
-%
-% Insert the name of "your journal" with
- \journalname{Computing}
-%
+\documentclass[prodmode,acmtoms]{acmsmall}
+
+% Package to generate and customize Algorithm as per ACM style
+\usepackage[ruled]{algorithm2e}
+\renewcommand{\algorithmcfname}{ALGORITHM}
+\SetAlFnt{\small}
+\SetAlCapFnt{\small}
+\SetAlCapNameFnt{\small}
+\SetAlCapHSkip{0pt}
+\IncMargin{-\parindent}
+
+% Metadata Information
+\acmVolume{?}
+\acmNumber{?}
+\acmArticle{??}
+\acmYear{???}
+\acmMonth{0}
 
@@
 
+\markboth{C. Goll et al.}{DOpElib: Differential Equations and Optimization Environment}
+
 \title{DOpElib: Differential Equations and Optimization Environment; A Goal Oriented Software Library for Solving PDEs and Optimization Problems with PDEs}
 
-\titlerunning{DOpElib}        % if too long for running head
-
-\author{C. Goll \and T. Wick \and W. Wollner}
-
-%\authorrunning{Short form of author list} % if too long for running head
-
-\institute{Christian Goll \at
-              University of Heidelberg,\\ Institute of Applied Mathematics, 
-              \\Im Neuenheimer Feld 293/294, 69120 Heidelberg, Germany\\
-              %Tel.: +49 622-154-6154\\
-              \email{christian.goll@iwr.uni-heidelberg.de}           %  \\
-           \and
-Thomas Wick \at
-              The Institute for Computational Engineering and Sciences,\\
-The University of Texas at Austin,\\
-Austin, Texas 78712, USA \\
-              %Tel.: +1 512-672-7763\\
-              \email{twick@ices.utexas.edu}   
- \and
-Winnifried Wollner \at
-              Universit\"at Hamburg\\
-              Department of Mathematics\\
-              Bundesstr. 55, 20146 Hamburg, Germany\\
-              \email{winnifried.wollner@math.uni-hamburg.de}   
-}
-
-\date{Received: date / Accepted: date}
-% The correct dates will be entered by the editor
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\maketitle
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\titlerunning{DOpElib}        % if too long for running head
+
+\author{CHRISTIAN GOLL
+\affil{University of Heidelberg}
+THOMAS WICK
+\affil{The University of Texas at Austin}
+WINNIFRIED WOLLNER
+\affil{University of Hamburg}}
+
 \begin{abstract}
@@
 numerical results as demonstration for the versatility of the software.
-
-\keywords{software environment \and goal oriented interface design \and numerical solution of optimization problems and partial differential equations}
-\subclass{49-04 \and  35-04}
 \end{abstract}
+
+\category{G.4}{Mathematical Software}{Optimization routines and interface to finite element software}
+\category{G.1.6}{Optimization}{Optimization with partial differential equations -- Constrained optimization}
+\category{G.1.8}{Partial Differential Equations}{Finite element method}
+%\todocg{ Siehe http://www.acm.org/about/class/1998/, ich komme damit aber nicht zurecht.}
+
+\terms{Algorithms, Design, Documentation}
+
+\keywords{goal oriented interface design, numerical solution of optimization problems and partial differential equations}
+
+\acmformat{Goll, C.; Wick, G.; Wollner, W.; DOpElib: Differential Equations and Optimization Environment.}
+
+\begin{bottomstuff}
+Author's addresses: C. Goll, Institute of Applied Mathematics, 
+              University of Heidelberg,
+              Im Neuenheimer Feld 293/294, 69120 Heidelberg, Germany,  christian.goll@iwr.uni-heidelberg.de;   
+T. Wick, The Institute for Computational Engineering and Sciences,
+The University of Texas at Austin,
+Austin, Texas 78712, USA, 
+              twick@ices.utexas.edu;   
+Winnifried Wollner,  
+              Universit\"at Hamburg,
+              Department of Mathematics,
+              Bundesstr. 55, 20146 Hamburg, Germany,winnifried.wollner@math.uni-hamburg.de;   
+\end{bottomstuff}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\maketitle
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 This is the Benchmark-Case 2D-1 described in \cite{TuSchae96}.
-\begin{figure}[h]
+\begin{figure}[hb]
 \centering
@@
 
-\begin{figure}[hbt]
+\begin{figure}[h]
 \centering
-{\includegraphics[trim = 50mm 50mm 50mm 50mm, clip, width=\textwidth]{Pictures/local_grid_NSv2.png}}
+{\includegraphics[trim = 50mm 50mm 50mm 50mm, clip, width=0.95\textwidth]{Pictures/local_grid_NSv2.png}}
 \caption{Locally adapted grid after eight refinement steps.}
@@
 % Bilder von FSI 2 
-\begin{figure}[h]
+\begin{figure}[hbt]
 \centering
-{\includegraphics[width=4.5cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_mesh8070_scale.png}}
-{\includegraphics[width=4.5cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_x_velo8070_scale.png}}
+{\includegraphics[width=5.5cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_mesh8070_scale.png}}
+{\includegraphics[width=5.5cm]{Pictures/visit_fsi_2_CNn_t_2e-2_global_3_biharmonic_x_velo8070_scale.png}}
 \caption{FSI 2 test case: mesh (left) and velocity profile in vertical 
@@
 
-\begin{figure}[H]
+\begin{figure}[hbt]
 \centering
-{\includegraphics[width=4.5cm]{Pictures/ux_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
-{\includegraphics[width=4.5cm]{Pictures/uy_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
-{\includegraphics[width=4.5cm]{Pictures/Drag_fluid_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
-{\includegraphics[width=4.5cm]{Pictures/Lift_fluid_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
+{\includegraphics[width=0.42\textwidth]{Pictures/ux_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
+{\includegraphics[width=0.42\textwidth]{Pictures/uy_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
+{\includegraphics[width=0.42\textwidth]{Pictures/Drag_fluid_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
+{\includegraphics[width=0.42\textwidth]{Pictures/Lift_fluid_FSI_2_FS_t_3e-2_t_15e-3_global_2_Hron_grid.pdf}}
 \caption{FSI 2: The deflections of the beam, $u_x(A)$ and $u_y(A)$ (in $cm$), and 
@@
 
-\newpage
 \subsection{Compliance Minimization}
@@
 thickness.
-\begin{figure}
+\begin{figure}[hbt]
 \centering
@@
 
-\section*{Thanks}
+%\section*{Thanks}
+\begin{acks}
 The \dope{} project makes use of various finite elements taken from  
@@
 possibility to finish this work.
+\end{acks}
 
@@
 \bibliography{lit}
-%
+% History dates
+\received{??}{??}{??}
 

### 2012-12-10 Paper_May_2012/Anschreiben/LetterToTheEditior_ACMTOMS.tex

--- 2012-12-05 Paper_May_2012/LetterToTheEditior.tex
@@
 }
-\begin{letter}{Editor-in-Chief of Computing\\Prof. Dr. Schahram Dustdar\\[\baselineskip]
+\begin{letter}{Editor-in-Chief of ACM Transactions on Mathematical Software\\Prof. Dr. Mike Heroux\\[\baselineskip]
     electronic}
@@
 
-\opening{Dear Professor Dustar,}
+\opening{Dear Professor Heroux,}
 
@@
 
-for publication in Computing by
+for publication in ACM Transactions on Mathematical Software by
 C.Goll, T. Wick and W. Wollner.
@@
 to PDEs with special emphasis on modular design of all components. 
-We hope that you deem this publication worthy of publication in Computing.
+We hope that you deem this publication worthy of publication in ACM Transactions on Mathematical Software.
 

### 2012-12-12 Examples/Documentation/examples.tex
Modified the Documentation according to suggestions by Francesco Ludovici

@@
 
-Adding a new regression test to the repository 
-needs some concentration of the user. In the 
-following, we explain how to do this and what 
-\textit{must} be taken into account:
+Before being able to change and compile the new code, the user must 
+follow some easy steps in order to modify the information related to the old code. In 
+this section we explain how to modify such information using as model 
+\texttt{PDE/StatPDE/Example1}.
+
+\begin{enumerate}
+ \item In a first step, we copy \texttt{Example1} and renamed it, e.g., 
+\texttt{ExampleNew}. At the same time it is important to remove the repository 
+information that it is stored in the directory \texttt{.svn.}. \\
+After having reached the folder of the example in question in the terminal, 
+\texttt{PDE/StatPDE} in our case, we perform these operation writing the following:
 \begin{verbatim}
@@
 \end{verbatim}
-It is important to remove the repository information,
-which is stored in the directory .svn.
+Please note that removing the \texttt{.svn} sub-directories is important,
+as otherwise your files may be replaced or changed during your next
+update. Also, if you can submit information to the subversion repository 
+you might accidentally overwrite the original example, here \texttt{Example1}.
 
-Then, we have to consider the following step. In order to 
-run the regression tests, we start the executable file
-of the present example. This executable comes from the 
-Makefile. For example in the Makefile of 
-Example \ref{PDE_Stat_Stokes}, 
+\item We now have to change the target in the \texttt{Makefile}, 
+  otherwise our new example will refer to the executable of the old one.
+
+  Thus we open the \texttt{Makefile} in the directory \texttt{ExampleNew}
+  to change the name of the executable. In our example you will find a line
 \begin{verbatim}
@@
 \end{verbatim}
-and concretely:
+Here we need to replace the target to a new name to avoid any 
+possible complication. Thus for example we change the line to 
 \begin{verbatim}
-DOpE-PDE-StatPDE-Example1-2d-2d
+target   = $(BINDIR)/MyExampleNew
 \end{verbatim}
-The aforementioned expression is the name of 
-the executable that is required to run the 
-regression test. Hence, in 
+This will build the executable file in the standard binary directory 
+of \texttt{DOpElib}, i.e., \texttt{\$(DOpE)/bin}, under the name 
+\texttt{MyExampleNew}.
+
+\textit{Note that you should use a different name for all your 
+examples as otherwise you may experience a lot of strange behavior, as
+the executable will be overwritten by all your examples!}
+
+\item If you want to run automated tests on you program so that you can 
+  verify whether your code is running as expected after updating the 
+  library you may want to update the sub-directory \texttt{Test} 
+  as well, see also Chapter~\ref{chap:test}. Otherwise you may skip this 
+  step.
+
+  Change to the \texttt{Test} sub-directory. And then modify the
+  test-script to contain the new name of the executable.
+  Assuming you want to use Emacs, open the file \texttt{test.sh}
 \begin{verbatim}
@@
 \end{verbatim}
-you find two times the line
+where, in our example you find two times the lines
 \begin{verbatim}
-echo "Running Program 
-     ../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm"
-     (../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm 2>&1) 
-     > /dev/null
+echo "Running Program ../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm"
+(../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm 2>&1) > /dev/null
 \end{verbatim} 
-It is really important when copying an existing example,
-to change this information. 
-First, we have to change the target in the Makefile
+if you made a copy of an other example the part \texttt{DOpE-PDE-StatPDE-Example1-2d-2d}
+may differ. These lines need to be replaced with the new name of the 
+executable, i.e., for our given example
 \begin{verbatim}
-target   = $(BINDIR)/DOpE-PDE-StatPDE-ExampleNew-$
-           (dope_dimension)d-$(deal_II_dimension)d
+echo "Running Program ../../../../../bin/MyExampleNew test.prm"
+(../../../../../bin/MyExampleNew test.prm 2>&1) > /dev/null
 \end{verbatim}
-Second, we have to replace 
-Example1 by ExampleNew such that 
-\begin{verbatim}
-echo "Running Program 
-     ../../../../../bin/DOpE-PDE-StatPDE-ExampleNew-2d-2d test.prm"
-     (../../../../../bin/DOpE-PDE-StatPDE-ExampleNew-2d-2d test.prm 2>&1) 
-     > /dev/null
-\end{verbatim} 
-Please, also change the dimension parameters 
-\begin{verbatim}
--2d-2d
-\end{verbatim} 
-Now, the user is prepared to change any information in the 
+
+\item  Now, you are prepared to change any of the problem
+  dependent data in information in the files 
 \begin{verbatim}
@@
 \end{verbatim}
+
+\item Once you have finished and are sure that your example is running correctly
+  and you want to use the automated test scripts --see 3) above-- You need 
+  to store new test information to account for your changes.
+
+  To do so, change to the \texttt{Test} sub-directory and run the test:
+\begin{verbatim}
+./test Test
+\end{verbatim}
+  Note that this should fail, otherwise you have not changed anything in the program, 
+  or forgot part 3) of this description.
+
+  If it failed have a look into the file \texttt{dope.log} and see whether you like the 
+  output. If you do not like it you may wish to update the file \texttt{test.prm} that 
+  takes care of the parameters for the test run.
+
+  Once your satisfied with what you see in the log-file \texttt{dope.log} you need to store 
+  that information using
+\begin{verbatim}
+./test Store
+\end{verbatim}
+\end{enumerate}

### 2013-01-08 Examples/PDE/StatPDE/Example11/content.tex
Added hp-variation of pde-statpde-example1

--- 2012-03-15 Examples/PDE/StatPDE/Example1/content.tex
@@
-\textbf{General problem description}\\
-
-\vspace{0.2cm}
+\subsubsection{General problem description}
 \index{stationary PDE}
-In this example we consider the stationary incompressible Stokes equation \index{stationary PDE!Stokes equation}. Here,
+In this example we consider the same setting as in subsection~\ref{PDE_Stat_Stokes}, the only
+difference is that we want to employ the hp-Finite-Elements. So the equation we solve is still 
+the stationary incompressible Stokes equation \index{stationary PDE!Stokes equation}. Here,
 we use the symmetric stress tensor which has a little consequence when using 
@@
 
-\vspace{0.2cm}
-
-\textbf{Program structure}\\
-
-\vspace{0.2cm}
-
-In all examples, the whole program is split up into several files for the sake of readability.
-These files are always denoted in the same way, so we only have to explain the general structure in this first example, whereas in the following examples, we will only point out differences to the current one.
-The content of the single files will be described in more detail below.\\
-
-If we do not use one of the standard grids given in the deal.II library, we can read a grid from an input file.
-In our example, the domain $\Omega = [-6,6]\times [0,2]$ is given in the \textit{channel.inp} file, where all nodes, cells and boundary lines are listed explicitly and the boundary is divided into disjoint parts by attributing different colors to the boundary lines.\index{grid}\\
-
-Certain parameters occurring during the solution process, e.g.
-error tolerances or the maximum number of iterations in an iterative solution procedure, are fixed in a parameter file called \textit{dope.prm}.
-This parameter file comprises several subsections corresponding to different solver components.\index{parameters}\\
-
-In the \textit{functionals.h} file we declare classes for different scalar quantities of interest (described mathematically as functionals) which we want to evaluate during the solution process.\index{functional}\\
-
-The \textit{localfunctional.h} file is relevant only if we want to solve an optimal control problem.
-In this case, it contains the cost functional, whereas the file is not needed for the forward solution of PDEs.
-We will get back to this later in the context of optimal control problems.\index{cost functional}\\
-
-All information about the PDE problem (in the optimal control case about the constraining PDE) is included in the \textit{localpde.h} file.
-In a class called \texttt{LocalPDE}, we build up the cell equation, the cell matrix and cell righthand side as well as the boundary equation, boundary matrix and boundary righthand side.
-Later on, the integrator collects this local information and creates the global vectors and matrices.\\
-
-The most important part of each example is the \textit{main.cc} file which contains the \texttt{int main()} function.
-Here we create objects of all classes described above and actually solve the respective problem.\\
-
-\vspace{0.2cm}
-
-\textbf{The} functionals.h \textbf{file}\\
-
-\vspace{0.2cm}
-\index{functional evaluation}
-Here, we declare all quantities of interest (functionals), e.g. point values, drag, lift, mean values of certain quantities over a subdomain etc. \\
-Each of these functionals is declared as a class of its own, but in DOpE all classes are derived from a so-called \texttt{FunctionalInterface} class.\\
-In the current example we declare functionals for point values of the velocity and for the flux at the outflow boundary of the channel.\index{functional!point value} \index{functional!boundary flux}\\
-
-\vspace{0.2cm}
-
-\textbf{The} localpde.h \textbf{file}\\
-
-\vspace{0.2cm}
-
-The \textit{LocalPDE} is derived from a \texttt{PDEInterface} class.
-It comprises several functions which build up the cell and boundary equations, matrices and righthand sides.
-The weak formulation of problem \eqref{eq:strong} with $f=0$ is
-\begin{equation} \label{eq:weak}
-   \frac{1}{2}(\nabla v, \nabla \varphi) + \frac{1}{2}(\nabla v^{T}, \nabla \varphi) - (p, \nabla \cdot \varphi) + (\nabla \cdot v, \psi) = 0.
-\end{equation}
-This problem is vector valued, i.e.
-the velocity variable $v$ has two components and the pressure variable $p$ is a scalar.
-For the implementation, we use a vector valued solution variable with three components, where the distinction between velocity and pressure is done by use of the deal.II \texttt{FEValuesExtractors} class.
-\index{vector-valued problem}\\
-Furthermore, in DOpE we always interpret the problems in the context of a Newton method. Usually, a PDE in its weak formulation is given as
-\begin{equation*}
-   a(u;\varphi) = f(\varphi).
-\end{equation*}
-The lefthand side is implemented in the \texttt{CellEquation} function, the righthand side is implemented in the \texttt{CellRightHandSide} function (which is unused in this example, because $f=0$).\\
-To apply Newton's method, this problem is linearized: on the lefthand side, we have the derivative of the (semilinear) form $a(\cdot;\cdot)$ with respect to the solution variable $u$, and the righthand side is the residual of the weak formulation:
-\begin{equation*}
-   a_u'(u;u^+,\varphi) = -a(u;\varphi) + f(\varphi).
-\end{equation*}
-In the \texttt{CellMatrix} function, we implement the following matrix $A$ as representation of the derivative on the lefthand side:
-\begin{equation*}
-  A = (a_u'(u;\varphi_i,\varphi_j))_{j,i=1}^N
-\end{equation*}
-with the number $N$ of the degrees of freedom. Similarly, the \texttt{CellEquation} contains the vector
-\begin{equation*}
-  a = a(u;\varphi_i)_{i=1}^{N},
-\end{equation*}
-and the \texttt{CellRightHandSide} in the case $f \neq 0$ would contain a vector
-\begin{equation*}
-  \tilde{f} = (f;\varphi_i)_{i=1}^N.
-\end{equation*}
-The system of equations which is then actually solved is
-\begin{equation*}
-  A\tilde{u}^+ = -a + \tilde{f}.
-\end{equation*}
-Because of the linearity of equation \eqref{eq:weak}, there is almost no difference between the two functions.\index{Newton's method}\\
-
-\vspace{0.1cm}
-
-\textit{At this point, it is important to note that DOpE interprets any given problem as a nonlinear one which is solved by Newton's method; the special case of linear problems is included into this general framework.}\\
-
-\vspace{0.2cm}
-
-\textbf{The} main.cc \textbf{file}\\
-
-\vspace{0.2cm}
-
-First of all, several header files have to be included that are needed during the solution process.
-We divide these includes into blocks corresponding to DOpE headers, deal.II headers, C++ headers and header files of the example itself (like the ones mentioned above).\\
-Furthermore, we define names for certain objects via \texttt{typedef} which act as abbreviations in order to keep the code readable.
-In our case, these are \texttt{OP, INTEGRATOR, LINEARSOLVER, NLS} and \texttt{SSOLVER}.\\
-In the \texttt{int main()} function, we first create a possibility to read the parameter values from the \textit{dope.prm} file. Then there are several standard steps for finite element codes like
-\begin{itemize}
-\item
-definition of a triangulation and create a grid object (which we read from the \textit{channel.inp} file)
-\item
-creation of finite element objects for the state and the control and of quadrature formula objects
-\end{itemize}
-and in addition, we 
-\begin{itemize}
-\item
-create objects of the \texttt{LocalPDE} class and of the different functional classes declared in the \textit{functionals.h} file (including an object of the \texttt{LocalFunctional} class, even if we do not regard an optimal control problem here).
-\end{itemize}
-\textit{REMARK:}\\
-Up to now  we have to create a pseudo time even for stationary problems.
-The \\\texttt{MethodOfLines\underline{ }StateSpaceTimeHandler} object (\texttt{DOFH}) which is needed for the initialization of \texttt{OP} requires a vector in which timepoints are specified.
-However, this is again merely a dummy variable, for we do not actually apply a time stepping method in the stationary case.
-This will also be removed in future versions of DOpE.\\
-
-Before we initialize the \texttt{SSolver} object and actually solve the problem, we have to set the correct boundary conditions.
-Via the \texttt{compmask} vector, we ensure that the boundary conditions are set only for the velocity components of our solution vector.
-We set homogeneous Dirichlet values at the upper and lower boundaries of the channel.
-The inflow is described by a parabolic profile at the left boundary (the corresponding function class is declared in the \textit{myfunctions.cc} file), whereas we do not prescribe anything at the outflow boundary (so-called do-nothing condition).\\
-The output of the program (the two functional values) is rather unspectacular; as the problem is linear, the solution is computed within one Newton step.
+\subsubsection{Adding hp-Elements}
+One sees by comparing  the \texttt{main.cc}-file of this problem with the one of subsection~\ref{PDE_Stat_Stokes} that the change to hp-Elements is really easy.
+TODO: continue, erklaere insbesondere indexsetter.

### 2013-01-08 Examples/description.tex
Adaptions in the nonstationary case for the dual-timesteppingscheme
Description um Opt Beispiel 9 erweitert
Docu of PDE/InstatPDE/Example1 has some flaws.
Update of the documentation, includes now higher order mappings.
deal.II -> Differential Equations and
TODO: Continue text for Exampe11

@@
 \begin{document}
-\title{DOpElib - Deal Optimization Environment}
+\title{DOpElib: 
+Differential Equations\\ and Optimization Environment}
 \author{
@@
 \clearpage
+\subsection{Usage of Higher Order Mappings: Approximation of $\pi$}
+\label{PDE_higher_order_mapping}
+\input{PDE/StatPDE/Example10/content.tex}
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Stationary Stokes Equations with hp-Elements} 
+\label{PDE_Stat_Stokes_hp}
+\input{PDE/StatPDE/Example11/content.tex}
+\clearpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Nonstationary Stokes Equations}
+\subsection{Nonstationary Navier-Stokes Equations}
 \label{PDE_Instat_Stokes}
@@
 \label{OPT_Stat_Param_Nonlin_FSI}
-%\input{OPT/StatPDE/Example5/content.tex}
+\input{OPT/StatPDE/Example9/content.tex}
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Subject to a Non-Stationary PDE}
+\label{OPT_Instat}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Control of a nonlinear heat equation via the initial values}
+\label{OPT_Instat initial-value end-time}
+\input{OPT/InstatPDE/Example1/content.tex}
 \clearpage

### 2013-01-10 Examples/PDE/StatPDE/Example11/content.tex
Forgot at last checking

@@
 \subsubsection{Adding hp-Elements}
-One sees by comparing  the \texttt{main.cc}-file of this problem with the one of subsection~\ref{PDE_Stat_Stokes} that the change to hp-Elements is really easy.
-TODO: continue, erklaere insbesondere indexsetter.
+One sees by comparing  the \texttt{main.cc}-file of this problem with the one of subsection~\ref{PDE_Stat_Stokes} that the change to hp-Elements is really easy, so we will keep the description short.
+
+The obvious things to change are some templa

### 2013-02-07 Examples/PDE/StatPDE/Example9/content.tex
Changed Interface of the method RefineSpace of the MOL-Spacetimehandlers, see their documentation for further details.

@@
 \end{verbatim}
-With these indicators, we are now able to refine our grid adaptively (there are several mesh adaption strategies implemented, like 'optimized', 'fixednumber' or 'fixedfraction')
+With these indicators, we are now able to refine our grid adaptively (there are several mesh adaption strategies implemented, like 'RefineOptimized', 'RefineFixedNumber' or 'RefineFixedFraction')
 \begin{verbatim}
-DOFH.RefineSpace("optimized", &error_ind);
+DOFH.RefineSpace(RefineOptimized(optimized, error_ind));
 \end{verbatim}

### 2013-02-07 Examples/OPT/StatPDE/Example6/main.cc

@@ -218,7 +218,7 @@ int main(int argc, char **argv)
     }
     if(i != niter-1)
     {
-      DOFH.RefineSpace("global");
+      DOFH.RefineSpace();
       {//Set Dirichlet Boundary!
 	for (Triangulation<2>::active_cell_iterator
 	       cell = triangulation.begin_active();

### 2013-04-10 Examples/OPT/StatPDE/Example1/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
Added DWR-Estimator for Optimization Problems (CostFunctional only, others will come in time...). See OPT/StatPDE/Example1 for details.

@@
  \overline{q} = \frac{1}{\alpha} \sin(\pi x) \sin(2 \pi y)\\
- \overline{u} = \sin(4 \pi x) \sin(2 \pi y)
+ \overline{u} = \sin(4 \pi x) \sin(2 \pi y).
+\end{gather*}
+Thus the exact optimal value of the cost functional can be calculated as 
+\begin{gather*}
+ J^* = J(\overline{q},\overline{u}) = \frac{1}{8}\Bigl(25\pi^4 + \frac{1}{\alpha}\Bigr).
 \end{gather*}
@@
 \end{gather*}
+
+The example shows how to estimate the error in the cost functional for stationary optimization
+problems.

### 2013-04-10 DOpEsrc/reducedproblems/statreducedproblem.h

@@ -1367,80 +1367,81 @@ namespace DOpE
             this->GetProblem()->GetSpaceTimeHandler()->GetStateDoFHandler().get_tria().n_active_cells();
         dwrc.ReInit(n_cells);
 
-        //If we need the dual solution, compute it
-        if (dwrc.NeedDual())
-          this->ComputeDualForErrorEstimation(q, dwrc.GetWeightComputation());
-
-        this->GetOutputHandler()->Write("Computing Error Indicators:",
+	//Estimation for Costfunctional
+	if(this->GetProblem()->EEFunctionalIsCost())
+	{
+	  this->GetOutputHandler()->Write("Computing Error Indicators:",
 					4 + this->GetBasePriority());
+	  this->GetProblem()->AddAuxiliaryToIntegrator(this->GetIntegrator());
+          
+          //add the primal and dual solution to the integrator
+	  this->GetIntegrator().AddDomainData("state",
+					      &(GetU().GetSpacialVector()));
+	  this->GetIntegrator().AddDomainData("adjoint_for_ee",
+					      &(GetZ().GetSpacialVector()));
 
-        this->GetProblem()->AddAuxiliaryToIntegrator(this->GetIntegrator());
-
-        //add the primal and (if needed) dual solution to the integrator
-        this->GetIntegrator().AddDomainData("state",
-            &(GetU().GetSpacialVector()));
-        if (dwrc.NeedDual())
-          this->GetIntegrator().AddDomainData("adjoint_for_ee",
-              &(GetZForEE().GetSpacialVector()));
-
-        if (dopedim == dealdim)
-        {
-          this->GetIntegrator().AddDomainData("control",
-              &(q.GetSpacialVector()));
-        }
-        else if (dopedim == 0)
-        {
-          this->GetIntegrator().AddParamData("control",
-              &(q.GetSpacialVectorCopy()));
-        }
-        else
-        {
-          throw DOpEException("dopedim not implemented",
+	  if (dopedim == dealdim)
+	  {
+	    this->GetIntegrator().AddDomainData("control",
+						&(q.GetSpacialVector()));
+	  }
+	  else if (dopedim == 0)
+	  {
+	    this->GetIntegrator().AddParamData("control",
+					       &(q.GetSpacialVectorCopy()));
+	  }
+	  else
+	  {
+	    throw DOpEException("dopedim not implemented",
               "StatReducedProblem::ComputeRefinementIndicators");
-        }
[…]

### 2013-05-08 Examples/Documentation/introduction.tex
Modified the Documentation according to suggestions by Francesco Ludovici
Additional documentation regardin the change of the RefineSpace-interface. Note that in order to just refine globally, the command DOFH.RefineSpace() suffices. If you want to use a local mesh refinement, you have to invoke DOFH.RefineSpace(ref) with ref being an object of the class RefineOptimized, RefineFixedFraction or RefineFixedNumber. See for example in Examples/PDE/StatPDE/Example9.
Added support for IPOPT

@@
   \item Michael Geiger (Examples for Plasticity, and Documentation of several PDE-Examples)
+  \item Francesco Ludovici (Augmenting the Documentation)
 \end{itemize}
@@
 \texttt{./configure --with-umfpack --with-blas --with-lapack}
+
+\subsubsection{IPOPT}
+If you would like to use the optimization routines offered by IPOPT
+\url{https://projects.coin-or.org/Ipopt} you can  
+install this yourself and add a symlink as described in \texttt{ThirdPartyLibs/README}.
+Alternatively, you can use installation script
+\texttt{ThirdPartyLibs/install-free-libs.sh}. Note that to use all 
+available linear solvers you may have to obtain a corresponding license 
+manually. This is true in particular for the HSL solvers MA27, $\ldots$.
+For information on these see the information provided by the installation
+script.
 
@@
   face data container.
+\item \texttt{container/refinementcontainer.h} The classes defined herein are given to the \texttt{RefineSpace} method of the \texttt{SpaceTimeHandler} and determine how we define the spatial mesh (i.e.
+globally or locally with a fixed fraction, fixed number or shomehow 'optimized' strategy).
 \end{itemize}

### 2013-05-08 Examples/OPT/StatPDE/Example7/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
Added support for IPOPT

@@
 The Problem is similar to that of {\tt OPT/StatPDE/Example1} except for the box control constraints.
-Another new feature is the use of the commercial optimization library SNOPT. 
+The problem is solved using the optimization library IPOPT that you can obtain for free. To use it 
+a correct link to the ipopt library needs to be created in  {\tt DOpE/ThirdPartyLibs} by the name 
+{\tt ipopt}, i.e., you should have the file
+{\tt DOpE/ThirdPartyLibs/ipopt} pointing to the ipopt directory. If you have not done this you can compile the
+example but when running the example you will only get an error message like\\
+{\tt Warning: During execution of `Reduced\underline{ }IpoptAlgorithm::Solve`\\ the following Problem occurred!\\
+To use this algorithm you need to have IPOPT installed! \\To use this set the WITH\underline{ }IPOPT CompilerFlag.}
+
+Alternatively the commercial optimization library SNOPT can be used in this example.
 In order to use this library you need to install SNOPT on your computer and then generate a symlink to 
@@
 {\tt Warning: During execution of `Reduced\underline{ }SnoptAlgorithm::Solve`\\ the following Problem occurred!\\
-To use this algorithm you need to have SNOPT installed! \\To use this set the WITH\underline{ }SNOP CompilerFlag.}
+To use this algorithm you need to have SNOPT installed! \\To use this set the WITH\underline{ }SNOPT CompilerFlag.}

### 2013-05-08 DOpEsrc/include/outputhandler.h

@@ -226,6 +226,10 @@ class DOpEOutputHandler
     void StartSaveCTypeOutputToLog();
     void StopSaveCTypeOutputToLog();
 
+    /**
+     * This function gives the current output directory
+     */
+    std::string GetResultsDir() const;
 
   protected:
     /**
@@ -275,6 +279,7 @@ class DOpEOutputHandler
      *                are beeing stored as a separate string in this list.
      */
     void ParseString(const std::string tmp, std::vector<std::string>& list);
+
   private:
     std::map<std::string, unsigned int> _iteration_type_pos;
     std::vector<unsigned int> _iteration_number;

### 2013-05-08 DOpEsrc/reducedproblems/ipopt_problem.h

@@ -1,0 +1,380 @@
+/**
+*
+* Copyright (C) 2012 by the DOpElib authors
+*
+* This file is part of DOpElib
+*
+* DOpElib is free software: you can redistribute it
+* and/or modify it under the terms of the GNU General Public
+* License as published by the Free Software Foundation, either
+* version 3 of the License, or (at your option) any later
+* version.
+*
+* DOpElib is distributed in the hope that it will be
+* useful, but WITHOUT ANY WARRANTY; without even the implied
+* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+* PURPOSE.  See the GNU General Public License for more
+* details.
+*
+* Please refer to the file LICENSE.TXT included in this distribution
+* for further information on this license.
+*
+**/
+
+#ifndef _IPOPT_PROBLEM_H_
+#define _IPOPT_PROBLEM_H_
+
+#ifdef WITH_IPOPT
+#include "IpTNLP.hpp"
+#endif
+
+#include "controlvector.h"
+
+#include <iostream>
+
+namespace DOpE
+{
+#ifdef WITH_IPOPT
+  template <typename RPROBLEM, typename VECTOR>
+    class Ipopt_Problem : public Ipopt::TNLP
+  {
+  public: 
+    Ipopt_Problem( int& ret_val, 
+		   RPROBLEM * OP, ControlVector<VECTOR>& q,
+		   const ControlVector<VECTOR>* q_min, 
+		   const ControlVector<VECTOR>* q_max, 
+		   const ConstraintVector<VECTOR>& c);
+    
+    virtual ~Ipopt_Problem() {}
+
+    /**@name Overloaded from TNLP */
+    //@{
+    /** Method to return some info about the nlp */
+    virtual bool get_nlp_info(Ipopt::Index& n, Ipopt::Index& m, Ipopt::Index& nnz_jac_g,
+			      Ipopt::Index& nnz_h_lag, IndexStyleEnum& index_style);
+
+    /** Method to return the bounds for my problem */
+    virtual bool get_bounds_info(Ipopt::Index n, Ipopt::Number* x_l, Ipopt::Number* x_u,
+				 Ipopt::Index m, Ipopt::Number* g_l, Ipopt::Number* g_u);
+    
+    /** Method to return the starting point for the algorithm */
[…]

### 2013-05-08 Examples/Documentation/introduction.tex
Updated the documentation

@@
   solve problems that have no PDE constraint.
+\item \texttt{reducedproblems/ipopt\underline{ }problem.h} A wrapper file that is used by the 
+  reduced\underline{ }ipopt\underline{ }algorithm when solving optimization problems. This file
+  hides the interface to IPOPT.
 \end{itemize}
@@
 \item \texttt{opt\underline{ }algorithms/reduced\underline{ }snopt\underline{ }algorithm.h}
+  An algorithm to solve reduced optimization problems with additional control constraints
+  using the third-party library SNOPT.
+  ((reduced) state constraints are not yet implemented.)
+\item \texttt{opt\underline{ }algorithms/reduced\underline{ }ipopt\underline{ }algorithm.h}
   An algorithm to solve reduced optimization problems with additional control constraints.
+  using the third-party library IPOPT.
   ((reduced) state constraints are not yet implemented.)

### 2013-05-09 Examples/PDE/InstatPDE/Example1/content.tex
Reworked section 4.2.1 of the documentation.

@@
 In this example we consider the nonstationary incompressible Navier-Stokes equation. As in the stationary PDE Example 3 \ref{PDE_Stat_FSI_STVK},
-we use the symmetric fluid stress tensor.
-
-\textbf{The rest of this parte is written for Stokes instead of Navier-Stokes. Needs to be adjusted.}
-
-In strong formulation we deal with 
+we use the symmetric fluid stress tensor, i.e. 
+in strong formulation we deal with 
 \begin{align*}
-\partial_t v -\nabla\cdot (\nabla v + \nabla v^{T}) + \nabla p &= f \\
-div\, v &= 0 
+\partial_t v -\nabla\cdot (\nabla v + \nabla v^{T}) + (v\cdot\nabla)v  +\nabla p &= f \\
+\nabla\cdot v &= 0 
 \end{align*}
@@
 
-As introduced earlier, 
-we formulate the time stepping scheme as \textit{One-step-$\theta$ scheme},
+As introduced earlier, we formulate the time stepping scheme as \textit{One-step-$\theta$ scheme},
 which are based on finite difference schemes.
+In order to keep the presentation simple, we describe the scheme using the stokes equation and thus neglecting the nonlinearity.
+Note that in the program we use the full Navier-Stokes operator.
+
 The time interval is given by $I=[0,T]$. Let $v^n,p^n$ and the time
-step $k=t^{n+1}-t^n$ be given. Find $v=v^{n+1}, p^{n+1}$ such that:
+step $k=t^{n+1}-t^n$ be given. Find $v^{n+1}, p^{n+1}$ such that:
 
 \begin{align*}
-v - k\theta (\nabla\cdot (\nabla v + \nabla v^{T}) + \nabla p) &=
-k\theta f^{n+1} + k(1-\theta)f^{n+1}\\
-&+  v^n + k(1-\theta) (\nabla\cdot (\nabla v^n + \nabla (v^n)^{T}) 
-+ \nabla p^n) \\
-div \,v &= 0 
+v^{n+1} - k\theta \left(\nabla\cdot (\nabla v^{n+1} + \nabla {v^{n+1}}^T) + \nabla p^{n+1}\right) &=
+k\theta f^{n+1} + k(1-\theta)f^{n}\\
+&+  v^n + k(1-\theta) \left(\nabla\cdot (\nabla v^n + \nabla (v^n)^{T}) 
++ \nabla p^n\right) \\
+\nabla \cdot v^{n+1} &= 0 
 \end{align*}
@@
 \begin{align*}
-v - k\theta (\nabla\cdot (\nabla v + \nabla v^{T}) + \nabla p) &=
+v^{n+1} - k (\nabla\cdot (\nabla v^{n+1} + \nabla {v^{n+1}}^{T}) + \nabla p^n) &=
 k\theta f^{n+1} +  v^n  \\
-div \,v &= 0
+\nabla \cdot v^{n+1} &= 0
 \end{align*}
@@
 \begin{Problem}[Backward Euler (BE) timestepping problem]
-Find $v:= v^{n+1}\in V$ and $p:= p^{n+1}\in L$:
-\begin{align*}
+Let $\theta =1$.Find $v:= v^{n+1}\in V$ and $p:= p^{n+1}\in L$:
+\begin{align}
 (v,\phi^v) + k\theta (\nabla v + \nabla v^{T}, \nabla\phi^v) 
@@
 k\theta (f^{n+1},\phi^v) +  (v^n,\phi^v)\\
-(div \,v,\phi^p) = 0  
-\end{align*}
+(\nabla \cdot v,\phi^p) = 0  \label{eq.divfreiheit}
+\end{align}
 \end{Problem}
 for all suitable test functions ${\phi^v , \phi^p} \in V\times L$.
-The parameter $\theta$ is chosen as $\theta =1$.
 
 Derivation of the other timestepping problems is analogous.
+\begin{remark}
+Note that because of the zero right hand side we are allowed to multiply \eqref{eq.divfreiheit} by $k\theta$.
+So that we solve $$k\theta(\nabla \cdot v,\phi^p) = 0$$ instead of $(\nabla \cdot v,\phi^p) = 0$.
+\end{remark}
 
-\textbf{Specific features for 
-solving nonstationary problems}\\
+\subsubsection{Specific features for 
+solving nonstationary problems}
 

### 2013-05-09 Examples/Documentation/introduction.tex
Updated docu to new wrapper files
Added and corrected Doxygen comments

@@
 install this yourself and add a symlink as described in \texttt{ThirdPartyLibs/README}.
+
 Alternatively, you can use installation script
@@
   solve problems that have no PDE constraint.
-\item \texttt{reducedproblems/ipopt\underline{ }problem.h} A wrapper file that is used by the 
-  reduced\underline{ }ipopt\underline{ }algorithm when solving optimization problems. This file
+\item \texttt{reducedproblems/ipopt\underline{ }problem.h} A wrapper file 
+ required when solving optimization problems using the 
+  reduced\underline{ }ipopt\underline{ }algorithm. This file
   hides the interface to IPOPT.
@@
 The standard implementation is just a wrapper for \texttt{dealii::DoFTools::} \texttt{make\_sparsity\_pattern}, but the user can reimplement this class to allow for more sophisticated sparsity patterns.
-\item \texttt{include/pointconstraintsmaker.h} This class is derived from \texttt{include/userdefineddofconstraints.h}
-  and allows to set homogeneous dirichlet values at given points/components.
+\item \texttt{include/pointconstraintsmaker.h} This class allows to set 
+  homogeneous dirichlet values at given points/components.
 \end{itemize}
@@
     by SNOPT using our library.
+  \item \texttt{wrapper/solutiontransfer\underline{ }wrapper.h} A wrapper for the SolutionTransfer class
+    from \texttt{deal.II}. 
+  \item \texttt{wrapper/dataout\underline{ }wrapper.h} A wrapper for the DataOut class from \texttt{deal.II}.
 \end{itemize}

### 2013-05-09 Examples/PDE/StatPDE/Example1/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
PDE StatPDE Example1 cleaned and reworked.

@@
-\textbf{General problem description}\\
-
-\vspace{0.2cm}
+\subsubsection{General problem description}
 \index{stationary PDE}
@@
 \end{align} 
-on the domain $\Omega = [-6,6]\times [0,2]$. We choose for simplicity $f=0$.\\
+on the domain $\Omega = [-6,6]\times [0,2]$.
+We split $\partial \Omega = \Gamma_D \cup \Gamma_{out}$.
+The right hand side of the channel is $\Gamma_{out}$ on which we describe the free outflow condition, on the rest of the boundary we prescribe dirichlet values (An parabolic inflow on the left hand side and zero on the upper and lower channel walls).
+We choose for simplicity $f=0$.\\
 
-\vspace{0.2cm}
-
-\textbf{Program structure}\\
-
-\vspace{0.2cm}
+\subsubsection{Program structure}
 
@@
 \begin{equation} \label{eq:weak}
-   \frac{1}{2}(\nabla v, \nabla \varphi) + \frac{1}{2}(\nabla v^{T}, \nabla \varphi) - (p, \nabla \cdot \varphi) + (\nabla \cdot v, \psi) = 0.
+   \frac{1}{2}(\nabla v, \nabla \varphi)_\Omega + \frac{1}{2}(\nabla v^{T}, \nabla \varphi)_\Omega - (p, \nabla \cdot \varphi)_\Omega + (\nabla \cdot v, \psi)_\Omega - (n\cdot \nabla v^T,\phi)_{\Gamma_{out}} = 0.
 \end{equation}
+\begin{remark}
+Note the additional term on $\Gamma_{out}$, which is a consequence of the use of the symmetric stress tensor together with the free outflow condition.
+\end{remark}
 This problem is vector valued, i.e.
@@
 \end{equation*}
-The lefthand side is implemented in the \texttt{CellEquation} function, the righthand side is implemented in the \texttt{CellRightHandSide} function (which is unused in this example, because $f=0$).\\
+The lefthand side is implemented in the \texttt{CellEquation} function, the righthand side is implemented in the \texttt{CellRightHandSide} function (which is unused in this example, because $f=0$).
+\begin{remark}
+The weak formulation might contain some terms on faces or (parts of) the boundary.
+DOpE is able to handle these via \texttt{BoundaryEquation}, \texttt{BoundaryRightHandSide} etc..
+To keep things simple, we neglect these terms in this introduction.
+\end{remark}
 To apply Newton's method, this problem is linearized: on the lefthand side, we have the derivative of the (semilinear) form $a(\cdot;\cdot)$ with respect to the solution variable $u$, and the righthand side is the residual of the weak formulation:
@@
 Furthermore, we define names for certain objects via \texttt{typedef} which act as abbreviations in order to keep the code readable.
-In our case, these are \texttt{OP, INTEGRATOR, LINEARSOLVER, NLS} and \texttt{SSOLVER}.\\
+In our case, these are \texttt{OP, IDC, INTEGRATOR, LINEARSOLVER, NLS, SSOLVER} and \texttt{STH}.\\
 In the \texttt{int main()} function, we first create a possibility to read the parameter values from the \textit{dope.prm} file. Then there are several standard steps for finite element codes like
@@
 \item
-create objects of the \texttt{LocalPDE} class and of the different functional classes declared in the \textit{functionals.h} file (including an object of the \texttt{LocalFunctional} class, even if we do not regard an optimal control problem here).
+create objects of the \texttt{LocalPDE} class and of the different functional classes declared in the \textit{functionals.h} file.
 \end{itemize}
-\textit{REMARK:}\\
+\begin{remark}
 Up to now  we have to create a pseudo time even for stationary problems.
@@
 This will also be removed in future versions of DOpE.\\
-
+\end{remark}
 Before we initialize the \texttt{SSolver} object and actually solve the problem, we have to set the correct boundary conditions.

### 2013-05-09 Examples/PDE/StatPDE/Example1/my_functions.h

@@ -27,15 +27,26 @@ using namespace dealii;
 
 
 /******************************************************/
+/**
+ * This function defines the inflowcondition.
+ */
 
 class BoundaryParabel : public DOpEWrapper::Function<2> 
 {
 public:
   BoundaryParabel () : DOpEWrapper::Function<2>(3) {}
   
+  /**
+   * Returns the value of the component 'component' of the function
+   * in point 'p'.
+   */
   virtual double value (const Point<2>   &p,
 			const unsigned int  component = 0) const;
   
+  /**
+   * Returns the value  of the function  in point 'p'.
+   */
+
   virtual void vector_value (const Point<2> &p, 
 			     Vector<double>   &value) const;
   

### 2013-05-09 Examples/PDE/StatPDE/Example2/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
PDE StatPDE Example2 cleaned and reworked

@@
-\textbf{General problem description}
-
-\vspace{0.2cm}
+\subsubsection{General problem description}
 
@@
 is given as an incompressible Newtonian fluid modelled by the Stokes equation.
-Here, we use the symmetric stress tensor which has a little consequence when using the do-nothing outflow condition.
+Here, we use the symmetric stress tensor which has a little consequence when using the do-nothing outflow condition, see also section \ref{PDE_Stat_Stokes}.
 \index{fluid} \index{fluid!incompressible} \index{fluid!Newtonian}
@@
 
-\vspace{0.2cm}
-
-\textbf{Program description}
-
-\vspace{0.2cm}
+\subsubsection{Program description}
 
 There is not much difference to the prior example, so we will keep our remarks rather short.
-The \textit{functionals.h, localfunctional.h} and \textit{myfunctions.cc} files did not change at all, so we simply refer to the corresponding sections in the last example.\\
+The \textit{functionals.h} and \textit{myfunctions.cc} files did not change at all, so we simply refer to the corresponding sections in the last example.\\
 In the \textit{localpde.h} file, all functions of the \texttt{LocalPDE} class have to be adjusted to the current FSI problem.

### 2013-05-09 Examples/PDE/StatPDE/Example6/content.tex
Further cleaning of PDE Statpde Exampe6 plus some bugfixes.

@@
 todo. short: this example shows how to use the constraintsmaker class by enforcing periodic boundary conditions in 2d.
+
+\subsubsection{General problem description}
+We solve the vector values laplace equation on a quadratic domain $\Omega$ with a circular hole in the middle, i.e. in strong formulation we look for $u=(u_1,u_2)$ s.t.
+\begin{align*}
+-\nabla \cdot \nabla u =& f &&\text{in }\Omega.
+\end{align*}
+We set zero dirichlet values on the circular boundary in the middle of the domain and periodic boundary conditions on the other parts of the boundary.
+We choose the flux over the right hand side boundary as functional.
+We choose
+\begin{align*}
+f(x,y) = \left(\cos\left(\exp(10  x)\right)  y^2x + \sin(y), \cos\left(\exp(10 * y)\right)x^2y + \sin(x)\right)
+\end{align*}
+for the right hand side.
+\subsubsection{Program description}
+This example show how to implement user defined DoF constraints.
+DOpE has an interface for this, namely \texttt{UserDefinedDoFConstraints}.
+In our case, we derive the class \texttt{PeriodicityConstraints}, overwrite the method \texttt{MakeStateDoFConstraints} and give an instance of this class to SpaceTimeHandler at hand:
+\begin{verbatim}
+ 	PeriodicityConstraints<DOFHANDLER, DIM> constraints_mkr;
+ 	STH DOFH(triangulation, state_fe);
+ 	DOFH.SetUserDefinedDoFConstraints(constraints_mkr);
+\end{verbatim}
+This is all it takes. We refer to \textit{myconstraintsmaker.h} for the details of the implementation of the periodicity-constraints.

### 2013-05-09 Examples/PDE/StatPDE/Example7/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
Further work on the examples

@@
-In this example we consider the stationary incompressible Stokes equation. Here,
-we use the symmetric stress tensor which has a little consequence when using 
-the do-nothing outflow condition. In strong formulation we have
-\begin{align*}
--\nabla\cdot (\nabla v + \nabla v^{T}) + \nabla p &= f \\
-\operatorname{div}\, v &= 0 
-\end{align*}
-on the domain $\Omega = [-6,6]\times [0,2]$. We choose for simplicity $f=0$.
+\subsubsection{General problem description}
+This is the simple vector valued laplace equation on the unitsquare with righthandside $(1,1)$ and zero dirichlet values. The classical example of a PDE.

### 2013-05-10 Examples/PDE/StatPDE/Example7/content.tex
Further work on the examples

@@
 \subsubsection{General problem description}
-This is the simple vector valued laplace equation on the unitsquare with righthandside $(1,1)$ and zero dirichlet values. The classical example of a PDE.
+In this problem we solve the simple vector valued laplace equation in 2d
+on the unit square $\Omega=[0,1]^2$, i.e. in strong formulation we look for $u=(u_1,u_2)$ s.t.
+\begin{align*}
+-\Delta u =& f &&\text{in }\Omega.
+\end{align*}
+We set zero dirichlet values on $\partial \Gamma$ and choose $f=(1,1)$. The classical example of a PDE.

### 2013-05-10 Examples/PDE/StatPDE/Example8/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
Further work on the examples

@@
-Laplace in 3D
-on locally refined grids with Kelly estimator
-with iterative solver (CG without preconditioning)
+\subsubsection{General problem description}
+In this problem we solve the simple vector valued laplace equation in 3d
+on the unit square $\Omega=[0,1]^3$, i.e. in strong formulation we look for $u=(u_1, u_2, u_3)$ s.t.
+\begin{align*}
+-\Delta u =& f &&\text{in }\Omega.
+\end{align*}
+We set zero dirichlet values on $\partial \Gamma$ and choose $f=(1,1,1)$.
+\subsubsection{Program description}
+The PDE is discretized with Q3-elements on a series of locally refined grids (we use the \texttt{KellyErrorEstimator}).
+The algebraic equations are solved with  different iterative linear solvers acting on different vector and matrix-structures (i.e.
+we use \texttt{dealii::BlockVector} and \texttt{dealii::Vector} plus the appropriate matrix classes).
+
+To switch the linear solver is pretty easy since the newton solver has a template for the linear solver. Thus changing this template is all that is required.
+
+To change the structure of the vectors and matrices involves also only the change of some template parameters.
+Our example programs are mostly build such that only a change of some typedefs is required, i.e.
+one has to interchange the lines
+\begin{verbatim}
+	typedef SparseMatrix<double> MATRIX;
+	typedef SparsityPattern SPARSITYPATTERN;
+	typedef Vector<double> VECTOR;
+\end{verbatim}
+with
+\begin{verbatim}
+	typedef BlockSparseMatrix<double> MATRIX;
+	typedef BlockSparsityPattern SPARSITYPATTERN;
+	typedef BlockVector<double> VECTOR;
+\end{verbatim}
+to switch between the block and non-block structures.
+
+After solving the equation, we want to apply local mesh refinement.
+So first we extract with the help of the \texttt{SolutionExatractr}-class the vector \texttt{solution} representing the finite element solution
+\begin{verbatim}
+  SolutionExtractor<SSolver1, VECTORBLOCK> a1(solver1);
+ 	const StateVector<VECTORBLOCK> &gu1 = a1.GetU();
+ 	solution = gu1.GetSpacialVector();
+\end{verbatim}
+With this vector we estiamte the error via \texttt{KellyErrorEstimator} and get a vector holding the estimated error per cell.
+After choosing a refinement criterion (see \textit{refinementcontainer.h}, we opt here for refining the top 20\% of the cells), we give the SpaceTimeHandler an object of type \texttt{RefinementContainer} which holds all the information needed for the local mesh refinement.
+This is done via the \texttt{RefineSpace} method.
+\begin{verbatim}
+        DOFH1.RefineSpace(
+               RefineFixedNumber(estimated_error_per_cell, 0.2, 0.0));
+\end{verbatim}
+This method transfers our solution onto the new mesh.
+The transferred solution is then taken as the starting guess of the newton method in the next solution cycle.
+This is especially helpful for nonlinear problems.

### 2013-05-10 Examples/PDE/InstatPDE/Example5/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
Changed DOpE to DOpElib in the documentation, finished overhaul of PDEStatPDE-Examples

@@
 
-\vspace{0.2cm}
-
-\textbf{Program description}
-
-\vspace{0.2cm}
+\subsubsection{Program description}
 
@@
 In examples \ref{PDE_Stat_Laplace_2D} and \ref{PDE_Stat_Laplace_3D}, we regarded stationary problems with nonhomogeneous righthand sides, but up to now, we never involved the time variable into the nonhomogeneity.
-To do this, DOpE yields a \texttt{SetTime()} function which has to be applied in the \textit{localpde.h} file as well as at the place where the \texttt{RightHandSideFunction} class is declared (here the \textit{myfunctions.h} file.
+To do this, \texttt{DOpElib} yields a \texttt{SetTime()} function which has to be applied in the \textit{localpde.h} file as well as at the place where the \texttt{RightHandSideFunction} class is declared (here the \textit{myfunctions.h} file.

### 2013-05-10 Examples/PDE/StatPDE/Example1/content.tex
Changed DOpE to DOpElib in the documentation, finished overhaul of PDEStatPDE-Examples

@@
 Here, we declare all quantities of interest (functionals), e.g. point values, drag, lift, mean values of certain quantities over a subdomain etc. \\
-Each of these functionals is declared as a class of its own, but in DOpE all classes are derived from a so-called \texttt{FunctionalInterface} class.\\
+Each of these functionals is declared as a class of its own, but in \texttt{DOpElib} all classes are derived from a so-called \texttt{FunctionalInterface} class.\\
 In the current example we declare functionals for point values of the velocity and for the flux at the outflow boundary of the channel.\index{functional!point value} \index{functional!boundary flux}\\
@@
 \index{vector-valued problem}\\
-Furthermore, in DOpE we always interpret the problems in the context of a Newton method. Usually, a PDE in its weak formulation is given as
+Furthermore, in \texttt{DOpElib} we always interpret the problems in the context of a Newton method. Usually, a PDE in its weak formulation is given as
 \begin{equation*}

### 2013-05-11 Examples/OPT/StatPDE/Example8/localfunctional.h
commit before talk of t and w

@@ -93,44 +93,44 @@ template<template<template<int, int> class DH, typename VECTOR, int dealdim> cla
       
     }
 
-     void BoundaryValue_Q(const FDC<DH, VECTOR, dealdim>& fdc __attribute__((unused)),
-			  dealii::Vector<double> &local_cell_vector __attribute__((unused)), 
-			  double scale __attribute__((unused)))
+     void BoundaryValue_Q(const FDC<DH, VECTOR, dealdim>& fdc ,
+			  dealii::Vector<double> &local_cell_vector , 
+			  double scale )
      {
      }
      
-    void BoundaryValue_UU(const FDC<DH, VECTOR, dealdim>& fdc __attribute__((unused)),
-			  dealii::Vector<double> &local_cell_vector __attribute__((unused)), 
-			  double scale __attribute__((unused)))
+    void BoundaryValue_UU(const FDC<DH, VECTOR, dealdim>& fdc ,
+			  dealii::Vector<double> &local_cell_vector , 
+			  double scale )
     {
     }
      
-    void BoundaryValue_QU(const FDC<DH, VECTOR, dealdim>& fdc __attribute__((unused)),
-			  dealii::Vector<double> &local_cell_vector __attribute__((unused)), 
-			  double scale __attribute__((unused)))
+    void BoundaryValue_QU(const FDC<DH, VECTOR, dealdim>& fdc ,
+			  dealii::Vector<double> &local_cell_vector , 
+			  double scale )
     {
     }
     
-    void BoundaryValue_UQ(const FDC<DH, VECTOR, dealdim>& fdc __attribute__((unused)),
-			  dealii::Vector<double> &local_cell_vector __attribute__((unused)), 
-			  double scale __attribute__((unused)))
+    void BoundaryValue_UQ(const FDC<DH, VECTOR, dealdim>& fdc ,
+			  dealii::Vector<double> &local_cell_vector , 
+			  double scale )
     {
     }
      
-    void BoundaryValue_QQ(const FDC<DH, VECTOR, dealdim>& fdc __attribute__((unused)),
-			  dealii::Vector<double> &local_cell_vector __attribute__((unused)), 
-			  double scale __attribute__((unused)))
+    void BoundaryValue_QQ(const FDC<DH, VECTOR, dealdim>& fdc ,
+			  dealii::Vector<double> &local_cell_vector , 
+			  double scale )
     {
     }
 
[…]

### 2013-05-11 Examples/OPT/StatPDE/Example1/main.cc
Code cleaned up to OPT StatPDE Ex2. TODO: content.tex

@@ -23,19 +23,15 @@
 #include <iostream>
 
 #include <grid/tria.h>
-//#include <dofs/dof_handler.h>
 #include <grid/grid_generator.h>
 #include <fe/fe_q.h>
-//#include <dofs/dof_tools.h>
 #include <base/quadrature_lib.h>
-//#include <base/function.h>
 #include <lac/precondition_block.h>
 
 #include "reducednewtonalgorithm.h"
 #include "reducedtrustregionnewton.h"
 #include "optproblemcontainer.h"
 #include "functionalinterface.h"
-//#include "pdeinterface.h"
 #include "statreducedproblem.h"
 #include "newtonsolver.h"
 #include "cglinearsolver.h"
@@ -45,8 +41,6 @@
 #include "mol_spacetimehandler.h"
 #include "simpledirichletdata.h"
 #include "noconstraints.h"
-//#include "sparsitymaker.h"
-//#include "userdefineddofconstraints.h"
 #include "preconditioner_wrapper.h"
 #include "integratordatacontainer.h"
 #include "higher_order_dwrc_control.h"
@@ -61,7 +55,7 @@ using namespace DOpE;
 
 //Some abbreviations for better readability
 const static int DIM = 2;
-const static int CDIM = 2
+const static int CDIM = 2;
 //stands for the dimension of the controlvariable;
 
 #define DOFHANDLER DoFHandler
@@ -111,7 +105,7 @@ typedef StatReducedProblem<NLS, NLS, INTEGRATOR, INTEGRATOR, OP, VECTOR, CDIM,
     DIM> SSolver;
 
 //The spacetimehandler manages all the things related to the degrees of
-//freedom in space and time (for the optimization as wella s the state variable!)
+//freedom in space and time (for the optimization as well as the state variable!)
 typedef MethodOfLines_SpaceTimeHandler<FE, DOFHANDLER, SPARSITYPATTERN, VECTOR,
     CDIM, DIM> STH;
 
@@ -122,7 +116,12 @@ int
 main(int argc, char **argv)
 {
   /**
+   * This class solves a distributed minimization problem with
+   * the laplacian as PDE constraint and the control on the right
+   * hand side of the PDE. The target functional is
+   * $\frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2$.
    *
[…]

### 2013-05-11 Examples/OPT/StatPDE/Example4/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
OPT StatPDE Example 3 and 4 finished, removed unnecessary template parameter in localpdes. TODOproper description of the opt examples, especially 1 and 2

@@
+\subsubsection{General problem description}
 This example solves the minimization problem
@@
 with $\alpha = 10$.
+\subsubsection{Program description}
+The control in the dirichlet boundary values is incorporated via the class \texttt{LocalDirichletData} which is defined in the file \textit{localdirichletdata.h}.
+The class is then given to the \texttt{OptProblemContainer} as a template argument.
+This is all that is needed to use the control in the dirichlet boundary values.

### 2013-05-11 Examples/OPT/StatPDE/Example3/localfunctional.h

@@ -50,18 +50,18 @@ template<
 
       LocalFunctional(ParameterReader &param_reader)
       {
-        // Control- and regulraization parameters
+        // Control- and regularization parameters
         mu_regularization = 1.0e+1;
         upper_bound_for_control_sum = 1.0e-2;
 
-        // Fluid- and material parameters
+        // Fluid parameters
         param_reader.SetSubsection("Local PDE parameters");
         _density_fluid = param_reader.get_double("density_fluid");
         _viscosity = param_reader.get_double("viscosity");
       }
 
       double
-      BoundaryValue(const FaceDataContainer<DH, VECTOR, dealdim>& fdc)
+      BoundaryValue(const FDC<DH, VECTOR, dealdim>& fdc)
       {
         const auto & state_fe_face_values = fdc.GetFEFaceValuesState();
         unsigned int n_q_points = fdc.GetNQPoints();
@@ -96,7 +96,6 @@ template<
             grad_v[1][0] = _ufacegrads[q_point][1][0];
             grad_v[1][1] = _ufacegrads[q_point][1][1];
 
-            // constitutive stress tensors for fluid
             Tensor<2, 2> cauchy_stress_fluid;
             cauchy_stress_fluid =
                 500.0
@@ -147,7 +146,7 @@ template<
       }
 
       void
-      BoundaryValue_U(const FaceDataContainer<DH, VECTOR, dealdim>& fdc,
+      BoundaryValue_U(const FDC<DH, VECTOR, dealdim>& fdc,
           dealii::Vector<double> &local_cell_vector, double scale)
       {
         const auto & state_fe_face_values = fdc.GetFEFaceValuesState();
@@ -190,7 +189,7 @@ template<
       }
 
       void
-      BoundaryValue_Q(const FaceDataContainer<DH, VECTOR, dealdim>& fdc,
+      BoundaryValue_Q(const FDC<DH, VECTOR, dealdim>& fdc,
           dealii::Vector<double> &local_cell_vector, double scale)
       {
         const auto & state_fe_face_values = fdc.GetFEFaceValuesState();
@@ -231,7 +230,7 @@ template<
       }
 
       void
-      BoundaryValue_QQ(const FaceDataContainer<DH, VECTOR, dealdim>& fdc,
[…]

### 2013-05-11 Examples/OPT/StatPDE/Example4/localfunctional.h

@@ -1,25 +1,25 @@
 /**
-*
-* Copyright (C) 2012 by the DOpElib authors
-*
-* This file is part of DOpElib
-*
-* DOpElib is free software: you can redistribute it
-* and/or modify it under the terms of the GNU General Public
-* License as published by the Free Software Foundation, either
-* version 3 of the License, or (at your option) any later
-* version.
-*
-* DOpElib is distributed in the hope that it will be
-* useful, but WITHOUT ANY WARRANTY; without even the implied
-* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
-* PURPOSE.  See the GNU General Public License for more
-* details.
-*
-* Please refer to the file LICENSE.TXT included in this distribution
-* for further information on this license.
-*
-**/
+ *
+ * Copyright (C) 2012 by the DOpElib authors
+ *
+ * This file is part of DOpElib
+ *
+ * DOpElib is free software: you can redistribute it
+ * and/or modify it under the terms of the GNU General Public
+ * License as published by the Free Software Foundation, either
+ * version 3 of the License, or (at your option) any later
+ * version.
+ *
+ * DOpElib is distributed in the hope that it will be
+ * useful, but WITHOUT ANY WARRANTY; without even the implied
+ * warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+ * PURPOSE.  See the GNU General Public License for more
+ * details.
+ *
+ * Please refer to the file LICENSE.TXT included in this distribution
+ * for further information on this license.
+ *
+ **/
 
 #ifndef _LOCALFunctional_
 #define _LOCALFunctional_
@@ -30,178 +30,215 @@ using namespace std;
 using namespace dealii;
 using namespace DOpE;
 
-template<template<int, int> class DH, typename VECTOR, int dopedim, int dealdim>
-  class LocalFunctional : public FunctionalInterface<CellDataContainer,FaceDataContainer,DH, VECTOR, dopedim,dealdim>
+template<
+    template<template<int, int> class DH, typename VECTOR, int dealdim> class CDC,
+    template<template<int, int> class DH, typename VECTOR, int dealdim> class FDC,
[…]

### 2013-05-11 Examples/OPT/StatPDE/Example5/content.tex
Added some todos
Overhaul of OPT StatPDe Example 5

@@
 In addition the following functionals are evaluated:
-\begin{gather*}
-  \text{MidPoint: } u_h(0.5 ; 0.5)\\[2mm]
-  \text{L1-Value: }\int_\Omega |u_h|
-  \text{QError: }\int_\Omega |q_h-\overline{q}|^2
-  \text{UError: }\int_\Omega |u_h-\overline{u}|^2
-\end{gather*}
+\begin{align*}
+  &\text{MidPoint: } u_h(0.5 ; 0.5)&&
+  \text{ L1-Value: }\int_\Omega |u_h|\\
+  &\text{QError: }\int_\Omega |q_h-\overline{q}|^2
+  &&\text{  UError: }\int_\Omega |u_h-\overline{u}|^2
+\end{align*}
 
 The important new feature is that we can now use two different meshes for control and state variable.
-This is tested first for globally refined meshes, and then for locally refined meshes with different refinements 
-for the control and state variable.
+This is tested first for globally refined meshes, and then for locally refined meshes with different refinements for the control and state variable.
+\subsubsection{Program description}
+In order to use different meshes for control and state we need to use the multimesh variants of \texttt{CellDataContainer}, \texttt{FaceDataContainer} and \texttt{Integrator} and we have to choose a space time handler capable of managing multiple meshes, so we use
+
+\texttt{MethodOfLines\_MultiMesh\_SpaceTimeHandler}.
+
+The requirement for the control and state mesh is that they have a common coarse grid, so the space time handler gets only one mesh (to ensure a common coarse grid), but this gets copied internally so that we have two seperate meshes for control and state.
+We can then separetely refine the mesh for control and state (see \texttt{RefineControlSpace} and  \texttt{RefineStateSpace}).

### 2013-05-11 Examples/OPT/StatPDE/Example5/functionals.h

@@ -1,25 +1,25 @@
 /**
-*
-* Copyright (C) 2012 by the DOpElib authors
-*
-* This file is part of DOpElib
-*
-* DOpElib is free software: you can redistribute it
-* and/or modify it under the terms of the GNU General Public
-* License as published by the Free Software Foundation, either
-* version 3 of the License, or (at your option) any later
-* version.
-*
-* DOpElib is distributed in the hope that it will be
-* useful, but WITHOUT ANY WARRANTY; without even the implied
-* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
-* PURPOSE.  See the GNU General Public License for more
-* details.
-*
-* Please refer to the file LICENSE.TXT included in this distribution
-* for further information on this license.
-*
-**/
+ *
+ * Copyright (C) 2012 by the DOpElib authors
+ *
+ * This file is part of DOpElib
+ *
+ * DOpElib is free software: you can redistribute it
+ * and/or modify it under the terms of the GNU General Public
+ * License as published by the Free Software Foundation, either
+ * version 3 of the License, or (at your option) any later
+ * version.
+ *
+ * DOpElib is distributed in the hope that it will be
+ * useful, but WITHOUT ANY WARRANTY; without even the implied
+ * warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+ * PURPOSE.  See the GNU General Public License for more
+ * details.
+ *
+ * Please refer to the file LICENSE.TXT included in this distribution
+ * for further information on this license.
+ *
+ **/
 
 #ifndef _LOCALFunctionalS_
 #define _LOCALFunctionalS_
@@ -30,176 +30,225 @@ using namespace std;
 using namespace dealii;
 using namespace DOpE;
 
-template<template<int, int> class DH,typename VECTOR, int dopedim, int dealdim>
-  class QErrorFunctional : public FunctionalInterface<Multimesh_CellDataContainer,Multimesh_FaceDataContainer,DH, VECTOR,dopedim,dealdim>
+template<
+    template<template<int, int> class DH, typename VECTOR, int dealdim> class CDC,
+    template<template<int, int> class DH, typename VECTOR, int dealdim> class FDC,
+    template<int, int> class DH, typename VECTOR, int dopedim, int dealdim =
+        dopedim>
+  class QErrorFunctional : public FunctionalInterface<CDC, FDC, DH, VECTOR,
[…]

### 2013-05-11 Examples/Documentation/introduction.tex
Added doxygen comments to opt_algorithms. Moved mma_algorithm and related files to OPT/StatPDE/Examples6 which is the only example where this is useful.

@@
   nonstationary PDE.
-\item \texttt{reducedproblems/voidreducedproblem.h} A wrapper file that eliminates $u$ 
-  if it is not present anyways. This is used so that we can use the same routines to 
-  solve problems that have no PDE constraint.
 \item \texttt{reducedproblems/ipopt\underline{ }problem.h} A wrapper file 
@@
   PDE constraint.)
-\item \texttt{opt\underline{ }algorithms/generalized\underline{ }mma\underline{ }algorithm.h}
-  An implementation of the MMA-Algorithm for structural optimization using an augmented
-  Lagrangian formulation for the subproblems. The subproblem is implemented using the 
-  special purpose file
-  \texttt{include/augmentedlagrangianproblem.h}.
 \end{itemize}

### 2013-05-11 Examples/OPT/StatPDE/Example6/content.tex
Example 6 done, changes to localnsontraintacc
Added doxygen comments to opt_algorithms. Moved mma_algorithm and related files to OPT/StatPDE/Examples6 which is the only example where this is useful.

@@
-This example implements the minimum compliance problem for the thickness optimization 
+\subsubsection{General problem description}
+1This example implements the minimum compliance problem for the thickness optimization 
 of an MBB-Beam. Using the MMA-Method of K. Svanberg together with an augmented Lagrangian 
 approach for the subproblems following M. Stingl.
+%\todo{Gleichung, reference}
+
+The implementation is done using the following three additional files:
+
+\begin{itemize}
+\item \texttt{generalized\underline{ }mma\underline{ }algorithm.h}
+  An implementation of the MMA-Algorithm for structural optimization using an augmented
+  Lagrangian formulation for the subproblems. The subproblem is implemented using the 
+  special purpose file \texttt{augmentedlagrangianproblem.h}.
+\item \texttt{augmentedlagrangianproblem.h} The problem container 
+  for the augmented Lagrangian problem.
+\item \texttt{voidreducedproblem.h} A wrapper file that eliminates $u$ 
+  if it is not present anyways. This is used so that we can use the same routines to 
+  solve problems that have no PDE constraint. This is used to fit the augmentedlagrangian 
+  problem into our framework.
+\end{itemize}

### 2013-05-11 Examples/OPT/StatPDE/Example6/voidreducedproblem.h

@@ -1081,19 +1081,5 @@ void VoidReducedProblem<CONTROLNONLINEARSOLVER, CONTROLINTEGRATOR, PROBLEM, VECT
 }
 /******************************************************/
 
-//  template<typename CONTROLNONLINEARSOLVER, typename CONTROLINTEGRATOR,
-//      typename PROBLEM, typename VECTOR, int dopedim, int dealdim>
-//    template<class DWRCONTAINER>
-//      void
-//      VoidReducedProblem<CONTROLNONLINEARSOLVER, CONTROLINTEGRATOR, PROBLEM,
-//          VECTOR, dopedim, dealdim>::ComputeRefinementIndicators(
-//          DWRCONTAINER& dwrc)
-//
-//      {
-//        throw DOpEException(
-//            "Function makes no sense in this context because we have no state variable.",
-//            "VoidReducedProblem::ComputeRefinementIndicators");
-//        return 0;
-//      }
 }
 #endif

### 2013-05-11 Examples/OPT/StatPDE/Example7/functionals.h
Opt StatPDE Example 7 done

@@ -1,108 +1,129 @@
 /**
-*
-* Copyright (C) 2012 by the DOpElib authors
-*
-* This file is part of DOpElib
-*
-* DOpElib is free software: you can redistribute it
-* and/or modify it under the terms of the GNU General Public
-* License as published by the Free Software Foundation, either
-* version 3 of the License, or (at your option) any later
-* version.
-*
-* DOpElib is distributed in the hope that it will be
-* useful, but WITHOUT ANY WARRANTY; without even the implied
-* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
-* PURPOSE.  See the GNU General Public License for more
-* details.
-*
-* Please refer to the file LICENSE.TXT included in this distribution
-* for further information on this license.
-*
-**/
+ *
+ * Copyright (C) 2012 by the DOpElib authors
+ *
+ * This file is part of DOpElib
+ *
+ * DOpElib is free software: you can redistribute it
+ * and/or modify it under the terms of the GNU General Public
+ * License as published by the Free Software Foundation, either
+ * version 3 of the License, or (at your option) any later
+ * version.
+ *
+ * DOpElib is distributed in the hope that it will be
+ * useful, but WITHOUT ANY WARRANTY; without even the implied
+ * warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+ * PURPOSE.  See the GNU General Public License for more
+ * details.
+ *
+ * Please refer to the file LICENSE.TXT included in this distribution
+ * for further information on this license.
+ *
+ **/
 
 #ifndef _LOCALFunctionalS_
 #define _LOCALFunctionalS_
 
-#include "pdeinterface.h"
+#include "functionalinterface.h"
 
 using namespace std;
 using namespace dealii;
 using namespace DOpE;
 
-template<template<int, int> class DH, typename VECTOR, int dopedim, int dealdim>
-  class LocalMeanValueFunctional : public FunctionalInterface<CellDataContainer,FaceDataContainer,DH, VECTOR,dopedim,dealdim>
+template<
+    template<template<int, int> class DH, typename VECTOR, int dealdim> class CDC,
+    template<template<int, int> class DH, typename VECTOR, int dealdim> class FDC,
+    template<int, int> class DH, typename VECTOR, int dopedim, int dealdim =
+        dopedim>
+  class LocalMeanValueFunctional : public FunctionalInterface<CDC, FDC, DH,
[…]

### 2013-05-11 Examples/OPT/StatPDE/Example9/content.tex
Finished overhaul of OPT StatPDE examples. 7 and 8 are untested due to missing snopt

@@
+\subsubsection{General program description}
 In this example we solve an optimization problem where 
 the equations come from fluid-structure interaction (FSI). The idea
-is to extend the steady FSI benchmark problem (FSI 1) to an optimization 
-problem where the drag is minimized over the cylinder plus beam.
+is to extend the steady FSI benchmark problem (FSI 1, proposed by Hron/Turek) to an optimization 
+problem where the drag is minimized over the cylinder and the beam.
 
@@
 
-Detailed description will follow later.
+Detailed description will follow later. %TODO

### 2013-05-11 Examples/OPT/StatPDE/Example9/localpde.h

@@ -1,25 +1,25 @@
 /**
-*
-* Copyright (C) 2012 by the DOpElib authors
-*
-* This file is part of DOpElib
-*
-* DOpElib is free software: you can redistribute it
-* and/or modify it under the terms of the GNU General Public
-* License as published by the Free Software Foundation, either
-* version 3 of the License, or (at your option) any later
-* version.
-*
-* DOpElib is distributed in the hope that it will be
-* useful, but WITHOUT ANY WARRANTY; without even the implied
-* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
-* PURPOSE.  See the GNU General Public License for more
-* details.
-*
-* Please refer to the file LICENSE.TXT included in this distribution
-* for further information on this license.
-*
-**/
+ *
+ * Copyright (C) 2012 by the DOpElib authors
+ *
+ * This file is part of DOpElib
+ *
+ * DOpElib is free software: you can redistribute it
+ * and/or modify it under the terms of the GNU General Public
+ * License as published by the Free Software Foundation, either
+ * version 3 of the License, or (at your option) any later
+ * version.
+ *
+ * DOpElib is distributed in the hope that it will be
+ * useful, but WITHOUT ANY WARRANTY; without even the implied
+ * warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+ * PURPOSE.  See the GNU General Public License for more
+ * details.
+ *
+ * Please refer to the file LICENSE.TXT included in this distribution
+ * for further information on this license.
+ *
+ **/
 
 #ifndef _LOCALPDE_
 #define _LOCALPDE_
@@ -31,2208 +31,2256 @@ using namespace std;
 using namespace dealii;
 using namespace DOpE;
 
-template<template<int, int> class DH,typename VECTOR, int dealdim>
-  class LocalPDE : public PDEInterface<CellDataContainer,FaceDataContainer,DH, VECTOR, dealdim>
+template<
+    template<template<int, int> class DH, typename VECTOR, int dealdim> class CDC,
+    template<template<int, int> class DH, typename VECTOR, int dealdim> class FDC,
+    template<int, int> class DH, typename VECTOR, int dealdim>
+  class LocalPDE : public PDEInterface<CDC, FaceDataContainer, DH, VECTOR,
[…]

### 2013-05-11 Examples/description.tex
Reworked section 4.2.1 of the documentation.
Example 6 done, changes to localnsontraintacc
Opt StatPDE Example 7 done
Updated name

@@
 \documentclass[a4paper,cleardoubleempty]{scrreprt}
-\usepackage{amsmath,amssymb, commath}
+\usepackage{amsmath,amssymb, amsthm, commath}
 \usepackage[pdftex,bookmarks=false,pdfpagelabels=true,plainpages=false]{hyperref}
@@
 %\usepackage{sfmath}
+\usepackage{color}
+\theoremstyle{plain}
 \newtheorem{Problem}{Problem}[section]
 
-\newcommand{\todo}[1]{\textbf{\textsc{\textcolor{black}{TODO: #1}}}}
+\theoremstyle{remark}
+\newtheorem{remark}[Problem]{Remark}
+
+\newcommand{\todo}[1]{\textbf{\textsc{\textcolor{red}{TODO: #1}}}}
 
@@
     {christian.goll@iwr.uni-heidelberg.de}}, 
-  Thomas Wick\thanks{Heidelberg University, 
-    {thomas.wick@iwr.uni-heidelberg.de}}, 
+  Thomas Wick\thanks{The University of Texas at Austin, 
+    {twick@ices.utexas.edu}}, 
   Winnifried Wollner\thanks{University of Hamburg, 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Distributed control with a linear elliptic PDE using SNOPT}
+\subsection{Distributed control with a linear elliptic PDE using IPOPT/SNOPT}
 \label{OPT_Stat_Box_controlconstraints_SNOPT}

### 2013-06-07 Examples/PDE/StatPDE/Example3/content.tex
Preparation for moving the examples to a new ordering.

@@
-\textbf{General problem description}
+\subsubsection{General problem description}
+\index{stationary PDE}
+In this example we consider the same setting as in subsection~\ref{PDE_Stat_Stokes}, the only
+difference is that we want to employ the hp-Finite-Elements. So the equation we solve is still 
+the stationary incompressible Stokes equation \index{stationary PDE!Stokes equation}. Here,
+we use the symmetric stress tensor which has a little consequence when using 
+the do-nothing outflow condition. In strong formulation we have
+\begin{align} 
+-\frac{1}{2}\nabla\cdot (\nabla v + \nabla v^{T}) + \nabla p &= f \\ 
+\nabla \cdot v &= 0 \notag
+\end{align} 
+on the domain $\Omega = [-6,6]\times [0,2]$.
+We split $\partial \Omega = \Gamma_D \cup \Gamma_{out}$.
+The right hand side of the channel is $\Gamma_{out}$ on which we describe the free outflow condition, on the rest of the boundary we prescribe dirichlet values (An parabolic inflow on the left hand side and zero on the upper and lower channel walls).
+We choose for simplicity $f=0$..\\
 
-\vspace{0.2cm}
+\subsubsection{Adding hp-Elements}
+One sees by comparing  the \texttt{main.cc}-file of this problem with the one of subsection~\ref{PDE_Stat_Stokes} that the change to hp-Elements is really easy, so we will keep the description short.
+In comparison to example \ref{PDE_Stat_Stokes} the \textit{localpde.h} and \textit{functionals.h} have not changed, but we have one additional file, namely \textit{indexsetter.h}, in which the class \texttt{ActiveFEIndexSetter} is defined.
 
-This example is an extension the previous one.
-We solve an stationary FSI problem either with 
-INH material (see Problem definition before) 
-or St. Venant Kirchhoff material STVK: \index{structure} \index{structure!compressible St.Venant-Kirchhoff (STVK) material}
-\begin{Problem}[Compressible Saint Venant-Kirchhoff, Lagrangian framework]
-  \label{prob:stvk:lagrange}
-  Find $\{\hat v_f,\hat u_f\} \in \{ \hat v_f^D + \hat V\} 
-  \times \{ \hat u_f^D + \hat V\}$, 
-  such that 
-  \begin{eqnarray}
-    \begin{aligned}   
-      (\hat J_s \hat\sigma_s \hat F_s^{-T},\hat\nabla\hat\phi^v)_{\hat\Omega_s}
-      &= \langle\hat J_s\hat n_s\cdot \hat g_s^\sigma\hat
-      F_s^{-T},\hat\phi^v  \rangle_{\hat\Gamma_i}
-      &&\quad\forall\hat \phi^v\in\hat V_s\\
-      %%%%%%%%%%%%%%%% 
-      (\hat v_s,\hat\phi^u)_{\hat\Omega_s}&=0
-      &&\quad\forall\hat \phi^u\in\hat V_s    ,\\
-    \end{aligned}
-  \end{eqnarray}
-  where $\rho_s$ is the density of the structure, $\mu_s$ and $\lambda_s$ 
-  the Lam\'e coefficients, $\hat n_s$ the outer normal vector at $\hat\Gamma_i$, 
-  $\hat g_s^\sigma$ some forces on the interface.\index{Lam\'e coefficients} The properties of the STVK material
-  is specified by the constitutive law
-  \begin{eqnarray}
-    \hat\sigma_s:=\hat J^{-1} \hat F (\lambda_s (tr\hat E) I + 2\mu_s \hat E)\hat F^{-T}.
-  \end{eqnarray}
-\end{Problem}
-Often, the elasticity properties of structure materials is characterized by 
-Poisson's ratio $\nu_s$ ($\nu_s < \frac{1}{2}$ \index{Poisson's ratio} for 
-compressible materials) and the Young modulus $E$ \index{Young modulus}. The 
-relationship to the Lam\'e coefficients $\mu_s$ and $\lambda_s$ is
-given by:
-\begin{eqnarray}
-\nu_s = \frac{\lambda_s}{2(\lambda_s + \mu_s)}, \qquad 
-E = \frac{\mu_s (\lambda_s + 2\mu_s)}{(\lambda_s + \mu_s )} . 
-\end{eqnarray}
-The whole equation system is solved on the benchmark configuration domain. For 
-details on parameters and geometry, we refer to the 
-numerical FSI benchmark \index{fluid-structure interaction (FSI)} \index{fluid-structure interaction (FSI)!FSI benchmark} proposal from Hron and Turek [2006].
+In the hp-framework we have a stack of finite elements (a \texttt{hp::FECollection}) given.
+We assign each cell of the triangulation an fe-index which determines which finite element we use on this cell.
+The \texttt{ActiveFEIndexSetter} class manages these indices, see there for more information.
 
-The code is established by computing the stationary FSI benchmark example FSI 1 with 
-the following values of interest: $x$-displacement, $y$-displacement, drag, and lift.\index{functional} \index{functional!drag} \index{functional!lift} \index{functional!deflection}
-
-\vspace{0.2cm}
-
-\textbf{Program description}
-
-\vspace{0.2cm}
-
-Compared to the two former examples, there are some differences which we will briefly discuss in the following.
-First of all, the problem is nonlinear in contrast to the former ones.
-We work on a different domain (given in the \textit{benchfst0100tw.inp} file), namely a channel with a cylinder put at half height near the inflow boundary; further \textit{.inp} files yield the possibility to vary the domain.\\
-Furthermore, in the \textit{dope.prm} parameter file there are two additional subsections which are added only for the current problem.
-From the denotation of these subsections one can immediately see where in the code the parameters are used.\\
-As we want to compute certain benchmark quantities, we have to regard corresponding functionals in the \textit{functionals.h} file.
-The pressure at a point as well as the displacement in $x$- and $y$-directions are point values; furthermore we implement the drag and lift functionals (for which we need the additionally defined problem parameters).\\
-As before, we build up the cell and boundary equations and matrices in the \textit{localpde.h} file.
-Apart from using the additionally defined problem parameters and modelling compressible STVK material instead of INH material (which leads to changes in the weak formulation of the equations), there are no major differences to the corresponding file in the last example.\\
-In the \textit{main.cc} file, we have to include additional header files from the deal.II library concerning error estimation and grid refinement.
-Further on, everything is pretty much the same as in the last example, but we have to use the \texttt{SetBoundaryFunctionalColors} function of the \texttt{PDEProblemContainer} class to be able to compute drag and lift in the respective functional classes in \textit{functionals.h}.\\
-The main innovation in contrast to the preceding examples is the refinement of the grid combined with a simple error estimator given in the deal.II \texttt{KellyErrorEstimator} class.
-If we look at the output of our program, everything is computed several times (once on each refinement level).
-Furthermore, we see that several Newton steps are needed on each refinement level; this is due to the nonlinearity of the current problem.
+The changes in \textit{main.cc} are also minimal and are highlighted in the source code. Obviously, we use \texttt{FECollection} and \texttt{QCollection} as well as a different DoFHandler.
+\begin{verbatim}
+  	#define DOFHANDLER hp::DoFHandler
+  	#define FE hp::FECollection
+  	...
+  	typedef hp::QCollection<DIM> QUADRATURE;
+  typedef hp::QCollection<DIM - 1> FACEQUADRATURE;
+\end{verbatim}
+Apart from that we have only to tell the space time handler the distribution of the finite element indices:
+\begin{verbatim}
+  ActiveFEIndexSetter<2> indexsetter(pr);
+  STH DOFH(triangulation, state_fe_collection, indexsetter);
+\end{verbatim}

### 2013-06-07 Examples/PDE/StatPDE/Example8/content.tex
Preparation for moving the examples to a new ordering.

--- 2012-03-15 Examples/PDE/StatPDE/Example8/content.tex
@@
-Laplace in 3D
-on locally refined grids with Kelly estimator
-with iterative solver (CG without preconditioning)
+\subsubsection{General problem description}
+
+Similar to the previous example, we consider the following benchmark problem from plasticity theory:
+\begin{equation}
+   (\Pi(\sigma(u)),\varepsilon(\varphi)) = (g,\varphi)_{\Gamma_N}.
+\end{equation}
+Here $\tilde{\Omega}$ is again the quadratic domain with a circular hole around the center cut out.
+Again, we restrict our actual computational domain $\Omega$ to the upper left quarter of $\tilde{\Omega}$ for reasons of symmetry.\\
+We use the symmetric strain tensor $\varepsilon(v) := \frac{1}{2}(\nabla v + \nabla v^T)$, and the symmetric stress tensor $\sigma$ is defined as
+\begin{equation*}
+   \sigma(v) := 2\mu \varepsilon(v)^D + \kappa \cdot tr (\varepsilon(v)) I
+\end{equation*}
+where $\tau^D$ is the deviatoric part of a tensor $\tau$, in two dimensions defined as
+\begin{equation*}
+   \tau^D := \tau - \frac{1}{2} tr(\tau) I.
+\end{equation*}
+The main difference with respect to the elastic case is the projection operator $\Pi$ in equation $(1)$. It is defined as follows:
+\begin{equation*}
+\Pi(\tau) = \left\{
+            \begin{array}{lr}
+            \tau & |\tau^D| \leq \sigma_0\\
+            \sigma_0 |\tau^D|^{-1} \tau^D + \frac{1}{2} tr(\tau) I & |\tau^D| > \sigma_0\\ 
+            \end{array}
+            \right. 
+\end{equation*}
+In our computations, we choose $\sigma_0 = \sqrt{\frac{2}{3}}\cdot 450$, and the above parameters $\mu$ and $\kappa$ as $\mu = 80193.800283$ resp.
+$\kappa = 190937.589172$.
+The corner points of our computational domain are the same as before, and the boundary conditions are not altered, either.\\
+The goal of our computations is to detect a subdomain in $\Omega$ where plastic behaviour occurs (compare \textit{E.
+Stein (editor), Error-controlled Adaptive Finite Elements in Solid Mechanics, Wiley (2003), pp.
+386 - 389}).
+This subdomain depends on the righthand side $g$ in equation $(1)$ which we write as $g = \lambda \cdot p$ with $p = 100$ and $\lambda \in [1.5;4.5]$.\\
+
+\subsubsection{Program description}
+
+The code of the current example is nearly identical to the code of the previous one.
+The only difference worth mentioning is the change of the equations which leads to different implementations of the \texttt{CellEquation, CellMatrix} and \texttt{BoundaryEquations} functions in \textit{localpde.h}.\\
+Furthermore, the elasticity equations solved in the last example are linear, whereas the plasticity equations are nonlinear; this difference is evident also from the output (here, we need several Newton steps until convergence).\\
+The functionals that appear in the output yield additional information and are not required in the above problem setting.
+The subdomain with plastic behaviour we want to detect can be visualized from the \textit{.vtk} files written to the \texttt{Results/Mesh} subfolders.

### 2013-07-04 Examples/PDE/InstatPDE/Example6/content.tex
First version of Biot-Lame-Navier problem

\subsubsection{General problem description}

In the present example, we solve the quasi-static 
Biot-Lame-Navier problem.

For literture on this field, we refer to 
Rujie Liu, Mary Wheeler and co-workers from UT Austin.

### 2013-07-04 Examples/PDE/InstatPDE/Example6/functionals.h

@@ -1,0 +1,149 @@
+/**
+*
+* Copyright (C) 2012 by the DOpElib authors
+*
+* This file is part of DOpElib
+*
+* DOpElib is free software: you can redistribute it
+* and/or modify it under the terms of the GNU General Public
+* License as published by the Free Software Foundation, either
+* version 3 of the License, or (at your option) any later
+* version.
+*
+* DOpElib is distributed in the hope that it will be
+* useful, but WITHOUT ANY WARRANTY; without even the implied
+* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+* PURPOSE.  See the GNU General Public License for more
+* details.
+*
+* Please refer to the file LICENSE.TXT included in this distribution
+* for further information on this license.
+*
+**/
+
+#ifndef _LOCALFunctionalS_
+#define _LOCALFunctionalS_
+
+#include "pdeinterface.h"
+
+using namespace std;
+using namespace dealii;
+using namespace DOpE;
+
+
+// pressure
+/****************************************************************************************/
+
+template<template<int, int> class DH, typename VECTOR, int dopedim, int dealdim>
+  class LocalPointFunctionalP1 : public FunctionalInterface<CellDataContainer,FaceDataContainer,DH, VECTOR, dopedim,dealdim>
+  {
+  private:
+    mutable double time;
+
+  public:
+
+    void SetTime(double t) const
+    {
+      time = t;
+    }
+
+    bool NeedTime() const
+    {
+      return true;
+    }
+
+    double PointValue(const DOpEWrapper::DoFHandler<dopedim, DH > & /*control_dof_handler*/,
+		    const DOpEWrapper::DoFHandler<dealdim, DH > & state_dof_handler,
+		      const std::map<std::string, const dealii::Vector<double>* > &/*param_values*/,
+		    const std::map<std::string, const VECTOR* > &domain_values)
+  {
+
+    Point<2> p1(0.0,0.0);
+
+    typename map<string, const VECTOR* >::const_iterator it = domain_values.find("state");
+    Vector<double> tmp_vector(3);
+
+
+    VectorTools::point_value (state_dof_handler, *(it->second), p1, tmp_vector);
+    double p1_value = tmp_vector(2);
+
+    // pressure analysis
+    return (p1_value);
+
+
+  }
+
+  string GetType() const
+  {
+    return "point timelocal";
+    // 1) point domain boundary face
[…]

### 2013-07-23 Examples/PDE/StatPDE/Example12/content.tex
New Example: Shows how to use real vector-valued elements such as Raviart-Thomas, and gives an example on a user given linear solver

\subsubsection{General problem description}
This example shows the use of the Raviart-Thomas (RT) element in \texttt{DOpElib}. The example 
is taken from \texttt{dealii} step-20 and shows how this step can be implemented in \texttt{DOpElib}.

Most things are identical to all the other programs. However there is a subtle difference in 
\texttt{localpde.h}: Although the element used has 3 components (two for the RT-Element and 1 for the pressure)
the block_component vector mapping blocks to components has only 2 entries. 1 for each finite element used!

The second difference is that we have to initialize the mapping explicitly. This is due to the fact that the 
default 
\begin{verbatim}
  DOpEWrapper::Mapping<DIM,DOFHANDLER> mapping(1,false);
\end{verbatim}
is not working with the RT-element and leads to elements on which the divergence is NaN.

An additional feature of this example is that for the solution of the PDE in mixed form we are using the 
schur complement solver provided in \texttt{dealii} step-20. Hence this example shows how simple it is to 
use self-made linear solvers within the \texttt{DOpElib} framework.

### 2013-07-23 Examples/PDE/StatPDE/Example12/main.cc

@@ -1,0 +1,196 @@
+/**
+ *
+ * Copyright (C) 2012 by the DOpElib authors
+ *
+ * This file is part of DOpElib
+ *
+ * DOpElib is free software: you can redistribute it
+ * and/or modify it under the terms of the GNU General Public
+ * License as published by the Free Software Foundation, either
+ * version 3 of the License, or (at your option) any later
+ * version.
+ *
+ * DOpElib is distributed in the hope that it will be
+ * useful, but WITHOUT ANY WARRANTY; without even the implied
+ * warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+ * PURPOSE.  See the GNU General Public License for more
+ * details.
+ *
+ * Please refer to the file LICENSE.TXT included in this distribution
+ * for further information on this license.
+ *
+ **/
+
+#include "pdeproblemcontainer.h"
+#include "statpdeproblem.h"
+#include "newtonsolver.h"
+#include "integrator.h"
+#include "parameterreader.h"
+#include "mol_statespacetimehandler.h"
+#include "simpledirichletdata.h"
+#include "sparsitymaker.h"
+#include "integratordatacontainer.h"
+#include "preconditioner_wrapper.h"
+
+#include <iostream>
+#include <fstream>
+
+#include <grid/tria.h>
+#include <grid/grid_in.h>
+#include <dofs/dof_handler.h>
+#include <grid/grid_generator.h>
+#include <grid/tria_boundary_lib.h>
+#include <deal.II/grid/grid_out.h>
+#include <fe/fe_dgq.h>
+#include <dofs/dof_tools.h>
+#include <base/quadrature_lib.h>
+#include <base/function.h>
+#include <numerics/vector_tools.h>
+#include <fe/fe_raviart_thomas.h>
+
+#include "localpde.h"
+#include "localfunctional.h"
+#include "schurlinearsolver.h"
+
+using namespace std;
+using namespace dealii;
+using namespace DOpE;
+
+const static int DIM = 2;
+
+#define DOFHANDLER DoFHandler
+#define FE FESystem
+#define CDC CellDataContainer
+#define FDC FaceDataContainer
+
+typedef QGauss<DIM> QUADRATURE;
+typedef QGauss<DIM - 1> FACEQUADRATURE;
[…]

### 2013-07-23 Examples/PDE/StatPDE/Example12/content.tex
Typo removed

@@
 \texttt{localpde.h}: Although the element used has 3 components (two for the RT-Element and 1 for the pressure)
-the block_component vector mapping blocks to components has only 2 entries. 1 for each finite element used!
+the block\underline{ }component vector mapping blocks to components has only 2 entries. 1 for each finite element used!
 

### 2013-11-10 Examples/PDE/InstatPDE/Example2/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
Content tex in PDE Instat Examples aufgeraeumt

@@
-\textbf{General problem description}
+\subsubsection{General problem description}
 
 In the present example, we solve a 
-nonstationary fluid-structure interaction problem. 
+nonstationary fluid-structure interaction problem in 
+arbitrary Lagrangian-Eulerian (ALE) coordinates. 
 The underlying equations are stated in the following:
@@
   \end{eqnarray*}  
-  with $\hat\rho_f$, $\hat\rho_s$, $\nu_f$, $\mu_s$, $\lambda_s$, $\hat F$, and $\hat J$. 
-  The stress tensors for the fluid and structure are
-implemented in  
-  $\hat\sigma_f$, $\hat\sigma_s$, and $\hat\sigma_g$ 
+  with the densities $\hat\rho_f$ and $\hat\rho_s$, 
+the viscosity $\nu_f$, the Lam\'e parameters 
+$\mu_s$, $\lambda_s$ and the deformation gradient $\hat F$, and its
+determinant $\hat J$. The stress tensors for the fluid and structure are
+implemented by 
+\[
+\hat\sigma_f = -\hat pI + \hat\rho_f\nu_f (\hat\nabla\hat v \hat F^{-1}
+  + \hat F^{-T} \hat\nabla\hat v^T)
+\], 
+and 
+\[
+\hat\sigma_s = \hat F (\lambda_s \text{tr}\hat E I + 2\mu_s \hat E)
+\]
+with $\hat E = \frac{1}{2}(\hat F^T \hat F - I)$.
+The mesh deformation
+is of harmonic type and given by 
+\[
+\hat\sigma_g = \alpha(x) \hat\nabla\hat u,
+\] 
+with a specific parameter $\alpha(x)$ (motivation by Stein/Tezduyar/Benney 2003).
 \end{Problem}
+The ALE approach belongs to interface-tracking methods in which 
+the mesh is moved such that it fits in all time steps with 
+the FSI-interface. However, this leads to 
+a degeneration of the ALE map. Methods to circument such as 
+degeneration as long as possible are re-meshing techniques or 
+to use a biharmonic mesh motion technique.
+
+\newpage
 
@@
 
+The major difference to the first nonstationary program
+is the introduction of the 
+\begin{verbatim}
+void CellTimeEquationExplicit (...) 
+\end{verbatim}
+to write all time derivative terms explicitely:
+\begin{equation*}
+(v^n - v^{n-1}, \phi)_{\Omega}.
+\end{equation*}
+This behavior is useful since (as shown in the above equations)
+other solutions variables have to be considered
+around $\partial_t v$ such as $J:=J(u)$. 
+The same holds for the corresponding matrix part.
+
 \vspace{0.2cm}

### 2013-11-10 Examples/PDE/InstatPDE/Example6/content.tex
Update description
Content tex in PDE Instat Examples aufgeraeumt

@@
 \subsubsection{General problem description}
+The modeling part of this example is based on the coupled
+Biot-Lam\'e-Navier system, 
+which is a standard model in subsurface modeling. 
+Here, a reservoir (the
+pay-zone) is modeled as a poroelastic medium with the help of Biot's
+equations. A surrounding medium (the non-pay zone) is modeled as a static
+elastic solid. 
+In fact, the
+Biot system is a multi-scale problem which 
+is identified on the micro-scale as 
+a fluid-structure interaction problem (details on the 
+interface law are found in Mikeli\'c and Wheeler (2011). 
+This system 
+is specifically suited for applications 
+in subsurface modeling for the poroelastic part, the 
+so-called pay-zone.  
+On the other hand, surrounding rock 
+(the non-pay zone) 
+is modeled with the help of linear elasticity (Ciarlet 1984). 
+Therefore, the final configuration 
+belongs to a multiphysics problem in non-overlapping domains.
+The nonstationary coupled system for the state is formulated 
+within a variational monolithically-coupled framework, 
+which is known 
+to be more robust than partitioned solutions algorithms.
+Its discretization 
+is carried out with help of the Rothe method in which we first 
+discretize in time and then in space. 
+The configuration 
+is based on the augmented Mandel problem which 
+shows the important Mandel-Cryer effect: First increasing pressure
+and then decreasing pressure in time while applying some 
+traction force on the top boundary.
 
-In the present example, we solve the quasi-static 
-Biot-Lame-Navier problem.
+We begin by describing the setting for a pure poroelastic setting, 
+the so-called pay-zone.
+Let $\Omega_B$ the domain of interest and $\partial\Omega_B$ its boundary
+with the partition:
+\begin{equation*}
+\partial\Omega_B = \Gamma_u \cup \Gamma_t = \Gamma_p \cup \Gamma_f,
+\end{equation*}
+where $\Gamma_u$ denotes the displacement boundary (Dirichlet), $\Gamma_t$ the 
+total stress or traction boundary (Neumann), $\Gamma_p$ the pore 
+pressure boundary (Dirichlet), and $\Gamma_f$ the fluid flux boundary
+(Neumann). Concretely, we have for a material 
+with the displacement variable $u$ and its 
+Cauchy stress tensor $\sigma$:
+\begin{align*}
+u &= \bar{u} \quad \text{on } \Gamma_u, \\
+\sigma n &= \bar{t} \quad \text{on } \Gamma_t, 
+\end{align*}
+for given $\bar{u}$ and $\bar{t}$, and the 
+normal vector $n$. 
+For the pressure with the permeability tensor 
+$K$ and fluid's viscosity $\eta_f$, we have the conditions:
+\begin{align*}
+p &= \bar{p} \quad \text{on } \Gamma_p, \\
+- \frac{K}{\eta_f} \Bigl( \nabla p - \rho_f g\Bigr) \cdot n &= \bar{q}\quad \text{on } \Gamma_f,
+\end{align*}
+for given $\bar{p}$ and $\bar{q}$; and the density $\rho_f$ and the gravity $g$.
+For the initial conditions at time $\tau=0$, we prescribe
+\begin{align*}
+p(\tau =0) = p_0 \quad \text{in } \Omega_B,\\
+\sigma(\tau =0) = \sigma_0 \quad \text{in } \Omega_B,
+\end{align*}
+In this case of this extension
+(if $\Omega_B$ is totally embedded in $\Omega_S$)
+the boundary conditions on $\partial\Omega_B$ reduce 
+to interface conditions $\partial\Omega_B :=\Gamma_i =
+\Omega_B \cap \Omega_S$.
+Let $I:=[0,T]$ denote the time interval.
+The underlying Biot system reads:
+Find the pressure $p_{\text{B}}$ and displacement 
+$u_{\text{B}}$ such that
+\begin{align*}
+\partial_t (c_B p_{\text{B}} + \alpha_B \text{div}u_{\text{B}})\\ -
+\frac{1}{\eta_f} \text{div}\, K 
+(\nabla p_{\text{B}} - \rho_f g) &= q \quad \text{in }\Omega_{B} \times I, \\
+- \text{div}\bigl( \sigma_B (u) \bigr)
++ \alpha_B\nabla p_{\text{B}} &= f_{\text{B}} 
+\quad \text{in }\Omega_{\text{B}} \times I,
+\end{align*}
+with
+\[
+\sigma_B(u_B) := \mu_B(\nabla u_{\text{B}} + \nabla u_{\text{B}}^T) +  
+\lambda_B\text{div}\, u_{\text{B}}I,
+\]
+and the coefficients $c_B\geq 0$, 
+the Biot-Willis constant $\alpha_B \in [0,1]$,
+(in fact, this constant relates to the amount of 
+coupling between the flow part and the elastic part) 
+and the permeability tensor $K$, 
+fluid's viscosity and its density 
+$\eta_f$ and $\rho_f$, gravity $g$ and a volume 
+source term $q$ (i.e., usually, wells for oil production and
+fluid injection). In the second equation,
+the Lam\'e coefficients are denoted 
+by $\lambda_B >0$ and $\mu_B >0$ 
+and $f_{\text{B}}$ is a volume force.
 
-For literture on this field, we refer to 
-Rujie Liu, Mary Wheeler and co-workers from UT Austin.
+The velocity $v_{\text{B}}$ 
+in the porous medium is obtained 
+with the help of Darcy's law (Darcy 1856)
+and the Darcy equations which are obtained 
+through homogenization of Stokes's equations
+It holds:
+\begin{equation*}
+v_{\text{B}} = -\frac{1}{\eta_f} K(\nabla p_{\text{B}} - \rho_f g).
+\end{equation*}
+Usually the non-pay zone is described 
+in terms of linear elasticity: Find 
+a displacement $u_S$ such that
+\begin{align*}
+- \text{div}\bigl( \sigma_S (u_S)  \bigr) = f_S \quad \text{in }\Omega_{\text{S}} \times I,
+\end{align*}
+with
+\begin{align*}
+\sigma_S (u_S) := \mu_S(\nabla u_{\text{S}} + \nabla u_{\text{S}}^T) +  
+\lambda_S\text{div}\, u_{\text{S}}I, 
+\end{align*}
+with the Lam\'e coefficients $\mu_S$ and $\lambda_S$
+and a volume force $f_S$. On the boundary 
+$\partial\Omega_S := \Gamma_D \cup \Gamma_N$, the 
+conditions
+\begin{equation*}
+u_S = \bar{u}_S \quad\text{on } \Gamma_D , \quad \sigma_S (u_S) n_S = \bar{t}_S \quad\text{on } \Gamma_N,
+\end{equation*}
+are prescribed with given $\bar{u}_S$ and $\bar{t}_S$.
+
+It finally remains to describe the interface 
+conditions on $\Gamma_{i}$ between the two sub-systems:
+\begin{equation}
+\label{interface_conditions}
+\begin{aligned}
+u_{\text{B}} &= u_S ,   \\
+\sigma_{B}(u_B)n_B - \sigma_S(u_S)n_S &= 
+\alpha p_{\text{B}} n_B  , \\
+-\frac{1}{\eta_f} K(\nabla p_{\text{B}} - \rho_f g) \cdot n_S &= 0. 
+\end{aligned} 
+\end{equation}
+It is important to notice that the second condition 
+in \eqref{interface_conditions}, requires careful 
+implementation on the interface, which is 
+carried out with the help of deal.II's FE Nothing element.

### 2013-11-15 Examples/Documentation/introduction.tex
contents tex in opt problems

@@
   \item \texttt{pdeproblemcontainer.h} Is used to describe  stationary PDE problems.
-  \item \texttt{instatpdeproblemcontainer.h} {\bf TODO is still missing but will be used for nonstationary problems.} This will be implemented once we have nonstationary optimization problems running to avoid error duplication in the coding process.
+  \item \texttt{instatpdeproblemcontainer.h} 
+%{\bf TODO is still missing but will be used for nonstationary problems.} 
+This will be implemented once we have nonstationary optimization problems running to avoid error duplication in the coding process.
   \item \texttt{optproblemcontainer.h} Is used to describe  OPT problems governed by 
@@
   an interface to the dimension independent functionality of all space time dof handlers.
-  {\bf TODO: Beispiele geben}
+  %{\bf TODO: Beispiele geben}
 \item \texttt{basic/statespacetimehandler.h} Another intermediate interface class which adds 

### 2013-11-15 Examples/OPT/StatPDE/Example3/content.tex
Changed DOpE to DOpElib in the documentation, finished overhaul of PDEStatPDE-Examples
Example 3 laeuft
contents tex in opt problems

@@
-In this example, we solve a optrimization problems from 
+\subsubsection{General problem description}
+In this example, we solve a optimization problems from 
 fluid dynamcis. The configuration is similar to 
 the fluid optimization problem proposed by Roland Becker 
-in 2000 ``Mesh adaption for stationary flow control''.
+``Mesh adaption for stationary flow control'' (2000).
 
-We describe the underlying equations later...
+The configuration
+    comes from the original fluid benchmark problem
+    and has been modified to reduce drag around the
+    cylinder. To gain the solvability of
+    the optimization problem we add a quadratic
+    regularization term to the cost functional.
+
+\begin{figure}[h]
+\centering
+    {\includegraphics[width=7cm]{Documentation/visit_Nov_15_2013_0000_a.png}}
+    {\includegraphics[width=7cm]{Documentation/visit_Nov_15_2013_0001_a.png}}
+  \caption{Configuration of the cylinder-drag minimization problem. 
+By sucking out the fluid (right the $y$-velocity), 
+the force on the cylinder is reduced. At left, the $x$-velocity field 
+is shown. Behind the cylinder, almost no fluid goes from left to right, 
+which is shown in blue color.}
+  \label{ex_2_plots}
+\end{figure}
+
+The computational domain is denoted by $\Omega$ with the 
+boundary $\partial\Omega
+= \Gamma_in \cup \Gamma_out \cup \Gamma_Q \cup \Gamma_w$, 
+where $\Gamma_in$ and $\Gamma_out$ denote the inflow and outflow
+boundaries, respectively. On $\Gamma_in$, we prescribe a fixed parabolic 
+inflow profile. The part(s) $\Gamma_w$ denote the top and bottom 
+boundaries. Finally, $\Gamma_Q$, represent Neumann control boundaries. Here,
+we prescribe 
+\[
+\rho\nu \partial_n v - pn = qn \quad\text{on } \Gamma_Q
+\] 
+where $n$ denotes the outer unit normal to $\Gamma_Q$.
+
+The state equations are given by: Find $v$ and $p$ such that
+\begin{align*}
+- \nabla\cdot \sigma(v,p) + v\cdot \nabla v = 0,\\
+\nabla\cdot v = 0
+\end{align*}
+with $\sigma(v,p) = -pI + \rho\nu(\nabla v + \nabla v^T)$.
+
+\textbf{Remark:} Since, we use the 
+symmetric stress tensor, we need to subtract the nonsymmetric part 
+on the outflow boundary, related to the do-nothing condition.
+$\diamond$
+
+The target functional is considered as 
+\[
+k(v,p) = \int_{\Gamma_O} n\cdot \sigma(v,p)\cdot d \, ds,
+\]
+where $\Gamma_O$ denotes the cylinder boundary, and $d$ is a vector in the
+direction
+of the mean flow. For theoretical and numerical reasons, this functional 
+needs to be regularized, including the control variable $q$, such that
+\[
+K(v,p,q) = k(v,p) + \frac{\alpha}{2}||q - q_0||,
+\] 
+where $\alpha$ is the Tikhonov parameter and $q_0$ some 
+reference control.
+
+The rest of the progam is similar to the previous optimization problems where
+we formulate the state equation in a weak form $a(v,p)(\phi)$ such that the 
+final problem reads
+\[
+K(v,p,q) \rightarrow 0 \quad \text{s.t.} \quad a(v,p)(\phi) = 0.
+\]
+
+%\subsubsection{Program description}

### 2013-11-15 Examples/OPT/StatPDE/Example9/content.tex
contents tex in opt problems

@@
 the equations come from fluid-structure interaction (FSI). The idea
-is to extend the steady FSI benchmark problem (FSI 1, proposed by Hron/Turek) to an optimization 
+is to extend the steady FSI benchmark problem (FSI 1, proposed by Hron/Turek) 
+to an optimization 
 problem where the drag is minimized over the cylinder and the beam. 
+The setting is similar to Opt Example 3. In fact, the only 
+novel things are to attach an elastic beam at the cylinder and to 
+extend the equation to fluid-structure interaction (instead of 
+pure fluid as in Example 3).
 
-The example has many similarities to Opt Example 3.
+\begin{figure}[h]
+\centering
+    {\includegraphics[width=7cm]{Documentation/visit_fsi_1_opt_control_p3_0002_bearbeitet_Aug_15.png}}
+    {\includegraphics[width=7cm]{Documentation/visit_fsi_1_opt_control_p3_0005_bearbeitet_Aug_15.png}}
+  \caption{Configuration of the fsi cylinder-beam-drag minimization problem. 
+By sucking out the fluid, 
+the force on the cylinder is reduced. At left, the $x$-velocity field 
+is shown. In right figure, we the corresponding adjoint solution is shown.}
+  \label{ex_2_plots}
+\end{figure}
 
-Detailed description will follow later. %TODO
+The state equation system reads:
+\begin{Problem}[Stationary Fluid-Structure Interaction with STVK material]
+Let $q$ denote the control variable. Find $U:= \{\hat v,\hat p,\hat u\}$ such that
+  \begin{equation*}
+    \begin{aligned}
+      (\hat J \rho_f \hat F^{-1}\hat  v\cdot\hat\nabla \hat v,
+      \hat\phi^v)_{\hat\Omega_f}
+      + (\hat J\hat\sigma_f\hat F^{-T},\hat \nabla\hat\phi^v)_{\hat\Omega_f}\\
+      + (\hat J\hat\sigma_s\hat F^{-T},\hat \nabla\hat\phi^v)_{\hat\Omega_s}
+      + \langle q,\hat\phi^v\cdot \hat n\rangle_{\Gamma_Q}
+      &= 0&&\forall\hat\phi^v\in \hat V,\\
+      %%%%%%%%%%%% 
+      (\hat v,\hat\phi^u)_{\hat\Omega_s}
+      + (\alpha_u \hat \nabla \hat u,\hat \nabla\hat\phi^u)_{\hat\Omega_f}
+      &=0&&\forall\hat\phi^u\in \hat V,\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+      (\widehat{\text{div}}\,(\hat J\hat F^{-1}\hat
+      v_f),\hat\phi^p)_{\hat\Omega_f} 
+      + (\hat p,\hat \phi^p)_{\hat\Omega_s}
+      &=0&&\forall\hat \phi^p\in \hat L,
+    \end{aligned}
+  \end{equation*}  
+\end{Problem}
+
+The target functional is considered as 
+\[
+k(U) = \int_{\hat\Gamma_O \cup \hat\Gamma_i} \hat n\cdot \hat J\hat\sigma(v,p)\hat
+F^{-T} \cdot \hat d \, ds,
+\]
+where $\Gamma_O$ denotes the cylinder boundary and $\Gamma_i$ the 
+interface between fluid and solid, and $\hat d$ is a vector in the
+direction
+of the mean flow. Moreover, $\hat J$ and $\hat F$ denote the deformation 
+gradient and its determinant as well known in fluid-structure interaction.
+For theoretical and numerical reasons, this functional 
+needs to be regularized, including the control variable $q$, such that
+\[
+K(U,q) = k(U) + \frac{\alpha}{2}||q - q_0||,
+\] 
+where $\alpha$ is the Tikhonov parameter and $q_0$ some 
+reference control.
+
+The rest of the progam is similar to the previous optimization problems where
+we formulate the state equation in a weak form $a(v,p)(\phi)$ such that the 
+final problem reads
+\[
+K(U,q) \rightarrow 0 \quad \text{s.t.} \quad a(v,p)(\phi) = 0.
+\]

### 2013-11-18 Examples/OPT/StatPDE/Example9/content.tex

@@
 \[
-K(U,q) = k(U) + \frac{\alpha}{2}||q - q_0||,
+K(q,U) = k(U) + \frac{\alpha}{2}||q - q_0||,
 \] 
@@
 The rest of the progam is similar to the previous optimization problems where
-we formulate the state equation in a weak form $a(v,p)(\phi)$ such that the 
+we formulate the state equation in a weak form $a(q,U)(\phi)$ such that the 
 final problem reads
 \[
-K(U,q) \rightarrow 0 \quad \text{s.t.} \quad a(v,p)(\phi) = 0.
+K(q,U) \rightarrow \text{min} \quad \text{s.t.} \quad a(q,U)(\phi) = 0.
 \]

### 2013-12-08 Examples/PDE/InstatPDE/Example2/content.tex
Typo in Gleichungen behoben und Test neu. FSI 1 ist eingecheckt als Test. FSI 2 nonstationary laueft aber auch. Die parameter stehen in prm und main file

@@
 arbitrary Lagrangian-Eulerian (ALE) coordinates. 
+For mesh motion model is based on solving a biharmonic equation
+\cite{Wi11} rather than a linear-elastic model. 
 The underlying equations are stated in the following:
-\begin{Problem}[Variational fluid-structure interaction framework]
-  \label{eq:fsi:ale:harmonic}
-  Find $\{\hat v,\hat u,\hat p\} \in \{ \hat v^D + \hat {\cal V}^0\} 
-\times \{ \hat u^D + \hat {\cal V}^0 \}\times \hat {\cal L}$, 
-  such that $\hat v (0) = \hat v^0$ and $\hat u(0) = \hat u^0$, for almost all
-  time steps $t$, and
+\begin{Problem}[FSI with biharmonic mesh motion]
+  \label{eq:fsi:ale:biharmonic}
+  Find $\{\hat v,\hat u,\hat w,\hat p\} \in \{ \hat v^D + \hat {\cal V}^0\} 
+\times \{ \hat u^D + \hat {\cal V}^0 \}\times \hat {\cal V} \times \hat {\cal
+  L}$, such that $\hat v (0) = \hat v^0$ and $\hat u(0) = \hat u^0$, 
+for almost all time steps $t$, and
   \begin{eqnarray*}
     \begin{aligned}
-      (\hat J\hat\rho_f  \partial_t \hat v,\hat\psi^v)_{\hat\Omega_f}  
+      (\hat J \hat\rho_f \partial_t \hat v,\hat\psi^v)_{\hat\Omega_f}  
       +(\hat\rho_f \hat J  (\hat F^{-1}(\hat
@@
       - (\hat\rho_s\hat f_s, \hat\psi^v)_{\hat\Omega_s}
-      &=0&&\forall\hat\psi^v\in \hat V^0,
+      &= 0&&\forall\hat\psi^v\in \hat V^0,
       \\
-      %%%%%%%%%%%% 
-      (\partial_t\hat u-\hat v,\hat\psi^u)_{\hat\Omega_s} 
-      + 
-      (\hat\sigma_g ,\hat \nabla\hat\psi^u)_{\hat\Omega_f}
-      -\langle \hat\sigma_g \hat n_f ,\hat\psi^u\rangle_{\hat \Gamma_i}
-      &=0&&\forall\hat\psi^u\in \hat V^0,\\
+      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+      (\hat\alpha_u\hat w, \hat\psi^w)_{\hat\Omega_f}  + ( \hat\alpha_u\hat\nabla\hat
+      u, \hat\nabla\hat\psi^w)_{\hat\Omega_f} +( \hat\alpha_u\hat\nabla\hat
+      w, \hat\nabla\hat\psi^w)_{\hat\Omega_s}  
+%-\langle \hat\alpha_u \hat n_f\hat \nabla \hat u,\hat\psi^w\rangle_{\hat \Gamma_i} 
+      & = 0 &&\forall\hat\psi^w\in \hat V , \\
+      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
+      \hat\rho_s (\partial_t\hat u-\hat v,\hat\psi^u)_{\hat\Omega_s}
+      + (\hat\alpha_u \hat \nabla \hat w,\hat \nabla\hat\psi^u)_{\hat\Omega_f}
+      %-\langle \hat\alpha_w \hat n_f\hat \nabla \hat w,\hat\psi^u\rangle_{\hat \Gamma_i}
+      &=0&&\forall\hat\psi^u\in \hat V^0 ,\\    
       %%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
@@
 with $\hat E = \frac{1}{2}(\hat F^T \hat F - I)$.
-The mesh deformation
-is of harmonic type and given by 
-\[
-\hat\sigma_g = \alpha(x) \hat\nabla\hat u,
-\] 
-with a specific parameter $\alpha(x)$ (motivation by Stein/Tezduyar/Benney 2003).
 \end{Problem}
@@
 
-\newpage
-
-{\bf Code validation for fluid problems with ALE}
+{\bf Code validation for ALE-fluid and FSI problems}
 
@@
 The code is validated by the well-known 
-fluid- and FSI benchmark problems. Here, 
-the results have been summarized in the table below.
+fluid- and FSI benchmark problems \cite{SchaeTu96, HrTu06b}. 
+For the FSI test cases, 
+the basic configuration is 
+sketched in Fig.~\ref{configuration_csm_and_fsi_2D} 
+at which an elastic beam is attached 
+behind the rigid cylinder.
 
- \begin{table}[h]
-   \small
+\begin{figure}[h]
 \centering
-     \begin{tabular}{llllcccc}    
-       \hline
-       Configuration & Date         & Refinement & Time dis. & $k$ & $\Delta p$& $F_D$ & $F_L$   \\ \hline\hline
-       BFAC 2D-1     & Mar 10, 2010 & 2 global   & BE        & 1.0    & 0.117412  &       &         \\
-       BFAC 2D-1     & Mar 10, 2010 & 2 global   & FS        & 1.0    & 0.117412  &       &         \\
-       BFAC 2D-1     & Mar 10, 2010 & 2 global   & CN(k)     & 1.0    & 0.117412  &       &         \\ 
-       BFAC 2D-1     & Mar 10, 2010 & 2 global   & CN        & 1.0    & 0.117383  &       &         \\ 
-       BFAC 2D-2     & Mar 10, 2010 & 2 global   & CN(k)     & 1.0    & 2.31541   &       &         \\ 
-       BFAC 2D-2     & Mar 10, 2010 & 3 global   & CN(k)     & 1.0    & 2.32011   &       &         \\   
-       BFAC 2D-2     & Mar 12, 2010 & 3 global   & CN(k)     & 1.0e-2 & 2.50288   &       &         \\   
-     \end{tabular}
-  \end{table}
+\input{PDE/InstatPDE/Example2/pic_12}
+\caption{Flow around cylinder with elastic beam with 
+circle-center $C=(0.2,0.2)$ and radius $r=0.05$.}
+\label{configuration_csm_and_fsi_2D}
+\end{figure}
 
-\subsubsection{Code validation for FSI with ALE}
+The elastic beam has length
+$l=0.35m$ and height $h=0.02m$. The right lower end is positioned at 
+$(0.6m,0.19m)$, and
+the left end is attached to the circle. 
+Control points $A(t)$ (with $A(0) = (0.6,0.2)$) are fixed at the 
+trailing edge of the structure, measuring $x$- and $y$-deflections of the beam.
+Details 
+on parameters and evaluation functionals and other results 
+can be found in \cite{HrTu06b,BuSc06,Wi11}. 
+The time-stepping scheme can be 
+very easily chosen in the \texttt{main.cc} function by choosing an appropriate 
+time-stepping scheme as explained at the beginning of this manual
+and detailed in the previous example.
 
-The results are summarized below:
+The quantities of interest are evaluations of 
+$x$- and $y$ displacement at the point $A(0) = (0.6,0.2)$
+and the drag and lift forces acting on the cylinder and the elastic beam:
+\begin{align}
+\label{drag_lift_forces}
+(F_D , F_L) 
+= {\int_{S_f} \sigma_f \cdot n_f \, ds + 
+\int_{\Gamma_i} \sigma_s \cdot n_s \, ds},
+\end{align}
+where $S_{f}$ denotes the path over the cylinder in the fluid part and
+$\Gamma_i$ the interface between the elastic beam and the 
+fluid.
 
- \begin{table}[h]
-   \small
-   \centering
-     \begin{tabular}{llllccccc}    
-       \hline
-       Config. & Date & Refinement & Time dis. & $k$ &  $u_x(A) [\times 10^{-5}]$ &$u_y(A) [\times 10^{-4}]$ & $F_D$ & $F_L$  \\ \hline
-       FSI 1   & Mar 12, 2010 & 3 global & BE   & 1.0 & 8.2003  & 2.2732  & &        \\
-       FSI 1   & Apr 08, 2010 & 2 global & BE   & 1.0 & 8.2258  & 2.2813  & &        \\
-       FSI 1   & Apr 08, 2010 & 2 global & CN(k)& 0.5 & 8.2268  & 2.2813  & &        \\
-       FSI 1   & Apr 08, 2010 & 2 global & CN   & 0.5 & 8.2268  & 2.2813  & &        \\
-     \end{tabular}
-  \end{table}
-
-\vspace{0.2cm}
-
+% \begin{table}[h]
+%   \small
+%   \centering
+%     \begin{tabular}{llllcccc}    
+%       \hline
+%       Configuration & Date         & Refinement & Time dis. & $k$ & $\Delta p$& $F_D$ & $F_L$   \\ \hline\hline
+%       BFAC 2D-1     & Mar 10, 2010 & 2 global   & BE        & 1.0    & 0.117412  &       &         \\
+%       BFAC 2D-1     & Mar 10, 2010 & 2 global   & FS        & 1.0    & 0.117412  &       &         \\
+%       BFAC 2D-1     & Mar 10, 2010 & 2 global   & CN(k)     & 1.0    & 0.117412  &       &         \\ 
+%       BFAC 2D-1     & Mar 10, 2010 & 2 global   & CN        & 1.0    & 0.117383  &       &         \\ 
+%       BFAC 2D-2     & Mar 10, 2010 & 2 global   & CN(k)     & 1.0    & 2.31541   &       &         \\ 
+%       BFAC 2D-2     & Mar 10, 2010 & 3 global   & CN(k)     & 1.0    & 2.32011   &       &         \\   
+%       BFAC 2D-2     & Mar 12, 2010 & 3 global   & CN(k)     & 1.0e-2 & 2.50288   &       &         \\   
+%     \end{tabular}
+%  \end{table}
+% 
+%\subsubsection{Code validation for FSI with ALE}
+%
+%The results are summarized below:
+%
+% \begin{table}[h]
+%   \small
+%   \centering
+%     \begin{tabular}{llllccccc}    
+%       \hline
+%       Config. & Date & Refinement & Time dis. & $k$ &  $u_x(A) [\times 10^{-5}]$ &$u_y(A) [\times 10^{-4}]$ & $F_D$ & $F_L$  \\ \hline
+%       FSI 1   & Mar 12, 2010 & 3 global & BE   & 1.0 & 8.2003  & 2.2732  & &        \\
+%       FSI 1   & Apr 08, 2010 & 2 global & BE   & 1.0 & 8.2258  & 2.2813  & &        \\
+%       FSI 1   & Apr 08, 2010 & 2 global & CN(k)& 0.5 & 8.2268  & 2.2813  & &        \\
+%       FSI 1   & Apr 08, 2010 & 2 global & CN   & 0.5 & 8.2268  & 2.2813  & &        \\
+%     \end{tabular}
+%  \end{table}
+%
+%\vspace{0.2cm}
+%
 \textbf{Program description}

### 2013-12-09 Examples/Documentation/introduction.tex

@@
   in addition require the use of the multimesh version for integrator and 
-  face- as well as celldatacontainer.
+  face- as well as elementdatacontainer.
 \end{itemize}
@@
 \begin{itemize}
-\item \texttt{container/celldatacontainer.h} This object is used to pass data 
-  given on the current element (cell) of the mesh to the functions in PDE, functional, 
+\item \texttt{container/elementdatacontainer.h} This object is used to pass data 
+  given on the current element of the mesh to the functions in PDE, functional, 
   $\ldots$. 
@@
   $\ldots$. 
-\item \texttt{container/multimesh\underline{ }celldatacontainer.h} This is the same as the 
-  celldatacontainer, but it
+\item \texttt{container/multimesh\underline{ }elementdatacontainer.h} This is the same as the 
+  elementdatacontainer, but it
   is capable to handle data defined on an alternative triangulation.
@@
 \item \texttt{container/integratordatacontainer.h} This contains some data that 
-  should be passed to the integrator like quadrature formulas and the above cell and 
+  should be passed to the integrator like quadrature formulas and the above element and 
   face data container.
@@
 \begin{itemize}
-\item \texttt{interfaces/active\underline{ }fe\underline{ }index\underline{ }setter\underline{ }interface.h} In the case of hp finite elements, one has to specify for each cell which finite element to use.
+\item \texttt{interfaces/active\underline{ }fe\underline{ }index\underline{ }setter\underline{ }interface.h} In the case of hp finite elements, one has to specify for each element which finite element to use.
 This is done via this interface.

### 2013-12-09 Examples/OPT/InstatPDE/Example1/content.tex
Slight corrections in the nonstat opt problem. In general it seems okay

--- 2012-05-05 Examples/OPT/InstatPDE/Example1/content.tex
@@
 This example is a modified version of \texttt{PDE/InstatPDE/Example5}. 
-Again, we consider the heat equation, this time with an additional nonlinear term
+Again, we consider the heat equation, this time with an additional nonlinear
+term and most important a time derivative leading to a 
+nonstationary optimimization problem. The governing equation is
 \begin{equation*}
@@
 \end{align*}
-and $J(\overline{q},\overline{u}) = 0$.
+and $J(\overline{q},\overline{u}) = 0$.\\[4mm]
+
+The following new things differ from the PDE-Eamples and the optimization examples with stationary 
+PDEs:\\[2mm]
+In the file \texttt{main.cc} the SpaceTimeHandler now gets an additional argument.
+In this case \texttt{DOpEtypes::initial} which specifies that the control is entering in the 
+initial value.\\[2mm]
+In the file \texttt{localpde.h} we now have to specify the Methods \texttt{Init\underline{ }ElementRhs}
+and \texttt{Init\underline{ }ElementRhs\underline{ }Q}. They need to be adapted, since usualy the InitialValue for the 
+PDE is auto generated from a deal.ii function in the ProblemContainer. This Value is set in the 
+\texttt{Init\underline{ }ElementRhs} hence we need to change this function to use the control instead. 
+Correspondingly we need to implement the first derivative of this with respect to the control. 
+--We don't need the second derivative since it is zero anyways.--\\[2mm]
+
+{\bf Note:} In contrast to the ,,normal'' Element-terms in the PDE we assume in the program 
+that the \texttt{Init\underline{ }ElementEquation} is linear in the state and no other solution variables are present.
+Thus no derivatives of the \texttt{Init\underline{ }ElementEquation} need to be implemented.

### 2013-12-09 Examples/OPT/StatPDE/Example5/content.tex

@@
 \subsubsection{Program description}
-In order to use different meshes for control and state we need to use the multimesh variants of \texttt{CellDataContainer}, \texttt{FaceDataContainer} and \texttt{Integrator} and we have to choose a space time handler capable of managing multiple meshes, so we use
+In order to use different meshes for control and state we need to use the multimesh variants of \texttt{ElementDataContainer}, \texttt{FaceDataContainer} and \texttt{Integrator} and we have to choose a space time handler capable of managing multiple meshes, so we use
 

### 2013-12-09 Examples/PDE/InstatPDE/Example1/content.tex
density in Beispiel nun dabei und auch Beschreibung angepasst

@@
 \begin{align*}
-\partial_t v -\nabla\cdot (\nabla v + \nabla v^{T}) + (v\cdot\nabla)v  +\nabla p &= f \\
+\rho \partial_t v - \rho \nabla\cdot (\nabla v + \nabla v^{T}) + \rho (v\cdot\nabla)v  +\nabla p &= f \\
 \nabla\cdot v &= 0 
 \end{align*}
-on time interval $I=[0,T]$ and the domain 
-$\Omega = [-6,6]\times [0,2]$. For simplicity,  we set $f=0$.
+on time interval $I=[0,T]$ and the fluid benchmark domain (Schaefer/Turek 1996) 
+Here, we set $f=0$ and the flow is driven by a Dirichlet inflow condition.
 
@@
 \begin{verbatim}
-void CellEquation (..., double scale, double scale_ico)
+void ElementEquation (..., double scale, double scale_ico)
 \end{verbatim}
@@
 \begin{verbatim}
-void CellMatrix (..., double scale, double scale_ico)
+void ElementMatrix (..., double scale, double scale_ico)
 \end{verbatim}
@@
 \begin{verbatim}
-void CellTimeEquation (...) 
+void ElementTimeEquation (...) 
 \end{verbatim}
@@
 \begin{verbatim}
-void CellTimeEquationExplicit (...) 
+void ElementTimeEquationExplicit (...) 
 \end{verbatim}
@@
 Consequently, the directional derivatives of the 
-cell terms are implemented in the corresponding
+element terms are implemented in the corresponding
 matrix functions, i.e., 
 \begin{verbatim}
-void CellTimeMatrix (...), void CellTimeMatrixExplicit 
+void ElementTimeMatrix (...), void ElementTimeMatrixExplicit 
 \end{verbatim}

### 2013-12-09 Examples/PDE/StatPDE/Example1/content.tex

@@
 If we do not use one of the standard grids given in the deal.II library, we can read a grid from an input file.
-In our example, the domain $\Omega = [-6,6]\times [0,2]$ is given in the \textit{channel.inp} file, where all nodes, cells and boundary lines are listed explicitly and the boundary is divided into disjoint parts by attributing different colors to the boundary lines.\index{grid}\\
+In our example, the domain $\Omega = [-6,6]\times [0,2]$ is given in the \textit{channel.inp} file, where all nodes, elements and boundary lines are listed explicitly and the boundary is divided into disjoint parts by attributing different colors to the boundary lines.\index{grid}\\
 
@@
 All information about the PDE problem (in the optimal control case about the constraining PDE) is included in the \textit{localpde.h} file.
-In a class called \texttt{LocalPDE}, we build up the cell equation, the cell matrix and cell righthand side as well as the boundary equation, boundary matrix and boundary righthand side.
+In a class called \texttt{LocalPDE}, we build up the element equation, the element matrix and element righthand side as well as the boundary equation, boundary matrix and boundary righthand side.
 Later on, the integrator collects this local information and creates the global vectors and matrices.\\
@@
 The \textit{LocalPDE} is derived from a \texttt{PDEInterface} class.
-It comprises several functions which build up the cell and boundary equations, matrices and righthand sides.
+It comprises several functions which build up the element and boundary equations, matrices and righthand sides.
 The weak formulation of problem \eqref{eq:strong} with $f=0$ is
@@
 \end{equation*}
-The lefthand side is implemented in the \texttt{CellEquation} function, the righthand side is implemented in the \texttt{CellRightHandSide} function (which is unused in this example, because $f=0$).
+The lefthand side is implemented in the \texttt{ElementEquation} function, the righthand side is implemented in the \texttt{ElementRightHandSide} function (which is unused in this example, because $f=0$).
 \begin{remark}
@@
 \end{equation*}
-In the \texttt{CellMatrix} function, we implement the following matrix $A$ as representation of the derivative on the lefthand side:
+In the \texttt{ElementMatrix} function, we implement the following matrix $A$ as representation of the derivative on the lefthand side:
 \begin{equation*}
@@
 \end{equation*}
-with the number $N$ of the degrees of freedom. Similarly, the \texttt{CellEquation} contains the vector
+with the number $N$ of the degrees of freedom. Similarly, the \texttt{ElementEquation} contains the vector
 \begin{equation*}
@@
 \end{equation*}
-and the \texttt{CellRightHandSide} in the case $f \neq 0$ would contain a vector
+and the \texttt{ElementRightHandSide} in the case $f \neq 0$ would contain a vector
 \begin{equation*}

### 2013-12-09 Examples/PDE/StatPDE/Example10/content.tex

@@
 \subsubsection{General problem description}
-This example shows the use higher order mappings in DOpE. To this end, we solve a simple laplace equation
-\begin{align*}
--\Delta u &= -4 && \text{ in } \Omega
-\end{align*}
-with the analytical solution
-\begin{align*}
-u = {x^2 + y^2},
-\end{align*}
-and the Dirichlet Conditions on $u=1$, where the domain is given by a circle with radius 1 and the center located in the origin.
 
-We compute the $L^2$-norm of the error and, additionally, we evaluate a functional which does not depend on the solution at all.
-We integrate the constant $\frac 12$ once over the boundary of the domain. The result is an approximation of $\pi$.
+This example is an extension the previous one.
+We solve an stationary FSI problem either with 
+INH material (see Problem definition before) 
+or St. Venant Kirchhoff material STVK: \index{structure} \index{structure!compressible St.Venant-Kirchhoff (STVK) material}
+\begin{Problem}[Compressible Saint Venant-Kirchhoff, Lagrangian framework]
+  \label{prob:stvk:lagrange}
+  Find $\{\hat v_f,\hat u_f\} \in \{ \hat v_f^D + \hat V\} 
+  \times \{ \hat u_f^D + \hat V\}$, 
+  such that 
+  \begin{eqnarray}
+    \begin{aligned}   
+      (\hat J_s \hat\sigma_s \hat F_s^{-T},\hat\nabla\hat\phi^v)_{\hat\Omega_s}
+      &= \langle\hat J_s\hat n_s\cdot \hat g_s^\sigma\hat
+      F_s^{-T},\hat\phi^v  \rangle_{\hat\Gamma_i}
+      &&\quad\forall\hat \phi^v\in\hat V_s\\
+      %%%%%%%%%%%%%%%% 
+      (\hat v_s,\hat\phi^u)_{\hat\Omega_s}&=0
+      &&\quad\forall\hat \phi^u\in\hat V_s    ,\\
+    \end{aligned}
+  \end{eqnarray}
+  where $\rho_s$ is the density of the structure, $\mu_s$ and $\lambda_s$ 
+  the Lam\'e coefficients, $\hat n_s$ the outer normal vector at $\hat\Gamma_i$, 
+  $\hat g_s^\sigma$ some forces on the interface.\index{Lam\'e coefficients} The properties of the STVK material
+  is specified by the constitutive law
+  \begin{eqnarray}
+    \hat\sigma_s:=\hat J^{-1} \hat F (\lambda_s (tr\hat E) I + 2\mu_s \hat E)\hat F^{-T}.
+  \end{eqnarray}
+\end{Problem}
+Often, the elasticity properties of structure materials is characterized by 
+Poisson's ratio $\nu_s$ ($\nu_s < \frac{1}{2}$ \index{Poisson's ratio} for 
+compressible materials) and the Young modulus $E$ \index{Young modulus}. The 
+relationship to the Lam\'e coefficients $\mu_s$ and $\lambda_s$ is
+given by:
+\begin{eqnarray}
+\nu_s = \frac{\lambda_s}{2(\lambda_s + \mu_s)}, \qquad 
+E = \frac{\mu_s (\lambda_s + 2\mu_s)}{(\lambda_s + \mu_s )} . 
+\end{eqnarray}
+The whole equation system is solved on the benchmark configuration domain. For 
+details on parameters and geometry, we refer to the 
+numerical FSI benchmark \index{fluid-structure interaction (FSI)} \index{fluid-structure interaction (FSI)!FSI benchmark} proposal from Hron and Turek [2006].
 
-All this is standard and would not justify an additional example, however we solve the equation and the functional not one but two
-times. First with the standard Q1-mapping, the second time we use a higher order mapping. The exact order can be determined by the 
-parameter file, the preset is Q2-mapping. At the end, we gather the errors and convergence rates over some refinement cycles
-in a nice table and notice the higher order of convergence for the higher order mapping solutions. This is due to the fact that
-we can approximate the circular boundary much better by a quadratic mapping.
+The code is established by computing the stationary FSI benchmark example FSI 1 with 
+the following values of interest: $x$-displacement, $y$-displacement, drag, and lift.\index{functional} \index{functional!drag} \index{functional!lift} \index{functional!deflection}
+
 \subsubsection{Program description}
-In this section we want to focus on what you have to do if you want to enhance your existing code to use higher order mappings,
-which is actually pretty simple.
 
-You have to include the file
-\begin{verbatim}
-mapping_wrapper.h
-\end{verbatim}
-
-and create a mapping of the desired order by
-\begin{verbatim}
-  DOpEWrapper::Mapping<dim, DOFHANDLER> mapping(order_of_mapping);
-\end{verbatim}
-
-The last step is to give the mapping to the DoFHandler:
-\begin{verbatim}
-MethodOfLines_StateSpaceTimeHandler<FE, DOFHANDLER, SPARSITYPATTERN,VECTOR, 2> 
-	  DOFH(triangulation, mapping, state_fe);
-\end{verbatim}
-The rest of the program is as usual.
+Compared to the two former examples, there are some differences which we will briefly discuss in the following.
+First of all, the problem is nonlinear in contrast to the former ones.
+We work on a different domain (given in the \textit{benchfst0100tw.inp} file), namely a channel with a cylinder put at half height near the inflow boundary; further \textit{.inp} files yield the possibility to vary the domain.\\
+Furthermore, in the \textit{dope.prm} parameter file there are two additional subsections which are added only for the current problem.
+From the denotation of these subsections one can immediately see where in the code the parameters are used.\\
+As we want to compute certain benchmark quantities, we have to regard corresponding functionals in the \textit{functionals.h} file.
+The pressure at a point as well as the displacement in $x$- and $y$-directions are point values; furthermore we implement the drag and lift functionals (for which we need the additionally defined problem parameters).\\
+As before, we build up the element and boundary equations and matrices in the \textit{localpde.h} file.
+Apart from using the additionally defined problem parameters and modelling compressible STVK material instead of INH material (which leads to changes in the weak formulation of the equations), there are no major differences to the corresponding file in the last example.\\
+In the \textit{main.cc} file, we have to include additional header files from the deal.II library concerning error estimation and grid refinement.
+Further on, everything is pretty much the same as in the last example, but we have to use the \texttt{SetBoundaryFunctionalColors} function of the \texttt{PDEProblemContainer} class to be able to compute drag and lift in the respective functional classes in \textit{functionals.h}.\\
+The main innovation in contrast to the preceding examples is the refinement of the grid combined with a simple error estimator given in the deal.II \texttt{KellyErrorEstimator} class.
+If we look at the output of our program, everything is computed several times (once on each refinement level).
+Furthermore, we see that several Newton steps are needed on each refinement level; this is due to the nonlinearity of the current problem.

### 2013-12-09 Examples/PDE/StatPDE/Example5/content.tex

@@
-\textbf{General problem description}
+\subsubsection{General problem description}
+This example shows the use of the adaptive grid refinement and error estimation by the \textit{DWR method} (For a description of the method, see \cite{BR03}.) applied to the laplace equation 
+\begin{align*}
+-\Delta u &= f && \text{ in } \Omega
+\end{align*}
+with the analytical solution
+\begin{align*}
+u = \sin\left(\frac{\pi}{x^2 + y^2}\right),
+\end{align*}
+ the corresponding right hand side $f = -\Delta u$ and appropriate Dirichlet Conditions on $\partial \Omega$, where the domain is given by
+\begin{align*}
+\Omega = [-2,2]^2\setminus \overline{B}_{0.5}(0).
+\end{align*}
+We want to estimate the error in the following functional of interest
+$$\fullfunction{J}{H^1(\Omega)}{\mathbb{R}}{u}{\frac 1 {\abs{\Gamma}} \int_\Gamma u \,dx}$$
+where
+$\Gamma = \left\{(x,y)\in \mathbb R ^2\,\big | \,x=0,-2<y<0.5 \right \}$.
 
-\vspace{0.2cm}
+For this setting, we have the error representation
+\begin{align}\label{dwr.error_identity}
+J(e)=\sum_{K\in \mathbb T _h} \left\{(R_h, z-\psi_h)_K + (r_h, z-\psi_h)_{\partial_K}\right\}
+\end{align}
+with the error $e=u-u_h$, the Triangulation $\mathbb T_h$, the dual solution $z$, arbitraty function $\phi_h \in V_h$ (the ansatz space) and the element- and edge-residuals: 
+\begin{align}\label{dwr.element_residual}
+\eval[1]{R_h}_{K} = f+\Delta u_h
+\end{align}
+resp. 
+\begin{align}\label{dwr.edge_residual}
+\eval[1]{r_h}_{\Sigma} =
+\begin{cases}
+\frac 12 [\partial_n u_h] ,&\text{if } \Sigma \subset \partial_K \setminus \partial \Omega,\\
+0,& \text{if } \Sigma \subset\partial \Omega.
+\end{cases}
+\end{align}
 
-Similar to the previous example, we consider the following benchmark problem from plasticity theory:
-\begin{equation}
-   (\Pi(\sigma(u)),\varepsilon(\varphi)) = (g,\varphi)_{\Gamma_N}.
-\end{equation}
-Here $\tilde{\Omega}$ is again the quadratic domain with a circular hole around the center cut out.
-Again, we restrict our actual computational domain $\Omega$ to the upper left quarter of $\tilde{\Omega}$ for reasons of symmetry.\\
-We use the symmetric strain tensor $\varepsilon(v) := \frac{1}{2}(\nabla v + \nabla v^T)$, and the symmetric stress tensor $\sigma$ is defined as
-\begin{equation*}
-   \sigma(v) := 2\mu \varepsilon(v)^D + \kappa \cdot tr (\varepsilon(v)) I
-\end{equation*}
-where $\tau^D$ is the deviatoric part of a tensor $\tau$, in two dimensions defined as
-\begin{equation*}
-   \tau^D := \tau - \frac{1}{2} tr(\tau) I.
-\end{equation*}
-The main difference with respect to the elastic case is the projection operator $\Pi$ in equation $(1)$. It is defined as follows:
-\begin{equation*}
-\Pi(\tau) = \left\{
-            \begin{array}{lr}
-            \tau & |\tau^D| \leq \sigma_0\\
-            \sigma_0 |\tau^D|^{-1} \tau^D + \frac{1}{2} tr(\tau) I & |\tau^D| > \sigma_0\\ 
-            \end{array}
-            \right. 
-\end{equation*}
-In our computations, we choose $\sigma_0 = \sqrt{\frac{2}{3}}\cdot 450$, and the above parameters $\mu$ and $\kappa$ as $\mu = 80193.800283$ resp.
-$\kappa = 190937.589172$.
-The corner points of our computational domain are the same as before, and the boundary conditions are not altered, either.\\
-The goal of our computations is to detect a subdomain in $\Omega$ where plastic behaviour occurs (compare \textit{E.
-Stein (editor), Error-controlled Adaptive Finite Elements in Solid Mechanics, Wiley (2003), pp.
-386 - 389}).
-This subdomain depends on the righthand side $g$ in equation $(1)$ which we write as $g = \lambda \cdot p$ with $p = 100$ and $\lambda \in [1.5;4.5]$.\\
+It holds $J(u) \approx 0.441956231972232$.
+\subsubsection{Program description}
+In this section we want to focus on what you have to do if you want to enhance your existing code to use the \textit{DWR method}.
 
-\vspace{0.2cm}
+First, additionally to all the things one has to do when just solving the equation, we have to include the file
+\begin{verbatim}
+higher_order_dwrc.h
+\end{verbatim}
 
-\textbf{Program description}
+As we approximate the so called 'weights' $z-\phi_h$ in the error representation by a patchwise higher order interpolation of $z_h$ (the computed dual solution), we have to enforce patch-wise refinement of the grid by giving the flag
+\begin{verbatim}
+Triangulation<2>::MeshSmoothing::patch_level_1
+\end{verbatim}
+to the triangulation.
 
-\vspace{0.2cm}
+To be able to solve the adjoint equation for the error estimation one needs to implement some methods regarding the equation as well es the functional of interest:
+\begin{itemize}
+\item In {\verb pdeinterface.h }
+\begin{itemize}
+\item {\verb ElementEquation_U }: Weak form of the adjoint equation.
+\item {\verb ElementMatrix_T }: The FE matrix for the adjoint problem.
+\item {\verb FaceEquation_U }: This one is needed in this case here because we have a functional of interest that lives on faces. 
+\end{itemize}
+\item {\verb functionalinterface.h }
+\begin{itemize}
+\item {\verb FaceValue_U }: This is the right hand side of the adjoint equation.
+\end{itemize}
+\end{itemize}
+During the  evaluation of \eqref{dwr.error_identity}, the following methods are needed
+\begin{itemize}
+\item {\verb StrongElementResidual }: The element residual, see \eqref{dwr.element_residual}.
+\item {\verb StrongFaceResidual }: The terms in \eqref{dwr.error_identity} that lies in the interiour (i.e. the jumps).
+\item {\verb StrongBoundaryResidual }:The terms in \eqref{dwr.error_identity} that lies on the boundary (There are none in this case).
+\end{itemize}
 
-The code of the current example is nearly identical to the code of the previous one.
-The only difference worth mentioning is the change of the equations which leads to different implementations of the \texttt{CellEquation, CellMatrix} and \texttt{BoundaryEquations} functions in \textit{localpde.h}.\\
-Furthermore, the elasticity equations solved in the last example are linear, whereas the plasticity equations are nonlinear; this difference is evident also from the output (here, we need several Newton steps until convergence).\\
-The functionals that appear in the output yield additional information and are not required in the above problem setting.
-The subdomain with plastic behaviour we want to detect can be visualized from the \textit{.vtk} files written to the \texttt{Results/Mesh} subfolders.
+Note that in the above three functions we always apply the method \texttt{ResidualModifier} both to the 
+residual as well as to the jumps on the faces. This is done to assert that we can apply both a 
+DWR-error estimator where the residual should be multiplied with the computed weights (then this function 
+does not do anything) as well as Residual Type error estimator for the $L^2$ or $H^1$ norm where we need to calculate element wise norms of the residual and the jumps.
+Then this function calculates the appropriate
+local terms, e.g., the square of the residual scaled with appropriate powers of the local mesh size.
+
+After this, we tell the problem which functional we want to use for the error estimation, this is done via
+\begin{verbatim}
+P.SetFunctionalForErrorEstimation(LFF.GetName())
+\end{verbatim}
+where \texttt{P} is of type \texttt{PDEProblemContainer} and \texttt{LFF} is the desired functional of interest.
+
+The next thing we need is an object of the type
+\begin{verbatim}
+HigherOrderDWRContainer
+\end{verbatim}
+This container takes care of the computation of the weights.
+
+To build this, we need the following:
+\begin{itemize}
+\item {\verb DOFH_higher_order }: With some higher order Finite Elemnts and the already defined triangulation, we build this \texttt{SpaceTimeHandler}.
+This is needed because  we want to use the patch-wise higher order interpolation of the weights.
+\item {\verb idc_high }: A \texttt{IntegratorDataContainer} in which we put some (face)quadrature formulas for the evaluation of the error Identity.
+\item A string which indicates how we want to store the weight-vectors (here: { \verb "fullmem" }).
+\item \texttt{pr}: The \texttt{ParameterReader} which we have already defined.
+\item A enum of type \texttt{EETerms} that tells the container, which error terms we want to compute (primal error indicators vs. dual error indicators, see \cite{BR03}).  
+\end{itemize}
+The last preparation step is now to initialize the \texttt{DWRDataContainer} with the problem in use:
+\begin{verbatim}
+  solver.InitializeHigherOrderDWRC(dwrc);
+\end{verbatim}
+
+Succeding the solution of the state equation 
+\begin{verbatim}
+solver.ComputeReducedFunctionals(); 
+\end{verbatim}
+we compute the error indicators by calling 
+\begin{verbatim}
+solver.ComputeRefinementIndicators(dwrc);
+\end{verbatim}
+We can now get the error indicators (with signs!) out of \texttt{dwrc} by 
+\begin{verbatim}
+dwrc.GetErrorIndicators();
+\end{verbatim}
+With these indicators, we are now able to refine our grid adaptively (there are several mesh adaption strategies implemented, like 'RefineOptimized', 'RefineFixedNumber' or 'RefineFixedFraction')
+\begin{verbatim}
+DOFH.RefineSpace(RefineOptimized(optimized, error_ind));
+\end{verbatim}
+Note, that one has to take the norm of each entry in the vector of the error indicators before feeding them into the \texttt{RefineSpace} method.

### 2013-12-09 Examples/PDE/StatPDE/Example6/content.tex
Further work on the examples

--- 2012-03-15 Examples/PDE/StatPDE/Example6/content.tex
@@
-todo. short: this example shows how to use the constraintsmaker class by enforcing periodic boundary conditions in 2d.
+\subsubsection{General problem description}
+In this problem we solve the simple vector valued laplace equation in 3d
+on the unit square $\Omega=[0,1]^3$, i.e. in strong formulation we look for $u=(u_1, u_2, u_3)$ s.t.
+\begin{align*}
+-\Delta u =& f &&\text{in }\Omega.
+\end{align*}
+We set zero dirichlet values on $\partial \Gamma$ and choose $f=(1,1,1)$.
+\subsubsection{Program description}
+The PDE is discretized with Q3-elements on a series of locally refined grids (we use the \texttt{KellyErrorEstimator}).
+The algebraic equations are solved with  different iterative linear solvers acting on different vector and matrix-structures (i.e.
+we use \texttt{dealii::BlockVector} and \texttt{dealii::Vector} plus the appropriate matrix classes).
+
+To switch the linear solver is pretty easy since the newton solver has a template for the linear solver. Thus changing this template is all that is required.
+
+To change the structure of the vectors and matrices involves also only the change of some template parameters.
+Our example programs are mostly build such that only a change of some typedefs is required, i.e.
+one has to interchange the lines
+\begin{verbatim}
+	typedef SparseMatrix<double> MATRIX;
+	typedef SparsityPattern SPARSITYPATTERN;
+	typedef Vector<double> VECTOR;
+\end{verbatim}
+with
+\begin{verbatim}
+	typedef BlockSparseMatrix<double> MATRIX;
+	typedef BlockSparsityPattern SPARSITYPATTERN;
+	typedef BlockVector<double> VECTOR;
+\end{verbatim}
+to switch between the block and non-block structures.
+
+After solving the equation, we want to apply local mesh refinement.
+So first we extract with the help of the \texttt{SolutionExatractr}-class the vector \texttt{solution} representing the finite element solution
+\begin{verbatim}
+  SolutionExtractor<SSolver1, VECTORBLOCK> a1(solver1);
+ 	const StateVector<VECTORBLOCK> &gu1 = a1.GetU();
+ 	solution = gu1.GetSpacialVector();
+\end{verbatim}
+With this vector we estiamte the error via \texttt{KellyErrorEstimator} and get a vector holding the estimated error per element.
+After choosing a refinement criterion (see \textit{refinementcontainer.h}, we opt here for refining the top 20\% of the elements), we give the SpaceTimeHandler an object of type \texttt{RefinementContainer} which holds all the information needed for the local mesh refinement.
+This is done via the \texttt{RefineSpace} method.
+\begin{verbatim}
+        DOFH1.RefineSpace(
+               RefineFixedNumber(estimated_error_per_element, 0.2, 0.0));
+\end{verbatim}
+This method transfers our solution onto the new mesh.
+The transferred solution is then taken as the starting guess of the newton method in the next solution cycle.
+This is especially helpful for nonlinear problems.

### 2013-12-09 Examples/PDE/StatPDE/Example7/content.tex

--- 2013-05-09 Examples/PDE/StatPDE/Example7/content.tex
@@
 \subsubsection{General problem description}
-This is the simple vector valued laplace equation on the unitsquare with righthandside $(1,1)$ and zero dirichlet values. The classical example of a PDE.
+
+In this example we consider the following benchmark problem from elasticity theory: \index{stationary PDE} \index{stationary PDE!elasticity equations}
+\begin{equation}
+   (\sigma(u),\varepsilon(\varphi)) = (g,\varphi)_{\Gamma_N}.
+\end{equation}
+Here $\tilde{\Omega}$ is a quadratic domain with side length $200$ mm, where a circular hole with radius $10$ mm around the center is cut out.
+Using symmetries of the domain, we restrict our actual computational domain $\Omega$ to the upper left quarter of $\tilde{\Omega}$.\\
+In the above equation, $\varepsilon(v) := \frac{1}{2}(\nabla v + \nabla v^T)$ is the symmetric strain tensor, and \index{strain tensor}
+\begin{equation*}
+   \sigma(v) := 2\mu \varepsilon(v)^D + \kappa \cdot tr (\varepsilon(v)) I
+\end{equation*}
+denotes the symmetric stress tensor. Here $\tau^D$ is the deviatoric part of a tensor $\tau$, in two dimensions defined as \index{deviator}
+\begin{equation*}
+   \tau^D := \tau - \frac{1}{2} tr(\tau) I,
+\end{equation*}
+and the parameters $\mu$ and $\kappa$ are chosen as $\mu = 80193.800283$ resp. $\kappa = 190937.589172$.\\
+The corner points of our computational domain are in anticlockwise order: $(0,0), (90,0)$, $(100,10), (100,100)$ and $(0,100)$.
+We prescribe homogeneous Dirichlet boundary conditions in $y$-direction between $(0,0)$ and $(90,0)$ (lower boundary part), homogeneous Dirichlet boundary conditions in $x$-direction between $(100,10)$ and $(100,100)$ (right boundary part), and we interpret the righthand side of equation $(1)$ with $g = 450$ as a boundary condition between $(0,100)$ and $(100,100)$ (upper boundary part).\\
+The goal of our computations is to match the following functional reference values taken from \textit{E.
+Stein (editor), Error-controlled Adaptive Finite Elements in Solid Mechanics, Wiley (2003), pp.
+386 - 387}:
+\begin{table}[h]
+\centering
+\begin{tabular}{lccc}    
+\hline
+ Functional & $u_1$ at $(90,0)$ & $\sigma_{22}$ at $(90,0)$ & $u_2$ at $(100,100)$\\
+\hline 
+ Reference value & 0.021290 & 1388.732343 & 0.20951\\
+\hline
+\end{tabular}
+\end{table}
+\begin{table}[h]
+\centering
+\begin{tabular}{lcc}    
+\hline
+ Functional & $u_1$ at $(0,100)$ & $\int_{(100,100)}^{(0,100)} u_2$\\
+\hline 
+ Reference value & 0.076758 & 20.40344\\
+\hline
+\end{tabular}
+\end{table}
+
+\subsubsection{Program description}
+
+From the previous examples we know how to read a grid from an \textit{.inp} file.
+The grid of our current example comes from the above mentioned benchmark problem.
+The parameter file (\textit{dope.prm}) contains the same variables as in the first two examples.\\
+Apart from different point values of derivatives of the solution, we want to evaluate an integral over part of the boundary.
+\index{functional} \index{functional!boundary integral} This is newly implemented in \textit{functionals.h}.\\
+The (linear) elasticity equation is much simpler than the FSI problems we treated before, which can be seen in \textit{localpde.h} where the \texttt{ElementEquation} and \texttt{ElementMatrix} are implemented.
+In principle, everything is clear from the preceding examples.\\
+The \textit{main.cc} file does not contain any innovation, either. Here, we refine the grid globally instead of using an error estimator for local refinement.\\
+The output of the program reflects again the linearity of the problem (only one Newton step is needed for solution).\\

### 2013-12-09 Examples/PDE/StatPDE/Example8/content.tex

@@
 The code of the current example is nearly identical to the code of the previous one.
-The only difference worth mentioning is the change of the equations which leads to different implementations of the \texttt{CellEquation, CellMatrix} and \texttt{BoundaryEquations} functions in \textit{localpde.h}.\\
+The only difference worth mentioning is the change of the equations which leads to different implementations of the \texttt{ElementEquation, ElementMatrix} and \texttt{BoundaryEquations} functions in \textit{localpde.h}.\\
 Furthermore, the elasticity equations solved in the last example are linear, whereas the plasticity equations are nonlinear; this difference is evident also from the output (here, we need several Newton steps until convergence).\\

### 2013-12-09 DOpEsrc/container/dwrdatacontainer.h

@@ -90,7 +90,7 @@ namespace DOpE
          * is released (see ReleaseLock()).
          */
         virtual void
-        ReInit(unsigned int n_cells);
+        ReInit(unsigned int n_elements);
 
         /**
          * Releases the lock and fills the vector of error indicators
@@ -527,17 +527,17 @@ namespace DOpE
 
   template<typename VECTOR>
     void
-    DWRDataContainerBase<VECTOR>::ReInit(unsigned int n_cells)
+    DWRDataContainerBase<VECTOR>::ReInit(unsigned int n_elements)
     {
-      _error_ind.reinit(n_cells);
-      GetPrimalErrorIndicators().reinit(n_cells);
-      GetDualErrorIndicators().reinit(n_cells);
-      GetControlErrorIndicators().reinit(n_cells);
+      _error_ind.reinit(n_elements);
+      GetPrimalErrorIndicators().reinit(n_elements);
+      GetDualErrorIndicators().reinit(n_elements);
+      GetControlErrorIndicators().reinit(n_elements);
       _lock = true;
     }
 
   /**
-   * Adds just the pure virtual functions GetCellWeight() and GetFaceWeight().
+   * Adds just the pure virtual functions GetElementWeight() and GetFaceWeight().
    * They have to get implemented in derived classes. These two methods are
    * excluded from DWRDataContainerBase() to save two template parameters.
    */
@@ -558,11 +558,11 @@ namespace DOpE
         ;
 
         /**
-         * Returns a CellDataContainer for the weights on the
-         * cells. Pure virtual.
+         * Returns a ElementDataContainer for the weights on the
+         * elements. Pure virtual.
          */
         virtual CDC&
-        GetCellWeight() const = 0;
+        GetElementWeight() const = 0;
 
         /**
          * Returns a FaceDataContainer for the weight on the
@@ -591,7 +591,7 @@ namespace DOpE
     CDC*
     ExtractCDC(const DWRDataContainer<STH, IDC, CDC, FDC, VECTOR>& dwrc)
     {
-      return &dwrc.GetCellWeight();
+      return &dwrc.GetElementWeight();
     }
   template<class FDC, class STH, class IDC, class CDC, typename VECTOR>
     FDC*

### 2013-12-09 DOpEsrc/container/multimesh_elementdatacontainer.h

@@ -21,8 +21,8 @@
 *
 **/
 
-#ifndef MULTIMESH_CELLDATACONTAINER_H_
-#define MULTIMESH_CELLDATACONTAINER_H_
+#ifndef MULTIMESH_ELEMENTDATACONTAINER_H_
+#define MULTIMESH_ELEMENTDATACONTAINER_H_
 
 #include "spacetimehandler.h"
 #include "statespacetimehandler.h"
@@ -42,7 +42,7 @@ namespace DOpE
    * DOFHANDLER, we
    * distinguish between the 'classic' and the 'hp' case.
    *
-   * The Multimesh_CellDataContainers, can deal with different meshes for the control
+   * The Multimesh_ElementDataContainers, can deal with different meshes for the control
    * and state variable as long as both are given as refinements of a common coarse grid
    * by calculation of the respective values on a common refinement.
    *
@@ -52,21 +52,21 @@ namespace DOpE
    */
 
   template<template<int, int> class DH, typename VECTOR, int dim>
-    class Multimesh_CellDataContainer
+    class Multimesh_ElementDataContainer
     {
       public:
-        Multimesh_CellDataContainer()
+        Multimesh_ElementDataContainer()
         {
           throw(DOpE::DOpEException(
               "Dummy class, this constructor should never get called.",
-              "Multimesh_CellDataContainer<dealii::DoFHandler , VECTOR, dim>::Multimesh_CellDataContainer"));
+              "Multimesh_ElementDataContainer<dealii::DoFHandler , VECTOR, dim>::Multimesh_ElementDataContainer"));
         }
         ;
     };
 
   /**
    * This two classes hold all the information we need in the integrator to
-   * integrate something over a cell (could be a functional, a PDE, etc.).
+   * integrate something over a element (could be a functional, a PDE, etc.).
    * Of particular importance: This class holds the FEValues objects.
    *
    * @template VECTOR     Type of the vector we use in our computations (i.e. Vector<double> or BlockVector<double>)
@@ -74,7 +74,7 @@ namespace DOpE
    */
 
   template<typename VECTOR, int dim>
-    class Multimesh_CellDataContainer<dealii::DoFHandler, VECTOR, dim>
+    class Multimesh_ElementDataContainer<dealii::DoFHandler, VECTOR, dim>
     {
 
       public:
@@ -89,7 +89,7 @@ namespace DOpE
          * @param quad                    Reference to the quadrature-rule which we use at the moment.
[…]

### 2013-12-09 DOpEsrc/container/residualestimator.h

@@ -46,7 +46,7 @@ namespace DOpE
         virtual void
         InitFace(double /*h*/) = 0;
         virtual void
-        InitCell(double /*h*/) = 0;
+        InitElement(double /*h*/) = 0;
 
     };
 
@@ -128,7 +128,7 @@ namespace DOpE
          * as well as the weight-vectors.
          */
         void
-        ReInit(unsigned int n_cells);
+        ReInit(unsigned int n_elements);
 
         StateVector<VECTOR>&
         GetPI_h_u()
@@ -199,7 +199,7 @@ namespace DOpE
         virtual DOpEtypes::WeightComputation
         GetWeightComputation() const
         {
-          return DOpEtypes::cell_diameter;
+          return DOpEtypes::element_diameter;
         }
 
         /**
@@ -234,7 +234,7 @@ namespace DOpE
           _weight = h * h * h;
         }
         void
-        InitCell(double h)
+        InitElement(double h)
         {
           _weight = h * h * h * h;
         }
@@ -268,9 +268,9 @@ namespace DOpE
 
   template<class STH, typename VECTOR, int dim>
     void
-    L2ResidualErrorContainer<STH, VECTOR, dim>::ReInit(unsigned int n_cells)
+    L2ResidualErrorContainer<STH, VECTOR, dim>::ReInit(unsigned int n_elements)
     {
-      DWRDataContainerBase<VECTOR>::ReInit(n_cells);
+      DWRDataContainerBase<VECTOR>::ReInit(n_elements);
 
       if (this->GetEETerms() == DOpEtypes::primal_only
           || this->GetEETerms() == DOpEtypes::mixed)
@@ -343,7 +343,7 @@ namespace DOpE
          * as well as the weight-vectors.
          */
         void
-        ReInit(unsigned int n_cells);
+        ReInit(unsigned int n_elements);
 
         StateVector<VECTOR>&
         GetPI_h_u()
@@ -414,7 +414,7 @@ namespace DOpE
         virtual DOpEtypes::WeightComputation
         GetWeightComputation() const
         {
-          return DOpEtypes::cell_diameter;
+          return DOpEtypes::element_diameter;
         }
 
         /**
@@ -448,7 +448,7 @@ namespace DOpE
           _weight = h;
         }
         void
-        InitCell(double h)
+        InitElement(double h)
         {
           _weight = h * h;
         }
[…]

### 2013-12-09 DOpEsrc/interfaces/constraintinterface.h

@@ -42,7 +42,7 @@ namespace DOpE
    * GlobalConstraints are dealt with as a Functional, 
    * hence all functions from Functionals are inherited.
    *
-   * @tparam <CDC>             The CellDataContainer object
+   * @tparam <CDC>             The ElementDataContainer object
    *                           needed by the base class.
    * @tparam <FDC>             The FaceDataContainer object
    *                           needed by the base class.

### 2013-12-09 DOpEsrc/source/outputhandler.cc

@@ -506,7 +506,7 @@ namespace DOpE
 
    template<typename VECTOR>
      void
-     DOpEOutputHandler<VECTOR>::WriteCellwise(const Vector<double>&q,
+     DOpEOutputHandler<VECTOR>::WriteElementwise(const Vector<double>&q,
          std::string name, std::string dof_type)
      {
        if (AllowWrite(name))
@@ -527,10 +527,10 @@ namespace DOpE
          Write("Writing [" + outfile.str() + "]", 4);
 
          if (dof_type == "control")
-           GetReducedProblem()->WriteToFileCellwise(q, name, outfile.str(),
+           GetReducedProblem()->WriteToFileElementwise(q, name, outfile.str(),
                dof_type, _control_ending);
          else if (dof_type == "state")
-           GetReducedProblem()->WriteToFileCellwise(q, name, outfile.str(),
+           GetReducedProblem()->WriteToFileElementwise(q, name, outfile.str(),
                dof_type, _ending);
        }
      }

### 2013-12-09 DOpEsrc/tsschemes/forward_euler_problem.h

@@ -101,35 +101,35 @@ namespace DOpE
       
       template<typename CDC>
         void
-        CellEquation(const CDC& cdc,
-		     dealii::Vector<double> &local_cell_vector, double scale, double /*scale_ico*/)
+        ElementEquation(const CDC& cdc,
+		     dealii::Vector<double> &local_vector, double scale, double /*scale_ico*/)
         {
           if (this->GetPart() == "New")
             {
-              dealii::Vector<double> tmp(local_cell_vector);
+              dealii::Vector<double> tmp(local_vector);
               tmp = 0.0;
-              this->GetProblem().CellEquation(cdc, tmp, 0., scale * this->GetProblem().GetBaseProblem().GetSpaceTimeHandler()->GetStepSize());
-              local_cell_vector += tmp;
+              this->GetProblem().ElementEquation(cdc, tmp, 0., scale * this->GetProblem().GetBaseProblem().GetSpaceTimeHandler()->GetStepSize());
+              local_vector += tmp;
 
               tmp = 0.0;
-              this->GetProblem().CellTimeEquation(cdc, tmp,
+              this->GetProblem().ElementTimeEquation(cdc, tmp,
                   scale );
-              local_cell_vector += tmp;
+              local_vector += tmp;
 
-              this->GetProblem().CellTimeEquationExplicit(cdc, local_cell_vector,
+              this->GetProblem().ElementTimeEquationExplicit(cdc, local_vector,
                   scale );
 
             }
           else if (this->GetPart() == "Old")
             {
-              dealii::Vector<double> tmp(local_cell_vector);
+              dealii::Vector<double> tmp(local_vector);
               tmp = 0.0;
-              this->GetProblem().CellEquation(cdc, tmp, scale * this->GetProblem().GetBaseProblem().GetSpaceTimeHandler()->GetStepSize(), 0.);
-              local_cell_vector += tmp;
+              this->GetProblem().ElementEquation(cdc, tmp, scale * this->GetProblem().GetBaseProblem().GetSpaceTimeHandler()->GetStepSize(), 0.);
+              local_vector += tmp;
 
-              this->GetProblem().CellTimeEquation(
+              this->GetProblem().ElementTimeEquation(
                   cdc,
-                  local_cell_vector,
+                  local_vector,
[…]

### 2013-12-09 Examples/OPT/StatPDE/Example4/localpde.h

@@ -48,13 +48,13 @@ template<
       }
 
       void
-      CellEquation(const CDC<DH, VECTOR, dealdim>& cdc,
-          dealii::Vector<double> &local_cell_vector, double scale,
+      ElementEquation(const CDC<DH, VECTOR, dealdim>& cdc,
+          dealii::Vector<double> &local_vector, double scale,
           double /*scale_ico*/)
       {
         const DOpEWrapper::FEValues<dealdim> & state_fe_values =
             cdc.GetFEValuesState();
-        unsigned int n_dofs_per_cell = cdc.GetNDoFsPerCell();
+        unsigned int n_dofs_per_element = cdc.GetNDoFsPerElement();
         unsigned int n_q_points = cdc.GetNQPoints();
         {
           //Reading data
@@ -74,9 +74,9 @@ template<
                   * sin(M_PI * state_fe_values.quadrature_point(q_point)(1)));
           _fvalues[q_point](1) = 1.;
 
-          for (unsigned int i = 0; i < n_dofs_per_cell; i++)
+          for (unsigned int i = 0; i < n_dofs_per_element; i++)
           {
-            local_cell_vector(i) += scale
+            local_vector(i) += scale
                 * (_ugrads[q_point][0]
                     * state_fe_values[comp_0].gradient(i, q_point)
                     + _ugrads[q_point][1]
@@ -91,13 +91,13 @@ template<
       }
 
       void
-      CellEquation_U(const CDC<DH, VECTOR, dealdim>& cdc,
-          dealii::Vector<double> &local_cell_vector, double scale,
+      ElementEquation_U(const CDC<DH, VECTOR, dealdim>& cdc,
+          dealii::Vector<double> &local_vector, double scale,
           double /*scale_ico*/)
       {
         const DOpEWrapper::FEValues<dealdim> & state_fe_values =
             cdc.GetFEValuesState();
-        unsigned int n_dofs_per_cell = cdc.GetNDoFsPerCell();
+        unsigned int n_dofs_per_element = cdc.GetNDoFsPerElement();
         unsigned int n_q_points = cdc.GetNQPoints();
         {
           assert(this->_problem_type == "adjoint");
@@ -110,9 +110,9 @@ template<
         const FEValuesExtractors::Scalar comp_1(1);
         for (unsigned int q_point = 0; q_point < n_q_points; q_point++)
         {
[…]

### 2013-12-09 Examples/OPT/StatPDE/Example7/localpde.h

@@ -45,13 +45,13 @@ template<
       }
 
       void
-      CellEquation(const CDC<DH, VECTOR, dealdim>& cdc,
-          dealii::Vector<double> &local_cell_vector, double scale,
+      ElementEquation(const CDC<DH, VECTOR, dealdim>& cdc,
+          dealii::Vector<double> &local_vector, double scale,
           double /*scale_ico*/)
       {
         const DOpEWrapper::FEValues<dealdim> & state_fe_values =
             cdc.GetFEValuesState();
-        unsigned int n_dofs_per_cell = cdc.GetNDoFsPerCell();
+        unsigned int n_dofs_per_element = cdc.GetNDoFsPerElement();
         unsigned int n_q_points = cdc.GetNQPoints();
         {
           //Reading data
@@ -67,9 +67,9 @@ template<
 
         for (unsigned int q_point = 0; q_point < n_q_points; q_point++)
         {
-          for (unsigned int i = 0; i < n_dofs_per_cell; i++)
+          for (unsigned int i = 0; i < n_dofs_per_element; i++)
           {
-            local_cell_vector(i) += scale
+            local_vector(i) += scale
                 * (_ugrads[q_point] * state_fe_values.shape_grad(i, q_point)
                     - _qvalues[q_point]
                         * state_fe_values.shape_value(i, q_point))
@@ -79,13 +79,13 @@ template<
       }
 
       void
-      CellEquation_U(const CDC<DH, VECTOR, dealdim>& cdc,
-          dealii::Vector<double> &local_cell_vector, double scale,
+      ElementEquation_U(const CDC<DH, VECTOR, dealdim>& cdc,
+          dealii::Vector<double> &local_vector, double scale,
           double /*scale_ico*/)
       {
         const DOpEWrapper::FEValues<dealdim> & state_fe_values =
             cdc.GetFEValuesState();
-        unsigned int n_dofs_per_cell = cdc.GetNDoFsPerCell();
+        unsigned int n_dofs_per_element = cdc.GetNDoFsPerElement();
         unsigned int n_q_points = cdc.GetNQPoints();
         {
           assert(this->_problem_type == "adjoint");
@@ -96,9 +96,9 @@ template<
 
         for (unsigned int q_point = 0; q_point < n_q_points; q_point++)
         {
-          for (unsigned int i = 0; i < n_dofs_per_cell; i++)
+          for (unsigned int i = 0; i < n_dofs_per_element; i++)
           {
[…]

### 2013-12-09 Examples/PDE/InstatPDE/Example2/localpde.h

@@ -81,18 +81,18 @@ template<
         return false;
       }
 
-      // The part of CellEquation scaled by scale contains all "normal" terms which
+      // The part of ElementEquation scaled by scale contains all "normal" terms which
       // can be treated by full "theta" time-discretization
       void
-      CellEquation(const CellDataContainer<DH, VECTOR, dealdim>& cdc,
-          dealii::Vector<double> &local_cell_vector, double scale,
+      ElementEquation(const ElementDataContainer<DH, VECTOR, dealdim>& cdc,
+          dealii::Vector<double> &local_vector, double scale,
           double scale_ico)
       {
         assert(this->_problem_type == "state");
 
         const DOpEWrapper::FEValues<dealdim> & state_fe_values =
             cdc.GetFEValuesState();
-        unsigned int n_dofs_per_cell = cdc.GetNDoFsPerCell();
+        unsigned int n_dofs_per_element = cdc.GetNDoFsPerElement();
         unsigned int n_q_points = cdc.GetNQPoints();
         unsigned int material_id = cdc.GetMaterialId();
 
@@ -172,7 +172,7 @@ template<
                 NSE_in_ALE::get_Incompressibility_ALE<dealdim>(q_point,
                     _ugrads);
 
-            for (unsigned int i = 0; i < n_dofs_per_cell; i++)
+            for (unsigned int i = 0; i < n_dofs_per_element; i++)
             {
               // Fluid, explicit
               const Tensor<1, dealdim> phi_i_v =
@@ -189,25 +189,25 @@ template<
 		state_fe_values[displacements_w].gradient(i, q_point);
 
 
-              local_cell_vector(i) += scale
+              local_vector(i) += scale
                   * (convection_fluid * phi_i_v
                       + scalar_product(stress_fluid, phi_i_grads_v))
                   * state_fe_values.JxW(q_point);
 
-              local_cell_vector(i) += scale_ico
+              local_vector(i) += scale_ico
 		* ( alpha_u
                       * scalar_product(grad_w, phi_i_grads_u))
                   * state_fe_values.JxW(q_point);
 
-              local_cell_vector(i) += scale
+              local_vector(i) += scale
                   * (scalar_product(fluid_pressure, phi_i_grads_v))
                   * state_fe_values.JxW(q_point);
[…]

### 2013-12-09 Examples/PDE/InstatPDE/Example5/localpde.h

@@ -46,17 +46,17 @@ template<
 
       }
 
-      // Domain values for cells
+      // Domain values for elements
       void
-      CellEquation(const CDC<DH, VECTOR, dealdim>& cdc,
-          dealii::Vector<double> &local_cell_vector, double scale,
+      ElementEquation(const CDC<DH, VECTOR, dealdim>& cdc,
+          dealii::Vector<double> &local_vector, double scale,
           double /*scale_ico*/)
       {
         assert(this->_problem_type == "state");
 
         const DOpEWrapper::FEValues<dealdim> & state_fe_values =
             cdc.GetFEValuesState();
-        unsigned int n_dofs_per_cell = cdc.GetNDoFsPerCell();
+        unsigned int n_dofs_per_element = cdc.GetNDoFsPerElement();
         unsigned int n_q_points = cdc.GetNQPoints();
 
         _uvalues.resize(n_q_points);
@@ -68,14 +68,14 @@ template<
         for (unsigned int q_point = 0; q_point < n_q_points; q_point++)
         {
 
-          for (unsigned int i = 0; i < n_dofs_per_cell; i++)
+          for (unsigned int i = 0; i < n_dofs_per_element; i++)
           {
 
             const double phi_i = state_fe_values.shape_value(i, q_point);
             const Tensor<1, dealdim> phi_i_grads = state_fe_values.shape_grad(i,
                 q_point);
 
-            local_cell_vector(i) += scale
+            local_vector(i) += scale
                 * ((_ugrads[q_point] * phi_i_grads)
                     + _uvalues[q_point] * _uvalues[q_point] * phi_i)
                 * state_fe_values.JxW(q_point);
@@ -85,31 +85,31 @@ template<
       }
 
       void
-      CellMatrix(const CDC<DH, VECTOR, dealdim>& cdc,
-          FullMatrix<double> &local_entry_matrix, double scale, double)
+      ElementMatrix(const CDC<DH, VECTOR, dealdim>& cdc,
+          FullMatrix<double> &local_matrix, double scale, double)
       {
         const DOpEWrapper::FEValues<dealdim> & state_fe_values =
             cdc.GetFEValuesState();
-        unsigned int n_dofs_per_cell = cdc.GetNDoFsPerCell();
+        unsigned int n_dofs_per_element = cdc.GetNDoFsPerElement();
         unsigned int n_q_points = cdc.GetNQPoints();
[…]

### 2013-12-09 Examples/PDE/StatPDE/Example10/main.cc

@@ -56,7 +56,7 @@ const static int DIM = 2;
 
 #define DOFHANDLER DoFHandler
 #define FE FESystem
-#define CDC CellDataContainer
+#define CDC ElementDataContainer
 #define FDC FaceDataContainer
 
 typedef QGauss<DIM> QUADRATURE;
@@ -220,7 +220,7 @@ main(int argc, char **argv)
       const StateVector<VECTOR> &gu = a.GetU();
       solution = 0;
       solution = gu.GetSpacialVector();
-      Vector<float> estimated_error_per_cell(triangulation.n_active_cells());
+      Vector<float> estimated_error_per_element(triangulation.n_active_cells());
 
       std::vector<bool> component_mask(5, false);
       component_mask[2] = true;
@@ -228,8 +228,8 @@ main(int argc, char **argv)
       KellyErrorEstimator<DIM>::estimate(
           static_cast<const DoFHandler<DIM>&>(DOFH.GetStateDoFHandler()),
           QGauss<1>(2), FunctionMap<DIM>::type(), solution,
-          estimated_error_per_cell, component_mask);
-      DOFH.RefineSpace(RefineFixedNumber(estimated_error_per_cell, 0.3, 0.0));
+          estimated_error_per_element, component_mask);
+      DOFH.RefineSpace(RefineFixedNumber(estimated_error_per_element, 0.3, 0.0));
       solver.ReInit();
       out.ReInit();
     }

### 2013-12-09 Examples/PDE/StatPDE/Example12/main.cc

@@ -60,7 +60,7 @@ const static int DIM = 2;
 
 #define DOFHANDLER DoFHandler
 #define FE FESystem
-#define CDC CellDataContainer
+#define CDC ElementDataContainer
 #define FDC FaceDataContainer
 
 typedef QGauss<DIM> QUADRATURE;

### 2013-12-09 Examples/PDE/StatPDE/Example8/main.cc

@@ -55,7 +55,7 @@ const static int DIM = 2;
 
 #define DOFHANDLER DoFHandler
 #define FE FESystem
-#define CDC CellDataContainer
+#define CDC ElementDataContainer
 #define FDC FaceDataContainer
 
 typedef QSimpson<DIM> QUADRATURE;

### 2013-12-09 Examples/Documentation/foreword.tex
Update Doku und Beseitigung kleinerer Doku-typos in zwei main files.

@@
 problems by interpreting every PDE problem as nonlinear and applying a
-Newton method to solve it. While deal.II leaves much of the work and many
+Newton method to solve it. 
+The focus is on the numerical solution of both stationary and nonstationary
+problems which come from different application fields, like elasticity and
+plasticity, fluid dynamics, and multiphysics problems such as 
+fluid-structure interactions.
+
+\twick{This is copied from the paper:
+The main feature of \dope{} is to give a unified interface to high level 
+algorithms such as 
+time-stepping methods, nonlinear solvers and optimization routines. 
+\dope{} is designed in such a way that the user only needs to write those parts
+of the code that are problem dependent while all invariant 
+parts of the algorithms
+are reusable without any need for further coding.
+In particular, the user is enabled to switch between various different 
+algorithms without the need to rewrite the problem dependent code, 
+though obviously he or she will
+have to replace the algorithm object with an other one. 
+This replacement can be done by replacing the appropriate object at only
+one point in the code.
+In addition to the finite element code provided by deal.II --which at present is the only FE-toolkit to which we provide an 
+interface-- 
+the presented library
+\dope{} is user-focused by delivering
+prefabricated tools which require adjustments by the user only for parts
+connected to his specific problem. 
+This is in contrast to deal.II which leaves the implementation of all high-level algorithms to the user.
+The possible solution of 
+optimization problems with PDE constraints is an innovative feature in \dope{}.
+}
+
+\twickremove{Please check if we can remove part and use the red part!
+While deal.II leaves much of the work and many
 decisions to the user, DOpE intends to be user-optimized by delivering
@@
 constraints is an innovation in the DOpE framework.
-The focus is on the numerical solution of both stationary and nonstationary
-problems which come from different application fields, like elasticity and
-plasticity, fluid dynamics, and fluid-structure interactions.
+}
 
@@
 \item Several examples showing the solution of several PDEs including
-   Poisson, Navier-Stokes, Plasticity and fluid-structure interaction problems. 
+   Poisson, Navier-Stokes, plasticity, fluid-structure interaction problems,
+and finally a coupled Biot-Lam\'e-Navier system
 \item Self written line search and trust region newton algorithms for the 

### 2013-12-09 Examples/OPT/StatPDE/Example1/content.tex
commit before talk of t and w
Code cleaned up to OPT StatPDE Ex2. TODO: content.tex
contents tex in opt problems
Update Doku und Beseitigung kleinerer Doku-typos in zwei main files.

@@
 \subsubsection{General problem description}
-This example solves the distributed minimization problem
+\twick{Dieses Beispiel ist nun extremst ausfuehrlich hinsichtlich der
+localpde. Da dies ein Manual ist, wuensche ich mir, dass wir dies (respective
+moeglicher Rechtschreibfehler und typos) auch so belassen. Mir hat das damals 
+bei FSI Opt sehr geholfen.}
+This example solves a distributed minimization problem
+\twick{and shows how to estimate the error in the cost functional for stationary optimization
+problems. The problem reads:}
 \begin{gather*}
@@
 
-The example shows how to estimate the error in the cost functional for stationary optimization
-problems.
+\twick{Ab hier ist alles neu!!}
+\subsubsection{Background information and program description}
+In the following, we describe all extensions to the previous problems 
+relevant to solving PDE-based optimization with \dope{}. So far, 
+we had only to implement the \texttt{ElementEquation} and the corresponding
+matrix \texttt{ElementMatrix}. Now, based on the idea of the reduced cost
+functional, we have to compute certain additional equations representing the 
+adjoint, tangent, and adjoint hessian equations
+\texttt{ElementEquation\underline{ }U, ElementEquation\underline{ }T,
+ElementEquation\underline{ }UTT}
+for the state equation and in the same terms arising from the functional
+itself. Let us shed some light into all equations by giving some background
+information and overview first.
+
+{
+In abstract form, we are given the following optimization problem:
+\begin{equation*}
+J(q,U) \, \rightarrow \, \min , \quad a(q,U)(\Psi) = 0 \quad \forall\Psi V
+\end{equation*}
+Lagrangian:
+\begin{equation*}
+{\cal L}(q,U,Z):=J(q,U) - a(q,U)(Z)
+\end{equation*}
+Optimality system (KKT system):
+\begin{align*}
+a_U'(q,U)(\Phi , Z) &= J'_U(q,U)(\Phi) \quad \forall \Phi\in V \\
+a_q'(q,U)(\chi , Z) &= J'_q(q,U)(\chi) \quad \forall \chi\in Q \\
+a(q,U)(\Psi) &= 0 \qquad \forall\Psi\in V
+\end{align*}
+or equivalently
+\begin{align*}
+{\cal L}'_U(q,U,Z)(\Phi) &= 0 \quad\forall \Phi\in V \quad\text{(Adjoint Equation)} \\
+{\cal L}'_q(q,U,Z)(\chi) &= 0 \quad\forall \chi\in V \quad\text{(Gradient Equation)} \\
+{\cal L}'_Z(q,U,Z)(\Psi) &= 0 \quad\forall \Psi\in V \quad\text{(State Equation)}
+\end{align*}
+The continuous problem is discretized by a standard Galerkin method using 
+finite dimensional subspaces $Q_h \times V_h \subset Q\times V$:
+\begin{equation*}
+J(q,U) \, \rightarrow \, \min , \quad a(q,U)(\Psi) = 0 \quad \forall\Psi V
+\end{equation*}
+Discrete saddle-point problems
+\begin{align*}
+a_U'(q_h,U_h)(\Phi_h , Z_h) &= J'_U(q_h,U_h)(\Phi_h) \quad \forall \Phi_h\in V_h \\
+a_q'(q_h,U_h)(\chi_h , Z_h) &= J'_q(q_h,U_h)(\chi_h) \quad \forall \chi_h\in Q_h \\
+a(q_h,U_h)(\Psi_h) &= 0 \qquad \forall\Psi_h\in V_h
+\end{align*}
+\subsubsection{Solution process}
+In this section, we briefly discuss the solution process for the 
+optimization problem. We further details, we refer to the standard literature.
+The unconstrained optimal control problem is reformulated as follows. 
+We introduce the solution operator $S:Q\rightarrow V$ of the state equation.
+Then:
+\begin{equation*}
+j(q):= J(q,S(q))\, \rightarrow \, \min , \quad a(q,S(q))(\Psi) = 0 \quad 
+\forall \Psi \in V.
+\end{equation*}
+The local existence and sufficient regularity of $S$ is assumed. The 
+necessary optimality conditions of first and second order are
+\begin{equation*}
+j'(q)(\delta q) = 0 , \quad j''(q)(\delta q,\delta q) \geq 0 \quad 
+\forall\delta q\in Q.
+\end{equation*} 
+The derivatives of the reduced functional can be computed using the 
+Lagrangian
+\begin{equation*}
+{\cal L}(q,U,Z) =J(q,U) - a(q,U)(Z)
+\end{equation*}
+as already introduced. Let $q\in Q$ be given. We derive
+ the corresponding state $U=S(q)\in V$ and the dual variable
+$Z\in V$. The following terms hold:
+}
+
+\textbf{Dual equation}
+
+\begin{equation*}
+{\cal L}_U'(q,U,Z)(\Psi) = 0 \quad\forall\Psi\in V
+\end{equation*}
+Then,
+\begin{equation*}
+j'(q)(\delta q) = {\cal L}_q' (q,U,Z)(\delta q) \quad \text{for }\delta q\in Q.
+\end{equation*}
+
+\textbf{Tangent equation}
+
+For the right-hand side of Newton's method, we also need further terms. Let
+$\delta q\in Q$ be given, and let $\delta U\in V$ and $\delta Z\in V$ fulfill
+the equation
+\begin{equation*}
+{\cal L}_{qZ}'' (q,U,Z)(\delta q,\Phi) + {\cal L}_{UZ}'' (q,U,Z)(\delta
+U,\Phi)
+\quad\forall \Phi\in V,
+\end{equation*}
+further
+
+\textbf{Dual Hessian equation}
+
+\begin{equation*}
+{\cal L}_{qU}'' (q,U,Z)(\delta q,\Phi) + {\cal L}_{UU}'' 
+(q,U,Z)(\delta U,\Phi) + {\cal L}_{ZU}'' (q,U,Z)(\delta Z,\Phi)
+\quad\forall \Phi\in V,
+\end{equation*}
+Then, for $\delta r\in Q$, it holds
+\begin{align*}
+j''(q)(\delta q, \delta r) 
+&= {\cal L}_{qq}'' (q,U,Z)(\delta q,\delta r) \\
+&+ {\cal L}_{Uq}'' (q,U,Z)(\delta U,\delta r) \\ 
+&+ {\cal L}_{Zq}'' (q,U,Z)(\delta Z,\delta r) \quad \text{for }\delta r\in Q.
+\end{align*}
+These terms build the the Newton scheme as follows:
+\begin{align*}
+j''(q^n )(\delta q,\chi ) =&\, -j'(q^n)(\chi) \quad\forall\chi \in Q,\\
+q^{n+1} :=&\, q^n + \delta q , \quad n=0,1,2,\ldots .
+\end{align*}
+
+Finally, we collect the representations of the forgoing work. First, we recall 
+the terms of the Lagrangian:
+
+\textbf{Adjoint:} Find $z\in X$ such that
+\begin{equation*}
+{\cal L}_u' (q,u,z)(\phi) = 0 \quad \forall\phi\in X. 
+\end{equation*}
+\textbf{Tangent (linearized state):} Find $\delta u\in X$ such that
+\begin{equation*}
+{\cal L}_{qz}'' (q,u,z)(\delta q, \phi) +
+{\cal L}_{uz}'' (q,u,z)(\delta u, \phi) = 0 \quad \forall\phi\in X. 
+\end{equation*}
+\textbf{Additional adjoint:} Find $\delta z\in X$ such that
+\begin{equation*}
+{\cal L}_{qu}'' (q,u,z)(\delta q, \phi) +
+{\cal L}_{uu}'' (q,u,z)(\delta u, \phi) +
+{\cal L}_{zu}'' (q,u,z)(\delta z, \phi) = 0 \quad \forall\phi\in X. 
+\end{equation*}
+
+Especially, the equations for stationary optmization problems are
+given as follows:
+
+\textbf{Adjoint:} Find $z\in X$ such that
+\begin{equation*}
+a_u'(q,u)(\phi,z) = J'(u)(\phi) \quad \forall\phi\in X. 
+\end{equation*}
+\textbf{Tangent:} Find $\delta u\in X$ such that
+\begin{equation*}
+a_u'(q,u)(\delta u,\phi) = -a_q'(q,u)(\delta q,\phi)  \quad \forall\phi\in X. 
+\end{equation*}
+\textbf{Additional adjoint:} Find $\delta z\in X$ such that
+\begin{equation*}
+a_u'(q,u)(\phi,\delta z) = a_{uu}''(q,u)(\delta u,\phi,z)
+-a_{qu}''(q,u)(\delta q,\phi,z) + J''(u)(\delta,\phi) \quad \forall\phi\in X. 
+\end{equation*}
+
+Using the Lagrangian, the previous expressions read
+
+\textbf{Lagrangian}
+\begin{equation*}
+{\cal L}(q,u,z) =J(q,u) - a(q,u)(z)
+\end{equation*}
+
+\textbf{State Equation}
+\begin{align*}
+a(q,u)(\phi) = 0 
+\end{align*}
+
+\textbf{Dual Equation}
+
+\begin{align*}
+{\cal L}_{u}' (q,u,z)(\phi ) &= 0 \\
+\Leftrightarrow a_u' (q,u)(\phi,z) &= J_u' (q,u)(\phi)
+\end{align*}
+
+\textbf{Tangent Equation}
+
+\begin{align*}
+{\cal L}_{qz}'' (q,u,z)(\delta q, \phi) 
++ {\cal L}_{uz}'' (q,u,z)(\delta u, \phi) &= 0 \\
+\Leftrightarrow a_u' (q,u)(\delta u,\phi) + a_q' (q,u)(\delta q,\phi) &= 0
+\end{align*}
+
+\textbf{Dual Hessian Equation}
+
+\begin{align*}
+{\cal L}_{qu}'' (q,u,z)(\delta q, \phi) 
+&+ {\cal L}_{uu}'' (q,u,z)(\delta u, \phi) \\
+&+ {\cal L}_{zu}'' (q,u,z)(\delta z, \phi) = 0 \\
+\Leftrightarrow a_u' (q,u)(\phi, \delta z) 
+&+ a_{uu}' (q,u)(\delta u,\phi, z) \\
+&+ a_{qu}' (q,u)(\delta q,\phi, z)  = J_{uu}''(q,u)(\cdot ,\cdot )
+\end{align*}
+
+\textbf{Newton left hand side: first derivative}
+
+\begin{align*}
+j'(q)(\delta q) &= {\cal L}_{q}' (q,u,z)(\delta q) \quad\text{(Gradient equation)} \\
+\Leftrightarrow j'(q)(\delta q) &= -a_q' (q,u)(\delta q,z) + J_{q}'(q,u)(\delta q)
+\end{align*}
+
+\textbf{Newton right hand side: second derivative}
+
+\begin{align*}
+j''(q)(\delta q, \delta r) &= {\cal L}_{qq}'' (q,u,z)(\delta q, \delta r) + 
+{\cal L}_{uq}'' (q,u,z)(\delta u, \delta r) \\
+&\quad +  {\cal L}_{zq}'' (q,u,z)(\delta z, \delta r) \\
+\Leftrightarrow  j''(q)(\delta q, \delta r) &= 
+J_{qq}''(q,u)(\delta q, \delta r) -a_{qq}'' (q,u)(\delta q,\delta r, z) \\
+&\quad - a_{uq}'' (q,u)(\delta u,\delta r, z) -a_{q}' (q,u)(\delta r,\delta z)  
+\end{align*}
+
+\subsubsection{Implementation in \dope{}}
+
+\begin{align*}
+\text{ElementEquation (state)} \quad &\Leftrightarrow \quad a(q,u)(\phi) \\
+\text{ElementEquationU (adjoint)} \quad 
+&\Leftrightarrow \quad a_u'(q,u)(\phi, z) \\ 
+\text{ElementEquationUT (tangent)} \quad 
+&\Leftrightarrow \quad a_u'(q,u)(\delta u, \phi) \\
+\text{ElementEquationUTT (adjoint hessian)} \quad 
+&\Leftrightarrow \quad  a_u'(q,u)(\phi, \delta z) \\
+\text{ElementEquationQ (gradient)} \quad 
+&\Leftrightarrow \quad a_q'(q,u)(\delta q,z) \\
+\text{ElementEquationQT (tangent)} \quad 
+&\Leftrightarrow \quad a_q'(q,u)(\delta q, \phi) \\
+\text{ElementEquationQTT (hessian)} \quad 
+&\Leftrightarrow \quad a_{q}'(q,u)(\delta q,\delta z) \\
+\text{ElementEquationUU (adjoint hessian)} \quad 
+&\Leftrightarrow \quad a_{uu}''(q,u)(\delta u, \phi, z) \\
+\text{ElementEquationQU (adjoint hessian)} \quad 
+&\Leftrightarrow \quad a_{qu}''(q,u)(\delta q, \phi, z) \\
+\text{ElementEquationUQ (hessian)} \quad 
+&\Leftrightarrow \quad a_{uq}''(q,u)(\delta u, \delta r, z)\\
+\text{ElementEquationQQ (hessian)} \quad 
+&\Leftrightarrow \quad a_{qq}''(q,u)(\delta q, \delta r, z)\\
+\text{ElementRightHandSide (state)} \quad 
+&\Leftrightarrow \quad f(\phi ) \\
+\text{ElementMatrix (all)} \quad 
+&\Leftrightarrow \quad a'(q,u)(\delta u) \\
+\text{ControlElementEquation (gradient or hessian)} \quad 
+&\Leftrightarrow \quad {\cal L}_q' (q,u,z)(\delta q) \\
+\text{ControlElementMatrix (all)} \quad 
+&\Leftrightarrow \quad a(q,u)(\phi)
+\end{align*}
+
+Functionals
+\begin{align*}
+\text{ElementValue (all)} \quad  & \Leftrightarrow \quad J(q,u) \\
+\text{ElementValueU (all)} \quad  & \Leftrightarrow \quad J_u'(q,u)(\phi) \\
+\text{ElementValueQ (all)} \quad  & \Leftrightarrow \quad J_q'(q,u)(\phi) \\
+\text{ElementValueUU (all)} \quad & \Leftrightarrow \quad J_{uu}''(q,u)(\psi, \phi) \\
+\text{ElementValueQU (all)} \quad & \Leftrightarrow \quad J_{qu}''(q,u)(\psi, \phi) \\
+\text{ElementValueUQ (all)} \quad & \Leftrightarrow \quad J_{uq}''(q,u)(\psi, \phi) \\
+\text{ElementValueQQ (all)} \quad & \Leftrightarrow \quad J_{qq}''(q,u)(\psi, \phi) \\
+\end{align*}
+
+\subsubsection{Back to the specific equations in this example}
+We have
+\[
+a(q,u)(\phi) = (\nabla u, \nabla\phi) - (q + f, \phi).
+\]
+and
+\[
+a_u'(q,u)(\phi, z) = (\nabla\phi, \nabla z)
+\]
+and
+\[
+a_u'(q,u)(\delta u, \phi) = (\nabla\delta u, \nabla\phi)
+\]
+and
+\[
+a_u'(q,u)(\phi, \delta z) = (\nabla\phi, \nabla\delta z).
+\]
+and 
+\[
+a_q'(q,u)(\delta q, \phi) = -(z,\psi^q)
+\]
+and 
+\[
+a_q'(q,u)(\delta q, \phi) = (\delta q,\phi)
+\]
+and finally
+\[
+a_{q}'(q,u)(\delta q,\delta z) = - (\delta z,\psi^q)
+\]
+On the functional level, we have the following terms:
+\[
+J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2d
+\]
+and 
+\[
+J_u'(q,u)(\phi) = (u-u^d, \phi)
+\]
+and
+\[
+J_q'(q,u)(\phi) = \alpha (q,\psi^q)
+\]
+and 
+\[
+J_{uu}''(q,u)(\psi, \phi) = (\delta u, \phi).
+\]
+All other terms, specifically mixed terms with $QU$ etc. are zero in this
+example.
+
+\subsubsection{\texttt{main.cc}}
+Finally, the main file of the optimization examples does not look very much 
+different than for pure PDE computations - which is one of the crucial aims 
+of our library. 
+Here, instead of using a \texttt{pdeproblemcontainer}, we use 
+now an \texttt{optproblemcontainer}. Furthermore, 
+we define \texttt{ReducedNewtonAlgorithm} to solve the optimization problem.
+Next, in the body of the main file, we introduce a second FE function for the 
+control variable (using the FE\underline{}Nothing function as we will see 
+in Example 2 also allows for 0d controls). Then, we define 
+a \texttt{COSTFUNCTIONAL}. Finally, the problem is either solved 
+by calling \texttt{Alg.Solve(q)} and/or the user might check if the 
+derivatives are implemented correctly by calling 
+\texttt{Alg.CheckGrads} or \texttt{Alg.CheckHessian}. The latter 
+two functionalities are highly recommended to check your initial 
+implementation before wondering about your results.

### 2013-12-09 Examples/OPT/StatPDE/Example2/content.tex
Code cleaned up to OPT StatPDE Ex2. TODO: content.tex
Further changes
Overhaul of OPT StatPDe Example 5
Update Doku und Beseitigung kleinerer Doku-typos in zwei main files.

@@
 \begin{align*}
-\min_{(q,u)\in \mathbb R^3 \times H_0^1(\Omega; \mathbb R^2)} J(q,u) &= \frac{1}{2} \sum_{i=0}^2\abs{(u-\overline u)(x_i)}^2 + \frac{\alpha}{2}\|q\|^2\\
-\text{s.t.} (\nabla u,\nabla \phi) &= (f(q),\phi)\;\;\forall\,\phi \in H^1_0(\Omega; \mathbb R^2)
+\min_{(q,u)\in \mathbb R^3 \times H_0^1(\Omega; \mathbb R^2)} J(q,u) &= \frac{1}{2} \sum_{i=0}^2\abs{(u-\overline u)(x_i)}^2 + \frac{\alpha}{2}\|q\|^2&&\\
+\text{s.t. } (\nabla u,\nabla \phi) &= (f(q),\phi) \quad\forall\,\phi \in H^1_0(\Omega; \mathbb R^2)
 \end{align*}
-on the domain $\Omega = [0,1]^2$, with
+on the domain $\Omega = [0,1]^2$, with zero dirichlet boundary conditions and
 \begin{itemize}
@@
 \subsubsection{Program description}
-Notice that the cost functional is of mixed type (from our computational point of view), i.e.
+\twick{
+In contrast to the first example, the control is now a discrete quantity (for
+the three observation points) where we use the \texttt{FE\underline{}Nothing}
+element to assigne the three controls. Here, the number of components equals
+the number of controls. 
+In addition, }notice that the cost functional is of mixed type (from our computational point of view), i.e.
 the first part is a pointfunctional whereas the regularization part requires the evaluation of a domain integral.
-So our \texttt{LocalFunctional} returns as his type:
+To handle this, we need a special integrator as well as a special newtonsolver.
+Additionally, our \texttt{LocalFunctional} returns as his type:
 \begin{verbatim}
@@
 \begin{itemize} 
-\item \texttt{Value},
-\item  \texttt{Value\_U},
-\item \texttt{Value\_Q}, 
-\item \texttt{Value\_UU},
-\item \texttt{Value\_UQ},
-\item \texttt{Value\_QU},
-\item \texttt{Value\_QQ}
+\item \texttt{ElementValue},
+\item \texttt{ElementValue\_U},
+\item \texttt{ElementValue\_Q}, 
+\item \texttt{ElementValue\_UU},
+\item \texttt{ElementValue\_UQ},
+\item \texttt{ElementValue\_QU},
+\item \texttt{ElementValue\_QQ}
 \end{itemize}
  from \texttt{FunctionalInterface}, but also all the aforementioned methods with a preceeding \text{Point} (\texttt{PointValue} etc.).
+%FE_Nothing, handle of the control erlaeutern.

### 2013-12-09 Examples/OPT/StatPDE/Example3/content.tex
Update Doku und Beseitigung kleinerer Doku-typos in zwei main files.

@@
 
+The control $q$ enters via the weak formulation. It reads,
+\begin{align*}
+a(q,v,p)(\phi) = (\sigma(v,p), \phi) + (v\cdot \nabla v, \phi) 
++ \langle q, \phi\cdot n \rangle + (\nabla\cdot v, \chi) = 0,
+\end{align*}
+
 The target functional is considered as 
@@
 \[
-K(v,p,q) = k(v,p) + \frac{\alpha}{2}||q - q_0||,
+K(q,v,p) = k(v,p) + \frac{\alpha}{2}||q - q_0||,
 \] 
@@
 \[
-K(v,p,q) \rightarrow 0 \quad \text{s.t.} \quad a(v,p)(\phi) = 0.
+K(q,v,p) \rightarrow \text{min} \quad \text{s.t.} \quad a(q,v,p)(\phi) = 0.
 \]
 
-%\subsubsection{Program description}
+\twick{
+\subsubsection{Program description}
+The implementation of this example does not introduce any new \dope{}-specific
+features but shows that more complicated equations such as the Navier-Stokes 
+system can be used as forward problem for the optimization process. 
+This example builds on the previous Example \ref{OPT_Stat_Param_Lin_Ellipt}.
+}

### 2013-12-09 Examples/OPT/StatPDE/Example7/content.tex
Update Doku und Beseitigung kleinerer Doku-typos in zwei main files.

@@
 
+\subsubsection{Program description}
+\twick{
 The Problem is similar to that of {\tt OPT/StatPDE/Example1} except for the box control constraints.
+The implementation of these constraints is taken care of in the main file
+where we add a constrained description \texttt{lcc}.
+}
+
+\twickremove{Winni: in dieser Beischreibung ist 
+\[
+-500 \leq q \leq 500.
+\]
+Allerdings finde ich die '500' nirgendwo in main, localpde, or localfunctional
+file. Wie funktioniert dieses Programm'?}
+
+\subsubsection{External optimization solver}
 The problem is solved using the optimization library IPOPT that you can obtain for free. To use it 

### 2013-12-09 Examples/OPT/StatPDE/Example8/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
Update Doku und Beseitigung kleinerer Doku-typos in zwei main files.

@@
+\subsubsection{General problem description}
 This example implements the topology optimization of an MBB-Beam given in \\
@@
 the control variable can be included into the optimization call.
+
+\subsubsection{Program description}
+\twick{
+The global constraint is inserted in the main file via 
+\texttt{constraints(lcc, 1)} instead of 
+\texttt{constraints(lcc, 0)} what we had in the previous two examples.
+}

### 2013-12-09 Examples/OPT/StatPDE/Example9/content.tex
Update Doku und Beseitigung kleinerer Doku-typos in zwei main files.

@@
 \]
+
+\subsubsection{Program description}
+\twick{
+This example is similar to Example \ref{OPT_Stat_Param_Nonlin_Fluid} (both 
+based on Example \ref{OPT_Stat_Param_Lin_Ellipt}), except that we 
+have again much more complicated (nonlinear) equations. A modification 
+of this example and several numerical tests are presented 
+in \cite{RiWi13_fsi_opt}.
+}

### 2013-12-09 Examples/PDE/InstatPDE/Example6/content.tex
Update Doku und Beseitigung kleinerer Doku-typos in zwei main files.

@@
 The modeling part of this example is based on the coupled
-Biot-Lam\'e-Navier system, 
-which is a standard model in subsurface modeling. 
+Biot-Lam\'e-Navier system. The Biot system itself   
+is a standard model in subsurface modeling \cite{Biot41a, Biot41b, Biot55}. 
 Here, a reservoir (the

### 2013-12-10 Examples/OPT/InstatPDE/Example1/content.tex
Update Doku

@@
 
+\subsubsection{Program description}
 The following new things differ from the PDE-Eamples and the optimization examples with stationary 
 PDEs:\\[2mm]
-In the file \texttt{main.cc} the SpaceTimeHandler now gets an additional argument.
+\twick{
+First, we need to introduce a reasonable dual time-stepping scheme in the main
+file since we have to
+compute 
+the adjoint equation backward in time. 
+Next,} in the file \texttt{main.cc} the SpaceTimeHandler now gets an additional argument.
 In this case \texttt{DOpEtypes::initial} which specifies that the control is entering in the 

### 2013-12-10 Examples/Documentation/foreword.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 the new code did not break anything.
+
+\paragraph{Contributors \& Thanks:}
+\added{We wish to express our gratitude to those people that have contributed to the library or
+  have been providing open software to which we have provided interface.
+
+These are the developers of the software toolkits \texttt{Gascoigne} and \texttt{RoDoBo}
+\cite{Gascoigne} and \cite{rodobo-web}, as their software as well as the discussions 
+have shaped some of the initial design ideas for this library.
+
+Further, we would like the thank the developers of the \texttt{deal.II} finite element library 
+\cite{deal,DealIIReference} for providing the finite elements to which we have interfaced. 
+Similar thanks go to the developers of \texttt{Ipopt},~\cite{ipopt}.
+
+Last, but not least, we would like to thank all contributors listed in Section~\ref{sec:contrib}
+for their respective contribution to the library.
+}
+
+\todo[inline]{Wollen wir so einen Abschnitt? Fehlt da noch wer ? gcc ...?}

### 2013-12-10 Examples/Documentation/introduction.tex
Update Doku und Beseitigung kleinerer Doku-typos in zwei main files.
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 \chapter{Introduction}\label{chap:intro}
+\section{How to get \dope{}}\label{sec:obtain}
+\added{There are two ways to obtain a copy of \dope{}:\\
+%
+\begin{enumerate}[A)]
+\item You can obtain a copy of \dope{} from the developers svn repository using\\
+\texttt{svn co --username=[your username] \textbackslash \\ 
+\hspace*{5mm}https://ganymed.iwr.uni-heidelberg.de/svn/dope}\\
+(The backslash in the first line means that the line will be continued as one!)
+
+Note that you need a valid username and password. If you have none you can contact the 
+maintainer by sending an EMail to \url{dope@dopelib.net}.
+%
+\item You can download the sources as a tar-ball from the project website\\ 
+\url{http://www.dopelib.net}.
+\end{enumerate}
+}
+
 \section{License information}
@@
 
+\section{References}
+\added{
+If you like the library and use it for your own projects, please give credits by 
+referencing the project \dope{}~\cite{dope} using the following bibtex entry:
+}
+\begin{lstlisting}
+@MISC{dope,
+  key   = {DOpElib},
+  title = {The {D}ifferential {E}quation and 
+               {O}ptimization {E}nvironment: \textsc{DOpElib}},
+  url   = {http://www.dopelib.net},
+  note  = {\texttt{http://www.dopelib.net}}
+}
+\end{lstlisting}
+
 \section{Developers}
+\label{sec:contrib}
 The library is currently maintained by 
@@
 Under MAC OSX, you need to install the \texttt{XCode} tools delivered with the operating system 
-or availiable for free \texttt{developer.apple.com/xcode}. Unfortunately, the delivered 
+or available for free \texttt{developer.apple.com/xcode}. Unfortunately, the delivered 
 version of \texttt{g++} is too old, so you need to install the real thing. To do so, 
 download \texttt{MacPorts} from \texttt{macports.org}. Once you have installed MacPorts 
-you can use it to install aditional linux software. 
+you can use it to install additional Linux software. 
 First, update the MacPorts installation \texttt{sudo port selfupdate}
-after that you can install a new version of \texttt{g++} using forinstance 
+after that you can install a new version of \texttt{g++} using for instance 
 \texttt{sudo port install gcc45} to install version $4.5$ of the compiler.
@@
 \section{Installation}
-I) Get a copy of \texttt{DOpElib} from the svn repository using\\
-\texttt{svn co --username=[your username] \textbackslash \\ \hspace*{5mm}https://ganymed.iwr.uni-heidelberg.de/svn/dope}\\
-(The backslash in the first line means that the line will be continued as one!)
-Note that you need a valid username and password. If you have none you can contact the 
-maintainer by sending an EMail to \url{dope@ganymed.iwr.uni-heidelberg.de}.\\[3mm]
+\label{sec_installation}
+I) Get a copy of \texttt{DOpElib}, see Section~\ref{sec:obtain} for details.
 %
@@
   This means that your computer does not find the new compiler. Try
-  \texttt{which g++} to see whether it appears in the list of availiable 
+  \texttt{which g++} to see whether it appears in the list of available 
   compilers (but is maybe too far in the back of the list.) Then you should 
@@
   If \texttt{which g++} only returns one g++ compiler, then probably you need to
-  set an appropriate symlink. Or more robust, you can configure \texttt{dealii}
+  set an appropriate symlink. Or more robust, you can configure \texttt{deal.II}
   to use the compiler you intend by configuring deal with the right compiler. To
   do so adjust  the \texttt{CC} and \texttt{CXX} environment variable appropriately
-  before configuring \texttt{dealii}
+  before configuring \texttt{deal.II}
 
@@
   which is in fact a symlink to \texttt{/usr/bin/g++-4.2}. 
-  So that you need to insall a newer compiler this may then be found ,e.g., using macports 
-  g++ 4.5 on OSX Lion, in \texttt{/opt/local/bin/g++-mp-4.5}.
+  So that you need to install a newer compiler. You can do so, for instance
+  using macports. Then you can find, e.g., g++ version 4.5 on OSX Lion 
+  in \texttt{/opt/local/bin/g++-mp-4.5}.
 
@@
 \item \texttt{container/refinementcontainer.h} The classes defined herein are given to the \texttt{RefineSpace} method of the \texttt{SpaceTimeHandler} and determine how we define the spatial mesh (i.e.
-globally or locally with a fixed fraction, fixed number or shomehow 'optimized' strategy).
+globally or locally with a fixed fraction, fixed number or 'optimized' strategy).
 \end{itemize}
@@
   \texttt{deal.II}. The solver will build and store the stiffness matrix for the PDE.
-\item \texttt{templates/gmreslinearsolver.h} This is a wrapper for the gmres solver 
+\item \texttt{templates/gmreslinearsolver.h} This is a wrapper for the GMRES solver 
   implemented in \texttt{deal.II}. The solver will build and store the stiffness matrix 
@@
 \item \texttt{templates/voidlinearsolver.h} This is a wrapper for certain cases when we 
-  know that the matrix to be inverted is the identity. It simply copies the rhs to the
-  lhs. This is only needed for compatibility reasons some other components.
+  know that the matrix to be inverted is the identity. It simply copies the right hand side to the
+  left hand side. This is only needed for compatibility reasons some other components.
 \end{itemize}
@@
 \item \texttt{include/pointconstraintsmaker.h} This class allows to set 
-  homogeneous dirichlet values at given points/components.
+  homogeneous Dirichlet values at given points/components.
 \end{itemize}
@@
     of the initial values. This is generated by the different time-stepping schemes based upon the defined 
-    representation by the pde, which is set to the component wise $L^2$ projection by default.
+    representation by the PDE, which is set to the component wise $L^2$ projection by default.
   \item \texttt{problemdata/primaldirichletdata.h} This class contains the Dirichlet data for the 
@@
     interfaces for their initialization.
-  \item \texttt{wrapper/snopt\underline{ }wrapper.h} An interface to the SNOPT fortran library. This is an additional
+  \item \texttt{wrapper/snopt\underline{ }wrapper.h} An interface to the FORTRAN library SNOPT. 
+    This is an additional
     wrapper to the one provided by SNOPT to allow automatic construction of the functions required 
@@
 \item \texttt{include/helper.h} Collection of various small helper functions.
-\item \texttt{reducedproblems/problemcontainer\underline{ }internal.h} Houses some functions und variables common in the various problemcontainer.
+\item \texttt{reducedproblems/problemcontainer\underline{ }internal.h} Houses some functions and variables common in the various problemcontainer.
 \item \texttt{tsschemes/primal\_ts\_base.h}  This class contains the methods which all primal time stepping schemes share.

### 2013-12-10 Examples/OPT/InstatPDE/Example1/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 term and most important a time derivative leading to a 
-nonstationary optimimization problem. The governing equation is
+nonstationary optimization problem. The governing equation is
 \begin{equation*}
@@
 \end{equation*}
-with homogeneous dirichlet-data.
+with homogeneous Dirichlet-data.
 The computational domain is $\Omega \times I = [0,\pi]^2 \times [0,1]$. From the known solution, we can compute the appropriate data 
@@
 \subsubsection{Program description}
-The following new things differ from the PDE-Eamples and the optimization examples with stationary 
+The following new things differ from the PDE-Examples and the optimization examples with stationary 
 PDEs:\\[2mm]
@@
 In the file \texttt{localpde.h} we now have to specify the Methods \texttt{Init\underline{ }ElementRhs}
-and \texttt{Init\underline{ }ElementRhs\underline{ }Q}. They need to be adapted, since usualy the InitialValue for the 
+and \texttt{Init\underline{ }ElementRhs\underline{ }Q}. They need to be adapted, since usually the InitialValue for the 
 PDE is auto generated from a deal.ii function in the ProblemContainer. This Value is set in the 

### 2013-12-10 Examples/OPT/StatPDE/Example1/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

--- 2013-04-10 Examples/OPT/StatPDE/Example1/content.tex
@@
 
-The example shows how to estimate the error in the cost functional for stationary optimization
-problems.
+\added{
+\subsubsection{Background information and program description}
+In the following, we describe all extensions to the previous problems 
+relevant to solving PDE-based optimization with \dope{}. So far, 
+we had only to implement the \texttt{ElementEquation} and the corresponding
+matrix \texttt{ElementMatrix}. Now, based on the idea of the reduced cost
+functional, we have to compute certain additional equations representing the 
+adjoint, tangent, and adjoint hessian equations
+\texttt{ElementEquation\underline{ }U, ElementEquation\underline{ }UT,
+ElementEquation\underline{ }UTT}
+for the state equation and in the same terms arising from the functional
+itself. Let us shed some light into all equations by giving some background
+information and overview first.
+
+In abstract form, we are given the following optimization problem:
+\begin{equation*}
+J(q,u) \, \rightarrow \, \min , \quad a(q,u)(\psi) = 0 \quad \forall\psi\in V
+\end{equation*}
+Lagrangian:
+\begin{equation*}
+{\cal L}(q,u,z):=J(q,u) - a(q,u)(z)
+\end{equation*}
+Optimality system (KKT system):
+\begin{align*}
+a_u'(q,u)(\phi , z) &= J'_u(q,u)(\phi) \quad \forall \phi\in V \\
+a_q'(q,u)(\chi , z) &= J'_q(q,u)(\chi) \quad \forall \chi\in Q \\
+a(q,u)(\psi) &= 0 \qquad \forall\psi\in V
+\end{align*}
+or equivalently, in terms of the Lagrangian
+\begin{align*}
+{\cal L}'_u(q,u,z)(\phi) &= 0 \quad\forall \phi\in V \quad\text{(Adjoint Equation)} \\
+{\cal L}'_q(q,u,z)(\chi) &= 0 \quad\forall \chi\in V \quad\text{(Gradient Equation)} \\
+{\cal L}'_z(q,u,z)(\psi) &= 0 \quad\forall \psi\in V \quad\text{(State Equation)}
+\end{align*}
+The continuous problem is discretized by a standard Galerkin method using 
+finite dimensional subspaces $Q_h \times V_h \subset Q\times V$:
+\begin{equation*}
+J(q,u) \, \rightarrow \, \min , \quad a(q,u)(\psi) = 0 \quad \forall\psi\in V
+\end{equation*}
+Discrete saddle-point problems
+\begin{align*}
+a_u'(q_h,u_h)(\phi_h , z_h) &= J'_u(q_h,u_h)(\phi_h) \quad \forall \phi_h\in V_h \\
+a_q'(q_h,u_h)(\chi_h , z_h) &= J'_q(q_h,u_h)(\chi_h) \quad \forall \chi_h\in Q_h \\
+a(q_h,u_h)(\psi_h) &= 0 \qquad \forall\psi_h\in V_h
+\end{align*}
+\subsubsection{Solution process}
+In this section, we briefly discuss the solution process for the 
+optimization problem. For further details, we refer to the standard literature.
+The unconstrained optimal control problem is reformulated as follows. 
+We introduce the solution operator $S:Q\rightarrow V$ of the state equation.
+Then:
+\begin{equation*}
+j(q):= J(q,S(q))\, \rightarrow \, \min , \quad a(q,S(q))(\Psi) = 0 \quad 
+\forall \Psi \in V.
+\end{equation*}
+The local existence and sufficient regularity of $S$ is assumed. The 
+necessary optimality conditions of first and second order are
+\begin{equation*}
+j'(q)(\delta q) = 0 , \quad j''(q)(\delta q,\delta q) \geq 0 \quad 
+\forall\delta q\in Q.
+\end{equation*} 
+The derivatives of the reduced functional can be computed using the 
+Lagrangian
+\begin{equation*}
+{\cal L}(q,u,z) =J(q,u) - a(q,u)(z)
+\end{equation*}
+as already introduced. Let $q\in Q$, and the corresponding state $u=S(q)\in V$ 
+be given. To calculate the derivative of the reduced cost functional $j$, 
+we introduce the dual variable
+$z\in V$ solving the\\[3mm]
+%
+\textbf{Dual equation}\\
+\begin{equation*}
+{\cal L}_U'(q,U,Z)(\Psi) = 0 \quad\forall\Psi\in V.
+\end{equation*}
+Then,
+\begin{equation*}
+j'(q)(\delta q) = {\cal L}_q' (q,U,Z)(\delta q) \quad \text{for }\delta q\in Q.
+\end{equation*}
+\\[2mm]
+%
+To calculate the second derivatives, we need to solve additional equations.\\
+%
+Let $\delta q\in Q$ be a given direction. Then we search $\delta U\in V$  
+solving the\\[2mm]
+%
+\textbf{Tangent equation}\\
+%
+\begin{equation*}
+{\cal L}_{qZ}'' (q,U,Z)(\delta q,\Phi) + {\cal L}_{UZ}'' (q,U,Z)(\delta
+U,\Phi) = 0
+\quad\forall \Phi\in V.
+\end{equation*}
+Further, we have an auxilliary\\[2mm]  
+%
+\textbf{Dual for Hessian equation}
+to find $\delta Z\in V$ solving
+\begin{equation*}
+{\cal L}_{qU}'' (q,U,Z)(\delta q,\Phi) + {\cal L}_{UU}'' 
+(q,U,Z)(\delta U,\Phi) + {\cal L}_{ZU}'' (q,U,Z)(\delta Z,\Phi) =0 
+\quad\forall \Phi\in V.
+\end{equation*}
+
+Then, for $\delta r\in Q$, we can express the second derivatives of $j$ by
+\begin{align*}
+j''(q)(\delta q, \delta r) 
+&= {\cal L}_{qq}'' (q,U,Z)(\delta q,\delta r) \\
+&\;\;\;\;+ {\cal L}_{Uq}'' (q,U,Z)(\delta U,\delta r) \\ 
+&\;\;\;\;+ {\cal L}_{Zq}'' (q,U,Z)(\delta Z,\delta r).
+\end{align*}
+With these terms, we can calculate the Newton direction $\delta q$, at a given 
+iterate $q^n$, as solution to the problem
+\begin{align*}
+j''(q^n )(\delta q,\chi ) =&\, -j'(q^n)(\chi) \quad\forall\chi \in Q.
+\end{align*}
+
+Moreover, we would like to work in the Hilbert space $Q$. However, the derivative 
+$j'(q) \in H^*$ only. Hence, we need to calculate the Riesz representation 
+for the gradient $\nabla j(q) \in H$ using the definition:
+\[
+ (\nabla j(q),\delta q)_Q = j'(q)(\delta q) \quad \forall\,\delta q\in Q.
+\]
+In the given example, $Q = L^2(\Omega)$ and hence the scalar product will be the
+standard $L^2$-inner product. Simillarly, we can define the Hessian operator 
+$H(q) \in mathcal L(Q,Q)$ by defining
+\[
+ (H(q)\tau q,\delta q)_Q = j''(q)(\tau q,\delta q) \quad \forall\,\delta q,\tau q\in Q.
+\]
+}
+\todo[inline]{Brauchen wir den n\"achsten Block wirklich? Das steht 1 Seite dar\"uber schon, 
+die Notation passt aber nicht immer, z.B. $J''(u)$, fehlende Argumente und Funktionen (z.N. Dual Hessian Equation)
+. 
+Ansonsten müssen wir den Block nochmal durchgehen um die Sachen zu beheben.}
+
+Finally, we recall the representations of the previous work. First, we recall 
+in terms of the Lagrangian:
+
+\textbf{Adjoint:} Find $z\in X$ such that
+\begin{equation*}
+{\cal L}_u' (q,u,z)(\phi) = 0 \quad \forall\phi\in X. 
+\end{equation*}
+\textbf{Tangent (linearized state):} Find $\delta u\in X$ such that
+\begin{equation*}
+{\cal L}_{qz}'' (q,u,z)(\delta q, \phi) +
+{\cal L}_{uz}'' (q,u,z)(\delta u, \phi) = 0 \quad \forall\phi\in X. 
+\end{equation*}
+
+\textbf{Additional adjoint:} Find $\delta z\in X$ such that
+\begin{equation*}
+{\cal L}_{qu}'' (q,u,z)(\delta q, \phi) +
+{\cal L}_{uu}'' (q,u,z)(\delta u, \phi) +
+{\cal L}_{zu}'' (q,u,z)(\delta z, \phi) = 0 \quad \forall\phi\in X. 
+\end{equation*}
+
+Especially, the equations for stationary optimization problems are
+given as follows:
+
+\textbf{Adjoint:} Find $z\in X$ such that
+\begin{equation*}
+a_u'(q,u)(\phi,z) = J'(u)(\phi) \quad \forall\phi\in X. 
+\end{equation*}
+\textbf{Tangent:} Find $\delta u\in X$ such that
+\begin{equation*}
+a_u'(q,u)(\delta u,\phi) = -a_q'(q,u)(\delta q,\phi)  \quad \forall\phi\in X. 
+\end{equation*}
+\textbf{Additional adjoint:} Find $\delta z\in X$ such that
+\begin{equation*}
+a_u'(q,u)(\phi,\delta z) = a_{uu}''(q,u)(\delta u,\phi,z)
+-a_{qu}''(q,u)(\delta q,\phi,z) + J''(u)(\delta,\phi) \quad \forall\phi\in X. 
+\end{equation*}
+
+Using the Lagrangian, the previous expressions read
+
+\textbf{Lagrangian}
+\begin{equation*}
+{\cal L}(q,u,z) =J(q,u) - a(q,u)(z)
+\end{equation*}
+
+\textbf{State Equation}
+\begin{align*}
+a(q,u)(\phi) = 0 
+\end{align*}
+
+\textbf{Dual Equation}
+
+\begin{align*}
+{\cal L}_{u}' (q,u,z)(\phi ) &= 0 \\
+\Leftrightarrow a_u' (q,u)(\phi,z) &= J_u' (q,u)(\phi)
+\end{align*}
+
+\textbf{Tangent Equation}
+
+\begin{align*}
+{\cal L}_{qz}'' (q,u,z)(\delta q, \phi) 
++ {\cal L}_{uz}'' (q,u,z)(\delta u, \phi) &= 0 \\
+\Leftrightarrow a_u' (q,u)(\delta u,\phi) + a_q' (q,u)(\delta q,\phi) &= 0
+\end{align*}
+
+\textbf{Dual Hessian Equation}
+
+\begin{align*}
+{\cal L}_{qu}'' (q,u,z)(\delta q, \phi) 
+&+ {\cal L}_{uu}'' (q,u,z)(\delta u, \phi) \\
+&+ {\cal L}_{zu}'' (q,u,z)(\delta z, \phi) = 0 \\
+\Leftrightarrow a_u' (q,u)(\phi, \delta z) 
+&+ a_{uu}' (q,u)(\delta u,\phi, z) \\
+&+ a_{qu}' (q,u)(\delta q,\phi, z)  = J_{uu}''(q,u)(\cdot ,\cdot )
+\end{align*}
+
+\textbf{Newton left hand side: first derivative}
+
+\begin{align*}
+j'(q)(\delta q) &= {\cal L}_{q}' (q,u,z)(\delta q) \quad\text{(Gradient equation)} \\
+\Leftrightarrow j'(q)(\delta q) &= -a_q' (q,u)(\delta q,z) + J_{q}'(q,u)(\delta q)
+\end{align*}
+
+\textbf{Newton right hand side: second derivative}
+
+\begin{align*}
+j''(q)(\delta q, \delta r) &= {\cal L}_{qq}'' (q,u,z)(\delta q, \delta r) + 
+{\cal L}_{uq}'' (q,u,z)(\delta u, \delta r) \\
+&\quad +  {\cal L}_{zq}'' (q,u,z)(\delta z, \delta r) \\
+\Leftrightarrow  j''(q)(\delta q, \delta r) &= 
+J_{qq}''(q,u)(\delta q, \delta r) -a_{qq}'' (q,u)(\delta q,\delta r, z) \\
+&\quad - a_{uq}'' (q,u)(\delta u,\delta r, z) -a_{q}' (q,u)(\delta r,\delta z)  
+\end{align*}
+\todo[inline]{end of delete comment}
+
+\subsubsection{Implementation in \dope{}}
+\added{
+From the previous details, and the definition of the Lagrangian and its 
+derivatives it is clear, that the user has to provide the respective derivatives. 
+Since the Lagrangian consists of the PDE and the cost functional it 
+is sufficient to provide the respective derivatives, while \texttt{DOpElib} 
+will assemble them as required.
+Test functions for vector valued terms will be denoted by an index $i$ while 
+matrix valued terms are indexed in $i$ and $j$. Test functions in the control 
+space $Q$ are denoted as $\phi^q_i$ while those in the state space $V$ are denoted 
+as $\phi_i$.
+
+To solve the linear equations the following matrices are needed
+\begin{align*}
+\text{ElementMatrix} \quad 
+&\Leftrightarrow \quad a_{i,j} = a'_u(q,u)(\phi_j,\phi_i), \\
+\text{ControlElementMatrix } \quad 
+&\Leftrightarrow \quad a_{i,j} = (\phi^q_j, \phi^q_i)_Q.
+\end{align*}
+The first one is required for all primal and dual PDE solves, while the second one is needed 
+to calculate the Riesz representation of the derivatives of $j$. 
+If desired, the matrix for the adjoint PDEs can be provided separately as 
+\texttt{ElementMatrix\underline{ }T}, but otherwise this will be 
+calculated automatically from the primal matrix.
+
+Additional terms are needed to calculate the corresponding 
+right hand sides. These are for the PDE
+\begin{align*}
+\text{ElementEquation (state)} \quad &\Leftrightarrow \quad a(q,u)(\phi_i), \\
+\text{ElementEquation\underline{ }U (adjoint)} \quad 
+&\Leftrightarrow \quad a_u'(q,u)(\phi_i, z), \\ 
+\text{ElementEquation\underline{ }UT (tangent)} \quad 
+&\Leftrightarrow \quad a_u'(q,u)(\delta u, \phi_i), \\
+\text{ElementEquation\underline{ }UTT (adjoint hessian)} \quad 
+&\Leftrightarrow \quad  a_u'(q,u)(\phi_i, \delta z), \\
+\text{ElementEquation\underline{ }Q (gradient)} \quad 
+&\Leftrightarrow \quad a_q'(q,u)(\phi^q_i,z), \\
+\text{ElementEquation\underline{ }QT (tangent)} \quad 
+&\Leftrightarrow \quad a_q'(q,u)(\delta q, \phi_i), \\
+\text{ElementEquation\underline{ }QTT (hessian)} \quad 
+&\Leftrightarrow \quad a_{q}'(q,u)(\phi^q_i,\delta z), \\
+\text{ElementEquation\underline{ }UU (adjoint hessian)} \quad 
+&\Leftrightarrow \quad a_{uu}''(q,u)(\delta u, \phi_i, z), \\
+\text{ElementEquation\underline{ }QU (adjoint hessian)} \quad 
+&\Leftrightarrow \quad a_{qu}''(q,u)(\delta q, \phi_i, z), \\
+\text{ElementEquation\underline{ }UQ (hessian)} \quad 
+&\Leftrightarrow \quad a_{uq}''(q,u)(\delta u, \phi^q_i, z),\\
+\text{ElementEquation\underline{ }QQ (hessian)} \quad 
+&\Leftrightarrow \quad a_{qq}''(q,u)(\delta q, \phi^q_i, z),\\
+\text{ElementRightHandSide (state)} \quad 
+&\Leftrightarrow \quad f(\phi_i), \\
+\text{ControlElementEquation (gradient or hessian)} \quad 
+&\Leftrightarrow \quad (\nabla j(q),\phi^q_i)_Q.
+\end{align*}
+As for PDE problems, it is up to the user to decide if the ElementRightHandSide
+is used, or if the terms are included in the ElementEquation.
+
+For the cost functional, we have to provide
+\begin{align*}
+\text{ElementValue (all)} \quad  & \Leftrightarrow \quad J(q,u), \\
+\text{ElementValue\underline{ }U (all)} \quad  & \Leftrightarrow \quad J_u'(q,u)(\phi_i), \\
+\text{ElementValue\underline{ }Q (all)} \quad  & \Leftrightarrow \quad J_q'(q,u)(\phi_i), \\
+\text{ElementValue\underline{ }UU (all)} \quad & \Leftrightarrow \quad J_{uu}''(q,u)(\delta u, \phi_i), \\
+\text{ElementValue\underline{ }QU (all)} \quad & \Leftrightarrow \quad J_{qu}''(q,u)(\delta q, \phi_i), \\
+\text{ElementValue\underline{ }UQ (all)} \quad & \Leftrightarrow \quad J_{uq}''(q,u)(\delta u, \phi^q_i), \\
+\text{ElementValue\underline{ }QQ (all)} \quad & \Leftrightarrow \quad J_{qq}''(q,u)(\delta q, \phi^q_i).
+\end{align*}
+
+Clearly, if the PDE or cost functional contains other terms, such as boundary or face integrals 
+corresponding derivatives must be provided as well.
+
+\subsubsection{Back to the specific equations in this example}
+We have
+\todo[inline]{Wollen wir die alle? Die Terme $a'_u$ sind ja alle Gleich bis auf das 
+jeweilige argument...}
+\begin{align*}
+a(q,u)(\phi) &= (\nabla u, \nabla\phi) - (q + f, \phi),\\
+a_u'(q,u)(\phi, z) &= (\nabla\phi, \nabla z),\\
+a_u'(q,u)(\delta u, \phi) &= (\nabla\delta u, \nabla\phi),\\
+a_u'(q,u)(\phi, \delta z) &= (\nabla\phi, \nabla\delta z),\\
+a_q'(q,u)(\delta q, \phi) &= -(z,\psi^q),\\
+a_q'(q,u)(\delta q, \phi) &= (\delta q,\phi),\\
+a_{q}'(q,u)(\delta q,\delta z) &= - (\delta z,\psi^q).
+\end{align*}
+%
+For the cost functional, we have the following terms:
+\begin{align*}
+J(q,u) &= \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2,\\
+J_u'(q,u)(\phi) = (u-u^d, \phi),\\
+J_q'(q,u)(\phi) = \alpha (q,\psi^q),\\
+J_{uu}''(q,u)(\psi, \phi) = (\delta u, \phi).
+\end{align*}
+All other terms, specifically mixed terms with $QU$ etc. are zero in this
+example.
+
+\subsubsection{\texttt{main.cc}}
+Finally, the main file of the optimization examples does not look very much 
+different than for pure PDE computations - which is one of the crucial aims 
+of our library. 
+Here, instead of using a \texttt{pdeproblemcontainer}, we use 
+now an \texttt{optproblemcontainer} which can assemble all additionally needed informations,
+such as adjoint and tangent PDEs. 
+Furthermore, we define \texttt{ReducedNewtonAlgorithm} and 
+\texttt{ReducedTrustregion\underline{ }NewtonAlgorithm}
+to solve the optimization problem with a linesearch and a trust-region Newton algorithm.
+Of course one would be sufficient, but we wanted to show how to change optimization 
+solvers easily using \texttt{DOpElib}.
+
+Next, in the body of the main file, we introduce a second FE function for the 
+control variable. Then, we define 
+a \texttt{COSTFUNCTIONAL}. Finally, the problem is either solved 
+by calling \texttt{Alg.Solve(q)} and/or the user might check if the 
+derivatives are implemented correctly by calling 
+\texttt{Alg.CheckGrads} or \texttt{Alg.CheckHessian}. The latter 
+two functionalities are highly recommended to check your
+implementation before wondering about your results.
+
+}

### 2013-12-10 Examples/OPT/StatPDE/Example2/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 \end{align*}
-on the domain $\Omega = [0,1]^2$, with zero dirichlet boundary conditions and
+on the domain $\Omega = [0,1]^2$, with zero Dirichlet boundary conditions and
 \begin{itemize}
@@
 the three observation points) where we use the \texttt{FE\underline{}Nothing}
-element to assigne the three controls. Here, the number of components equals
+element to assign the three controls. Here, the number of components equals
 the number of controls. 
@@
 \end{itemize}
- from \texttt{FunctionalInterface}, but also all the aforementioned methods with a preceeding \text{Point} (\texttt{PointValue} etc.).
+ from \texttt{FunctionalInterface}, but also all the aforementioned methods with a preceding \text{Point} (\texttt{PointValue} etc.).
 %FE_Nothing, handle of the control erlaeutern.

### 2013-12-10 Examples/OPT/StatPDE/Example3/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 In this example, we solve a optimization problems from 
-fluid dynamcis. The configuration is similar to 
+fluid dynamics. The configuration is similar to 
 the fluid optimization problem proposed by Roland Becker 
@@
 
-The rest of the progam is similar to the previous optimization problems where
+The rest of the program is similar to the previous optimization problems where
 we formulate the state equation in a weak form $a(v,p)(\phi)$ such that the 

### 2013-12-10 Examples/OPT/StatPDE/Example4/content.tex
Update Doku und Beseitigung kleinerer Doku-typos in zwei main files.
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

--- 2012-03-15 Examples/OPT/StatPDE/Example4/content.tex
@@
 \end{gather*}
-on the domain $\Omega = [0,1]^2$. In addition we set the dirichlet data of the state on the boundary as follows
+on the domain $\Omega = [0,1]^2$. In addition we set the Dirichlet data of the state on the boundary as follows
 \begin{gather*}
@@
 with $\alpha = 10$.
+\subsubsection{Program description}
+The control in the Dirichlet boundary values is incorporated via the
+class \texttt{LocalDirichletData} which is defined in the
+file \textit{localdirichletdata.h}. The class is then given to
+the \texttt{OptProblemContainer} as a template argument. This is all that is
+needed to use the control in the Dirichlet boundary values. 
+\twick{
+In the main file and the localpde program, we work still with the 
+\texttt{FE\underline{}Nothing} element to assign the five controls.
+}

### 2013-12-10 Examples/OPT/StatPDE/Example5/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 
-The requirement for the control and state mesh is that they have a common coarse grid, so the space time handler gets only one mesh (to ensure a common coarse grid), but this gets copied internally so that we have two seperate meshes for control and state.
-We can then separetely refine the mesh for control and state (see \texttt{RefineControlSpace} and  \texttt{RefineStateSpace}).
+The requirement for the control and state mesh is that they have a common coarse grid, so the space time handler gets only one mesh (to ensure a common coarse grid), but this gets copied internally so that we have two separate meshes for control and state.
+We can then separately refine the mesh for control and state (see \texttt{RefineControlSpace} and  \texttt{RefineStateSpace}).

### 2013-12-10 Examples/OPT/StatPDE/Example9/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
     {\includegraphics[width=7cm]{Documentation/visit_fsi_1_opt_control_p3_0005_bearbeitet_Aug_15.png}}
-  \caption{Configuration of the fsi cylinder-beam-drag minimization problem. 
+  \caption{Configuration of the FSI cylinder-beam-drag minimization problem. 
 By sucking out the fluid, 
@@
 
-The rest of the progam is similar to the previous optimization problems where
+The rest of the program is similar to the previous optimization problems where
 we formulate the state equation in a weak form $a(q,U)(\phi)$ such that the 

### 2013-12-10 Examples/PDE/InstatPDE/Example1/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 
-After discretization in time, the space is treated, as ususally, with 
+After discretization in time, the space is treated, as usually, with 
 a Galerkin finite element scheme, here based on the Taylor-Hood element 
@@
 The two arguments 
-are used to distigunish between explicit 
+are used to distinguish between explicit 
 components and fully implicit components. 
@@
 \end{verbatim} 
-are used to distigungish between fully implicit 
+are used to distinguish between fully implicit 
 are other behavior.
@@
 \end{equation*}
-This behavior is useful for multi-phyics problems
+This behavior is useful for multi-physics problems
 where other solutions variables have to be considered

### 2013-12-10 Examples/PDE/InstatPDE/Example2/content.tex
Update Doku und Beseitigung kleinerer Doku-typos in zwei main files.
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 arbitrary Lagrangian-Eulerian (ALE) coordinates. 
-For mesh motion model is based on solving a biharmonic equation
+The mesh motion model is based on solving a biharmonic equation
 \cite{Wi11} rather than a linear-elastic model. 
@@
 the FSI-interface. However, this leads to 
-a degeneration of the ALE map. Methods to circument such as 
+a degeneration of the ALE map. Methods to circumvent such as 
 degeneration as long as possible are re-meshing techniques or 
@@
 \begin{verbatim}
-void CellTimeEquationExplicit (...) 
+void ElementTimeEquationExplicit (...) 
 \end{verbatim}
-to write all time derivative terms explicitely:
+to write all time derivative terms explicitly:
 \begin{equation*}

### 2013-12-10 Examples/PDE/InstatPDE/Example3/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
Content tex in PDE Instat Examples aufgeraeumt
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 we choose $\overline x \in \mathbb{R}^d_+$ and consider the computational domain $\Omega := (x_1, \overline x_1)\times \dots \times (x_d, \overline x_d)$.
-On the new part of the boundary $\Gamma$ with $\Gamma :=\{x\in \partial\Omega |\exists_{1\leq i\leq d}  x_i=\overline x_i\}$) we impose asymptotic values as dirichlet conditions.
+On the new part of the boundary $\Gamma$ with $\Gamma :=\{x\in \partial\Omega |\exists_{1\leq i\leq d}  x_i=\overline x_i\}$) we impose asymptotic values as Dirichlet conditions.
 For a put, we take $u|_\Gamma=0$.
@@
 
-\vspace{0.2cm}
-
-\textbf{Program description}
-
-\vspace{0.2cm}
+%\subsubsection{Program description}

### 2013-12-10 Examples/PDE/InstatPDE/Example5/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 
-The new feature of this example is the nonhomogeneous righthand side.
-In examples \ref{PDE_Stat_Laplace_2D} and \ref{PDE_Stat_Laplace_3D}, we regarded stationary problems with nonhomogeneous righthand sides, but up to now, we never involved the time variable into the nonhomogeneity.
+The new feature of this example is the nonhomogeneous right hand side.
+In examples \ref{PDE_Stat_Laplace_2D} and \ref{PDE_Stat_Laplace_3D}, we regarded stationary problems with nonhomogeneous right hand sides, but up to now, we never involved the time variable into the nonhomogeneity.
 To do this, \texttt{DOpElib} yields a \texttt{SetTime()} function which has to be applied in the \textit{localpde.h} file as well as at the place where the \texttt{RightHandSideFunction} class is declared (here the \textit{myfunctions.h} file.

### 2013-12-10 Examples/PDE/StatPDE/Example1/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 We split $\partial \Omega = \Gamma_D \cup \Gamma_{out}$.
-The right hand side of the channel is $\Gamma_{out}$ on which we describe the free outflow condition, on the rest of the boundary we prescribe dirichlet values (An parabolic inflow on the left hand side and zero on the upper and lower channel walls).
+The right hand side of the channel is $\Gamma_{out}$ on which we describe the free outflow condition, on the rest of the boundary we prescribe Dirichlet values (An parabolic inflow on the left hand side and zero on the upper and lower channel walls).
 We choose for simplicity $f=0$.\\
@@
 All information about the PDE problem (in the optimal control case about the constraining PDE) is included in the \textit{localpde.h} file.
-In a class called \texttt{LocalPDE}, we build up the element equation, the element matrix and element righthand side as well as the boundary equation, boundary matrix and boundary righthand side.
+In a class called \texttt{LocalPDE}, we build up the element equation, the element matrix and element right hand side as well as the boundary equation, boundary matrix and boundary right hand side.
 Later on, the integrator collects this local information and creates the global vectors and matrices.\\
@@
 The \textit{LocalPDE} is derived from a \texttt{PDEInterface} class.
-It comprises several functions which build up the element and boundary equations, matrices and righthand sides.
+It comprises several functions which build up the element and boundary equations, matrices and right hand sides.
 The weak formulation of problem \eqref{eq:strong} with $f=0$ is
@@
 \end{equation*}
-The lefthand side is implemented in the \texttt{ElementEquation} function, the righthand side is implemented in the \texttt{ElementRightHandSide} function (which is unused in this example, because $f=0$).
+The left hand side is implemented in the \texttt{ElementEquation} function, the right hand side is implemented in the \texttt{ElementRightHandSide} function (which is unused in this example, because $f=0$).
 \begin{remark}
@@
 \end{remark}
-To apply Newton's method, this problem is linearized: on the lefthand side, we have the derivative of the (semilinear) form $a(\cdot;\cdot)$ with respect to the solution variable $u$, and the righthand side is the residual of the weak formulation:
+To apply Newton's method, this problem is linearized: on the left hand side, we have the derivative of the (semilinear) form $a(\cdot;\cdot)$ with respect to the solution variable $u$, and the right hand side is the residual of the weak formulation:
 \begin{equation*}
@@
 \end{equation*}
-In the \texttt{ElementMatrix} function, we implement the following matrix $A$ as representation of the derivative on the lefthand side:
+In the \texttt{ElementMatrix} function, we implement the following matrix $A$ as representation of the derivative on the left hand side:
 \begin{equation*}

### 2013-12-10 Examples/PDE/StatPDE/Example10/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 As before, we build up the element and boundary equations and matrices in the \textit{localpde.h} file.
-Apart from using the additionally defined problem parameters and modelling compressible STVK material instead of INH material (which leads to changes in the weak formulation of the equations), there are no major differences to the corresponding file in the last example.\\
+Apart from using the additionally defined problem parameters and modeling compressible STVK material instead of INH material (which leads to changes in the weak formulation of the equations), there are no major differences to the corresponding file in the last example.\\
 In the \textit{main.cc} file, we have to include additional header files from the deal.II library concerning error estimation and grid refinement.

### 2013-12-10 Examples/PDE/StatPDE/Example11/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

--- 2013-01-08 Examples/PDE/StatPDE/Example11/content.tex
@@
 \subsubsection{General problem description}
-\index{stationary PDE}
-In this example we consider the same setting as in subsection~\ref{PDE_Stat_Stokes}, the only
-difference is that we want to employ the hp-Finite-Elements. So the equation we solve is still 
-the stationary incompressible Stokes equation \index{stationary PDE!Stokes equation}. Here,
-we use the symmetric stress tensor which has a little consequence when using 
-the do-nothing outflow condition. In strong formulation we have
-\begin{align} \label{eq:strong}
--\frac{1}{2}\nabla\cdot (\nabla v + \nabla v^{T}) + \nabla p &= f \\ 
-\nabla \cdot v &= 0 \notag
-\end{align} 
-on the domain $\Omega = [-6,6]\times [0,2]$. We choose for simplicity $f=0$.\\
+This example shows the use higher order mappings in \texttt{DOpElib}. To this end, we solve a simple Laplace equation
+\begin{align*}
+-\Delta u &= -4 && \text{ in } \Omega
+\end{align*}
+with the analytical solution
+\begin{align*}
+u = {x^2 + y^2},
+\end{align*}
+and the Dirichlet Conditions on $u=1$, where the domain is given by a circle with radius 1 and the center located in the origin.
 
-\subsubsection{Adding hp-Elements}
-One sees by comparing  the \texttt{main.cc}-file of this problem with the one of subsection~\ref{PDE_Stat_Stokes} that the change to hp-Elements is really easy.
-TODO: continue, erklaere insbesondere indexsetter.
+We compute the $L^2$-norm of the error and, additionally, we evaluate a functional which does not depend on the solution at all.
+We integrate the constant $\frac 12$ once over the boundary of the domain. The result is an approximation of $\pi$.
+
+All this is standard and would not justify an additional example, however we solve the equation and the functional not one but two
+times. First with the standard Q1-mapping, the second time we use a higher order mapping. The exact order can be determined by the 
+parameter file, the preset is Q2-mapping. At the end, we gather the errors and convergence rates over some refinement cycles
+in a nice table and notice the higher order of convergence for the higher order mapping solutions. This is due to the fact that
+we can approximate the circular boundary much better by a quadratic mapping.
+\subsubsection{Program description}
+In this section we want to focus on what you have to do if you want to enhance your existing code to use higher order mappings,
+which is actually pretty simple.
+
+You have to include the file
+\begin{verbatim}
+mapping_wrapper.h
+\end{verbatim}
+
+and create a mapping of the desired order by
+\begin{verbatim}
+  DOpEWrapper::Mapping<dim, DOFHANDLER> mapping(order_of_mapping);
+\end{verbatim}
+
+The last step is to give the mapping to the DoFHandler:
+\begin{verbatim}
+MethodOfLines_StateSpaceTimeHandler<FE, DOFHANDLER, SPARSITYPATTERN,VECTOR, 2> 
+	  DOFH(triangulation, mapping, state_fe);
+\end{verbatim}
+The rest of the program is as usual.

### 2013-12-10 Examples/PDE/StatPDE/Example12/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 An additional feature of this example is that for the solution of the PDE in mixed form we are using the 
-schur complement solver provided in \texttt{dealii} step-20. Hence this example shows how simple it is to 
+Schur complement solver provided in \texttt{dealii} step-20. Hence this example shows how simple it is to 
 use self-made linear solvers within the \texttt{DOpElib} framework.

### 2013-12-10 Examples/PDE/StatPDE/Example2/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

--- 2012-03-15 Examples/PDE/StatPDE/Example2/content.tex
@@
-\textbf{General problem description}
-
-\vspace{0.2cm}
-
-In this example we consider a simple stationary FSI problem. \index{stationary PDE} \index{stationary PDE!FSI problem} The fluid
-is given as an incompressible Newtonian fluid modelled by the Stokes equation.
-Here, we use the symmetric stress tensor which has a little consequence when using the do-nothing outflow condition.
-\index{fluid} \index{fluid!incompressible} \index{fluid!Newtonian}
-The computation domain is $\Omega = [-6,6]\times [0,2]$ and we choose for simplicity $f=0$.
-
-The fluid reads: \index{fluid!Eulerian framework}
-\begin{Problem}[Variational fluid problem, Eulerian framework]
- Find $\{v_f,p_f\} \in \{ v_f^D + V\} \times L_f$, such that,
-  \begin{equation*}
-    \begin{aligned} 
-      (\rho_f v\cdot\nabla v_f,\phi^v)_{\Omega_f} +
-      (\sigma_f,\nabla\phi^v)_{\Omega_f} 
-      &= \langle n_f\cdot g_s^\sigma ,\phi^v \rangle_{\Gamma_i}
-      &&\forall \phi^v\in V_f, \\
-      %%%
-      (\operatorname{div}\, v_f,\phi^p)_{\Omega_f} &= 0
-      &&\forall\phi^p\in L_f.
-    \end{aligned}
-  \end{equation*}
-  The Cauchy stress tensor $\sigma_f$ is given by \index{Cauchy stress tensor}
-  \begin{eqnarray}
-    \sigma_f:=-p_f I + \rho_f\nu_f (\nabla v_f+\nabla v_f^T),
-  \end{eqnarray}
-  with the fluid's density $\rho_f$  and the kinematic viscosity
-  $\nu_f$. By $n_f$ we denote the outer normal vector on $\Gamma_i$
-  and by $g^\sigma_f$ is a function which describes forces acting on the interface. These
-  will be specified in the context of fluid-structure interaction
-  models. \index{fluid-structure interaction (FSI)} \index{fluid-structure interaction (FSI)!interface}
-\end{Problem}
-We define:
-\begin{equation*}
-\hat T:=\text{id}+\hat u,\quad
-\hat F:=I+\hat\nabla \hat u,\quad
-\hat J:=\text{det}(I+\hat\nabla \hat u).
-\end{equation*}
-
-The structure equations are given by incompressible neo-Hookean material
-\index{structure} \index{structure!incompressible neo-Hookean (INH) material} \index{structure!Lagrangian framework}
-\begin{Problem}[Incompressible neo-Hookean Model (Lagrangian)]
-  \begin{equation*}
-    \begin{aligned}
-      (\hat J_s \hat\sigma_s \hat F_s^{-T},\hat\nabla\hat\phi^v)_{\hat\Omega_s}
-      &= \langle\hat J_s\hat n_s\cdot \hat g_s^\sigma\hat
-      F_s^{-T},\hat\phi^v  \rangle_{\hat\Gamma_i}
-      &&\quad\forall\hat \phi^v\in\hat V_s\\
-      %%%%%%%%%%%%%%%% 
-      (\hat v_s,\hat\phi^u)_{\hat\Omega_s}&=0
-      &&\quad\forall\hat \phi^u\in\hat V_s    ,\\
-      (\text{det}\,\hat F_s-1,\hat\phi^p)_{\hat\Omega_s} &=0
-      &&\quad\forall\hat \phi^p\in\hat L_s,\\
-    \end{aligned}
-  \end{equation*}
-  where $\rho_s$ is the solid's density, $\mu_s$ the Lam\'e
-  coefficient, $\hat n_s$ the outer normal vector at $\hat\Gamma_i$, 
-  $\hat g_s^\sigma$ the force on the interface and with  
-  \[
-  \hat\sigma_s:=-\hat p_sI+\mu_s(\hat F_s\hat F_s^T-I).
-  \]
-\end{Problem}
-
-The resulting FSI problem is then given by: \index{fluid-structure interaction (FSI)!ALE model}
-\begin{Problem}[Stationary Fluid-Structure Interaction (ALE)]
-  \begin{equation*}
-    \begin{aligned}
-      (\hat J \rho_f \hat F^{-1}\hat  v\cdot\hat\nabla \hat v,
-      \hat\phi^v)_{\hat\Omega_f}
-      + (\hat J\hat\sigma_f\hat F^{-T},\hat \nabla\hat\phi^v)_{\hat\Omega_f}\\
-      + (\hat J\hat\sigma_s\hat F^{-T},\hat \nabla\hat\phi^v)_{\hat\Omega_s}
-      &= 0&&\forall\hat\phi^v\in \hat V,\\
-      %%%%%%%%%%%% 
-      (\hat v,\hat\phi^u)_{\hat\Omega_s}
-      + (\alpha_u \hat \nabla \hat u,\hat \nabla\hat\phi^u)_{\hat\Omega_f}
-      &=0&&\forall\hat\phi^u\in \hat V,\\
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-      (\widehat{\text{div}}\,(\hat J\hat F^{-1}\hat
-      v_f),\hat\phi^p)_{\hat\Omega_f} 
-      + (\hat J - 1,\hat \phi^p)_{\hat\Omega_s}
-      &=0&&\forall\hat \phi^p\in \hat L,
-    \end{aligned}
-  \end{equation*}  
-\end{Problem}
-
-\vspace{0.2cm}
-
-\textbf{Program description}
-
-\vspace{0.2cm}
-
-There is not much difference to the prior example, so we will keep our remarks rather short.
-The \textit{functionals.h, localfunctional.h} and \textit{myfunctions.cc} files did not change at all, so we simply refer to the corresponding sections in the last example.\\
-In the \textit{localpde.h} file, all functions of the \texttt{LocalPDE} class have to be adjusted to the current FSI problem.
-This only makes the equations and matrices a little more complicated, and our solution vector now consists of five components (two velocity components of the fluid, the pressure component, and two additional displacement components for the structure variables).
-Otherwise, everything is analogous to the former example.\\
-In the \textit{main.cc} we only have to add two components to the \texttt{compmask} vector and prescribe boundary conditions for the structure variables.
-Apart from that, we define objects for the same classes as before that are even named equally and use the same solvers.\\
-Again, the solution is reached within one Newton step, and all we see from the program output is the values of the functionals.
+\subsubsection{General problem description}
+We solve the vector values Laplace equation on a quadratic domain $\Omega$ with a circular hole in the middle, i.e. in strong formulation we look for $u=(u_1,u_2)$ s.t.
+\begin{align*}
+-\nabla \cdot \nabla u =& f &&\text{in }\Omega.
+\end{align*}
+We set zero Dirichlet values on the circular boundary in the middle of the domain and periodic boundary conditions on the other parts of the boundary.
+We choose the flux over the right hand side boundary as functional.
+We choose
+\begin{align*}
+f(x,y) = \left(\cos\left(\exp(10  x)\right)  y^2x + \sin(y), \cos\left(\exp(10 * y)\right)x^2y + \sin(x)\right)
+\end{align*}
+for the right hand side.
+\subsubsection{Program description}
+This example show how to implement user defined DoF constraints.
+\texttt{DOpElib} has an interface for this, namely \texttt{UserDefinedDoFConstraints}.
+In our case, we derive the class \texttt{PeriodicityConstraints}, overwrite the method \texttt{MakeStateDoFConstraints} and give an instance of this class to SpaceTimeHandler at hand:
+\begin{verbatim}
+ 	PeriodicityConstraints<DOFHANDLER, DIM> constraints_mkr;
+ 	STH DOFH(triangulation, state_fe);
+ 	DOFH.SetUserDefinedDoFConstraints(constraints_mkr);
+\end{verbatim}
+This is all it takes. We refer to \textit{myconstraintsmaker.h} for the details of the implementation of the periodicity-constraints.

### 2013-12-10 Examples/PDE/StatPDE/Example4/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
-\textbf{General problem description}
-
-\vspace{0.2cm}
-
-In this example we consider the following benchmark problem from elasticity theory: \index{stationary PDE} \index{stationary PDE!elasticity equations}
-\begin{equation}
-   (\sigma(u),\varepsilon(\varphi)) = (g,\varphi)_{\Gamma_N}.
-\end{equation}
-Here $\tilde{\Omega}$ is a quadratic domain with side length $200$ mm, where a circular hole with radius $10$ mm around the center is cut out.
-Using symmetries of the domain, we restrict our actual computational domain $\Omega$ to the upper left quarter of $\tilde{\Omega}$.\\
-In the above equation, $\varepsilon(v) := \frac{1}{2}(\nabla v + \nabla v^T)$ is the symmetric strain tensor, and \index{strain tensor}
-\begin{equation*}
-   \sigma(v) := 2\mu \varepsilon(v)^D + \kappa \cdot tr (\varepsilon(v)) I
-\end{equation*}
-denotes the symmetric stress tensor. Here $\tau^D$ is the deviatoric part of a tensor $\tau$, in two dimensions defined as \index{deviator}
-\begin{equation*}
-   \tau^D := \tau - \frac{1}{2} tr(\tau) I,
-\end{equation*}
-and the parameters $\mu$ and $\kappa$ are chosen as $\mu = 80193.800283$ resp. $\kappa = 190937.589172$.\\
-The corner points of our computational domain are in anticlockwise order: $(0,0), (90,0)$, $(100,10), (100,100)$ and $(0,100)$.
-We prescribe homogeneous Dirichlet boundary conditions in $y$-direction between $(0,0)$ and $(90,0)$ (lower boundary part), homogeneous Dirichlet boundary conditions in $x$-direction between $(100,10)$ and $(100,100)$ (right boundary part), and we interpret the righthand side of equation $(1)$ with $g = 450$ as a boundary condition between $(0,100)$ and $(100,100)$ (upper boundary part).\\
-The goal of our computations is to match the following functional reference values taken from \textit{E.
-Stein (editor), Error-controlled Adaptive Finite Elements in Solid Mechanics, Wiley (2003), pp.
-386 - 387}:
-\begin{table}[h]
-\centering
-\begin{tabular}{lccc}    
-\hline
- Functional & $u_1$ at $(90,0)$ & $\sigma_{22}$ at $(90,0)$ & $u_2$ at $(100,100)$\\
-\hline 
- Reference value & 0.021290 & 1388.732343 & 0.20951\\
-\hline
-\end{tabular}
-\end{table}
-\begin{table}[h]
-\centering
-\begin{tabular}{lcc}    
-\hline
- Functional & $u_1$ at $(0,100)$ & $\int_{(100,100)}^{(0,100)} u_2$\\
-\hline 
- Reference value & 0.076758 & 20.40344\\
-\hline
-\end{tabular}
-\end{table}
-
-\vspace{0.2cm}
-
-\textbf{Program description}
-
-\vspace{0.2cm}
-
-From the previous examples we know how to read a grid from an \textit{.inp} file.
-The grid of our current example comes from the above mentioned benchmark problem.
-The parameter file (\textit{dope.prm}) contains the same variables as in the first two examples.\\
-Apart from different point values of derivatives of the solution, we want to evaluate an integral over part of the boundary.
-\index{functional} \index{functional!boundary integral} This is newly implemented in \textit{functionals.h}.\\
-The (linear) elasticity equation is much simpler than the FSI problems we treated before, which can be seen in \textit{localpde.h} where the \texttt{CellEquation} and \texttt{CellMatrix} are implemented.
-In principle, everything is clear from the preceding examples.\\
-The \textit{main.cc} file does not contain any innovation, either. Here, we refine the grid globally instead of using an error estimator for local refinement.\\
-The output of the program reflects again the linearity of the problem (only one Newton step is needed for solution).\\
+\subsubsection{General problem description}
+In this problem we solve the simple vector valued Laplace equation in 2d
+on the unit square $\Omega=[0,1]^2$, i.e. in strong formulation we look for $u=(u_1,u_2)$ s.t.
+\begin{align*}
+-\Delta u =& f &&\text{in }\Omega.
+\end{align*}
+We set zero Dirichlet values on $\partial \Gamma$ and choose $f=(1,1)$. The classical example of a PDE.

### 2013-12-10 Examples/PDE/StatPDE/Example5/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 \subsubsection{General problem description}
-This example shows the use of the adaptive grid refinement and error estimation by the \textit{DWR method} (For a description of the method, see \cite{BR03}.) applied to the laplace equation 
+This example shows the use of the adaptive grid refinement and error estimation by the \textit{DWR method} (For a description of the method, see \cite{BR03}.) applied to the Laplace equation 
 \begin{align*}
@@
 \end{align}
-with the error $e=u-u_h$, the Triangulation $\mathbb T_h$, the dual solution $z$, arbitraty function $\phi_h \in V_h$ (the ansatz space) and the element- and edge-residuals: 
+with the error $e=u-u_h$, the Triangulation $\mathbb T_h$, the dual solution $z$, arbitrary function $\phi_h \in V_h$ (the ansatz space) and the element- and edge-residuals: 
 \begin{align}\label{dwr.element_residual}
@@
 \item {\verb StrongElementResidual }: The element residual, see \eqref{dwr.element_residual}.
-\item {\verb StrongFaceResidual }: The terms in \eqref{dwr.error_identity} that lies in the interiour (i.e. the jumps).
+\item {\verb StrongFaceResidual }: The terms in \eqref{dwr.error_identity} that lies in the interior (i.e. the jumps).
 \item {\verb StrongBoundaryResidual }:The terms in \eqref{dwr.error_identity} that lies on the boundary (There are none in this case).
@@
 \begin{itemize}
-\item {\verb DOFH_higher_order }: With some higher order Finite Elemnts and the already defined triangulation, we build this \texttt{SpaceTimeHandler}.
+\item {\verb DOFH_higher_order }: With some higher order Finite Elements and the already defined triangulation, we build this \texttt{SpaceTimeHandler}.
 This is needed because  we want to use the patch-wise higher order interpolation of the weights.
@@
 
-Succeding the solution of the state equation 
+Succeeding the solution of the state equation 
 \begin{verbatim}

### 2013-12-10 Examples/PDE/StatPDE/Example6/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 \subsubsection{General problem description}
-In this problem we solve the simple vector valued laplace equation in 3d
+In this problem we solve the simple vector valued Laplace equation in 3d
 on the unit square $\Omega=[0,1]^3$, i.e. in strong formulation we look for $u=(u_1, u_2, u_3)$ s.t.
@@
 \end{align*}
-We set zero dirichlet values on $\partial \Gamma$ and choose $f=(1,1,1)$.
+We set zero Dirichlet values on $\partial \Gamma$ and choose $f=(1,1,1)$.
 \subsubsection{Program description}
@@
 \end{verbatim}
-With this vector we estiamte the error via \texttt{KellyErrorEstimator} and get a vector holding the estimated error per element.
+With this vector we estimate the error via \texttt{KellyErrorEstimator} and get a vector holding the estimated error per element.
 After choosing a refinement criterion (see \textit{refinementcontainer.h}, we opt here for refining the top 20\% of the elements), we give the SpaceTimeHandler an object of type \texttt{RefinementContainer} which holds all the information needed for the local mesh refinement.

### 2013-12-10 Examples/PDE/StatPDE/Example7/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 The corner points of our computational domain are in anticlockwise order: $(0,0), (90,0)$, $(100,10), (100,100)$ and $(0,100)$.
-We prescribe homogeneous Dirichlet boundary conditions in $y$-direction between $(0,0)$ and $(90,0)$ (lower boundary part), homogeneous Dirichlet boundary conditions in $x$-direction between $(100,10)$ and $(100,100)$ (right boundary part), and we interpret the righthand side of equation $(1)$ with $g = 450$ as a boundary condition between $(0,100)$ and $(100,100)$ (upper boundary part).\\
+We prescribe homogeneous Dirichlet boundary conditions in $y$-direction between $(0,0)$ and $(90,0)$ (lower boundary part), homogeneous Dirichlet boundary conditions in $x$-direction between $(100,10)$ and $(100,100)$ (right boundary part), and we interpret the right hand side of equation $(1)$ with $g = 450$ as a boundary condition between $(0,100)$ and $(100,100)$ (upper boundary part).\\
 The goal of our computations is to match the following functional reference values taken from \textit{E.

### 2013-12-10 Examples/PDE/StatPDE/Example8/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

@@
 The corner points of our computational domain are the same as before, and the boundary conditions are not altered, either.\\
-The goal of our computations is to detect a subdomain in $\Omega$ where plastic behaviour occurs (compare \textit{E.
+The goal of our computations is to detect a subdomain in $\Omega$ where plastic behavior occurs (compare \textit{E.
 Stein (editor), Error-controlled Adaptive Finite Elements in Solid Mechanics, Wiley (2003), pp.
 386 - 389}).
-This subdomain depends on the righthand side $g$ in equation $(1)$ which we write as $g = \lambda \cdot p$ with $p = 100$ and $\lambda \in [1.5;4.5]$.\\
+This subdomain depends on the right hand side $g$ in equation $(1)$ which we write as $g = \lambda \cdot p$ with $p = 100$ and $\lambda \in [1.5;4.5]$.\\
 
@@
 The functionals that appear in the output yield additional information and are not required in the above problem setting.
-The subdomain with plastic behaviour we want to detect can be visualized from the \textit{.vtk} files written to the \texttt{Results/Mesh} subfolders.
+The subdomain with plastic behavior we want to detect can be visualized from the \textit{.vtk} files written to the \texttt{Results/Mesh} subfolders.

### 2013-12-10 Examples/PDE/StatPDE/Example9/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

--- 2012-04-17 Examples/PDE/StatPDE/Example9/content.tex
@@
 \subsubsection{General problem description}
-This example shows the use of the adaptive grid refinement and error estimation by the \textit{DWR method} (For a description of the method, see \cite{BR03}.) applied to the laplace equation 
-\begin{align*}
--\Delta u &= f && \text{ in } \Omega
-\end{align*}
-with the analytical solution
-\begin{align*}
-u = \sin\left(\frac{\pi}{x^2 + y^2}\right),
-\end{align*}
- the corresponding right hand side $f = -\Delta u$ and appropriate Dirichlet Conditions on $\partial \Omega$, where the domain is given by
-\begin{align*}
-\Omega = [-2,2]^2\setminus \overline{B}_{0.5}(0).
-\end{align*}
-We want to estimate the error in the following functional of interest
-$$\fullfunction{J}{H^1(\Omega)}{\mathbb{R}}{u}{\frac 1 {\abs{\Gamma}} \int_\Gamma u \,dx}$$
-where
-$\Gamma = \left\{(x,y)\in \mathbb R ^2\,\big | \,x=0,-2<y<0.5 \right \}$.
 
-For this setting, we have the error representation
-\begin{align}\label{dwr.error_identity}
-J(e)=\sum_{K\in \mathbb T _h} \left\{(R_h, z-\psi_h)_K + (r_h, z-\psi_h)_{\partial_K}\right\}
-\end{align}
-with the error $e=u-u_h$, the Triangulation $\mathbb T_h$, the dual solution $z$, arbitraty function $\phi_h \in V_h$ (the ansatz space) and the cell- and edge-residuals: 
-\begin{align}\label{dwr.cell_residual}
-\eval[1]{R_h}_{K} = f+\Delta u_h
-\end{align}
-resp. 
-\begin{align}\label{dwr.edge_residual}
-\eval[1]{r_h}_{\Sigma} =
-\begin{cases}
-\frac 12 [\partial_n u_h] ,&\text{if } \Sigma \subset \partial_K \setminus \partial \Omega,\\
-0,& \text{if } \Sigma \subset\partial \Omega.
-\end{cases}
-\end{align}
+In this example we consider a simple stationary FSI problem. \index{stationary PDE} \index{stationary PDE!FSI problem} The fluid
+is given as an incompressible Newtonian fluid modeled by the Stokes equation.
+Here, we use the symmetric stress tensor which has a little consequence when using the do-nothing outflow condition, see also section \ref{PDE_Stat_Stokes}.
+\index{fluid} \index{fluid!incompressible} \index{fluid!Newtonian}
+The computation domain is $\Omega = [-6,6]\times [0,2]$ and we choose for simplicity $f=0$.
 
-It holds $J(u) \approx 0.441956231972232$.
+The fluid reads: \index{fluid!Eulerian framework}
+\begin{Problem}[Variational fluid problem, Eulerian framework]
+ Find $\{v_f,p_f\} \in \{ v_f^D + V\} \times L_f$, such that,
+  \begin{equation*}
+    \begin{aligned} 
+      (\rho_f v\cdot\nabla v_f,\phi^v)_{\Omega_f} +
+      (\sigma_f,\nabla\phi^v)_{\Omega_f} 
+      &= \langle n_f\cdot g_s^\sigma ,\phi^v \rangle_{\Gamma_i}
+      &&\forall \phi^v\in V_f, \\
+      %%%
+      (\operatorname{div}\, v_f,\phi^p)_{\Omega_f} &= 0
+      &&\forall\phi^p\in L_f.
+    \end{aligned}
+  \end{equation*}
+  The Cauchy stress tensor $\sigma_f$ is given by \index{Cauchy stress tensor}
+  \begin{eqnarray}
+    \sigma_f:=-p_f I + \rho_f\nu_f (\nabla v_f+\nabla v_f^T),
+  \end{eqnarray}
+  with the fluid's density $\rho_f$  and the kinematic viscosity
+  $\nu_f$. By $n_f$ we denote the outer normal vector on $\Gamma_i$
+  and by $g^\sigma_f$ is a function which describes forces acting on the interface. These
+  will be specified in the context of fluid-structure interaction
+  models. \index{fluid-structure interaction (FSI)} \index{fluid-structure interaction (FSI)!interface}
+\end{Problem}
+We define:
+\begin{equation*}
+\hat T:=\text{id}+\hat u,\quad
+\hat F:=I+\hat\nabla \hat u,\quad
+\hat J:=\text{det}(I+\hat\nabla \hat u).
+\end{equation*}
+
+The structure equations are given by incompressible neo-Hookean material
+\index{structure} \index{structure!incompressible neo-Hookean (INH) material} \index{structure!Lagrangian framework}
+\begin{Problem}[Incompressible neo-Hookean Model (Lagrangian)]
+  \begin{equation*}
+    \begin{aligned}
+      (\hat J_s \hat\sigma_s \hat F_s^{-T},\hat\nabla\hat\phi^v)_{\hat\Omega_s}
+      &= \langle\hat J_s\hat n_s\cdot \hat g_s^\sigma\hat
+      F_s^{-T},\hat\phi^v  \rangle_{\hat\Gamma_i}
+      &&\quad\forall\hat \phi^v\in\hat V_s\\
+      %%%%%%%%%%%%%%%% 
+      (\hat v_s,\hat\phi^u)_{\hat\Omega_s}&=0
+      &&\quad\forall\hat \phi^u\in\hat V_s    ,\\
+      (\text{det}\,\hat F_s-1,\hat\phi^p)_{\hat\Omega_s} &=0
+      &&\quad\forall\hat \phi^p\in\hat L_s,\\
+    \end{aligned}
+  \end{equation*}
+  where $\rho_s$ is the solid's density, $\mu_s$ the Lam\'e
+  coefficient, $\hat n_s$ the outer normal vector at $\hat\Gamma_i$, 
+  $\hat g_s^\sigma$ the force on the interface and with  
+  \[
+  \hat\sigma_s:=-\hat p_sI+\mu_s(\hat F_s\hat F_s^T-I).
+  \]
+\end{Problem}
+
+The resulting FSI problem is then given by: \index{fluid-structure interaction (FSI)!ALE model}
+\begin{Problem}[Stationary Fluid-Structure Interaction (ALE)]
+  \begin{equation*}
+    \begin{aligned}
+      (\hat J \rho_f \hat F^{-1}\hat  v\cdot\hat\nabla \hat v,
+      \hat\phi^v)_{\hat\Omega_f}
+      + (\hat J\hat\sigma_f\hat F^{-T},\hat \nabla\hat\phi^v)_{\hat\Omega_f}\\
+      + (\hat J\hat\sigma_s\hat F^{-T},\hat \nabla\hat\phi^v)_{\hat\Omega_s}
+      &= 0&&\forall\hat\phi^v\in \hat V,\\
+      %%%%%%%%%%%% 
+      (\hat v,\hat\phi^u)_{\hat\Omega_s}
+      + (\alpha_u \hat \nabla \hat u,\hat \nabla\hat\phi^u)_{\hat\Omega_f}
+      &=0&&\forall\hat\phi^u\in \hat V,\\
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+      (\widehat{\text{div}}\,(\hat J\hat F^{-1}\hat
+      v_f),\hat\phi^p)_{\hat\Omega_f} 
+      + (\hat J - 1,\hat \phi^p)_{\hat\Omega_s}
+      &=0&&\forall\hat \phi^p\in \hat L,
+    \end{aligned}
+  \end{equation*}  
+\end{Problem}
+
 \subsubsection{Program description}
-In this section we want to focus on what you have to do if you want to enhance your existing code to use the \textit{DWR method}.
 
-First, additionally to all the things one has to do when just solving the equation, we have to include the file
-\begin{verbatim}
-higher_order_dwrc.h
-\end{verbatim}
-
-As we approximate the so called 'weights' $z-\phi_h$ in the error representation by a patchwise higher order interpolation of $z_h$ (the computed dual solution), we have to enforce patch-wise refinement of the grid by giving the flag
-\begin{verbatim}
-Triangulation<2>::MeshSmoothing::patch_level_1
-\end{verbatim}
-to the triangulation.
-
-To be able to solve the adjoint equation for the error estimation one needs to implement some methods regarding the equation as well es the functional of interest:
-\begin{itemize}
-\item In {\verb pdeinterface.h }
-\begin{itemize}
-\item {\verb CellEquation_U }: Weak form of the adjoint equation.
-\item {\verb CellMatrix_T }: The FE matrix for the adjoint problem.
-\item {\verb FaceEquation_U }: This one is needed in this case here because we have a functional of interest that lives on faces. 
-\end{itemize}
-\item {\verb functionalinterface.h }
-\begin{itemize}
-\item {\verb FaceValue_U }: This is the right hand side of the adjoint equation.
-\end{itemize}
-\end{itemize}
-During the  evaluation of \eqref{dwr.error_identity}, the following methods are needed
-\begin{itemize}
-\item {\verb StrongCellResidual }: The cell residual, see \eqref{dwr.cell_residual}.
-\item {\verb StrongFaceResidual }: The terms in \eqref{dwr.error_identity} that lies in the interiour (i.e. the jumps).
-\item {\verb StrongBoundaryResidual }:The terms in \eqref{dwr.error_identity} that lies on the boundary (There are none in this case).
-\end{itemize}
-
-After this, we tell the problem which functional we want to use for the error estimation, this is done via
-\begin{verbatim}
-P.SetFunctionalForErrorEstimation(LFF.GetName())
-\end{verbatim}
-where \texttt{P} is of type \texttt{PDEProblemContainer} and \texttt{LFF} is the desired functional of interest.
-
-The next thing we need is an object of the type
-\begin{verbatim}
-HigherOrderDWRContainer
-\end{verbatim}
-This container takes care of the computation of the weights.
-
-To build this, we need the following:
-\begin{itemize}
-\item {\verb DOFH_higher_order }: With some higher order Finite Elemnts and the already defined triangulation, we build this \texttt{SpaceTimeHandler}.
-This is needed because  we want to use the patch-wise higher order interpolation of the weights.
-\item {\verb idc_high }: A \texttt{IntegratorDataContainer} in which we put some (face)quadrature formulas for the evaluation of the error Identity.
-\item A string which indicates how we want to store the weight-vectors (here: { \verb "fullmem" }).
-\item \texttt{pr}: The \texttt{ParameterReader} which we have already defined.
-\item A enum of type \texttt{EETerms} that tells the container, which error terms we want to compute (primal error indicators vs. dual error indicators, see \cite{BR03}).  
-\end{itemize}
-The last preparation step is now to initialize the \texttt{DWRDataContainer} with the problem in use:
-\begin{verbatim}
-  solver.InitializeHigherOrderDWRC(dwrc);
-\end{verbatim}
-
-Succeding the solution of the state equation 
-\begin{verbatim}
-solver.ComputeReducedFunctionals(); 
-\end{verbatim}
-we compute the error indicators by calling 
-\begin{verbatim}
-solver.ComputeRefinementIndicators(dwrc);
-\end{verbatim}
-We can now get the error indicators (with signs!) out of \texttt{dwrc} by 
-\begin{verbatim}
-dwrc.GetErrorIndicators();
-\end{verbatim}
-With these indicators, we are now able to refine our grid adaptively (there are several mesh adaption strategies implemented, like 'optimized', 'fixednumber' or 'fixedfraction')
-\begin{verbatim}
-DOFH.RefineSpace("optimized", &error_ind);
-\end{verbatim}
-Note, that one has to take the norm of each entry in the vector of the error indicators before feeding them into the \texttt{RefineSpace} method.
+There is not much difference to the prior example, so we will keep our remarks rather short.
+The \textit{functionals.h} and \textit{myfunctions.cc} files did not change at all, so we simply refer to the corresponding sections in the last example.\\
+In the \textit{localpde.h} file, all functions of the \texttt{LocalPDE} class have to be adjusted to the current FSI problem.
+This only makes the equations and matrices a little more complicated, and our solution vector now consists of five components (two velocity components of the fluid, the pressure component, and two additional displacement components for the structure variables).
+Otherwise, everything is analogous to the former example.\\
+In the \textit{main.cc} we only have to add two components to the \texttt{compmask} vector and prescribe boundary conditions for the structure variables.
+Apart from that, we define objects for the same classes as before that are even named equally and use the same solvers.\\
+Again, the solution is reached within one Newton step, and all we see from the program output is the values of the functionals.

### 2013-12-10 Examples/OPT/StatPDE/Example8/content.tex
Finished exchange of the examples

@@
 This example implements the topology optimization of an MBB-Beam given in \\
-{\tt OPT/StatPDE/Example6} using the SIMP method.
+{\tt OPT/StatPDE/Example7} using the SIMP method.
 
-The solution is computed using the commercial optimization library SNOPT as in 
-{\tt OPT/StatPDE/Example7}. This Example demonstrate how global constraints on
+The solution is computed using the commercial optimization library SNOPT, similar 
+to {\tt OPT/StatPDE/Example6} where IPOPT is used. 
+This Example demonstrate how global constraints on
 the control variable can be included into the optimization call.

### 2013-12-10 Examples/Documentation/foreword.tex
Update Doku. Habe Winnis-Superfunktionen fuer den Moment auskommentiert, da ich trotz tex-Update sonst nicht kompilieren kann

@@
 
-\twick{This is copied from the paper:
 The main feature of \dope{} is to give a unified interface to high level 
@@
 optimization problems with PDE constraints is an innovative feature in \dope{}.
-}
-
-\twickremove{Please check if we can remove part and use the red part!
-While deal.II leaves much of the work and many
-decisions to the user, DOpE intends to be user-optimized by delivering
-prefabricated tools which require from the user only adjustments connected
-to his specific problem. The solution of optimal control problems with PDE
-constraints is an innovation in the DOpE framework.
-}
 
@@
 \paragraph{Contributors \& Thanks:}
-\added{We wish to express our gratitude to those people that have contributed to the library or
+{We wish to express our gratitude to those people that have contributed to the library or
   have been providing open software to which we have provided interface.
@@
 }
-
-\todo[inline]{Wollen wir so einen Abschnitt? Fehlt da noch wer ? gcc ...?}

### 2013-12-10 Examples/Documentation/introduction.tex
Update Doku. Habe Winnis-Superfunktionen fuer den Moment auskommentiert, da ich trotz tex-Update sonst nicht kompilieren kann

@@
 \section{How to get \dope{}}\label{sec:obtain}
-\added{There are two ways to obtain a copy of \dope{}:\\
-%
+There are two ways to obtain a copy of \dope{}:\\
+
 \begin{enumerate}[A)]
@@
 \end{enumerate}
-}
 
@@
 \section{References}
-\added{
 If you like the library and use it for your own projects, please give credits by 
 referencing the project \dope{}~\cite{dope} using the following bibtex entry:
-}
+
 \begin{lstlisting}
@@
 \label{sec_installation}
-I) Get a copy of \texttt{DOpElib}, see Section~\ref{sec:obtain} for details.
+I) Get a copy of \texttt{DOpElib}, see Section~\ref{sec:obtain} for details.\\[1mm]
 %

### 2013-12-10 Examples/OPT/InstatPDE/Example1/content.tex
Update Doku. Habe Winnis-Superfunktionen fuer den Moment auskommentiert, da ich trotz tex-Update sonst nicht kompilieren kann

@@
 PDEs:\\[2mm]
-\twick{
 First, we need to introduce a reasonable dual time-stepping scheme in the main
@@
 the adjoint equation backward in time. 
-Next,} in the file \texttt{main.cc} the SpaceTimeHandler now gets an additional argument.
+Next, in the file \texttt{main.cc} the SpaceTimeHandler now gets an additional argument.
 In this case \texttt{DOpEtypes::initial} which specifies that the control is entering in the 

### 2013-12-10 Examples/OPT/StatPDE/Example1/content.tex
Update Doku. Habe Winnis-Superfunktionen fuer den Moment auskommentiert, da ich trotz tex-Update sonst nicht kompilieren kann

@@
 \subsubsection{General problem description}
-\twick{Dieses Beispiel ist nun extremst ausfuehrlich hinsichtlich der
-localpde. Da dies ein Manual ist, wuensche ich mir, dass wir dies (respective
-moeglicher Rechtschreibfehler und typos) auch so belassen. Mir hat das damals 
-bei FSI Opt sehr geholfen.}
 This example solves a distributed minimization problem
-\twick{and shows how to estimate the error in the cost functional for stationary optimization
-problems. The problem reads:}
+and shows how to estimate the error in the cost functional for stationary optimization
+problems. The problem reads:
 \begin{gather*}
@@
 
-\added{
+{
 \subsubsection{Background information and program description}
@@
 }
-\todo[inline]{Brauchen wir den n\"achsten Block wirklich? Das steht 1 Seite dar\"uber schon, 
-die Notation passt aber nicht immer, z.B. $J''(u)$, fehlende Argumente und Funktionen (z.N. Dual Hessian Equation)
-. 
-Ansonsten müssen wir den Block nochmal durchgehen um die Sachen zu beheben.}
-
-Finally, we recall the representations of the previous work. First, we recall 
-in terms of the Lagrangian:
-
-\textbf{Adjoint:} Find $z\in X$ such that
-\begin{equation*}
-{\cal L}_u' (q,u,z)(\phi) = 0 \quad \forall\phi\in X. 
-\end{equation*}
-\textbf{Tangent (linearized state):} Find $\delta u\in X$ such that
-\begin{equation*}
-{\cal L}_{qz}'' (q,u,z)(\delta q, \phi) +
-{\cal L}_{uz}'' (q,u,z)(\delta u, \phi) = 0 \quad \forall\phi\in X. 
-\end{equation*}
-
-\textbf{Additional adjoint:} Find $\delta z\in X$ such that
-\begin{equation*}
-{\cal L}_{qu}'' (q,u,z)(\delta q, \phi) +
-{\cal L}_{uu}'' (q,u,z)(\delta u, \phi) +
-{\cal L}_{zu}'' (q,u,z)(\delta z, \phi) = 0 \quad \forall\phi\in X. 
-\end{equation*}
-
-Especially, the equations for stationary optimization problems are
-given as follows:
-
-\textbf{Adjoint:} Find $z\in X$ such that
-\begin{equation*}
-a_u'(q,u)(\phi,z) = J'(u)(\phi) \quad \forall\phi\in X. 
-\end{equation*}
-\textbf{Tangent:} Find $\delta u\in X$ such that
-\begin{equation*}
-a_u'(q,u)(\delta u,\phi) = -a_q'(q,u)(\delta q,\phi)  \quad \forall\phi\in X. 
-\end{equation*}
-\textbf{Additional adjoint:} Find $\delta z\in X$ such that
-\begin{equation*}
-a_u'(q,u)(\phi,\delta z) = a_{uu}''(q,u)(\delta u,\phi,z)
--a_{qu}''(q,u)(\delta q,\phi,z) + J''(u)(\delta,\phi) \quad \forall\phi\in X. 
-\end{equation*}
-
-Using the Lagrangian, the previous expressions read
-
-\textbf{Lagrangian}
-\begin{equation*}
-{\cal L}(q,u,z) =J(q,u) - a(q,u)(z)
-\end{equation*}
-
-\textbf{State Equation}
-\begin{align*}
-a(q,u)(\phi) = 0 
-\end{align*}
-
-\textbf{Dual Equation}
-
-\begin{align*}
-{\cal L}_{u}' (q,u,z)(\phi ) &= 0 \\
-\Leftrightarrow a_u' (q,u)(\phi,z) &= J_u' (q,u)(\phi)
-\end{align*}
-
-\textbf{Tangent Equation}
-
-\begin{align*}
-{\cal L}_{qz}'' (q,u,z)(\delta q, \phi) 
-+ {\cal L}_{uz}'' (q,u,z)(\delta u, \phi) &= 0 \\
-\Leftrightarrow a_u' (q,u)(\delta u,\phi) + a_q' (q,u)(\delta q,\phi) &= 0
-\end{align*}
-
-\textbf{Dual Hessian Equation}
-
-\begin{align*}
-{\cal L}_{qu}'' (q,u,z)(\delta q, \phi) 
-&+ {\cal L}_{uu}'' (q,u,z)(\delta u, \phi) \\
-&+ {\cal L}_{zu}'' (q,u,z)(\delta z, \phi) = 0 \\
-\Leftrightarrow a_u' (q,u)(\phi, \delta z) 
-&+ a_{uu}' (q,u)(\delta u,\phi, z) \\
-&+ a_{qu}' (q,u)(\delta q,\phi, z)  = J_{uu}''(q,u)(\cdot ,\cdot )
-\end{align*}
-
-\textbf{Newton left hand side: first derivative}
-
-\begin{align*}
-j'(q)(\delta q) &= {\cal L}_{q}' (q,u,z)(\delta q) \quad\text{(Gradient equation)} \\
-\Leftrightarrow j'(q)(\delta q) &= -a_q' (q,u)(\delta q,z) + J_{q}'(q,u)(\delta q)
-\end{align*}
-
-\textbf{Newton right hand side: second derivative}
-
-\begin{align*}
-j''(q)(\delta q, \delta r) &= {\cal L}_{qq}'' (q,u,z)(\delta q, \delta r) + 
-{\cal L}_{uq}'' (q,u,z)(\delta u, \delta r) \\
-&\quad +  {\cal L}_{zq}'' (q,u,z)(\delta z, \delta r) \\
-\Leftrightarrow  j''(q)(\delta q, \delta r) &= 
-J_{qq}''(q,u)(\delta q, \delta r) -a_{qq}'' (q,u)(\delta q,\delta r, z) \\
-&\quad - a_{uq}'' (q,u)(\delta u,\delta r, z) -a_{q}' (q,u)(\delta r,\delta z)  
-\end{align*}
-\todo[inline]{end of delete comment}
 
 \subsubsection{Implementation in \dope{}}
-\added{
+{
 From the previous details, and the definition of the Lagrangian and its 
@@
 We have
-\todo[inline]{Wollen wir die alle? Die Terme $a'_u$ sind ja alle Gleich bis auf das 
-jeweilige argument...}
 \begin{align*}

### 2013-12-10 Examples/OPT/StatPDE/Example2/content.tex
Update Doku. Habe Winnis-Superfunktionen fuer den Moment auskommentiert, da ich trotz tex-Update sonst nicht kompilieren kann

@@
 \subsubsection{Program description}
-\twick{
+{
 In contrast to the first example, the control is now a discrete quantity (for

### 2013-12-10 Examples/OPT/StatPDE/Example6/content.tex
Finished overhaul of OPT StatPDE examples. 7 and 8 are untested due to missing snopt
Update Doku. Habe Winnis-Superfunktionen fuer den Moment auskommentiert, da ich trotz tex-Update sonst nicht kompilieren kann

--- 2012-03-15 Examples/OPT/StatPDE/Example6/content.tex
@@
-This example implements the minimum compliance problem for the thickness optimization 
-of an MBB-Beam. Using the MMA-Method of K. Svanberg together with an augmented Lagrangian 
-approach for the subproblems following M. Stingl.
+\subsubsection{General problem description}
+This example solves the distributed minimization problem
+\begin{gather*}
+\min J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2\\
+\text{s.t.} (\nabla u,\nabla \phi) = (q+f,\phi)\;\;\forall\,\phi \in H^1_0(\Omega)\\
+\text{s.t.} -500 \le q \le 500\;\text{a.e. in }\Omega
+\end{gather*}
+on the domain $\Omega = [0,1]^2$, and the data is chosen as follows:
+\begin{gather*}
+ f = \left(20\pi^2  \sin(4 \pi x) - \frac{1}{\alpha}  \sin(\pi x)\right) \sin(2 \pi y)\\
+ u^d = \left( 5 \pi^2 \sin(\pi x) + \sin(4 \pi x)\right)  \sin(2\pi y)
+\end{gather*}
+and $\alpha = 10^{-3}$.
+
+In addition the following functionals are evaluated:
+\begin{gather*}
+  \text{MidPoint: } u(0.5 ; 0.5)\\[2mm]
+  \text{MeanValue: }\int_\Omega u
+\end{gather*}
+
+\subsubsection{Program description}
+{
+The Problem is similar to that of {\tt OPT/StatPDE/Example1} except for the box control constraints.
+The implementation of these constraints is taken care of in the main file
+where we add a constrained description \texttt{lcc}.
+}
+
+\textcolor{blue}{Winni: Aus den numerischen 
+Ergebnissen lese ich ab, das die Kontrolle zwischen 0 und 1 liegt.
+\[
+0 \leq q \leq 1.
+\]
+ Aber ich habe keine Ahnung was die Assoziierung zu 
+\texttt{
+lcc[0][0] = 1; //each component is constrained individualy
+  lcc[0][1] = 2; // each two constraints (lower and upper bound)
+}
+ist. Auch die zugehoerigen Kommentare im main.cc file bringen mich nicht
+  wirklich
+weiter. Wieso each component? Das ist nur eine Kontrolle?! Oder meint
+  component die state-equation component? 
+Also konkret: ich kann nicht sagen, wofuer bei \texttt{lcc[0][0] = 1}  die 
+einzelnen Eintraege stehen.
+}
+
+\subsubsection{External optimization solver}
+The problem is solved using the optimization library IPOPT that you can obtain for free. To use it 
+a correct link to the ipopt library needs to be created in  {\tt DOpE/ThirdPartyLibs} by the name 
+{\tt ipopt}, i.e., you should have the file
+{\tt DOpE/ThirdPartyLibs/ipopt} pointing to the ipopt directory. If you have not done this you can compile the
+example but when running the example you will only get an error message like\\
+{\tt Warning: During execution of `Reduced\underline{ }IpoptAlgorithm::Solve`\\ the following Problem occurred!\\
+To use this algorithm you need to have IPOPT installed! \\To use this set the WITH\underline{ }IPOPT CompilerFlag.}
+
+Alternatively the commercial optimization library SNOPT can be used in this example.
+In order to use this library you need to install SNOPT on your computer and then generate a symlink to 
+the snopt directory (where you have the libs and the header files) 
+in the {\tt DOpE/ThirdPartyLibs} directory named {\tt snopt}, i.e., you should have the file
+{\tt DOpE/ThirdPartyLibs/snopt} pointing to the snopt directory. If you have not done this you can compile the
+example but when running the example you will only get an error message like\\
+{\tt Warning: During execution of `Reduced\underline{ }SnoptAlgorithm::Solve`\\ the following Problem occurred!\\
+To use this algorithm you need to have SNOPT installed! \\To use this set the WITH\underline{ }SNOPT CompilerFlag.}

### 2013-12-10 Examples/OPT/StatPDE/Example9/content.tex
Update Doku. Habe Winnis-Superfunktionen fuer den Moment auskommentiert, da ich trotz tex-Update sonst nicht kompilieren kann

@@
 \subsubsection{Program description}
-\twick{
 This example is similar to Example \ref{OPT_Stat_Param_Nonlin_Fluid} (both 
@@
 in \cite{RiWi13_fsi_opt}.
-}

### 2013-12-12 Examples/OPT/StatPDE/Example6/content.tex
Doku: Beschreibung der constraints und Rechtschreib-typo im Kommentar des Quellcodes

@@
 \subsubsection{Program description}
-{
+
+{\textcolor{red}{
 The Problem is similar to that of {\tt OPT/StatPDE/Example1} except for the box control constraints.
 The implementation of these constraints is taken care of in the main file
-where we add a constrained description \texttt{lcc}.
-}
-
-\textcolor{blue}{Winni: Aus den numerischen 
-Ergebnissen lese ich ab, das die Kontrolle zwischen 0 und 1 liegt.
+(where we add a constrained description \texttt{lcc}) and in the
+\texttt{localconstraints.h} file. This files serves to implement the actual constraints. 
+In this example, we introduce the handling of \textit{local} constraints
+whereas the mixture of local and \textit{global} constraints will be discussed
+in the next Example \ref{OPT_Stat_MBB-Beam}. 
+%
+First, we implement the upper and lower control bounds in 
+\texttt{localconstraints.h}, i.e,
 \[
-0 \leq q \leq 1.
+q_{min} \leq q \leq q_{max}.
 \]
- Aber ich habe keine Ahnung was die Assoziierung zu 
-\texttt{
-lcc[0][0] = 1; //each component is constrained individualy
-  lcc[0][1] = 2; // each two constraints (lower and upper bound)
-}
-ist. Auch die zugehoerigen Kommentare im main.cc file bringen mich nicht
-  wirklich
-weiter. Wieso each component? Das ist nur eine Kontrolle?! Oder meint
-  component die state-equation component? 
-Also konkret: ich kann nicht sagen, wofuer bei \texttt{lcc[0][0] = 1}  die 
-einzelnen Eintraege stehen.
+with $q_{min} = -500$ and $q_{max} = 500$.
+%
+Second, the \texttt{lcc} vector describes the meaning and the applicability of
+these constraints. Further information can be found
+in \texttt{basic/constraints.h}. We give here the necessary information for
+the 
+present example. We initialize \texttt{lcc(1)}, i.e., we have 1 control block.
+Then, this block is given the constant size $2$ (in the case of more than one 
+control block, each block would be given a size of $2$). The first
+entry \texttt{lcc[0][\textcolor{red}{0}] = 1} tells 
+you how many local entries in the present block are locally constrained, here 
+it is $1$ local entry. The second entry \texttt{lcc[0][\textcolor{red}{1}] =
+2}, determines the number of constraints on this local entry. Here, we impose 
+a lower and an upper bound, i.e., we give 2 constraints.
 }

### 2013-12-12 Examples/OPT/StatPDE/Example7/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented
Doku: Beschreibung der constraints und Rechtschreib-typo im Kommentar des Quellcodes

--- 2012-03-15 Examples/OPT/StatPDE/Example7/content.tex
@@
-This example solves the distributed minimization problem
-\begin{gather*}
-\min J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2\\
-\text{s.t.} (\nabla u,\nabla \phi) = (q+f,\phi)\;\;\forall\,\phi \in H^1_0(\Omega)\\
-\text{s.t.} -500 \le q \le 500\;\text{a.e. in }\Omega
-\end{gather*}
-on the domain $\Omega = [0,1]^2$, and the data is chosen as follows:
-\begin{gather*}
- f = \left(20\pi^2  \sin(4 \pi x) - \frac{1}{\alpha}  \sin(\pi x)\right) \sin(2 \pi y)\\
- u^d = \left( 5 \pi^2 \sin(\pi x) + \sin(4 \pi x)\right)  \sin(2\pi y)
-\end{gather*}
-and $\alpha = 10^{-3}$.
+\subsubsection{General problem description}
+This example implements the minimum compliance problem for the thickness optimization 
+of an MBB-Beam. Using the MMA-Method of K. Svanberg together with an augmented Lagrangian 
+approach for the subproblems following M. Stingl.
+%\todo{Gleichung, reference}
 
-In addition the following functionals are evaluated:
-\begin{gather*}
-  \text{MidPoint: } u(0.5 ; 0.5)\\[2mm]
-  \text{MeanValue: }\int_\Omega u
-\end{gather*}
+The implementation is done using the following three additional files:
 
-The Problem is similar to that of {\tt OPT/StatPDE/Example1} except for the box control constraints.
-Another new feature is the use of the commercial optimization library SNOPT. 
-In order to use this library you need to install SNOPT on your computer and then generate a symlink to 
-the snopt directory (where you have the libs and the header files) 
-in the {\tt DOpE/ThirdPartyLibs} directory named {\tt snopt}, i.e., you should have the file
-{\tt DOpE/ThirdPartyLibs/snopt} pointing to the snopt directory. If you have not done this you can compile the
-example but when running the example you will only get an error message like\\
-{\tt Warning: During execution of `Reduced\underline{ }SnoptAlgorithm::Solve`\\ the following Problem occurred!\\
-To use this algorithm you need to have SNOPT installed! \\To use this set the WITH\underline{ }SNOP CompilerFlag.}
+\begin{itemize}
+\item \texttt{generalized\underline{ }mma\underline{ }algorithm.h}
+  An implementation of the MMA-Algorithm for structural optimization using an augmented
+  Lagrangian formulation for the subproblems. The subproblem is implemented using the 
+  special purpose\\ file \texttt{augmentedlagrangianproblem.h}.
+\item \texttt{augmentedlagrangianproblem.h} The problem container 
+  for the augmented Lagrangian problem.
+\item \texttt{voidreducedproblem.h} A wrapper file that eliminates $u$ 
+  if it is not present anyways. This is used so that we can use the same routines to 
+  solve problems that have no PDE constraint. This is used to fit the augmentedlagrangian 
+  problem into our framework.
+\end{itemize}
+
+\subsubsection{Program description}
+\textcolor{red}{
+In addition to the previous Example \ref{OPT_Stat_MBB-Beam}, we consider now 
+in addition 1 global constraint, which is realized by setting $1$ in the 
+second argument of \texttt{constraints(lcc, 1)}. 
+We use 
+\texttt{localconstraints.h} and \texttt{localconstraintaccessor.h} to impose 
+all constraints. First, we have again one control block with a lower and 
+an upper bound, 
+\[
+\rho_{min} \leq q \leq \rho_{max}
+\]
+with $\rho_{min} = 10^{-4}$ and $\rho_{max} = 1$ ($\rho$ denotes the density
+of the material). These are implemented in \texttt{localconstraintaccessor.h}.
+The global constraint is the maximum volume of the material,
+which should remain constant with the value $0.54$. Its implementation is 
+provided in \texttt{localconstraints.h}.
+}

### 2013-12-12 Examples/OPT/StatPDE/Example8/content.tex
Update Doku. Habe Winnis-Superfunktionen fuer den Moment auskommentiert, da ich trotz tex-Update sonst nicht kompilieren kann
Doku: Beschreibung der constraints und Rechtschreib-typo im Kommentar des Quellcodes

@@
 \subsubsection{Program description}
-\twick{
-The global constraint is inserted in the main file via 
-\texttt{constraints(lcc, 1)} instead of 
-\texttt{constraints(lcc, 0)} what we had in the previous two examples.
+\textcolor{red}{
+In this example, we have now some local (point) constraints, local constraints and a global
+constraint. In constrast to the previous example, these 
+constraints are all taken care of in the 
+\texttt{localconstraints.h} file. 
+Specifically, the point constraints are handled by the 
+\texttt{DOpE::PointConstraints<...>} function, which is also initializad 
+in the main file. 
 }

### 2013-12-12 Examples/Documentation/examples.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented
Update Doku

@@
  \item In a first step, we copy \texttt{Example1} and renamed it, e.g., 
-\texttt{ExampleNew}. At the same time it is important to remove the repository 
-information that it is stored in the directory \texttt{.svn.}. \\
+\texttt{MyWonderfulFirstExample}. At the same time it is important to remove the repository 
+information that it is stored in the directory \texttt{.svn/}. \\
 After having reached the folder of the example in question in the terminal, 
@@
 \begin{verbatim}
-cp -r Example1 ExampleNew
-cd ExampleNew
+cp -r Example1 MyWonderfulFirstExample
+cd MyWonderfulFirstExample
 rm -rf .svn
@@
 
-  Thus we open the \texttt{Makefile} in the directory \texttt{ExampleNew}
+  Thus we open the \texttt{Makefile} in the directory \texttt{MyWonderfulFirstExample}
   to change the name of the executable. In our example you will find a line
@@
 \begin{verbatim}
-target   = $(BINDIR)/MyExampleNew
+target   = $(BINDIR)/MyWonderfulFirstExample
 \end{verbatim}
@@
 of \texttt{DOpElib}, i.e., \texttt{\$(DOpE)/bin}, under the name 
-\texttt{MyExampleNew}.
+\texttt{MyWonderfulFirstExample}.
 
@@
 \begin{verbatim}
-echo "Running Program ../../../../../bin/MyExampleNew test.prm"
-(../../../../../bin/MyExampleNew test.prm 2>&1) > /dev/null
+echo "Running Program ../../../../../bin/MyWonderfulFirstExample test.prm"
+(../../../../../bin/MyWonderfulFirstExample test.prm 2>&1) > /dev/null
 \end{verbatim}

### 2013-12-12 Examples/Documentation/foreword.tex
Update Doku

@@
 \chapter{Foreword}
-In this report, we describe the DOpE
+\textcolor{red}{Ich habe Section 2.8 etc. 'how this library is structured' zu 
+einem eigenen chapter gemacht}
+In this report, we describe the \dope{}
 \textit{Differential Equations and  Optimization Environment} project. 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The \textit{Differential Equations and  Optimization Environment} (DOpE) project is 
-mainly based on the \textit{deal.II} finite element library which has been developed
- initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{deal}.
-Its main feature is to give a unified interface to high level algorithms such as 
+\textcolor{blue}{Eigentlich kann dieser ganze paragraph weg. Im naechsten
+  Paragraphen steht das
+im Prinzip nochmal!
+Its key feature is to give a unified interface to high level algorithms such as 
 time stepping methods, nonlinear solvers and optimization routines. 
 We aim that the user should only need to write those parts
-of the code that are problem dependent while all invariant parts of the algorithms
+of code that are problem dependent while all invariant parts of the algorithms
 should be reusable without any need for further coding.
@@
 have to replace the algorithm object with an other one. 
-
-The authors acknowledge their past experience as well as discussions with 
-the authors of the libraries 
-Gascoigne/RoDoBo project, which was initiated by 
-Roland Becker, Dominik Meidner,  and Boris Vexler \cite{rodobo}. 
-From which some of the ideas to modularize the algorithms have arisen.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The aim of DOpE is to provide a software toolkit to solve forward PDE
+Another aim of \dope{} is to provide a software toolkit to solve forward PDE
 problems as well as optimal control problems constrained by PDE. The
 solution of a broad variety of PDE is possible in deal.II as well, but
-DOpE concentrates on a unified approach for both linear and nonlinear
+\dope{} concentrates on a unified approach for both linear and nonlinear
 problems by interpreting every PDE problem as nonlinear and applying a
@@
 fluid-structure interactions.
+}
 
@@
 This is in contrast to deal.II which leaves the implementation of all high-level algorithms to the user.
-The possible solution of 
-optimization problems with PDE constraints is an innovative feature in \dope{}.
+
+An innovative feature of \dope{} is to provide a software toolkit to solve forward PDE
+problems as well as optimal control problems constrained by PDE. 
+\dope{} concentrates on a unified approach for both linear and nonlinear
+problems by interpreting every PDE problem as nonlinear and applying a
+Newton method to solve it. 
+The focus is on the numerical solution of both stationary and nonstationary
+problems which come from different application fields, like elasticity and
+plasticity, fluid dynamics, and multiphysics problems such as 
+fluid-structure interactions.
 
@@
    Poisson, Navier-Stokes, plasticity, fluid-structure interaction problems,
-and finally a coupled Biot-Lam\'e-Navier system
-\item Self written line search and trust region newton algorithms for the 
+and finally a coupled Biot-Lam\'e-Navier system.
+\item Self-written line search and trust region newton algorithms for the 
    solution of optimization problems with PDEs \cite{NoWr00}
@@
   involving stationary PDE constraints.
-\item Mesh adaptation.
+\item Mesh adaptation and \textcolor{red}{goal-oriented error estimation}.
 \item Different spatial triangulations for control and state variables.
@@
 The rest of this document is structured as follows: We start with an introduction in
-Chapter~\ref{chap:intro} where you will learn what is needed to run {\tt DOpElib}. 
+Chapter~\ref{chap:intro} where you will learn what is needed to run \dope{}. 
 Further you will learn what problems we can solve and how all the different classes 
@@
 \paragraph{Contributors \& Thanks:}
-{We wish to express our gratitude to those people that have contributed to the library or
-  have been providing open software to which we have provided interface.
-
-These are the developers of the software toolkits \texttt{Gascoigne} and \texttt{RoDoBo}
-\cite{Gascoigne} and \cite{rodobo-web}, as their software as well as the discussions 
-have shaped some of the initial design ideas for this library.
-
-Further, we would like the thank the developers of the \texttt{deal.II} finite element library 
-\cite{deal,DealIIReference} for providing the finite elements to which we have interfaced. 
+\textcolor{red}{Ich habe das von oben hierhin verschoben und etwas 
+umgeschrieben:
+The \dope{} project is 
+mainly based on the \textit{deal.II} finite element library which has been developed
+ initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{deal} and 
+now maintained by \cite{deal2}. Special thanks go to them!
+In addition to deal.II, the authors 
+gratefully acknowledge their past experience as well as discussions with 
+the authors of  
+the software toolkits \texttt{Gascoigne} and \texttt{RoDoBo}
+\cite{Gascoigne} and \cite{rodobo-web}, 
+from which some of the ideas to modularize the algorithms have arisen.
 Similar thanks go to the developers of \texttt{Ipopt},~\cite{ipopt}.
-
-Last, but not least, we would like to thank all contributors listed in Section~\ref{sec:contrib}
+Last, but not least, we would like to express our gratitude 
+to all contributors listed in Section~\ref{sec:contrib}
 for their respective contribution to the library.
 }
+
+%{We wish to express our gratitude to those people that have contributed to the library or
+%  have been providing open software to which we have provided interface.
+%
+%
+%These are the developers of the software toolkits \texttt{Gascoigne} and \texttt{RoDoBo}
+%\cite{Gascoigne} and \cite{rodobo-web}, as their software as well as the discussions 
+%have shaped some of the initial design ideas for this library. 
+%
+%Further, we would like the thank the developers of the \texttt{deal.II} finite element library 
+%\cite{deal,DealIIReference} for providing the finite elements to which we have interfaced. 
+%

### 2013-12-12 Examples/Documentation/introduction.tex
Update Doku

@@
 
-Note that you need a valid username and password. If you have none you can contact the 
-maintainer by sending an EMail to \url{dope@dopelib.net}.
+Note that you need a valid username and password. If you have none, please contact the 
+maintainers by sending an EMail to \url{dope@dopelib.net}.
 %
@@
 
+\newpage
 \section{References}
@@
 an extensive tutorial where many features of deal.II are discussed in a
-well-documented example framework. In order to use DOpE, it is highly
+well-documented example framework. In order to use DOpE, it is
 recommendable to be roughly acquainted with deal.II.
@@
 \texttt{./configure --with-umfpack}
+in version up to 7.3.0. From 8.0.0 on, deal.II is compited with 
+cmake, which makes the inclusion of external packages easier.
 
@@
 
-\section{How this library is structured}
+%%%%%%%%%%%%%%%%%%
+\chapter{The Structure of \dope{}}
 This library is designed to allow easy implementation and numerical solutions 

### 2013-12-12 Examples/Documentation/tests.tex
Update Doku

@@
-\chapter{Testing examples}\label{chap:test}
+\chapter{Regression tests}\label{chap:test}
+\textcolor{red}{Ich habe die Ueberschrift geandert}
 The DOpE test suite consists of regression tests. They are run to compare 
@@
 Every command is computed via a Makefile. In the basic example 
-directory
+directory.
+
+\textcolor{red}{
+\begin{remark}
+Please keep in mind that there are two optimization examples that 
+need external packages to be able to run successfully. These tests will
+fail if these libraries are not installed!
+\end{remark}
+}
 

### 2013-12-12 Examples/OPT/StatPDE/Example1/content.tex
cosmetic changes in examples

@@
 standard $L^2$-inner product. Simillarly, we can define the Hessian operator 
-$H(q) \in mathcal L(Q,Q)$ by defining
+$H(q) \in \mathcal L(Q,Q)$ by defining
 \[
@@
 \subsubsection{Implementation in \dope{}}
-{
 From the previous details, and the definition of the Lagrangian and its 
@@
 
-To solve the linear equations the following matrices are needed
+To solve the linear equations the following matrices are needed{\color{red}
 \begin{align*}
-\text{ElementMatrix} \quad 
-&\Leftrightarrow \quad a_{i,j} = a'_u(q,u)(\phi_j,\phi_i), \\
-\text{ControlElementMatrix } \quad 
-&\Leftrightarrow \quad a_{i,j} = (\phi^q_j, \phi^q_i)_Q.
-\end{align*}
+&\texttt{ElementMatrix} 
+&\Leftrightarrow \quad a_{i,j} &= a'_u(q,u)(\phi_j,\phi_i), \\
+&\texttt{ControlElementMatrix}  
+&\Leftrightarrow \quad a_{i,j} &= (\phi^q_j, \phi^q_i)_Q.
+\end{align*}}
 The first one is required for all primal and dual PDE solves, while the second one is needed 
@@
 Additional terms are needed to calculate the corresponding 
-right hand sides. These are for the PDE
+right hand sides. These are for the PDE the following: 
+{\color{red}
 \begin{align*}
-\text{ElementEquation (state)} \quad &\Leftrightarrow \quad a(q,u)(\phi_i), \\
-\text{ElementEquation\underline{ }U (adjoint)} \quad 
-&\Leftrightarrow \quad a_u'(q,u)(\phi_i, z), \\ 
-\text{ElementEquation\underline{ }UT (tangent)} \quad 
-&\Leftrightarrow \quad a_u'(q,u)(\delta u, \phi_i), \\
-\text{ElementEquation\underline{ }UTT (adjoint hessian)} \quad 
-&\Leftrightarrow \quad  a_u'(q,u)(\phi_i, \delta z), \\
-\text{ElementEquation\underline{ }Q (gradient)} \quad 
-&\Leftrightarrow \quad a_q'(q,u)(\phi^q_i,z), \\
-\text{ElementEquation\underline{ }QT (tangent)} \quad 
-&\Leftrightarrow \quad a_q'(q,u)(\delta q, \phi_i), \\
-\text{ElementEquation\underline{ }QTT (hessian)} \quad 
-&\Leftrightarrow \quad a_{q}'(q,u)(\phi^q_i,\delta z), \\
-\text{ElementEquation\underline{ }UU (adjoint hessian)} \quad 
-&\Leftrightarrow \quad a_{uu}''(q,u)(\delta u, \phi_i, z), \\
-\text{ElementEquation\underline{ }QU (adjoint hessian)} \quad 
-&\Leftrightarrow \quad a_{qu}''(q,u)(\delta q, \phi_i, z), \\
-\text{ElementEquation\underline{ }UQ (hessian)} \quad 
-&\Leftrightarrow \quad a_{uq}''(q,u)(\delta u, \phi^q_i, z),\\
-\text{ElementEquation\underline{ }QQ (hessian)} \quad 
-&\Leftrightarrow \quad a_{qq}''(q,u)(\delta q, \phi^q_i, z),\\
-\text{ElementRightHandSide (state)} \quad 
-&\Leftrightarrow \quad f(\phi_i), \\
-\text{ControlElementEquation (gradient or hessian)} \quad 
-&\Leftrightarrow \quad (\nabla j(q),\phi^q_i)_Q.
+&\texttt{ElementEquation}\text{ (state)} &\Leftrightarrow &\quad a(q,u)(\phi_i), \\
+&\texttt{ElementRightHandSide}\text{ (state)}  &\Leftrightarrow &\quad f(\phi_i), \\
+&\texttt{ControlElementEquation}\text{ (gradient or hessian)}  &\Leftrightarrow &\quad (\nabla j(q),\phi^q_i)_Q,
 \end{align*}
-As for PDE problems, it is up to the user to decide if the ElementRightHandSide
-is used, or if the terms are included in the ElementEquation.
+as well as 
+\begin{align*}
+&\texttt{ElementEquation\underline{ }U}\text{ (adjoint)} 
+&\Leftrightarrow &\quad a_u'(q,u)(\phi_i, z),\\
+&\texttt{ElementEquation\underline{ }Q}\text{ (gradient)} 
+&\Leftrightarrow &\quad a_q'(q,u)(\phi^q_i,z), \\
+\end{align*}
+the terms
+\begin{align*}
+&\texttt{ElementEquation\underline{ }UU}\text{ (adjoint hessian)}  
+&\Leftrightarrow &\quad a_{uu}''(q,u)(\delta u, \phi_i, z),\\
+&\texttt{ElementEquation\underline{ }UQ}\text{ (hessian)}
+&\Leftrightarrow &\quad a_{uq}''(q,u)(\delta u, \phi^q_i, z),\\
+&\texttt{ElementEquation\underline{ }QU}\text{ (adjoint hessian)} 
+&\Leftrightarrow &\quad a_{qu}''(q,u)(\delta q, \phi_i, z), \\
+&\texttt{ElementEquation\underline{ }QQ}\text{ (hessian)} 
+&\Leftrightarrow &\quad a_{qq}''(q,u)(\delta q, \phi^q_i, z),\\
+&\texttt{ElementEquation\underline{ }UT}\text{ (tangent)}  
+&\Leftrightarrow &\quad a_u'(q,u)(\delta u, \phi_i), \\
+&\texttt{ElementEquation\underline{ }QT}\text{ (tangent)}  
+&\Leftrightarrow &\quad a_q'(q,u)(\delta q, \phi_i),
+\end{align*}
+and finally
+\begin{align*}
+&\texttt{ElementEquation\underline{ }UTT}\text{ (adjoint hessian)} \quad 
+&\Leftrightarrow &\quad  a_u'(q,u)(\phi_i, \delta z), \\
+&\texttt{ElementEquation\underline{ }QTT}\text{ (hessian)} \quad 
+&\Leftrightarrow &\quad a_{q}'(q,u)(\phi^q_i,\delta z).
+\end{align*}}
+As for PDE problems, it is up to the user to decide if the \texttt{ElementRightHandSide}
+is used, or if the terms are included in the \texttt{ElementEquation}.
 
-For the cost functional, we have to provide
+For the cost functional, we have to provide{\color{red}
 \begin{align*}
-\text{ElementValue (all)} \quad  & \Leftrightarrow \quad J(q,u), \\
-\text{ElementValue\underline{ }U (all)} \quad  & \Leftrightarrow \quad J_u'(q,u)(\phi_i), \\
-\text{ElementValue\underline{ }Q (all)} \quad  & \Leftrightarrow \quad J_q'(q,u)(\phi_i), \\
-\text{ElementValue\underline{ }UU (all)} \quad & \Leftrightarrow \quad J_{uu}''(q,u)(\delta u, \phi_i), \\
-\text{ElementValue\underline{ }QU (all)} \quad & \Leftrightarrow \quad J_{qu}''(q,u)(\delta q, \phi_i), \\
-\text{ElementValue\underline{ }UQ (all)} \quad & \Leftrightarrow \quad J_{uq}''(q,u)(\delta u, \phi^q_i), \\
-\text{ElementValue\underline{ }QQ (all)} \quad & \Leftrightarrow \quad J_{qq}''(q,u)(\delta q, \phi^q_i).
-\end{align*}
+&\texttt{ElementValue}\text{ (all)}   &\Leftrightarrow &\quad J(q,u), \\
+&\texttt{ElementValue\underline{ }U}\text{ (all)}    &\Leftrightarrow &\quad J_u'(q,u)(\phi_i), \\
+&\texttt{ElementValue\underline{ }Q}\text{ (all)}    &\Leftrightarrow &\quad J_q'(q,u)(\phi_i), \\
+&\texttt{ElementValue\underline{ }UU}\text{ (all)}  &\Leftrightarrow &\quad J_{uu}''(q,u)(\delta u, \phi_i), \\
+&\texttt{ElementValue\underline{ }UQ}\text{ (all)}  &\Leftrightarrow &\quad J_{uq}''(q,u)(\delta u, \phi^q_i), \\
+&\texttt{ElementValue\underline{ }QU}\text{ (all)}   &\Leftrightarrow &\quad J_{qu}''(q,u)(\delta q, \phi_i), \\
+&\texttt{ElementValue\underline{ }QQ}\text{ (all)}  &\Leftrightarrow &\quad J_{qq}''(q,u)(\delta q, \phi^q_i).
+\end{align*}}
 
@@
 J(q,u) &= \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2,\\
-J_u'(q,u)(\phi) = (u-u^d, \phi),\\
-J_q'(q,u)(\phi) = \alpha (q,\psi^q),\\
-J_{uu}''(q,u)(\psi, \phi) = (\delta u, \phi).
+J_u'(q,u)(\phi) &= (u-u^d, \phi),\\
+J_q'(q,u)(\phi) &= \alpha (q,\psi^q),\\
+J_{uu}''(q,u)(\psi, \phi) &= (\delta u, \phi).
 \end{align*}
@@
 implementation before wondering about your results.
-
-}

### 2013-12-12 Examples/OPT/StatPDE/Example2/content.tex
cosmetic changes in examples

@@
 \end{verbatim}
-This brings along that we have not only to implement 
-\begin{itemize} 
-\item \texttt{ElementValue},
-\item \texttt{ElementValue\_U},
-\item \texttt{ElementValue\_Q}, 
-\item \texttt{ElementValue\_UU},
-\item \texttt{ElementValue\_UQ},
-\item \texttt{ElementValue\_QU},
-\item \texttt{ElementValue\_QQ}
-\end{itemize}
+This brings along that we have not only to implement {\color{red}the methods
+\begin{align*} 
+ &\texttt{ElementValue},&&
+ \texttt{ElementValue\_U},\\
+ &\texttt{ElementValue\_Q}, &&
+ \texttt{ElementValue\_UU},\\
+ &\texttt{ElementValue\_UQ},&&
+ \texttt{ElementValue\_QU},\\
+ &\texttt{ElementValue\_QQ},
+\end{align*}}
  from \texttt{FunctionalInterface}, but also all the aforementioned methods with a preceding \text{Point} (\texttt{PointValue} etc.).

### 2013-12-12 Examples/Documentation/foreword.tex
Kleine Aenderungen, siehe Mail

@@
 
-\paragraph{Contributors \& Thanks:}
+%\paragraph{Contributors \& Thanks:}
+\paragraph{\color{red}Thanks:}
 \textcolor{red}{Ich habe das von oben hierhin verschoben und etwas 
@@
  initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{deal} and 
-now maintained by \cite{deal2}. Special thanks go to them!
-In addition to deal.II, the authors 
+now maintained by \cite{deal2}. Special thanks go to them!}
+
+\textcolor{red}{In addition to deal.II, the authors 
 gratefully acknowledge their past experience as well as discussions with 
@@
 from which some of the ideas to modularize the algorithms have arisen.
-Similar thanks go to the developers of \texttt{Ipopt},~\cite{ipopt}.
+Similar thanks go to the developers of \texttt{Ipopt},~\cite{ipopt}.}
+
+\textcolor{red}{
 Last, but not least, we would like to express our gratitude 

### 2013-12-12 Examples/OPT/StatPDE/Example6/content.tex
Doku anpassung

@@
 
-{\textcolor{red}{
-The Problem is similar to that of {\tt OPT/StatPDE/Example1} except for the box control constraints.
+The Problem is similar to that of {\tt OPT/StatPDE/Example1} except for the box 
+control constraints.
 The implementation of these constraints is taken care of in the main file
@@
 \texttt{localconstraints.h} file. This files serves to implement the actual constraints.
+
 In this example, we introduce the handling of \textit{local} constraints
 whereas the mixture of local and \textit{global} constraints will be discussed
-in the next Example \ref{OPT_Stat_MBB-Beam}. 
+in the Example \ref{OPT_Stat_MBB-Beam}. 
 %
@@
 \[
-q_{min} \leq q \leq q_{max}.
+q_{\min} \leq q \leq q_{\max}.
 \]
-with $q_{min} = -500$ and $q_{max} = 500$.
+with $q_{\min} = -500$ and $q_{\max} = 500$.
+The constraints are `local', by which we mean the constraints are imposed 
+on the nodal values of the control vector. Thus, in the constraint 
+description \texttt{localconstraints.h},
+these vectors are manipulated directly without additional integration. We 
+note that the constraints need to be written such, that a feasible control
+generates non positive entried, i.e., we 
+calculate the vector
+\[
+ \mathcal C(q) = \begin{pmatrix}
+   q_{\min} - q\\q-q_{\max}
+ \end{pmatrix}.
+ \]
 %
-Second, the \texttt{lcc} vector describes the meaning and the applicability of
-these constraints. Further information can be found
-in \texttt{basic/constraints.h}. We give here the necessary information for
-the 
-present example. We initialize \texttt{lcc(1)}, i.e., we have 1 control block.
-Then, this block is given the constant size $2$ (in the case of more than one 
-control block, each block would be given a size of $2$). The first
-entry \texttt{lcc[0][\textcolor{red}{0}] = 1} tells 
+Second, the \texttt{lcc} vector is used to describe the amount of unknowns that 
+need to be reserved to store the constraints, and, eventually, corresponding 
+Lagrange multipliers. 
+Further information can be found in \texttt{basic/constraints.h}. 
+In our example, we have only one block in the control (The control is stored in 
+a \texttt{deal.II::BlockVector}). Hence, the \texttt{lcc} vector has a length of 
+one. 
+The only entry \texttt{lcc[0]} is a vector of size two (in the case of more than one 
+control block, each block would be given a size of $2$). 
+Each of the two entries has a specific meaning. 
+The first entry \texttt{lcc[0][\textcolor{red}{0}] = 1} tells 
 you how many local entries in the present block are locally constrained, here 
-it is $1$ local entry. The second entry \texttt{lcc[0][\textcolor{red}{1}] =
-2}, determines the number of constraints on this local entry. Here, we impose 
+it is one local entry. ({\em Note that this entry will typically be one. 
+it is not one  would be if we have constraints of the type $q_{i} + q_{i+1} \le 1$ for each 
+even $i$, or similar combinations of multiple entries in the control vector.})
+
+The second entry \texttt{lcc[0][\textcolor{red}{1}] = 2}, determines the number of 
+constraints on this local entry. Here, we impose 
 a lower and an upper bound, i.e., we give 2 constraints.
-}
+This information tells the SpaceTimeHandler, that the vector 
+$\mathcal C(q)$ needs exactely twice the amount of unknowns as the vector $q$.
+In general, the space needed for $\mathcal C(q)$ is given as 
+\[
+\frac{\text{\texttt{lcc[0][1]}}}{\text{\texttt{lcc[0][0]}}}
+\]
+times the unknowns for the control.
 

### 2013-12-12 Examples/OPT/StatPDE/Example7/content.tex
typo
Doku anpassung

@@
 \subsubsection{Program description}
-\textcolor{red}{
-In addition to the previous Example \ref{OPT_Stat_MBB-Beam}, we consider now 
-in addition 1 global constraint, which is realized by setting $1$ in the 
-second argument of \texttt{constraints(lcc, 1)}. 
+In addition to the previous Example \ref{OPT_Stat_Box_controlconstraints_SNOPT}, 
+we consider now in addition one global constraint. To calculate the correct 
+storrage needed we use the 
+second argument of \texttt{constraints(lcc, 1)}, which is now one.
+
 We use 
@@
 The global constraint is the maximum volume of the material,
-which should remain constant with the value $0.54$. Its implementation is 
-provided in \texttt{localconstraints.h}.
-}
+which should remain constant with the value $0.5$. Its implementation is 
+provided in \texttt{localconstraints.h} where the global constraint is handled as 
+a functional, which again is normalized to be nonpositive if the control is feasible.

### 2013-12-12 Examples/Documentation/foreword.tex
Update Doku in Einklang mit Christians und Winnis Bemerkungen/Zustimmung

--- 2013-12-12 Examples/Documentation/foreword.tex
@@
 \chapter{Foreword}
-\textcolor{red}{Ich habe Section 2.8 etc. 'how this library is structured' zu 
-einem eigenen chapter gemacht}
 In this report, we describe the \dope{}
@@
 group of Rolf Rannacher.
-
-\textcolor{blue}{Eigentlich kann dieser ganze paragraph weg. Im naechsten
-  Paragraphen steht das
-im Prinzip nochmal!
-Its key feature is to give a unified interface to high level algorithms such as 
-time stepping methods, nonlinear solvers and optimization routines. 
-We aim that the user should only need to write those parts
-of code that are problem dependent while all invariant parts of the algorithms
-should be reusable without any need for further coding.
-In particular, the user should be able to switch between various different 
-algorithms without the need to rewrite the problem dependent code, though he or she will
-have to replace the algorithm object with an other one. 
-Another aim of \dope{} is to provide a software toolkit to solve forward PDE
-problems as well as optimal control problems constrained by PDE. The
-solution of a broad variety of PDE is possible in deal.II as well, but
-\dope{} concentrates on a unified approach for both linear and nonlinear
-problems by interpreting every PDE problem as nonlinear and applying a
-Newton method to solve it. 
-The focus is on the numerical solution of both stationary and nonstationary
-problems which come from different application fields, like elasticity and
-plasticity, fluid dynamics, and multiphysics problems such as 
-fluid-structure interactions.
-}
 
@@
   involving stationary PDE constraints.
-\item Mesh adaptation and \textcolor{red}{goal-oriented error estimation}.
+\item Mesh adaptation and goal-oriented error estimation with
+the dual-weighted residual (DWR) method.
 \item Different spatial triangulations for control and state variables.
@@
 
-\paragraph{Contributors \& Thanks:}
-\textcolor{red}{Ich habe das von oben hierhin verschoben und etwas 
-umgeschrieben:
+%\paragraph{Contributors \& Thanks:}
+\paragraph{Thanks:}
 The \dope{} project is 
@@
 now maintained by \cite{deal2}. Special thanks go to them!
+
 In addition to deal.II, the authors 
@@
 Similar thanks go to the developers of \texttt{Ipopt},~\cite{ipopt}.
+
 Last, but not least, we would like to express our gratitude 
@@
 for their respective contribution to the library.
-}
 

### 2013-12-12 Examples/Documentation/tests.tex
typo
Update Doku in Einklang mit Christians und Winnis Bemerkungen/Zustimmung

@@
-\chapter{Regression tests}\label{chap:test}
-\textcolor{red}{Ich habe die Ueberschrift geandert}
-The DOpE test suite consists of regression tests. They are run to compare 
+\chapter{Regression Tests}\label{chap:test}
+The \dope{} test suite consists of regression tests. They are run to compare 
 the output to previous outputs. This is useful (necessary) after 
-changing programming code anywhere in the library. If a test
-succeeds, everything is fine in the library. If not, you should not
-check in your code into DOpE. Please make sure what is going wrong and WHY!
+changing programming code anywhere in the library.
 
-Every command is computed via a Makefile. In the basic example 
+\begin{itemize}
+\item If a test
+succeeds, everything is fine in the library. 
+\item If not, you should not
+check in your code into \dope{}. 
+Please make sure what is going wrong and WHY!
+\item 
+Every command is computed via a Makefile in the basic example 
 directory.
+\end{itemize}
 
-\textcolor{red}{
 \begin{remark}
@@
 \end{remark}
-}
 

### 2013-12-12 Examples/OPT/StatPDE/Example1/content.tex
Doku anpassung
Update Doku in Einklang mit Christians und Winnis Bemerkungen/Zustimmung

@@
 \begin{equation*}
-{\cal L}_U'(q,U,Z)(\Psi) = 0 \quad\forall\Psi\in V.
+{\cal L}_u'(q,u,z)(\psi) = 0 \quad\forall\psi\in V.
 \end{equation*}
@@
 \begin{equation*}
-j'(q)(\delta q) = {\cal L}_q' (q,U,Z)(\delta q) \quad \text{for }\delta q\in Q.
+j'(q)(\delta q) = {\cal L}_q' (q,u,z)(\delta q) \quad \text{for }\delta q\in Q.
 \end{equation*}
@@
 %
-Let $\delta q\in Q$ be a given direction. Then we search $\delta U\in V$  
+Let $\delta q\in Q$ be a given direction. Then we search $\delta u\in V$  
 solving the\\[2mm]
@@
 \begin{equation*}
-{\cal L}_{qZ}'' (q,U,Z)(\delta q,\Phi) + {\cal L}_{UZ}'' (q,U,Z)(\delta
-U,\Phi) = 0
-\quad\forall \Phi\in V.
+{\cal L}_{qz}'' (q,u,z)(\delta q,\phi) + {\cal L}_{uz}'' (q,u,z)(\delta
+u,\phi) = 0
+\quad\forall \phi\in V.
 \end{equation*}
@@
 \textbf{Dual for Hessian equation}
-to find $\delta Z\in V$ solving
+to find $\delta z\in V$ solving
 \begin{equation*}
-{\cal L}_{qU}'' (q,U,Z)(\delta q,\Phi) + {\cal L}_{UU}'' 
-(q,U,Z)(\delta U,\Phi) + {\cal L}_{ZU}'' (q,U,Z)(\delta Z,\Phi) =0 
-\quad\forall \Phi\in V.
+{\cal L}_{qu}'' (q,u,z)(\delta q,\phi) + {\cal L}_{uu}'' 
+(q,u,z)(\delta u,\phi) + {\cal L}_{zu}'' (q,u,z)(\delta z,\phi) =0 
+\quad\forall \phi\in V.
 \end{equation*}
@@
 j''(q)(\delta q, \delta r) 
-&= {\cal L}_{qq}'' (q,U,Z)(\delta q,\delta r) \\
-&\;\;\;\;+ {\cal L}_{Uq}'' (q,U,Z)(\delta U,\delta r) \\ 
-&\;\;\;\;+ {\cal L}_{Zq}'' (q,U,Z)(\delta Z,\delta r).
+&= {\cal L}_{qq}'' (q,u,z)(\delta q,\delta r) \\
+&\;\;\;\;+ {\cal L}_{uq}'' (q,u,z)(\delta u,\delta r) \\ 
+&\;\;\;\;+ {\cal L}_{zq}'' (q,u,z)(\delta z,\delta r).
 \end{align*}
@@
 
-To solve the linear equations the following matrices are needed{\color{red}
+To solve the linear equations the following matrices are needed
 \begin{align*}
@@
 &\Leftrightarrow \quad a_{i,j} &= (\phi^q_j, \phi^q_i)_Q.
-\end{align*}}
+\end{align*}
 The first one is required for all primal and dual PDE solves, while the second one is needed 
@@
 right hand sides. These are for the PDE the following: 
-{\color{red}
+{
 \begin{align*}
@@
 
-For the cost functional, we have to provide{\color{red}
+For the cost functional, we have to provide{
 \begin{align*}

### 2013-12-12 Examples/OPT/StatPDE/Example3/content.tex
Update Doku. Habe Winnis-Superfunktionen fuer den Moment auskommentiert, da ich trotz tex-Update sonst nicht kompilieren kann
Update Doku in Einklang mit Christians und Winnis Bemerkungen/Zustimmung

@@
 
-\textbf{Remark:} Since, we use the 
+\begin{remark} Since, we use the 
 symmetric stress tensor, we need to subtract the nonsymmetric part 
@@
 $\diamond$
+\end{remark}
 
@@
 
-\twick{
 \subsubsection{Program description}
@@
 This example builds on the previous Example \ref{OPT_Stat_Param_Lin_Ellipt}.
-}

### 2013-12-12 Examples/OPT/StatPDE/Example6/content.tex
Update Doku in Einklang mit Christians und Winnis Bemerkungen/Zustimmung

@@
 Each of the two entries has a specific meaning. 
-The first entry \texttt{lcc[0][\textcolor{red}{0}] = 1} tells 
+The first entry \texttt{lcc[0][\textcolor{blue}{0}] = 1} tells 
 you how many local entries in the present block are locally constrained, here 
@@
 
-The second entry \texttt{lcc[0][\textcolor{red}{1}] = 2}, determines the number of 
+The second entry \texttt{lcc[0][\textcolor{blue}{1}] = 2}, determines the number of 
 constraints on this local entry. Here, we impose 

### 2013-12-13 Examples/PDE/StatPDE/Example7/content.tex
Ueberarbeitung PDE Statpde examples mit Ausnahme von 9 und 10.

@@
 
-From the previous examples we know how to read a grid from an \textit{.inp} file.
-The grid of our current example comes from the above mentioned benchmark problem.
-The parameter file (\textit{dope.prm}) contains the same variables as in the first two examples.\\
+From the previous examples we know how to read a grid from an \textit{.inp} file. The grid of our current example comes from the above mentioned benchmark problem.\\
 Apart from different point values of derivatives of the solution, we want to evaluate an integral over part of the boundary.
-\index{functional} \index{functional!boundary integral} This is newly implemented in \textit{functionals.h}.\\
-The (linear) elasticity equation is much simpler than the FSI problems we treated before, which can be seen in \textit{localpde.h} where the \texttt{ElementEquation} and \texttt{ElementMatrix} are implemented.
-In principle, everything is clear from the preceding examples.\\
-The \textit{main.cc} file does not contain any innovation, either. Here, we refine the grid globally instead of using an error estimator for local refinement.\\
+\index{functional} \index{functional!boundary integral}This is newly implemented in \textit{functionals.h}.\\
+In principle, everything is clear from the preceding examples. We refine the grid globally instead of using an error estimator for local refinement.
 The output of the program reflects again the linearity of the problem (only one Newton step is needed for solution).\\

### 2013-12-13 Examples/PDE/StatPDE/Example9/content.tex
Ueberarbeitung PDE Statpde examples mit Ausnahme von 9 und 10.

@@
 \subsubsection{General problem description}
-
-In this example we consider a simple stationary FSI problem. \index{stationary PDE} \index{stationary PDE!FSI problem} The fluid
+\index{stationary PDE}\index{stationary PDE!FSI problem} \index{fluid} \index{fluid!incompressible} \index{fluid!Newtonian}
+In this example we consider a simple stationary FSI problem. The fluid
 is given as an incompressible Newtonian fluid modeled by the Stokes equation.
 Here, we use the symmetric stress tensor which has a little consequence when using the do-nothing outflow condition, see also section \ref{PDE_Stat_Stokes}.
-\index{fluid} \index{fluid!incompressible} \index{fluid!Newtonian}
 The computation domain is $\Omega = [-6,6]\times [0,2]$ and we choose for simplicity $f=0$.

### 2013-12-13 Examples/PDE/StatPDE/Example2/functionals.h

@@ -47,7 +47,6 @@ template<
         assert(dealdim == 2);
       }
 
-      // compute drag value around cylinder
       double
       BoundaryValue(const FDC<DH, VECTOR, dealdim>& fdc)
       {
@@ -80,7 +79,6 @@ template<
         return mass_flux_stokes;
       }
 
-      //Achtung, hier kein gradient update
       UpdateFlags
       GetFaceUpdateFlags() const
       {

### 2013-12-13 Examples/PDE/InstatPDE/Example1/content.tex
PDE/InstatPDE Example 1, 3, 4, 5 ueberarbeitet

@@
 \subsubsection{General problem description}
-In this example we consider the nonstationary incompressible Navier-Stokes equation. As in the stationary PDE Example 3 \ref{PDE_Stat_FSI_STVK},
+In this example we consider the nonstationary incompressible Navier-Stokes equation. As in the stationary PDE Example \ref{PDE_Stat_Stokes},
 we use the symmetric fluid stress tensor, i.e. 
@@
 \end{align*}
-on time interval $I=[0,T]$ and the fluid benchmark domain (Schaefer/Turek 1996) 
+on time interval $I=[0,T]$ (with $T=80$) and the fluid benchmark domain (Schaefer/Turek 1996). 
 Here, we set $f=0$ and the flow is driven by a Dirichlet inflow condition.
@@
 \begin{align*}
-v^{n+1} - k\theta \left(\nabla\cdot (\nabla v^{n+1} + \nabla {v^{n+1}}^T) + \nabla p^{n+1}\right) &=
+v^{n+1} - k\theta \left(\nabla\cdot (\nabla v^{n+1} + \nabla {v^{n+1}}^T) + \nabla p^{n+1}\right) =&
 k\theta f^{n+1} + k(1-\theta)f^{n}\\
@@
 v^{n+1} - k (\nabla\cdot (\nabla v^{n+1} + \nabla {v^{n+1}}^{T}) + \nabla p^n) &=
-k\theta f^{n+1} +  v^n  \\
+kf^{n+1} +  v^n  \\
 \nabla \cdot v^{n+1} &= 0
@@
 \begin{Problem}[Backward Euler (BE) timestepping problem]
-Let $\theta =1$.Find $v:= v^{n+1}\in V$ and $p:= p^{n+1}\in L$:
+Let $\theta =1$. Find $v:= v^{n+1}\in V$ and $p:= p^{n+1}\in L$:
 \begin{align}

### 2013-12-13 Examples/PDE/InstatPDE/Example3/content.tex
PDE/InstatPDE Example 1, 3, 4, 5 ueberarbeitet

@@
 
-%\subsubsection{Program description}
+\subsubsection{Program description}
+Note that the initial conditions are only $H^1$-regular.
+Because of this, the Crank-Nicolson scheme, which is not strongly A-stable, is not suited to solve this problem.
+As we want to use a second order accurate time stepping scheme, we use the shifted Crank-Nicolson method.
+The rest of the program is as before.

### 2013-12-13 Examples/PDE/InstatPDE/Example4/content.tex
1) Minor changes in the formatting of the documentation. 2) Added documentation to PDE/StatPDE/Example9
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented
PDE/InstatPDE Example 1, 3, 4, 5 ueberarbeitet

@@
-\textbf{General problem description}
+\subsubsection{General problem description}
 
-\vspace{0.2cm}
-
-In this example we consider one of the prototypical instationary equations, the parabolic heat equation
+In this example we consider one of the prototypical instationary equations, the parabolic heat equation, i.e.
+for $x \in \Omega \subset \mathbb{R}^d, t\in I =[0,T], T\in \mathbb{R}^+,$ we search for the unknown solution $u:I \times \Omega \rightarrow \mathbb{R}$
 \begin{align*}
@@
 u(t,x)|_{\partial \Omega} &= g(t,x),\\
-u(0,x) &= u_0(x)
-\end{align*}
-with unknown solution $u:\Omega \rightarrow \mathbb{R}$, where $\Omega \subset \mathbb{R}^d$.
-In our example, we consider the simplest case $d=1$, where the Laplacian $\Delta$ reduces to $\partial_x^2$.
-The computational domain is $\Omega \times I = [0,1] \times [0,1]$.
-For further simplification, we choose the righthand side as $f=0$ as well as homogeneous Dirichlet boundary conditions ($g=0$).
+u(0,x) &= u_0(x).
+\end{align*} In our example, we consider the simplest case $d=1$, where the Laplacian $\Delta$ reduces to $\partial_x^2$.
+The computational domain is $ I\times \Omega = [0,1] \times [0,1]$.
+For further simplification, we choose the right hand side as $f=0$ as well as homogeneous Dirichlet boundary conditions ($g=0$).
 The initial condition is given by $u_0(x) = \min(x,1-x)$.
-\vspace{0.2cm}
 
-\textbf{Program description}
+\subsubsection{Program description}
 
-\vspace{0.2cm}
-
-There are few new things compared to the first two nonstationary examples.
+There are few new things compared to the other nonstationary examples.
 This is the first time we solve an equation in one spatial dimension.

### 2013-12-13 Examples/Documentation/examples.tex
Update Doku

@@
 \label{chap:howtoex}
+\section{Getting started}
+Beside the fact that \dope{} is still under development,
+it offers already various different (linear and nonlinear) 
+examples for a lots 
+of different applications in two and three dimensions; 
+we refer the reader to the 
+next two Chapters \ref{PDE} and \ref{OPT}.
+
 To implement new examples or to use existing examples 
@@
 can simply copy an existing example. In this 
-new example, own code and changes can be compiled.
+new example, own code and changes can be compiled. Here is some 
+advice to get started:
+\begin{itemize}
+\item If you are a first time user of \dope{} 
+with some numerics background, you 
+might be familiar with the Poisson (or more general Laplace) equation.
+\dope{} has it too. Check-out Example \ref{PDE_Stat_Laplace_2D},
+to see how \dope{} implements this well-known equation in two
+dimensions or \ref{PDE_Stat_Laplace_3D} for its three-dimensional version.
+\item Before you implement a new example, please check which 
+existing example might be similar to your goals and get familiar 
+to it. Then, proceed as described in Section \ref{getting_started}. 
+\end{itemize}
+\begin{remark}[Why two examples for the Laplace equation?]
+\dope{} offers dimensional-independent programming! So, we could have 
+implemented the 2d and 3d case in one file. The reason why we
+present
+to separate programs \ref{PDE_Stat_Laplace_2D} and \ref{PDE_Stat_Laplace_3D}
+is, that in the 3d case, we show additional 
+capabilities of the library itself; specifically how to use different
+iterative linear solvers.
+\end{remark}
 
+\section{Creating new examples}
+\label{getting_started}
 Before being able to change and compile the new code, the user must 

### 2013-12-13 Examples/Documentation/foreword.tex
Update Doku

@@
 \item All finite elements of from deal.II including hp-support.
-\item Several examples showing the solution of several PDEs including
-   Poisson, Navier-Stokes, plasticity, fluid-structure interaction problems,
-and finally a coupled Biot-Lam\'e-Navier system.
 \item Self-written line search and trust region newton algorithms for the 
@@
 \item Different spatial triangulations for control and state variables.
+\item Several examples showing the solution of several PDEs including
+   Poisson, Navier-Stokes, plasticity, fluid-structure interaction problems,
+a coupled Biot-Lam\'e-Navier system, and finally from financial mathematics
+the Black-Scholes equations.
 \end{itemize}

### 2013-12-13 Examples/PDE/InstatPDE/Example2/content.tex
Update Doku

@@
       F^{-T},\hat\nabla\hat\psi^v)_{\hat\Omega_f}
+      - \langle \hat g, \hat\psi^v \rangle_{\hat\Gamma_N}&\\ 
       %
-      + (\hat\rho_s \partial_t \hat v,\hat\psi^v)_{\Omega_s}  
-      + (\hat J\hat\sigma_s\hat F^{-T},\hat\nabla\hat\psi^v)_{\hat\Omega_s}&\\
-      - \langle \hat g, \hat\psi^v \rangle_{\hat\Gamma_N} -
-      (\hat\rho_f \hat J\hat f_f, \hat\psi^v)_{\hat\Omega_f}
-      - (\hat\rho_s\hat f_s, \hat\psi^v)_{\hat\Omega_s}
+      + (\hat\rho_s \partial_t \hat v,\hat\psi^v)_{\hat\Omega_s}  
+      + (\hat J\hat\sigma_s\hat F^{-T},\hat\nabla\hat\psi^v)_{\hat\Omega_s}
+%      - (\hat\rho_f \hat J\hat f_f, \hat\psi^v)_{\hat\Omega_f}
+%      - (\hat\rho_s\hat f_s, \hat\psi^v)_{\hat\Omega_s}
       &= 0&&\forall\hat\psi^v\in \hat V^0,
@@
 \hat\sigma_f = -\hat pI + \hat\rho_f\nu_f (\hat\nabla\hat v \hat F^{-1}
-  + \hat F^{-T} \hat\nabla\hat v^T)
-\], 
+  + \hat F^{-T} \hat\nabla\hat v^T),
+\] 
 and 
@@
 with $\hat E = \frac{1}{2}(\hat F^T \hat F - I)$.
+Finally, we notice that this problem is driven by a Dirichlet 
+inflow condition. It is possible to add a gravity term $\hat f_f$ or 
+$\hat f_s$, which 
+would enter as a right hand side force 
+\[
+- (\hat\rho_f \hat J\hat f_f, \hat\psi^v)_{\hat\Omega_f}
+      - (\hat\rho_s\hat f_s, \hat\psi^v)_{\hat\Omega_s}
+\]
+into the problem.
 \end{Problem}
@@
 degeneration as long as possible are re-meshing techniques or 
-to use a biharmonic mesh motion technique.
+to use (as suggested here) a biharmonic mesh motion technique.
 
@@
 \begin{equation*}
-F:= I , \quad \det F = J = 1.
+\hat F:= I , \quad \det \hat F = \hat J = 1.
 \end{equation*}
@@
 (F_D , F_L) 
-= {\int_{S_f} \sigma_f \cdot n_f \, ds + 
-\int_{\Gamma_i} \sigma_s \cdot n_s \, ds},
+= {\int_{S_f} \hat\sigma_f \cdot \hat n_f \, ds + 
+\int_{\hat\Gamma_i} \hat\sigma_s \cdot \hat n_s \, ds},
 \end{align}
@@
 %
-\textbf{Program description}
+\subsubsection{Program description}
 

### 2013-12-13 Examples/PDE/StatPDE/Example10/content.tex
Update Doku

@@
 \end{eqnarray}
+On fluid side, we extend the problem from Stokes flow 
+to stationary Navier-Stokes flow to demonstrate how to 
+implement the nonlinear convection term
+\[
+             (\hat J \rho_f \hat F^{-1}\hat  v\cdot\hat\nabla \hat v,
+             \hat\phi^v)_{\hat\Omega_f}.
+\]
+
 The whole equation system is solved on the benchmark configuration domain. For 
@@
 
-Compared to the two former examples, there are some differences which we will briefly discuss in the following.
+Compared to the previous Example \ref{PDE_Stat_FSI_STVK}, 
+there are some differences which we will briefly discuss in the following.
 First of all, the problem is nonlinear in contrast to the former ones.

### 2013-12-13 Examples/PDE/StatPDE/Example4/content.tex
Update Doku

@@
 \end{align*}
-We set zero Dirichlet values on $\partial \Gamma$ and choose $f=(1,1)$. The classical example of a PDE.
+We set zero Dirichlet values on $\partial \Gamma$ and choose $f=(1,1)$. The
+classical example of a PDE.
+
+\begin{remark}[Why this example?]
+Originally, \dope{} was designed for coupled and nonlinear problems
+with possible PDE-based optimization extensions. Later, we decided 
+to add the most simplest PDE (the Laplace/Poisson equation) 
+in order to demonstrate how DOpE treats
+this well-known example. In addition, the first-time user might 
+start here to get a feeling for \dope{} and its capabilities.
+\end{remark}

### 2013-12-13 Examples/PDE/StatPDE/Example9/content.tex
Update Doku

@@
     \begin{aligned} 
-      (\rho_f v\cdot\nabla v_f,\phi^v)_{\Omega_f} +
+  %    (\rho_f v\cdot\nabla v_f,\phi^v)_{\Omega_f} +
       (\sigma_f,\nabla\phi^v)_{\Omega_f} 
@@
       &&\quad\forall\hat \phi^u\in\hat V_s    ,\\
-      (\text{det}\,\hat F_s-1,\hat\phi^p)_{\hat\Omega_s} &=0
+      (\hat J-1,\hat\phi^p)_{\hat\Omega_s} &=0
       &&\quad\forall\hat \phi^p\in\hat L_s,\\
@@
     \begin{aligned}
-      (\hat J \rho_f \hat F^{-1}\hat  v\cdot\hat\nabla \hat v,
-      \hat\phi^v)_{\hat\Omega_f}
-      + (\hat J\hat\sigma_f\hat F^{-T},\hat \nabla\hat\phi^v)_{\hat\Omega_f}\\
+%      (\hat J \rho_f \hat F^{-1}\hat  v\cdot\hat\nabla \hat v,
+%      \hat\phi^v)_{\hat\Omega_f}
+      (\hat J\hat\sigma_f\hat F^{-T},\hat \nabla\hat\phi^v)_{\hat\Omega_f}
       + (\hat J\hat\sigma_s\hat F^{-T},\hat \nabla\hat\phi^v)_{\hat\Omega_s}
@@
 
-There is not much difference to the prior example, so we will keep our remarks rather short.
+%%% Unsinn! Das example hat nicht viel mit dem davor zu tun!
+%There is not much difference to the prior example, so we will keep our remarks rather short.
 The \textit{functionals.h} and \textit{myfunctions.cc} files did not change at all, so we simply refer to the corresponding sections in the last example.\\

### 2013-12-13 Examples/description.tex
Reordered the Examples in PDE/StatPDE according to the suggestion by Thomas.
Update description
New Example: Shows how to use real vector-valued elements such as Raviart-Thomas, and gives an example on a user given linear solver
contents tex in opt problems
Update Doku und Beseitigung kleinerer Doku-typos in zwei main files.
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented
Finished exchange of the examples
Update Doku. Habe Winnis-Superfunktionen fuer den Moment auskommentiert, da ich trotz tex-Update sonst nicht kompilieren

@@
 \usepackage{color}
+\usepackage{graphicx} 
+\usepackage{enumerate}
+\usepackage{listings}
+
 \theoremstyle{plain}
@@
 
-\newcommand{\todo}[1]{\textbf{\textsc{\textcolor{red}{TODO: #1}}}}
+%%% Geht bei mir nicht: habe tex neu installiert, geht aber trotzdem nicht
+%\usepackage{changes}
+%\usepackage{todonotes}
+
+%\newcommand{\twick}[1]{\added{#1}}
+%\newcommand{\twickremove}[1]{\deleted{#1}}
+
+\newcommand{\dope}{\texttt{DOpElib}}
 
@@
 \begin{document}
-\title{DOpElib: 
-Differential Equations\\ and Optimization Environment}
+\title{DOpElib:\\ 
+Differential Equations and Optimization Environment}
 \author{
@@
 
-\date{\today}
+\date{\today
+\vspace*{2cm}
+\begin{figure}[h]
+\centering
+{\includegraphics[width=8cm]{Documentation/logo.png}}
+\end{figure}
+}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \input{Documentation/foreword.tex}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary FSI with INH Material} 
-\label{PDE_Stat_FSI_INH}
+\subsection{Laplace equation with periodic BC} 
+\label{PDE_Stat_Periodic_Stokes}
 \input{PDE/StatPDE/Example2/content.tex}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary FSI with STVK Material} 
-\label{PDE_Stat_FSI_STVK}
+\subsection{Stationary Stokes Equations with hp-Elements} 
+\label{PDE_Stat_Stokes_hp}
 \input{PDE/StatPDE/Example3/content.tex}
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Laplace Equation in 2D} 
+\label{PDE_Stat_Laplace_2D}
+\input{PDE/StatPDE/Example4/content.tex}
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Adaptive Solution of Laplace Equation in 2D}
+\label{PDE_adap_Stat_Laplace}
+\input{PDE/StatPDE/Example5/content.tex}
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Laplace Equation in 3D}
+\label{PDE_Stat_Laplace_3D}
+\input{PDE/StatPDE/Example6/content.tex}
 \clearpage
@@
 \label{PDE_Stat_Elasticity}
-\input{PDE/StatPDE/Example4/content.tex}
+\input{PDE/StatPDE/Example7/content.tex}
 \clearpage
@@
 \label{PDE_Stat_Plasticity}
-\input{PDE/StatPDE/Example5/content.tex}
+\input{PDE/StatPDE/Example8/content.tex}
 \clearpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary Stokes Equations with periodic BC} 
-\label{PDE_Stat_Periodic_Stokes}
-\input{PDE/StatPDE/Example6/content.tex}
+\subsection{Stationary FSI with Stokes and INH Material} 
+\label{PDE_Stat_FSI_INH}
+\input{PDE/StatPDE/Example9/content.tex}
 \clearpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Laplace Equation in 2D} 
-\label{PDE_Stat_Laplace_2D}
-\input{PDE/StatPDE/Example7/content.tex}
-\clearpage
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Laplace Equation in 3D}
-\label{PDE_Stat_Laplace_3D}
-\input{PDE/StatPDE/Example8/content.tex}
-\clearpage
-\subsection{Adaptive Solution of Laplace Equation in 2D}
-\label{PDE_adap_Stat_Laplace}
-\input{PDE/StatPDE/Example9/content.tex}
-\clearpage
-\subsection{Usage of Higher Order Mappings: Approximation of $\pi$}
-\label{PDE_higher_order_mapping}
+\subsection{Stationary FSI with Navier-Stokes and STVK Material} 
+\label{PDE_Stat_FSI_STVK}
 \input{PDE/StatPDE/Example10/content.tex}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Stationary Stokes Equations with hp-Elements} 
-\label{PDE_Stat_Stokes_hp}
+\subsection{Usage of Higher Order Mappings: Approximation of $\pi$}
+\label{PDE_higher_order_mapping}
 \input{PDE/StatPDE/Example11/content.tex}
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Use of Raviart-Thomas element; Special linear solvers}
+\label{PDE_RT}
+\input{PDE/StatPDE/Example12/content.tex}
 \clearpage
@@
 \input{PDE/InstatPDE/Example5/content.tex}
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Biot-Lam\'e-Navier problem}
+\label{PDE_Instat_biot_lame_navier}
+\input{PDE/InstatPDE/Example6/content.tex}
 \cleardoublepage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Compliance Minimization of a variable Thickness MBB-Beam}
-\label{OPT_Stat_MBB-Beam}
+\subsection{Distributed control with a linear elliptic PDE using IPOPT/SNOPT}
+\label{OPT_Stat_Box_controlconstraints_SNOPT}
 \input{OPT/StatPDE/Example6/content.tex}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Distributed control with a linear elliptic PDE using IPOPT/SNOPT}
-\label{OPT_Stat_Box_controlconstraints_SNOPT}
+\subsection{Compliance Minimization of a variable Thickness MBB-Beam}
+\label{OPT_Stat_MBB-Beam}
 \input{OPT/StatPDE/Example7/content.tex}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Subject to a Non-Stationary PDE}
+\section{Subject to a Nonstationary PDE}
 \label{OPT_Instat}

### 2013-12-13 Examples/PDE/StatPDE/Example10/content.tex
Christians und Winnis Anmerkungen zu 5.1.9 und 5.1.10

@@
   \label{prob:stvk:lagrange}
-  Find $\{\hat v_f,\hat u_f\} \in \{ \hat v_f^D + \hat V\} 
-  \times \{ \hat u_f^D + \hat V\}$, 
+  Find $\{\hat u_s\} \in \{ \hat u_s^D + \hat V\}$, 
   such that 
@@
     \begin{aligned}   
-      (\hat J_s \hat\sigma_s \hat F_s^{-T},\hat\nabla\hat\phi^v)_{\hat\Omega_s}
+      (\hat J_s \hat\sigma_s(\hat u_s) \hat F_s^{-T},\hat\nabla\hat\phi^v)_{\hat\Omega_s}
       &= \langle\hat J_s\hat n_s\cdot \hat g_s^\sigma\hat
       F_s^{-T},\hat\phi^v  \rangle_{\hat\Gamma_i}
-      &&\quad\forall\hat \phi^v\in\hat V_s\\
-      %%%%%%%%%%%%%%%% 
-      (\hat v_s,\hat\phi^u)_{\hat\Omega_s}&=0
-      &&\quad\forall\hat \phi^u\in\hat V_s    ,\\
+      &&\quad\forall\hat \phi^v\in\hat V_s
     \end{aligned}
@@
   \begin{eqnarray}
-    \hat\sigma_s:=\hat J^{-1} \hat F (\lambda_s (tr\hat E) I + 2\mu_s \hat E)\hat F^{-T}.
+    \hat\sigma_s(\hat u_s):=\hat J^{-1} \hat F (\lambda_s (tr\hat E) I + 2\mu_s \hat E)\hat F^{-T}.
   \end{eqnarray}
 \end{Problem}
+\begin{remark}
+In the code, we also implement 
+\[
+(\hat v_s, \hat\phi),
+\]
+although this is not physically necessary. It is first for computational
+convenience in order to extend the fluid velocity variable to the whole
+domain.
+This could be resolved by using the FE Nothing element. Second, 
+using $\hat v_s$ here makes it easier to understand the 
+nonstationary FSI problem. The same holds for the (artificial) 
+pressure variable 
+in the STVK case.
+\end{remark}
+
 Often, the elasticity properties of structure materials is characterized by 
@@
 On fluid side, we extend the problem from Stokes flow 
-to stationary Navier-Stokes flow to demonstrate how to 
-implement the nonlinear convection term
+to stationary Navier-Stokes flow considering the convection term
+\[
+v\cdot \nabla v
+\]
+which reads in transformed form \cite{Wi11}
 \[

### 2013-12-13 Examples/PDE/StatPDE/Example9/content.tex
Christians und Winnis Anmerkungen zu 5.1.9 und 5.1.10

@@
 
+\begin{remark}
+In the problems above and the code, we implement the term
+\[
+(\hat v_s, \hat\phi^u),
+\]
+although this is not physically necessary. It is first for computational
+convenience in order to extend the fluid velocity variable to the whole
+domain.
+This could be resolved by using the FE Nothing element. Second, 
+using $\hat v_s$ here makes it easier to understand the 
+nonstationary FSI problem.
+\end{remark}
+
 \subsubsection{Program description}

### 2013-12-14 Examples/Documentation/introduction.tex
Kleine Aenderungen, siehe Mail
Update Doku in Einklang mit Christians und Winnis Bemerkungen/Zustimmung
Bemerkung 3.0.1 zu fullmem und store_on_disc

@@
 
-\section{Developers}
+\section{Contributors \& Developers}
 \label{sec:contrib}
@@
 \end{itemize}
+\begin{remark}
+\label{remark_state_behavior}
+We notice that the behavior of the statevector can be chosen as 
+\texttt{fullmem} or \texttt{store\underline{ }on\underline{ }disc}. In the
+first state, the RAM memory of the computer is used. In the second state,
+all vectors are written on disc, to avoid the RAM. This might take some 
+time at the beginning of a new executing program (cleary depending 
+on the number of spatial and temporal unknowns and the capabilities of your 
+local machine). In addition, if the program aborts abnormally in the 
+using \texttt{store\underline{ }on\underline{ }disc} behavior, please make 
+sure to remove manually the \texttt{tmp\underline{ }state} folder in your 
+\texttt{Results} folder. 
+\end{remark}
 

### 2013-12-14 Examples/PDE/InstatPDE/Example2/content.tex
Bemerkung 3.0.1 zu fullmem und store_on_disc

@@
 
+In addition, we notice that specifically the FSI 3 version of this 
+code introduces a large amount of total DOFs. Probably, your machine 
+will terminate in \texttt{fullmem} modus. Here you might change the 
+statevector behavior to \texttt{store\underline{ }on\underline{ }disc}.
+However, it takes some time to write all vectors locally on your disc.
+On a Lenovo X230, it takes $3$ minutes. In a future version of the library,
+we plan to introduce a less expensive statevector behavior.
+We also refer to Remark \ref{remark_state_behavior}.
+
 \vspace{0.2cm}

### 2013-12-16 Examples/Documentation/examples.tex

@@
 \end{itemize}
-\begin{remark}[Why two examples for the Laplace equation?]
-\dope{} offers dimensional-independent programming! So, we could have 
-implemented the 2d and 3d case in one file. The reason why we
-present
-to separate programs \ref{PDE_Stat_Laplace_2D} and \ref{PDE_Stat_Laplace_3D}
-is, that in the 3d case, we show additional 
-capabilities of the library itself; specifically how to use different
-iterative linear solvers.
-\end{remark}
+%\begin{remark}[Why two examples for the Laplace equation?]
+%\dope{} offers dimensional-independent programming! So, we could have 
+%implemented the 2d and 3d case in one file. The reason why we
+%present
+%to separate programs \ref{PDE_Stat_Laplace_2D} and \ref{PDE_Stat_Laplace_3D}
+%is, that in the 3d case, we show additional 
+%capabilities of the library itself; specifically how to use different
+%iterative linear solvers.
+%\end{remark}
 

### 2013-12-16 Examples/Documentation/introduction.tex

@@
 
-\subsection{Problem description}
+\section{Problem description}
 In order to allow our algorithms the automatic assembly of all required 
@@
 
-\subsection{Numeric components}
+\section{Numeric components}
 These are the components from which a user needs to select some in order to actually 
@@
 
-\subsubsection{Space-time handler}
+\subsection{Space-time handler}
 First we need to select a method how to handle all dofs in space and time.
@@
 
-\subsubsection{Container classes}
+\subsection{Container classes}
 Second you will need to specify some container classes to be used to 
@@
 
-\subsubsection{Time stepping schemes}
+\subsection{Time stepping schemes}
 Third, at least for nonstationary PDEs we need to select a time stepping scheme
@@
 
-\subsubsection{Integrator routines}
+\subsection{Integrator routines}
 Finally, we need to select a way how to integrate and solve linear and nonlinear equations
@@
 
-\subsubsection{Nonlinear solvers}
+\subsection{Nonlinear solvers}
 \begin{itemize}
@@
 
-\subsubsection{Linear solvers}
+\subsection{Linear solvers}
 \begin{itemize}
@@
 
-\subsection{Problem specific classes}
+\section{Problem specific classes}
 The following classes are used to describe the problem and will usually require 
@@
 
-\subsection{Reduced problems (Solve the PDE)}
+\section{Reduced problems (Solve the PDE)}
 At times it is nice to remove the PDE constraint in (OPT). 
@@
 
-\subsection{Optimization algorithms}
+\section{Optimization algorithms}
 Now, in order to solve optimization algorithms we need to define some algorithms.
@@
 
-\subsection{Other Components}
+\section{Other Components}
 Beyond these clearly structured groups before there are some classes remaining that
@@
 
-\subsubsection{Vectors}
+\subsection{Vectors}
 \begin{itemize}
@@
 
-\subsubsection{Parameter handling}
+\subsection{Parameter handling}
 \begin{itemize}
@@
 
-\subsubsection{Exception handling}
+\subsection{Exception handling}
 \begin{itemize}
@@
 
-\subsubsection{Output handling}
+\subsection{Output handling}
 \begin{itemize}
@@
 
-\subsection{Data Access}
+\section{Data Access}
 \begin{itemize}
@@
 
-\subsubsection{Constraints and system matrix}
+\subsection{Constraints and system matrix}
 \begin{itemize}
@@
 
-\subsubsection{HP components}
+\subsection{HP components}
 \begin{itemize}
@@
 
-\subsection{Internal structures}
-\subsubsection{Interface Classes}
+\section{Internal structures}
+\subsection{Interface Classes}
 \begin{itemize}
@@
 
-\subsubsection{Default Classes}
+\subsection{Default Classes}
 \begin{itemize}
@@
 
-\subsubsection{Auto-generated Problem Descriptions}
+\subsection{Auto-generated Problem Descriptions}
 \begin{itemize}
@@
 \end{itemize}
-\subsubsection{Management of Time Dependent Problems}
+\subsection{Management of Time Dependent Problems}
 \begin{itemize}
@@
 \end{itemize}
-\subsection{Wrapper classes}
+\section{Wrapper classes}
 \begin{itemize}
@@
 
-\subsection{Other}
+\section{Other}
 \begin{itemize}

### 2013-12-16 Examples/PDE/InstatPDE/Example3/content.tex

@@
 \subsubsection{Program description}
-Note that the initial conditions are only $H^1$-regular.
-Because of this, the Crank-Nicolson scheme, which is not strongly A-stable, is not suited to solve this problem.
-As we want to use a second order accurate time stepping scheme, we use the shifted Crank-Nicolson method.
-The rest of the program is as before.
+Note that the initial conditions are only $H^1$-regular. Because of this, 
+the Crank-Nicolson scheme, which is not strongly A-stable, is not suited to solve 
+this problem. As we want to use a second order accurate time stepping scheme, we use 
+the shifted Crank-Nicolson method. The rest of the program is as before.
+
+In addition, this program shows how to change the vector behavior, from our
+default option \texttt{fullmem}, where the whole vector
+is stored in the computers main memory. Here the storage behavior is set to 
+\texttt{store\underline{ }on\underline{ }disc}, where only those unknowns are loaded into 
+main memory that are needed in the current time step, while all other 
+unknowns are stored on the hard drive.
+
+This behavior is particularly useful, if many time steps are taken; so that the 
+whole set of unknowns can no longer be stored in main memory. Note that all vectors 
+will allocate the required memory when the vector is reinitialized to a new size.
+Hence, for large vectors this may need some time.
+
+To avoid multiple programs accessing the same files on the hard drive, a 
+\texttt{lock} file is initialized. Under normal conditions, this will be deleted 
+once the program terminates. However, should the program exit exceptionally, the 
+lock file will still exist. Calling the program in this case will produce an 
+exception, with the following text:
+\begin{verbatim}
+Warning: During execution of `StateVector<VECTOR>::StateVector` 
+         the following Problem occurred!
+The directory Results/tmp_state/ is propaply already in use.
+\end{verbatim}
+To resolve the issue, you have to delete the named directory with the 
+temporary storage files manually.

### 2014-01-02 Examples/Documentation/introduction.tex
Added only_recent behavior to the StateVector

@@
 We notice that the behavior of the statevector can be chosen as 
-\texttt{fullmem} or \texttt{store\underline{ }on\underline{ }disc}. In the
+\texttt{fullmem}, \texttt{only_recent}, or 
+\texttt{store\underline{ }on\underline{ }disc}. In the
 first state, the RAM memory of the computer is used. In the second state,
+only the spatial vectors at the current time step (and the preious one)
+are stored. This reduces memory requirements, but also prohibits access to the
+whole space-time trajectory after the computation.
+In the third state,
 all vectors are written on disc, to avoid the RAM. This might take some 

### 2014-01-02 Examples/PDE/InstatPDE/Example2/content.tex
Added only_recent behavior to the StateVector

--- 2013-12-13 Examples/PDE/InstatPDE/Example2/content.tex
@@
 
+Further, this example shows, how to change the vector behavior, from 
+our default option \texttt{fullmem}, where the whole vector is stored in 
+the computers main memory. Here, we are only interested in calculating the 
+solution once, hence two vectors, one for the current time point and one for the
+previous one are sufficient. Hence we choose the option 
+\texttt{only_recent} so that we don't have to reserve unneccessary memory.
+If we need to store the whole trajectory for some reason another option is available 
+to circumvent the restrictions due to the size of the main memory, it will be 
+described in Example~\ref{PDE_Instat_Black_Scholes}.
+
 \vspace{0.2cm}

### 2014-01-02 Examples/PDE/InstatPDE/Example3/content.tex
typo
Added only_recent behavior to the StateVector

@@
 This behavior is particularly useful, if many time steps are taken; so that the 
-whole set of unknowns can no longer be stored in main memory. Note that all vectors 
+whole set of unknowns can no longer be stored in main memory, but access to all
+parts of the solution is required after the solution process and thus the 
+option \texttt{only_recent} used in Example~\ref{PDE_Instat_FSI} is not sufficient.
+Note that all vectors 
 will allocate the required memory when the vector is reinitialized to a new size.
@@
          the following Problem occurred!
-The directory Results/tmp_state/ is propaply already in use.
+The directory Results/tmp_state/ is probably already in use.
 \end{verbatim}

### 2014-01-23 Examples/PDE/InstatPDE/Example3/content.tex
1) Sehr verwirrende (und falsche) Materialparameter in plasticity aufgeraeumt, 2) Typos in Doku beseitigt, da nicht mehr kompiliert (only_recent -> only\underline{}recent

--- 2013-12-16 Examples/PDE/InstatPDE/Example3/content.tex
@@
 This behavior is particularly useful, if many time steps are taken; so that the 
-whole set of unknowns can no longer be stored in main memory. Note that all vectors 
+whole set of unknowns can no longer be stored in main memory, but access to all
+parts of the solution is required after the solution process and thus the 
+option \texttt{only\underline{ }recent} used in Example~\ref{PDE_Instat_FSI} is not sufficient.
+Note that all vectors 
 will allocate the required memory when the vector is reinitialized to a new size.

### 2014-01-23 Examples/PDE/StatPDE/Example7/content.tex
Parameterkorrekturen ebenfalls in elasticity Beispiel

@@
 \begin{equation*}
-   \sigma(v) := 2\mu \varepsilon(v)^D + \kappa \cdot tr (\varepsilon(v)) I
+   \sigma(v) := 2\mu \varepsilon(v)^D + \rho \, tr (\varepsilon(v)) I = 2\mu\varepsilon(v) + \lambda tr(\varepsilon(v)) I,
 \end{equation*}
@@
 \end{equation*}
-and the parameters $\mu$ and $\kappa$ are chosen as $\mu = 80193.800283$ resp. $\kappa = 190937.589172$.\\
+and the parameters $\mu$ and $\rho$ are chosen as $\mu = 80193.800283$
+resp. $\rho = (\mu + \lambda) = 190937.589172$. We notice that $\mu$ and
+$\lambda$ denote the usual Lam\'e parameters.\\
 The corner points of our computational domain are in anticlockwise order: $(0,0), (90,0)$, $(100,10), (100,100)$ and $(0,100)$.

### 2014-01-23 Examples/PDE/StatPDE/Example7/localpde.h

@@ -64,8 +64,11 @@ template<
         const FEValuesExtractors::Vector displacements(0);
 
         const double mu = 80193.800283;
-        const double kappa = 271131.389455;
         const double lambda = 110743.788889;
+
+	// kappa = 2*mu + lambda
+        const double kappa = 271131.389455;
+
 
         for (unsigned int q_point = 0; q_point < n_q_points; q_point++)
         {
@@ -111,8 +114,11 @@ template<
         const FEValuesExtractors::Vector displacements(0);
 
         const double mu = 80193.800283;
-        const double kappa = 271131.389455;
         const double lambda = 110743.788889;
+
+	// kappa = 2*mu + lambda
+        const double kappa = 271131.389455;
+
 
         std::vector<Tensor<2, 2> > phi_grads_v(n_dofs_per_element);
         std::vector<Tensor<2, 2> > phi_grads_real(n_dofs_per_element);

### 2014-01-27 DOpEsrc/opt_algorithms/reduced_snopt_algorithm.h
Updated DOpE to have the timestepsize hidden in the optproblemcontainer; New ControlVectorType::stationary for controls that are independent of time, but act at all timesteps.

@@ -26,6 +26,7 @@
 
 #include "reducedalgorithm.h"
 #include "parameterreader.h"
+#include "dopetypes.h"
 
 #include "snopt_wrapper.h"
 
@@ -154,6 +155,12 @@ namespace DOpE
       _vector_behavior = vector_behavior;
       _postindex = "_" + this->GetProblem()->GetName();
 
+      DOpEtypes::ControlType ct = S->GetProblem()->GetSpaceTimeHandler()->GetControlType();
+      if((ct != DOpEtypes::ControlType::initial) && (ct != DOpEtypes::ControlType::stationary))
+      {
+	throw DOpEException("The ControlType: "+ DOpEtypesToString(ct) + " is not supported.",
+              "Reduced_SnoptAlgorithm::Reduced_SnoptAlgorithm");
+      }
     }
 
   /******************************************************/

### 2014-01-28 Examples/OPT/InstatPDE/Example2/content.tex
Added global time refinement; Fixed missing initializations in InstatOptProblemContainer; New Example OPT/InstatPDE/Example2 for ControlType::stationary

\subsubsection{General problem description}

This example is a modified version of \texttt{OPT/InstatPDE/Example1}. 
Again, we consider the heat equation, this time with an additional nonlinear
term and most important a time derivative leading to a 
nonstationary optimization problem. The governing equation is
\begin{equation*}
\partial_t u(t,x,y) - \Delta u(t,x,y) = f(t)\cdot q(x,y),
\end{equation*}
with homogeneous Dirichlet-data.
The computational domain is $\Omega \times I = [0,\pi]^2 \times [0,1]$. From the known solution, we can compute the appropriate data 
\begin{align*}
f(t) &= 2,\\
u_0(x,y) &= \sin(x)\sin(y).
\end{align*}

With the cost functional 
\begin{align*}
 \min_{q,u} J(q,u) =&\; \frac{1}{2}\int_{\Omega} \Biggl(u(1,x,y) - \biggl(\frac{2e^2-1}{e^2-1}\biggr)\sin(x) \sin(y)\Biggr)^2\,d(x,y) + \frac{1}{2} \int_{\Omega} q(x,y)^2\,d(x,y).
\end{align*}
It has the optimal solution 
\begin{align*}
\overline{u}(t,x,y) &= \sin(x) \sin(y),\\
\overline{q}(t,x,y) &= \sin(x) \sin(y)
\end{align*}
together with the optimal adjoint state 
\[
 \overline{z}(t,x,y) = \frac{e^2}{1-e^2}e^{2(t-1)}\sin(x) \sin(y).
\]
and the corresponding cost functional value 
\[
J(\overline{q},\overline{u}) = \frac{1}{2} \Bigl(\frac{e^4}{(e^2-1)^2}+1\Bigr)\frac{\pi^2}{4} \approx 2.88382.
\]

\subsubsection{Program description}
In this example, we demonstrate how to implement a control that acts distributed in space and time,
but has no temporal dependence. For this, the control vector type is set to be 
{\texttt ControlType::stationary} (the default for stationary equations). In contrast to 
the case of control in the initial values, the control vector can be accessed at a 
all times.

Since the control has no time dependence \texttt{DOpElib} assumes that the control part of 
the cost functional, here 
\[
 \frac{1}{2}\int_\Omega q^2\,dx
\]
is evaluated at initial time (i.e., $t=0$ in this example).

### 2014-01-28 DOpEsrc/basic/spacetimehandler.h

@@ -77,7 +77,7 @@ namespace DOpE
                 dealii::numbers::invalid_unsigned_int)
         {
         }
-        SpaceTimeHandler(const dealii::Triangulation<1> & times,
+        SpaceTimeHandler(dealii::Triangulation<1> & times,
             DOpEtypes::ControlType type) :
             SpaceTimeHandlerBase<VECTOR>(times, type), _control_index(
                 dealii::numbers::invalid_unsigned_int), _state_index(
@@ -92,7 +92,7 @@ namespace DOpE
                 &index_setter)
         {
         }
-        SpaceTimeHandler(const dealii::Triangulation<1> & times,
+        SpaceTimeHandler(dealii::Triangulation<1> & times,
             DOpEtypes::ControlType type,
             const ActiveFEIndexSetterInterface<dopedim, dealdim>& index_setter) :
             SpaceTimeHandlerBase<VECTOR>(times, type), _control_index(

### 2014-02-26 Examples/OPT/InstatPDE/Example3/content.tex
Added support for time dependent controls and included OPT/InstatPDE/Example3 to show how to use it. Further more, a bug in the functional evaluation has been fixed, up to now it was always assumed that the type ist timelocal. This is now fixed

\subsubsection{General problem description}

This example is a modified version of \texttt{OPT/InstatPDE/Example2}. 
The governing equation is
\begin{equation*}
\partial_t u(t,x,y) - \Delta u(t,x,y) = q(t)\cdot f(x,y),
\end{equation*}
with homogeneous Dirichlet-data. Hence, we now allow for a time dependence of the control.

The computational domain is $\Omega \times I = [0,\pi]^2 \times [0,1]$. From the known solution, we can compute the appropriate data 
\begin{align*}
f(x,y) &= \sin(x)\sin(y),\\
u_0(x,y) &= 0,\\
u^d(t,x,y)&= \sin(x)\sin(y) \bigl(\tfrac{1}{4}(3-2t-3e^{-2t}) + \tfrac{4}{\pi^2} + (1-t)\tfrac{8}{\pi^2}\bigr).
\end{align*}

With the cost functional 
\begin{align*}
 \min_{q,u} J(q,u) =&\; \frac{1}{2}\int_I\int_{\Omega} \bigl(u(t,x,y) - u^d(t,x,y)\bigr)^2\,d(x,y)\,dt + \frac{1}{2} \int_{I} q(t)^2\,d(t).
\end{align*}
It has the optimal solution 
\begin{align*}
\overline{u}(t,x,y) &= \tfrac{1}{4}(3-2t-3e^{-2t})\sin(x) \sin(y),\\
\overline{q}(t) &= 1-t 
\end{align*}
together with the optimal adjoint state 
\[
 \overline{z}(t,x,y) = \sin(x) \sin(y) \frac{4(t-1)}{\pi^2}.
\]
Additionally, we evaluate the following functionals
\begin{align*}
  \overline{u}_h(1,0.5\pi,0.5\pi) &= \frac{1-3e^{-2}}{4} \approx 0.148499,\\
  \| \overline{u}_h - \overline{u}\|^2_{\Omega \times I},& \\
  \| \overline{q}_h - \overline{q}\|^2_I. &  
\end{align*}

\subsubsection{Program description}
In this example, we demonstrate how to implement a control that acts distributed in time. 
For this, the control vector type is set to be 
{\texttt ControlType::nonstationary}. 
Obviously one can implement space and time dependence by considering a ControlVector in
dimension $\ne 0$.

### 2014-02-26 DOpEsrc/reducedproblems/statpdeproblem.h

@@ -235,13 +235,10 @@ namespace DOpE
          *
          *  @param v           The ControlVector<VECTOR> to write to a file.
          *  @param name        The names of the variables, e.g., in a fluid problem: v1, v2, p.
-         *  @param outfile     The basic name for the output file to print.
          *  @param dof_type    Has the DoF type: state or control.
-         *  @param filetype    The filetype. Actually, *.vtk outputs are possible.
          */
         void
-        WriteToFile(const ControlVector<VECTOR> &v, std::string name,
-            std::string outfile, std::string dof_type, std::string filetype);
+        WriteToFile(const ControlVector<VECTOR> &v, std::string name, std::string dof_type);
 
         /**
          * Basic function to write a std::vector to a file.
@@ -921,10 +918,9 @@ namespace DOpE
       typename VECTOR, int dealdim>
     void
     StatPDEProblem<NONLINEARSOLVER, INTEGRATOR, PROBLEM, VECTOR, dealdim>::WriteToFile(
-        const ControlVector<VECTOR> &v, std::string name, std::string outfile,
-        std::string dof_type, std::string filetype)
+        const ControlVector<VECTOR> &v, std::string name, std::string dof_type)
     {
-      WriteToFile(v.GetSpacialVector(), name, outfile, dof_type, filetype);
+      throw DOpEException("This Problem does not support ControlVectors","StatPDEProblem::WriteToFile");
     }
 }
 #endif

### 2014-10-14 Examples/Documentation/examples.tex

@@
 \end{verbatim}
-where, in our example you find two times the lines
+where, in our example you find the line
 \begin{verbatim}
-echo "Running Program ../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm"
-(../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d test.prm 2>&1) > /dev/null
+PROGRAM=../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d
 \end{verbatim} 
@@
 \begin{verbatim}
-echo "Running Program ../../../../../bin/MyWonderfulFirstExample test.prm"
-(../../../../../bin/MyWonderfulFirstExample test.prm 2>&1) > /dev/null
+PROGRAM=../../../../../bin/MyWonderfulFirstExample
 \end{verbatim}
@@
 \begin{verbatim}
-./test Test
+./test.sh Test
 \end{verbatim}
@@
 \begin{verbatim}
-./test Store
+./test.sh Store
 \end{verbatim}

### 2014-10-14 Examples/OPT/StatPDE/Example8/Test/test.sh

@@ -5,6 +5,8 @@ if [ $# -ne 1 ]
     exit 1
 fi
 
+PROGRAM=../../../../../bin/DOpE-OPT-StatPDE-Example8-2d-2d
+
 if [ -f dope.log ]
 then
 	rm dope.log
@@ -14,8 +16,8 @@ if [ $1 == "Test" ]
 then
     if [ -f test.dlog ]
     then
-	echo "Running Program ../../../../../bin/DOpE-OPT-StatPDE-Example8-2d-2d test.prm"
-	(../../../../../bin/DOpE-OPT-StatPDE-Example8-2d-2d test.prm 2>&1) > /dev/null
+	echo "Running Program ${PROGRAM} test.prm"
+	(${PROGRAM} test.prm 2>&1) > /dev/null
 	echo "Comparing Results:"
 	(diff dope.log test.dlog 2>&1) > /dev/null
 	if [ $? -eq 0 ]
@@ -39,8 +41,8 @@ then
 else
     if [ $1 == "Store" ]
     then
-	echo "Running Program ../../../../../bin/DOpE-OPT-StatPDE-Example8-2d-2d test.prm"
-	(../../../../../bin/DOpE-OPT-StatPDE-Example8-2d-2d test.prm 2>&1) > /dev/null
+	echo "Running Program ${PROGRAM} test.prm"
+	(${PROGRAM} test.prm 2>&1) > /dev/null
 	echo "Run completed. Cleaning up ..."
 	mv dope.log test.dlog
 	rm -r Mesh0/

### 2014-10-14 Examples/OPT/StatPDE/Example9/Test/test.sh

@@ -5,6 +5,8 @@ if [ $# -ne 1 ]
     exit 1
 fi
 
+PROGRAM=../../../../../bin/DOpE-OPT-StatPDE-Example9-0d-2d
+
 if [ -f dope.log ]
 then
 	rm dope.log
@@ -14,8 +16,8 @@ if [ $1 == "Test" ]
 then
     if [ -f test.dlog ]
     then
-	echo "Running Program ../../../../../bin/DOpE-OPT-StatPDE-Example9-0d-2d test.prm"
-	(../../../../../bin/DOpE-OPT-StatPDE-Example9-0d-2d test.prm 2>&1) > /dev/null
+	echo "Running Program ${PROGRAM} test.prm"
+	(${PROGRAM} test.prm 2>&1) > /dev/null
 	echo "Comparing Results:"
 	(diff dope.log test.dlog 2>&1) > /dev/null
 	if [ $? -eq 0 ]
@@ -37,8 +39,8 @@ then
 else
     if [ $1 == "Store" ]
     then
-	echo "Running Program ../../../../../bin/DOpE-OPT-StatPDE-Example9-0d-2d test.prm"
-	(../../../../../bin/DOpE-OPT-StatPDE-Example9-0d-2d test.prm 2>&1) > /dev/null
+	echo "Running Program ${PROGRAM} test.prm"
+	(${PROGRAM} test.prm 2>&1) > /dev/null
 	echo "Run completed. Cleaning up ..."
 	mv dope.log test.dlog
 	rm -r Mesh0/

### 2014-10-14 Examples/description.tex
Updated DOpE to have the timestepsize hidden in the optproblemcontainer; New ControlVectorType::stationary for controls that are independent of time, but act at all timesteps.
Added support for time dependent controls and included OPT/InstatPDE/Example3 to show how to use it. Further more, a bug in the functional evaluation has been fixed, up to now it was always assumed that the type ist timelocal. This is now fixed

@@
     {christian.goll@iwr.uni-heidelberg.de}}, 
-  Thomas Wick\thanks{The University of Texas at Austin, 
-    {twick@ices.utexas.edu}}, 
+  Thomas Wick\thanks{Johann Radon Institute for Computational and Applied Mathematics (RICAM), 
+    {thomas.wick@ricam.oeaw.ac.at}}, 
   Winnifried Wollner\thanks{University of Hamburg, 
-    {winnifried.wollner@math.uni-hamburg.de}}
+    {winnifried.wollner@uni-hamburg.de}}
 }
@@
 \clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Control of the heat equation via a space dependent right hand side}
+\label{OPT_Instat stationary control}
+\input{OPT/InstatPDE/Example2/content.tex}
+\clearpage
+\subsection{Control of the heat equation via a time dependent right hand side}
+\label{OPT_Instat nonstationary control}
+\input{OPT/InstatPDE/Example3/content.tex}
+\clearpage
 

### 2014-10-18 Examples/OPT/StatPDE/Example6/Test/test.sh

@@ -5,7 +5,7 @@ if [ $# -ne 1 ]
     exit 1
 fi
 
-PROGRAM=../../../../../bin/DOpE-OPT-StatPDE-Example6-2d-2d
+PROGRAM=../DOpE-OPT-StatPDE-Example6-2d-2d
 
 if [ -f dope.log ]
 then

### 2014-10-18 Examples/PDE/InstatPDE/Example5/Test/test.sh

@@ -5,7 +5,7 @@ if [ $# -ne 1 ]
     exit 1
 fi
 
-PROGRAM=../../../../../bin/DOpE-PDE-InstatPDE-Example5-2d-2d
+PROGRAM=../DOpE-PDE-InstatPDE-Example5-2d-2d
 
 if [ -f dope.log ]
 then

### 2014-10-18 Examples/PDE/StatPDE/Example4/Test/test.sh

@@ -5,7 +5,7 @@ if [ $# -ne 1 ]
     exit 1
 fi
 
-PROGRAM=../../../../../bin/DOpE-PDE-StatPDE-Example4-2d-2d
+PROGRAM=../DOpE-PDE-StatPDE-Example4-2d-2d
 
 if [ -f dope.log ]
 then

### 2014-10-18 Examples/PDE/StatPDE/Example9/Test/test.sh

@@ -5,7 +5,7 @@ if [ $# -ne 1 ]
     exit 1
 fi
 
-PROGRAM=../../../../../bin/DOpE-PDE-StatPDE-Example9-2d-2d
+PROGRAM=../DOpE-PDE-StatPDE-Example9-2d-2d
 
 if [ -f dope.log ]
 then

### 2014-10-22 Examples/Documentation/introduction.tex
1) Sehr verwirrende (und falsche) Materialparameter in plasticity aufgeraeumt, 2) Typos in Doku beseitigt, da nicht mehr kompiliert (only_recent -> only\underline{}recent
Software review: Punkte 1,6,7,8 bearbeitet (siehe TODO in welcher Form ich diese bearbeitet habe)

@@
   \item Francesco Ludovici (Augmenting the Documentation)
+  \item Masoud Ghaderi (Augmenting the Documentation)
 \end{itemize}
@@
 \section{Installation}
+To work with DOpElib, your need necessarily deal.II, which installation 
+we describe first. Afterwards, the DOpElib installation is described and finally
+other optional packages might be installed.
 \label{sec_installation}
-I) Get a copy of \texttt{DOpElib}, see Section~\ref{sec:obtain} for details.\\[1mm]
-%
-II) Install \texttt{deal.II} to your home directory, i.e., it should be located in \\
+\begin{itemize}
+\item Install \texttt{deal.II} to your home directory, i.e., it should be located in \\
 \texttt{\textasciitilde/deal.II}. \\[1mm]
@@
 \texttt{D = \$(HOME)/deal.II} with the appropriate path.\\[1mm]
-Configure it properly and make the whole \texttt{deal.II} library.\\[3mm]
+Detailed installing instructions for deal.II (here the last version
+8.1.0) can be found on 
+\texttt{http://www.dealii.org/8.1.0/readme.html}. The deal.II 
+installation instructions are descriptive enough and we omit 
+any further comments and refer to their webpage.
 %
-III) If you want to use some of the supported third party libraries install them and follow 
- the instructions in \texttt{ThirdPartyLibs/README}. There may be further information 
- in some \texttt{ThirdPartyLibs/*.INSTALLNOTES} that you may want to consider.\\[3mm]
+
 %
-IV) To build the \texttt{DOpElib} you have to change to the directory 
-\texttt{DOpEsrc} where you can call\\ 
-\texttt{make all}\\
-to build the library. If you want to generate some documentation you may use
-\texttt{make pdf-doc} or \texttt{make html-doc} which will require either 
-latex or doxygen to be installed on your computer.\\[3mm]
+\item Get a copy of \texttt{DOpElib}, see Section~\ref{sec:obtain} for details.
 %
-V) To build the Examples change to the \texttt{Examples} directory and \texttt{make all}.
-%
-VI) Finally you may wish to test if everything worked. To do so you can 
+\item Unpack in your preferred directory:
+\begin{lstlisting}
+~/Software/> tar -xzf dopelib-1.0.tar.gz
+\end{lstlisting}
+to get
+\begin{lstlisting}
+~/Software/dopelib-1.0
+\end{lstlisting}
+To build the \texttt{DOpElib} you have to change to the directory 
+\texttt{DOpEsrc} where you can call 
+\begin{lstlisting}
+~/Software/dopelib-1.0/DOpEsrc> make all
+\end{lstlisting}
+to build the library. You also get various installation options by just typing 
+\texttt{make}.
+\item To generate documentation, please go into:
+\begin{lstlisting}
+~/Software/dopelib-1.0/Examples
+\end{lstlisting}
+Here, you get all options by typing:
+\begin{lstlisting}
+make
+\end{lstlisting}
+Specifically, you can create and run all examples, running a test suite,
+and generating documentation:
+{\footnotesize
+\begin{lstlisting}
+===========================================================================
+=              Makefile for the DOpE documentation                        =
+===========================================================================
+=                                                                         =
+= The following targets exist:                                            =
+=    all       :  Make all examples (in optimized mode)                   =
+=    allignore :  Make all examples (in optimized mode), ignore errors.   =
+=    clean     :  Cleaning up all examples                                =
+=    tests     :  Run all test param data.                                =
+=    cat       :  Run clean, run all, run tests (combine these commands)  =
+=    catignore :  Same as cat, but ignores compiler errors.               =
+=    pdf-doc   :  Create documentation in pdf file format via latexmk     =
+=    doc       :  Create documentation in pdf file format via latexmk     =
+=    distclean :  Cleaning up, including documentation                    =
+=    warncheck :  Checks whether all Examples compile without warnings    =
+===========================================================================
+\end{lstlisting}
+}
+For html documentation, change into:
+\begin{lstlisting}
+~/Software/dopelib-1.0/doxygen
+\end{lstlisting}
+and then typing \texttt{make}. Doxygen documentation
+will require either 
+latex or doxygen to be installed on your computer.
+\item If you wish to test if everything worked. To do so you can 
 change to the \texttt{Examples} directory and \texttt{make tests} which will give you a 
@@
 Chapter~\ref{chap:test}.
+\item  If you want to use some of the supported third party libraries install them and follow 
+ the instructions in \texttt{ThirdPartyLibs/README}. There may be further information 
+ in some \texttt{ThirdPartyLibs/*.INSTALLNOTES} that you may want to
+ consider.\\[3mm]
+ As example, the installation of ipopt works as follows:\\
+In the path \texttt{~/dope/ThirdPartyLibs>} type in the terminal\\
+\texttt{./install-free-libs.sh}
+
+\end{itemize}
 
@@
 We notice that the behavior of the statevector can be chosen as 
-\texttt{fullmem}, \texttt{only_recent}, or 
+\texttt{fullmem}, \texttt{only\underline{ }recent}, or 
 \texttt{store\underline{ }on\underline{ }disc}. In the

### 2014-10-24 Examples/Documentation/introduction.tex

@@
 \section{License information}
-Copyright (C) 2012 by the DOpElib authors\\[2mm]
+Copyright (C) 2012--2014 by the DOpElib authors\\[2mm]
 %
@@
 \begin{lstlisting}
-~/Software/> tar -xzf dopelib-1.0.tar.gz
+~/Software/> tar -xzf dopelib-2.0.tar.gz
 \end{lstlisting}
@@
 \begin{lstlisting}
-~/Software/dopelib-1.0
+~/Software/dopelib-2.0
 \end{lstlisting}
@@
 \begin{lstlisting}
-~/Software/dopelib-1.0/DOpEsrc> make all
+~/Software/dopelib-2.0/DOpEsrc> make all
 \end{lstlisting}
@@
 \begin{lstlisting}
-~/Software/dopelib-1.0/Examples
+~/Software/dopelib-2.0/Examples
 \end{lstlisting}
@@
 \begin{lstlisting}
-~/Software/dopelib-1.0/doxygen
+~/Software/dopelib-2.0/doxygen
 \end{lstlisting}

### 2014-10-24 Examples/Documentation/introduction.tex
Typos und mini-updates in description Section 2.5

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Software requirements}
@@
 \texttt{./configure --with-umfpack}
-in version up to 7.3.0. From 8.0.0 on, deal.II is compited with 
+in version up to 7.3.0. From 8.0.0 on, deal.II is compiled with 
 cmake, which makes the inclusion of external packages easier.
+Detailed installing instructions for deal.II (here the last version
+8.1.0) can be found on\\ 
+\texttt{http://www.dealii.org/8.1.0/readme.html}.
 
@@
 \texttt{deal.II} with at least the following options:\\
-\texttt{./configure --with-umfpack --with-blas --with-lapack}
+\texttt{./configure --with-umfpack --with-blas --with-lapack} (up to 7.3.0)
+or enabling UMFPACK in the ccmake options from version 8.0.0 on.
 

### 2014-10-31 Examples/Documentation/introduction.tex
A bit more information on the flags and LD_LIBRARY_PATH for ipopt installation

@@
 script.
+
+The installation is straighforward and has been tested on openSUSE 12.1
+machines as well as on MAC. At the end of the installtion do not forget to
+add ipopt to your {\tt LD\underline{ }LIBRARY\underline{ }PATH}:
+\begin{lstlisting}
+**************************************************************
+                 Installation complete!
+Add /home/..../dopelib-2.0/ThirdPartyLibs/ipopt/lib64
+    to your $LD_LIBRARY_PATH variable
+**************************************************************
+\end{lstlisting}
 

### 2014-10-31 Examples/OPT/StatPDE/Example6/content.tex
A bit more information on the flags and LD_LIBRARY_PATH for ipopt installation

@@
 \begin{gather*}
-  \text{MidPoint: } u(0.5 ; 0.5)\\[2mm]
-  \text{MeanValue: }\int_\Omega u
+  \text{MidPoint: } u(0.125 ; 0.75)\\[2mm]
+  \text{MeanValue: }\int_\Omega |u|
 \end{gather*}
@@
 To use this algorithm you need to have IPOPT installed! \\To use this set the WITH\underline{ }IPOPT CompilerFlag.}
+If you receive this message and have the ipopt installation complete, you
+might have overseen to add ipopt to your {\tt LD\underline{ }LIBRARY\underline{ }PATH}:
+\begin{lstlisting}
+**************************************************************
+                 Installation complete!
+Add /home/..../dopelib-2.0/ThirdPartyLibs/ipopt/lib64
+    to your $LD_LIBRARY_PATH variable
+**************************************************************
+\end{lstlisting}
 

### 2015-01-16 Examples/Documentation/introduction.tex
Updated Documentation
Added CMake build system to the examples. Now in the Examples-directory the new 'make c-all' and 'make c-cat' can be used to test this new functionality

@@
 \texttt{svn co --username=[your username] \textbackslash \\ 
-\hspace*{5mm}https://ganymed.iwr.uni-heidelberg.de/svn/dope}\\
+\hspace*{5mm}https://svn.dopelib.net/dope}\\
 (The backslash in the first line means that the line will be continued as one!)
@@
   \item Michael Geiger (Examples for Plasticity, and Documentation of several PDE-Examples)
+  \item Masoud Ghaderi (Augmenting the Documentation)
+  \item Uwe K{\"o}cher (Makefile compatibility)
   \item Francesco Ludovici (Augmenting the Documentation)
-  \item Masoud Ghaderi (Augmenting the Documentation)
+  \item Matthias Maier (CMake)
 \end{itemize}
@@
 Add /home/..../dopelib-2.0/ThirdPartyLibs/ipopt/lib64
-    to your $LD_LIBRARY_PATH variable
+    to your LD\_LIBRARY\_PATH variable
 **************************************************************
@@
 \section{Installation}
+\subsection{Until Version 2.0}
 To work with DOpElib, your need necessarily deal.II, which installation 
@@
 \end{itemize}
+\subsection{After Version 2.0 - CMake build system}
+To work with DOpElib, your need necessarily deal.II, which installation 
+we describe first. Afterwards, the DOpElib installation is described and finally
+other optional packages might be installed.
+\label{sec_installation}
+\begin{itemize}
+\item Install \texttt{deal.II} to your home directory, i.e., it should be located in \\
+\texttt{\textasciitilde/deal.II}. \\[1mm]
+If you prefer another position you can install it anywhere but need 
+to set the \texttt{DEAL\_II\_DIR} environment variable, so that it 
+will be found by cmake, i.e.,
+\begin{lstlisting}
+export DEAL_II_DIR=[path to deal.ii]
+\end{lstlisting}
+Detailed installing instructions for deal.II (here the last version
+8.1.0) can be found on 
+\texttt{http://www.dealii.org/8.1.0/readme.html}. The deal.II 
+installation instructions are descriptive enough and we omit 
+any further comments and refer to their webpage.
+%
 
+%
+\item Get a copy of \texttt{DOpElib}, see Section~\ref{sec:obtain} for details.
+%
+\item Unpack in your preferred directory:
+\begin{lstlisting}
+~/Software/> tar -xzf dopelib-3.0.tar.gz
+\end{lstlisting}
+to get
+\begin{lstlisting}
+~/Software/dopelib-3.0
+\end{lstlisting}
+To build the \texttt{DOpElib} you have to change to the directory 
+\texttt{DOpEsrc} where you can call 
+\begin{lstlisting}
+~/Software/dopelib-3.0/DOpEsrc> make c-all
+\end{lstlisting}
+to build and configure the library using cmake. 
+You also get various installation options by just typing 
+\texttt{make}.
+\item To generate documentation, please go into:
+\begin{lstlisting}
+~/Software/dopelib-3.0/Examples
+\end{lstlisting}
+Here, you get all options by typing:
+\begin{lstlisting}
+make
+\end{lstlisting}
+Specifically, you can create and run all examples, running a test suite,
+and generating documentation:
+{\footnotesize
+\begin{lstlisting}
+===========================================================================
+=              Makefile for the DOpE documentation                        =
+===========================================================================
+=                                                                         =
+= The following targets exist:                                            =
+=    all       :  Make all examples (in optimized mode)                   =
+=    c-all     :  Make all examples using cmake                           =
+=    allignore :  Make all examples (in optimized mode), ignore errors.   =
+=    clean     :  Cleaning up all examples                                =
+=    tests     :  Run all test param data.                                =
+=    cat       :  Run clean, run all, run tests (combine these commands)  =
+=    c-cat     :  Run clean, run c-all, run tests (combine these commands)=
+=    catignore :  Same as cat, but ignores compiler errors.               =
+=    pdf-doc   :  Create documentation in pdf file format via latexmk     =
+=    doc       :  Create documentation in pdf file format via latexmk     =
+=    distclean :  Cleaning up, including documentation                    =
+=    warncheck :  Checks whether all Examples compile without warnings    =
+===========================================================================
+\end{lstlisting}
+}
+The new options are $c-all$ and $c-cat$ utilizing the cmake build system.
+For html documentation, change into:
+\begin{lstlisting}
+~/Software/dopelib-3.0/doxygen
+\end{lstlisting}
+and then typing \texttt{make}. Doxygen documentation
+will require either 
+latex or doxygen to be installed on your computer.
+\item If you wish to test if everything worked. To do so you can 
+change to the \texttt{Examples} directory and \texttt{make tests} which will give you a 
+list of all the examples and whether they behave as expected by the library, see also 
+Chapter~\ref{chap:test}.
+\item  If you want to use some of the supported third party libraries install them and follow 
+ the instructions in \texttt{ThirdPartyLibs/README}. There may be further information 
+ in some \texttt{ThirdPartyLibs/*.INSTALLNOTES} that you may want to
+ consider.\\[3mm]
+ As example, the installation of ipopt works as follows:\\
+In the path \texttt{~/dope/ThirdPartyLibs>} type in the terminal\\
+\texttt{./install-free-libs.sh}
+
+\end{itemize}
 \section{FAQs}

### 2015-11-27 Examples/PDE/StatPDE/Example13/content.tex
Started Work on DG-Example (unfinished)

\subsubsection{General problem description}
Within this example, we demonstrate how to use the
dG (discontinuous Galerkin) method for the solution of 
a transport equation. The example corresponds to 
the dealii step-12.

\subsubsection{Implementational Details}

### 2015-11-27 Examples/PDE/StatPDE/Example13/functionals.h

@@ -1,0 +1,94 @@
+/**
+*
+* Copyright (C) 2012-2014 by the DOpElib authors
+*
+* This file is part of DOpElib
+*
+* DOpElib is free software: you can redistribute it
+* and/or modify it under the terms of the GNU General Public
+* License as published by the Free Software Foundation, either
+* version 3 of the License, or (at your option) any later
+* version.
+*
+* DOpElib is distributed in the hope that it will be
+* useful, but WITHOUT ANY WARRANTY; without even the implied
+* warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+* PURPOSE.  See the GNU General Public License for more
+* details.
+*
+* Please refer to the file LICENSE.TXT included in this distribution
+* for further information on this license.
+*
+**/
+
+
+#ifndef FUNCTIONALS_H_
+#define FUNCTIONALS_H_
+
+#include "functionalinterface.h"
+
+using namespace std;
+using namespace dealii;
+using namespace DOpE;
+
+/****************************************************************************************/
+
+template<
+template<template<int, int> class DH, typename VECTOR, int dealdim> class EDC,
+  template<template<int, int> class DH, typename VECTOR, int dealdim> class FDC,
+  template<int, int> class DH, typename VECTOR, int dealdim>
+  class MeanValueFunctional : public FunctionalInterface<EDC, FDC, DH, VECTOR, dealdim>
+  {
+    public:
+      MeanValueFunctional()
+      {
+      }
+
+      double
+      ElementValue(const EDC<DH,VECTOR,dealdim>& edc)
+      {
+        unsigned int n_q_points = edc.GetNQPoints();
+
+        double mean = 0;
+
+	vector<double> uvalues;
+	uvalues.resize(n_q_points);
+	edc.GetValuesState("state", uvalues);
+	
+	for (unsigned int q_point = 0; q_point < n_q_points; q_point++)
+	{
+	  double v;
+	  
+	  v = uvalues[q_point];
+	  
+	  mean += v * edc.GetFEValuesState().JxW(q_point);
+	}
+        return mean;
+      }
+
+      UpdateFlags
+      GetUpdateFlags() const
+      {
+        return update_values | update_quadrature_points;
+      }
+
+      string
+      GetType() const
+      {
+        return "domain";
+      }
+
+      bool HasFaces() const
+      {
+        return false;
+      }
+
+      string
+      GetName() const
+      {
+        return "Mean-value";
+      }
+
+    private:
+  };
+#endif /* FUNCTIONALS_H_ */

### 2015-11-27 Examples/PDE/StatPDE/Example13/myfunctions.h

@@ -1,0 +1,65 @@
+/**
+ *
+ * Copyright (C) 2012-2014 by the DOpElib authors
+ *
+ * This file is part of DOpElib
+ *
+ * DOpElib is free software: you can redistribute it
+ * and/or modify it under the terms of the GNU General Public
+ * License as published by the Free Software Foundation, either
+ * version 3 of the License, or (at your option) any later
+ * version.
+ *
+ * DOpElib is distributed in the hope that it will be
+ * useful, but WITHOUT ANY WARRANTY; without even the implied
+ * warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+ * PURPOSE.  See the GNU General Public License for more
+ * details.
+ *
+ * Please refer to the file LICENSE.TXT included in this distribution
+ * for further information on this license.
+ *
+ **/
+
+#ifndef MYFUNCTIONS_H_
+#define MYFUNCTIONS_H_
+
+using namespace std;
+using namespace dealii;
+
+#include <deal.II/base/numbers.h>
+#include "function_wrapper.h"
+
+namespace DOpE
+{
+  class ExactSolution : public DOpEWrapper::Function<2>
+  {
+    public:
+      ExactSolution() :
+          DOpEWrapper::Function<2>(1)
+      {
+      }
+
+      virtual double
+      value(const Point<2> &p, const unsigned int component = 0) const;
+
+  };
+
+  /******************************************************/
+
+  double
+    ExactSolution::value(const Point<2> &p, const unsigned int /*component*/) const
+  {
+    const double x = p[0];
+    const double y = p[1];
+    if (x*x+y*y <= 0.25)
+      return 1.;
+    else
+      return 0.;
+  }
+
+  /******************************************************/
+
+}
+
+#endif /* MYFUNCTIONS_H_ */

### 2015-12-02 Examples/Documentation/introduction.tex
Updated Documentation

@@
   \item Christian Goll (Heidelberg University)
-  \item Thomas Wick (The University of Texas at Austin)
-  \item Winnifried Wollner (University of Hamburg)
+  \item Thomas Wick (Johann Radon Institute for Computational and Applied Mathematics (RICAM))
+  \item Winnifried Wollner (Technische Universität Darmstadt)
+\end{itemize}
+
+We would like to express our gratitude to the former maintainer
+\begin{itemize}
+\item Christian Goll (Maintainer from 2009--2015)
 \end{itemize}
@@
 machines as well as on MAC. At the end of the installtion do not forget to
-add ipopt to your {\tt LD\underline{ }LIBRARY\underline{ }PATH}:
+add ipopt to your \texttt{LD\underline{ }LIBRARY\underline{ }PATH}:
 \begin{lstlisting}
@@
 
-\section{Installation}
+\section{Installation}\label{sec_installation}
 \subsection{Until Version 2.0}
@@
 other optional packages might be installed.
-\label{sec_installation}
 \begin{itemize}
@@
 \begin{itemize}
-\item {\bf unrecognized command line option "-std=gnu++0x"}\\
+\item \textbf{unrecognized command line option "-std=gnu++0x"}\\
   This means that your compiler is too old. You can check the 
@@
   \item \texttt{instatpdeproblemcontainer.h} 
-%{\bf TODO is still missing but will be used for nonstationary problems.} 
+%\textbf{TODO is still missing but will be used for nonstationary problems.} 
 This will be implemented once we have nonstationary optimization problems running to avoid error duplication in the coding process.
@@
   an interface to the dimension independent functionality of all space time dof handlers.
-  %{\bf TODO: Beispiele geben}
+  %\textbf{TODO: Beispiele geben}
 \item \texttt{basic/statespacetimehandler.h} Another intermediate interface class which adds 
@@
   \item \texttt{problemdata/stateproblem.h} This is the problem description for the (forward/primal) PDE constraint.
-    {\bf Similar descriptors will be build for the other problems (adjoint, tangent, $\ldots$) when time allows.}
+    \textbf{Similar descriptors will be build for the other problems (adjoint, tangent, $\ldots$) when time allows.}
   \item \texttt{problemdata/initialproblem.h} This is the problem descriptor to compute the finite element representation
@@
     DoFHandler and HPDoFHandler.
-  \item \texttt{wrapper/fevalues\underline{ }wrapper.h} {\bf Will be removed soon!}
+  \item \texttt{wrapper/fevalues\underline{ }wrapper.h} \textbf{Will be removed soon!}
   \item \texttt{wrapper/function\underline{ }wrapper.h} An interface that allows to use functions that depend 

### 2015-12-02 Examples/OPT/InstatPDE/Example1/content.tex
Updated Documentation

@@
 
-{\bf Note:} In contrast to the ,,normal'' Element-terms in the PDE we assume in the program 
+\textbf{Note:} In contrast to the ,,normal'' Element-terms in the PDE we assume in the program 
 that the \texttt{Init\underline{ }ElementEquation} is linear in the state and no other solution variables are present.

### 2015-12-02 Examples/OPT/StatPDE/Example6/content.tex
Updated Documentation

@@
 
-The Problem is similar to that of {\tt OPT/StatPDE/Example1} except for the box 
+The Problem is similar to that of \texttt{OPT/StatPDE/Example1} except for the box 
 control constraints.
@@
 The problem is solved using the optimization library IPOPT that you can obtain for free. To use it 
-a correct link to the ipopt library needs to be created in  {\tt DOpE/ThirdPartyLibs} by the name 
-{\tt ipopt}, i.e., you should have the file
-{\tt DOpE/ThirdPartyLibs/ipopt} pointing to the ipopt directory. If you have not done this you can compile the
+a correct link to the ipopt library needs to be created in  \texttt{DOpE/ThirdPartyLibs} by the name 
+\texttt{ipopt}, i.e., you should have the file
+\texttt{DOpE/ThirdPartyLibs/ipopt} pointing to the ipopt directory. If you have not done this you can compile the
 example but when running the example you will only get an error message like\\
-{\tt Warning: During execution of `Reduced\underline{ }IpoptAlgorithm::Solve`\\ the following Problem occurred!\\
+\texttt{Warning: During execution of `Reduced\underline{ }IpoptAlgorithm::Solve`\\ the following Problem occurred!\\
 To use this algorithm you need to have IPOPT installed! \\To use this set the WITH\underline{ }IPOPT CompilerFlag.}
 If you receive this message and have the ipopt installation complete, you
-might have overseen to add ipopt to your {\tt LD\underline{ }LIBRARY\underline{ }PATH}:
+might have overseen to add ipopt to your \texttt{LD\underline{ }LIBRARY\underline{ }PATH}:
 \begin{lstlisting}
@@
 the snopt directory (where you have the libs and the header files) 
-in the {\tt DOpE/ThirdPartyLibs} directory named {\tt snopt}, i.e., you should have the file
-{\tt DOpE/ThirdPartyLibs/snopt} pointing to the snopt directory. If you have not done this you can compile the
+in the \texttt{DOpE/ThirdPartyLibs} directory named \texttt{snopt}, i.e., you should have the file
+\texttt{DOpE/ThirdPartyLibs/snopt} pointing to the snopt directory. If you have not done this you can compile the
 example but when running the example you will only get an error message like\\
-{\tt Warning: During execution of `Reduced\underline{ }SnoptAlgorithm::Solve`\\ the following Problem occurred!\\
+\texttt{Warning: During execution of `Reduced\underline{ }SnoptAlgorithm::Solve`\\ the following Problem occurred!\\
 To use this algorithm you need to have SNOPT installed! \\To use this set the WITH\underline{ }SNOPT CompilerFlag.}

### 2015-12-02 Examples/OPT/StatPDE/Example8/content.tex
Update Doku in Einklang mit Christians und Winnis Bemerkungen/Zustimmung
Updated Documentation

@@
 This example implements the topology optimization of an MBB-Beam given in \\
-{\tt OPT/StatPDE/Example7} using the SIMP method.
+\texttt{OPT/StatPDE/Example7} using the SIMP method.
 
 The solution is computed using the commercial optimization library SNOPT, similar 
-to {\tt OPT/StatPDE/Example6} where IPOPT is used. 
+to \texttt{OPT/StatPDE/Example6} where IPOPT is used. 
 This Example demonstrate how global constraints on
@@
 \subsubsection{Program description}
-\textcolor{red}{
 In this example, we have now some local (point) constraints, local constraints and a global
@@
 in the main file.
-}

### 2015-12-02 Examples/PDE/InstatPDE/Example3/content.tex
Updated Documentation

@@
 \index{instationary PDE} \index{instationary PDE!Black-Scholes equation}\\
-To state the general form of the equation we need some nomenclature: We consider an option on $d$ risky assets with {\it maturity} $T>0$ and {\it strikeprice} $K>0$.
-For the sake of simplicity we assume the {\it interest rate} $r>0$ and the {\it volatility} of the i-th asset $\sigma_i>0$, $1\leq i\leq d$, to be constant.
-Besides, we assume the matrix $\rho=(\rho_{ij})$ of the {\it correlation factors} $\rho_{ij}$ with  $-1\leq \rho_{ij}\leq 1$ for  $1\leq i,j \leq d$, to be positive definite.
+To state the general form of the equation we need some nomenclature: We consider an option on $d$ risky assets with \textit{maturity} $T>0$ and \textit{strikeprice} $K>0$.
+For the sake of simplicity we assume the \textit{interest rate} $r>0$ and the \textit{volatility} of the i-th asset $\sigma_i>0$, $1\leq i\leq d$, to be constant.
+Besides, we assume the matrix $\rho=(\rho_{ij})$ of the \textit{correlation factors} $\rho_{ij}$ with  $-1\leq \rho_{ij}\leq 1$ for  $1\leq i,j \leq d$, to be positive definite.
 Of course $\rho$ is symmetric with $\rho_{ii}=1$.\\
@@
 \end{subequations}
-The initial condition $u_0\in C^{0}\left({\mathbb{R}_+^d}\right) $ (i.e. the {\it payoff}) is given depending of the type of the option. For example
+The initial condition $u_0\in C^{0}\left({\mathbb{R}_+^d}\right) $ (i.e. the \textit{payoff}) is given depending of the type of the option. For example
 \begin{equation}

### 2015-12-02 Examples/description.tex
Started Work on DG-Example (unfinished)
Updated Documentation

@@
 \author{
-  Christian Goll\thanks{Heidelberg University, 
-    {christian.goll@iwr.uni-heidelberg.de}}, 
   Thomas Wick\thanks{Johann Radon Institute for Computational and Applied Mathematics (RICAM), 
     {thomas.wick@ricam.oeaw.ac.at}}, 
-  Winnifried Wollner\thanks{University of Hamburg, 
-    {winnifried.wollner@uni-hamburg.de}}
+  Winnifried Wollner\thanks{Technische Universität Darmstadt, 
+    {wollner@mathematik.tu-darmstadt.de}}
 }
@@
 \clearpage
+\subsection{Discontinuous Galerkin}
+\label{PDE_dG}
+\input{PDE/StatPDE/Example13/content.tex}
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-12-03 Examples/PDE/InstatPDE/Example7/content.tex
Added Nonstationary Example (isothermal Euler) with dG discretization in time + Bugfixes

\subsubsection{General problem description}
This example solves the isothermal Euler equations
\begin{align*}
 \partial_t(\rho v) + \partial_x (p(\rho)+ \rho v^2) &= - \frac{\lambda}{2D} v|v| - gh' \rho\\
 \partial_t \rho + \partial_x (\rho v) &=0
\end{align*}
on the set $t \in (0,1)$, $x \in (0,2)$.
The relation 
\[
p(\rho) = \frac{RT\, \rho}{1-\alpha RT\,\rho}
\]
depends on the user provided data for gas-constant $R$ and temperature $T$ as well as a parameter 
$\alpha$ ($\alpha = 0$ for ideal gases).
The other parameters in the system to be provided by the user are 
the friction parameter $\lambda$, the diameter $D$ of the pipe, the gravity $g$ and the slope $h'$ of the 
pipe.

The discretization is done by a dG-method in space.

### 2015-12-03 Examples/PDE/InstatPDE/Example7/localpde.h

@@ -1,0 +1,777 @@
+/**
+ *
+ * Copyright (C) 2012-2014 by the DOpElib authors
+ *
+ * This file is part of DOpElib
+ *
+ * DOpElib is free software: you can redistribute it
+ * and/or modify it under the terms of the GNU General Public
+ * License as published by the Free Software Foundation, either
+ * version 3 of the License, or (at your option) any later
+ * version.
+ *
+ * DOpElib is distributed in the hope that it will be
+ * useful, but WITHOUT ANY WARRANTY; without even the implied
+ * warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+ * PURPOSE.  See the GNU General Public License for more
+ * details.
+ *
+ * Please refer to the file LICENSE.TXT included in this distribution
+ * for further information on this license.
+ *
+ **/
+
+#ifndef LOCALPDE_H_
+#define LOCALPDE_H_
+
+#include "pdeinterface.h"
+#include "myfunctions.h"
+#include <deal.II/base/numbers.h>
+
+using namespace std;
+using namespace dealii;
+using namespace DOpE;
+
+/***********************************************************************************************/
+template<
+    template<template<int, int> class DH, typename VECTOR, int dealdim> class EDC,
+    template<template<int, int> class DH, typename VECTOR, int dealdim> class FDC,
+    template<int, int> class DH, typename VECTOR, int dealdim>
+  class LocalPDE : public PDEInterface<EDC, FDC, DH, VECTOR, dealdim>
+  {
+  public:
+  LocalPDE(ParameterReader &param_reader) :
+    state_block_component_(2, 0)
+    {
+      param_reader.SetSubsection("localpde parameters");
+      R_      = param_reader.get_double("R");
+      T_      = param_reader.get_double("T");
+      alpha_  = param_reader.get_double("alpha");
+      lambda_ = param_reader.get_double("lambda");
+      D_      = param_reader.get_double("D");
+      g_      = param_reader.get_double("g");
+      hprime_ = param_reader.get_double("hprime");
+
+      stab_param_ = 1.;
+    }
+
+      void
[…]

### 2016-04-01 Examples/Documentation/introduction.tex
Removed warnings when compiling with deal 8.4.0

@@
 cmake, which makes the inclusion of external packages easier.
-Detailed installing instructions for deal.II (here the last version
-8.1.0) can be found on\\ 
-\texttt{http://www.dealii.org/8.1.0/readme.html}.
+Detailed installing instructions for deal.II can be found on\\ 
+\texttt{http://www.dealii.org/}.
+
+In deal.II 8 and newer a configuration of deal.II working with \texttt{DOpElib} can be done with
+the following commands assuming that one is currently in the 
+directory where the deal.II sources are located.
+\begin{lstlisting}
+mkdir build
+mkdir install
+cd build
+cmake -DCMAKE_INSTALL_PREFIX=../install -DDEAL_II_WITH_UMFPACK=true -DDEAL_II_FORCE_BUNDLED_UMFPACK=true  ../
+make install
+\end{lstlisting}
+If in addition, features are desired corresponding flags need to be passed to cmake. If, for 
+example Trilinos solvers are desired, the above cmake command needs to have the additional flags
+
+\begin{lstlisting}
+-DDEAL_II_WITH_TRILINOS=true -DTRILINOS_DIR=[path to trilinos]
+\end{lstlisting}
 

### 2016-04-01 DOpEsrc/container/pdeproblemcontainer.h

@@ -2222,25 +2222,6 @@ namespace DOpE
 
   template<typename PDE, typename DD, typename SPARSITYPATTERN, typename VECTOR,
       int dealdim, template<int, int> class FE, template<int, int> class DH>
-    unsigned int
-    PDEProblemContainer<PDE, DD, SPARSITYPATTERN, VECTOR, dealdim, FE, DH>::GetDoFsPerBlock(
-        unsigned int b) const
-    {
-      if ((this->GetType() == "state") || (this->GetType() == "adjoint_for_ee"))
-      {
-        return GetSpaceTimeHandler()->GetStateDoFsPerBlock(b);
-      }
-      else
-      {
-        throw DOpEException("Unknown Type:" + this->GetType(),
-            "PDEProblemContainer::GetDoFsPerBlock");
-      }
-    }
-
-  /******************************************************/
-
-  template<typename PDE, typename DD, typename SPARSITYPATTERN, typename VECTOR,
-      int dealdim, template<int, int> class FE, template<int, int> class DH>
     const std::vector<unsigned int>&
     PDEProblemContainer<PDE, DD, SPARSITYPATTERN, VECTOR, dealdim, FE, DH>::GetDoFsPerBlock() const
     {

### 2016-04-01 Examples/Documentation/introduction.tex
Removed old makefiles parts

@@
 
-When installing \texttt{deal.II} you should take care to configure 
-it to use UMFPACK, i.e.,\\
-\texttt{./configure --with-umfpack}
-in version up to 7.3.0. From 8.0.0 on, deal.II is compiled with 
-cmake, which makes the inclusion of external packages easier.
-Detailed installing instructions for deal.II can be found on\\ 
-\texttt{http://www.dealii.org/}.
+When installing \texttt{deal.II} (at least version 8.0) you should take care to configure 
+it to use UMFPACK.
 
-In deal.II 8 and newer a configuration of deal.II working with \texttt{DOpElib} can be done with
+A configuration of deal.II working with \texttt{DOpElib} can be obtained with
 the following commands assuming that one is currently in the 

### 2016-04-05 Examples/PDE/StatPDE/Example10/content.tex
Added Trilinos Direct Solvers and demonstrate them in example PDE/StatPDE/Example10

@@
 Furthermore, we see that several Newton steps are needed on each refinement level; this is due to the nonlinearity of the current problem.
+
+Finally this example demonstrates how to use the direct solvers provided by trilinos. 
+This is done by the line 
+\begin{verbatim}
+typedef TrilinosDirectLinearSolverWithMatrix
+         <SPARSITYPATTERN, MATRIX, VECTOR> LINEARSOLVER;
+\end{verbatim}
+interfacing to the trilinos library. The selection of the precise direct solver can 
+be done using the parameter file.

### 2016-04-08 Examples/Documentation/introduction.tex
Minor Typo in description

@@
 cd build
-cmake -DCMAKE_INSTALL_PREFIX=../install -DDEAL_II_WITH_UMFPACK=true -DDEAL_II_FORCE_BUNDLED_UMFPACK=true  ../
+cmake -DCMAKE_INSTALL_PREFIX=../install \ 
+      -DDEAL_II_WITH_UMFPACK=true \
+      -DDEAL_II_FORCE_BUNDLED_UMFPACK=true  ../
 make install

### 2016-05-13 Examples/OPT/StatPDE/Example10/content.tex

\subsubsection{General problem description}
This example is similar to Example~\ref{OPT_Stat_Param_Nonlin_Fluid}.
The notable difference in the setting is that now, the boundary control areas, 
where the fluid can be sucked out of the domain, are in front of the 
circular inclusion. Hense the drag
\[
k(v,p) = \int_{\Gamma_O} n\cdot \sigma(v,p)\cdot d \, ds,
\]
where $\Gamma_O$ denotes the cylinder boundary, and $d$ is a vector in the
direction of the mean in-flow can become negative, i.e., minimizing the 
value of $k(v,p)$ is no longer viable.

Instead, we consider the objective 
\[
K(q,v,p) = |k(v,p)| + \frac{\alpha}{2}||q - q_0||^2
\] 
is to be minimized.

In contrast to all previous examples, this means that the functional can
no longer be calculated by one integration, but instead the values of the 
integration (for the drag) need to be post-processed.

### 2016-05-13 Examples/OPT/StatPDE/Example10/main.cc

@@ -1,0 +1,291 @@
+/**
+ *
+ * Copyright (C) 2012-2014 by the DOpElib authors
+ *
+ * This file is part of DOpElib
+ *
+ * DOpElib is free software: you can redistribute it
+ * and/or modify it under the terms of the GNU General Public
+ * License as published by the Free Software Foundation, either
+ * version 3 of the License, or (at your option) any later
+ * version.
+ *
+ * DOpElib is distributed in the hope that it will be
+ * useful, but WITHOUT ANY WARRANTY; without even the implied
+ * warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+ * PURPOSE.  See the GNU General Public License for more
+ * details.
+ *
+ * Please refer to the file LICENSE.TXT included in this distribution
+ * for further information on this license.
+ *
+ **/
+
+#include <iostream>
+#include <fstream>
+
+#include <deal.II/grid/tria.h>
+#include <deal.II/grid/grid_in.h>
+#include <deal.II/grid/tria_boundary_lib.h>
+#include <deal.II/grid/grid_generator.h>
+#include <deal.II/fe/fe_q.h>
+#include <deal.II/fe/fe_dgp.h>
+#include <deal.II/fe/fe_nothing.h>
+#include <deal.II/base/quadrature_lib.h>
+#include <deal.II/numerics/error_estimator.h>
+#include <deal.II/grid/grid_refinement.h>
+
+#include "reducednewtonalgorithm.h"
+#include "optproblemcontainer.h"
+#include "functionalinterface.h"
+#include "statreducedproblem.h"
+#include "newtonsolver.h"
+#include "directlinearsolver.h"
+#include "voidlinearsolver.h"
+#include "integrator.h"
+#include "noconstraints.h"
+#include "solutionextractor.h"
+#include "integratormixeddims.h"
+#include "newtonsolvermixeddims.h"
+#include "parameterreader.h"
+#include "mol_spacetimehandler.h"
+#include "simpledirichletdata.h"
+#include "integratordatacontainer.h"
+
+#include "localpde.h"
+#include "localfunctional.h"
+#include "functionals.h"
[…]

### 2016-05-13 Examples/OPT/StatPDE/Example10/content.tex
Further work on optimization with functionals not given by a single integration. Calculation of such functionals is now working, derivative calculation still missing

@@
 \[
-K(q,v,p) = |k(v,p)| + \frac{\alpha}{2}||q - q_0||^2
+K(q,v,p) = |k(v,p)|^2 + \frac{\alpha}{2}||q - q_0||^2
 \] 
@@
 integration (for the drag) need to be post-processed.
+
+To do so, the calculation of the functional and its derivatives
+is reordered in to two steps. First the value of $k(v,p)$ is
+calculated (and stored) then in a second sweep. The value of
+$K$ is calculated.
+
+To this end the following modifications are needed in the
+code:
+
+\textbf{\texttt{localpde.h}}
+There is a new method \texttt{unsigned int NeedPrecomputations() const}
+returning the value $1$ as we need one calculation of $k$ before we can
+assemble the value of the cost-functional.
+
+This pre-iteration has the Type \texttt{cost\_functional\_pre} and a corresponding
+number (here $0$ as only one pre-iteration is performed.
+
+In the cost functional, for the pre-iteration we set the type to
+\texttt{boundary} since $k$ is a boundary functional.
+For the evaluation of $K$ itself the type is \texttt{boundary algebraic}
+as we calculate the boundary integral $\|q\|^2$ and the algebraic calculation
+$|{k}|^2$.
+
+For higher derivatives, we notice that for a differentiable functions $g,f\colon \mathbb R\rightarrow \mathbb R$
+it holds for the directional derivative in direction $\delta u$
+\begin{align*}
+ \left((g\left(\int f(u(x))\,\mathrm{d}x\right)\right)'\delta u
+  &= g'(\int f(u(x))\,\mathrm{d}x) \int f'(u(x)) \delta u(x) \mathrm{d}x \\
+  &= \int g'(\int f(u(x))\,\mathrm{d}x) f'(u(x)) \delta u(x) \mathrm{d}x 
+\end{align*}
+Consequently, the first derivative can be calculated with only a single integration
+-- as a boundary integral in the present example.

### 2016-05-18 Examples/OPT/InstatPDE/Example4/content.tex

\subsubsection{General problem description}

This example is a modified version of \texttt{OPT/InstatPDE/Example1}. 
Again, we consider the heat equation, this time with an additional nonlinear
term and most important a time derivative leading to a 
nonstationary optimization problem. The governing equation is
\begin{equation*}
\partial_t u(t,x,y) - \Delta u(t,x,y) + u(t,x,y)^2 = f(t,x,y),
\end{equation*}
with homogeneous Dirichlet-data.
The computational domain is $\Omega \times I = [0,\pi]^2 \times [0,1]$ with the data
\begin{align*}
f(t,x,y) &= (3-2t)e^{t-t^2} \sin(x) \sin(y) + e^{(t-t^2)^2} \sin^2(x) \sin^2(y),\\
u_0(x,y) &= q(x,y).
\end{align*}

In contrast to \texttt{OPT/InstatPDE/Example1}, we modify the cost-functional
such that a (spatial) mean-value of zero is desired for the state.
The corresponding functional is
\[
 \min_{q,u} J(q,u) = \int_0^1 \frac{1}{2}\left|\int_{\Omega} u(t,x,y) \,\mathrm{d}(x,y)\right|^2 \,\mathrm{d}t + \frac{1}{2} \int_{\Omega} (q(x,y) - \sin(x) \sin(y))^2\,\mathrm{d}(x,y).
\]
\subsubsection{Program description}
The following things differ from the previous examples:\\[2mm]
We are now considering a cost-functional, that can not be evaluated as one single integral. Instead,
we need to evaluate the integral $\int_{\Omega} u(t,x,y) \,d(x,y)$ in space first in each time step and 
then integrate over its absolute value in time. We have seen a similar structure in 
\texttt{OPT/StatPDE/Example10}~\ref{OPT_Stat_Param_Nonlin_Fluid_extension}. Here, the integral 
in question is of the form
\[
\int_I g(\int f(u(t,x)) \mathrm{d}x) \mathrm{d}t
\]
Consequently, we need to set \texttt{unsigned int NeedPrecomputations() const} to the value $1$ 
as we need to calculate -- in each time point -- the values $\tilde{f}(t) = \int f(u(t,x)) \mathrm{d}x$ 
prior to the evaluation of the cost functional.
Then the first derivative of the functional in the direction $\delta u$ is given as 
\begin{align*}
   \left(\int_I(g\left(\tilde{f}(t)\right)\mathrm{d}t\right)'\delta u
  &= \int_I g'(\tilde{f}(t)) \int f'(u(t,x)) \delta u(t,x) \mathrm{d}x\;\mathrm{d}t \\
  &= \int_I \int g'(\tilde{f}(t)) f'(u(t,x)) \delta u(t,x) \mathrm{d}x\;\mathrm{d}t
\end{align*}
And for the second derivative, we also calculate the tangential derivatives 
\[
\tilde{f}_{u}(t) = \int f'(u(t,x)) \delta u(t,x) \mathrm{d}x = \int_{\Omega} \delta u(t,x,y) \,\mathrm{d}(x,y).
\]
With this the second derivative in direction $\delta u $ and $\tau u$ is given as 
\begin{align*}
   \Biggl(\int_Ig&\left(\tilde{f}(t)\right)\mathrm{d}t\Biggr)''(\delta u,\tau u)\\
  &= \int_I g''(\tilde{f}(t)) \tilde{f}_{u}(t) \int f'(u(t,x)) \tau u(t,x) \mathrm{d}x\;\mathrm{d}t \\
  &\;\;\;\;+ \int_I \int g'(\tilde{f}(t)) f''(u(t,x)) \delta u(t,x) \tau u(t,x) \mathrm{d}x\;\mathrm{d}t\\
&= \int_I \int \left( g''(\tilde{f}(t)) \tilde{f}_{u}(t)f'(u(t,x))+ g'(\tilde{f}(t)) f''(u(t,x)) \delta u(t,x) \right) \tau u(t,x) \mathrm{d}x\;\mathrm{d}t.
\end{align*}
As in \texttt{OPT/StatPDE/Example10} the higher derivatives can be calculated by one space-time 
integral if the needed values $\tilde{f}$ and $\tilde{f}_u$ are available at the 
different time-points. These values can be accessed as ParamValues with the respective 
labels \texttt{cost\_functional\_pre} and \texttt{cost\_functional\_pre\_tangent}.

One further thing differs from the previous examples. This is that the cost functional 
now has one part being integrated over space-time while the other 
is acting in space (at initial-time) only. The two types of functionals can at present 
not be mixed. Hence both terms are evaluated as a distributed integral 
over space-time by adding an additional weight to the control-cost integral.
This weight needs to be specially modified for the derivatives w.r.t. the control
which are evaluated at the initial time only!
This modification is not exact and consequently the derivatives are not exactly calculated. 
The accuracy is increased as the temporal meshes are refined. Hence, if more accuracy 
in the residual is desired the temporal mesh needs to be further refined.

### 2016-05-18 Examples/PDE/InstatPDE/Example2/localfunctional.h

@@ -38,21 +38,14 @@ template<
       dopedim, dealdim>
   {
     public:
-      LocalFunctional() : time_(0)
+      LocalFunctional() 
       {
-      }
-
-      // include NeedTime
-      void
-      SetTime(double t) const
-      {
-        time_ = t;
       }
 
       bool
       NeedTime() const
       {
-        if (fabs(time_ - 1.) < 1.e-13)
+        if (fabs(this->GetTime() - 1.) < 1.e-13)
           return true;
         return false;
       }
@@ -80,10 +73,6 @@ template<
       {
         return "dummy functional";
       }
-
-    private:
-
-      mutable double time_;
 
   };
 #endif

### 2016-05-19 Examples/Documentation/introduction.tex
Updated Documentation - removed the details for early dope versions

@@
 \subsection{g++}
-\texttt{DOpElib} requires a recent \texttt{g++} (at least 4.4.3) due to 
-some new \texttt{C++} features implemented in \texttt{C++0x} and \texttt{C++11}. 
+\texttt{DOpElib} requires a recent \texttt{g++} (at least 4.8) due to 
+some new \texttt{C++} features implemented in \texttt{C++11}. 
 You can check the version number using the command-line argument \texttt{g++ -v}.
@@
 after that you can install a new version of \texttt{g++} using for instance 
-\texttt{sudo port install gcc45} to install version $4.5$ of the compiler.
+\texttt{sudo port install gcc48} to install version $4.8$ of the compiler.
 Afterwards, you need to set the search path appropriate to find the macports version
@@
                  Installation complete!
-Add /home/..../dopelib-2.0/ThirdPartyLibs/ipopt/lib64
+Add /home/..../dopelib-4.0/ThirdPartyLibs/ipopt/lib64
     to your LD\_LIBRARY\_PATH variable
@@
 \section{Installation}\label{sec_installation}
-\subsection{Until Version 2.0}
+\subsection{After Version 2.0 - CMake build system}
 To work with DOpElib, your need necessarily deal.II, which installation 
@@
 \texttt{\textasciitilde/deal.II}. \\[1mm]
-If you would like to have another path you will either have to manually edit the files
-\texttt{DOpEsrc/source/Makefile} and \texttt{Examples/Make.global\underline{ }options} and replace the 
-line \\
-\texttt{D = \$(HOME)/deal.II} with the appropriate path.\\[1mm]
+If you prefer another position you can install it anywhere but need 
+to set the \texttt{DEAL\_II\_DIR} environment variable, so that it 
+will be found by cmake, i.e.,
+\begin{lstlisting}
+export DEAL_II_DIR=[path to deal.ii]
+\end{lstlisting}
 Detailed installing instructions for deal.II (here the last version
-8.1.0) can be found on 
-\texttt{http://www.dealii.org/8.1.0/readme.html}. The deal.II 
+8.4.0) can be found on 
+\texttt{http://www.dealii.org/8.4.0/readme.html}. The deal.II 
 installation instructions are descriptive enough and we omit 
@@
 \begin{lstlisting}
-~/Software/> tar -xzf dopelib-2.0.tar.gz
+~/Software/> tar -xzf dopelib-4.0.tar.gz
 \end{lstlisting}
@@
 \begin{lstlisting}
-~/Software/dopelib-2.0
+~/Software/dopelib-4.0
 \end{lstlisting}
@@
 \begin{lstlisting}
-~/Software/dopelib-2.0/DOpEsrc> make all
+~/Software/dopelib-4.0/DOpEsrc> make c-all
 \end{lstlisting}
-to build the library. You also get various installation options by just typing 
+to build and configure the library using cmake. 
+You also get various installation options by just typing 
 \texttt{make}.
@@
 \begin{lstlisting}
-~/Software/dopelib-2.0/Examples
+~/Software/dopelib-4.0/Examples
 \end{lstlisting}
@@
 = The following targets exist:                                            =
-=    all       :  Make all examples (in optimized mode)                   =
-=    allignore :  Make all examples (in optimized mode), ignore errors.   =
+=    c-all     :  Make all examples using cmake                           =
 =    clean     :  Cleaning up all examples                                =
 =    tests     :  Run all test param data.                                =
-=    cat       :  Run clean, run all, run tests (combine these commands)  =
-=    catignore :  Same as cat, but ignores compiler errors.               =
-=    pdf-doc   :  Create documentation in pdf file format via latexmk     =
+=    c-cat     :  Run clean, run c-all, run tests (combine these commands)=
 =    doc       :  Create documentation in pdf file format via latexmk     =
@@
 }
+The options are $c-all$ and $c-cat$ utilizing the cmake build system.
 For html documentation, change into:
 \begin{lstlisting}
-~/Software/dopelib-2.0/doxygen
-\end{lstlisting}
-and then typing \texttt{make}. Doxygen documentation
-will require either 
-latex or doxygen to be installed on your computer.
-\item If you wish to test if everything worked. To do so you can 
-change to the \texttt{Examples} directory and \texttt{make tests} which will give you a 
-list of all the examples and whether they behave as expected by the library, see also 
-Chapter~\ref{chap:test}.
-\item  If you want to use some of the supported third party libraries install them and follow 
- the instructions in \texttt{ThirdPartyLibs/README}. There may be further information 
- in some \texttt{ThirdPartyLibs/*.INSTALLNOTES} that you may want to
- consider.\\[3mm]
- As example, the installation of ipopt works as follows:\\
-In the path \texttt{~/dope/ThirdPartyLibs>} type in the terminal\\
-\texttt{./install-free-libs.sh}
-
-\end{itemize}
-\subsection{After Version 2.0 - CMake build system}
-To work with DOpElib, your need necessarily deal.II, which installation 
-we describe first. Afterwards, the DOpElib installation is described and finally
-other optional packages might be installed.
-\begin{itemize}
-\item Install \texttt{deal.II} to your home directory, i.e., it should be located in \\
-\texttt{\textasciitilde/deal.II}. \\[1mm]
-If you prefer another position you can install it anywhere but need 
-to set the \texttt{DEAL\_II\_DIR} environment variable, so that it 
-will be found by cmake, i.e.,
-\begin{lstlisting}
-export DEAL_II_DIR=[path to deal.ii]
-\end{lstlisting}
-Detailed installing instructions for deal.II (here the last version
-8.1.0) can be found on 
-\texttt{http://www.dealii.org/8.1.0/readme.html}. The deal.II 
-installation instructions are descriptive enough and we omit 
-any further comments and refer to their webpage.
-%
-
-%
-\item Get a copy of \texttt{DOpElib}, see Section~\ref{sec:obtain} for details.
-%
-\item Unpack in your preferred directory:
-\begin{lstlisting}
-~/Software/> tar -xzf dopelib-3.0.tar.gz
-\end{lstlisting}
-to get
-\begin{lstlisting}
-~/Software/dopelib-3.0
-\end{lstlisting}
-To build the \texttt{DOpElib} you have to change to the directory 
-\texttt{DOpEsrc} where you can call 
-\begin{lstlisting}
-~/Software/dopelib-3.0/DOpEsrc> make c-all
-\end{lstlisting}
-to build and configure the library using cmake. 
-You also get various installation options by just typing 
-\texttt{make}.
-\item To generate documentation, please go into:
-\begin{lstlisting}
-~/Software/dopelib-3.0/Examples
-\end{lstlisting}
-Here, you get all options by typing:
-\begin{lstlisting}
-make
-\end{lstlisting}
-Specifically, you can create and run all examples, running a test suite,
-and generating documentation:
-{\footnotesize
-\begin{lstlisting}
-===========================================================================
-=              Makefile for the DOpE documentation                        =
-===========================================================================
-=                                                                         =
-= The following targets exist:                                            =
-=    all       :  Make all examples (in optimized mode)                   =
-=    c-all     :  Make all examples using cmake                           =
-=    allignore :  Make all examples (in optimized mode), ignore errors.   =
-=    clean     :  Cleaning up all examples                                =
-=    tests     :  Run all test param data.                                =
-=    cat       :  Run clean, run all, run tests (combine these commands)  =
-=    c-cat     :  Run clean, run c-all, run tests (combine these commands)=
-=    catignore :  Same as cat, but ignores compiler errors.               =
-=    pdf-doc   :  Create documentation in pdf file format via latexmk     =
-=    doc       :  Create documentation in pdf file format via latexmk     =
-=    distclean :  Cleaning up, including documentation                    =
-=    warncheck :  Checks whether all Examples compile without warnings    =
-===========================================================================
-\end{lstlisting}
-}
-The new options are $c-all$ and $c-cat$ utilizing the cmake build system.
-For html documentation, change into:
-\begin{lstlisting}
-~/Software/dopelib-3.0/doxygen
+~/Software/dopelib-4.0/doxygen
 \end{lstlisting}
@@
 \begin{itemize}
-\item \textbf{unrecognized command line option "-std=gnu++0x"}\\
+\item \textbf{unrecognized command line option "-std=c++11"}\\
   This means that your compiler is too old. You can check the 
   version of your compiler using \texttt{g++ -v}. If the version is lower than
-  \texttt{4.5} you need to get a newer compiler version.
+  \texttt{4.8} you need to get a newer compiler version.
 \end{itemize}
@@
   So that you need to install a newer compiler. You can do so, for instance
-  using macports. Then you can find, e.g., g++ version 4.5 on OSX Lion 
-  in \texttt{/opt/local/bin/g++-mp-4.5}.
+  using macports. Then you can find, e.g., g++ version 4.8 on OSX Lion 
+  in \texttt{/opt/local/bin/g++-mp-4.8}.
 

### 2016-05-26 DOpEsrc/basic/mol_multimesh_spacetimehandler.h

@@ -24,13 +24,13 @@
 #ifndef MOL_MULTIMESH_SPACE_TIME_HANDLER_H_
 #define MOL_MULTIMESH_SPACE_TIME_HANDLER_H_
 
-#include "mol_spacetimehandler.h"
-#include "constraints.h"
-#include "sparsitymaker.h"
-#include "userdefineddofconstraints.h"
-#include "sth_internals.h"
-#include "refinementcontainer.h"
-#include "solutiontransfer_wrapper.h"
+#include <basic/mol_spacetimehandler.h>
+#include <basic/constraints.h>
+#include <include/sparsitymaker.h>
+#include <include/userdefineddofconstraints.h>
+#include <basic/sth_internals.h>
+#include <container/refinementcontainer.h>
+#include <wrapper/solutiontransfer_wrapper.h>
 
 #include <deal.II/dofs/dof_handler.h>
 #include <deal.II/dofs/dof_renumbering.h>

### 2016-05-26 DOpEsrc/container/higher_order_dwrc.h

@@ -24,7 +24,7 @@
 #ifndef HIGHER_ORDER_DWRC_H_
 #define HIGHER_ORDER_DWRC_H_
 
-#include "dwrdatacontainer.h"
+#include <container/dwrdatacontainer.h>
 #include <deal.II/fe/fe_tools.h>
 
 namespace DOpE

### 2016-05-26 DOpEsrc/container/refinementcontainer.h

@@ -25,7 +25,7 @@
 #define REFINEMENTCONTAINER_H_
 
 #include <deal.II/lac/vector.h>
-#include "dopetypes.h"
+#include <basic/dopetypes.h>
 #include <limits>
 
 namespace DOpE

### 2016-05-26 DOpEsrc/opt_algorithms/reduced_snopt_algorithm.h

@@ -24,11 +24,11 @@
 #ifndef REDUCED_SNOPT__ALGORITHM_H_
 #define REDUCED_SNOPT__ALGORITHM_H_
 
-#include "reducedalgorithm.h"
-#include "parameterreader.h"
-#include "dopetypes.h"
+#include <opt_algorithms/reducedalgorithm.h>
+#include <include/parameterreader.h>
+#include <basic/dopetypes.h>
 
-#include "snopt_wrapper.h"
+#include <wrapper/snopt_wrapper.h>
 
 #include <iostream>
 #include <assert.h>

### 2016-05-26 DOpEsrc/source/mapping_wrapper.cc

@@ -22,7 +22,7 @@
 **/
 
 
-#include"mapping_wrapper.h"
+#include<wrapper/mapping_wrapper.h>
 #include <deal.II/dofs/dof_handler.h>
 //#include <deal.II/multigrid/mg_dof_handler.h>
 #include <deal.II/hp/dof_handler.h>

### 2016-05-26 DOpEsrc/tsschemes/crank_nicolson_problem.h

@@ -24,8 +24,8 @@
 #ifndef CRANKNICOLSONProblem_H_
 #define CRANKNICOLSONProblem_H_
 
-#include "initialproblem.h" 
-#include "primal_ts_base.h"
+#include <problemdata/initialproblem.h> 
+#include <tsschemes/primal_ts_base.h>
 
 namespace DOpE
 {

### 2016-05-26 Examples/OPT/InstatPDE/Example2/main.cc

@@ -36,24 +36,24 @@
 #include <deal.II/grid/grid_generator.h>
 
 //DOpE includes
-#include "parameterreader.h"
-#include "directlinearsolver.h"
-#include "integrator.h"
-#include "mol_spacetimehandler.h"
-#include "simpledirichletdata.h"
-#include "integratordatacontainer.h"
-#include "newtonsolver.h"
-#include "functionalinterface.h"
-#include "noconstraints.h"
+#include <include/parameterreader.h>
+#include <templates/directlinearsolver.h>
+#include <templates/integrator.h>
+#include <basic/mol_spacetimehandler.h>
+#include <problemdata/simpledirichletdata.h>
+#include <container/integratordatacontainer.h>
+#include <templates/newtonsolver.h>
+#include <interfaces/functionalinterface.h>
+#include <problemdata/noconstraints.h>
 
 //DOpE includes for instationary problems
-#include "instatreducedproblem.h"
-#include "instat_step_newtonsolver.h"
-#include "reducednewtonalgorithm.h"
-#include "instatoptproblemcontainer.h"
+#include <reducedproblems/instatreducedproblem.h>
+#include <templates/instat_step_newtonsolver.h>
+#include <opt_algorithms/reducednewtonalgorithm.h>
+#include <container/instatoptproblemcontainer.h>
 
 //various timestepping schemes
-#include "backward_euler_problem.h"
+#include <tsschemes/backward_euler_problem.h>
 
 #include "localpde.h"
 #include "localfunctional.h"

### 2016-05-26 Examples/OPT/InstatPDE/Example3/main.cc

@@ -36,26 +36,26 @@
 #include <deal.II/grid/grid_generator.h>
 
 //DOpE includes
-#include "parameterreader.h"
-#include "directlinearsolver.h"
-#include "integrator.h"
-#include "mol_spacetimehandler.h"
-#include "simpledirichletdata.h"
-#include "integratordatacontainer.h"
-#include "functionalinterface.h"
-#include "noconstraints.h"
+#include <include/parameterreader.h>
+#include <templates/directlinearsolver.h>
+#include <templates/integrator.h>
+#include <basic/mol_spacetimehandler.h>
+#include <problemdata/simpledirichletdata.h>
+#include <container/integratordatacontainer.h>
+#include <interfaces/functionalinterface.h>
+#include <problemdata/noconstraints.h>
 
-#include "integratormixeddims.h" // for mixed dim opt. control
-#include "newtonsolvermixeddims.h" // for mixed dim opt. control
+#include <templates/integratormixeddims.h> // for mixed dim opt. control
+#include <templates/newtonsolvermixeddims.h> // for mixed dim opt. control
 
 //DOpE includes for instationary problems
-#include "instatreducedproblem.h"
-#include "instat_step_newtonsolver.h"
-#include "reducednewtonalgorithm.h"
-#include "instatoptproblemcontainer.h"
+#include <reducedproblems/instatreducedproblem.h>
+#include <templates/instat_step_newtonsolver.h>
+#include <opt_algorithms/reducednewtonalgorithm.h>
+#include <container/instatoptproblemcontainer.h>
 
 //various timestepping schemes
-#include "backward_euler_problem.h"
+#include <tsschemes/backward_euler_problem.h>
 
 #include "localpde.h"
 #include "localfunctional.h"

### 2016-05-26 Examples/OPT/StatPDE/Example6/main.cc

@@ -30,20 +30,20 @@
 #include <deal.II/base/quadrature_lib.h>
 #include <deal.II/lac/precondition_block.h>
 
-#include "reduced_ipopt_algorithm.h"
-#include "reduced_snopt_algorithm.h"
-#include "optproblemcontainer.h"
-#include "functionalinterface.h"
-#include "statreducedproblem.h" 
-#include "newtonsolver.h"
-#include "cglinearsolver.h"
-#include "integrator.h"
-#include "parameterreader.h"
-#include "mol_spacetimehandler.h"
-#include "simpledirichletdata.h"
-#include "noconstraints.h"
-#include "preconditioner_wrapper.h"
-#include "integratordatacontainer.h"
+#include <opt_algorithms/reduced_ipopt_algorithm.h>
+#include <opt_algorithms/reduced_snopt_algorithm.h>
+#include <container/optproblemcontainer.h>
+#include <interfaces/functionalinterface.h>
+#include <reducedproblems/statreducedproblem.h> 
+#include <templates/newtonsolver.h>
+#include <templates/cglinearsolver.h>
+#include <templates/integrator.h>
+#include <include/parameterreader.h>
+#include <basic/mol_spacetimehandler.h>
+#include <problemdata/simpledirichletdata.h>
+#include <problemdata/noconstraints.h>
+#include <wrapper/preconditioner_wrapper.h>
+#include <container/integratordatacontainer.h>
 
 #include "localpde.h"
 #include "localfunctional.h"

### 2016-05-26 Examples/PDE/StatPDE/Example10/main.cc

@@ -33,17 +33,17 @@
 #include <deal.II/numerics/error_estimator.h>
 #include <deal.II/grid/grid_refinement.h>
 
-#include "pdeproblemcontainer.h"
-#include "statpdeproblem.h"
-#include "newtonsolver.h"
-#include "trilinosdirectlinearsolver.h"
-#include "directlinearsolver.h"
-#include "integrator.h"
-#include "parameterreader.h"
-#include "mol_statespacetimehandler.h"
-#include "simpledirichletdata.h"
-#include "solutionextractor.h"
-#include "integratordatacontainer.h"
+#include <container/pdeproblemcontainer.h>
+#include <reducedproblems/statpdeproblem.h>
+#include <templates/newtonsolver.h>
+#include <templates/trilinosdirectlinearsolver.h>
+#include <templates/directlinearsolver.h>
+#include <templates/integrator.h>
+#include <include/parameterreader.h>
+#include <basic/mol_statespacetimehandler.h>
+#include <problemdata/simpledirichletdata.h>
+#include <include/solutionextractor.h>
+#include <container/integratordatacontainer.h>
 
 #include "localpde.h"
 #include "functionals.h"

### 2016-05-26 Examples/PDE/StatPDE/Example12/main.cc

@@ -21,16 +21,16 @@
  *
  **/
 
-#include "pdeproblemcontainer.h"
-#include "statpdeproblem.h"
-#include "newtonsolver.h"
-#include "integrator.h"
-#include "parameterreader.h"
-#include "mol_statespacetimehandler.h"
-#include "simpledirichletdata.h"
-#include "sparsitymaker.h"
-#include "integratordatacontainer.h"
-#include "preconditioner_wrapper.h"
+#include <container/pdeproblemcontainer.h>
+#include <reducedproblems/statpdeproblem.h>
+#include <templates/newtonsolver.h>
+#include <templates/integrator.h>
+#include <include/parameterreader.h>
+#include <basic/mol_statespacetimehandler.h>
+#include <problemdata/simpledirichletdata.h>
+#include <include/sparsitymaker.h>
+#include <container/integratordatacontainer.h>
+#include <wrapper/preconditioner_wrapper.h>
 
 #include <iostream>
 #include <fstream>

### 2016-11-17 Examples/Documentation/examples.tex

@@
 
-\item We now have to change the target in the \texttt{Makefile}, 
-  otherwise our new example will refer to the executable of the old one.
+\item You will find a file \texttt{CMakeLists.txt} in the folder 
+\texttt{MyWonderfulFirstExample}. Open this file, in it you can 
+find the line 
+\begin{verbatim}
+  SET(TARGET "DOpE-PDE-StatPDE-Example1")
+\end{verbatim}
+the string \texttt{DOpE-PDE-StatPDE-Example1} will be the name of the
+executable of your program. Change it to something suitable, e.g.,
+\begin{verbatim}
+  SET(TARGET "MyWonderfulFirstExample")
+\end{verbatim}
+Moreover, this file contains the lines 
+\begin{verbatim}
+SET(dope_dimension 2)
+SET(deal_dimension 2)
+\end{verbatim}
+which define the dimension of the domain for the control-variable
+(\texttt{dope\_dimension})
+and the PDE solution (\texttt{deal\_dimension}). If for your example 
+one of these differs from $2$ adjust the number accordingly.
+This file will not need any further modifications.
 
-  Thus we open the \texttt{Makefile} in the directory \texttt{MyWonderfulFirstExample}
-  to change the name of the executable. In our example you will find a line
+\item  Now, you are prepared to change any of the problem
+  dependent data in information in the files 
 \begin{verbatim}
-target   = $(BINDIR)/DOpE-PDE-StatPDE-Example1-$
-           (dope_dimension)d-$(deal_II_dimension)d
+main.cc, localpde.h, functionals.h, localfunctional.h, etc
 \end{verbatim} 
-Here we need to replace the target to a new name to avoid any 
-possible complication. Thus for example we change the line to 
+If above you have changed the dimensions, make sure to adjust all
+files accordingly!
+
+\item The cmake system can - in principle be used
+  ''in-source'' to create the executable. 
+  However, this may break some of the automated tests (later),
+  so you are encouraged to proceed differently. To do this end
+  proceed as follows in the directory \texttt{MyWonderfulFirstExample}:
   \begin{verbatim}
-target   = $(BINDIR)/MyWonderfulFirstExample
+MyWonderfulFirstExample$ mkdir build
+MyWonderfulFirstExample$ cd build
+MyWonderfulFirstExample/build$ cmake -DCMAKE_BUILD_TYPE=Release ../
 \end{verbatim}
-This will build the executable file in the standard binary directory 
-of \texttt{DOpElib}, i.e., \texttt{\$(DOpE)/bin}, under the name 
-\texttt{MyWonderfulFirstExample}.
+which will configure the build (if you want to debug it is useful to
+replace the string \texttt{Release} with \texttt{Debug}). 
+Once this is done, you can compile and run the code:
+\begin{verbatim}
+MyWonderfulFirstExample/build$ make 
+MyWonderfulFirstExample/build$ cd ..
+MyWonderfulFirstExample$ ./MyWonderfulFirstExample
+\end{verbatim}
+(Assuming that no errors occurred during the make call)
 
-\textit{Note that you should use a different name for all your 
-examples as otherwise you may experience a lot of strange behavior, as
-the executable will be overwritten by all your examples!}
+\item The \texttt{Makefile} in the directory is present only to
+  preserve backward compatibility. If you wish to use the automated
+  build/test routines in DOpE lib, you need to make sure that it its
+  configured correctly. If you just want to use the cmake capabilities 
+  you can safely ignore this passage.
+
+  Furthermore, if you have followed the instructions above, then no
+  changes will be needed.
+
+  If you have moved the directory to some other place, i.e., it is not
+  in the same folder as Example1, then open the \texttt{Makefile} 
+  in the directory \texttt{MyWonderfulFirstExample}
+  you will find the line 
+  \begin{verbatim}
+  DOpE = ../../../../
+  \end{verbatim}
+  which points to the root directory of your DOpE installation.
+  Adjust the path to match accordingly.
 
@@
 \begin{verbatim}
-PROGRAM=../../../../../bin/DOpE-PDE-StatPDE-Example1-2d-2d
+PROGRAM=../DOpE-PDE-StatPDE-Example1-2d-2d
 \end{verbatim} 
@@
 \begin{verbatim}
-PROGRAM=../../../../../bin/MyWonderfulFirstExample
-\end{verbatim}
-
-\item  Now, you are prepared to change any of the problem
-  dependent data in information in the files 
-\begin{verbatim}
-main.cc, localpde.h, functionals.h, localfunctional.h, etc
+PROGRAM=../MyWonderfulFirstExample
 \end{verbatim}

### 2016-11-17 Examples/Documentation/introduction.tex

@@
 \section{License information}
-Copyright (C) 2012--2014 by the DOpElib authors\\[2mm]
+Copyright (C) 2012--2016 by the DOpElib authors\\[2mm]
 %
@@
 \begin{itemize}
-  \item Christian Goll (Heidelberg University)
-  \item Thomas Wick (Johann Radon Institute for Computational and Applied Mathematics (RICAM))
+  \item Thomas Wick (Centre de Mathématiques Appliquées (CMAP) École Polytechnique)
   \item Winnifried Wollner (Technische Universität Darmstadt)
@@
 
+for the use of MPI use
+
+\begin{lstlisting}
+-DDEAL_II_WITH_MPI=true
+\end{lstlisting}
+
+and for the use of distributed meshes over several computers the p4est
+library is needed in deal.II to use this use the flags (in addition to the use of MPI)
+\begin{lstlisting}
+-DDEAL_II_WITH_P4EST=true -DP4EST_DIR=[path to p4est]
+\end{lstlisting}
+
 \subsection{ThirdPartyLibraries}
@@
 \texttt{ThirdPartyLibs/SNOPT.INSTALLNOTES}. In particular you need to configure 
-\texttt{deal.II} with at least the following options:\\
-\texttt{./configure --with-umfpack --with-blas --with-lapack} (up to 7.3.0)
-or enabling UMFPACK in the ccmake options from version 8.0.0 on.
+\texttt{deal.II} with at least with enabled UMFPACK.
 
@@
 \section{Installation}\label{sec_installation}
-\subsection{After Version 2.0 - CMake build system}
 To work with DOpElib, your need necessarily deal.II, which installation 

### 2016-11-17 Examples/description.tex
Added Nonstationary Example (isothermal Euler) with dG discretization in time + Bugfixes
Further work on optimization with functionals not given by a single integration. Calculation of such functionals is now working, derivative calculation still missing
Corrected umlaute in documentation

@@
 \documentclass[a4paper,cleardoubleempty]{scrreprt}
+\usepackage[utf8]{inputenc}
 \usepackage{amsmath,amssymb, amsthm, commath}
@@
 \author{
-  Thomas Wick\thanks{Johann Radon Institute for Computational and Applied Mathematics (RICAM), 
-    {thomas.wick@ricam.oeaw.ac.at}}, 
+  Thomas Wick\thanks{École Polytechnique, 
+    {thomas.wick@polytechnique.edu}}, 
   Winnifried Wollner\thanks{Technische Universität Darmstadt, 
@@
 \cleardoublepage
+\subsection{Isothermal Euler equations}
+\label{PDE_isothermal_euler}
+\input{PDE/InstatPDE/Example7/content.tex}
+\cleardoublepage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Parameter control with a nonlinear PDE from fluid dynamics - with cost functional
+not given as an integral but as function of an integral}
+\label{OPT_Stat_Param_Nonlin_Fluid_extension}
+\input{OPT/StatPDE/Example10/content.tex}
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Subject to a Nonstationary PDE}
@@
 \clearpage
+\subsection{Minimizing the spatial mean-value - cost functional involves functions of integrals}
+\label{OPT_Instat initial-value extension}
+\input{OPT/InstatPDE/Example4/content.tex}
+\clearpage
 

### 2017-04-12 Examples/Documentation/introduction.tex
Description modified w.r.t. to make c-all

@@
   in \texttt{/opt/local/bin/g++-mp-4.8}.
+
+\subsection*{3.) Problem using make c-all in the Examples folder :}
+
+\begin{lstlisting}
+wick@linux:~/Software/dopelib/Examples> make c-all
+cd OPT/StatPDE/Example1; make -s c-all
+make[2]: *** No targets specified and no makefile found.  Stop.
+../../../..//Examples/Make.global_options:2: recipe for target 'c-all' failed
+make[1]: *** [c-all] Error 2
+Makefile:24: recipe for target 'OPT/StatPDE/Example1' failed
+make: *** [OPT/StatPDE/Example1] Error 2
+\end{lstlisting}
+
+This means that the Makefile in the specific examples folder has been 
+modified in some illegal way. Remove this Makefile and get the original 
+one from DOpElib. Then try again make c-all (make always sure that before 
+typing make c-all that the deal.II path is exported - see above).
 

### 2017-04-12 Examples/Documentation/introduction.tex
optproblemcontainer modified to fit with deal.II 8.5.0

@@
 The options are $c-all$ and $c-cat$ utilizing the cmake build system.
+
+It is \textbf{important} to generate the examples via make c-all 
+in the home folder of the examples and \textbf{not} as in deal.II by 
+using cmake . in the specific example itself. With this you
+would currently destroy the delivered Makefile. See also the FAQ below.
+
 For html documentation, change into:
@@
 This means that the Makefile in the specific examples folder has been 
-modified in some illegal way. Remove this Makefile and get the original 
-one from DOpElib. Then try again make c-all (make always sure that before 
+modified in some illegal way. Remove this Makefile 
+\begin{lstlisting}
+rm Makefile
+rm -r CMakeFiles
+rm CMakeCache.txt
+rm -r autobuild
+\end{lstlisting}
+and get the original 
+one from DOpElib:
+\begin{lstlisting}
+git checkout Makefile
+\end{lstlisting}
+Then try again make c-all (make always sure that before 
 typing make c-all that the deal.II path is exported - see above).

### 2017-04-12 Examples/Documentation/introduction.tex
removed more svn stuff from the documentation, explained to use candi for thirdparty deal.II libries

@@
 \begin{enumerate}[A)]
-\item You can obtain a copy of \dope{} from the developers svn repository using\\
-\texttt{svn co --username=[your username] \textbackslash \\ 
-\hspace*{5mm}https://svn.dopelib.net/dope}\\
-(The backslash in the first line means that the line will be continued as one!)
+\item You can obtain a copy of \dope{} from the developers git repository using\\
+\texttt{git clone git://git.mathematik.tu-darmstadt.de/dopelib} \\
+in the command line of your terminal.
 
-Note that you need a valid username and password. If you have none, please contact the 
-maintainers by sending an EMail to \url{dope@dopelib.net}.
+%Note that you need a valid username and password. If you have none, please contact the 
+%maintainers by sending an EMail to \url{dope@dopelib.net}.
 %
@@
 \end{itemize}
+%%% TODO: add Daniel Jodlbauer and Bernhard Endtmayer
 
@@
 
+\begin{remark}
+Our current DOpE installation has been successfully tested for deal.II version 8.5.0.
+\end{remark}
+
 A configuration of deal.II working with \texttt{DOpElib} can be obtained with
@@
 mkdir build
-mkdir install
 cd build 
-cmake -DCMAKE_INSTALL_PREFIX=../install \ 
-      -DDEAL_II_WITH_UMFPACK=true \
-      -DDEAL_II_FORCE_BUNDLED_UMFPACK=true  ../
+cmake -DCMAKE_INSTALL_PREFIX=/path/to/install/dir ../deal.II
 make install
 \end{lstlisting}
-If in addition, features are desired corresponding flags need to be passed to cmake. If, for 
-example Trilinos solvers are desired, the above cmake command needs to have the additional flags
-
+If in addition, features are desired corresponding flags need to be passed to
+cmake. An excellent description is provided on the deal.II webpage
 \begin{lstlisting}
--DDEAL_II_WITH_TRILINOS=true -DTRILINOS_DIR=[path to trilinos]
+http://www.dealii.org/8.5.0/readme.html
 \end{lstlisting}
 
-for the use of MPI use
-
+\subsection{deal.II ThirdPartyLibraties}
+To install third party libraries used for deal.II we recommend to use
+the candi script:
 \begin{lstlisting}
--DDEAL_II_WITH_MPI=true
+https://github.com/koecher/candi
 \end{lstlisting}
-
-and for the use of distributed meshes over several computers the p4est
-library is needed in deal.II to use this use the flags (in addition to the use of MPI)
-\begin{lstlisting}
--DDEAL_II_WITH_P4EST=true -DP4EST_DIR=[path to p4est]
-\end{lstlisting}
+On this webpage, you will find also enough information how to use 
+the candi script.
 

### 2017-04-12 Examples/Documentation/examples.tex

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Changing from Release mode to Debug mode}
+In a specific example go into the autobuild folder 
+and change therein:
+\begin{lstlisting}
+CMAKE_BUILD_TYPE:STRING=Release
+\end{lstlisting}
+to 
+\begin{lstlisting}
+Debug Release RelWithDebInfo MinSizeRel
+\end{lstlisting}
+Then go one folder back and type as usual `make' to build the executable file.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Creating new examples}
@@
 \begin{enumerate}
- \item In a first step, we copy \texttt{Example1} and renamed it, e.g., 
+\item \textbf{New: in git from March 2017} In a first step, we copy \texttt{Example1} and renamed it, e.g., 
+\texttt{MyWonderfulFirstExample}. 
+After having reached the folder of the example in question in the terminal, 
+\texttt{PDE/StatPDE} in our case, we perform these operation writing the following:
+\begin{verbatim}
+cp -r Example1 MyWonderfulFirstExample
+cd MyWonderfulFirstExample
+\end{verbatim}
+
+ \item \textbf{Old: in the svn up to version 8.3} In a first step, we copy \texttt{Example1} and renamed it, e.g., 
 \texttt{MyWonderfulFirstExample}. At the same time it is important to remove the repository 

### 2017-04-12 Examples/Documentation/introduction.tex
added full deal.ii config line to docu

@@
 http://www.dealii.org/8.5.0/readme.html
+\end{lstlisting}
+Currently, for all features of \texttt{DOpElib} the following settings
+are useful to be passed to cmake
+\begin{lstlisting}
+  cmake -DCMAKE_INSTALL_PREFIX=[path to install dir] \ 
+      -DDEAL_II_WITH_UMFPACK=true \
+      -DDEAL_II_FORCE_BUNDLED_UMFPACK=true \
+      -DDEAL_II_WITH_TRILINOS=true \
+      -DTRILINOS_DIR=[path to trilinos] \
+      -DDEAL_II_WITH_MPI=true \
+      -DDEAL_II_WITH_P4EST=true \
+      -DP4EST_DIR=[path to p4est] \
+      [Path to deal.II sources]
 \end{lstlisting}

### 2017-04-12 Examples/Documentation/examples.tex
Section 4.2 in the docu: How to run existing examples

@@
 %\end{remark}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{How to run existing examples}
+The easiest way is to first build all examples. Go into 
+\begin{lstlisting}
+dopelib/Examples/
+\end{lstlisting}
+Herein, type 
+\begin{lstlisting}
+make c-all
+\end{lstlisting}
+By typing only `make' you will see all options.
+This procedure will take some minutes. Afterwards,
+go into the examples folder of your choice. For instance:
+\begin{lstlisting}
+dopelib/Examples/PDE/InstatPDE/Example1
+\end{lstlisting}
+Herein you create the executable via
+\begin{lstlisting}
+make 
+\end{lstlisting}
+and then 
+\begin{lstlisting}
+./DOpE-PDE-InstatPDE-Example1 
+\end{lstlisting}
+Or both commands together:
+\begin{lstlisting}
+make && ./DOpE-PDE-InstatPDE-Example1 
+\end{lstlisting}
+You will find the results on the screen in the terminal as well
+as some graphical output in the `Results' folder.
 

### 2017-04-12 Examples/Documentation/introduction.tex

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
 \section{Installation}\label{sec_installation}
@@
 \end{lstlisting}
+\begin{remark}
+\textbf{Be careful} if you work in several terminal windows and have several deal.II versions. In this case, the
+export-command might have to be done again in that new window. Otherwise
+another deal.II version might be linked resulting in linker errors or
+segmentation faults without useful error message. In order to check to which
+deal.II version DOpE is linked, type 
+\begin{lstlisting}
+echo $DEAL_II_DIR
+\end{lstlisting}
+in the current terminal window.
+\end{remark}
 Detailed installing instructions for deal.II (here the last version

### 2017-04-13 Examples/Documentation/examples.tex
Updated docu w.r.t. to linker errors while using the wrong deal.II library
Updated documentation at several places

@@
 \section{How to run existing examples}
+\subsection{The global way}
 The easiest way is to first build all examples. Go into 
@@
 \end{lstlisting}
-By typing only `make' you will see all options.
+By typing only `make' you will see all options 
+(we also refer to Section \ref{sec_installation}).
 This procedure will take some minutes. Afterwards,
@@
 \begin{lstlisting}
-dopelib/Examples/PDE/InstatPDE/Example1
+dopelib/Examples/PDE/InstatPDE/Example1/autobuild
 \end{lstlisting}
@@
 \end{lstlisting}
-and then 
+and then go back via ../ and 
 \begin{lstlisting}
@@
 \begin{lstlisting}
-make && ./DOpE-PDE-InstatPDE-Example1 
+autobuild> make && cd .. &&  ./DOpE-PDE-InstatPDE-Example1 
 \end{lstlisting}
@@
 
+\subsection{Building, making and running in the local folder}
+In case we have run first globally as previsously described, 
+in each examples folder we find an autobuild subfolder. 
+If we now want to work locally (which is the usual way) then 
+we have to type `make' in the autobuild folder:
+\begin{lstlisting}
+autobuild> make
+\end{lstlisting}
+In case we need to modify the Makefile, we need first to do:
+\begin{lstlisting}
+autobuild> cmake ..
+\end{lstlisting}
+For instance one reason to modify the Makefile could be 
+to change to debug mode as described in Section \ref{sec_release_debug}.
+Finally, go one folder back and run the object file.
+
+In case you work and test out new things (modifiying the 
+main.cc and localpde.h files etc.), one command for instance 
+in Example 9 is:
+\begin{lstlisting}
+Example9> cd autobuild/ && make && cd .. &&  ./DOpE-PDE-StatPDE-Example9
+\end{lstlisting}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Changing from Release mode to Debug mode}
+\label{sec_release_debug}
 In a specific example go into the autobuild folder 
@@
 \begin{lstlisting}
-CMAKE_BUILD_TYPE:STRING=Release
+cmake CMAKE_BUILD_TYPE=Release ..
 \end{lstlisting}
-to 
+to some other behavior:
 \begin{lstlisting}
@@
 \end{lstlisting}
-Then go one folder back and type as usual `make' to build the executable file.
+Then type `make' to build the executable file and go back 
+into the parent folder and execute the object file.
 

### 2017-04-13 Examples/PDE/StatPDE/Example9/content.tex
Updated documentation at several places

@@
 In this example we consider a simple stationary FSI problem. The fluid
-is given as an incompressible Newtonian fluid modeled by the Stokes equation.
-Here, we use the symmetric stress tensor which has a little consequence when using the do-nothing outflow condition, see also section \ref{PDE_Stat_Stokes}.
-The computation domain is $\Omega = [-6,6]\times [0,2]$ and we choose for simplicity $f=0$.
+is given as an incompressible Newtonian fluid modeled by the Stokes
+equation. Here, we use the symmetric stress tensor which has a little
+consequence when using the do-nothing outflow condition, see also
+section \ref{PDE_Stat_Stokes}. The flow is driven by non-homogeneous Dirichlet 
+condition on the left bounddary.
+
+The computation domain is $\Omega = [-6,6]\times [0,2]$ and we choose for
+simplicity $f=0$. 
+We add in the subdomain $\Omega_s = [0,2]\times [0,1]$ a solid obstacle. In
+this 
+solid we prescribe an incompressible neo-Hookean material law.
 
@@
 \end{Problem}
+\begin{remark}
+At our developer meeting on Apr 13, 2017, we also added 
+a simplified STVK material for testing purposes. This STVK 
+material is implemented in a different header file (see also 
+further comments below).
+\end{remark}
 
@@
 
-%%% Unsinn! Das example hat nicht viel mit dem davor zu tun!
-%There is not much difference to the prior example, so we will keep our remarks rather short.
-The \textit{functionals.h} and \textit{myfunctions.cc} files did not change at all, so we simply refer to the corresponding sections in the last example.\\
 In the \textit{localpde.h} file, all functions of the \texttt{LocalPDE} class have to be adjusted to the current FSI problem.
@@
 Apart from that, we define objects for the same classes as before that are even named equally and use the same solvers.\\
-Again, the solution is reached within one Newton step, and all we see from the program output is the values of the functionals.
+Again, the solution is reached within one Newton step, and all we see from the
+program output is the values of the functionals.
+
+We also demonstrate another very convenient feature. Since we have 
+two different material laws (INH and simplified STVK), we do not implement 
+them together in one \textit{localpde.h} file by using if-conditions etc. 
+But we implement them seperately in two different *.h files, namely
+\begin{lstlisting}
+localpde.h
+localpde_stvk_material.h
+\end{lstlisting}
+In the \texttt{main.cc} function we can now simply comment 
+or uncomment the respective file we want to work with.
+This allows us to keep a clean file for a running example and experiment 
+in other files (possibly more than $2$) and by just changing 
+two lines in the main.cc in order to change the equations.

### 2017-05-15 Examples/Documentation/introduction.tex
Updated description with Daniel and Bernhard
Fehlerbeschreibung compilieren von Bernhard eingecheckt

@@
 \begin{itemize}
+  \item Daniel Jodlbauer (current work on parallelization)
+  \item Bernhard Endtmayer (currently working on phase-field fracture)
   \item Michael Geiger (Examples for Plasticity, and Documentation of several PDE-Examples)
@@
 typing make c-all that the deal.II path is exported - see above).
+\newline
+\subsection*{4.) Different g++ compilers are used for compiling DOpElib and the
+Examples folder}
+
+\begin{lstlisting}
+bernhard@ebook:~/Schreibtisch/Prog/DOpe/Dope/dopelib/Examples> make c-all
+cd OPT/StatPDE/Example1; make -s c-all
+[ 50%] Building CXX object CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/main.cc.o
+c++: error: unrecognized command line option '-fopenmp-simd'
+c++: error: unrecognized command line option '-std=c++14'
+CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/build.make:62: recipe for target
+'CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/main.cc.o' failed
+make[4]: *** [CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/main.cc.o] Error 1
+CMakeFiles/Makefile2:131: recipe for target
+'CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/all' failed
+make[3]: *** [CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/all] Error 2
+Makefile:83: recipe for target 'all' failed
+make[2]: *** [all] Error 2
+../../../..//Examples/Make.global_options:2: recipe for target 'c-all' failed
+make[1]: *** [c-all] Error 2
+Makefile:24: recipe for target 'OPT/StatPDE/Example1' failed
+make: *** [OPT/StatPDE/Example1] Error 2
+\end{lstlisting}
+The reason for this problem might be that the variables c++ and g++ as well as
+cc and gcc do not match.
+You can remove the old Examples folder and replace it by a new one and set the
+compilers manually by
+\begin{lstlisting}
+export CXX=path/to/g++-compiler
+export CC=path/to/gcc-compiler
+\end{lstlisting}
+where path/to/compiler should be the path to compiler DOpElib was compiled with.
+You can check the current version of your gcc- and g++-compiler with
+\begin{lstlisting}
+gcc --version
+g++ --version
+\end{lstlisting}
+Alternatively you can set the variable c++ to your g++ and the variable cc to
+gcc using update-alternatives
+as for example
+\begin{lstlisting}
+sudo update-alternatives --install /usr/bin/c++ c++ /usr/bin/g++-6 60
+sudo update-alternatives --install /usr/bin/cc cc /usr/bin/gcc-6 60
+\end{lstlisting}
+Thanks to Florian Tischler (Admin at RICAM Linz) for the help.
 

### 2017-05-17 Examples/Documentation/introduction.tex

@@
 typing make c-all that the deal.II path is exported - see above).
-\newline
-\subsection*{4.) Different g++ compilers are used for compiling DOpElib and the
-Examples folder}
 
+\subsection*{4.) Different g++ compilers are used for compiling DOpElib and the Examples folder}
 \begin{lstlisting}
@@
 c++: error: unrecognized command line option '-std=c++14'
-CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/build.make:62: recipe for target
-'CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/main.cc.o' failed
+CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/build.make:62: recipe for target 'CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/main.cc.o' failed
 make[4]: *** [CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/main.cc.o] Error 1
-CMakeFiles/Makefile2:131: recipe for target
-'CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/all' failed
+CMakeFiles/Makefile2:131: recipe for target 'CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/all' failed
 make[3]: *** [CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/all] Error 2
@@
 \end{lstlisting}
-The reason for this problem might be that the variables c++ and g++ as well as
-cc and gcc do not match.
-You can remove the old Examples folder and replace it by a new one and set the
-compilers manually by
+
+The reason for this problem might be that the variables c++ and g++ as well as cc and gcc do not match.
+\newline
+You can remove the old Examples folder and replace it by a new one and set the compilers manually by
 \begin{lstlisting}
@@
 \end{lstlisting}
-Alternatively you can set the variable c++ to your g++ and the variable cc to
-gcc using update-alternatives
+Alternatively you can set the variable c++ to your g++ and the variable cc to gcc using update-alternatives
 as for example

### 2017-05-28 Examples/PDE/InstatPDE/Example8/content.tex
Example 8: phase-field - Endtmayer / Wick

\subsubsection{General problem description}
We consider phase-field fracture propagation on a slit domain, with timedependet 
Dirichlet data and homogenous Neumann data.
The main novelties are:
\begin{itemize}
\item Imposing a variational inequality constraint ($\frac{\partial \varphi}{\partial t} \ge 0$)
via penalization;
%\item Implementation of an augmented Lagrangian loop (not yet)
\item Stress splitting into tension and compression \`a
la \cite{MieWelHof10a}, see as well \cite{AmorMarigoMaurini2009}, located 
in 
\begin{verbatim}
stress_splitting.h
\end{verbatim}
\item Implementation of a Newton solver that 
allows for an increase of the residual for which, however,
no convergence results is available:
\begin{verbatim}
DOpEsrc/templates/instat_step_modified_newtonsolver.h
\end{verbatim}
\item Two localpde* files:
\begin{verbatim}
localpde_quasi_monolithic.h
localpde_fully_implicit.h
\end{verbatim}
that provide implementations of a quasi-monolithic scheme (similar, but not
the same as \cite{HeWheWi15}) and a fully implicit implementation.
The fully implicit implementation requires a modified Newton scheme that 
allows for a temporary increase of the Newton residual \cite{Wi17_pff_error_oriented_Newton}.
\end{itemize}
Let $u$ be the displacement and $\varphi$ be the phase field variable which should highlight the crack.
Furthermore we denote the $L^2(\Omega)$ inner product of $v_1$ and $v_2$ with $(v_1,v_2)$.
A material is supposed to be undamaged at position $x$ if $\varphi(x)$ is close to 1 and comletely cracked
if it is close to $0$.
We want to find a stationary point of the energy functional
\begin{align*}
E(u, \varphi):=& \frac{1}{2}(\sigma(u,\varphi),\varepsilon(u)) \\
+&\frac{G_c}{2}(\frac{1}{\varepsilon} \Vert 1 - \varphi \Vert^2)+\varepsilon \Vert \nabla \varphi  \Vert\\
+&\underbrace{\frac{\gamma_{pen}}{2} ((\varphi(t_n)-\varphi(t_{n-1}))_+)^2}_{\text{penalization term}},
\end{align*}
with
\begin{align*}
\sigma(u,\varphi)&:= (1-\kappa)\varphi^2 \sigma(u)^+ + \sigma(u)^-,\\
\sigma(u)^+ &:= 2*\mu \varepsilon(u)^+ +\lambda (trace(\varepsilon(u)))_+ I,\\
\sigma(u)^- &:= 2*\mu \varepsilon(u)^- +\lambda (trace(\varepsilon(u)))_- I,\\
(c)_+ &:= max(0, c), \\
(c)_- &:= c-(c)_+, \\
\varepsilon(u)^+ &:= \sum_{i=1}^{d} (\lambda_i)_+ n_i^T n_i,\\
\varepsilon(u)^- &:= \sum_{i=1}^{d} (\lambda_i)_- n_i^T n_i,\\
\varepsilon(u)   &:= \frac{1}{2}(\nabla u +(\nabla u)^T),
\end{align*}
where $c \in \mathbb{R}$, $t^n$ denotes the time at timestep $n$, $\lambda_i$ denotes the $i-th$ eigenvalue of $\varepsilon(u)$ and $n_i$ the corresponding eigenvector.
The idea for the choice of the energy functional is presented
in \cite{MieHofWel2010} (based on the original work \cite{BourFraMar00,BourFraMar08,FraMar98}).
This leads to the following problem: Find $(u, \varphi) \in V_u \times V_\varphi$ such that

\begin{align*}
A(u,\varphi)(\phi_u,\phi_\varphi):=& (\sigma(u,\varphi),\varepsilon(\phi_u)) \\
+& (((1-\kappa)\varphi \sigma(u)^+,\varepsilon(u)),\phi_\varphi)\\
 +& G_c(\frac{1}{\varepsilon} (1 - \varphi,\phi_\varphi)+\varepsilon (\nabla \varphi, \nabla \phi_\varphi)\\
+&\gamma_{pen} ((\varphi(t_n)-\varphi(t_{n-1}))_+,\phi_\varphi)=0 \quad \forall (\phi_u, \phi_\varphi) \in V_u \times V_\varphi.
\end{align*}

The test case is the so-called single edged notched 
shear test. For further details about this problem it is referred
to \cite{MieHofWel2010,BoVeScoHuLa12}, and \cite{WeWiWo2014}.

\subsubsection{Future improvements}
Possible future improvements are:
\begin{itemize}
\item Implement the phase-field extrapolation in the displacement equation proposed in \cite{HeWheWi15};
\item Implement the augmented Lagrangian loop proposed by
Wheeler/Wick/Wollner, CMAME 2014 \cite{WeWiWo2014}. 
\end{itemize}

% Need to implement extrapolation with two 
% previous time steps
% To be changed: 
% instatpdeproblem.h
% instat_step_newton_solver.h
%
% augmented Lagrangian solver loop
% -> currently simple penalization
% has to be implemented in instat_step_newton_solver.h
%
% Dokumentation in content.tex

### 2017-05-28 Examples/Documentation/introduction.tex
Docu modifications related to the new Example 8

@@
   \item Daniel Jodlbauer (current work on parallelization)
-  \item Bernhard Endtmayer (currently working on phase-field fracture)
+  \item Bernhard Endtmayer (phase-field fracture, PDE Instat Example 8 - in cooperation with T. Wick)
   \item Michael Geiger (Examples for Plasticity, and Documentation of several PDE-Examples)
@@
 \end{lstlisting}
-where path/to/compiler should be the path to compiler DOpElib was compiled with.
+where path/to/compiler should be the path to the compiler, DOpElib was compiled with.
 You can check the current version of your gcc- and g++-compiler with 

### 2017-07-12 Examples/Documentation/introduction.tex
FAQ for MPI errors in SUSE versions

@@
 \subsection*{4.) Different g++ compilers are used for compiling DOpElib and the Examples folder}
+{\footnotesize
 \begin{lstlisting}
@@
 c++: error: unrecognized command line option '-std=c++14'
-CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/build.make:62: recipe for target 'CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/main.cc.o' failed
+CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/build.make:62: 
+    recipe for target 'CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/main.cc.o' failed
 make[4]: *** [CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/main.cc.o] Error 1
-CMakeFiles/Makefile2:131: recipe for target 'CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/all' failed
+CMakeFiles/Makefile2:131: recipe for target 
+    'CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/all' failed
 make[3]: *** [CMakeFiles/DOpE-OPT-StatPDE-Example1.dir/all] Error 2
@@
 \end{lstlisting}
+}
 
@@
 Thanks to Florian Tischler (Admin at RICAM Linz) for the help.
+
+\subsection*{5.) Different g++, cc, cxx compilers are used for compiling
+  DOpElib and the Examples folder}
+
+The error message is as before:
+{\footnotesize
+\begin{lstlisting}
+Linking CXX executable ../DOpE-PDE-StatPDE-Example1
+/usr/lib64/gcc/x86_64-suse-linux/4.8/../../../../x86_64-suse-linux/bin/ld: cannot find -lmpi_usempi
+/usr/lib64/gcc/x86_64-suse-linux/4.8/../../../../x86_64-suse-linux/bin/ld: cannot find -lmpi_mpifh
+/usr/lib64/gcc/x86_64-suse-linux/4.8/../../../../x86_64-suse-linux/bin/ld: cannot find -lmpi
+collect2: error: ld returned 1 exit status
+\end{lstlisting}
+}
+
+An alternative to the previous solution is:
+\begin{lstlisting}
+export DEAL_II_DIR =your_mpi_deal_II_version // only necessary 
+      // when not in the standard path
+export PATH=/usr/lib64/mpi/gcc/openmpi/bin:${PATH}  // obtain your MPI path
+      // with `whereis mpicxx' etc.
+export CC=mpicc; export CXX=mpicxx; export FC=mpif90; export FF=mpif77
+\end{lstlisting}
+Then compile first the DOpElib library and then the Examples folder with these 
+settings. Possibly pay attention that also the deal.II version has been
+compiled 
+with this MPI path.
+
+\begin{remark}
+Be careful that you work in the same shell when only using `export'. 
+Or include the export lines into the 
+\begin{lstlisting}
+.bash_rc 
+\end{lstlisting}
+or 
+\begin{lstlisting}
+.profile
+\end{lstlisting}
+file to assert that the correct
+compiler is always used.
+\end{remark}
 

### 2017-08-07 Examples/Documentation/introduction.tex
Updated Documentation

--- 2017-05-28 Examples/Documentation/introduction.tex
@@
 \section{License information}
-Copyright (C) 2012--2016 by the DOpElib authors\\[2mm]
+Copyright (C) 2012--2017 by the DOpElib authors\\[2mm]
 %
@@
 If you like the library and use it for your own projects, please give credits by 
-referencing the project \dope{}~\cite{dope} using the following bibtex entry:
+referencing the project \dope{}~\cite{dope} using the following bibtex entries:
 
 \begin{lstlisting}
+@Article{dopelib,
+  Title   = {{DOpElib}: {D}ifferential Equations and 
+    {O}ptimization {E}nvironment; {A} Goal Oriented 
+    Software Library for Solving {PDE}s and Optimization 
+    Problems with {PDE}s},
+  Author  = {C. Goll and T. Wick and W. Wollner},
+  Journal = {Archive of Numerical Software},
+  Year    = {2017},
+  Number  = {2},
+  Pages   = {1--14},
+  Volume  = {5},
+  Doi     = {10.11588/ans.2017.2.11815}
+}
+
 @MISC{dope,
@@
 version of \texttt{g++} is too old, so you need to install the real thing. To do so, 
-download \texttt{MacPorts} from \texttt{macports.org}. Once you have installed MacPorts 
+download, e.g., \texttt{MacPorts} from \texttt{macports.org}. Once you have installed MacPorts 
 you can use it to install additional Linux software. 
@@
 \texttt{ThirdPartyLibs/SNOPT.INSTALLNOTES}. In particular you need to configure 
-\texttt{deal.II} with at least with enabled UMFPACK.
+\texttt{deal.II} at least with enabled UMFPACK.
 
@@
 
-The installation is straighforward and has been tested on openSUSE 12.1
-machines as well as on MAC. At the end of the installtion do not forget to
-add ipopt to your \texttt{LD\underline{ }LIBRARY\underline{ }PATH}:
+The installation is straightforward and has been tested on OpenSUSE
+12.1, Ubuntu 14.04 LTS and 16.04 LTS
+machines as well as on MAC. At the end of the installation do not forget to
+add \texttt{IPOPT} to your \texttt{LD\underline{ }LIBRARY\underline{ }PATH}:
 \begin{lstlisting}
@@
 \begin{itemize}
-\item Install \texttt{deal.II} to your home directory, i.e., it should be located in \\
-\texttt{\textasciitilde/deal.II}. \\[1mm]
-If you prefer another position you can install it anywhere but need 
+\item We automatically search for \texttt{deal.II} to your home
+  directory, i.e., if \texttt{deal.II} is located in \\
+\texttt{\textasciitilde/deal.II} you don't have to do anything. 
+\item Alternatively, if you prefer another position you can install it anywhere but need 
 to set the \texttt{DEAL\_II\_DIR} environment variable, so that it 
@@
 \begin{lstlisting}
-echo $DEAL_II_DIR
+echo ${DEAL_II_DIR}
 \end{lstlisting}
@@
 %
-\item Unpack in your preferred directory:
+\item If you obtained a tar-ball, unpack in your preferred directory:
 \begin{lstlisting}
-~/Software/> tar -xzf dopelib-4.0.tar.gz
+ tar -xzf dopelib-4.0.tar.gz
 \end{lstlisting}
-to get
+to get a directory 
 \begin{lstlisting}
-~/Software/dopelib-4.0
+dopelib-4.0/
 \end{lstlisting}
-To build the \texttt{DOpElib} you have to change to the directory 
-\texttt{DOpEsrc} where you can call 
+containing the source files. If you cloned the git repository you
+already have a directory; \texttt{dopelib} by default, containing the sources.
+\item Change to the directory with the sources, and then to the
+  subdirectory \texttt{DOpEsrc}.
+Once there, to build \texttt{DOpElib} you can call 
 \begin{lstlisting}
-~/Software/dopelib-4.0/DOpEsrc> make c-all
+ make c-all
 \end{lstlisting}
 to build and configure the library using cmake. 
-You also get various installation options by just typing 
-\texttt{make}.
-\item To generate documentation, please go into:
-\begin{lstlisting}
-~/Software/dopelib-4.0/Examples
-\end{lstlisting}
-Here, you get all options by typing:
+You also get various alternatives by just typing 
+\texttt{make}, including the generation of the documentation.
+Building the documentation will require  
+latex or doxygen to be installed on your computer.
+\item If your would like to check the examples, change to the
+  subfolder \texttt{Examples} from the main source directory.
+Typing 
 \begin{lstlisting}
@@
 \end{lstlisting}
-Specifically, you can create and run all examples, running a test suite,
-and generating documentation:
+will again provide you with the available options:
 {\footnotesize
@@
 }
-The options are $c-all$ and $c-cat$ utilizing the cmake build system.
+The options are $c-all$ and $c-cat$ utilizing the cmake build system
+to create the executables (and optionally run the tests if all went well).
 
@@
 in the home folder of the examples and \textbf{not} as in deal.II by 
-using cmake . in the specific example itself. With this you
-would currently destroy the delivered Makefile. See also the FAQ below.
+using cmake in the specific example itself. With this you
+would currently destroy the delivered Makefile; which is fine if you
+don't intend to use it but some automated tests will fail. See also the FAQ below.
 
-For html documentation, change into:
-\begin{lstlisting}
-~/Software/dopelib-4.0/doxygen
-\end{lstlisting}
-and then typing \texttt{make}. Doxygen documentation
-will require either 
-latex or doxygen to be installed on your computer.
 \item If you wish to test if everything worked. To do so you can 
@@
  consider.\\[3mm]
- As example, the installation of ipopt works as follows:\\
-In the path \texttt{~/dope/ThirdPartyLibs>} type in the terminal\\
+ As example, the installation of \texttt{IPOPT} works as follows:\\
+In the path \texttt{ThirdPartyLibs>} type in the terminal\\
 \texttt{./install-free-libs.sh}
@@
 \subsection*{3.) Problem using make c-all in the Examples folder :}
-
+Sometimes people encounter problems when calling \texttt{make c-all}. 
+This can be due to several reasons:
+\begin{itemize}
+\item If you abort the script for the c-make auto-configuration prematurely,
+you might encounter the following error:
+{\small
 \begin{lstlisting}
-wick@linux:~/Software/dopelib/Examples> make c-all
+[...]> make c-all
 cd OPT/StatPDE/Example1; make -s c-all
 make[2]: *** No targets specified and no makefile found.  Stop.
-../../../..//Examples/Make.global_options:2: recipe for target 'c-all' failed
 make[1]: *** [c-all] Error 2
-Makefile:24: recipe for target 'OPT/StatPDE/Example1' failed
 make: *** [OPT/StatPDE/Example1] Error 2
 \end{lstlisting}
+}
+This is due to the fact, that some unfinished files from the previous
+run are in the way of the script. Go to the corresponding examples
+folder, and delete the folder \texttt{autobuild}.
 
-This means that the Makefile in the specific examples folder has been 
-modified in some illegal way. Remove this Makefile 
+\item Some times the \texttt{Makefiles} in the example folders are
+  damaged, e.g., if an in-source cmake build was attempted in the
+  example folders. This gives an error message 
+{\small
 \begin{lstlisting}
-rm Makefile
-rm -r CMakeFiles
-rm CMakeCache.txt
-rm -r autobuild
+[...]> make c-all
+cd OPT/StatPDE/Example1; make -s c-all
+make[1]: *** No rule to make target `c-all'.  Stop.
+make: *** [OPT/StatPDE/Example1] Error 2   
 \end{lstlisting}
-and get the original 
-one from DOpElib:
-\begin{lstlisting}
-git checkout Makefile
-\end{lstlisting}
+}
+If this happens, go to the corresponding example directory, and replace
+the corresponding Makefile from a clean source, e.g., by a new
+checkout of the Makefile if you are using the git version.
+\end{itemize}
+
 Then try again make c-all (make always sure that before 
@@
 
-\subsection*{4.) Different g++ compilers are used for compiling DOpElib and the Examples folder}
+\subsection*{4.) Different g++ compilers are used for compiling
+  DOpElib and the Examples folder}
+On OpenSUSE and Fedora systems we have observed the following error 
+when calling \texttt{make c-all} in the examples folder:
+{\footnotesize
 \begin{lstlisting}
-bernhard@ebook:~/Schreibtisch/Prog/DOpe/Dope/dopelib/Examples> make c-all
+[...]> make c-all
 cd OPT/StatPDE/Example1; make -s c-all
@@
 \end{lstlisting}
+}
 
-The reason for this problem might be that the variables c++ and g++ as well as cc and gcc do not match.
+The reason for this problem might be that the variables c++ and g++ as
+well as cc and gcc used for compiling \texttt{deal.II} and
+\texttt{DOpElib} do not match.
 \newline
-You can remove the old Examples folder and replace it by a new one and set the compilers manually by
+You can remove the old Examples/autobuild folder and set the compilers manually by
 \begin{lstlisting}
@@
 \end{lstlisting}
-where path/to/compiler should be the path to the compiler, DOpElib was compiled with.
+where path/to/compiler should be the path to the compiler, DOpElib
+(and deal.II) was compiled with.
 You can check the current version of your gcc- and g++-compiler with 
@@
 \end{lstlisting}
+It is even better if you automatically set these variables; e.g.,
+using your \texttt{\${HOME}.profile} file.
+
 Alternatively you can set the variable c++ to your g++ and the variable cc to gcc using update-alternatives
 as for example
+\footnote{Thanks to Florian Tischler (Admin at RICAM Linz) for the
+  help.}
+{\small
 \begin{lstlisting}
@@
 sudo update-alternatives --install /usr/bin/cc cc /usr/bin/gcc-6 60
+\end{lstlisting}}
+
+\subsection*{5.) Different g++, cc, cxx compilers are used for compiling
+  DOpElib and the Examples folder}
+
+Similar to the previous error, using mpi-compilers bears a similar
+problem if the needed libraries are not in your \texttt{\${PATH}}.
+Again this problem seems to appear on OpenSUSE and Fedora systems
+where this happens by default.
+When running \texttt{make c-all} this gives a message like 
+{\small
+\begin{lstlisting}
+Linking CXX executable ../DOpE-PDE-StatPDE-Example1
+/usr/[...]/x86_64-suse-linux/bin/ld: cannot find -lmpi_usempi
+/usr/[...]/x86_64-suse-linux/bin/ld: cannot find -lmpi_mpifh
+/usr/[...]/x86_64-suse-linux/bin/ld: cannot find -lmpi
+collect2: error: ld returned 1 exit status
 \end{lstlisting}
-Thanks to Florian Tischler (Admin at RICAM Linz) for the help.
+}
+
+To fix this issue, you need to set the correct environment variables.
+We assume you have a standard installation of openmpi done. If not you
+can find the correct replacement for the path
+\texttt{/usr/lib64/mpi/gcc/openmpi/bin} 
+using \texttt{whereis mpicxx}.
+{\small
+\begin{lstlisting}
+export DEAL_II_DIR =your_mpi_deal_II_version 
+export PATH=/usr/lib64/mpi/gcc/openmpi/bin:${PATH}  
+export CC=mpicc; export CXX=mpicxx; export FC=mpif90; export FF=mpif77
+\end{lstlisting}}
+Then compile first the DOpElib library and then the Examples folder with these 
+settings. Possibly pay attention that also the deal.II version has been
+compiled 
+with this MPI path.
+
+\begin{remark}
+Be careful that you work in the same shell when only using `export'. 
+Or include the export lines into the \texttt{\${HOME}/.bash\_rc} 
+\texttt{\${HOME}/.profile}
+file to assert that the correct
+compiler is always used.
+\end{remark}
 

### 2017-08-09 Examples/PDE/StatPDE/Example14/content.tex
First running version of the network code

\subsubsection{General problem description}
Within this example, we demonstrate how to solve a system of PDEs on a
network\index{PDE on network}.
The example contains the trivial network consisting of the two lines 
\[
(0,50) \qquad (50,100)
\]
and we solve the PDEs 
\begin{align*}
  \partial_x w &= 1& w(0)&=10,\\
  -\partial_x p &= -1& p(100)&=1.
\end{align*}
together with the coupling conditions 
\[
w^-(50) = w^+(50),\qquad p^-(50) = p^+(50).
\]
The functionals evaluate the $L^1$-error between the computed discrete
solutions and the known exact solution for this problem.

\subsubsection{Implementational Details}
In order to realize the coupling across multiple domains we no longer
use a standard \texttt{dealii::DoFHandler} but rather one for each
domain. Internally, the data is stored in a
\texttt{dealii::BlockVector}, thus only a standard
\texttt{dealii::Vector} may be used on each domain, otherwise the
linker fails since, e.g., no implementation of the needed 
\texttt{dealii::BlockVector<BlockVector>} exists.
The first blocks contain each the variables for one of the PDEs on the
given domains. The last block stores the boundary values for each of
the given domains, first all `left' conditions than all `right'
conditions (corresponding to the boundary\_id $0$ and $1$
respectively).

\paragraph{main.cc}
In the main, not much is changing, for many of the already known
classes there is an update given in the \texttt{namespace
  DOpE::Networks} which internally handle the correct selection of
subdomains.
Further, we have to initialize triangulations for each of the domains
and provide a description of the network topology given in the class \texttt{LocalNetwork}.

\paragraph{localpde.h}
In this class, as always we implement the local element and boundary
integrals (this is unchanged), and the same object is used on all
subdomains, hence if different PDEs need to be coupled these need to
be selected, e.g., by setting the \texttt{material\_id} accordingly.
The only difference is that now, we have to access the externally
given boundary values from the other domains (these values are 
now called fluxes). This can be done by
calling \texttt{GetFluxValues} from 
the \texttt{Networks::Network\_FaceDataContainer}.

For the network, there are some new methods to be implemented.

\texttt{BoundaryEquation\_BV} contains the derivative of the
boundary equation with respect to the given flux variables and store
them according to the residual line in which they appear.

\texttt{BoundaryMatrix\_BV} implements the corresponding
matrix. Notice, that it is implicitly assumed that the boundary values
depend linearly on the fluxes!

\texttt{OutflowValues} implements a selection routine to specify which
of the fluxes are `OutflowValues', i.e., those values that are not
considered in the PDE and thus the corresponding flux must be adjusted
in order to match the given outflow.
In addition, this method must set Boolean flags indicating if a value
is to be considered as outflow.

\texttt{OutflowMatrix} the matrix corresponding to the previous
function. Again, it is assumed that the OutflowValues depend linearly
on the solution variable.

\texttt{PipeCouplingResidual} and \texttt{CouplingMatrix} just pass
arguments to the network.

\paragraph{localnetwork.h} This class implements the topology
information needed on the network. In particular:

\texttt{PipeCouplingResidual} needs to evaluate the residual in the 
coupling conditions. These conditions are always \texttt{NPipes}
outflow conditions and an additional \texttt{NPipes} coupling
conditions between the different flux variables, i.e., the boundary
values for $w$ and $p$ and the continuity condition at the point $50$.
It is assumed, again, that these conditions are (affine)-linear.

\texttt{CouplingMatrix} calculates the matrix corresponding to the
coupling conditions.

\texttt{GetFluxSparsityPattern} is needed to calculate the 
sparsity pattern for the coupling ($2\cdot \text{NPipes} \cdot
\text{NComp} \times 2\cdot \text{NPipes} \cdot
\text{NComp}$) matrix.

### 2018-08-06 Examples/Documentation/introduction.tex
Some modifications in the documention: institute TW, more remarks about draps when installing dope, etc.

@@
 If you like the library and use it for your own projects, please give credits by 
-referencing the project \dope{}~\cite{dope} using the following bibtex entries:
+referencing the project \dope{} using the following bibtex entries \cite{dope,dopelib}
 
@@
 \begin{itemize}
-  \item Thomas Wick (Centre de Mathématiques Appliquées (CMAP) École Polytechnique)
+  \item Thomas Wick (Leibniz Universit\"at Hannover)
   \item Winnifried Wollner (Technische Universität Darmstadt)
@@
 \end{remark}
+\begin{remark}
+\textbf{Furthermore:} please make sure that you point in the above path 
+to the deal.II \texttt{install} folder and NOT in the main directory of your deal.II 
+installation.
+\end{remark}
+
 Detailed installing instructions for deal.II (here the last version
@@
 to build and configure the library using cmake. 
+\begin{remark}
+Due to very frequent problems at this stage, we strongly recommend 
+to read first now Section \ref{sec_FAQ} before yielding to despair.
+\end{remark}
+
 You also get various alternatives by just typing 
@@
 \section{FAQs}
+\label{sec_FAQ}
 \paragraph{1.) When building the library I get an error message:}

### 2018-08-06 Examples/description.tex
Docu modifications related to the new Example 8
First running version of the network code
Testing access to git
Some modifications in the documention: institute TW, more remarks about draps when installing dope, etc.

@@
 
+\usepackage{wallpaper} % Hintergrundbild ueber ganze Titelseite
+\usepackage{eso-pic,transparent}
+
 \newcommand{\dope}{\texttt{DOpElib}}
@@
 \author{
-  Thomas Wick\thanks{École Polytechnique, 
-    {thomas.wick@polytechnique.edu}}, 
+  Thomas Wick\thanks{Leibniz Universit\"at Hannover, 
+    {thomas.wick@ifam.uni-hannover.de}}, 
   Winnifried Wollner\thanks{Technische Universität Darmstadt, 
@@
 \date{\today
-\vspace*{2cm}
-\begin{figure}[h]
-\centering
-{\includegraphics[width=8cm]{Documentation/logo.png}}
-\end{figure}
+\vspace*{8.1cm}
+%\begin{figure}[h]
+%\centering
+%{\includegraphics[width=8cm]{Documentation/logo.png}}
+%\end{figure}
 }
+
+\transparent{1.0}{\ThisCenterWallPaper{0.5}{Documentation/logo.png}}
+%\ThisTileWallPaper
 
@@
 \clearpage
+\subsection{PDEs on Networks}
+\label{PDE_network}
+\input{PDE/StatPDE/Example14/content.tex}
+\clearpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \input{PDE/InstatPDE/Example7/content.tex}
+\cleardoublepage
+\subsection{Fully monolithic phase-field fracture propagation}
+\label{PDE_phase_field_fracture}
+\input{PDE/InstatPDE/Example8/content.tex}
 \cleardoublepage

### 2018-08-07 Examples/Documentation/introduction.tex
Changes in the introduction: cleaning typos, cleaning test in foreword and intro, new Section 2.3, new Chapter 8

@@
 for further information on this license.
+
+\section{Contact / Mailinglist}
+We encourage you to contact us in case you have questions:
+\begin{verbatim}
+dope@dopelib.net 
+\end{verbatim}
 
@@
 \begin{remark}
-Our current DOpE installation has been successfully tested for deal.II version 8.5.0.
+Our current \dope{} installation has been successfully tested for deal.II version 9.0.0.
 \end{remark}
@@
 \section{Installation}\label{sec_installation}
-To work with DOpElib, your need necessarily deal.II, which installation 
-we describe first. Afterwards, the DOpElib installation is described and finally
-other optional packages might be installed.
+Having deal.II and other possible libraries installed, we turn now our focus 
+to the installation of \dope{} itself:
 \begin{itemize}
@@
 Detailed installing instructions for deal.II (here the last version
-8.4.0) can be found on 
-\texttt{http://www.dealii.org/8.4.0/readme.html}. The deal.II 
+9.0.0) can be found on 
+\url{http://www.dealii.org/9.0.0/readme.html}. These
 installation instructions are descriptive enough and we omit 

### 2018-08-07 Examples/Documentation/terminal_output.tex
added missing file

\chapter{Terminal Output  and Parameter Files when Running the Examples}
\label{sec_terminal_output}
We have implemented various options for terminal output and 
data output into files that can be controlled with the help of 
the respective parameter files
\begin{verbatim}
dope.prm
\end{verbatim}
in each Examples folder.

In the following we briefly show such output and briefly 
explain where to control it.
\section{Example of a terminal output: \texttt{PDE/InstatPDE/Example8}}

\begin{verbatim}
Computing State Solution:
Computing Initial Values:
                 Newton step: 0  Residual (abs.):   9.7656e-04
                 Newton step: 0  Residual (rel.):     1.0000e+00
                 Newton step: 1  Residual (rel.): < 1.0000e-11   LineSearch {0} 
Writing [Results/Mesh0/State_InstatPDEProblemContainer.00000.vtk]
                 Precalculating functional values 
        StressX: 0
         Timestep: 1 (0 -> 0.0001) using Crank-Nicolson
                 Newton step: 0  Residual (abs.):   4.0385e-04
                 Newton step: 0  Residual (rel.):     1.0000e+00
                 Newton step: 1  Residual (rel.):   1.8656e-03   LineSearch {0} 
                 Newton step: 2  Residual (rel.):   3.4959e-06   LineSearch {0} 
                 Newton step: 3  Residual (rel.):   1.3608e-10   LineSearch {0} 
Writing [Results/Mesh0/State_InstatPDEProblemContainer.00001.vtk]
                 Precalculating functional values 
        StressX: 6.42408
\end{verbatim}

We observe the respective Newtons step, its current (relative) residual. 
Then, we see at which place the result is written and the respective 
name of that file. Afterwards, we calculate the functional values, here the 
\texttt{StressX} value (see the detailed description of that example before).
Then, we increment the time step number and proceed further.

\section{Controlling the output in the parameter file}
We have implemented various options to control the output via the 
parameter file. Specifically in the part \texttt{subsection output
  parameters}.
Taking the parameter file from the previous example it looks like:
\begin{verbatim}
subsection output parameters
  # File format for the output of solution variables
  set file_format       = .vtk

  # Iteration Counters that should not reflect in the outputname, seperated by
  # `;`
  set ignore_iterations = PDENewton;Cg

  # Name of the logfile
  set logfile           = dope.log

  # Do not write files whose name contains a substring given here by a list of
  # `;` separated words
  set never_write_list  = Gradient;Residual;Hessian;Tangent;Adjoint;Update
  #set never_write_list  = Gradient;Hessian;Tangent;Adjoint

  # Defines what strings should be printed, the higher the number the more
  # output
  set printlevel        = 6

  # Set the precision of the newton output
  set number_precision	 = 4

  # Set manually the machine tolarance for the output
  set eps_machine_set_by_user	 = 1.0e-11

  # Directory where the output goes to
  set results_dir       = Results/
end
\end{verbatim}
Most settings are self-explaining with the help of their respective comments
provided here. In the following, we discuss some of them in more detail though.

\newpage
\subsection{Printlevel}
Let us first look to 
\begin{verbatim}
 set printlevel        = 6
\end{verbatim}
Taking $5$ rather than $6$ yields the terminal output:
\begin{verbatim}
Computing State Solution:
Computing Initial Values:
Writing [Results/Mesh0/State_InstatPDEProblemContainer.00000.vtk]
         Timestep: 1 (0 -> 0.0001) using Crank-Nicolson
Writing [Results/Mesh0/State_InstatPDEProblemContainer.00001.vtk]
\end{verbatim}
Reducing further the number yields less and less output on the screen. 
For debugging purposes in the developement stage, it might be helpful to 
work with print level $6$.

\begin{remark}
In the \texttt{test.prm} file for running the regression tests, the print
level 
should be sufficiently high in order to compare `something'. If nothing 
is printed, nothing can be compared whether \dope{} still works correctly.
\end{remark}

\subsection{Set never write list}
With respect to graphical file output (here *.vtk), specifically in
optimization many steps in the algorithm may be useful when checking 
new developements. For instance, in the previous list, we never write 
\begin{verbatim}
 set never_write_list  = Gradient;Residual;Hessian;Tangent;Adjoint;Update
\end{verbatim}
On the other hand, if \texttt{set never write list} is empty, 
we would write for each vector the respective *.vtk file
which requires for fine discretizations a lot of memory on the hard disk.

Let us give an example: we set 
\begin{verbatim}
 set never_write_list  = Gradient;Residual;Hessian;Tangent;Adjoint;Update
\end{verbatim}
Then, only the physical solution is written as *.vtk:
\begin{verbatim}
Writing [Results/Mesh0/State_InstatPDEProblemContainer.00000.vtk]
\end{verbatim}
If we remove \texttt{Residual} from that list, we obtain:
\begin{verbatim}
 set never_write_list  = Gradient;Hessian;Tangent;Adjoint;Update
\end{verbatim}
and we write the residual of the solution after each Newton step (using
\texttt{printlevel 6} here):
\begin{verbatim}
         Timestep: 1 (0 -> 0.0001) using Crank-Nicolson
Writing [Results/Mesh0/Residualstate.00001.vtk]
                 Newton step: 0  Residual (abs.):   4.0385e-04
                 Newton step: 0  Residual (rel.):     1.0000e+00
Writing [Results/Mesh0/Residualstate.00001.vtk]
                 Newton step: 1  Residual (rel.):   1.8656e-03   LineSearch {0} 
Writing [Results/Mesh0/Residualstate.00001.vtk]
                 Newton step: 2  Residual (rel.):   3.4959e-06   LineSearch {0} 
Writing [Results/Mesh0/Residualstate.00001.vtk]
                 Newton step: 3  Residual (rel.):   1.3608e-10   LineSearch {0} 
Writing [Results/Mesh0/State_InstatPDEProblemContainer.00001.vtk]
                 Precalculating functional values 
        StressX: 6.42408
\end{verbatim}

\begin{remark}
By removing \texttt{Update}, we would also write each Newton update into a
vtk-file.
These options allow detailed debugging of the code.
\end{remark}

\subsection{dope.log}
The \texttt{dope.log} file contains exactly the output of the terminal 
during and after running an example. This allows to double-check the output 
when necessary.

\subsection{Functional output into \texttt{gnuplot}}
At the end of a simulation (in particular a time-dependent problem), we 
are often interested in the temporal evolution of the functionals of interest.
Here, in Example 8 it is \texttt{StressX}. If too much output would appear,
\dope{} passes it directly into a file:
\begin{verbatim}
Computing Functionals:
StressX too large. Writing to file instead: 
Writing [Results/Mesh0/StressX_InstatPDEProblemContainer.00140.gpl]
\end{verbatim}
This file can then be plotted using gnuplot.
The behavior for time-dependent problems is indeed usually to write 
all functional values into a file at the end of a computation.
For stationary problems, the functional value will be printed in the terminal.

### 2018-10-15 Examples/Documentation/examples.tex

@@
 
+  When working with cmake, the easiest option is to set 
+  the DOpElib environment variable, so that it will be found 
+  be cmake (similar to Section \ref{sec_installation}):
+  \begin{verbatim}
+  export DOPE_DIR=path_to_DOpElib
+  \end{verbatim}
+  or alternatively, passing directly a flag to cmake:
+  \begin{verbatim}
+  -DDOPE_DIR=path_to_DOpElib
+  \end{verbatim}
+
 \item If you want to run automated tests on you program so that you can 

### 2018-10-15 Examples/description.tex

@@
   Thomas Wick\thanks{Leibniz Universit\"at Hannover, 
-    {thomas.wick@ifam.uni-hannover.de}}, 
+    {thomas.wick@ifam.uni-hannover.de}} 
+   $\,$ and $\,$
   Winnifried Wollner\thanks{Technische Universität Darmstadt, 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\input{Documentation/terminal_output.tex}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2018-10-15 Examples/Documentation/foreword.tex
Updated Documentation

@@
 \textit{Differential Equations and  Optimization Environment} project. 
-Originally, the project was initiated in the year 2009 at the 
-University of Heidelberg (Germany) in the numerical analysis
+Originally, the project was initiated in the year 2009 at 
+Heidelberg University (Germany) in the numerical analysis
 group of Rolf Rannacher.
@@
 plasticity, fluid dynamics, and multiphysics problems such as 
-fluid-structure interactions.
+fluid-structure interactions, porous media flow or phase-field fracture problems.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-At the present stage the following features are supported by the library
+At the present stage the following features are supported by the library:
 \begin{itemize}
@@
 
+\paragraph{Outline:} 
 The rest of this document is structured as follows: We start with an introduction in
@@
 
+Special thanks go to Christian Goll who has been a great help in 
+maintaining and developing this library from 2009 until 2015.
+
 %{We wish to express our gratitude to those people that have contributed to the library or

### 2018-10-15 Examples/Documentation/introduction.tex

@@
 
+\section{Overview}
+The main library can be roughly split into three parts:
+\begin{verbatim}
+DOpEsrc, Examples, ThirdPartyLibs
+\end{verbatim}
+Here, \texttt{ThirdPartyLibs} was already previously discussed.
+The \texttt{Examples} will be discussed in the chapters \ref{chap:howtoex}
+to \ref{sec_terminal_output}. In this chapter, we concentrate on 
+\texttt{DOpEsrc}. Therein, the library implementation is divided into the following 
+subfolders:
+\begin{verbatim}
+container       reducedproblems    wrapper
+network         source             templates
+basic           include            opt_algorithms  
+interfaces      problemdata        tsschemes
+\end{verbatim}
+
+%\newpage
 \section{Problem description}

### 2018-10-15 Examples/PDE/StatPDE/Example15/main.cc
Example 15: mpi

@@ -1,0 +1,349 @@
+/**
+ *
+ * Copyright (C) 2012-2018 by the DOpElib authors
+ *
+ * This file is part of DOpElib
+ *
+ * DOpElib is free software: you can redistribute it
+ * and/or modify it under the terms of the GNU General Public
+ * License as published by the Free Software Foundation, either
+ * version 3 of the License, or (at your option) any later
+ * version.
+ *
+ * DOpElib is distributed in the hope that it will be
+ * useful, but WITHOUT ANY WARRANTY; without even the implied
+ * warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+ * PURPOSE.  See the GNU General Public License for more
+ * details.
+ *
+ * Please refer to the file LICENSE.TXT included in this distribution
+ * for further information on this license.
+ *
+ **/
+#include <fstream>
+#include <iostream>
+
+#pragma GCC diagnostic ignored "-Wdeprecated-declarations"
+
+#include <deal.II/base/exceptions.h>
+#include <deal.II/base/quadrature_lib.h>
+#include <deal.II/distributed/tria.h>
+#include <deal.II/fe/fe_q.h>
+#include <deal.II/grid/grid_generator.h>
+#include <deal.II/grid/tria.h>
+#include <deal.II/numerics/error_estimator.h>
+
+#include <deal.II/lac/trilinos_block_sparse_matrix.h>
+#include <deal.II/lac/trilinos_parallel_block_vector.h>
+#include <deal.II/lac/trilinos_sparsity_pattern.h>
+
+#include <basic/dopetypes.h>
+#include <basic/mol_statespacetimehandler.h>
+#include <container/integratordatacontainer.h>
+#include <container/pdeproblemcontainer.h>
+#include <include/parameterreader.h>
+#include <interfaces/functionalinterface.h>
+#include <interfaces/pdeinterface.h>
+#include <problemdata/simpledirichletdata.h>
+#include <reducedproblems/statpdeproblem.h>
+#include <templates/gmreslinearsolver.h>
+#include <templates/integrator.h>
+#include <templates/newtonsolver.h>
+#include <wrapper/preconditioner_wrapper.h>
+
+#include "functionals.h"
+#include "localpde.h"
+
+using namespace std;
+using namespace dealii;
[…]

### 2018-10-17 Examples/Documentation/tests.tex
add examples to parallel-dev

@@
 
-\begin{remark}
-Please keep in mind that there are two optimization examples that 
-need external packages to be able to run successfully. These tests will
-fail if these libraries are not installed!
-\end{remark}
+\section{Tests with third-party packages}
+Please keep in mind that DOpElib does not 
+only link to deal.II but also to third party libraries.
+If these libraries are not installed, the respective regression tests 
+will fail.
+
+Currently the following examples link to third-party examples:
+\begin{itemize}
+\item \texttt{OPT/StatPDE/Example6}: IPOPT
+\item \texttt{OPT/StatPDE/Example8}: SNOPT
+\item \texttt{PDE/StatPDE/Example10}: Trilinos (via deal.II)
+\end{itemize}
 

### 2018-10-17 Examples/OPT/StatPDE/Example1/content.tex
add examples to parallel-dev

@@
 Of course one would be sufficient, but we wanted to show how to change optimization 
-solvers easily using \texttt{DOpElib}.
-
+solvers easily using \texttt{DOpElib}.\\
 Next, in the body of the main file, we introduce a second FE function for the 
@@
 two functionalities are highly recommended to check your
-implementation before wondering about your results.
+implementation before wondering about your results.\\
+Finally, this example uses a DWR-error estimator to estimate the 
+error made in the cost functional. In contrast to the 
+error estimation for PDEs here, we have to include the error in the
+control by using the \texttt{HigherOrderDWRContainerControl}.

### 2018-10-17 Examples/OPT/StatPDE/Example10/content.tex
add examples to parallel-dev

@@
 where the fluid can be sucked out of the domain, are in front of the 
-circular inclusion. Hense the drag
+circular inclusion. Hence the drag
 \[
-k(v,p) = \int_{\Gamma_O} n\cdot \sigma(v,p)\cdot d \, ds,
+k(v,p) = \int_{\Gamma_O} n\cdot \sigma(v,p)\cdot d \, \mathrm{d}s,
 \]
@@
 \[
-K(q,v,p) = |k(v,p)|^2 + \frac{\alpha}{2}||q - q_0||^2
+K(q,v,p) = \frac{1}{2}|k(v,p)|^2 + \frac{\alpha}{2}||q - q_0||^2
 \] 
@@
 Consequently, the first derivative can be calculated with only a single integration
--- as a boundary integral in the present example.
+-- as a boundary integral in the present example where the 
+factor $g'$ can be calculated using the drag value in the 
+last iterate  
+\[
+g'(k(v,p)) \quad\text{and}\quad k'(v,p)\delta u(x) = k(\delta v, \delta p)
+\]
+since the drag is linear in $(v,p)$.
+
+For the second derivative, we can calculate in the directions $\delta u$ and $\tau u$
+by the following observation 
+\begin{align*}
+ \left(g\left(\int f(u(x))\,\mathrm{d}x\right)\right)''&(\delta u,\tau u)\\
+  &= g''(\int f(u(x))\,\mathrm{d}x) \int f'(u(x)) \delta u(x) \mathrm{d}x\int f'(u(x)) \tau u(x) \mathrm{d}x \\
+  &\;\;\;\;+ g'(\int f(u(x))\,\mathrm{d}x) \int f''(u(x)) \delta u(x) \tau u(x) \mathrm{d}x \\
+  &= \int \left(g''(\int f(u(x))\,\mathrm{d}x) \int f'(u(x)) \delta u(x) \mathrm{d}x\right)
+ f'(u(x)) \tau u(x) \mathrm{d}x \\
+ &\;\;\;\;+ \int g'(\int f(u(x))\,\mathrm{d}x)f''(u(x)) \delta u(x) \tau u(x) \mathrm{d}x.
+\end{align*}
+Consequently, the second derivative can be calculated as one boundary integral, if 
+the values of 
+\[
+\int f(u(x))\,\mathrm{d}x \quad \text{and} \quad \int f'(u(x)) \delta u(x) \mathrm{d}x
+\]
+in the tangent direction $\delta u$ are available by pre-computations.
+
+To do so in the present example, the following modifications are needed in the cost functional:
+\begin{itemize}
+\item There is a method \texttt{AlgebraicValue} which calculates $0.5 x^2$ for a given value $x$
+- here the pre-computed value of the drag.
+\item In the method \texttt{BoundaryValue}, we have to distinguish several cases.
+  based upon the problem type evaluated. The current value can be accessed by
+  \texttt{GetProblemType()} and if this matches
+  \begin{description}
+    \item[cost\_functional\_pre] then we must calculate the drag-value, i.e., $k(u)$.
+    \item[cost\_functional\_pre\_tangent] then we must calculate the derivative of 
+      the drag in direction $\delta u$ (the provided tangent-direction), i.e., $k(\delta u)$ 
+      since $k$ is linear in its argument.
+    \item[cost\_functional] then we must calculate the rest of the functional. Here only the integral
+      over the control costs is needed.
+  \end{description}
+\item In the methods \texttt{BoundaryValue\_U} and \texttt{BoundaryValue\_UU} we calculate
+  the entire derivative w.r.t $u$ (or second derivative respectively) as one integral 
+  using the formulas for the first and second derivative, respectively. The needed
+  values, i.e.,
+  \[
+  \int f(u)\mathrm{d}x = \int_{\Gamma_O} n \cdot \sigma(v,p)\,d \mathrm{d}s 
+  \]
+  and 
+  \[
+  \int f'(u(x)) \delta u(x) \mathrm{d}x = \int_{\Gamma_O} n \cdot \sigma(\delta v,\delta p)\,d \mathrm{d}s 
+  \]
+  are available as \texttt{ParamValues} with the respective labels 
+  \texttt{cost\_functional\_pre} and \texttt{cost\_functional\_pre\_tangent}.
+
+  Notice, that both \texttt{ParamValues} are a vector of the size given 
+  by \texttt{unsigned int NeedPrecomputations() const}. I.e. if multiple functional parts 
+  need a pre-integration these can be calculated in an arbitrary number of pre-integration runs.
+  In the case of multiple pre-integrations the method \texttt{*Value} needs not only to
+  consider the value of \texttt{GetProblemType()} but also the respective \texttt{GetProblemNum()}
+  running between $0$ and \texttt{NeedPrecomputations()}$-1$. 
+  The order in the ParamValues Vector corresponds to the chosen order of integration.
+\end{itemize}

### 2018-10-17 Examples/OPT/StatPDE/Example2/content.tex
Update Doku in Einklang mit Christians und Winnis Bemerkungen/Zustimmung
add examples to parallel-dev

@@
 \end{verbatim}
-This brings along that we have not only to implement {\color{red}the methods
+Indicating, we have both some point values in the functional, as well as a domain contribution, 
+i.e., we calculate $\|q\| = \int_\Omega |q|\,dx$,
+even though this in not necessary, since $\alpha =0$ and the euclidean 
+norm of $q\in \mathbb R^3$ could be evaluated more easily using an \text{AlgebraicValue} in the functional.
+
+This brings along that we have not only to implement {the methods
 \begin{align*} 

### 2018-10-17 Examples/OPT/StatPDE/Example7/content.tex
add examples to parallel-dev

@@
 The global constraint is the maximum volume of the material,
-which should remain constant with the value $0.5$. Its implementation is 
+which should remain constant with the value $V_{\max} = 0.5$, i.e.,
+\[
+ \int_\Omega q - V_{\max} \,dx \le 0.
+\]
+Its implementation is 
 provided in \texttt{localconstraints.h} where the global constraint is handled as 

### 2018-10-17 Examples/PDE/InstatPDE/Example6/content.tex
add examples to parallel-dev

@@
 Let $I:=[0,T]$ denote the time interval.
-The underlying Biot system reads:
+
+\begin{Problem}[The Biot system]
 Find the pressure $p_{\text{B}}$ and displacement 
@@
 and $f_{\text{B}}$ is a volume force.
+\end{Problem}
 
@@
 \end{equation*}
+
 Usually the non-pay zone is described 
-in terms of linear elasticity: Find 
+in terms of linear elasticity: 
+\begin{Problem}
+Find 
 a displacement $u_S$ such that
@@
 are prescribed with given $\bar{u}_S$ and $\bar{t}_S$.
+\end{Problem}
 
@@
 \end{equation}
-It is important to notice that the second condition 
+
+\subsubsection{Program description}
+In this example, the crucial aspect (from mathematical point of view 
+as well as from the implementation) are the interface conditions
+\eqref{interface_conditions}.
+Here, it is important to notice that the second condition 
 in \eqref{interface_conditions}, requires careful 
@@
 carried out with the help of deal.II's FE Nothing element.
+Second, please do not forget to activate the flag
+\begin{verbatim}
+      HasFaces() const
+      {
+        return true;
+      }
+\end{verbatim}
+
+The problem is driven by traction forces (Neumann conditions), which 
+are imposed via the 
+\begin{verbatim}
+      void
+      BoundaryEquation(...)
+      {
+        ...   
+      }
+\end{verbatim}
+
+As functionals, we evaluate the pressure in two different 
+points of the domain. The observation should be that the 
+pressure first starts increasing reaching a maximum and then
+starts decaying. This is the so-called Mandel-Creyer effect.

### 2018-10-17 Examples/PDE/InstatPDE/Example8/content.tex
add examples to parallel-dev

@@
 \subsubsection{General problem description}
-We consider phase-field fracture propagation on a slit domain, with timedependet 
+We consider phase-field fracture propagation on a slit domain, with time dependent 
 Dirichlet data and homogenous Neumann data.
@@
 \begin{itemize}
-\item Imposing a variational inequality constraint ($\frac{\partial \varphi}{\partial t} \ge 0$)
+\item Imposing a variational inequality constraint ($\frac{\partial \varphi}{\partial t} \leq 0$)
 via penalization;
@@
 allows for an increase of the residual for which, however,
-no convergence results is available:
+no convergence results are available:
 \begin{verbatim}
@@
 the same as \cite{HeWheWi15}) and a fully implicit implementation.
-The fully implicit implementation requires a modified Newton scheme that 
-allows for a temporary increase of the Newton residual \cite{Wi17_pff_error_oriented_Newton}.
+The fully implicit implementation requires a modified Newton scheme (see the
+previous bullet point) that 
+allows for a temporary increase of the Newton residual (similar to \cite{Wi17_pff_error_oriented_Newton}).
 \end{itemize}

### 2018-10-17 Examples/PDE/StatPDE/Example1/content.tex
add examples to parallel-dev

@@
 We choose for simplicity $f=0$.\\
+
+As code verification, we evaluate two different types of functionals. 
+First a point functional measuring the $x$-velocity and 
+a flux functional 
+\[
+\int_{\Gamma_{out}} v\cdot n \, ds,
+\]
+on the outflow boundary. Both a described in the 
+\texttt{functionals.h} file as described below.
 
@@
 Each of these functionals is declared as a class of its own, but in \texttt{DOpElib} all classes are derived from a so-called \texttt{FunctionalInterface} class.\\
-In the current example we declare functionals for point values of the velocity and for the flux at the outflow boundary of the channel.\index{functional!point value} \index{functional!boundary flux}\\
+As already mentioned previsously, in the current example we declare functionals for point values of the velocity and for the flux at the outflow boundary of the channel.\index{functional!point value} \index{functional!boundary flux}\\
 

### 2018-10-17 Examples/PDE/StatPDE/Example13/content.tex
add examples to parallel-dev

@@
 a transport equation. The example corresponds to 
-the dealii step-12.
+the dealii step-12. Here we want to solve the transport equation 
+\begin{align*}
+ \nabla \cdot (\beta u) &= 0 & \text{in }&\Omega,\\
+ u &= g &\text{on }&\Gamma_- := \{ x \in \partial \Omega\,|\, \beta(x) \cdot n(x) < 0\}.
+\end{align*}
+Where $n$ is the outward unit normal, $\Omega = (0,1)^2$ and 
+\[
+\beta(x) = \frac{1}{|x|} \begin{pmatrix} -x_2\\x_1
+  \end{pmatrix}.
+\]
+
+For the numerical solution, as in dealii step~12, we use the upwind discontinuous 
+Galerkin \index{discontinuous Galerkin}. Hence we solve the problem of finding $u_h$ 
+such that 
+\[
+ -\sum_{T \in \mathcal T_h} (u_h \beta \cdot  \nabla v_h)_T 
+ + \sum_{F \in \mathcal F_h} (u_h^-,[\beta \cdot n v_h])_F + (u_h,\beta \cdot n v_h)_{\Gamma_+} = -(g,\beta \cdot n v_h)_{\Gamma_-}
+\]
+where $\Gamma_+ =  \{ x \in \partial \Omega\,|\, \beta(x) \cdot n(x) > 0\}$, $\mathcal T_h$ and $\mathcal F_h$ denote the 
+elements and interior faces of the mesh, respectively. The jump is defined as 
+\[
+[\beta \cdot n v_h] = (v^+ - v^-) \beta \cdot n^+
+\]
+where the superscript $+$ or $-$ denotes the dependence on the upstream $+$ or downstream $-$ element.
 
 \subsubsection{Implementational Details}
+Within this program, we need to make use of the additional \texttt{Face*} and \texttt{Interface*} methods as given in the  
+\texttt{PDEInterface} class. The Face* methods define all integrals on $\mathcal F$ in which the element interacts 
+with it self. The Interface* methods are used for the coupling between the two neighboring elements over the 
+given face.
+
+The program requires the following changes in contrast to the prior examples:
+
+\paragraph{main.cc} We utilize Block Preconditioners for the solution of the resulting system. 
+To this end we included the line
+\begin{verbatim}
+typedef DOpEWrapper::PreconditionBlockSSOR_Wrapper<MATRIX,4> 
+        PRECONDITIONERSSOR; 
+\end{verbatim}
+In contrast to all other preconditioners, we need to specify the block size. This number needs 
+to correspond to the number of unknowns per elements; here 4 since we use Q1-elements. Note that this works for dG
+elements only.
+
+The next important change is that we now not only use a discontinuous element, but we will need to assemble 
+terms on faces between elements that couple the unknowns in the different elements together. For this,
+the matrix needs to have the corresponding non-zero entries specified in the sparsity pattern.  
+To do this, the space time handler has an argument \texttt{bool flux\_pattern} in the constructor that needs to be set to true, i.e., we instantiate as follows: 
+\begin{verbatim}
+STH DOFH(triangulation, state_fe, true);
+\end{verbatim}
+
+Finally, the integration will utilize a special function to be declared in the \texttt{LocalPDE}, 
+hence all objects must use the \texttt{LocalPDE}
+and not the \texttt{PDEInterface}.
+To make sure that this is the case, the \texttt{PDEProblemContainer} needs to be initialized with the following arguments
+\begin{verbatim}
+typedef PDEProblemContainer<LocalPDE<CDC, FDC, DOFHANDLER, VECTOR, DIM>,
+    SimpleDirichletData<VECTOR, DIM>, SPARSITYPATTERN, VECTOR, DIM> OP;
+\end{verbatim}
+
+\paragraph{localpde.h}
+In order to integrate the PDE above, we have to deal with one term that has not been considered before
+\[
+ \sum_{F \in \mathcal F_h} (u_h^-,[\beta \cdot n v_h])_F 
+\]
+Since internally all terms by sums over elements we split this term into contributions on the 
+element edges $\partial K$. On an element $K$ a face $F$, with outward normal $n$ 
+connects to another element $K'$, depending on the sign of $\beta \cdot n$ we have two cases. 
+$\beta \cdot n > 0$ in which case $u^-_h = u_h$ or $\beta \cdot n < 0$ in which case $u^-_h = u_h^* = u_h\lvert_{K'}$, i.e.
+the value from the neighbor. The jump always contains the values $v_h$ and $v_h^*$.
+
+Let now $\beta \cdot n > 0$. Then we assemble the contributions coming only from this element in the \texttt{FaceEquation} 
+(and \texttt{FaceMatrix}),
+i.e.,
+\[
+ (u_h,\beta \cdot n v_h)_F
+\]
+The other part of the jump, namely 
+\[
+ -(u_h,\beta \cdot n v_h^*)_F
+\]
+is not assembled here, since the test functions do not live on the selected element $K$. These contributions will
+be assembled once the element $K'$ is selected (and hence on the same face $\beta \cdot n < 0$.
+Once this is the case, i.e. $\beta \cdot n < 0$, we assemble the other part of the jump,
+which is now 
+\[
+(u_h^*,\beta \cdot n v_h)_F.
+\]
+This is done in the \texttt{InterfaceEquation} (and \texttt{InterfaceMatrix}) since we couple unknowns for the neighboring element $K^*$ (the values of $u_h^*$) with those on $K$ (the values of $v_h$).
+Note that in contrast to the view on the element with
+$\beta \cdot n > 0$ the term has apparently switched signs. This is no typo, but due to the fact, that the outward normal 
+has changed direction.
+
+The precise assembly is analogous to the usual integrals, hence we don't provide more details. The only 
+thing different in the \texttt{InterfaceEquation} and \texttt{InterfaceMatrix} we need to access the 
+values on the element on the other side of the face. To this end, all \texttt{Get*} functions used, 
+such as \texttt{GetFEFaceValuesState},
+have a counterpart \texttt{GetNbr*}, i.e., \texttt{GetNbrFEFaceValuesState}, to access the corresponding values 
+on the neighboring element.
+
+Naturally the two functions
+\begin{verbatim}
+bool HasFaces() const;
+bool HasInterfaces() const;
+\end{verbatim}
+need to return \texttt{true}.
+
+A last and important change is that we now need to implement the 
+method 
+\begin{verbatim}
+template<typename ELEMENTITERATOR>
+ bool AtInterface(ELEMENTITERATOR& element, unsigned int face) const
+ {
+   if (element[0]->neighbor_index(face) != -1) 
+     return true;
+   return false;
+ }
+\end{verbatim}
+that returns true whenever we are on an interior face and false otherwise.

### 2018-10-17 Examples/PDE/StatPDE/Example3/content.tex
add examples to parallel-dev

@@
 We split $\partial \Omega = \Gamma_D \cup \Gamma_{out}$.
-The right hand side of the channel is $\Gamma_{out}$ on which we describe the free outflow condition, on the rest of the boundary we prescribe dirichlet values (An parabolic inflow on the left hand side and zero on the upper and lower channel walls).
+The right hand side of the channel is $\Gamma_{out}$ on which we describe the free outflow condition, on the rest of the boundary we prescribe Dirichlet values (An parabolic inflow on the left hand side and zero on the upper and lower channel walls).
 We choose for simplicity $f=0$..\\
@@
 In the hp-framework we have a stack of finite elements (a \texttt{hp::FECollection}) given.
-We assign each cell of the triangulation an fe-index which determines which finite element we use on this cell.
+We assign each element of the triangulation an fe-index which determines which finite element we use on this element.
 The \texttt{ActiveFEIndexSetter} class manages these indices, see there for more information.

### 2018-10-17 Examples/PDE/StatPDE/Example8/content.tex
add examples to parallel-dev

@@
 We use the symmetric strain tensor $\varepsilon(v) := \frac{1}{2}(\nabla v + \nabla v^T)$, and the symmetric stress tensor $\sigma$ is defined as
-\begin{equation*}
-   \sigma(v) := 2\mu \varepsilon(v)^D + \kappa \cdot tr (\varepsilon(v)) I
-\end{equation*}
+\begin{align*}
+   \sigma(v) &:= 2\mu \varepsilon(v)^D + \rho \, tr (\varepsilon(v)) I = 2\mu\varepsilon(v) + \lambda tr(\varepsilon(v)) I,
+\end{align*}
 where $\tau^D$ is the deviatoric part of a tensor $\tau$, in two dimensions defined as
@@
 \end{equation*}
+Furthermore, the (standard) Lam\'e parameters are denoted by $\mu$ and
+$\lambda$ and which are more conveniently (here and in the code) expressed through
+$\rho = \mu + \lambda$ and $\kappa = 2\mu + \lambda$.
 The main difference with respect to the elastic case is the projection operator $\Pi$ in equation $(1)$. It is defined as follows:
@@
 \end{equation*}
-In our computations, we choose $\sigma_0 = \sqrt{\frac{2}{3}}\cdot 450$, and the above parameters $\mu$ and $\kappa$ as $\mu = 80193.800283$ resp.
-$\kappa = 190937.589172$.
-The corner points of our computational domain are the same as before, and the boundary conditions are not altered, either.\\
+In our computations, we choose $\sigma_0 = \sqrt{\frac{2}{3}}\cdot 450$, and
+the above parameters $\mu, \lambda$ and $\rho$ as $\mu = 80193.800283, \lambda
+= 110743.788889, $ and $\rho = 190937.589172$, respectively. The corner points of our computational domain are the same as before, and the boundary conditions are not altered, either.\\
 The goal of our computations is to detect a subdomain in $\Omega$ where plastic behavior occurs (compare \textit{E.

### 2018-10-17 Examples/PDE/StatPDE/Example10/Test/test.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+if [ $# -ne 1 ]
+    then
+    echo "Usage: "$0" [Test|Store]"
+    exit 1
+fi
+
+PROGRAM=../DOpE-PDE-StatPDE-Example10
+
+../../../../test-single.sh $1 $PROGRAM
+    

### 2018-10-18 Examples/PDE/StatPDE/Example15/content.tex
Example 15: mpi
Example 15: description

--- 2013-05-09 Examples/PDE/StatPDE/Example6/content.tex
@@
-todo. short: this example shows how to use the constraintsmaker class by enforcing periodic boundary conditions in 2d.
+\subsubsection{General problem description}
+The setup of this example is the same as Example 6, i.e. we solve a vector-valued Laplace equation in 3d. The major difference is the use of MPI to parallelize and speedup our program.
 
-\subsubsection{General problem description}
-We solve the vector values laplace equation on a quadratic domain $\Omega$ with a circular hole in the middle, i.e. in strong formulation we look for $u=(u_1,u_2)$ s.t.
-\begin{align*}
--\nabla \cdot \nabla u =& f &&\text{in }\Omega.
-\end{align*}
-We set zero dirichlet values on the circular boundary in the middle of the domain and periodic boundary conditions on the other parts of the boundary.
-We choose the flux over the right hand side boundary as functional.
-We choose
-\begin{align*}
-f(x,y) = \left(\cos\left(\exp(10  x)\right)  y^2x + \sin(y), \cos\left(\exp(10 * y)\right)x^2y + \sin(x)\right)
-\end{align*}
-for the right hand side.
 \subsubsection{Program description}
-This example show how to implement user defined DoF constraints.
-DOpE has an interface for this, namely \texttt{UserDefinedDoFConstraints}.
-In our case, we derive the class \texttt{PeriodicityConstraints}, overwrite the method \texttt{MakeStateDoFConstraints} and give an instance of this class to SpaceTimeHandler at hand:
+MPI based parallelization in DOpE happens mainly behind the scenes. For simple examples like this one, there are only few changes required in the user provided source code.
+
+The first one is to use data structures that allow for MPI based parallelization.
+This is not supported by the previously used dealii::Vector and dealii::SparseMatrix structures, but we have to switch to the corresponding vectors in the dealii::TrilinosWrappers namespace.
+
 \begin{verbatim}
- 	PeriodicityConstraints<DOFHANDLER, DIM> constraints_mkr;
- 	STH DOFH(triangulation, state_fe);
- 	DOFH.SetUserDefinedDoFConstraints(constraints_mkr);
+	using MATRIXBLOCK          = TrilinosWrappers::BlockSparseMatrix;
+	using SPARSITYPATTERNBLOCK = TrilinosWrappers::BlockSparsityPattern;
+	using VECTORBLOCK          = TrilinosWrappers::MPI::BlockVector;
 \end{verbatim}
-This is all it takes. We refer to \textit{myconstraintsmaker.h} for the details of the implementation of the periodicity-constraints.
+
+\begin{remark}
+	\begin{verbatim} using \end{verbatim} is just the more modern version of \begin{verbatim} typedef \end{verbatim}
+\end{remark}
+
+MPI needs to be initialized, this is done via the line 
+\begin{verbatim}
+	dealii::Utilities::MPI::MPI_InitFinalize mpi(argc, argv);
+\end{verbatim}
+which is a deal.II class handling the initialization for us.
+This object has to be created before any call to a MPI function, hence it is advised to put it right at the beginning of the main function.
+MPI also needs to be "finalized", i.e.
+closed explicitly.
+This is included in the previous class and happens when the MPI_InitFinalize object gets out of scope.
+
+% TODO pcout
+
+This are already all the changes required to run our application in parallel. In order to run the application with MPI, we have to call it via 
+\begin{verbatim}
+	mpirun -np 4 application
+\end{verbatim}
+This example call spawns 4 instances of our application, which then proceed to jointly solve our problem.
+The number of cores used can be controlled via the "-np x" option.
+"-np 1" runs the application without MPI parallelization.
+
+\begin{remark}
+	Depending on your system, the exact command might be different, i.e. "srun -n 4 application" for slurm based systems.
+\end{remark}
+
+\begin{remark}
+	Note that on standard personal computers you may not actually see any speedup. This is due to limitations of the memory bandwidth. 
+\end{remark}

### 2018-11-15 Examples/PDE/StatPDE/Example15/content.tex
updated documentation
Added reference to description

@@
 \subsubsection{General problem description}
-The setup of this example is the same as Example 6, i.e. we solve a vector-valued Laplace equation in 3d. The major difference is the use of MPI to parallelize and speedup our program.
+The setup of this example is the same as Example~\ref{PDE_Stat_Laplace_3D}, i.e.
+we solve a vector-valued Laplace equation in 3d.
+The major difference is the use of MPI to parallelize and speedup our program.
 
@@
 closed explicitly.
-This is included in the previous class and happens when the MPI_InitFinalize object gets out of scope.
+This is included in the previous class and happens when the MPI\_InitFinalize object gets out of scope.
 

### 2018-12-16 Examples/OPT/InstatPDE/Example2/content.tex
Added support for time dependent controls and included OPT/InstatPDE/Example3 to show how to use it. Further more, a bug in the functional evaluation has been fixed, up to now it was always assumed that the type ist timelocal. This is now fixed
Updated ControlVector to new common storage Class

@@
 This example is a modified version of \texttt{OPT/InstatPDE/Example1}. 
-Again, we consider the heat equation, this time with an additional nonlinear
-term and most important a time derivative leading to a 
-nonstationary optimization problem. The governing equation is
+Again, we consider the heat equation, this time without an additional nonlinear
+term. The governing equation is
 \begin{equation*}
@@
 \overline{u}(t,x,y) &= \sin(x) \sin(y),\\
-\overline{q}(t,x,y) &= \sin(x) \sin(y)
+\overline{q}(x,y) &= \sin(x) \sin(y)
 \end{align*}
@@
 but has no temporal dependence. For this, the control vector type is set to be 
-{\texttt ControlType::stationary} (the default for stationary equations). In contrast to 
+{\texttt VectorAction::stationary} (the default for stationary equations). In contrast to 
 the case of control in the initial values, the control vector can be accessed at a 

### 2018-12-16 DOpEsrc/basic/mol_multimesh_spacetimehandler.h

@@ -61,7 +61,7 @@ namespace DOpE
     MethodOfLines_MultiMesh_SpaceTimeHandler(
       dealii::Triangulation<dim> &triangulation,
       const FE<dim, dim> &control_fe, const FE<dim, dim> &state_fe,
-      DOpEtypes::ControlType type,
+      DOpEtypes::VectorAction type,
       bool flux_pattern = false,
       const ActiveFEIndexSetterInterface<dim, dim> &index_setter =
         ActiveFEIndexSetterInterface<dim, dim>()) :
@@ -81,7 +81,7 @@ namespace DOpE
     MethodOfLines_MultiMesh_SpaceTimeHandler(
       dealii::Triangulation<dim> &triangulation,
       const FE<dim, dim> &control_fe, const FE<dim, dim> &state_fe,
-      dealii::Triangulation<1> &times, DOpEtypes::ControlType type,
+      dealii::Triangulation<1> &times, DOpEtypes::VectorAction type,
       bool flux_pattern = false,
       const ActiveFEIndexSetterInterface<dim, dim> &index_setter =
         ActiveFEIndexSetterInterface<dim, dim>()) :
@@ -101,7 +101,7 @@ namespace DOpE
     MethodOfLines_MultiMesh_SpaceTimeHandler(
       dealii::Triangulation<dim> &triangulation,
       const FE<dim, dim> &control_fe, const FE<dim, dim> &state_fe,
-      const Constraints &c, DOpEtypes::ControlType type,
+      const Constraints &c, DOpEtypes::VectorAction type,
       bool flux_pattern = false,
       const ActiveFEIndexSetterInterface<dim, dim> &index_setter =
         ActiveFEIndexSetterInterface<dim, dim>()) :
@@ -123,7 +123,7 @@ namespace DOpE
       dealii::Triangulation<dim> &triangulation,
       const FE<dim, dim> &control_fe, const FE<dim, dim> &state_fe,
       dealii::Triangulation<1> &times, const Constraints &c,
-      DOpEtypes::ControlType type,
+      DOpEtypes::VectorAction type,
       bool flux_pattern = false,
       const ActiveFEIndexSetterInterface<dim, dim> &index_setter =
         ActiveFEIndexSetterInterface<dim, dim>()) :

### 2018-12-16 DOpEsrc/include/controlvector.h

@@ -28,18 +28,7 @@
 // TODO remove ...
 //#pragma GCC diagnostic ignored "-Wterminate"
 
-#include <basic/spacetimehandler_base.h>
-#include <basic/dopetypes.h>
-#include <include/parallel_vectors.h>
-#include <include/helper.h>
-
-#include <deal.II/lac/vector.h>
-#include <deal.II/lac/block_vector_base.h>
-#include <deal.II/lac/block_vector.h>
-
-#include <vector>
-#include <iostream>
-#include <sstream>
+#include <include/spacetimevector.h>
 
 
 namespace DOpE
@@ -52,7 +41,7 @@ namespace DOpE
    *                      (i.e. dealii::Vector<double> or dealii::BlockVector<double>)
    */
   template<typename VECTOR>
-  class ControlVector
+    class ControlVector : public SpaceTimeVector<VECTOR>
   {
   public:
     //TODO: Currently we only consider one fixed control
@@ -62,347 +51,24 @@ namespace DOpE
     //      with different meshes for Vectors.
     //      Note that this requires to keep track of the interpolation
     //      between state and control time points...
-    ControlVector(const ControlVector &ref);
-    ControlVector(const SpaceTimeHandlerBase<VECTOR> *STH, DOpEtypes::VectorStorageType behavior);
-    ~ControlVector();
-
-//    /**
-//     * Sets the time in the vector. This Function or SetTimeDoFNumber
-//     * must be called before calling GetSpacialVector
-//     * in order to load the required data.
-//     *
-//     * @param t            A double containing the time we are interested in. If t doesn't match the time given by
-//     *                     time_point, then an interpolation between the corresponding time_points is
-//     *                     computed.
-//     * @param interval      An TimeIterator. The interval containing t.
-//     *
-//     */
-//    void SetTime(double t, const TimeIterator& interval) const;
-    /**
-     * Sets the time in the vector. This Function or SetTime
-     * must be called before calling GetSpacialVector
-     * in order to load the required data.
-     *
-     * @param time_point   An unsigned integer. This gives the number of the point in the  time mesh.
-     *
-     */
-    void SetTimeDoFNumber(unsigned int time_point) const;
-    /**
-     * Returns a reference to the spacial vector associated to the last time given by SetTime*
-     * If the vecor behavior is initial this generates an error if we are not in the
-     * initial time point.
-     */
-    VECTOR &GetSpacialVector();
-    /**
[…]

### 2018-12-16 Examples/OPT/InstatPDE/Example4/main.cc

@@ -168,7 +168,7 @@ main(int argc, char **argv)
 
   //Note that we give DOpEtypes::initial as the type of control.
   MethodOfLines_SpaceTimeHandler<FE, DOFHANDLER, SPARSITYPATTERN, VECTOR, CDIM,
-                                 DIM> DOFH(triangulation, control_fe, state_fe, times, DOpEtypes::ControlType::initial);
+                                 DIM> DOFH(triangulation, control_fe, state_fe, times, DOpEtypes::VectorAction::initial);
 
   NoConstraints<ElementDataContainer, FaceDataContainer, DOFHANDLER, VECTOR, CDIM,
                 DIM> Constraints;
@@ -196,7 +196,7 @@ main(int argc, char **argv)
       if (cases == "check")
         {
           Alg.ReInit();
-          ControlVector<VECTOR> q(&DOFH, DOpEtypes::VectorStorageType::fullmem);
+          ControlVector<VECTOR> q(&DOFH, DOpEtypes::VectorStorageType::fullmem,pr);
           q = 0.1;
           ControlVector<VECTOR> dq(q);
           // eps: step size for difference quotient
@@ -209,7 +209,7 @@ main(int argc, char **argv)
         {
           Alg.ReInit();
 
-          ControlVector<VECTOR> q(&DOFH, DOpEtypes::VectorStorageType::fullmem);
+          ControlVector<VECTOR> q(&DOFH, DOpEtypes::VectorStorageType::fullmem,pr);
 
           Alg.Solve(q);
         }

### 2018-12-16 Examples/OPT/StatPDE/Example6/main.cc

@@ -174,7 +174,7 @@ main(int argc, char **argv)
   IPOPT_Alg Alg(&P, &solver, DOpEtypes::VectorStorageType::fullmem, pr);
 
   Alg.ReInit();
-  ControlVector<BlockVector<double> > q(&DOFH, DOpEtypes::VectorStorageType::fullmem);
+  ControlVector<BlockVector<double> > q(&DOFH, DOpEtypes::VectorStorageType::fullmem,pr);
 
   for (int i = 0; i < niter; i++)
     {

### 2019-01-25 Examples/PDE/StatPDE/Example5/content.tex
add examples to parallel-dev
Updated dwrdatacontainer to work with space-time indicator vectors

@@
 \begin{align*}
--\Delta u &= f && \text{ in } \Omega
+-\Delta u &= f \quad\text{ in } \Omega
 \end{align*}
@@
 \end{verbatim}
-We can now get the error indicators (with signs!) out of \texttt{dwrc} by 
+We can now get the error indicators out of \texttt{dwrc} by 
 \begin{verbatim}
-dwrc.GetErrorIndicators();
+dwrc.GetErrorIndicators()[0];
 \end{verbatim}
-With these indicators, we are now able to refine our grid adaptively (there are several mesh adaption strategies implemented, like 'RefineOptimized', 'RefineFixedNumber' or 'RefineFixedFraction')
+With these indicators\footnote{In contrast to DOpElib 3.0 and earlier, GetErrorIndicators()
+  already returns non-negative values, so no additional manipulations
+  are needed. However, note that the values of GetError() have been
+  obtained by summing over the signed indices, hence the value of
+  GetError() can be smaller than the sum of the ErrorIndicators.
+}, we are now able to refine our grid adaptively (there are several mesh adaption strategies implemented, like 'RefineOptimized', 'RefineFixedNumber' or 'RefineFixedFraction')
 \begin{verbatim}
-DOFH.RefineSpace(RefineOptimized(optimized, error_ind));
+DOFH.RefineSpace(RefineOptimized(error_ind));
 \end{verbatim}
-Note, that one has to take the norm of each entry in the vector of the error indicators before feeding them into the \texttt{RefineSpace} method.

### 2019-01-25 Examples/PDE/StatPDE/Example5/main.cc

@@ -286,9 +286,7 @@ main(int argc, char **argv)
           //For global mesh refinement, uncomment the next line
           // DOFH.RefineSpace(DOpEtypes::RefinementType::global); //or just DOFH.RefineSpace()
 
-          Vector<float> error_ind(dwrc.GetErrorIndicators());
-          for (unsigned int i = 0; i < error_ind.size(); i++)
-            error_ind(i) = std::fabs(error_ind(i));
+          Vector<float> error_ind(dwrc.GetErrorIndicators()[0]);
 
           DOFH.RefineSpace(RefineOptimized(error_ind));
           //There are other mesh refinement strategies implemented, for example

### 2019-03-20 Examples/PDE/InstatPDE/Example9/content.tex
Renamed Example4_TimeSpace to Example8 and started Documentation

\subsubsection{General problem description}

In this example we consider one of the prototypical nonstationary equations, the parabolic heat equation, i.e. for $x \in \Omega \subset \mathbb{R}^d, t\in I =[0,T], T\in \mathbb{R}^+,$ we search for the unknown solution $u:I \times \Omega \rightarrow \mathbb{R}$
\begin{align*}
\partial_t u(t,x) - \Delta u(t,x) &= f(t,x),\\
u(t,x)|_{\partial \Omega} &= g(t,x),\\
u(0,x) &= u_0(x).
\end{align*} In our example, we consider the simplest case $d=1$, where the Laplacian $\Delta$ reduces to $\partial_x^2$. The computational domain is $ I\times \Omega = [0,1] \times [0,1]$. For further simplification, we choose the right hand side as $f=0$ as well as homogeneous Dirichlet boundary conditions ($g=0$). The initial condition is given by $u_0(x) = \min(x,1-x)$.

\subsubsection{Program description}
The novelty of this program is the use of varying spatial meshes in
time, i.e., we use a \index{RotheDiscretization} Rothe-discretization,
where different spatial meshes at different time-points are allowed.

To use the Rothe-Discretization, we use a different DoF-Handler, i.e.,
we need to include
\begin{verbatim}
#include <basic/rothe_statespacetimehandler.h>
\end{verbatim}
To initialize the different meshes, we prepare a vector
\texttt{Rothe\_time\_to\_dof} which stores for each time-point a number
indicating the numer of the DoF-Handler to be used. In this example,
we want a different DoF-Handler at each time-point, hence we
initialize
\begin{verbatim}
std::vector<unsigned int> Rothe_time_to_dof(n_time_steps+1,0);

for(unsigned int i = 0; i < Rothe_time_to_dof.size(); i++)
    Rothe_time_to_dof[i]=i;

Rothe_StateSpaceTimeHandler<FE, DOFHANDLER, SPARSITYPATTERN, VECTOR,
			      DIM> DOFH(triangulation, state_fe, times, Rothe_time_to_dof);
\end{verbatim}
Note, that the vector \texttt{Rothe\_time\_to\_dof} requires its indices
to satisfy the conditions
\begin{itemize}
\item \verb|Rothe_time_to_dof[0] == 0|
\item If \verb|Rothe_time_to_dof[i] == n| for some $n > 0$, then there
  must be an index $j < i$ with \verb|Rothe_time_to_dof[j] == n-1|. 
\end{itemize}
This allows, to use, e.g. only two DoF-Handlers, with number 0 and 1,
that are used on even and odd time-points by setting 
\begin{verbatim}
for(unsigned int i = 0; i < Rothe_time_to_dof.size(); i++)
    Rothe_time_to_dof[i]=i%2;
\end{verbatim}
Similarly, one could use one DoF-Handler for the first 10 time-points
and then a different DoF-Handler by appropriately assigning values to \texttt{Rothe\_time\_to\_dof}.

The only other change to other programms is that now the refinement
needs to be given error indicators for each time-point, i.e., a
\verb|std::vector<dealii::Vector<float> >| where the outer
\texttt{std::vector} needs to have the size \texttt{n\_time\_steps+1}.
This is immeadiately given by the
\texttt{ResidualErrorContainer}. Hence, for refinement of the
different spatial meshes, we just have to call
\begin{verbatim}
const std::vector<dealii::Vector<float> > 
               error_ind(h1resc.GetErrorIndicators());
DOFH.RefineSpace(SpaceTimeRefineOptimized(error_ind));
\end{verbatim}

### 2019-03-20 Examples/PDE/InstatPDE/Example9/main.cc

@@ -50,6 +50,7 @@
 #include <reducedproblems/instatpdeproblem.h>
 #include <templates/instat_step_newtonsolver.h>
 #include <container/instatpdeproblemcontainer.h>
+#include <container/residualestimator.h>
 
 #include <tsschemes/forward_euler_problem.h>
 #include <tsschemes/backward_euler_problem.h>
@@ -119,6 +120,13 @@ typedef InstatStepNewtonSolver<INTEGRATOR, LINEARSOLVER, VECTOR> NLS;
 typedef InstatPDEProblem<NLS, INTEGRATOR, OP, VECTOR,
         DIM> RP;
 
+typedef H1ResidualErrorContainer<Rothe_StateSpaceTimeHandler<FE,
+							     DOFHANDLER,
+							     SPARSITYPATTERN,
+							     VECTOR,
+							     DIM>,
+				 VECTOR, DIM> H1_RESC;
+
 int
 main(int argc, char **argv)
 {
@@ -167,10 +175,12 @@ main(int argc, char **argv)
 
   //Time grid of [0,1]
   Triangulation<1> times;
-  GridGenerator::subdivided_hyper_cube(times, 50);
+  unsigned int n_time_steps=50;
+  GridGenerator::subdivided_hyper_cube(times,n_time_steps);
   triangulation.refine_global(4);
-  std::vector<unsigned int> Rothe_time_to_dof(51,0);
-  Rothe_time_to_dof[25]=1;
+  std::vector<unsigned int> Rothe_time_to_dof(n_time_steps+1,0);
+  for(unsigned int i = 0; i < Rothe_time_to_dof.size(); i++)
+    Rothe_time_to_dof[i]=i%2;
   Rothe_StateSpaceTimeHandler<FE, DOFHANDLER, SPARSITYPATTERN, VECTOR,
 			      DIM> DOFH(triangulation, state_fe, times, Rothe_time_to_dof);
   
@@ -203,6 +213,8 @@ main(int argc, char **argv)
   solver.RegisterOutputHandler(&output);
   solver.RegisterExceptionHandler(&ex);
 
+  H1_RESC h1resc(DOFH, DOpEtypes::VectorStorageType::fullmem, pr, DOpEtypes::primal_only);
+
   for (int j = 0; j < 2; j++)
     {
     try
@@ -225,6 +237,9 @@ main(int argc, char **argv)
         //We compute the value of the functionals. To this end, we have to solve
         //the PDE at hand.
         solver.ComputeReducedFunctionals();
[…]

### 2019-03-20 Examples/PDE/InstatPDE/Example9/content.tex
Changed PDE/InstatPDE/Example9 to 2d

@@
 \end{align*} In our example, we consider the simplest case $d=1$, where the Laplacian $\Delta$ reduces to $\partial_x^2$.
-The computational domain is $ I\times \Omega = [0,1] \times [0,1]$.
+The computational domain is $ I\times \Omega = [0,1] \times [0,1]^2$.
 For further simplification, we choose the right hand side as $f=0$ as well as homogeneous Dirichlet boundary conditions ($g=0$).
-The initial condition is given by $u_0(x) = \min(x,1-x)$.
+The initial condition is given by $u_0(x) = \min(x_1,1-x_1)$.
 
@@
 \end{verbatim}
+
+To evaluate the error indicator, we write the methods
+\begin{verbatim}
+void StrongElementResidual;
+void StrongFaceResidual;
+\end{verbatim}
+in \texttt{localpde.h} analogous to
+Example~\ref{PDE_adap_Stat_Laplace},
+except that the ElementResidual now needs to contain the
+discrete time-derivative, see, e.g.,~\cite{Verfuehrt:2003}.
+Note, that in this example no indicators for the temporal error and
+the mesh-change error are included.

### 2019-03-20 Examples/PDE/InstatPDE/Example9/my_functions.h

@@ -28,18 +28,18 @@
 
 using namespace dealii;
 
-class InitialData : public DOpEWrapper::Function<1>
+class InitialData : public DOpEWrapper::Function<2>
 {
 public:
   InitialData() :
-    DOpEWrapper::Function<1>()
+    DOpEWrapper::Function<2>()
   {
 
   }
   virtual double
-  value(const Point<1> &p, const unsigned int component = 0) const;
+  value(const Point<2> &p, const unsigned int component = 0) const;
   virtual void
-  vector_value(const Point<1> &p, Vector<double> &value) const;
+  vector_value(const Point<2> &p, Vector<double> &value) const;
 
 private:
 
@@ -48,7 +48,7 @@ private:
 /******************************************************/
 
 double
-InitialData::value(const Point<1> &p, const unsigned int /*component*/) const
+InitialData::value(const Point<2> &p, const unsigned int /*component*/) const
 {
   double x = p[0];
 
@@ -59,7 +59,7 @@ InitialData::value(const Point<1> &p, const unsigned int /*component*/) const
 /******************************************************/
 
 void
-InitialData::vector_value(const Point<1> &p, Vector<double> &values) const
+InitialData::vector_value(const Point<2> &p, Vector<double> &values) const
 {
   for (unsigned int c = 0; c < this->n_components; ++c)
     values(c) = InitialData::value(p, c);

### 2019-03-20 Examples/PDE/InstatPDE/Example9/content.tex
Added inhomogeneous Dirichlet-data to Example/InstatPDE/Example9 to show Bug with hanging nodes on inhomogeneous Dirichlet Boundaries

@@
 u(0,x) &= u_0(x).
-\end{align*} In our example, we consider the simplest case $d=1$, where the Laplacian $\Delta$ reduces to $\partial_x^2$.
-The computational domain is $ I\times \Omega = [0,1] \times [0,1]^2$.
-For further simplification, we choose the right hand side as $f=0$ as well as homogeneous Dirichlet boundary conditions ($g=0$).
-The initial condition is given by $u_0(x) = \min(x_1,1-x_1)$.
+\end{align*} In our example, we consider the simplest case $d=1$,
+where the Laplacian $\Delta$ reduces to $\partial_x^2$. The
+computational domain is $ I\times \Omega = [0,1] \times [0,1]^2$. For
+further simplification, we choose the right hand side as $f=0$ and the
+initial condition is given by $u_0(x) = \min(x_1,1-x_1)$.
+The Dirichlet-data are
+\[
+  g(t,x) = \begin{cases} tx_2 & x_1 = 1,\\
+    0 & \text{otherwise.}
+  \end{cases}
+\]
 

### 2019-04-22 Examples/PDE/StatPDE/Example16/content.tex
Added Example for the Obstacle problem

\subsubsection{General problem description}
Within this example, we solve the obstacle problem\index{Obstacle problem}
of finding $u \in \mathcal K$ solving 
\begin{equation}\label{ex16:vi}
  (\nabla u, \nabla (\phi-u)) \ge (f,\phi-u) \quad \forall v \in
  \mathcal K
\end{equation}
where
\[
  \mathcal K = \{v \in H^1_0(\Omega)\,|\, v \ge \chi \text{ a.e. in }\Omega\}
\]
on $\Omega = (-1,1)^2$ with
\[
 \chi(x) = \operatorname{dist}(x,\partial \Omega) - 2
 \operatorname{dist}\Bigl(x,\Omega \setminus (-\tfrac{1}{4},\tfrac{1}{4})\Bigr) -\frac{1}{5}
\]
and $f = -5$ as inspired
by~\cite[Example~7.5]{NochettoSiebertVeeser:2003}.
To handle the variational inequality \index{variational inequality},
we introduce a Lagrange multiplier $\lambda$ such that~\eqref{ex16:vi}
is equivalent to finding $u$ and $\lambda$ solving 
\begin{equation}\label{ex16:lagrange}
  \begin{aligned}
    (\nabla u, \nabla \phi) - (f,\phi) - (\lambda,\phi) &= 0 & \forall
    v &\in H^1_0(\Omega),\\
    \lambda & \ge 0,\\
    u-\chi & \ge 0,\\
    (\lambda,u-\chi) &=0.
  \end{aligned}
\end{equation}
Then, we replace the last three inequalities by a complementarity
function, i.e., we notice that for any $c > 0$ 
\[
  x-\max(0,x-cy) = 0 \quad\Leftrightarrow \quad x \ge 0, y \ge 0, xy
  = 0,
\]
and obtain the formulation
\begin{equation}\label{ex16:complementarity}
  \begin{aligned}
    (\nabla u, \nabla \phi) - (f,\phi) - (\lambda,\phi) &= 0 & \forall
    v &\in H^1_0(\Omega),\\
    \lambda - \max(0,\lambda-c(u-\chi)) &= 0.
  \end{aligned}
\end{equation}
For its discretization, we let $\chi_h = I_h \chi$ be the $\mathcal
Q_1$ interpolation of the obstacle and take $u_h \in \mathcal Q_1$.
For the multiplier $\lambda$, we utilize the dual basis of $\mathcal
Q_1$ to define $\mathcal Q_1^*$, i.e., if $\phi^i$ are the nodal basis
functions for $\mathcal Q_1$, then we define the basis $\psi^i$ of
$\mathcal Q_1^*$ from
\[
 (\phi^i,\psi^j) = \delta_{ij} = \begin{cases} 1 & i = j,\\ 0 &
   \text{otherwise.}
 \end{cases}
\] 
This has the advantage, that
the first equation in \eqref{ex16:complementarity} just gets
\[
  (\nabla u, \nabla \phi^i) - (f,\phi^i) - \lambda_i = 0 \quad \forall
    \phi^i \;\text{nodal basis function of $\mathcal Q_1$.}
\]
The second equation, we enforce in the corresponding vertices $x_i$ to
$\phi^i$, only, and get
\[
    \lambda_i - \max(0,\lambda_i-c(u(x_i)-\chi(x_i))) = 0.
\]
 The advantage of this formulation is that we don't need to actually
 evaluate the basis functions $\psi^i$, since we only need the values
 of $\lambda_i$ which we simply store in a $\mathcal Q_1$
 Finite Element at the vertices since then
 \[
   \lambda_i = \lambda(x_i) 
 \]
 and we use a Gauss-Lobatto quadrature to actually evaluate the
 functions at the vertices. The ugly part is that we need to assert
 that we never evaluate $\lambda$ at a non-vertex quadrature point.

 Now, when running over the mesh, we will have each vertex $x_i$ in
 multiple elements, thus counting its contribution more than once.
 To normalize it by the number of times it is counted, we need to
 divide by this number of elements. We can get it from the
 \texttt{ElementDataContainer} using
 \texttt{GetNNeighbourElementsOfVertex}.
 To use it we need to have the function \texttt{HasVertices} in
 \texttt{LocalPDE} to return true, so that the information is generated.

### 2019-04-22 Examples/PDE/StatPDE/Example16/localpde.h

@@ -1,0 +1,304 @@
+/**
+ *
+ * Copyright (C) 2012-2014 by the DOpElib authors
+ *
+ * This file is part of DOpElib
+ *
+ * DOpElib is free software: you can redistribute it
+ * and/or modify it under the terms of the GNU General Public
+ * License as published by the Free Software Foundation, either
+ * version 3 of the License, or (at your option) any later
+ * version.
+ *
+ * DOpElib is distributed in the hope that it will be
+ * useful, but WITHOUT ANY WARRANTY; without even the implied
+ * warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+ * PURPOSE.  See the GNU General Public License for more
+ * details.
+ *
+ * Please refer to the file LICENSE.TXT included in this distribution
+ * for further information on this license.
+ *
+ **/
+
+#ifndef LOCALPDE_H_
+#define LOCALPDE_H_
+
+#include<limits>
+
+#include <interfaces/pdeinterface.h>
+#include <deal.II/base/numbers.h>
+#include "functions.h"
+
+using namespace std;
+using namespace dealii;
+using namespace DOpE;
+
+/***********************************************************************************************/
+template<
+  template<template<int, int> class DH, typename VECTOR, int dealdim> class EDC,
+  template<template<int, int> class DH, typename VECTOR, int dealdim> class FDC,
+  template<int, int> class DH, typename VECTOR, int dealdim>
+class LocalPDELaplace : public PDEInterface<EDC, FDC, DH, VECTOR, dealdim>
+{
+public:
+  
+  LocalPDELaplace() :
+    state_block_component_(2, 0)
+  {
+    state_block_component_[1] =1;
+  }
+
+  void
+  ElementEquation(const EDC<DH, VECTOR, dealdim> &edc,
+                  dealii::Vector<double> &local_vector, double scale,
+                  double/*scale_ico*/)
+  {
+    unsigned int n_dofs_per_element = edc.GetNDoFsPerElement();
+    unsigned int n_q_points = edc.GetNQPoints();
+    const DOpEWrapper::FEValues<dealdim> &state_fe_values =
+      edc.GetFEValuesState();
+
+    assert(this->problem_type_ == "state");
+
+    ugrads_.resize(n_q_points, std::vector<Tensor<1, dealdim> >(2));
[…]

### 2019-04-24 Examples/PDE/StatPDE/Example16/content.tex
Added error estimation for obstacle problem

@@
  \texttt{LocalPDE} to return true, so that the information is generated.
+
+ Further, in this example we show a residual based error estimator
+ for the obstacle problem, as given in~\cite{Walloth:2017}. In
+ principle this is similar to the calculations needed in
+ Example~\ref{PDE_adap_Stat_Laplace}, i.e., defining the
+ \texttt{Strong*Residual} terms in the \texttt{LocalPDE} class.
+ However, for the estimate of the obstacle we need too special
+ information. First, for a given vertex $x_i$, we need to know if
+ $x_i$ is a \textit{full contact} node, meaning that $u = \chi$ on
+ $\omega_i = \cup_{T \mid x_i \in T} T$ the patch around $v$. Further,
+ we need to have $|\omega_i| = \int_{\omega_i} \phi_i\,\mathrm{d}x$ for all
+ vertices to correctly scale the complementarity residual.
+ Unfortunately, both data are non-local, i.e., we can not compute it
+ on any given $T$ alone.
+
+ However, we can compute these values if we have one additional
+ integration over the domain. I.e. to get $|\omega_i|$ we simply
+ define a function $\omega \in \mathcal Q_1$ by $\omega(v) = \omega_v$
+ which we can calculate by computing the nodal vector corresponding to
+ the functional
+ \[
+   \int_\Omega \phi\,\mathrm{d}x.
+ \]
+ In a similar manner, we can calculate if a vertex is in full contact
+ by adding $1$ to a nodal vector whenever its support is contained in
+ a full contact element, and rescaling with the number of elements
+ adjacent to the corresponding node.
+
+ To execute these computations the \texttt{ObstacleResidualErrorContainer}
+ has a function \texttt{NPrecomputedNodalValues} returning $1$ since we
+ can compute both vectors in a single integration. The actual
+ integration is then implemented in \texttt{LocalPDE} in the
+ \texttt{*AuxRhs} which will be evaluated prior to error estimation.
+ The results are then accessible via the vector
+ \texttt{aux\_errror\_0} from the \texttt{*DataContainers} in the
+ evaluation of the \texttt{Strong*Residual}.
+
+ Note, if more than one precomputation is requested the results will
+ be accessible in \texttt{aux\_error\_i} for an appropriate index $i$.

### 2019-04-24 Examples/PDE/StatPDE/Example16/content.tex
Fixed references and license

@@
  Further, in this example we show a residual based error estimator
- for the obstacle problem, as given in~\cite{Walloth:2017}. In
+ for the obstacle problem, as given in~\cite{Walloth:2017}; see
+ also~\cite{MoonNochettoPetersdorffZhang:2007,KrauseVeeserWalloth:2015} for the original idea for parabolic VIs and the analysis
+ for a Signorini problem. In
  principle this is similar to the calculations needed in

### 2019-04-29 Examples/OPT/StatPDE/Example10/content.tex
Fixed Typo in Docu

@@
 
-\textbf{\texttt{localpde.h}}
+\textbf{\texttt{localfunctional.h}}
 There is a new method \texttt{unsigned int NeedPrecomputations() const}

### 2020-02-23 Examples/OPT/StatPDE/Example4/content.tex
Update Doku. Habe Winnis-Superfunktionen fuer den Moment auskommentiert, da ich trotz tex-Update sonst nicht kompilieren kann
add examples to parallel-dev
Added possibility for geometry based mesh refinement and corresponding example update for OPT/StatPDE/Example4

@@
 \end{gather*}
-on the domain $\Omega = [0,1]^2$. In addition we set the Dirichlet data of the state on the boundary as follows
+on the domain $\Omega = [0,1]^2$. In addition, we set the Dirichlet data of the state on the boundary as follows
 \begin{gather*}
@@
 needed to use the control in the Dirichlet boundary values. 
-\twick{
 In the main file and the localpde program, we work still with the 
 \texttt{FE\underline{}Nothing} element to assign the five controls.
-}
+
+Finally, we refine only the boundary of the domain, to demonstrate how
+refinement driven by geometric features can be realized. To this end,
+the class \texttt{BoundaryRefinement} in \texttt{boundaryrefinement.h}
+implements the marking base upon local information on the individual elements.

### 2021-07-23 Examples/PDE/StatPDE/Example17/content.tex
Added Example17 - Use of FEInterpolatedValues

\subsubsection{General problem description}
 In this example we show how we use the FEInterpolatedValues class in \texttt{DOpElib}.
 We consider a small variation to the nearly incompressible linear elasticity problem given 
 in ~\cite{Linke:2014}. The displacement and pressure are given as follows
 \begin{align*}
   \textbf{u}(x, y) & = \begin{bmatrix}
 		    200 x^2 (1-x)^2y(1-y)(1-2y) \\
 		    -200 y^2(10y)^2x(1-x)(1-2x)
 	  	   \end{bmatrix} \\
   p(x,y) & = 10\left(x - \frac{1}{2}\right)^3y^2 + (1-x)^3\left(y-\frac{1}{2}\right)^3 + \frac{1}{8}.
 \end{align*}
 respectively for the incompressible linear elasticity equation
 \begin{equation}
   \begin{aligned}[b] \label{eq:strong_form}
     -2\mu \nabla \! \cdot \! \varepsilon(\textbf{u}) + \nabla p = f, \\
     \nabla \cdot \textbf{u} + \frac{1}{\lambda} = 0,
   \end{aligned}
 \end{equation}
 where $\lambda \neq \infty$.

 The modified weak form of \eqref{eq:strong_form} is given as
  \begin{align}
   2\mu \int\limits_\Omega \varepsilon(\textbf{u}_h) : \varepsilon(\textbf{v}_h) \; \mathrm{d} x - 
   \int\limits_\Omega p_h \nabla \cdot \textbf{v}_h \; \mathrm{d} x & = \int\limits_\Omega \textbf{f} \cdot  \bm{\pi}^{\rm{div}} \textbf{v}_h \; \mathrm{d} x  \label{eq:modified_weak_form} , \\
   \int\limits_\Omega q_h \nabla \cdot \textbf{u}_h \; \mathrm{d} x  + \dfrac{1}{\lambda}\int\limits_\Omega p_hq_h \; \mathrm{d} x  & = 0  \label{eq:nearly_incompress_modified}
 \end{align}
 where, $\bm{\pi}^{\rm{div}}$ is the interpolation operator that maps the displacement finite element space 
 to $H^{\rm{div}}$ conforming space.

 Most things are identical to all the other programs. However, we discuss the minor changes needed here. 
 Firstly, we need to include the \texttt{interpolatedintegratordatacontainer.h} instead of \texttt{integratordatacontainer.h}.
 Secondly, we define \texttt{IDC idc(velocity\_component, map, fe\_interpolate, quadrature\_formula, face\_quadrature\_formula);}
 according to the syntax of interpolatedintegratordatacontainer. \par
 Now, we can move on to \texttt{localpde.h}. Since we need the interpolated values on the right hand
 side of the equation, we only need to change in \texttt{ElementRightHandSide} class. We use \texttt{InterpEDC} instead of 
 \texttt{EDC}. In addition to initializing the \texttt{state\_fe\_values}, we also initialize \texttt{fe\_values\_interpolated} as \\
 \texttt{InterpolatedFEValues<dealdim> fe\_values\_interpolated = edc.GetInterpolatedFEValues();} 
 When adding the \texttt{local\_vector}, we use \\ \texttt{fe\_values\_interpolated.value(i, q\_point)} instead of 
 \texttt{state\_fe\_values[disp].value(i, q\_point)}.

### 2021-07-23 DOpEsrc/container/interpolatedintegratordatacontainer.h

@@ -94,7 +94,7 @@ namespace DOpE
        interp_edc_ = new InterpolatedElementDataContainer<DH, VECTOR, dim>(selected_component_,
 									   map_, 
 									   fe_interpolate_,
-									   this->GetQuad(), 
+									   quad, 
 									   update_flags,
 									   sth,
 									   element,
@@ -145,7 +145,8 @@ namespace DOpE
       interp_fdc_ = new InterpolatedFaceDataContainer<DH, VECTOR, dim>(selected_component_,
 								       map_, 
 								       fe_interpolate_,
-								       this->GetFaceQuad(),
+								       this->GetQuad(),
+								       fquad,
 								       update_flags, sth,
 								       element, param_values,
 								       domain_values,

### 2021-07-25 Examples/PDE/StatPDE/Example17/content.tex
Removed tmp files, added Copyright, added FIXMEs, updated description of Example 17

@@
 \subsubsection{General problem description}
- In this example we show how we use the FEInterpolatedValues class in \texttt{DOpElib}.
- We consider a small variation to the nearly incompressible linear elasticity problem given 
- in ~\cite{Linke:2014}. The displacement and pressure are given as follows
+ In this example we show how to implement pressure robust finite elements
+ as proposed by~\cite{Linke:2014} using the
+ FEInterpolatedValues class in \texttt{DOpElib}.
+ For this, we consider a small variation to the incompressible
+ Stokes problem in~\cite{Linke:2014} towards quasi incompressible
+ linear elasticity. The displacement and pressure are given as follows
  \begin{align*}
@@
  \end{align*}
- respectively for the incompressible linear elasticity equation
+ respectively for the quasi incompressible linear elasticity equation
  \begin{equation}
@@
      -2\mu \nabla \! \cdot \! \varepsilon(\textbf{u}) + \nabla p = f, \\
-     \nabla \cdot \textbf{u} + \frac{1}{\lambda} = 0,
+     \nabla \cdot \textbf{u} + \frac{1}{\lambda} p = 0,
    \end{aligned}
  \end{equation}
- where $\lambda \neq \infty$.
+ for given parameters $\lambda, \mu > 0$ and $\lambda$ ``large''.
 
- The modified weak form of \eqref{eq:strong_form} is given as
+ For the discretization we search $(\textbf{u}_h,p_h) \in V_h \times Q_h$,
+ here given by continuous $Q_2$ finite elements for $V_h$ and discontinuous
+ $Q_1$ finite elements for $Q_h$.
+ The gradient robust discretization of \eqref{eq:strong_form} is given as
   \begin{align}
@@
  \end{align}
- where, $\bm{\pi}^{\rm{div}}$ is the interpolation operator that maps the displacement finite element space 
- to $H^{\rm{div}}$ conforming space.
+ for all $\textbf{v}_h \in V_h$, $q_h \in Q_h$
+ where, $\bm{\pi}^{\rm{div}} \colon V_h\rightarrow X_h \subset H^{\rm{div}}$
+ is a suitable interpolation operator satisfying
+ \[
+ (q_h, \nabla \cdot \textbf{v}_h) = (q_h \nabla \cdot \bm{\pi}^{\rm{div}}\textbf{v}_h) 
+ \]
+ for all $q_h \in Q_h$ and $\textbf{v}_h \in V_h$. For the choice of $V_h,Q_h$ made here a
+ suitable choice for this space $X_h$ is the space of Raviart-Thomas finite elements
+ with the canonical interpolation.
 
- Most things are identical to all the other programs. However, we discuss the minor changes needed here. 
- Firstly, we need to include the \texttt{interpolatedintegratordatacontainer.h} instead of \texttt{integratordatacontainer.h}.
- Secondly, we define \texttt{IDC idc(velocity\_component, map, fe\_interpolate, quadrature\_formula, face\_quadrature\_formula);}
- according to the syntax of interpolatedintegratordatacontainer. \par
- Now, we can move on to \texttt{localpde.h}. Since we need the interpolated values on the right hand
- side of the equation, we only need to change in \texttt{ElementRightHandSide} class. We use \texttt{InterpEDC} instead of 
- \texttt{EDC}. In addition to initializing the \texttt{state\_fe\_values}, we also initialize \texttt{fe\_values\_interpolated} as \\
- \texttt{InterpolatedFEValues<dealdim> fe\_values\_interpolated = edc.GetInterpolatedFEValues();} 
- When adding the \texttt{local\_vector}, we use \\ \texttt{fe\_values\_interpolated.value(i, q\_point)} instead of 
- \texttt{state\_fe\_values[disp].value(i, q\_point)}.
+ In contrast to standard mixed discretizations as discussed, e.g. in Section~\ref{PDE_Stat_Stokes} only some minor changes are needed which we discuss below.
+
+In \textit{main.cc}, we need to utilize modified versions of the ElementDataContainer,
+FaceDataContainer and IntegratorContainer that give access to the values of
+$\bm{\pi}^{\rm{div}}\textbf{v}_h$ in the local integrals for the righthandside.  
+These classes are provided by the files
+\texttt{interpolatedintegratordatacontainer.h}, 
+\texttt{interpolatedelementdatacontainer.h} and \texttt{interpolatedfacedatacontainer.h}
+For the instantiation, we only need to take care of the InterpolatedIntegratorDataContainer
+via
+\texttt{IDC idc(velocity\_component, map, fe\_interpolate, quadrature\_formula, face\_quadrature\_formula);}
+here the first three arguments are new and needed for the interpolation. The \texttt{velocity\_component} is a \texttt{FEValuesExtractpor::Vector} that indicates which components of the finite
+element $V_h\times Q_h$ should be interpolated onto $X_h$ consisting of the element
+\texttt{fe\_interpolate} here initialized as \texttt{RaviartThomasNodal<2>} of degree $1$.
+The \texttt{map} indicates the Mapping to be used for the transformation of the reference
+RaviartThomas element to the real element.
+
+In \texttt{localpde.h}, the only substantial change occures in the
+\texttt{ElementRightHandSide} where we need to multiply $f$ with
+$\bm{\pi}^{\rm{div}} \textbf{v}_h$.
+Here we utilize that the \texttt{InterpolatedElementDataContainer} gives access to $\bm{\pi}^{\rm{div}} \textbf{v}_h$ by the method \texttt{GetInterpolatedFEValues()}.
+More precisely, initializing 
+\begin{verbatim}
+        InterpolatedFEValues<dealdim> fe_values_interpolated 
+                                      = edc.GetInterpolatedFEValues();   
+\end{verbatim}                
+allows to access $\bm{\pi}^{\rm{div}} \textbf{v}_h$ transparently via
+\begin{verbatim}
+        fe_values_interpolated.value(i, q_point)
+\end{verbatim}
+whereas the choice 
+\begin{verbatim}
+        FEValues<dealdim> fe_values = edc.GetFEValuesState(); 
+        const FEValuesExtractors::Vector disp(0);
+\end{verbatim}                
+allows to access $\textbf{v}_h$ via
+\begin{verbatim}
+        fe_values[disp].value(i, q_point)
+\end{verbatim}

### 2021-07-27 Examples/PDE/InstatPDE/Example8/content.tex
Updated PDE/InstatPDE/Example8 to use an exact complementarity formulation rather than a penalty method.

--- 2017-05-28 Examples/PDE/InstatPDE/Example8/content.tex
@@
 \begin{itemize}
-\item Imposing a variational inequality constraint ($\frac{\partial \varphi}{\partial t} \ge 0$)
-via penalization;
-%\item Implementation of an augmented Lagrangian loop (not yet)
+\item Imposing a variational inequality constraint ($\frac{\partial \varphi}{\partial t} \leq 0$)
+via a Lagrange multiplier and complementarity formulation ;
 \item Stress splitting into tension and compression \`a
@@
 \end{verbatim}
-\item Two localpde* files:
+\item The localpde* file:
 \begin{verbatim}
-localpde_quasi_monolithic.h
 localpde_fully_implicit.h
 \end{verbatim}
-that provide implementations of a quasi-monolithic scheme (similar, but not
-the same as \cite{HeWheWi15}) and a fully implicit implementation.
-The fully implicit implementation requires a modified Newton scheme that 
-allows for a temporary increase of the Newton residual \cite{Wi17_pff_error_oriented_Newton}.
+that provides a fully implicit implementation.
+The fully implicit implementation requires a modified Newton scheme (see the
+previous bullet point) that 
+allows for a temporary increase of the Newton residual (similar to \cite{Wi17_pff_error_oriented_Newton}).
 \end{itemize}
-Let $u$ be the displacement and $\varphi$ be the phase field variable which should highlight the crack.
+We define the function spaces $\mathcal{V}:= H_0^1(\Omega)^2$ and
+$\mathcal{W}:=  H^1(\Omega)$. 
+Further, we define $\mathcal{X}:=\{\tau \in \mathcal{W}^*\, \vert\, \tau \geq 0\}$, 
+where $\mathcal{W}^*$ is a dual space of $\mathcal{W}$ and $\mathcal{K}$ as the convex set
+\[
+\mathcal{K}:= \mathcal{K}^n = \{w\in \mathcal{W} |\, w\leq \varphi^{n-1} \leq 1 \text{ a.e. on }
+\Omega\}.
+\] 
+Let $u \in \mathcal{V}$ be the displacement and $\varphi \in \mathcal{W}$ be the phase field variable which should highlight the crack.
+To realize the inequality constraint, we introduce a Lagrange multiplier $\tau \in \mathcal{X}$ as a third unknown variable. \\
 Furthermore we denote the $L^2(\Omega)$ inner product of $v_1$ and $v_2$ with $(v_1,v_2)$.
-A material is supposed to be undamaged at position $x$ if $\varphi(x)$ is close to 1 and comletely cracked
+A material is supposed to be undamaged at position $x$ if $\varphi(x)$ is close to $1$ and completely cracked
 if it is close to $0$.
@@
 +&\frac{G_c}{2}(\frac{1}{\varepsilon} \Vert 1 - \varphi \Vert^2)+\varepsilon \Vert \nabla \varphi  \Vert\\
-+&\underbrace{\frac{\gamma_{pen}}{2} ((\varphi(t_n)-\varphi(t_{n-1}))_+)^2}_{\text{penalization term}},
+%+&\underbrace{\frac{\gamma_{pen}}{2} ((\varphi(t_n)-\varphi(t_{n-1}))_+)^2}_{\text{penalization term}},
 \end{align*}
-with
+under the constraint 
+\begin{align*}
+\partial_t \varphi \leq 0.
+ \end{align*}
+The constraint realizes the crack irreversibility. Physically-speaking: the
+crack cannot heal. To derive an incremental version, the constraint is discretized in time via:
+\begin{align*}
+ \frac{\varphi(t^{n+1})-\varphi(t^n)}{t^{n+1}-t^n} \leq 0.
+\end{align*}
+Further it is defined
 \begin{align*}
 \sigma(u,\varphi)&:= (1-\kappa)\varphi^2 \sigma(u)^+ + \sigma(u)^-,\\
-\sigma(u)^+ &:= 2*\mu \varepsilon(u)^+ +\lambda (trace(\varepsilon(u)))_+ I,\\
-\sigma(u)^- &:= 2*\mu \varepsilon(u)^- +\lambda (trace(\varepsilon(u)))_- I,\\
-(c)_+ &:= max(0, c), \\
+\sigma(u)^+ &:= 2\mu E(u)^+ +\lambda (\operatorname{trace}(E(u)))_+ \textbf{I},\\
+\sigma(u)^- &:= 2\mu E(u)^- +\lambda (\operatorname{trace}(E(u)))_- \textbf{I},\\
+(c)_+ &:= \max(0, c), \\
 (c)_- &:= c-(c)_+, \\
-\varepsilon(u)^+ &:= \sum_{i=1}^{d} (\lambda_i)_+ n_i^T n_i,\\
-\varepsilon(u)^- &:= \sum_{i=1}^{d} (\lambda_i)_- n_i^T n_i,\\
-\varepsilon(u)   &:= \frac{1}{2}(\nabla u +(\nabla u)^T),
+E(u)^+ &:= \sum_{i=1}^{d} (\lambda_i)_+ n_i^T n_i,\\
+E(u)^- &:= \sum_{i=1}^{d} (\lambda_i)_- n_i^T n_i,\\
+E(u)   &:= \frac{1}{2}(\nabla u +(\nabla u)^T),
 \end{align*}
-where $c \in \mathbb{R}$, $t^n$ denotes the time at timestep $n$, $\lambda_i$ denotes the $i-th$ eigenvalue of $\varepsilon(u)$ and $n_i$ the corresponding eigenvector.
+where $c \in \mathbb{R}$, $t^n$ denotes the time at timestep $n$, $\lambda_i$ denotes the $i-\rm{th}$ eigenvalue of $E(u)$ and $n_i$ the corresponding eigenvector.
 The idea for the choice of the energy functional is presented
 in \cite{MieHofWel2010} (based on the original work \cite{BourFraMar00,BourFraMar08,FraMar98}).
-This leads to the following problem: Find $(u, \varphi) \in V_u \times V_\varphi$ such that
-
+To realize the inequality constraint $\varphi 
+\leq \varphi^{n-1}$, we introduce
+a Lagrange multiplier $\tau_h \in \mathcal{X}_h$ as proposed in \cite[Section~4.1]{MangWickWollner:2020}.
+This leads to the following discrete problem: \\
+Choose discrete function spaces $\mathcal{V}_h \subset \mathcal{V}$, $\mathcal{U}_h \subset \mathcal{U}$, $\mathcal{W}_h \subset \mathcal{W}$ and a proper subset $\mathcal{X}_h \subset \mathcal{X}$.
+Given the initial data $\varphi_h^{n-1} \in \mathcal{W}_h$.
+For the loading steps $n=1,2,\ldots, N$ solve the following system of equations:
+Find $u_h \in \mathcal{V}_h$, $p_h \in \mathcal{U}_h$, $\varphi_h:= \varphi_h^n \in \mathcal{W}_{n,h}$ and $\tau_h \in \mathcal{X}_h$ such that 
  \begin{align*}
-A(u,\varphi)(\phi_u,\phi_\varphi):=& (\sigma(u,\varphi),\varepsilon(\phi_u)) \\
-+& (((1-\kappa)\varphi \sigma(u)^+,\varepsilon(u)),\phi_\varphi)\\
- +& G_c(\frac{1}{\varepsilon} (1 - \varphi,\phi_\varphi)+\varepsilon (\nabla \varphi, \nabla \phi_\varphi)\\
-+&\gamma_{pen} ((\varphi(t_n)-\varphi(t_{n-1}))_+,\phi_\varphi)=0 \quad \forall (\phi_u, \phi_\varphi) \in V_u \times V_\varphi.
+ \begin{aligned}
+   \left(g(\varphi_h^{n-1})\left[2\mu E^+(u_h) + \lambda (\operatorname{trace}(E(u)))_+ \textbf{I}\right],\nabla w_h \right)\\
+  + \left(2\mu \left(E(u_h)-E^+(u_h)\right),E(w_h)\right)\\
+ + \left(\lambda (\operatorname{trace}(E(u)))_- \textbf{I},E(w_h)\right) =&\ 0,\\
+ %\forall w_h\in&\ \mathcal{V}_h,\\
+ %\left(g(\varphi_h) \nabla \cdot u_h,q_h\right) - \frac{1}{\lambda} \left(g(\varphi_h) p_h,q_h\right)=&\ 0,\\
+ %\forall q_h\in&\ \mathcal{U}_h,\\
+  (1- \kappa)\left(\varphi_h 2\mu E^+(u_h) + \lambda (\operatorname{trace}(E(u)))_+ \textbf{I} : E(u_h),\psi_h\right)\\
+  + G_c \left(-\frac{1}{\epsilon} (1-\varphi_h),\psi_h\right)\\ 
+  + G_c \epsilon \left(\nabla \varphi_h,\nabla\psi_h \right) - \left(\tau_h,\psi_h\right) =&\ 0,\\
+  %\forall \psi_h \in&\ \mathcal{W}_h,\\
+%\tau_h\geq&\ 0,\\
+%\varphi_h^n - \varphi_h^{n-1} \leq&\ 0,\\
+%\left(\tau_h,\varphi_h^n -\varphi_h^{n-1}\right)=&\ 0,
+\left(\tau_h - \chi_h,\varphi_h -\varphi_h^{n-1}\right)\geq&\ 0,
+\end{aligned}
 \end{align*}
+for all $\left\{w_h,\psi_h,\chi_h\right\} \in \mathcal{V}_h \times
+\mathcal{K}_h \times \mathcal{X}_h$.
+For the implementation, the last inequality is replaced by a
+non-smooth complementarity function, see, e.g.,~\cite[Section~4.1]{MangWallothWickWollner:2019}.
 
@@
 to \cite{MieHofWel2010,BoVeScoHuLa12}, and \cite{WeWiWo2014}.
-
-\subsubsection{Future improvements}
-Possible future improvements are:
-\begin{itemize}
-\item Implement the phase-field extrapolation in the displacement equation proposed in \cite{HeWheWi15};
-\item Implement the augmented Lagrangian loop proposed by
-Wheeler/Wick/Wollner, CMAME 2014 \cite{WeWiWo2014}. 
-\end{itemize}
-
-% Need to implement extrapolation with two 
-% previous time steps
-% To be changed: 
-% instatpdeproblem.h
-% instat_step_newton_solver.h
-%
-% augmented Lagrangian solver loop
-% -> currently simple penalization
-% has to be implemented in instat_step_newton_solver.h
-%
-% Dokumentation in content.tex

### 2021-07-27 Examples/PDE/InstatPDE/Example10/content.tex
Added PDE/InstatPDE Examples 10, 11, 12

\subsubsection{General problem description}
We consider phase-field fracture propagation on a slit domain, with time dependent 
Dirichlet data and homogeneous Neumann data.
The main novelties are:
\begin{itemize}
\item Imposing a variational inequality constraint ($\frac{\partial \varphi}{\partial t} \leq 0$)
via a Langrage multiplier similar to  \cite{mang2020phase};
%\item Implementation of an augmented Lagrangian loop (not yet)
\item Stress splitting into tension and compression \`a
la \cite{MieWelHof10a}, see as well \cite{AmorMarigoMaurini2009}, located 
in 
\begin{verbatim}
stress_splitting.h
\end{verbatim}
\item Implementation of a Newton solver that 
allows for an increase of the residual for which, however,
no convergence results are available:
\begin{verbatim}
instat_step_modified_newtonsolver.h
\end{verbatim}
\item The localpde* files:
\begin{verbatim}
localpde_quasi_monolithic.h
\end{verbatim}
which provides the implementation of a quasi-monolithic scheme
(similar to\cite{HeWheWi15}).
\end{itemize}

The problem formulation as well as the handling of the irreversibility constraint is exactly the same as in Example 5.2.8.

The test case is the so-called single edged notched 
tension test. For further details about this problem it is referred
to \cite{MieHofWel2010,BoVeScoHuLa12}, and \cite{WeWiWo2014}.

% \subsubsection{Future improvements}
% Possible future improvements are:
% \begin{itemize}
% \item Implement the phase-field extrapolation in the displacement equation proposed in \cite{HeWheWi15};
% \item Implement the augmented Lagrangian loop proposed by
% Wheeler/Wick/Wollner, CMAME 2014 \cite{WeWiWo2014}. 
% \end{itemize}

% Need to implement extrapolation with two 
% previous time steps
% To be changed: 
% instatpdeproblem.h
% instat_step_newton_solver.h
%
% augmented Lagrangian solver loop
% -> currently simple penalization
% has to be implemented in instat_step_newton_solver.h
%
% Dokumentation in content.tex

### 2021-07-27 Examples/PDE/InstatPDE/Example11/content.tex
Added PDE/InstatPDE Examples 10, 11, 12

\subsubsection{General problem description}

In this example, we consider a lower-dimensional fracture in a
  $2$-dimensional domain. This fracture has 
a constant length and varying width.
The driving force in this example is a given constant pressure prescribed in the fracture.
The setting is motivated by the book of Sneddon and Lowengrub
\cite{sneddon1969crack} and therefore known as `Sneddon' benchmark or `pressure-driven cavity'. Analytical solutions are derived in \cite{sneddon1969crack} and are also discussed in 
\cite{bourdin2012variational}. Subsequently, 
\cite{wheeler2014augmented,heister2015primal} coin the proposed benchmark, 
and provide numerical results.\\

Using the notation and problem setup of Example 5.2.8 and combine it with a a given pressure $p:\Omega \to \mathbb{R}$, we get the
following regularized energy functional \cite{MiWheWi19}:

\begin{align*}
\begin{aligned}
 E_{\epsilon}(u,\varphi)=&\ \frac{1}{2}\left(\left((1-\kappa)\varphi^2+\kappa\right) \sigma(u),e(u)\right)+(\varphi^2 p,\Div u)\\ +&\ G_C\left(\frac{1}{2\epsilon}\|1-\varphi\|^2 
 + \frac{\epsilon}{2}\|\nabla\varphi\|^2\right),
\end{aligned}
\end{align*}
where $\kappa$ is a positive regularization parameter for the elastic energy with $\kappa \ll \epsilon$. Further the stress tensor $\sigma(u)$ is given by $\sigma(u) := 2 \mu e(u)+ \lambda \text{tr} (e(u)) \textbf{I}$ 
with the Lam\'e coefficients $\mu,\lambda > 0$.
The linearized strain tensor therein is defined as $e(u):=\frac{1}{2} (\nabla u + \nabla u^T)$.\\

The two-dimensional domain $\Omega = (-10,10)^2$ is sketched in the following figure: 
\begin{figure}[htbp!]
 \centering
\begin{tikzpicture}
\draw[gray] (0,0)  -- (0,4) -- (4,4)  -- (4,0) -- cycle;
\node at (-0.3,-0.3) {\scriptsize{$(-10,-10)$}};
\node at (4.3,-0.3) {\scriptsize{$(10,-10)$}};
\node at (4.3,4.3) {\scriptsize{$(10,10)$}};
\node at (-0.3,4.3) {\scriptsize{$(-10,10)$}};
\node[gray] at (-0.4,0.5) {$\partial\Omega$};
\node at (3.5,3.5) {$\Omega$};
\draw [fill,opacity=0.1, blue] plot [smooth] coordinates { (1.7,2) (1.8,1.9) (2.2,1.9) (2.3,2) (2.2,2.1) (1.8,2.1) (1.7,2)};
\draw [thick, blue] plot [smooth] coordinates { (1.7,2) (1.8,1.9) (2.2,1.9) (2.3,2) (2.2,2.1) (1.8,2.1) (1.7,2)};
\draw[thick, draw=red] (1.8,2)--(2.2,2);
\node[red] at (3,2) {\scriptsize{crack $C$}};
\draw[->,blue] (2,1.6)--(2,1.8);
\node[blue] at (2,1.4) {\scriptsize{transition zone of size $\epsilon$}};
\end{tikzpicture}
\caption{Domain $\Omega$ (in 2D) with Dirichlet boundaries $\partial\Omega$, an initial crack $C$ of length $2l_0$ and a zone of width $\epsilon$, where the phase-field function $\varphi$ is defined.}
 \end{figure}

 An initial crack with length $2l_0 = 2.0$
and thickness $d$ of two cells
on 
$\Omega_c=[-1,1] \times [-d, d] \subset \Omega$ 
is prescribed by help of the phase-field function $\varphi$, i.e.,
$\varphi = 0$ in $\Omega_c$ and $\varphi = 1$ in $\Omega\setminus\Omega_c$.
Note that the thickness of $2d$ corresponds to $2h/\sqrt{2}$.\\
As boundary conditions, the displacements $u$ are set to zero on $\partial
\Omega$.
For the phase-field variable, we use homogeneous Neumann conditions (traction
free), i.e., $\epsilon \partial_n \varphi = 0$ on $\partial \Omega$.\\

With
\[
  \mathcal{V}:=H_0^1(\Omega), \quad  \mathcal{W}_{\text{in}}:=\{w \in H^1(\Omega)| w\leq
  \varphi^{n-1} \leq 1\text{ a.e. on }\Omega\}, \quad\text{and}\quad\mathcal{ W}:=H^1(\Omega),
\]
we obtain the following weak formulation:

\begin{Problem}[Euler-Lagrange System]
Find $(u,\varphi) \in \mathcal{V} \times \mathcal{W}$ with
\begin{align*}
\left( \left((1-\kappa)\varphi^2 + \kappa\right) \sigma(u),e(w)\right)+(\varphi^2 p,\Div w)=0\quad \forall w \in \mathcal{V},
\end{align*}
and
\begin{align*}
(1- \kappa)(&\varphi \sigma(u):e(u),\psi-\varphi) + 2(\varphi p \Div u,\psi-\varphi)\\
+&\ G_C\left(-\frac{1}{\epsilon}(1-\varphi,\psi-\varphi)+\epsilon\left(\nabla \varphi,\nabla(\psi-\varphi)\right) \right)\geq 0\quad \forall \psi \in \mathcal{W}_{\text{in}} \cap L^{\infty}(\Omega).
\end{align*}
\end{Problem}
The crack irreversibilty condition is handled via a Lagrange multiplier as in Example 5.2.8.\\

In this example, we are especially interested in three computed quantities:
\begin{itemize}
 \item The \textbf{total crack volume} (TCV) can be computed numerically using
\begin{align}
\label{eq_TCV_num}
\text{TCV}_{h,\epsilon}=\int_{\Omega} u(x,y) \cdot \nabla \varphi(x,y)\ \mathrm{d}{(x,y)}.
\end{align}
A formula for the limit can be obtained using \cite{sneddon1969crack}.
Using symmetry of the configuration, i.e., $u_y(x,0^+) = - u_y(x,0^-)$,
and the known crack location $[-1,1]\times \{0\}$, one obtains
\begin{align*}
\text{TCV}_{2D}=2 \int_{-\infty}^{\infty} u_y (x,0^+)\ \mathrm{d}x,
\end{align*}
where $0^\pm$ denotes the respective limit from above or below, and
$u_y$ denotes the second ($y$) component of the displacement.

Using the exact representation of $u_y$ (cf. \cite{sneddon1969crack}, page
29)
\[
u_y(x,0^+) =  \frac{p l_0}{E^{\prime}} \left(1-\frac{x^2}{l_0^2}\right)^{1/2}
\]
we obtain:
\begin{align}
\label{eq_TCV_exact}
 \text{TCV}_{2D}  
= \int_{-\infty}^{\infty} 2 u_y (x,0^+)\ \mathrm{d}x 
= \frac{2\pi p l_0^2}{E^{\prime}}.
\end{align}
Applied to our parameter settings, we consequently
obtain the reference value for an infinite domain as:
\[
\text{TCV}_{2D} \approx 6.03186\times 10^{-3}.
\]

\item The \textbf{bulk energy} $E_B$ is given by
\begin{align}
\label{eq_bulk}
E_B= \int_{\Omega} ((1-\kappa)\varphi^2 + \kappa) \psi(e)\, \mathrm{d}{(x,y)}.
\end{align}
The strain energy functional $\psi(e)$ in equation (\ref{eq_bulk}) is defined as
\begin{align*}
 \psi(e):= \mu \tr (e(u)^2) + \frac{1}{2} \lambda \tr (e(u))^2.
\end{align*}
Here, no manufactured reference values are provided and we only present
values computed numerically.
\item The \textbf{crack energy} can be computed via
\begin{align}
\label{eq_crack}
 E_C = \frac{G_C}{2} \int_{\Omega} \left( \frac{(\varphi-1)^2}{\epsilon} +  \epsilon |\nabla \varphi|^2\right) \, \mathrm{d}{(x,y)}.
\end{align}
Again, no manufactured reference values are provided and we only present
values computed numerically.
\end{itemize}

Further details on this numerical example can be found in \cite[Chapter 7]{schroder2020selection}.

%
% Dokumentation in content.tex

### 2021-07-27 Examples/PDE/InstatPDE/Example12/content.tex
Added PDE/InstatPDE Examples 10, 11, 12

\subsubsection{General problem description}

Based on the setup of the previous example, we are interested in simulating a pressure-driven cavity in a nearly incompressible material. To avoid locking effects considering incompressible solids, 
a mixed form for the solid equations is proposed.\\

In the previous example in terms of the total crack volume, 
for $\nu_s = 0.49999$, the fracture in incompressible 
solids would not open anymore and the TCV is almost $0$. On the other hand, 
the formulae in \cite{SneddLow69}[Section 2.4] suggest a value greater than zero. 
The reason being that therein an infinite domain was assumed.
To study incompressible solids in larger domains, we use a trick and add a compressible 
layer as surrounding area. Considering the figure in the previous example, now we work in a domain $(-20,20)^2$ which contains the previously defined domain $(-10,10)^2$.
The surrounding layer of width $10$ is defined as a compressible material with $\nu = 0.2$.
All other parameters, namely $E$, $G_c$, $\kappa$ and $\Omega_c$ are
kept as before.
The same compressible material is used inside of the prescribed
fracture on the set $(-1,1)\times (-d,d)$.

\begin{Problem}[Euler-Lagrange System in Mixed Form]
\label{form_mixed}
Let $p_g\in W^{1,\infty}(\Omega)$ be given. 
For the loading steps $n=1,2,3,\ldots, N$: Find vector-valued displacements,
a scalar-valued pressure, 
and a
scalar-valued phase-field variable 
$\{\bu,p,\varphi\} := \{\bu^n,p^n,\varphi^n\} \in \mathcal{V} \times \mathcal{U} \times \mathcal{W}$ such that
\begin{equation}\label{eq_u_mixed}
\begin{aligned}
\left(g(\varphi)\ \sigma(\bu,p)\,e({\bv})\right) 
+({\varphi}^{2} p_g, \Div  {\bv}) 
+ ({\varphi}^{2} \nabla p_g,  {\bv}) 
&=0 \quad \forall \bv\in \mathcal{V},
\end{aligned}
\end{equation}
and 
\begin{equation}
\begin{aligned}
(\operatorname{tr}e( {\bu})\,, q) - \frac{1}{\lambda} (p,q)
=0 \quad \forall q\in \mathcal{U},
\end{aligned}
\end{equation}
and
%%%%
\begin{equation}
\begin{aligned}
 (1-\kappa) &({\varphi} \;\sigma(\bu,p):e( \bu)\,, \psi {-\varphi}) \\
&+  2 ({\varphi}\;  p_g\; \Div  \bu,\psi{-\varphi})
+ 2\, ({\varphi} \nabla p_g\cdot  {\bu},\psi{-\varphi}) 
\\
&+  G_c  \left( -\frac{1}{\epsilon} (1-\varphi,\psi{-\varphi}) + \epsilon (\nabla
\varphi, \nabla (\psi - {\varphi}))   \right)  \geq  0
\quad \forall \psi \in \mathcal{K},
\end{aligned}
\end{equation}
including the inequality constraint. The stress tensor $\sigma(u,p)$ is given by $\sigma(u,p) := 2 \mu e(u) + p \textbf{I}$ with the Lam\'e coefficients $\mu,\lambda > 0$.
The linearized strain tensor therein is defined as in the previous example as $e(u):=\frac{1}{2} (\nabla u + \nabla u^T)$. 
By $\textbf{I}$, the two-dimensional identity matrix is denoted.
\end{Problem}

\begin{Problem}[Discrete formulation of Problem~\ref{form_mixed}]\label{form_mixed_h}
Let $p_g\in W^{1,\infty}(\Omega)$ be given. 
For the loading steps $n=1,2,3,\ldots, N$: Find vector-valued displacements,
a scalar-valued pressure, 
and a
scalar-valued phase-field variable 
$\{\bu_h,p_h,\varphi_h\} := \{\bu_h^n,p_h^n,\varphi_h^n\} \in \mathcal{V}_h \times \mathcal{U}_h \times \mathcal{W}_h$ such that
\begin{equation}\label{eq_u_mixed_disc}
\begin{aligned}
\left(g(\varphi_h)\;\sigma(\bu_h,p_h)\,, e( \bv_h)\right)  
&+({\varphi_h}^{2} p_g, \Div  \bv_h) \\
&+ ({\varphi_h}^{2} \nabla p_g,  \bv_h) 
=0 \quad \forall \bv_h\in \mathcal{V}_h ,
\end{aligned}
\end{equation}
and 
\begin{equation}\label{eq_p_mixed_disc}
\begin{aligned}
(\operatorname{tr}e(\bu_h)\,, q_h) - \frac{1}{\lambda} (p_h,q_h)
=0 \quad \forall q_h\in \mathcal{U}_h,
\end{aligned}
\end{equation}
and
%%%%
\begin{equation} \label{eq_varphi_mixed_disc}
\begin{aligned}
 (1&-\kappa) (\varphi_h \;\sigma(\bu_h,p_h):e( \bu_h)\,, \psi_h -\varphi_h) \\
&+  2 (\varphi_h\;  p_g\; \Div  \bu_h,\psi_h-\varphi_h)
+ 2\, (\varphi_h \nabla p_g\cdot  \bu_h,\psi_h-\varphi_h) 
\\
&+  G_c  \left( -\frac{1}{\epsilon} (1-\varphi_h,\psi_h-\varphi_h) + \epsilon (\nabla
\varphi_h, \nabla (\psi_h - \varphi_h))   \right)  \geq  0
\quad \forall \psi_h \in \mathcal{K}_h.
\end{aligned}
\end{equation}
\end{Problem}

The mixed problem formulation allows to run test cases with Poisson ratios $\nu=0.3, 0.49, 0.4999$ up to the incompressible limit $\nu=0.5$. 
Further details and results can be found in \cite{basava2020adaptive}.
%later refer to the journal
%
% Dokumentation in content.tex

### 2021-07-27 Examples/PDE/InstatPDE/Example11/localpde.h

@@ -1,0 +1,1150 @@
+/**
+ *
+ * Copyright (C) 2012-2017 by the DOpElib authors
+ *
+ * This file is part of DOpElib
+ *
+ * DOpElib is free software: you can redistribute it
+ * and/or modify it under the terms of the GNU General Public
+ * License as published by the Free Software Foundation, either
+ * version 3 of the License, or (at your option) any later
+ * version.
+ *
+ * DOpElib is distributed in the hope that it will be
+ * useful, but WITHOUT ANY WARRANTY; without even the implied
+ * warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR
+ * PURPOSE.  See the GNU General Public License for more
+ * details.
+ *
+ * Please refer to the file LICENSE.TXT included in this distribution
+ * for further information on this license.
+ *
+ **/
+
+#ifndef SNEDDONPDE_
+#define SNEDDONPDE_
+
+#include <interfaces/pdeinterface.h>
+#include "problem_data.h"
+// Aufspalten des Spannungstensors in sigma+ und sigma-
+// Nicht notwendig bei Sneddon, auch nicht bei Miehe tension
+#include "functions.h"
+
+using namespace DOpE;
+using namespace std;
+using namespace dealii;
+
+/**
+ * This class describes elementwise the weak formulation of the PDE.
+ * See pdeinterface.h for more information.
+ */
+
+template<
+template<template<int, int> class DH, typename VECTOR, int dealdim> class EDC,
+template<template<int, int> class DH, typename VECTOR, int dealdim> class FDC,
+template<int, int> class DH, typename VECTOR, int dealdim>
+class SneddonPDE : public PDEInterface<EDC, FDC, DH, VECTOR, dealdim>
+{
+public:
+
+	static void
+	declare_params(ParameterReader &param_reader)
+	{
+		param_reader.SetSubsection("Local PDE parameters");
+
+		param_reader.declare_entry("constant_k", "0.0", Patterns::Double(0));
+		param_reader.declare_entry("G_c", "0.0", Patterns::Double(0));
+		param_reader.declare_entry("Young_modulus", "1.0", Patterns::Double(0));
[…]

### 2021-07-27 Examples/OPT/StatPDE/Example8/content.tex
Added Benchmark fracture examples - and spellcheck in content files

@@
 In this example, we have now some local (point) constraints, local constraints and a global
-constraint. In constrast to the previous example, these 
+constraint. In contrast to the previous example, these 
 constraints are all taken care of in the 
@@
 Specifically, the point constraints are handled by the 
-\texttt{DOpE::PointConstraints<...>} function, which is also initializad 
+\texttt{DOpE::PointConstraints<...>} function, which is also initialized 
 in the main file.

### 2021-07-27 Examples/PDE/InstatPDE/Example11/content.tex
Added Benchmark fracture examples - and spellcheck in content files

@@
 The setting is motivated by the book of Sneddon and Lowengrub
-\cite{sneddon1969crack} and therefore known as `Sneddon' benchmark or `pressure-driven cavity'. Analytical solutions are derived in \cite{sneddon1969crack} and are also discussed in 
+\cite{SneddLow69} and therefore known as `Sneddon' benchmark or `pressure-driven cavity'. Analytical solutions are derived in \cite{SneddLow69} and are also discussed in 
 \cite{bourdin2012variational}. Subsequently, 
-\cite{wheeler2014augmented,heister2015primal} coin the proposed benchmark, 
+\cite{WeWiWo2014,HeWheWi15} coin the proposed benchmark, 
 and provide numerical results.\\
@@
 \end{Problem}
-The crack irreversibilty condition is handled via a Lagrange multiplier as in Example 5.2.8.\\
+The crack irreversibility condition is handled via a Lagrange multiplier as in Example~\ref{PDE_phase_field_fracture}\\
 
@@
 \end{align}
-A formula for the limit can be obtained using \cite{sneddon1969crack}.
+A formula for the limit can be obtained using \cite{SneddLow69}.
 Using symmetry of the configuration, i.e., $u_y(x,0^+) = - u_y(x,0^-)$,
@@
 
-Using the exact representation of $u_y$ (cf. \cite{sneddon1969crack}, page
+Using the exact representation of $u_y$ (cf. \cite{SneddLow69}, page
 29)
@@
 
-Further details on this numerical example can be found in \cite[Chapter 7]{schroder2020selection}.
+Further details on this numerical example can be found in \cite[Chapter~7]{SchroederEtAl:2021}.
 

### 2021-07-27 Examples/PDE/InstatPDE/Example12/functionals.h

@@ -1,6 +1,6 @@
 /**
  *
- * Copyright (C) 2012-2017 by the DOpElib authors
+ * Copyright (C) 2012-2018 by the DOpElib authors
  *
  * This file is part of DOpElib
  *
@@ -33,10 +33,17 @@ using namespace DOpE;
 
 /****************************************************************************************/
 
+#if DEAL_II_VERSION_GTE(9,3,0)
+template<
+  template<bool DH, typename VECTOR, int dealdim> class EDC,
+  template<bool DH, typename VECTOR, int dealdim> class FDC,
+  bool DH, typename VECTOR, int dealdim>
+#else
 template<
 template<template<int, int> class DH, typename VECTOR, int dealdim> class EDC,
 template<template<int, int> class DH, typename VECTOR, int dealdim> class FDC,
 template<int, int> class DH, typename VECTOR, int dealdim>
+#endif
 class LocalFunctionalTCV : public FunctionalInterface<EDC, FDC, DH, VECTOR, dealdim>
 {
 public:
@@ -123,10 +130,17 @@ public:
 
 /****************************************************************************************/
 
+#if DEAL_II_VERSION_GTE(9,3,0)
+template<
+  template<bool DH, typename VECTOR, int dealdim> class EDC,
+  template<bool DH, typename VECTOR, int dealdim> class FDC,
+  bool DH, typename VECTOR, int dealdim>
+#else
 template<
 template<template<int, int> class DH, typename VECTOR, int dealdim> class EDC,
 template<template<int, int> class DH, typename VECTOR, int dealdim> class FDC,
 template<int, int> class DH, typename VECTOR, int dealdim>
+#endif 
 class LocalFunctionalBulk : public FunctionalInterface<EDC, FDC, DH, VECTOR, dealdim>
 {
 	double constant_k_, lame_coefficient_mu_, lame_coefficient_lambda_;
@@ -246,10 +260,17 @@ public:
 
 /****************************************************************************************/
 
+#if DEAL_II_VERSION_GTE(9,3,0)
+template<
+  template<bool DH, typename VECTOR, int dealdim> class EDC,
+  template<bool DH, typename VECTOR, int dealdim> class FDC,
+  bool DH, typename VECTOR, int dealdim>
+#else
 template<
 template<template<int, int> class DH, typename VECTOR, int dealdim> class EDC,
[…]

### 2021-07-28 Examples/PDE/StatPDE/Example17/content.tex
Added Benchmark fracture examples - and spellcheck in content files
Normalized the signs in the mixed form

@@
  	  	   \end{bmatrix} \\
-   p(x,y) & = 10\left(x - \frac{1}{2}\right)^3y^2 + (1-x)^3\left(y-\frac{1}{2}\right)^3 + \frac{1}{8}.
+   p(x,y) & = -10\left(x - \frac{1}{2}\right)^3y^2 + (1-x)^3\left(y-\frac{1}{2}\right)^3 - \frac{1}{8}.
  \end{align*}
- respectively for the quasi incompressible linear elasticity equation
+ in the incompressible limit $\lambda \rightarrow \infty$ of the 
+ the quasi incompressible linear elasticity equation
  \begin{equation}
    \begin{aligned}[b] \label{eq:strong_form}
-     -2\mu \nabla \! \cdot \! \varepsilon(\textbf{u}) + \nabla p = f, \\
-     \nabla \cdot \textbf{u} + \frac{1}{\lambda} p = 0,
+     -2\mu \nabla \! \cdot \! \varepsilon(\textbf{u}) - \nabla p = f, \\
+     \nabla \cdot \textbf{u} - \frac{1}{\lambda} p = 0,
    \end{aligned}
@@
   \begin{align}
-   2\mu \int\limits_\Omega \varepsilon(\textbf{u}_h) : \varepsilon(\textbf{v}_h) \; \mathrm{d} x - 
+   2\mu \int\limits_\Omega \varepsilon(\textbf{u}_h) : \varepsilon(\textbf{v}_h) \; \mathrm{d} x + 
    \int\limits_\Omega p_h \nabla \cdot \textbf{v}_h \; \mathrm{d} x & = \int\limits_\Omega \textbf{f} \cdot  \bm{\pi}^{\rm{div}} \textbf{v}_h \; \mathrm{d} x  \label{eq:modified_weak_form} , \\
-   \int\limits_\Omega q_h \nabla \cdot \textbf{u}_h \; \mathrm{d} x  + \dfrac{1}{\lambda}\int\limits_\Omega p_hq_h \; \mathrm{d} x  & = 0  \label{eq:nearly_incompress_modified}
+   \int\limits_\Omega q_h \nabla \cdot \textbf{u}_h \; \mathrm{d} x  - \dfrac{1}{\lambda}\int\limits_\Omega p_hq_h \; \mathrm{d} x  & = 0  \label{eq:nearly_incompress_modified}
  \end{align}
@@
 
-In \texttt{localpde.h}, the only substantial change occures in the
+In \texttt{localpde.h}, the only substantial change occurs in the
 \texttt{ElementRightHandSide} where we need to multiply $f$ with

### 2021-09-17 Examples/OPT/StatPDE/Example1/content.tex

@@
 \end{equation*}
-Further, we have an auxilliary\\[2mm]  
+Further, we have an auxiliary\\[2mm]  
 %
@@
 In the given example, $Q = L^2(\Omega)$ and hence the scalar product will be the
-standard $L^2$-inner product. Simillarly, we can define the Hessian operator 
+standard $L^2$-inner product. Similarly, we can define the Hessian operator 
 $H(q) \in \mathcal L(Q,Q)$ by defining
@@
 Here, instead of using a \texttt{pdeproblemcontainer}, we use 
-now an \texttt{optproblemcontainer} which can assemble all additionally needed informations,
+now an \texttt{optproblemcontainer} which can assemble all additionally needed information,
 such as adjoint and tangent PDEs. 

### 2021-09-17 Examples/PDE/InstatPDE/Example10/content.tex

@@
 \subsubsection{General problem description}
-We consider phase-field fracture propagation on a slit domain, with time dependent 
-Dirichlet data and homogeneous Neumann data.
-The main novelties are:
-\begin{itemize}
-\item Imposing a variational inequality constraint ($\frac{\partial \varphi}{\partial t} \leq 0$)
-via a Langrage multiplier similar to  \cite{mang2020phase};
-%\item Implementation of an augmented Lagrangian loop (not yet)
-\item Stress splitting into tension and compression \`a
-la \cite{MieWelHof10a}, see as well \cite{AmorMarigoMaurini2009}, located 
-in 
-\begin{verbatim}
-stress_splitting.h
-\end{verbatim}
-\item Implementation of a Newton solver that 
-allows for an increase of the residual for which, however,
-no convergence results are available:
-\begin{verbatim}
-instat_step_modified_newtonsolver.h
-\end{verbatim}
-\item The localpde* files:
-\begin{verbatim}
-localpde_quasi_monolithic.h
-\end{verbatim}
-which provides the implementation of a quasi-monolithic scheme
-(similar to\cite{HeWheWi15}).
-\end{itemize}
-
-The problem formulation as well as the handling of the irreversibility constraint is exactly the same as in Example 5.2.8.
-
+In this example, we provide an implementation if the
 The test case is the so-called single edged notched 
@@
 to \cite{MieHofWel2010,BoVeScoHuLa12}, and \cite{WeWiWo2014}.
+For the choice of the boundary conditions see also~\cite[Section~5]{MangWallothWickWollner:2019}.
 
-% \subsubsection{Future improvements}
-% Possible future improvements are:
-% \begin{itemize}
-% \item Implement the phase-field extrapolation in the displacement equation proposed in \cite{HeWheWi15};
-% \item Implement the augmented Lagrangian loop proposed by
-% Wheeler/Wick/Wollner, CMAME 2014 \cite{WeWiWo2014}. 
-% \end{itemize}
-
-% Need to implement extrapolation with two 
-% previous time steps
-% To be changed: 
-% instatpdeproblem.h
-% instat_step_newton_solver.h
-%
-% augmented Lagrangian solver loop
-% -> currently simple penalization
-% has to be implemented in instat_step_newton_solver.h
-%
-% Dokumentation in content.tex
+The formulation of the irreversibility constraint  is exactly the same
+as in Example~\ref{PDE_phase_field_fracture}, however, the fracture is
+not handled fully implicit. Instead the elasticity equation
+couples with the phase-field from the previous time step, i.e., in
+contrast to Example~\ref{PDE_phase_field_fracture} the following
+system 
+\begin{align*}
+ \begin{aligned}
+   \left(g(\varphi_h^{n-1})\left[2\mu E^+(u_h) + \lambda (\operatorname{trace}(E(u)))_+ \textbf{I}\right],\nabla w_h \right)\\
+  + \left(2\mu \left(E(u_h)-E^+(u_h)\right),E(w_h)\right)\\
+ + \left(\lambda (\operatorname{trace}(E(u)))_- \textbf{I},E(w_h)\right) =&\ 0,\\
+ %\forall w_h\in&\ \mathcal{V}_h,\\
+ %\left(g(\varphi_h) \nabla \cdot u_h,q_h\right) - \frac{1}{\lambda} \left(g(\varphi_h) p_h,q_h\right)=&\ 0,\\
+ %\forall q_h\in&\ \mathcal{U}_h,\\
+  (1- \kappa)\left(\varphi_h 2\mu E^+(u_h) + \lambda (\operatorname{trace}(E(u)))_+ \textbf{I} : E(u_h),\psi_h\right)\\
+  + G_c \left(-\frac{1}{\epsilon} (1-\varphi_h),\psi_h\right)\\ 
+  + G_c \epsilon \left(\nabla \varphi_h,\nabla\psi_h \right) - \left(\tau_h,\psi_h\right) =&\ 0,\\
+  %\forall \psi_h \in&\ \mathcal{W}_h,\\
+%\tau_h\geq&\ 0,\\
+%\varphi_h^n - \varphi_h^{n-1} \leq&\ 0,\\
+%\left(\tau_h,\varphi_h^n -\varphi_h^{n-1}\right)=&\ 0,
+\left(\tau_h - \chi_h,\varphi_h -\varphi_h^{n-1}\right)\geq&\ 0,
+\end{aligned}
+\end{align*}
+is solved using a complementarity formulation for the variational inequality.

### 2021-09-17 Examples/PDE/InstatPDE/Example9/content.tex

@@
 \texttt{Rothe\_time\_to\_dof} which stores for each time-point a number
-indicating the numer of the DoF-Handler to be used. In this example,
+indicating the number of the DoF-Handler to be used. In this example,
 we want a different DoF-Handler at each time-point, hence we
@@
 
-The only other change to other programms is that now the refinement
+The only other change to other programs is that now the refinement
 needs to be given error indicators for each time-point, i.e., a
@@
 \texttt{std::vector} needs to have the size \texttt{n\_time\_steps+1}.
-This is immeadiately given by the
+This is immediately given by the
 \texttt{ResidualErrorContainer}. Hence, for refinement of the

### 2021-09-17 Examples/description.tex
updated documentation
Renamed Example4_TimeSpace to Example8 and started Documentation
Added Example for the Obstacle problem
Added Example17 - Use of FEInterpolatedValues
Removed tmp files, added Copyright, added FIXMEs, updated description of Example 17

@@
 \usepackage[utf8]{inputenc}
-\usepackage{amsmath,amssymb, amsthm, commath}
+\usepackage{amsmath,amssymb, amsthm, commath, bm}
 \usepackage[pdftex,bookmarks=false,pdfpagelabels=true,plainpages=false]{hyperref}
@@
 \usepackage{listings}
+\usepackage{tikz}
 
@@
 \newtheorem{remark}[Problem]{Remark}
+
+\DeclareMathOperator{\tr}{tr} % new
+\DeclareMathOperator{\Div}{div} % new
+\newcommand{\bu}{\mathbf{u}}
+\newcommand{\bv}{\mathbf{v}}
 
@@
 \clearpage
+\subsection{Parallel Solve for Laplace}
+\label{PDE_parallel}
+\input{PDE/StatPDE/Example15/content.tex}
+\clearpage
+\subsection{Obstacle Problem}
+\label{PDE_obstacle}
+\input{PDE/StatPDE/Example16/content.tex}
+\clearpage
+\subsection{Gradient Robust Discretization of Nearly Incompressible Elasticity}
+\label{PDE_pressurerobust}
+\input{PDE/StatPDE/Example17/content.tex}
+\clearpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \cleardoublepage
-\subsection{Fully monolithic phase-field fracture propagation}
+\subsection{Fully monolithic phase-field fracture propagation: Single Edged Notched Shear Test}
 \label{PDE_phase_field_fracture}
@@
 \cleardoublepage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Heat Equation in 2D with Rothe-discretization}
+\label{PDE_Instat_Heat_Rothe}
+\input{PDE/InstatPDE/Example9/content.tex}
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Time lagged phase-field fracture propagation: Single Edged Notched Tension Test}
+\label{PDE_SENT}
+\input{PDE/InstatPDE/Example10/content.tex}
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Pressure-driven cavity - Sneddon's benchmark}
+\label{PDE_Sneddon}
+\input{PDE/InstatPDE/Example11/content.tex}
+\clearpage
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Sneddon's benchmark layered for incompressible solids}
+\label{PDE_Sneddon-incompressible}
+\input{PDE/InstatPDE/Example12/content.tex}
+\clearpage
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-07-11 Examples/PDE/StatPDE/Example12/content.tex
Ueberarbeitung PDE Statpde examples mit Ausnahme von 9 und 10.
Added Benchmark fracture examples - and spellcheck in content files
Now the spacetimehandlers utilize the userprovided component-block mapping also in the renumbering of the dofs

@@
 This example shows the use of the Raviart-Thomas (RT) element in \texttt{DOpElib}. The example 
-is taken from \texttt{dealii} step-20 and shows how this step can be implemented in \texttt{DOpElib}.
+is taken from \texttt{dealii} step-20 and shows how this step can be implemented in \texttt{DOpElib}. The vector valued Laplace equation is solved in the mixed formulation.
 
 Most things are identical to all the other programs. However there is a subtle difference in 
-\texttt{localpde.h}: Although the element used has 3 components (two for the RT-Element and 1 for the pressure)
-the block\underline{ }component vector mapping blocks to components has only 2 entries. 1 for each finite element used!
+\texttt{localpde.h}: The block\underline{ }component vector mapping
+blocks to components has 3 entries, but since the first finite element
+is non-primitive we need to assign both components (0 and 1) to the
+same block.
 

### 2023-06-18 Examples/Documentation/introduction.tex
fixed old file header
Added Contributors Mang, Walloth

@@
 \section{License information}
-Copyright (C) 2012--2017 by the DOpElib authors\\[2mm]
+Copyright (C) 2012--2018 by the DOpElib authors\\[2mm]
 %
@@
 \end{itemize}
-%%% TODO: add Daniel Jodlbauer and Bernhard Endtmayer
 
@@
 \begin{itemize}
-  \item Daniel Jodlbauer (current work on parallelization)
+  \item Seshadri Basava (Gradient robust methods for vector valued problems)
   \item Bernhard Endtmayer (phase-field fracture, PDE Instat Example 8 - in cooperation with T. Wick)
@@
   \item Masoud Ghaderi (Augmenting the Documentation)
+  \item Daniel Jodlbauer (Current work on parallelization)
   \item Uwe K{\"o}cher (Makefile compatibility)
@@
   \item Matthias Maier (CMake)
+  \item Katrin Mang (Rothe method for non-stationary problems,
+    phase-field damage examples)
+  \item Mirjam Walloth (Error estimation in non-stationary problems)
   \end{itemize}

### 2023-06-18 Examples/OPT/StatPDE/Example2/content.tex

@@
 the number of controls. 
-In addition, }notice that the cost functional is of mixed type (from our computational point of view), i.e.
-the first part is a pointfunctional whereas the regularization part requires the evaluation of a domain integral.
-To handle this, we need a special integrator as well as a special newtonsolver.
-Additionally, our \texttt{LocalFunctional} returns as his type:
+In addition, }notice that the cost functional is of mixed type (from
+our computational point of view), i.e. the first part is a
+point-functional whereas the regularization part requires the
+evaluation of a domain integral. To handle this, we need a special
+integrator as well as a special Newton-solver. Additionally, our \texttt{LocalFunctional} returns as his type:
 \begin{verbatim}

### 2023-06-18 Examples/OPT/StatPDE/Example6/content.tex

@@
 note that the constraints need to be written such, that a feasible control
-generates non positive entried, i.e., we 
+generates non positive entries, i.e., we 
 calculate the vector
@@
 This information tells the SpaceTimeHandler, that the vector 
-$\mathcal C(q)$ needs exactely twice the amount of unknowns as the vector $q$.
+$\mathcal C(q)$ needs exactly twice the amount of unknowns as the vector $q$.
 In general, the space needed for $\mathcal C(q)$ is given as 

### 2023-06-18 Examples/OPT/StatPDE/Example7/content.tex

@@
 we consider now in addition one global constraint. To calculate the correct 
-storrage needed we use the 
+storage needed we use the 
 second argument of \texttt{constraints(lcc, 1)}, which is now one.
@@
 provided in \texttt{localconstraints.h} where the global constraint is handled as 
-a functional, which again is normalized to be nonpositive if the control is feasible.
+a functional, which again is normalized to be non-positive if the control is feasible.

### 2023-06-18 Examples/PDE/InstatPDE/Example5/content.tex
add examples to parallel-dev

--- 2013-05-10 Examples/PDE/InstatPDE/Example5/content.tex
@@
 \end{equation*}
-The computational domain is $\Omega \times I = [0,\pi]^2 \times [0,1]$. From the known solution, we can compute the appropriate data 
+The computational domain is $I\times\Omega = [0,1]\times [0,\pi]^2$. From the known solution, we can compute the appropriate data 
 \begin{align*}
@@
 
-The new feature of this example is the nonhomogeneous righthand side.
-In examples \ref{PDE_Stat_Laplace_2D} and \ref{PDE_Stat_Laplace_3D}, we regarded stationary problems with nonhomogeneous righthand sides, but up to now, we never involved the time variable into the nonhomogeneity.
+The new feature of this example is the non-homogeneous right hand side.
+In examples \ref{PDE_Stat_Laplace_2D} and \ref{PDE_Stat_Laplace_3D}, we regarded stationary problems with non-homogeneous right hand sides, but up to now, we never involved the time variable into the non-homogeneity.
 To do this, \texttt{DOpElib} yields a \texttt{SetTime()} function which has to be applied in the \textit{localpde.h} file as well as at the place where the \texttt{RightHandSideFunction} class is declared (here the \textit{myfunctions.h} file.

### 2023-06-18 Examples/PDE/StatPDE/Example1/content.tex

@@
 \[
-\int_{\Gamma_{out}} v\cdot n \, ds,
+\int_{\Gamma_{out}} v\cdot n \, \mathrm{d}s,
 \]
@@
 Each of these functionals is declared as a class of its own, but in \texttt{DOpElib} all classes are derived from a so-called \texttt{FunctionalInterface} class.\\
-As already mentioned previsously, in the current example we declare functionals for point values of the velocity and for the flux at the outflow boundary of the channel.\index{functional!point value} \index{functional!boundary flux}\\
+As already mentioned previously, in the current example we declare functionals for point values of the velocity and for the flux at the outflow boundary of the channel.\index{functional!point value} \index{functional!boundary flux}\\
 

### 2023-06-18 Examples/PDE/StatPDE/Example13/content.tex

@@
 \begin{verbatim}
-typedef PDEProblemContainer<LocalPDE<CDC, FDC, DOFHANDLER, VECTOR, DIM>,
+typedef PDEProblemContainer<LocalPDE<EDC, FDC, DOFHANDLER, VECTOR, DIM>,
     SimpleDirichletData<VECTOR, DIM>, SPARSITYPATTERN, VECTOR, DIM> OP;

### 2023-06-18 Examples/PDE/StatPDE/Example17/content.tex

@@
  here given by continuous $Q_2$ finite elements for $V_h$ and discontinuous
- $Q_1$ finite elements for $Q_h$.
+ $P_1$ finite elements for $Q_h$.
  The gradient robust discretization of \eqref{eq:strong_form} is given as
@@
  for all $q_h \in Q_h$ and $\textbf{v}_h \in V_h$. For the choice of $V_h,Q_h$ made here a
- suitable choice for this space $X_h$ is the space of Raviart-Thomas finite elements
+ suitable choice for this space $X_h$ is the space of Brezzi-Douglas-Marini $\mathcal{BDM}_2$ elements
  with the canonical interpolation.

### 2023-06-18 Examples/PDE/StatPDE/Example5/content.tex

@@
 
-As we approximate the so called 'weights' $z-\phi_h$ in the error representation by a patchwise higher order interpolation of $z_h$ (the computed dual solution), we have to enforce patch-wise refinement of the grid by giving the flag
+As we approximate the so called 'weights' $z-\phi_h$ in the error representation by a patch-wise higher order interpolation of $z_h$ (the computed dual solution), we have to enforce patch-wise refinement of the grid by giving the flag
 \begin{verbatim}
@@
 \item \texttt{pr}: The \texttt{ParameterReader} which we have already defined.
-\item A enum of type \texttt{EETerms} that tells the container, which error terms we want to compute (primal error indicators vs. dual error indicators, see \cite{BR03}).  
+\item An enum of type \texttt{EETerms} that tells the container, which error terms we want to compute (primal error indicators vs. dual error indicators, see \cite{BR03}).  
 \end{itemize}
